### 2024-10-10 abstract/abstract.tex
updated

\documentclass{article}

\usepackage{luatexja} % 日本語対応のパッケージ
\usepackage[utf8]{inputenc} % UTF-8の文字コードを使用
\usepackage{amsmath} % 数式用のパッケージ
\usepackage{titling} % タイトルカスタマイズ用

\title{{\Large ZF $+\, \neg$AC の相対的無矛盾性の\\Isabelle/ZF による形式的証明}}
\author{舟根 大喜 \\ 東北大学大学院 情報科学研究科 M2}
\date{} % 日付は空白に

\begin{document}

\maketitle

\vspace{-2em}

We formalize the relative consistency proof of ZF$+\neg$AC using Isabelle/ZF proof assistant.
Our approach assumes the existence of a transitive countable model of ZF and uses forcing to construct a symmetric extension which is a model of ZF$+\neg$AC.
We show that the symmetric extension satisfies ZF$+\neg$AC by formalizing a relativized forcing relation based on the formalization of forcing by Gunther et al.

\end{document}