### 2024-11-01 code/output/document/Arities.tex
updated

%
\begin{isabellebody}%
\setisabellecontext{Arities}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Arities of internalized formulas%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Arities\isanewline
\ \ \isakeyword{imports}\ FrecR\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ \ t{\isadigit{1}}{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isadigit{2}}{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ up{\isasymin}nat\ \ {\isasymrbrakk}\ {\isasymLongrightarrow}\ \isanewline
\ \ arity{\isacharparenleft}{\kern0pt}upair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}t{\isadigit{1}}{\isacharcomma}{\kern0pt}t{\isadigit{2}}{\isacharcomma}{\kern0pt}up{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymUnion}\ {\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}t{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}t{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}up{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ \ upair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un\ \ \ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ \ t{\isadigit{1}}{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isadigit{2}}{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ p{\isasymin}nat\ \ {\isasymrbrakk}\ {\isasymLongrightarrow}\ \isanewline
\ \ arity{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}t{\isadigit{1}}{\isacharcomma}{\kern0pt}t{\isadigit{2}}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymUnion}\ {\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}t{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}t{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}composition{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ s{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}composition{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}s{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymUnion}\ {\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}s{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ composition{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \ \ \ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ z{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}domain{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}range{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ z{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}range{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ range{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ y{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ z{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymUnion}\ {\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ y{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ z{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymUnion}\ {\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ y{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ z{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymUnion}\ {\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ y{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ y{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ f{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ arity{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ succ{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}field{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ z{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}field{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}range{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ empty{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat{\isacharsemicolon}{\kern0pt}y{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ cons{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ number{\isadigit{1}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}number{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ number{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}relation{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}relation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}restriction{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ r{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ z{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ A{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}restriction{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ restriction{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}typed{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ y{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ f{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ arity{\isacharparenleft}{\kern0pt}typed{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymUnion}\ {\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ typed{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}relation{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ y{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ subset{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}transset{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}transset{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ transset{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}transset{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}limit{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}limit{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ limit{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}finite{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}finite{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}limit{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}omega{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}omega{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ omega{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}limit{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}cartprod{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ A{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ B{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ z{\isasymin}nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}cartprod{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}B{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ fst{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ snd{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ftype{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ name{\isadigit{1}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ name{\isadigit{2}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ singleton{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ cons{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Memrel{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ \ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}quasinat{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}quasinat{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ quasinat{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ cons{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\isanewline
\ \ \ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}p{\isasymin}formula\ {\isacharsemicolon}{\kern0pt}\ v{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ n{\isasymin}nat{\isacharsemicolon}{\kern0pt}\ Z{\isasymin}nat{\isacharsemicolon}{\kern0pt}i{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ \ arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ i\ {\isasymLongrightarrow}\ \isanewline
\ \ arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}v{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}Z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}Z{\isacharparenright}{\kern0pt}\ {\isasymunion}\ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}restriction{\isacharunderscore}{\kern0pt}fm\isanewline
\ \ \ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}p{\isasymin}formula\ {\isacharsemicolon}{\kern0pt}\ v{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ n{\isasymin}nat{\isacharsemicolon}{\kern0pt}\ Z{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ i{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ i\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}v{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}Z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}Z{\isacharparenright}{\kern0pt}\ {\isasymunion}\ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm\ \ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}case{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}p{\isasymin}formula\ {\isacharsemicolon}{\kern0pt}\ v{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ n{\isasymin}nat{\isacharsemicolon}{\kern0pt}\ Z{\isasymin}nat{\isacharsemicolon}{\kern0pt}\ i{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ i\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}case{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}v{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}Z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}Z{\isacharparenright}{\kern0pt}\ {\isasymunion}\ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}case{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}quasinat{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}isF{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}v{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}g{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}i{\isasymin}nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}isF{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ i{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}isF{\isacharcomma}{\kern0pt}v{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymunion}\ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymphi}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Forall{\isacharparenleft}{\kern0pt}Implies{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ isF{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}ar\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}{\isacharquery}{\kern0pt}ar{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm\isanewline
\ \ \ \ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\ Un{\isacharunderscore}{\kern0pt}assoc{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}case{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ assms\ pred{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}eq\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}v{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}Z{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}i{\isasymin}nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ i{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}v{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}Z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}Z{\isacharparenright}{\kern0pt}\ {\isasymunion}\ \isanewline
\ \ \ \ \ \ \ \ \ \ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymphi}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}v{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasympsi}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}Z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}v{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{8}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isasymunion}\ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ arity{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}Z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isasymunion}\ \isanewline
\ \ \ \ \ \ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}fm\ assms\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}t{\isasymin}nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymphi}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}t{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \ \isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymphi}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{3}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}finite{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}fm\ name{\isadigit{1}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}fm\ name{\isadigit{2}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ecloseN{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ t{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}ecloseN{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}frecR{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}a{\isasymin}nat{\isacharsemicolon}{\kern0pt}b{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}frecR{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}fm\ name{\isadigit{1}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ name{\isadigit{2}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ \ \ number{\isadigit{1}}arity{\isacharunderscore}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}Collect{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Collect{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymunion}\ pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Arities.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%35=4%:%
%:%38=5%:%
%:%39=6%:%
%:%40=6%:%
%:%41=7%:%
%:%44=8%:%
%:%48=8%:%
%:%49=8%:%
%:%50=9%:%
%:%51=9%:%
%:%52=10%:%
%:%53=10%:%
%:%58=10%:%
%:%61=11%:%
%:%62=12%:%
%:%63=13%:%
%:%64=13%:%
%:%65=14%:%
%:%68=15%:%
%:%72=15%:%
%:%73=15%:%
%:%74=16%:%
%:%75=16%:%
%:%76=17%:%
%:%77=17%:%
%:%82=17%:%
%:%85=18%:%
%:%86=19%:%
%:%87=19%:%
%:%88=20%:%
%:%91=21%:%
%:%95=21%:%
%:%96=21%:%
%:%97=22%:%
%:%98=22%:%
%:%99=23%:%
%:%100=23%:%
%:%105=23%:%
%:%108=24%:%
%:%109=25%:%
%:%110=25%:%
%:%111=26%:%
%:%114=27%:%
%:%118=27%:%
%:%119=27%:%
%:%120=28%:%
%:%121=28%:%
%:%122=29%:%
%:%123=29%:%
%:%128=29%:%
%:%131=30%:%
%:%132=31%:%
%:%133=31%:%
%:%134=32%:%
%:%137=33%:%
%:%141=33%:%
%:%142=33%:%
%:%143=34%:%
%:%144=34%:%
%:%145=35%:%
%:%146=35%:%
%:%151=35%:%
%:%154=36%:%
%:%155=37%:%
%:%156=37%:%
%:%157=38%:%
%:%160=39%:%
%:%164=39%:%
%:%165=39%:%
%:%166=40%:%
%:%167=40%:%
%:%168=41%:%
%:%169=41%:%
%:%174=41%:%
%:%177=42%:%
%:%178=43%:%
%:%179=43%:%
%:%180=44%:%
%:%183=45%:%
%:%187=45%:%
%:%188=45%:%
%:%189=46%:%
%:%190=46%:%
%:%191=47%:%
%:%192=47%:%
%:%197=47%:%
%:%200=48%:%
%:%201=49%:%
%:%202=49%:%
%:%203=50%:%
%:%206=51%:%
%:%210=51%:%
%:%211=51%:%
%:%212=52%:%
%:%213=52%:%
%:%214=53%:%
%:%215=53%:%
%:%220=53%:%
%:%223=54%:%
%:%224=55%:%
%:%225=56%:%
%:%226=56%:%
%:%227=57%:%
%:%230=58%:%
%:%234=58%:%
%:%235=58%:%
%:%236=59%:%
%:%237=59%:%
%:%238=60%:%
%:%239=60%:%
%:%244=60%:%
%:%247=61%:%
%:%248=62%:%
%:%249=62%:%
%:%250=63%:%
%:%251=64%:%
%:%254=65%:%
%:%258=65%:%
%:%259=65%:%
%:%260=66%:%
%:%261=66%:%
%:%262=67%:%
%:%263=67%:%
%:%268=67%:%
%:%271=68%:%
%:%272=69%:%
%:%273=69%:%
%:%274=70%:%
%:%277=71%:%
%:%281=71%:%
%:%282=71%:%
%:%283=72%:%
%:%284=72%:%
%:%285=73%:%
%:%286=74%:%
%:%287=74%:%
%:%292=74%:%
%:%295=75%:%
%:%296=76%:%
%:%297=76%:%
%:%298=77%:%
%:%301=78%:%
%:%305=78%:%
%:%306=78%:%
%:%307=79%:%
%:%308=79%:%
%:%309=80%:%
%:%310=80%:%
%:%315=80%:%
%:%318=81%:%
%:%319=82%:%
%:%320=82%:%
%:%321=83%:%
%:%324=84%:%
%:%328=84%:%
%:%329=84%:%
%:%330=85%:%
%:%331=85%:%
%:%332=86%:%
%:%333=86%:%
%:%338=86%:%
%:%341=87%:%
%:%342=88%:%
%:%343=89%:%
%:%344=89%:%
%:%345=90%:%
%:%348=91%:%
%:%352=91%:%
%:%353=91%:%
%:%354=92%:%
%:%355=92%:%
%:%356=93%:%
%:%357=93%:%
%:%362=93%:%
%:%365=94%:%
%:%366=95%:%
%:%367=96%:%
%:%368=96%:%
%:%369=97%:%
%:%372=98%:%
%:%376=98%:%
%:%377=98%:%
%:%378=99%:%
%:%379=99%:%
%:%380=100%:%
%:%381=100%:%
%:%386=100%:%
%:%389=101%:%
%:%390=102%:%
%:%391=102%:%
%:%392=103%:%
%:%395=104%:%
%:%399=104%:%
%:%400=104%:%
%:%401=105%:%
%:%402=105%:%
%:%403=106%:%
%:%404=106%:%
%:%409=106%:%
%:%412=107%:%
%:%413=108%:%
%:%414=108%:%
%:%415=109%:%
%:%418=110%:%
%:%422=110%:%
%:%423=110%:%
%:%424=111%:%
%:%425=111%:%
%:%426=112%:%
%:%427=112%:%
%:%432=112%:%
%:%435=113%:%
%:%436=114%:%
%:%437=114%:%
%:%438=115%:%
%:%439=116%:%
%:%442=117%:%
%:%446=117%:%
%:%447=117%:%
%:%448=118%:%
%:%449=118%:%
%:%450=119%:%
%:%451=120%:%
%:%452=120%:%
%:%457=120%:%
%:%460=121%:%
%:%461=122%:%
%:%462=123%:%
%:%463=123%:%
%:%464=124%:%
%:%467=125%:%
%:%471=125%:%
%:%472=125%:%
%:%473=126%:%
%:%474=126%:%
%:%475=127%:%
%:%476=127%:%
%:%481=127%:%
%:%484=128%:%
%:%485=129%:%
%:%486=129%:%
%:%487=130%:%
%:%490=131%:%
%:%494=131%:%
%:%495=131%:%
%:%496=132%:%
%:%497=132%:%
%:%498=133%:%
%:%499=133%:%
%:%504=133%:%
%:%507=134%:%
%:%508=135%:%
%:%509=135%:%
%:%510=136%:%
%:%513=137%:%
%:%517=137%:%
%:%518=137%:%
%:%519=138%:%
%:%520=138%:%
%:%521=139%:%
%:%522=139%:%
%:%527=139%:%
%:%530=140%:%
%:%531=141%:%
%:%532=141%:%
%:%533=142%:%
%:%536=143%:%
%:%540=143%:%
%:%541=143%:%
%:%542=144%:%
%:%543=144%:%
%:%544=145%:%
%:%545=145%:%
%:%550=145%:%
%:%553=146%:%
%:%554=147%:%
%:%555=147%:%
%:%556=148%:%
%:%559=149%:%
%:%563=149%:%
%:%564=149%:%
%:%565=150%:%
%:%566=150%:%
%:%567=151%:%
%:%568=152%:%
%:%569=152%:%
%:%574=152%:%
%:%577=153%:%
%:%578=154%:%
%:%579=154%:%
%:%580=155%:%
%:%583=156%:%
%:%587=156%:%
%:%588=156%:%
%:%589=157%:%
%:%590=157%:%
%:%591=158%:%
%:%592=158%:%
%:%597=158%:%
%:%600=159%:%
%:%601=160%:%
%:%602=160%:%
%:%603=161%:%
%:%606=162%:%
%:%610=162%:%
%:%611=162%:%
%:%612=163%:%
%:%613=163%:%
%:%614=164%:%
%:%615=164%:%
%:%620=164%:%
%:%623=165%:%
%:%624=166%:%
%:%625=166%:%
%:%626=167%:%
%:%629=168%:%
%:%633=168%:%
%:%634=168%:%
%:%635=169%:%
%:%636=169%:%
%:%637=170%:%
%:%638=170%:%
%:%643=170%:%
%:%646=171%:%
%:%647=172%:%
%:%648=172%:%
%:%649=173%:%
%:%652=174%:%
%:%656=174%:%
%:%657=174%:%
%:%658=175%:%
%:%659=175%:%
%:%660=176%:%
%:%661=176%:%
%:%666=176%:%
%:%669=177%:%
%:%670=178%:%
%:%671=178%:%
%:%672=179%:%
%:%675=180%:%
%:%679=180%:%
%:%680=180%:%
%:%681=181%:%
%:%682=181%:%
%:%683=182%:%
%:%684=182%:%
%:%689=182%:%
%:%692=183%:%
%:%693=184%:%
%:%694=184%:%
%:%695=185%:%
%:%698=186%:%
%:%702=186%:%
%:%703=186%:%
%:%704=187%:%
%:%705=187%:%
%:%706=188%:%
%:%707=188%:%
%:%712=188%:%
%:%715=189%:%
%:%716=190%:%
%:%717=190%:%
%:%718=191%:%
%:%721=192%:%
%:%725=192%:%
%:%726=192%:%
%:%727=193%:%
%:%728=193%:%
%:%729=194%:%
%:%730=194%:%
%:%735=194%:%
%:%738=195%:%
%:%739=196%:%
%:%740=196%:%
%:%741=197%:%
%:%744=198%:%
%:%748=198%:%
%:%749=198%:%
%:%750=199%:%
%:%751=199%:%
%:%752=200%:%
%:%753=200%:%
%:%758=200%:%
%:%761=201%:%
%:%762=202%:%
%:%763=202%:%
%:%764=203%:%
%:%767=204%:%
%:%771=204%:%
%:%772=204%:%
%:%773=205%:%
%:%774=205%:%
%:%775=206%:%
%:%776=206%:%
%:%781=206%:%
%:%784=207%:%
%:%785=208%:%
%:%786=208%:%
%:%787=209%:%
%:%790=210%:%
%:%794=210%:%
%:%795=210%:%
%:%796=211%:%
%:%797=211%:%
%:%798=212%:%
%:%799=212%:%
%:%804=212%:%
%:%807=213%:%
%:%808=214%:%
%:%809=214%:%
%:%810=215%:%
%:%813=216%:%
%:%817=216%:%
%:%818=216%:%
%:%819=217%:%
%:%820=217%:%
%:%821=218%:%
%:%822=218%:%
%:%827=218%:%
%:%830=219%:%
%:%831=220%:%
%:%832=220%:%
%:%833=221%:%
%:%836=222%:%
%:%840=222%:%
%:%841=222%:%
%:%842=223%:%
%:%843=223%:%
%:%844=224%:%
%:%845=225%:%
%:%846=225%:%
%:%851=225%:%
%:%854=226%:%
%:%855=227%:%
%:%856=227%:%
%:%857=228%:%
%:%858=229%:%
%:%861=230%:%
%:%865=230%:%
%:%866=230%:%
%:%867=231%:%
%:%868=231%:%
%:%869=232%:%
%:%870=233%:%
%:%871=233%:%
%:%876=233%:%
%:%879=234%:%
%:%880=235%:%
%:%881=235%:%
%:%882=236%:%
%:%883=237%:%
%:%886=238%:%
%:%890=238%:%
%:%891=238%:%
%:%892=239%:%
%:%893=239%:%
%:%894=240%:%
%:%895=241%:%
%:%896=241%:%
%:%901=241%:%
%:%904=242%:%
%:%905=243%:%
%:%906=243%:%
%:%907=244%:%
%:%908=245%:%
%:%911=246%:%
%:%915=246%:%
%:%916=246%:%
%:%917=247%:%
%:%918=247%:%
%:%919=248%:%
%:%920=249%:%
%:%921=249%:%
%:%926=249%:%
%:%929=250%:%
%:%930=251%:%
%:%931=251%:%
%:%932=252%:%
%:%933=253%:%
%:%934=254%:%
%:%935=255%:%
%:%942=256%:%
%:%943=256%:%
%:%944=257%:%
%:%945=257%:%
%:%946=258%:%
%:%947=258%:%
%:%948=259%:%
%:%949=259%:%
%:%950=260%:%
%:%951=260%:%
%:%952=261%:%
%:%953=261%:%
%:%954=262%:%
%:%955=263%:%
%:%956=263%:%
%:%957=264%:%
%:%958=264%:%
%:%959=265%:%
%:%960=265%:%
%:%961=266%:%
%:%962=266%:%
%:%963=267%:%
%:%964=267%:%
%:%965=268%:%
%:%966=268%:%
%:%967=269%:%
%:%973=269%:%
%:%976=270%:%
%:%977=271%:%
%:%978=271%:%
%:%979=272%:%
%:%980=273%:%
%:%981=274%:%
%:%982=275%:%
%:%989=276%:%
%:%990=276%:%
%:%991=277%:%
%:%992=277%:%
%:%993=278%:%
%:%994=278%:%
%:%995=279%:%
%:%996=279%:%
%:%997=280%:%
%:%998=280%:%
%:%999=281%:%
%:%1000=281%:%
%:%1001=282%:%
%:%1002=282%:%
%:%1003=283%:%
%:%1004=283%:%
%:%1005=284%:%
%:%1006=284%:%
%:%1007=285%:%
%:%1008=286%:%
%:%1009=286%:%
%:%1010=287%:%
%:%1011=287%:%
%:%1012=288%:%
%:%1013=288%:%
%:%1014=289%:%
%:%1015=289%:%
%:%1016=290%:%
%:%1017=290%:%
%:%1018=291%:%
%:%1019=291%:%
%:%1020=292%:%
%:%1021=292%:%
%:%1022=293%:%
%:%1028=293%:%
%:%1031=294%:%
%:%1032=295%:%
%:%1033=295%:%
%:%1034=296%:%
%:%1035=297%:%
%:%1042=298%:%
%:%1043=298%:%
%:%1044=299%:%
%:%1045=299%:%
%:%1046=300%:%
%:%1047=300%:%
%:%1048=301%:%
%:%1049=301%:%
%:%1050=302%:%
%:%1051=302%:%
%:%1052=303%:%
%:%1053=303%:%
%:%1054=304%:%
%:%1055=304%:%
%:%1056=305%:%
%:%1057=305%:%
%:%1058=306%:%
%:%1059=306%:%
%:%1060=307%:%
%:%1061=307%:%
%:%1062=308%:%
%:%1068=308%:%
%:%1071=309%:%
%:%1072=310%:%
%:%1073=310%:%
%:%1074=311%:%
%:%1075=312%:%
%:%1082=313%:%
%:%1083=313%:%
%:%1084=314%:%
%:%1085=314%:%
%:%1086=315%:%
%:%1087=315%:%
%:%1088=316%:%
%:%1089=316%:%
%:%1090=317%:%
%:%1091=317%:%
%:%1092=318%:%
%:%1093=318%:%
%:%1094=319%:%
%:%1095=319%:%
%:%1096=320%:%
%:%1097=320%:%
%:%1098=321%:%
%:%1099=321%:%
%:%1100=322%:%
%:%1101=322%:%
%:%1102=323%:%
%:%1103=323%:%
%:%1104=324%:%
%:%1110=324%:%
%:%1113=325%:%
%:%1114=326%:%
%:%1115=326%:%
%:%1116=327%:%
%:%1119=328%:%
%:%1123=328%:%
%:%1124=328%:%
%:%1125=329%:%
%:%1126=329%:%
%:%1127=330%:%
%:%1128=330%:%
%:%1133=330%:%
%:%1136=331%:%
%:%1137=332%:%
%:%1138=332%:%
%:%1139=333%:%
%:%1142=334%:%
%:%1146=334%:%
%:%1147=334%:%
%:%1148=335%:%
%:%1149=335%:%
%:%1150=336%:%
%:%1151=336%:%
%:%1156=336%:%
%:%1159=337%:%
%:%1160=338%:%
%:%1161=338%:%
%:%1162=339%:%
%:%1165=340%:%
%:%1169=340%:%
%:%1170=340%:%
%:%1171=341%:%
%:%1172=341%:%
%:%1173=342%:%
%:%1174=342%:%
%:%1179=342%:%
%:%1182=343%:%
%:%1183=344%:%
%:%1184=344%:%
%:%1185=345%:%
%:%1188=346%:%
%:%1192=346%:%
%:%1193=346%:%
%:%1194=347%:%
%:%1195=347%:%
%:%1196=348%:%
%:%1197=348%:%
%:%1202=348%:%
%:%1205=349%:%
%:%1206=350%:%
%:%1207=350%:%
%:%1208=351%:%
%:%1211=352%:%
%:%1215=352%:%
%:%1216=352%:%
%:%1217=353%:%
%:%1218=353%:%
%:%1219=354%:%
%:%1220=355%:%
%:%1221=355%:%
%:%1226=355%:%
%:%1229=356%:%
%:%1230=357%:%
%:%1231=357%:%
%:%1232=358%:%
%:%1233=359%:%
%:%1236=360%:%
%:%1240=360%:%
%:%1241=360%:%
%:%1242=361%:%
%:%1243=361%:%
%:%1244=362%:%
%:%1245=362%:%
%:%1250=362%:%
%:%1255=363%:%
%:%1260=364%:%