### 2024-11-01 code/output/document/Automorphism_Definition.tex

%
\begin{isabellebody}%
\setisabellecontext{Automorphism{\isacharunderscore}{\kern0pt}Definition}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Automorphism{\isacharunderscore}{\kern0pt}Definition\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\ \isanewline
\ \ \ \ P{\isacharunderscore}{\kern0pt}Names{\isacharunderscore}{\kern0pt}M\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{locale}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial\ {\isacharequal}{\kern0pt}\ \isanewline
\ \ forcing{\isacharunderscore}{\kern0pt}data\ {\isacharplus}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ leq{\isacharunderscore}{\kern0pt}relation{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}P\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}leq\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ leq{\isacharunderscore}{\kern0pt}partial{\isacharunderscore}{\kern0pt}order\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}partial{\isacharunderscore}{\kern0pt}order{\isacharunderscore}{\kern0pt}on{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}antisym\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}antisym{\isacharparenleft}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}partial{\isacharunderscore}{\kern0pt}order\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ partial{\isacharunderscore}{\kern0pt}order{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}unique{\isacharunderscore}{\kern0pt}max\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P\ {\isasymLongrightarrow}\ one\ {\isasympreceq}\ x\ {\isasymLongrightarrow}\ x\ {\isacharequal}{\kern0pt}\ one{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}one\ {\isasympreceq}\ x{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasympreceq}\ one{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}max\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ one{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ leq{\isacharunderscore}{\kern0pt}antisym\ \isacommand{unfolding}\isamarkupfalse%
\ antisym{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasympi}\ {\isasymin}\ M\ {\isasymand}\ {\isasympi}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ P{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}\ p\ {\isasymin}\ P{\isachardot}{\kern0pt}\ {\isasymforall}q\ {\isasymin}\ P{\isachardot}{\kern0pt}\ p\ {\isasympreceq}\ q\ {\isasymlongleftrightarrow}\ {\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasympreceq}\ {\isasympi}{\isacharbackquote}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto\ \isakeyword{where}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}auto\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isasympi}\ {\isasymin}\ P\ {\isasymrightarrow}\ P{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasympi}\ {\isasymin}\ P\ {\isasymrightarrow}\ P{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}function\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ P\ {\isasymrightarrow}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p\ {\isasymin}\ P\ {\isasymLongrightarrow}\ {\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ P\ {\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ P\ {\isasymrightarrow}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in\ {\isacartoucheopen}p\ {\isasymin}\ P{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}leq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p\ {\isasymin}\ P\ {\isasymLongrightarrow}\ q\ {\isasymin}\ P\ {\isasymLongrightarrow}\ p\ {\isasympreceq}\ q\ {\isasymLongrightarrow}\ {\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasympreceq}\ {\isasympi}{\isacharbackquote}{\kern0pt}q{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}leq{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p\ {\isasymin}\ P\ {\isasymLongrightarrow}\ q\ {\isasymin}\ P\ {\isasymLongrightarrow}\ {\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasympreceq}\ {\isasympi}{\isacharbackquote}{\kern0pt}q\ {\isasymLongrightarrow}\ p\ {\isasympreceq}\ q{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}idP\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}closed\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}bij\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ conjI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ converse{\isacharunderscore}{\kern0pt}closed\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ conjI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}bij\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ p\ q\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ p\ {\isasympreceq}\ q\ {\isasymlongleftrightarrow}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasympreceq}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ q{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ pisurj\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p{\isacharprime}{\kern0pt}\ \isakeyword{where}\ p{\isacharprime}{\kern0pt}h\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ q{\isacharprime}{\kern0pt}\ \isakeyword{where}\ q{\isacharprime}{\kern0pt}h\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}q{\isacharprime}{\kern0pt}\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}q{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ q{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ pisurj\ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isacharequal}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ converse{\isacharunderscore}{\kern0pt}apply\ assms\ p{\isacharprime}{\kern0pt}h\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ q\ {\isacharequal}{\kern0pt}\ q{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ converse{\isacharunderscore}{\kern0pt}apply\ assms\ q{\isacharprime}{\kern0pt}h\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasympreceq}\ q\ {\isasymlongleftrightarrow}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasympreceq}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ q{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}\ p\ {\isasympreceq}\ q{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasympreceq}\ q{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isacharprime}{\kern0pt}h\ q{\isacharprime}{\kern0pt}h\ assms\ assms{\isacharprime}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasympreceq}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ q{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H{\isadigit{1}}\ H{\isadigit{2}}\ H{\isadigit{3}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasympreceq}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ q\ {\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasympreceq}\ q{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isacharprime}{\kern0pt}h\ q{\isacharprime}{\kern0pt}h\ assms\ assms{\isacharprime}{\kern0pt}\ H{\isadigit{1}}\ H{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasympreceq}\ q{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ p{\isacharprime}{\kern0pt}h\ q{\isacharprime}{\kern0pt}h\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}one\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasympi}{\isacharbackquote}{\kern0pt}one\ {\isacharequal}{\kern0pt}\ one{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ pp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharequal}{\kern0pt}one{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ surj{\isacharunderscore}{\kern0pt}def\ assm\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasympreceq}\ one{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}max\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasympreceq}\ {\isasympi}{\isacharbackquote}{\kern0pt}one{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}leq\ pp\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ p{\isadigit{1}}\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}one\ {\isasympreceq}\ {\isasympi}{\isacharbackquote}{\kern0pt}one{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}one\ {\isacharequal}{\kern0pt}\ one{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assm\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ one{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}unique{\isacharunderscore}{\kern0pt}max\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}density\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ D\ {\isasymsubseteq}\ P\ {\isasymLongrightarrow}\ q\ {\isasymin}\ P\ {\isasymLongrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}D{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ p\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}P{\isachardot}{\kern0pt}\ p\ {\isasympreceq}\ q\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}d{\isasymin}D{\isachardot}{\kern0pt}\ d\ {\isasymin}\ P\ {\isasymand}\ d\ {\isasympreceq}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasympreceq}\ {\isasympi}{\isacharbackquote}{\kern0pt}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}D\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p{\isacharprime}{\kern0pt}\ \isakeyword{where}\ p{\isacharprime}{\kern0pt}h{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasympreceq}\ q{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ {\isasympi}{\isacharequal}{\kern0pt}{\isasympi}\ \isakeyword{in}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}leq{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ d\ \isakeyword{where}\ dh{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}d\ {\isasymin}\ D{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ {\isasympreceq}\ p{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isacharprime}{\kern0pt}h\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}d\ {\isasymin}\ {\isasympi}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}D{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}d\ \isakeyword{in}\ imageI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}function\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}d\ {\isasympreceq}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}leq\ dh\ p{\isacharprime}{\kern0pt}h\ assms\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}d\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ assms\ dh\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}d{\isasymin}{\isasympi}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ D{\isachardot}{\kern0pt}\ d\ {\isasymin}\ P\ {\isasymand}\ d\ {\isasympreceq}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ H{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}density{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ D\ {\isasymsubseteq}\ P\ {\isasymLongrightarrow}\ q\ {\isasymin}\ P\ {\isasymLongrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}D{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}density\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ D{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ {\isacharbackquote}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}D\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ D{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ image{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}image{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add\ {\isacharcolon}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}q\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ {\isacharbackquote}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ converse{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add\ {\isacharcolon}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}\ D\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}q\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isasympi}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ D{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ q{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ D{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ {\isacharbackquote}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}density{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ image{\isacharunderscore}{\kern0pt}subset{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ local{\isachardot}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ comp{\isacharunderscore}{\kern0pt}closed{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ comp{\isacharunderscore}{\kern0pt}bij{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharcomma}{\kern0pt}\ rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharcomma}{\kern0pt}\ rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ p\ q{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharbackquote}{\kern0pt}p\ {\isasympreceq}\ {\isasymtau}{\isacharbackquote}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ p\ q{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymtau}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ P\ {\isasymand}\ {\isasymtau}{\isacharbackquote}{\kern0pt}q\ {\isasymin}\ P{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}H{\isacharbackquote}{\kern0pt}fst{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}snd{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ v\ {\isasymin}\ x{\isacharcomma}{\kern0pt}\ {\isasymexists}y\ p{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ \ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}function\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ F\ \isakeyword{where}\ {\isachardoublequoteopen}F\ {\isasymequiv}\ {\isasymlambda}y\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ xin{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ E{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}value\ assm\ \ \isacommand{unfolding}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ E{\isadigit{2}}{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ F{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ wftrec{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ wf{\isacharunderscore}{\kern0pt}trancl{\isacharcomma}{\kern0pt}\ rule\ wf{\isacharunderscore}{\kern0pt}Memrel{\isacharcomma}{\kern0pt}\ rule\ trans{\isacharunderscore}{\kern0pt}trancl{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ E{\isadigit{3}}{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}F{\isacharbackquote}{\kern0pt}fst{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}snd{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ v\ {\isasymin}\ x{\isacharcomma}{\kern0pt}\ {\isasymexists}y\ p{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ E{\isadigit{4}}{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}F{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ \ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ w\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}F{\isacharbackquote}{\kern0pt}fst{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}snd{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ v\ {\isasymin}\ x{\isacharcomma}{\kern0pt}\ {\isasymexists}y\ p{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ v\ \isakeyword{where}\ {\isachardoublequoteopen}w\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}F{\isacharbackquote}{\kern0pt}fst{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}snd{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymexists}y\ p{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ {\isachardoublequoteopen}w\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}F{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}F\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}F\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}F\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assm\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}F\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}F{\isacharbackquote}{\kern0pt}fst{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}snd{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ v\ {\isasymin}\ x{\isacharcomma}{\kern0pt}\ {\isasymexists}y\ p{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ SepReplace{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ E{\isadigit{5}}{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ xin\ assm\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ allI{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ y\ p{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y{\isasymin}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ vimageI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assm\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ E{\isadigit{6}}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ xin\ assm\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ eq{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}value{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ assm\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ E{\isadigit{1}}\ E{\isadigit{2}}\ E{\isadigit{3}}\ E{\isadigit{4}}\ E{\isadigit{5}}\ E{\isadigit{6}}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Automorphism_Definition.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=11%:%
%:%30=12%:%
%:%31=13%:%
%:%32=13%:%
%:%35=14%:%
%:%39=14%:%
%:%40=14%:%
%:%41=15%:%
%:%42=15%:%
%:%43=16%:%
%:%44=16%:%
%:%49=16%:%
%:%52=17%:%
%:%53=18%:%
%:%54=18%:%
%:%61=19%:%
%:%62=19%:%
%:%63=20%:%
%:%64=20%:%
%:%65=21%:%
%:%66=21%:%
%:%67=21%:%
%:%68=21%:%
%:%69=21%:%
%:%70=22%:%
%:%71=22%:%
%:%72=22%:%
%:%73=22%:%
%:%74=22%:%
%:%75=22%:%
%:%76=23%:%
%:%82=23%:%
%:%85=24%:%
%:%86=25%:%
%:%87=25%:%
%:%88=26%:%
%:%89=27%:%
%:%90=28%:%
%:%91=28%:%
%:%92=29%:%
%:%93=30%:%
%:%94=30%:%
%:%96=30%:%
%:%100=30%:%
%:%101=30%:%
%:%102=30%:%
%:%109=30%:%
%:%110=31%:%
%:%111=32%:%
%:%112=32%:%
%:%115=33%:%
%:%119=33%:%
%:%120=33%:%
%:%121=34%:%
%:%122=34%:%
%:%123=34%:%
%:%128=34%:%
%:%131=35%:%
%:%132=36%:%
%:%133=36%:%
%:%140=37%:%
%:%141=37%:%
%:%142=38%:%
%:%143=38%:%
%:%144=39%:%
%:%145=39%:%
%:%146=39%:%
%:%147=39%:%
%:%148=39%:%
%:%149=39%:%
%:%150=40%:%
%:%151=40%:%
%:%152=40%:%
%:%153=40%:%
%:%154=40%:%
%:%155=41%:%
%:%161=41%:%
%:%164=42%:%
%:%165=43%:%
%:%166=43%:%
%:%173=44%:%
%:%174=44%:%
%:%175=45%:%
%:%176=45%:%
%:%177=46%:%
%:%178=46%:%
%:%179=46%:%
%:%180=47%:%
%:%181=47%:%
%:%182=47%:%
%:%183=47%:%
%:%184=48%:%
%:%185=48%:%
%:%186=48%:%
%:%187=48%:%
%:%188=48%:%
%:%189=49%:%
%:%195=49%:%
%:%198=50%:%
%:%199=51%:%
%:%200=51%:%
%:%201=52%:%
%:%204=53%:%
%:%208=53%:%
%:%209=53%:%
%:%210=53%:%
%:%215=53%:%
%:%218=54%:%
%:%219=55%:%
%:%220=55%:%
%:%221=56%:%
%:%224=57%:%
%:%228=57%:%
%:%229=57%:%
%:%230=57%:%
%:%235=57%:%
%:%238=58%:%
%:%239=59%:%
%:%240=59%:%
%:%243=60%:%
%:%247=60%:%
%:%248=60%:%
%:%249=61%:%
%:%250=61%:%
%:%251=61%:%
%:%252=61%:%
%:%253=61%:%
%:%254=62%:%
%:%255=62%:%
%:%256=62%:%
%:%257=62%:%
%:%263=62%:%
%:%266=63%:%
%:%267=64%:%
%:%268=64%:%
%:%271=65%:%
%:%275=65%:%
%:%276=65%:%
%:%277=65%:%
%:%278=65%:%
%:%279=65%:%
%:%280=66%:%
%:%281=66%:%
%:%282=66%:%
%:%283=66%:%
%:%284=66%:%
%:%285=67%:%
%:%286=67%:%
%:%287=68%:%
%:%288=68%:%
%:%289=69%:%
%:%290=69%:%
%:%291=70%:%
%:%292=70%:%
%:%293=70%:%
%:%294=70%:%
%:%295=70%:%
%:%296=71%:%
%:%297=71%:%
%:%298=71%:%
%:%299=71%:%
%:%300=71%:%
%:%301=71%:%
%:%302=72%:%
%:%303=72%:%
%:%304=72%:%
%:%305=72%:%
%:%306=72%:%
%:%307=72%:%
%:%308=73%:%
%:%309=73%:%
%:%310=73%:%
%:%311=73%:%
%:%312=74%:%
%:%313=74%:%
%:%314=74%:%
%:%315=74%:%
%:%316=75%:%
%:%317=75%:%
%:%318=76%:%
%:%319=76%:%
%:%320=77%:%
%:%321=77%:%
%:%322=78%:%
%:%323=78%:%
%:%324=79%:%
%:%325=79%:%
%:%326=79%:%
%:%327=79%:%
%:%328=80%:%
%:%329=80%:%
%:%330=80%:%
%:%331=80%:%
%:%332=81%:%
%:%333=81%:%
%:%334=82%:%
%:%335=82%:%
%:%336=83%:%
%:%337=83%:%
%:%338=83%:%
%:%339=83%:%
%:%340=84%:%
%:%341=84%:%
%:%342=84%:%
%:%343=84%:%
%:%344=84%:%
%:%345=85%:%
%:%346=85%:%
%:%347=86%:%
%:%353=86%:%
%:%356=87%:%
%:%357=88%:%
%:%358=88%:%
%:%365=89%:%
%:%366=89%:%
%:%367=90%:%
%:%368=90%:%
%:%369=91%:%
%:%370=91%:%
%:%371=92%:%
%:%372=92%:%
%:%373=92%:%
%:%374=92%:%
%:%375=93%:%
%:%376=93%:%
%:%377=93%:%
%:%378=93%:%
%:%379=93%:%
%:%380=94%:%
%:%381=94%:%
%:%382=94%:%
%:%383=95%:%
%:%384=95%:%
%:%385=95%:%
%:%386=96%:%
%:%387=96%:%
%:%388=96%:%
%:%389=96%:%
%:%390=96%:%
%:%391=97%:%
%:%392=97%:%
%:%393=97%:%
%:%394=97%:%
%:%395=97%:%
%:%396=98%:%
%:%402=98%:%
%:%405=99%:%
%:%406=100%:%
%:%407=100%:%
%:%408=101%:%
%:%411=102%:%
%:%415=102%:%
%:%416=102%:%
%:%417=102%:%
%:%418=103%:%
%:%419=103%:%
%:%420=104%:%
%:%421=104%:%
%:%422=105%:%
%:%423=105%:%
%:%424=106%:%
%:%425=107%:%
%:%426=107%:%
%:%427=107%:%
%:%428=107%:%
%:%429=107%:%
%:%430=108%:%
%:%431=108%:%
%:%432=108%:%
%:%433=108%:%
%:%434=108%:%
%:%435=108%:%
%:%436=109%:%
%:%437=109%:%
%:%438=109%:%
%:%439=109%:%
%:%440=109%:%
%:%441=109%:%
%:%442=109%:%
%:%443=110%:%
%:%444=110%:%
%:%445=110%:%
%:%446=110%:%
%:%447=110%:%
%:%448=111%:%
%:%449=111%:%
%:%450=111%:%
%:%451=111%:%
%:%452=112%:%
%:%453=112%:%
%:%454=113%:%
%:%455=113%:%
%:%456=113%:%
%:%457=113%:%
%:%458=114%:%
%:%459=114%:%
%:%460=114%:%
%:%461=114%:%
%:%462=114%:%
%:%463=115%:%
%:%464=115%:%
%:%465=115%:%
%:%466=115%:%
%:%467=116%:%
%:%468=116%:%
%:%469=116%:%
%:%470=116%:%
%:%471=116%:%
%:%472=117%:%
%:%478=117%:%
%:%481=118%:%
%:%482=119%:%
%:%483=119%:%
%:%484=120%:%
%:%487=121%:%
%:%491=121%:%
%:%492=121%:%
%:%493=121%:%
%:%494=121%:%
%:%495=122%:%
%:%496=122%:%
%:%497=123%:%
%:%498=123%:%
%:%499=124%:%
%:%500=124%:%
%:%501=124%:%
%:%502=125%:%
%:%503=125%:%
%:%504=126%:%
%:%505=126%:%
%:%506=127%:%
%:%507=127%:%
%:%508=127%:%
%:%509=128%:%
%:%510=128%:%
%:%511=129%:%
%:%512=129%:%
%:%513=130%:%
%:%514=130%:%
%:%515=131%:%
%:%516=131%:%
%:%517=132%:%
%:%518=132%:%
%:%519=132%:%
%:%520=132%:%
%:%521=133%:%
%:%522=133%:%
%:%523=133%:%
%:%524=133%:%
%:%525=133%:%
%:%526=134%:%
%:%527=134%:%
%:%528=134%:%
%:%529=135%:%
%:%535=135%:%
%:%538=136%:%
%:%539=137%:%
%:%540=137%:%
%:%543=138%:%
%:%547=138%:%
%:%548=138%:%
%:%549=139%:%
%:%550=139%:%
%:%551=140%:%
%:%552=140%:%
%:%553=141%:%
%:%554=141%:%
%:%555=142%:%
%:%556=142%:%
%:%557=143%:%
%:%558=143%:%
%:%559=144%:%
%:%560=144%:%
%:%561=145%:%
%:%562=145%:%
%:%563=146%:%
%:%564=146%:%
%:%565=147%:%
%:%566=147%:%
%:%567=148%:%
%:%568=148%:%
%:%569=149%:%
%:%570=149%:%
%:%571=150%:%
%:%577=150%:%
%:%580=151%:%
%:%581=152%:%
%:%582=152%:%
%:%583=153%:%
%:%584=154%:%
%:%585=155%:%
%:%586=155%:%
%:%587=156%:%
%:%588=157%:%
%:%589=158%:%
%:%590=159%:%
%:%591=159%:%
%:%594=160%:%
%:%598=160%:%
%:%599=160%:%
%:%600=160%:%
%:%605=160%:%
%:%608=161%:%
%:%609=161%:%
%:%612=162%:%
%:%616=162%:%
%:%617=162%:%
%:%618=162%:%
%:%623=162%:%
%:%626=163%:%
%:%627=164%:%
%:%628=164%:%
%:%635=165%:%
%:%636=165%:%
%:%637=166%:%
%:%638=166%:%
%:%639=167%:%
%:%640=168%:%
%:%641=168%:%
%:%642=169%:%
%:%643=169%:%
%:%644=169%:%
%:%645=169%:%
%:%646=169%:%
%:%647=170%:%
%:%648=170%:%
%:%649=170%:%
%:%650=171%:%
%:%651=171%:%
%:%652=171%:%
%:%653=171%:%
%:%654=172%:%
%:%655=172%:%
%:%656=173%:%
%:%657=174%:%
%:%658=174%:%
%:%659=175%:%
%:%660=175%:%
%:%661=176%:%
%:%662=176%:%
%:%663=177%:%
%:%664=177%:%
%:%665=178%:%
%:%666=178%:%
%:%667=179%:%
%:%668=179%:%
%:%669=179%:%
%:%670=180%:%
%:%671=180%:%
%:%672=181%:%
%:%673=182%:%
%:%674=182%:%
%:%675=183%:%
%:%676=183%:%
%:%677=183%:%
%:%678=184%:%
%:%679=184%:%
%:%680=184%:%
%:%681=184%:%
%:%682=185%:%
%:%683=185%:%
%:%684=185%:%
%:%685=185%:%
%:%686=186%:%
%:%687=186%:%
%:%688=186%:%
%:%689=187%:%
%:%690=187%:%
%:%691=188%:%
%:%692=188%:%
%:%693=189%:%
%:%694=189%:%
%:%695=190%:%
%:%696=190%:%
%:%697=191%:%
%:%698=191%:%
%:%699=191%:%
%:%700=192%:%
%:%701=192%:%
%:%702=192%:%
%:%703=193%:%
%:%704=193%:%
%:%705=194%:%
%:%706=194%:%
%:%707=195%:%
%:%708=195%:%
%:%709=196%:%
%:%710=196%:%
%:%711=196%:%
%:%712=196%:%
%:%713=197%:%
%:%714=197%:%
%:%715=197%:%
%:%716=198%:%
%:%717=198%:%
%:%718=199%:%
%:%719=199%:%
%:%720=200%:%
%:%721=200%:%
%:%722=201%:%
%:%723=201%:%
%:%724=202%:%
%:%725=202%:%
%:%726=203%:%
%:%727=204%:%
%:%728=204%:%
%:%729=205%:%
%:%730=205%:%
%:%731=206%:%
%:%732=206%:%
%:%733=207%:%
%:%734=207%:%
%:%735=208%:%
%:%736=208%:%
%:%737=209%:%
%:%738=209%:%
%:%739=210%:%
%:%740=210%:%
%:%741=211%:%
%:%742=211%:%
%:%743=212%:%
%:%744=212%:%
%:%745=213%:%
%:%746=213%:%
%:%747=214%:%
%:%748=214%:%
%:%749=215%:%
%:%750=215%:%
%:%751=216%:%
%:%752=216%:%
%:%753=217%:%
%:%754=218%:%
%:%755=218%:%
%:%756=219%:%
%:%757=219%:%
%:%758=220%:%
%:%759=220%:%
%:%760=221%:%
%:%761=221%:%
%:%762=222%:%
%:%763=222%:%
%:%764=223%:%
%:%765=223%:%
%:%766=224%:%
%:%767=224%:%
%:%768=225%:%
%:%769=225%:%
%:%770=226%:%
%:%771=226%:%
%:%772=227%:%
%:%773=227%:%
%:%774=228%:%
%:%775=228%:%
%:%776=229%:%
%:%782=229%:%
%:%785=230%:%
%:%786=231%:%
%:%787=231%:%
%:%794=232%:%