### 2024-11-01 code/output/document/Automorphism_M.tex

%
\begin{isabellebody}%
\setisabellecontext{Automorphism{\isacharunderscore}{\kern0pt}M}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Automorphism{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\ \isanewline
\ \ \ \ P{\isacharunderscore}{\kern0pt}Names{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ Automorphism{\isacharunderscore}{\kern0pt}Definition\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasympi}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ leq{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}bijection{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}q\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}p{\isacharunderscore}{\kern0pt}q\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}{\isasympi}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}{\isasympi}q\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}{\isasympi}p{\isacharunderscore}{\kern0pt}{\isasympi}q\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ p\ {\isasymin}\ P\ {\isasymlongrightarrow}\ q\ {\isasymin}\ P\ {\isasymlongrightarrow}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ {\isasympi}p{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharcomma}{\kern0pt}\ q{\isacharcomma}{\kern0pt}\ {\isasympi}q{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\isanewline
\ \ \ \ \ \ \ \ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ q{\isacharcomma}{\kern0pt}\ p{\isacharunderscore}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasympi}p{\isacharcomma}{\kern0pt}\ {\isasympi}q{\isacharcomma}{\kern0pt}\ {\isasympi}p{\isacharunderscore}{\kern0pt}{\isasympi}q{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}p{\isacharunderscore}{\kern0pt}q\ {\isasymin}\ leq\ \ {\isasymlongleftrightarrow}\ {\isasympi}p{\isacharunderscore}{\kern0pt}{\isasympi}q\ {\isasymin}\ leq{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}bijection{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Implies\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Implies\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Implies\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Implies\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Implies{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Implies{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \ {\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial\ \isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}bijection{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}q\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}p{\isacharunderscore}{\kern0pt}q\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}{\isasympi}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}{\isasympi}q\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}{\isasympi}p{\isacharunderscore}{\kern0pt}{\isasympi}q\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ p\ {\isasymin}\ P\ {\isasymlongrightarrow}\ q\ {\isasymin}\ P\ {\isasymlongrightarrow}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ {\isasympi}p{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharcomma}{\kern0pt}\ q{\isacharcomma}{\kern0pt}\ {\isasympi}q{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\isanewline
\ \ \ \ \ \ \ \ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ q{\isacharcomma}{\kern0pt}\ p{\isacharunderscore}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasympi}p{\isacharcomma}{\kern0pt}\ {\isasympi}q{\isacharcomma}{\kern0pt}\ {\isasympi}p{\isacharunderscore}{\kern0pt}{\isasympi}q{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}p{\isacharunderscore}{\kern0pt}q\ {\isasymin}\ leq\ \ {\isasymlongleftrightarrow}\ {\isasympi}p{\isacharunderscore}{\kern0pt}{\isasympi}q\ {\isasymin}\ leq{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ p\ q{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasympi}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasymin}\ M{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasympi}\ {\isacharbackquote}{\kern0pt}\ q\ {\isasymin}\ M{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}auto\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}{\isasympi}{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ bijection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ injection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ surjection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}{\isasympi}{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}{\isasympi}{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ {\isasympi}\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ separation{\isacharunderscore}{\kern0pt}notation{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ M{\isacharunderscore}{\kern0pt}powerset{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ cartprod{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}auto\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto\ \isakeyword{and}\ a\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ {\isasympi}\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharsemicolon}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}y{\isacharunderscore}{\kern0pt}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}y{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ {\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ y{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ y{\isacharunderscore}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ y{\isacharunderscore}{\kern0pt}p\ {\isasymin}\ x\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}function{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ pi{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ y{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ {\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ elem\ {\isasymin}\ M\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ x{\isacharunderscore}{\kern0pt}pi\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ elem\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}pi{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ pi{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ function{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ {\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{and}\ inM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}elem\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p\ y{\isacharunderscore}{\kern0pt}p\ y{\isacharunderscore}{\kern0pt}pi\ y\ p\ x\ pi\ \isakeyword{where}\ \isanewline
\ \ \ \ \ \ H\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isacharunderscore}{\kern0pt}p\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ y{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ y{\isacharunderscore}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}y{\isacharunderscore}{\kern0pt}p\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}function{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ y{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ elemeq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}elem\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ pair{\isacharunderscore}{\kern0pt}abs\ inM\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ x{\isacharunderscore}{\kern0pt}pi{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}{\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ pair{\isacharunderscore}{\kern0pt}abs\ inM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ y{\isacharunderscore}{\kern0pt}pi{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y{\isacharunderscore}{\kern0pt}pi\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ pair{\isacharunderscore}{\kern0pt}abs\ inM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ y{\isacharunderscore}{\kern0pt}p{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y{\isacharunderscore}{\kern0pt}p\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ pair{\isacharunderscore}{\kern0pt}abs\ inM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ xin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ x{\isacharunderscore}{\kern0pt}pi{\isacharunderscore}{\kern0pt}eq\ inM\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ yin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ y{\isacharunderscore}{\kern0pt}pi{\isacharunderscore}{\kern0pt}eq\ H\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ pin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ y{\isacharunderscore}{\kern0pt}pi{\isacharunderscore}{\kern0pt}eq\ H\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ piin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ y{\isacharunderscore}{\kern0pt}pi{\isacharunderscore}{\kern0pt}eq\ H\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ pifun\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}abs\ piin\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p{\isacharunderscore}{\kern0pt}val\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}pi{\isacharbackquote}{\kern0pt}p\ {\isacharequal}{\kern0pt}\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}abs\ piin\ pin\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}val\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}pi{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isachardoublequoteclose}\ \ \ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ y{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isacharequal}{\kern0pt}g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}y{\isacharunderscore}{\kern0pt}pi{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ apply{\isacharunderscore}{\kern0pt}abs{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ H\ inM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ x{\isacharunderscore}{\kern0pt}pi\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ elem\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}pi{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ pi{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ function{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ {\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}y\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}p\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}pi\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}val\ pi{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}p{\isacharunderscore}{\kern0pt}val\ H\ elemeq\ pifun\ piin\ xin\ pin\ yin\ y{\isacharunderscore}{\kern0pt}p{\isacharunderscore}{\kern0pt}eq\ H\ x{\isacharunderscore}{\kern0pt}pi{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}x{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}pi{\isasymin}M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ x{\isacharunderscore}{\kern0pt}pi\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}\ {\isasymand}\ elem\ {\isacharequal}{\kern0pt}\ {\isasymlangle}g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ pi{\isasymrangle}{\isacharcomma}{\kern0pt}\ pi\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}\ {\isasymand}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymand}\ function{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{and}\ inM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}elem\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ x\ pi\ \isakeyword{where}\ H\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}elem\ {\isacharequal}{\kern0pt}\ {\isasymlangle}g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ pi{\isasymrangle}{\isacharcomma}{\kern0pt}\ pi\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}pi{\isacharbackquote}{\kern0pt}p{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}pi{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}y{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}p{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}x{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}pi{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ inM\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ H\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ {\isasymexists}A\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymforall}\ elem\ {\isasymin}\ M{\isachardot}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ elem\ {\isasymin}\ A{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ elem\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ inM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}elem\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ gfun\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ cond\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ x{\isacharunderscore}{\kern0pt}pi\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ elem\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}pi{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ pi{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ function{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ {\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ x\ pi\ \isakeyword{where}\ H\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}elem\ {\isacharequal}{\kern0pt}\ {\isasymlangle}g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ pi{\isasymrangle}{\isacharcomma}{\kern0pt}\ pi\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \isacommand{have}\isamarkupfalse%
\ P{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}\ {\isasymin}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ F{\isacharequal}{\kern0pt}g\ \isakeyword{and}\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}or{\isacharunderscore}{\kern0pt}not{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ gfun\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MVset\ Ord{\isacharunderscore}{\kern0pt}rank\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ pi{\isasymrangle}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ w\ \isakeyword{where}\ wh\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ w{\isachargreater}{\kern0pt}\ {\isasymin}\ g{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ w{\isachargreater}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ transM\ inM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ winM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}w\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ weq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ w{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ wh\ gfun\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}pi{\isachargreater}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ winM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}\ {\isasymin}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ MVsetI{\isacharsemicolon}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ lt{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ w{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ weq\ rank{\isacharunderscore}{\kern0pt}pair{\isadigit{2}}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ wh\ rank{\isacharunderscore}{\kern0pt}lt\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}pi\ {\isacharbackquote}{\kern0pt}\ p\ {\isasymin}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}or{\isacharunderscore}{\kern0pt}not{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MVset\ Ord{\isacharunderscore}{\kern0pt}rank\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ w\ \isakeyword{where}\ wh\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}p{\isacharcomma}{\kern0pt}\ w{\isachargreater}{\kern0pt}\ {\isasymin}\ pi{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ weq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}pi{\isacharbackquote}{\kern0pt}p\ {\isacharequal}{\kern0pt}\ w{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}p{\isacharcomma}{\kern0pt}\ w{\isachargreater}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ transM\ wh\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}pi\ {\isacharbackquote}{\kern0pt}\ p\ {\isasymin}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ MVsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ weq\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ lt{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}w\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ weq\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}p{\isacharcomma}{\kern0pt}\ w{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ rank{\isacharunderscore}{\kern0pt}pair{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}lt\ wh\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ rank{\isacharunderscore}{\kern0pt}pair{\isadigit{2}}\ \isanewline
\ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}elem\ {\isasymin}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isadigit{1}}\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ cartprod{\isacharunderscore}{\kern0pt}closed{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ MVset{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}closed\ Ord{\isacharunderscore}{\kern0pt}rank\ succ{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MI\ Ord{\isacharunderscore}{\kern0pt}succ\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ MVset{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}closed\ Ord{\isacharunderscore}{\kern0pt}rank\ succ{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MI\ Ord{\isacharunderscore}{\kern0pt}succ\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ MVset{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ elem{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharunderscore}{\kern0pt}pi{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm\ {\isacharequal}{\kern0pt}\ \ Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{9}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And\isanewline
{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial\ \isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ z{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharunderscore}{\kern0pt}pi{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}elem\ {\isasymin}\ M{\isachardot}{\kern0pt}\ elem\ {\isasymin}\ z\ {\isasymlongleftrightarrow}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ And{\isacharparenleft}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isanewline
\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial\ \isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ elem\ {\isasymin}\ M{\isachardot}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}elem\ {\isasymin}\ M\ {\isasymLongrightarrow}\ x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ \isanewline
\ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}elem{\isacharcomma}{\kern0pt}\ x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ z\ {\isasymin}\ M\ {\isasymLongrightarrow}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \isanewline
\ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}z{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ x{\isacharunderscore}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ z\ {\isacharequal}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}elem\ {\isasymin}\ M{\isachardot}{\kern0pt}\ elem\ {\isasymin}\ z\ {\isasymlongleftrightarrow}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ transM\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymAnd}x{\isacharunderscore}{\kern0pt}pi\ g{\isachardot}{\kern0pt}\ x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x{\isacharunderscore}{\kern0pt}pi\ g\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isacharunderscore}{\kern0pt}pi\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}A{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymforall}elem{\isasymin}M{\isachardot}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ elem\ {\isasymin}\ A{\isachardoublequoteclose}\ \isanewline
\ \ \ \isacommand{using}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ A\ \isakeyword{where}\ Ah{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}\ elem\ {\isasymin}\ M{\isachardot}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ elem\ {\isasymin}\ A{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ S\ \isakeyword{where}\ {\isachardoublequoteopen}S\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ elem\ {\isasymin}\ A{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}elem{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}elem\ {\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}elem{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add\ {\isacharcolon}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add\ {\isacharcolon}{\kern0pt}\ assms{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ sinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}S\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ S{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ separation{\isacharunderscore}{\kern0pt}notation{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ah\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}S\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ elem\ {\isasymin}\ A{\isachardot}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ assms\ Ah\ transM\ Ah\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ S{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ \ elem\ {\isasymin}\ M{\isachardot}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ elem{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharsemicolon}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transM\ Ah\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}S\ {\isacharequal}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharunderscore}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sinM\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isasymAnd}h\ g\ x\ {\isasympi}{\isachardot}{\kern0pt}\ {\isasympi}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto\ {\isasymLongrightarrow}\ h\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}{\isasympi}{\isacharbraceright}{\kern0pt}\ {\isasymrightarrow}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ h{\isacharbackquote}{\kern0pt}y\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ h\ g\ x\ {\isasympi}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isachardoublequoteclose}\ {\isachardoublequoteopen}h\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}{\isasympi}{\isacharbraceright}{\kern0pt}\ {\isasymrightarrow}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ h\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ piinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms{\isadigit{1}}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ u\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}u\ {\isasymin}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ueq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}h{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ ypin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ypin\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ hyeq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}h{\isacharbackquote}{\kern0pt}y\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ yinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ ypin\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ pinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ transM{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ ypin\ range{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ uinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ ueq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ hyeq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ apply{\isacharunderscore}{\kern0pt}closed{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ yinM\ piinM\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ apply{\isacharunderscore}{\kern0pt}closed{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ piinM\ pinM\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}h{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ piinM\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ subst\ hyeq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ apply{\isacharunderscore}{\kern0pt}closed{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ yinM\ piinM\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ apply{\isacharunderscore}{\kern0pt}closed{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ piinM\ pinM\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}y\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}p\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isasympi}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ hyeq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ypin\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ piinM\ yinM\ pinM\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}u\ {\isasymin}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ueq\ uinM\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ u\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}u\ {\isasymin}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}{\isasymlangle}x{\isacharcomma}{\kern0pt}\ {\isasympi}{\isasymrangle}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ uH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ u{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}xx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}xx{\isacharcomma}{\kern0pt}\ pi{\isachargreater}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ u\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}pi{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ pi{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ xx\ {\isasymand}\ \isanewline
\ \ \ \ \ \ function{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ {\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms{\isadigit{1}}\ piinM\ uH\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ueq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ ypin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}u\ {\isasymin}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ And{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial\ \isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ p\ pi\ x\ v\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}pi\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ p\ pi\ x\ v\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}pi\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ pi{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}pi{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ x\ {\isasympi}\ v\ env\ i\ j\ k\ l\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasympi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}l{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ piinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isasympi}\ \isakeyword{and}\ G{\isacharequal}{\kern0pt}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M\ \isakeyword{in}\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ G{\isacharequal}{\kern0pt}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M\ \isakeyword{and}\ Gfm\ {\isacharequal}{\kern0pt}\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isasympi}\ \isakeyword{in}\ memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ HPn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Automorphism_M.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%16=6%:%
%:%21=6%:%
%:%24=7%:%
%:%25=8%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=11%:%
%:%30=12%:%
%:%31=13%:%
%:%32=14%:%
%:%33=15%:%
%:%37=19%:%
%:%40=20%:%
%:%44=20%:%
%:%45=20%:%
%:%50=20%:%
%:%53=21%:%
%:%54=22%:%
%:%55=22%:%
%:%56=23%:%
%:%72=39%:%
%:%73=40%:%
%:%74=41%:%
%:%75=41%:%
%:%76=42%:%
%:%77=43%:%
%:%78=44%:%
%:%79=44%:%
%:%80=45%:%
%:%83=46%:%
%:%84=47%:%
%:%88=47%:%
%:%89=47%:%
%:%90=48%:%
%:%91=48%:%
%:%93=50%:%
%:%94=51%:%
%:%95=51%:%
%:%96=52%:%
%:%97=52%:%
%:%98=53%:%
%:%99=53%:%
%:%100=54%:%
%:%101=54%:%
%:%102=55%:%
%:%103=55%:%
%:%104=56%:%
%:%105=56%:%
%:%106=57%:%
%:%107=57%:%
%:%108=58%:%
%:%109=58%:%
%:%110=59%:%
%:%111=59%:%
%:%112=60%:%
%:%113=60%:%
%:%114=61%:%
%:%115=61%:%
%:%116=62%:%
%:%117=62%:%
%:%118=63%:%
%:%119=63%:%
%:%120=64%:%
%:%121=64%:%
%:%122=65%:%
%:%123=65%:%
%:%128=65%:%
%:%131=66%:%
%:%132=67%:%
%:%133=67%:%
%:%140=68%:%
%:%141=68%:%
%:%142=69%:%
%:%143=69%:%
%:%144=70%:%
%:%145=70%:%
%:%146=71%:%
%:%147=71%:%
%:%148=72%:%
%:%149=72%:%
%:%150=73%:%
%:%151=73%:%
%:%152=74%:%
%:%153=74%:%
%:%154=75%:%
%:%155=75%:%
%:%156=76%:%
%:%157=76%:%
%:%158=76%:%
%:%159=77%:%
%:%160=77%:%
%:%161=78%:%
%:%162=78%:%
%:%163=79%:%
%:%164=79%:%
%:%165=80%:%
%:%166=80%:%
%:%167=81%:%
%:%168=81%:%
%:%169=81%:%
%:%170=82%:%
%:%171=82%:%
%:%172=83%:%
%:%173=83%:%
%:%174=84%:%
%:%175=84%:%
%:%176=85%:%
%:%177=85%:%
%:%178=86%:%
%:%179=86%:%
%:%180=87%:%
%:%181=87%:%
%:%182=87%:%
%:%183=88%:%
%:%184=88%:%
%:%185=89%:%
%:%186=89%:%
%:%187=90%:%
%:%188=90%:%
%:%189=91%:%
%:%190=91%:%
%:%191=92%:%
%:%192=92%:%
%:%193=93%:%
%:%194=93%:%
%:%195=94%:%
%:%196=94%:%
%:%197=95%:%
%:%198=95%:%
%:%199=96%:%
%:%200=96%:%
%:%201=97%:%
%:%202=97%:%
%:%203=98%:%
%:%209=98%:%
%:%212=99%:%
%:%213=100%:%
%:%214=100%:%
%:%215=101%:%
%:%226=112%:%
%:%227=113%:%
%:%228=114%:%
%:%229=114%:%
%:%230=115%:%
%:%237=122%:%
%:%244=123%:%
%:%245=123%:%
%:%246=124%:%
%:%247=124%:%
%:%248=125%:%
%:%249=126%:%
%:%250=126%:%
%:%251=126%:%
%:%252=127%:%
%:%253=128%:%
%:%254=129%:%
%:%255=130%:%
%:%256=131%:%
%:%257=132%:%
%:%258=133%:%
%:%259=134%:%
%:%260=135%:%
%:%261=136%:%
%:%262=137%:%
%:%263=137%:%
%:%264=137%:%
%:%265=138%:%
%:%266=138%:%
%:%267=138%:%
%:%268=138%:%
%:%269=139%:%
%:%270=139%:%
%:%271=139%:%
%:%272=139%:%
%:%273=140%:%
%:%274=140%:%
%:%275=140%:%
%:%276=140%:%
%:%277=141%:%
%:%278=141%:%
%:%279=141%:%
%:%280=141%:%
%:%281=142%:%
%:%282=142%:%
%:%283=142%:%
%:%284=142%:%
%:%285=143%:%
%:%286=143%:%
%:%287=143%:%
%:%288=143%:%
%:%289=144%:%
%:%290=144%:%
%:%291=144%:%
%:%292=144%:%
%:%293=145%:%
%:%294=145%:%
%:%295=145%:%
%:%296=145%:%
%:%297=146%:%
%:%298=146%:%
%:%299=146%:%
%:%300=146%:%
%:%301=147%:%
%:%302=147%:%
%:%303=147%:%
%:%304=147%:%
%:%305=148%:%
%:%306=148%:%
%:%307=149%:%
%:%308=149%:%
%:%309=150%:%
%:%310=150%:%
%:%311=151%:%
%:%312=151%:%
%:%313=151%:%
%:%314=151%:%
%:%315=152%:%
%:%316=152%:%
%:%321=157%:%
%:%322=158%:%
%:%323=158%:%
%:%324=158%:%
%:%325=159%:%
%:%326=159%:%
%:%327=159%:%
%:%328=160%:%
%:%329=160%:%
%:%330=160%:%
%:%331=161%:%
%:%332=161%:%
%:%333=162%:%
%:%334=162%:%
%:%335=163%:%
%:%336=164%:%
%:%337=165%:%
%:%338=166%:%
%:%339=166%:%
%:%340=166%:%
%:%341=167%:%
%:%342=168%:%
%:%343=169%:%
%:%344=169%:%
%:%345=170%:%
%:%346=171%:%
%:%347=171%:%
%:%348=172%:%
%:%349=172%:%
%:%350=173%:%
%:%351=173%:%
%:%352=174%:%
%:%353=174%:%
%:%354=175%:%
%:%355=175%:%
%:%356=176%:%
%:%357=176%:%
%:%358=177%:%
%:%359=177%:%
%:%360=178%:%
%:%361=178%:%
%:%362=179%:%
%:%363=179%:%
%:%364=180%:%
%:%365=180%:%
%:%366=181%:%
%:%367=181%:%
%:%368=181%:%
%:%369=182%:%
%:%375=182%:%
%:%378=183%:%
%:%379=184%:%
%:%380=184%:%
%:%381=185%:%
%:%382=186%:%
%:%385=187%:%
%:%389=187%:%
%:%390=187%:%
%:%391=188%:%
%:%392=188%:%
%:%393=189%:%
%:%394=189%:%
%:%395=190%:%
%:%396=190%:%
%:%397=191%:%
%:%398=192%:%
%:%399=193%:%
%:%400=194%:%
%:%401=194%:%
%:%402=194%:%
%:%407=199%:%
%:%408=199%:%
%:%409=199%:%
%:%410=200%:%
%:%411=200%:%
%:%412=200%:%
%:%413=201%:%
%:%414=202%:%
%:%415=203%:%
%:%416=203%:%
%:%417=204%:%
%:%418=205%:%
%:%419=205%:%
%:%420=206%:%
%:%421=206%:%
%:%422=207%:%
%:%423=207%:%
%:%424=208%:%
%:%425=208%:%
%:%426=209%:%
%:%427=209%:%
%:%428=210%:%
%:%429=210%:%
%:%430=211%:%
%:%431=211%:%
%:%432=212%:%
%:%433=212%:%
%:%434=213%:%
%:%435=213%:%
%:%436=213%:%
%:%437=213%:%
%:%438=214%:%
%:%439=214%:%
%:%440=214%:%
%:%441=214%:%
%:%442=214%:%
%:%443=215%:%
%:%444=215%:%
%:%445=215%:%
%:%446=215%:%
%:%447=215%:%
%:%448=216%:%
%:%449=216%:%
%:%450=216%:%
%:%451=216%:%
%:%452=216%:%
%:%453=217%:%
%:%454=217%:%
%:%455=217%:%
%:%456=217%:%
%:%457=217%:%
%:%458=218%:%
%:%459=218%:%
%:%460=218%:%
%:%461=219%:%
%:%462=219%:%
%:%463=220%:%
%:%464=220%:%
%:%465=221%:%
%:%466=221%:%
%:%467=222%:%
%:%468=222%:%
%:%469=223%:%
%:%470=223%:%
%:%471=224%:%
%:%472=224%:%
%:%473=225%:%
%:%474=225%:%
%:%475=226%:%
%:%476=226%:%
%:%477=226%:%
%:%478=227%:%
%:%479=227%:%
%:%480=228%:%
%:%481=229%:%
%:%482=229%:%
%:%483=230%:%
%:%484=230%:%
%:%485=231%:%
%:%486=231%:%
%:%487=232%:%
%:%488=232%:%
%:%489=233%:%
%:%490=233%:%
%:%491=234%:%
%:%492=234%:%
%:%493=235%:%
%:%494=235%:%
%:%495=236%:%
%:%496=236%:%
%:%497=237%:%
%:%498=237%:%
%:%499=237%:%
%:%500=237%:%
%:%501=238%:%
%:%502=238%:%
%:%503=238%:%
%:%504=238%:%
%:%505=238%:%
%:%506=239%:%
%:%507=239%:%
%:%508=239%:%
%:%509=239%:%
%:%510=240%:%
%:%511=240%:%
%:%512=240%:%
%:%513=240%:%
%:%514=240%:%
%:%515=241%:%
%:%516=241%:%
%:%517=241%:%
%:%518=242%:%
%:%519=242%:%
%:%520=243%:%
%:%521=243%:%
%:%522=244%:%
%:%523=244%:%
%:%524=245%:%
%:%525=245%:%
%:%526=246%:%
%:%527=246%:%
%:%528=247%:%
%:%529=247%:%
%:%530=248%:%
%:%531=248%:%
%:%532=249%:%
%:%533=249%:%
%:%534=250%:%
%:%535=250%:%
%:%536=251%:%
%:%537=251%:%
%:%538=252%:%
%:%539=252%:%
%:%540=253%:%
%:%541=253%:%
%:%542=254%:%
%:%543=254%:%
%:%544=255%:%
%:%545=255%:%
%:%546=256%:%
%:%547=256%:%
%:%548=257%:%
%:%549=257%:%
%:%550=258%:%
%:%551=258%:%
%:%552=259%:%
%:%553=260%:%
%:%554=260%:%
%:%555=260%:%
%:%556=261%:%
%:%557=261%:%
%:%558=261%:%
%:%559=262%:%
%:%560=262%:%
%:%561=263%:%
%:%562=263%:%
%:%563=264%:%
%:%564=264%:%
%:%565=264%:%
%:%566=265%:%
%:%567=265%:%
%:%568=266%:%
%:%569=266%:%
%:%570=267%:%
%:%571=267%:%
%:%572=268%:%
%:%573=268%:%
%:%574=269%:%
%:%575=269%:%
%:%576=270%:%
%:%577=270%:%
%:%578=271%:%
%:%579=271%:%
%:%580=272%:%
%:%581=272%:%
%:%582=273%:%
%:%583=273%:%
%:%584=274%:%
%:%585=274%:%
%:%586=274%:%
%:%587=274%:%
%:%588=275%:%
%:%594=275%:%
%:%597=276%:%
%:%598=277%:%
%:%599=277%:%
%:%600=278%:%
%:%601=279%:%
%:%602=280%:%
%:%603=281%:%
%:%604=282%:%
%:%605=283%:%
%:%606=284%:%
%:%607=285%:%
%:%610=286%:%
%:%614=286%:%
%:%615=286%:%
%:%616=287%:%
%:%617=287%:%
%:%622=287%:%
%:%625=288%:%
%:%626=289%:%
%:%627=289%:%
%:%628=290%:%
%:%629=291%:%
%:%630=291%:%
%:%631=292%:%
%:%647=308%:%
%:%648=309%:%
%:%649=310%:%
%:%650=310%:%
%:%651=311%:%
%:%652=312%:%
%:%653=313%:%
%:%654=313%:%
%:%655=314%:%
%:%656=315%:%
%:%657=316%:%
%:%658=317%:%
%:%659=318%:%
%:%660=319%:%
%:%661=320%:%
%:%662=321%:%
%:%665=322%:%
%:%669=322%:%
%:%670=322%:%
%:%671=323%:%
%:%672=323%:%
%:%677=323%:%
%:%680=324%:%
%:%681=325%:%
%:%682=325%:%
%:%683=326%:%
%:%684=327%:%
%:%685=327%:%
%:%686=328%:%
%:%702=344%:%
%:%703=345%:%
%:%704=346%:%
%:%705=347%:%
%:%706=347%:%
%:%707=348%:%
%:%708=349%:%
%:%709=350%:%
%:%710=350%:%
%:%711=351%:%
%:%712=352%:%
%:%713=353%:%
%:%714=353%:%
%:%715=354%:%
%:%716=355%:%
%:%719=356%:%
%:%723=356%:%
%:%724=356%:%
%:%725=357%:%
%:%726=357%:%
%:%727=358%:%
%:%728=358%:%
%:%729=359%:%
%:%730=359%:%
%:%735=359%:%
%:%738=360%:%
%:%739=361%:%
%:%740=361%:%
%:%741=362%:%
%:%742=363%:%
%:%745=364%:%
%:%749=364%:%
%:%750=364%:%
%:%751=365%:%
%:%752=365%:%
%:%753=366%:%
%:%754=366%:%
%:%755=367%:%
%:%756=367%:%
%:%757=368%:%
%:%758=368%:%
%:%759=369%:%
%:%760=369%:%
%:%761=370%:%
%:%762=370%:%
%:%763=371%:%
%:%764=371%:%
%:%769=371%:%
%:%772=372%:%
%:%773=373%:%
%:%774=373%:%
%:%775=374%:%
%:%782=375%:%
%:%783=375%:%
%:%784=376%:%
%:%785=376%:%
%:%786=377%:%
%:%787=377%:%
%:%788=378%:%
%:%789=379%:%
%:%790=379%:%
%:%791=379%:%
%:%792=380%:%
%:%793=380%:%
%:%794=380%:%
%:%795=381%:%
%:%796=381%:%
%:%797=381%:%
%:%798=381%:%
%:%799=382%:%
%:%800=383%:%
%:%801=383%:%
%:%802=384%:%
%:%803=385%:%
%:%804=385%:%
%:%805=386%:%
%:%806=386%:%
%:%807=387%:%
%:%808=387%:%
%:%809=388%:%
%:%810=388%:%
%:%811=389%:%
%:%812=389%:%
%:%813=390%:%
%:%814=390%:%
%:%815=391%:%
%:%816=391%:%
%:%817=392%:%
%:%818=392%:%
%:%819=392%:%
%:%820=393%:%
%:%821=393%:%
%:%822=394%:%
%:%823=394%:%
%:%824=395%:%
%:%825=395%:%
%:%826=396%:%
%:%827=396%:%
%:%828=397%:%
%:%829=398%:%
%:%830=398%:%
%:%831=399%:%
%:%832=399%:%
%:%833=400%:%
%:%834=400%:%
%:%835=401%:%
%:%836=401%:%
%:%837=402%:%
%:%838=402%:%
%:%839=402%:%
%:%840=403%:%
%:%841=403%:%
%:%842=404%:%
%:%843=404%:%
%:%844=405%:%
%:%845=405%:%
%:%846=406%:%
%:%847=406%:%
%:%848=406%:%
%:%849=407%:%
%:%850=407%:%
%:%851=407%:%
%:%852=408%:%
%:%853=408%:%
%:%854=408%:%
%:%855=408%:%
%:%856=409%:%
%:%857=410%:%
%:%858=410%:%
%:%859=410%:%
%:%860=411%:%
%:%861=411%:%
%:%862=411%:%
%:%863=412%:%
%:%869=412%:%
%:%872=413%:%
%:%873=414%:%
%:%874=414%:%
%:%875=415%:%
%:%882=416%:%
%:%883=416%:%
%:%884=417%:%
%:%885=417%:%
%:%886=418%:%
%:%887=418%:%
%:%888=419%:%
%:%889=420%:%
%:%890=421%:%
%:%891=421%:%
%:%892=422%:%
%:%893=422%:%
%:%894=423%:%
%:%895=423%:%
%:%896=424%:%
%:%897=425%:%
%:%898=425%:%
%:%899=426%:%
%:%900=426%:%
%:%901=427%:%
%:%902=427%:%
%:%903=428%:%
%:%904=428%:%
%:%905=429%:%
%:%906=430%:%
%:%907=430%:%
%:%908=431%:%
%:%909=431%:%
%:%910=432%:%
%:%911=432%:%
%:%912=432%:%
%:%913=433%:%
%:%914=433%:%
%:%915=433%:%
%:%916=433%:%
%:%917=433%:%
%:%918=434%:%
%:%919=434%:%
%:%920=435%:%
%:%921=435%:%
%:%922=436%:%
%:%923=436%:%
%:%924=437%:%
%:%925=437%:%
%:%926=438%:%
%:%927=438%:%
%:%928=438%:%
%:%929=438%:%
%:%930=438%:%
%:%931=439%:%
%:%932=440%:%
%:%933=440%:%
%:%934=441%:%
%:%935=441%:%
%:%936=442%:%
%:%937=442%:%
%:%938=443%:%
%:%939=443%:%
%:%940=444%:%
%:%941=445%:%
%:%942=445%:%
%:%943=446%:%
%:%944=446%:%
%:%945=447%:%
%:%946=447%:%
%:%947=448%:%
%:%948=448%:%
%:%949=449%:%
%:%950=450%:%
%:%951=450%:%
%:%952=451%:%
%:%953=451%:%
%:%954=452%:%
%:%955=452%:%
%:%956=453%:%
%:%957=453%:%
%:%958=454%:%
%:%959=454%:%
%:%960=455%:%
%:%961=455%:%
%:%962=456%:%
%:%963=456%:%
%:%964=457%:%
%:%965=457%:%
%:%966=458%:%
%:%967=458%:%
%:%968=459%:%
%:%969=459%:%
%:%970=460%:%
%:%971=461%:%
%:%972=461%:%
%:%973=462%:%
%:%974=462%:%
%:%975=463%:%
%:%976=463%:%
%:%977=464%:%
%:%978=464%:%
%:%979=465%:%
%:%980=465%:%
%:%981=466%:%
%:%982=466%:%
%:%983=467%:%
%:%984=467%:%
%:%985=468%:%
%:%986=468%:%
%:%987=469%:%
%:%988=469%:%
%:%989=470%:%
%:%990=470%:%
%:%991=471%:%
%:%992=471%:%
%:%993=472%:%
%:%994=472%:%
%:%995=473%:%
%:%996=473%:%
%:%997=474%:%
%:%998=474%:%
%:%999=475%:%
%:%1000=475%:%
%:%1001=476%:%
%:%1002=476%:%
%:%1003=477%:%
%:%1004=477%:%
%:%1005=478%:%
%:%1006=478%:%
%:%1007=479%:%
%:%1008=480%:%
%:%1009=480%:%
%:%1010=480%:%
%:%1011=481%:%
%:%1012=481%:%
%:%1013=482%:%
%:%1014=482%:%
%:%1015=483%:%
%:%1016=483%:%
%:%1017=484%:%
%:%1018=484%:%
%:%1019=485%:%
%:%1020=485%:%
%:%1021=485%:%
%:%1022=486%:%
%:%1023=486%:%
%:%1024=486%:%
%:%1025=486%:%
%:%1026=486%:%
%:%1027=487%:%
%:%1028=487%:%
%:%1033=492%:%
%:%1034=493%:%
%:%1035=493%:%
%:%1036=494%:%
%:%1037=494%:%
%:%1038=495%:%
%:%1039=495%:%
%:%1040=496%:%
%:%1041=496%:%
%:%1042=496%:%
%:%1043=496%:%
%:%1044=497%:%
%:%1045=497%:%
%:%1046=497%:%
%:%1047=498%:%
%:%1048=498%:%
%:%1049=499%:%
%:%1050=499%:%
%:%1051=500%:%
%:%1052=500%:%
%:%1053=501%:%
%:%1054=501%:%
%:%1055=502%:%
%:%1056=502%:%
%:%1057=503%:%
%:%1058=503%:%
%:%1059=504%:%
%:%1060=504%:%
%:%1061=505%:%
%:%1062=505%:%
%:%1063=506%:%
%:%1064=506%:%
%:%1065=507%:%
%:%1071=507%:%
%:%1074=508%:%
%:%1075=509%:%
%:%1076=509%:%
%:%1077=510%:%
%:%1078=511%:%
%:%1079=511%:%
%:%1080=512%:%
%:%1081=513%:%
%:%1082=514%:%
%:%1083=514%:%
%:%1084=515%:%
%:%1085=516%:%
%:%1086=517%:%
%:%1087=517%:%
%:%1088=518%:%
%:%1089=519%:%
%:%1090=520%:%
%:%1093=521%:%
%:%1097=521%:%
%:%1098=521%:%
%:%1099=522%:%
%:%1100=522%:%
%:%1101=523%:%
%:%1102=523%:%
%:%1103=524%:%
%:%1104=524%:%
%:%1105=525%:%
%:%1106=525%:%
%:%1107=526%:%
%:%1108=526%:%
%:%1113=526%:%
%:%1116=527%:%
%:%1117=528%:%
%:%1118=528%:%
%:%1119=529%:%
%:%1120=530%:%
%:%1121=531%:%
%:%1124=532%:%
%:%1125=533%:%
%:%1129=533%:%
%:%1130=533%:%
%:%1131=534%:%
%:%1132=534%:%
%:%1133=535%:%
%:%1134=535%:%
%:%1135=536%:%
%:%1136=536%:%
%:%1137=537%:%
%:%1138=537%:%
%:%1139=538%:%
%:%1140=538%:%
%:%1141=539%:%
%:%1142=539%:%
%:%1143=540%:%
%:%1144=540%:%
%:%1145=541%:%
%:%1146=541%:%
%:%1147=542%:%
%:%1148=542%:%
%:%1149=543%:%
%:%1150=543%:%
%:%1151=544%:%
%:%1152=544%:%
%:%1153=545%:%
%:%1154=545%:%
%:%1155=546%:%
%:%1156=546%:%
%:%1157=547%:%
%:%1158=547%:%
%:%1159=548%:%
%:%1160=548%:%
%:%1161=549%:%
%:%1167=549%:%
%:%1170=550%:%
%:%1171=551%:%
%:%1172=551%:%
%:%1173=552%:%
%:%1174=553%:%
%:%1175=554%:%
%:%1176=555%:%
%:%1177=556%:%
%:%1184=557%:%
%:%1185=557%:%
%:%1186=558%:%
%:%1187=559%:%
%:%1188=559%:%
%:%1189=560%:%
%:%1190=560%:%
%:%1191=561%:%
%:%1192=561%:%
%:%1193=562%:%
%:%1194=563%:%
%:%1195=563%:%
%:%1196=564%:%
%:%1197=564%:%
%:%1198=565%:%
%:%1199=565%:%
%:%1200=566%:%
%:%1201=566%:%
%:%1202=567%:%
%:%1203=567%:%
%:%1204=568%:%
%:%1205=568%:%
%:%1206=569%:%
%:%1207=569%:%
%:%1208=570%:%
%:%1209=570%:%
%:%1210=571%:%
%:%1211=571%:%
%:%1212=572%:%
%:%1213=572%:%
%:%1214=573%:%
%:%1215=573%:%
%:%1216=574%:%
%:%1217=574%:%
%:%1218=575%:%
%:%1219=575%:%
%:%1220=576%:%
%:%1221=576%:%
%:%1222=577%:%
%:%1223=577%:%
%:%1224=578%:%
%:%1225=578%:%
%:%1226=579%:%
%:%1227=579%:%
%:%1228=580%:%
%:%1229=580%:%
%:%1230=581%:%
%:%1231=581%:%
%:%1232=582%:%
%:%1233=582%:%
%:%1234=583%:%
%:%1235=583%:%
%:%1236=584%:%
%:%1237=584%:%
%:%1238=585%:%
%:%1239=585%:%
%:%1240=585%:%
%:%1241=586%:%
%:%1242=586%:%
%:%1243=587%:%
%:%1244=587%:%
%:%1245=588%:%
%:%1246=588%:%
%:%1247=589%:%
%:%1248=589%:%
%:%1249=590%:%
%:%1250=590%:%
%:%1251=591%:%
%:%1252=591%:%
%:%1253=591%:%
%:%1254=591%:%
%:%1255=592%:%
%:%1261=592%:%
%:%1264=593%:%
%:%1265=594%:%
%:%1266=594%:%
%:%1267=595%:%
%:%1270=596%:%
%:%1271=597%:%
%:%1275=597%:%
%:%1276=597%:%
%:%1277=598%:%
%:%1278=598%:%
%:%1279=599%:%
%:%1280=599%:%
%:%1281=600%:%
%:%1282=600%:%
%:%1283=601%:%
%:%1284=601%:%
%:%1285=602%:%
%:%1286=602%:%
%:%1287=603%:%
%:%1288=603%:%
%:%1289=604%:%
%:%1290=604%:%
%:%1291=605%:%
%:%1292=605%:%
%:%1293=606%:%
%:%1294=606%:%
%:%1295=607%:%
%:%1296=607%:%
%:%1297=608%:%
%:%1298=608%:%
%:%1299=609%:%
%:%1300=609%:%
%:%1301=610%:%
%:%1302=610%:%
%:%1303=611%:%
%:%1304=611%:%
%:%1305=612%:%
%:%1306=612%:%
%:%1307=613%:%
%:%1308=613%:%
%:%1309=614%:%
%:%1310=614%:%
%:%1311=615%:%
%:%1312=615%:%
%:%1313=616%:%
%:%1314=616%:%
%:%1315=617%:%
%:%1321=617%:%
%:%1324=618%:%
%:%1325=619%:%
%:%1326=619%:%
%:%1333=620%:%