### 2024-11-01 code/output/document/Automorphism_Theorems.tex

%
\begin{isabellebody}%
\setisabellecontext{Automorphism{\isacharunderscore}{\kern0pt}Theorems}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Automorphism{\isacharunderscore}{\kern0pt}Theorems\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\ \isanewline
\ \ \ \ Automorphism{\isacharunderscore}{\kern0pt}M\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial\ \isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ Q\ \isakeyword{where}\ {\isachardoublequoteopen}Q\ {\isasymequiv}\ {\isasymlambda}x\ a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ trans{\isacharunderscore}{\kern0pt}induct{\isadigit{3}}{\isacharunderscore}{\kern0pt}raw{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Q{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add\ {\isacharcolon}{\kern0pt}\ Q{\isacharunderscore}{\kern0pt}def\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ b\ x\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ bord\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isakeyword{and}\ ih\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isakeyword{and}\ xpname\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ xinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ pinP\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardot}{\kern0pt}\ p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}forallI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ ih\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}y\ a{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ih\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bord\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xinM\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardot}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}forallI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ blast\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{and}\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}y\ p{\isachardot}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ pair{\isacharunderscore}{\kern0pt}forallE{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ xrel\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ yph{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ yph\ pinP\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yph\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}forallI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{and}\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}y\ p{\isachardot}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ pair{\isacharunderscore}{\kern0pt}forallE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ y\ p{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ ih{\isacharparenright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ blast\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ bord\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}forallI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{and}\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}y\ p{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ pair{\isacharunderscore}{\kern0pt}forallE{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ ih{\isacharparenright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ blast\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ bord\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{and}\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}y\ p{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ pair{\isacharunderscore}{\kern0pt}forallE{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ {\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinP\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ subsetI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\ \ \ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assm{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg\ assm{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ yph\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yph\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pinP\ yph\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ \ {\isachardoublequoteopen}v\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ yph\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\ \ \ \ \ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xpname\ Pn{\isacharunderscore}{\kern0pt}auto\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ \ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}x{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}forallI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ xpname\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ p\ \isacommand{assume}\isamarkupfalse%
\ ypin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ xpname\ Pn{\isacharunderscore}{\kern0pt}auto\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ b\ x\ \isacommand{assume}\isamarkupfalse%
\ blim\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ ih\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}a{\isasymin}b{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ psetb\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}a{\isacharless}{\kern0pt}b{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}lim\ blim\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ impI{\isacharsemicolon}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ ah\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isacommand{using}\isamarkupfalse%
\ psetb\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ih\ ltD\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}a{\isacharless}{\kern0pt}b{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ah\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ psetb\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ ah\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isacommand{using}\isamarkupfalse%
\ psetb\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ih\ ltD\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}a{\isacharless}{\kern0pt}b{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ah\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ psetb\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ main\ \isacommand{unfolding}\isamarkupfalse%
\ Q{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ ah\ {\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ ah\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymrightarrow}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}function\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ abH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ abH\ assm\ \isacommand{unfolding}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality\ abH\ assm\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}function\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ H\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymtimes}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ abH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ xpname\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ ih{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ ih{\isacharparenright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{using}\isamarkupfalse%
\ xpname\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P{\isacharunderscore}{\kern0pt}names\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P{\isacharunderscore}{\kern0pt}names\ \isakeyword{in}\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}range\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharsemicolon}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ yph\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isasymrangle}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}relI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasymtau}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ b\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ xpname\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}relI{\isacharparenright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \ \isacommand{using}\isamarkupfalse%
\ yph\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharprime}{\kern0pt}{\isasymrangle}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y{\isacharprime}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ xpname\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y{\isacharprime}{\kern0pt}\ p{\isacharprime}{\kern0pt}\ \isakeyword{where}\ y{\isacharprime}{\kern0pt}p{\isacharprime}{\kern0pt}H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasymtau}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ assms\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ {\isacharless}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasymtau}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymtau}{\isacharbackquote}{\kern0pt}p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}relI{\isacharparenright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ y{\isacharprime}{\kern0pt}p{\isacharprime}{\kern0pt}H\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{thm}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ eq{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ assms\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P{\isacharunderscore}{\kern0pt}names\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P{\isacharunderscore}{\kern0pt}names\ \isakeyword{in}\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ xpname\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ conjI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ comp{\isacharunderscore}{\kern0pt}closed\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ conjI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ comp{\isacharunderscore}{\kern0pt}bij\ \isacommand{apply}\isamarkupfalse%
\ blast\ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharbackquote}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasympreceq}\ {\isasympi}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharbackquote}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharbackquote}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasympreceq}\ {\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharbackquote}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymtau}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ P\ {\isasymand}\ {\isasymtau}{\isacharbackquote}{\kern0pt}q\ {\isasymin}\ P{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}\ p\ {\isacharequal}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}\ q\ {\isacharequal}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ {\isacharbackquote}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ comp{\isacharunderscore}{\kern0pt}type{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymrightarrow}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ range{\isacharunderscore}{\kern0pt}subset{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ domain{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ xH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ v{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ veq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}function{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ xH\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ veq\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{thm}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}eq\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ comp{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}comp{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ comp{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ comp{\isacharunderscore}{\kern0pt}function{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}function{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ domain{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ rule\ range{\isacharunderscore}{\kern0pt}subset{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ mp{\isacharcomma}{\kern0pt}\ rule\ main{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ \ x{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ ih\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ xpname\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ih\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ xpname\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{unfolding}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ xpname\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}range\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \ \ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharsemicolon}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}relI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}relation{\isacharunderscore}{\kern0pt}def\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ id{\isacharunderscore}{\kern0pt}def\ relation{\isacharunderscore}{\kern0pt}lam\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}function{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}id{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ rule\ function{\isacharunderscore}{\kern0pt}lam{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}id{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ mp{\isacharcomma}{\kern0pt}\ rule\ main{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}bij\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ g{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ fg{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}bijective{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasympi}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ right{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ {\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ {\isasympi}{\isachardoublequoteclose}\ \isakeyword{and}\ a\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ left{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ right{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ eq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P{\isacharunderscore}{\kern0pt}names\ \isakeyword{in}\ left{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}id{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ Pi{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ left{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}bij\ assms\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ comp{\isacharunderscore}{\kern0pt}assoc\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ \ \ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ right{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ right{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}id{\isacharcomma}{\kern0pt}\ rule\ converse{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ Pi{\isacharunderscore}{\kern0pt}def\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q{\isachardot}{\kern0pt}\ {\isasymforall}x{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}lemma\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q\ R{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}P\ {\isasymLongrightarrow}\ Q\ {\isasymlongleftrightarrow}\ R{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ P\ {\isasymand}\ Q\ {\isasymlongleftrightarrow}\ P\ {\isasymand}\ R{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ least{\isacharunderscore}{\kern0pt}iff{\isacharsemicolon}{\kern0pt}\ rule\ allI{\isacharsemicolon}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ iff{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}fixpoint\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ piauto{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mainlemma{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}v{\isachardot}{\kern0pt}\ v\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isasymforall}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{thm}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}induct\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}v{\isachardot}{\kern0pt}\ {\isasymforall}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}y{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isasymin}M{\isachardot}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ def{\isacharunderscore}{\kern0pt}check{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}one{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ assms\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ p\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isacharequal}{\kern0pt}\ one{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isacharequal}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}one{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ piauto\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}one\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ \ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ assms\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ p\ \isacommand{assume}\isamarkupfalse%
\ assm{\isadigit{1}}{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ yy\ \isakeyword{where}\ p{\isadigit{0}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}yy\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}yy{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}yy{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}yy\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ transM\ assms\ p{\isadigit{0}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ assm{\isadigit{1}}\ p{\isadigit{1}}\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
{\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ RepFun{\isacharunderscore}{\kern0pt}eq{\isacharsemicolon}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ yp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}\ y\ {\isasymin}\ x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymlambda}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ one{\isasymrangle}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}v\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlangle}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymlambda}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ one{\isasymrangle}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mainlemma\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Automorphism_Theorems.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=10%:%
%:%30=11%:%
%:%37=12%:%
%:%38=12%:%
%:%39=13%:%
%:%40=13%:%
%:%41=14%:%
%:%42=14%:%
%:%43=15%:%
%:%44=15%:%
%:%45=16%:%
%:%46=17%:%
%:%47=17%:%
%:%48=18%:%
%:%49=18%:%
%:%50=19%:%
%:%51=19%:%
%:%52=20%:%
%:%53=20%:%
%:%54=21%:%
%:%55=21%:%
%:%56=21%:%
%:%57=22%:%
%:%58=22%:%
%:%59=23%:%
%:%60=23%:%
%:%61=24%:%
%:%62=24%:%
%:%63=25%:%
%:%64=26%:%
%:%65=27%:%
%:%66=28%:%
%:%67=28%:%
%:%68=28%:%
%:%69=28%:%
%:%70=29%:%
%:%71=30%:%
%:%72=30%:%
%:%73=31%:%
%:%74=31%:%
%:%75=32%:%
%:%76=32%:%
%:%77=32%:%
%:%78=33%:%
%:%79=34%:%
%:%80=34%:%
%:%81=35%:%
%:%82=35%:%
%:%83=35%:%
%:%84=36%:%
%:%85=37%:%
%:%86=37%:%
%:%87=37%:%
%:%88=37%:%
%:%89=38%:%
%:%90=38%:%
%:%91=38%:%
%:%92=38%:%
%:%93=38%:%
%:%94=39%:%
%:%95=39%:%
%:%96=39%:%
%:%97=40%:%
%:%98=40%:%
%:%99=41%:%
%:%100=41%:%
%:%101=41%:%
%:%102=41%:%
%:%103=41%:%
%:%104=42%:%
%:%105=42%:%
%:%106=43%:%
%:%107=43%:%
%:%108=43%:%
%:%109=43%:%
%:%110=44%:%
%:%111=44%:%
%:%112=45%:%
%:%113=45%:%
%:%114=46%:%
%:%115=46%:%
%:%116=46%:%
%:%117=47%:%
%:%118=47%:%
%:%119=47%:%
%:%120=47%:%
%:%121=48%:%
%:%122=48%:%
%:%123=48%:%
%:%124=48%:%
%:%125=48%:%
%:%126=48%:%
%:%127=49%:%
%:%128=49%:%
%:%129=49%:%
%:%130=49%:%
%:%131=49%:%
%:%132=50%:%
%:%133=50%:%
%:%134=50%:%
%:%135=50%:%
%:%136=51%:%
%:%137=51%:%
%:%138=51%:%
%:%139=51%:%
%:%140=51%:%
%:%141=52%:%
%:%142=52%:%
%:%143=53%:%
%:%144=53%:%
%:%145=53%:%
%:%146=54%:%
%:%147=54%:%
%:%148=55%:%
%:%149=55%:%
%:%150=56%:%
%:%151=56%:%
%:%152=56%:%
%:%153=57%:%
%:%154=57%:%
%:%155=58%:%
%:%156=58%:%
%:%157=58%:%
%:%158=58%:%
%:%159=59%:%
%:%160=59%:%
%:%161=60%:%
%:%162=60%:%
%:%163=61%:%
%:%164=61%:%
%:%165=61%:%
%:%166=61%:%
%:%167=62%:%
%:%168=62%:%
%:%169=62%:%
%:%170=62%:%
%:%171=62%:%
%:%172=63%:%
%:%173=63%:%
%:%174=64%:%
%:%175=64%:%
%:%176=65%:%
%:%177=65%:%
%:%178=65%:%
%:%179=66%:%
%:%180=66%:%
%:%181=67%:%
%:%182=67%:%
%:%183=67%:%
%:%184=68%:%
%:%185=68%:%
%:%186=69%:%
%:%187=69%:%
%:%188=70%:%
%:%189=70%:%
%:%190=70%:%
%:%191=70%:%
%:%192=71%:%
%:%193=71%:%
%:%194=71%:%
%:%195=71%:%
%:%196=71%:%
%:%197=72%:%
%:%198=72%:%
%:%199=73%:%
%:%200=73%:%
%:%201=74%:%
%:%202=74%:%
%:%203=74%:%
%:%204=75%:%
%:%205=75%:%
%:%206=76%:%
%:%207=76%:%
%:%208=77%:%
%:%209=77%:%
%:%210=77%:%
%:%211=77%:%
%:%212=78%:%
%:%213=78%:%
%:%214=79%:%
%:%215=79%:%
%:%216=80%:%
%:%217=80%:%
%:%218=81%:%
%:%219=81%:%
%:%220=81%:%
%:%221=82%:%
%:%222=82%:%
%:%223=82%:%
%:%224=82%:%
%:%225=82%:%
%:%226=83%:%
%:%227=83%:%
%:%228=83%:%
%:%229=84%:%
%:%230=84%:%
%:%231=85%:%
%:%232=85%:%
%:%233=86%:%
%:%234=86%:%
%:%235=87%:%
%:%236=87%:%
%:%237=88%:%
%:%238=88%:%
%:%239=88%:%
%:%240=88%:%
%:%241=89%:%
%:%242=89%:%
%:%243=89%:%
%:%244=90%:%
%:%245=90%:%
%:%246=90%:%
%:%247=91%:%
%:%248=91%:%
%:%249=91%:%
%:%250=91%:%
%:%251=92%:%
%:%252=92%:%
%:%253=92%:%
%:%254=92%:%
%:%255=92%:%
%:%256=93%:%
%:%257=93%:%
%:%258=93%:%
%:%259=93%:%
%:%260=94%:%
%:%261=94%:%
%:%262=94%:%
%:%263=94%:%
%:%264=94%:%
%:%265=95%:%
%:%266=95%:%
%:%267=95%:%
%:%268=95%:%
%:%269=95%:%
%:%270=96%:%
%:%271=96%:%
%:%272=97%:%
%:%273=97%:%
%:%274=97%:%
%:%275=97%:%
%:%276=97%:%
%:%277=98%:%
%:%278=98%:%
%:%279=98%:%
%:%280=98%:%
%:%281=99%:%
%:%282=99%:%
%:%283=99%:%
%:%284=99%:%
%:%285=99%:%
%:%286=100%:%
%:%287=100%:%
%:%288=100%:%
%:%289=100%:%
%:%290=100%:%
%:%291=101%:%
%:%292=101%:%
%:%293=102%:%
%:%294=102%:%
%:%295=103%:%
%:%296=103%:%
%:%297=104%:%
%:%298=104%:%
%:%299=105%:%
%:%300=105%:%
%:%301=105%:%
%:%302=106%:%
%:%303=106%:%
%:%304=107%:%
%:%305=107%:%
%:%306=107%:%
%:%307=108%:%
%:%308=108%:%
%:%309=108%:%
%:%310=109%:%
%:%311=109%:%
%:%312=109%:%
%:%313=109%:%
%:%314=110%:%
%:%315=110%:%
%:%316=110%:%
%:%317=111%:%
%:%318=111%:%
%:%319=111%:%
%:%320=112%:%
%:%321=113%:%
%:%322=113%:%
%:%323=113%:%
%:%324=113%:%
%:%325=114%:%
%:%326=114%:%
%:%327=114%:%
%:%328=114%:%
%:%329=114%:%
%:%330=115%:%
%:%331=115%:%
%:%332=115%:%
%:%333=115%:%
%:%334=115%:%
%:%335=116%:%
%:%336=116%:%
%:%337=117%:%
%:%338=117%:%
%:%339=118%:%
%:%340=118%:%
%:%341=118%:%
%:%342=118%:%
%:%343=119%:%
%:%344=119%:%
%:%345=120%:%
%:%346=120%:%
%:%347=120%:%
%:%348=121%:%
%:%349=122%:%
%:%350=123%:%
%:%351=123%:%
%:%352=123%:%
%:%353=123%:%
%:%354=124%:%
%:%355=125%:%
%:%356=125%:%
%:%357=126%:%
%:%358=126%:%
%:%359=127%:%
%:%360=127%:%
%:%361=128%:%
%:%362=128%:%
%:%363=128%:%
%:%364=128%:%
%:%365=128%:%
%:%366=129%:%
%:%367=129%:%
%:%368=129%:%
%:%369=129%:%
%:%370=129%:%
%:%371=130%:%
%:%372=130%:%
%:%373=130%:%
%:%374=130%:%
%:%375=130%:%
%:%376=131%:%
%:%377=131%:%
%:%378=131%:%
%:%379=131%:%
%:%380=131%:%
%:%381=132%:%
%:%382=132%:%
%:%383=133%:%
%:%384=133%:%
%:%385=134%:%
%:%386=134%:%
%:%387=134%:%
%:%388=134%:%
%:%389=134%:%
%:%390=135%:%
%:%391=135%:%
%:%392=135%:%
%:%393=135%:%
%:%394=135%:%
%:%395=136%:%
%:%396=136%:%
%:%397=136%:%
%:%398=136%:%
%:%399=136%:%
%:%400=137%:%
%:%401=137%:%
%:%402=137%:%
%:%403=137%:%
%:%404=137%:%
%:%405=138%:%
%:%406=138%:%
%:%407=139%:%
%:%408=139%:%
%:%409=140%:%
%:%410=140%:%
%:%411=140%:%
%:%412=141%:%
%:%413=141%:%
%:%414=141%:%
%:%415=141%:%
%:%416=142%:%
%:%422=142%:%
%:%425=143%:%
%:%426=144%:%
%:%427=144%:%
%:%434=145%:%
%:%435=145%:%
%:%436=146%:%
%:%437=146%:%
%:%438=147%:%
%:%439=147%:%
%:%440=147%:%
%:%441=147%:%
%:%442=147%:%
%:%443=148%:%
%:%444=148%:%
%:%445=148%:%
%:%446=148%:%
%:%447=148%:%
%:%448=149%:%
%:%449=149%:%
%:%450=149%:%
%:%451=150%:%
%:%452=150%:%
%:%453=150%:%
%:%454=150%:%
%:%455=151%:%
%:%461=151%:%
%:%464=152%:%
%:%465=153%:%
%:%466=153%:%
%:%469=154%:%
%:%473=154%:%
%:%474=154%:%
%:%475=154%:%
%:%476=154%:%
%:%477=154%:%
%:%478=155%:%
%:%479=155%:%
%:%480=156%:%
%:%481=156%:%
%:%482=156%:%
%:%483=157%:%
%:%484=157%:%
%:%485=157%:%
%:%486=157%:%
%:%487=157%:%
%:%488=158%:%
%:%489=158%:%
%:%490=158%:%
%:%491=158%:%
%:%492=158%:%
%:%493=159%:%
%:%494=159%:%
%:%495=159%:%
%:%496=159%:%
%:%497=160%:%
%:%498=160%:%
%:%499=160%:%
%:%500=160%:%
%:%501=160%:%
%:%502=161%:%
%:%503=161%:%
%:%504=161%:%
%:%505=161%:%
%:%506=161%:%
%:%507=162%:%
%:%513=162%:%
%:%516=163%:%
%:%517=164%:%
%:%518=164%:%
%:%525=165%:%
%:%526=165%:%
%:%527=166%:%
%:%528=166%:%
%:%529=167%:%
%:%530=167%:%
%:%531=168%:%
%:%532=168%:%
%:%533=169%:%
%:%534=169%:%
%:%535=170%:%
%:%536=170%:%
%:%537=170%:%
%:%538=171%:%
%:%539=172%:%
%:%540=173%:%
%:%541=173%:%
%:%542=173%:%
%:%543=173%:%
%:%544=174%:%
%:%545=175%:%
%:%546=175%:%
%:%547=176%:%
%:%548=176%:%
%:%549=176%:%
%:%550=177%:%
%:%551=177%:%
%:%552=177%:%
%:%553=178%:%
%:%554=178%:%
%:%555=178%:%
%:%556=178%:%
%:%557=179%:%
%:%558=179%:%
%:%559=179%:%
%:%560=179%:%
%:%561=179%:%
%:%562=179%:%
%:%563=179%:%
%:%564=180%:%
%:%565=180%:%
%:%566=180%:%
%:%567=181%:%
%:%568=181%:%
%:%569=181%:%
%:%570=181%:%
%:%571=182%:%
%:%572=182%:%
%:%573=182%:%
%:%574=182%:%
%:%575=183%:%
%:%576=183%:%
%:%577=183%:%
%:%578=184%:%
%:%579=184%:%
%:%580=184%:%
%:%581=184%:%
%:%582=184%:%
%:%583=185%:%
%:%584=185%:%
%:%585=185%:%
%:%586=186%:%
%:%587=186%:%
%:%588=186%:%
%:%589=187%:%
%:%590=187%:%
%:%591=188%:%
%:%592=188%:%
%:%593=189%:%
%:%594=189%:%
%:%595=189%:%
%:%596=190%:%
%:%597=190%:%
%:%598=190%:%
%:%599=191%:%
%:%600=191%:%
%:%601=191%:%
%:%602=191%:%
%:%603=191%:%
%:%604=192%:%
%:%605=192%:%
%:%606=192%:%
%:%607=192%:%
%:%608=193%:%
%:%609=193%:%
%:%610=193%:%
%:%611=194%:%
%:%612=194%:%
%:%613=195%:%
%:%614=195%:%
%:%615=196%:%
%:%616=196%:%
%:%617=196%:%
%:%618=197%:%
%:%619=197%:%
%:%620=197%:%
%:%621=198%:%
%:%622=198%:%
%:%623=198%:%
%:%624=198%:%
%:%625=198%:%
%:%626=199%:%
%:%627=199%:%
%:%628=200%:%
%:%629=200%:%
%:%630=200%:%
%:%631=201%:%
%:%632=201%:%
%:%633=201%:%
%:%634=202%:%
%:%635=202%:%
%:%636=202%:%
%:%637=203%:%
%:%638=203%:%
%:%639=203%:%
%:%640=204%:%
%:%641=204%:%
%:%642=204%:%
%:%643=204%:%
%:%644=205%:%
%:%645=205%:%
%:%646=205%:%
%:%647=205%:%
%:%648=205%:%
%:%649=206%:%
%:%650=206%:%
%:%651=206%:%
%:%652=207%:%
%:%653=207%:%
%:%654=207%:%
%:%655=207%:%
%:%656=208%:%
%:%657=208%:%
%:%658=208%:%
%:%659=208%:%
%:%660=209%:%
%:%661=209%:%
%:%662=209%:%
%:%663=210%:%
%:%664=210%:%
%:%665=210%:%
%:%666=211%:%
%:%667=211%:%
%:%668=211%:%
%:%669=212%:%
%:%670=212%:%
%:%671=212%:%
%:%672=213%:%
%:%673=213%:%
%:%674=214%:%
%:%675=214%:%
%:%676=214%:%
%:%677=215%:%
%:%678=215%:%
%:%679=216%:%
%:%680=216%:%
%:%681=217%:%
%:%682=217%:%
%:%683=217%:%
%:%684=217%:%
%:%685=218%:%
%:%686=218%:%
%:%687=218%:%
%:%688=219%:%
%:%689=219%:%
%:%690=219%:%
%:%691=220%:%
%:%692=220%:%
%:%693=220%:%
%:%694=221%:%
%:%695=221%:%
%:%696=221%:%
%:%697=221%:%
%:%698=222%:%
%:%699=222%:%
%:%700=223%:%
%:%701=224%:%
%:%702=224%:%
%:%703=225%:%
%:%704=225%:%
%:%705=226%:%
%:%706=226%:%
%:%707=226%:%
%:%708=227%:%
%:%709=227%:%
%:%710=227%:%
%:%711=227%:%
%:%712=228%:%
%:%713=228%:%
%:%714=228%:%
%:%715=228%:%
%:%716=228%:%
%:%717=229%:%
%:%718=229%:%
%:%719=229%:%
%:%720=230%:%
%:%721=230%:%
%:%722=230%:%
%:%723=230%:%
%:%724=230%:%
%:%725=231%:%
%:%726=231%:%
%:%727=231%:%
%:%728=232%:%
%:%729=232%:%
%:%730=233%:%
%:%731=233%:%
%:%732=233%:%
%:%733=233%:%
%:%734=234%:%
%:%735=234%:%
%:%736=234%:%
%:%737=234%:%
%:%738=234%:%
%:%739=235%:%
%:%740=235%:%
%:%741=235%:%
%:%742=235%:%
%:%743=235%:%
%:%744=236%:%
%:%745=237%:%
%:%746=237%:%
%:%747=238%:%
%:%748=238%:%
%:%749=239%:%
%:%750=239%:%
%:%751=239%:%
%:%752=240%:%
%:%753=240%:%
%:%754=240%:%
%:%755=240%:%
%:%756=241%:%
%:%757=241%:%
%:%758=241%:%
%:%759=242%:%
%:%760=242%:%
%:%761=243%:%
%:%762=243%:%
%:%763=244%:%
%:%764=244%:%
%:%765=245%:%
%:%766=245%:%
%:%767=246%:%
%:%768=246%:%
%:%769=247%:%
%:%770=247%:%
%:%771=248%:%
%:%772=248%:%
%:%773=249%:%
%:%774=249%:%
%:%775=249%:%
%:%776=249%:%
%:%777=249%:%
%:%778=250%:%
%:%779=250%:%
%:%780=250%:%
%:%781=250%:%
%:%782=250%:%
%:%783=251%:%
%:%784=251%:%
%:%785=252%:%
%:%786=253%:%
%:%787=253%:%
%:%788=253%:%
%:%789=254%:%
%:%790=254%:%
%:%791=255%:%
%:%792=255%:%
%:%793=256%:%
%:%794=256%:%
%:%795=257%:%
%:%796=257%:%
%:%797=258%:%
%:%798=258%:%
%:%799=259%:%
%:%800=259%:%
%:%801=260%:%
%:%802=260%:%
%:%803=261%:%
%:%804=261%:%
%:%805=262%:%
%:%806=262%:%
%:%807=263%:%
%:%808=263%:%
%:%809=264%:%
%:%810=264%:%
%:%811=265%:%
%:%812=265%:%
%:%813=266%:%
%:%814=266%:%
%:%815=267%:%
%:%816=267%:%
%:%817=268%:%
%:%818=268%:%
%:%819=269%:%
%:%825=269%:%
%:%828=270%:%
%:%829=271%:%
%:%830=271%:%
%:%837=272%:%
%:%838=272%:%
%:%839=273%:%
%:%840=273%:%
%:%841=274%:%
%:%842=274%:%
%:%843=275%:%
%:%844=275%:%
%:%845=276%:%
%:%846=276%:%
%:%847=276%:%
%:%848=277%:%
%:%849=278%:%
%:%850=279%:%
%:%851=279%:%
%:%852=280%:%
%:%853=280%:%
%:%854=280%:%
%:%855=281%:%
%:%856=281%:%
%:%857=281%:%
%:%858=282%:%
%:%859=282%:%
%:%860=282%:%
%:%861=282%:%
%:%862=283%:%
%:%863=283%:%
%:%864=283%:%
%:%865=284%:%
%:%866=284%:%
%:%867=284%:%
%:%868=284%:%
%:%869=284%:%
%:%870=284%:%
%:%871=285%:%
%:%872=285%:%
%:%873=285%:%
%:%874=285%:%
%:%875=285%:%
%:%876=285%:%
%:%877=285%:%
%:%878=286%:%
%:%879=286%:%
%:%880=286%:%
%:%881=287%:%
%:%882=287%:%
%:%883=288%:%
%:%884=288%:%
%:%885=289%:%
%:%886=289%:%
%:%887=289%:%
%:%888=290%:%
%:%889=290%:%
%:%890=290%:%
%:%891=290%:%
%:%892=290%:%
%:%893=290%:%
%:%894=291%:%
%:%895=291%:%
%:%896=291%:%
%:%897=291%:%
%:%898=292%:%
%:%899=292%:%
%:%900=292%:%
%:%901=292%:%
%:%902=293%:%
%:%903=293%:%
%:%904=294%:%
%:%905=294%:%
%:%906=294%:%
%:%907=294%:%
%:%908=294%:%
%:%909=295%:%
%:%910=295%:%
%:%911=295%:%
%:%912=295%:%
%:%913=296%:%
%:%914=296%:%
%:%915=296%:%
%:%916=296%:%
%:%917=296%:%
%:%918=296%:%
%:%919=297%:%
%:%920=297%:%
%:%921=297%:%
%:%922=297%:%
%:%923=297%:%
%:%924=298%:%
%:%925=298%:%
%:%926=299%:%
%:%927=299%:%
%:%928=299%:%
%:%929=299%:%
%:%930=299%:%
%:%931=299%:%
%:%932=300%:%
%:%933=300%:%
%:%934=300%:%
%:%935=300%:%
%:%936=301%:%
%:%937=301%:%
%:%938=302%:%
%:%939=303%:%
%:%940=303%:%
%:%941=304%:%
%:%942=304%:%
%:%943=305%:%
%:%944=305%:%
%:%945=306%:%
%:%946=306%:%
%:%947=307%:%
%:%948=307%:%
%:%949=308%:%
%:%950=308%:%
%:%951=309%:%
%:%952=309%:%
%:%953=310%:%
%:%954=310%:%
%:%955=311%:%
%:%956=311%:%
%:%957=312%:%
%:%958=312%:%
%:%959=313%:%
%:%960=313%:%
%:%961=314%:%
%:%967=314%:%
%:%970=315%:%
%:%971=316%:%
%:%972=316%:%
%:%975=317%:%
%:%979=317%:%
%:%980=317%:%
%:%981=317%:%
%:%982=318%:%
%:%983=318%:%
%:%984=319%:%
%:%985=319%:%
%:%986=319%:%
%:%987=320%:%
%:%988=320%:%
%:%989=321%:%
%:%990=321%:%
%:%991=321%:%
%:%992=322%:%
%:%993=322%:%
%:%994=323%:%
%:%995=323%:%
%:%996=323%:%
%:%997=323%:%
%:%998=324%:%
%:%999=324%:%
%:%1000=324%:%
%:%1001=325%:%
%:%1002=325%:%
%:%1003=326%:%
%:%1004=326%:%
%:%1005=326%:%
%:%1006=327%:%
%:%1007=327%:%
%:%1008=328%:%
%:%1009=328%:%
%:%1010=328%:%
%:%1011=328%:%
%:%1012=329%:%
%:%1013=329%:%
%:%1014=329%:%
%:%1015=330%:%
%:%1016=330%:%
%:%1017=331%:%
%:%1018=331%:%
%:%1019=332%:%
%:%1020=332%:%
%:%1021=332%:%
%:%1022=332%:%
%:%1023=332%:%
%:%1024=333%:%
%:%1030=333%:%
%:%1033=334%:%
%:%1034=335%:%
%:%1035=335%:%
%:%1042=336%:%
%:%1043=336%:%
%:%1044=337%:%
%:%1045=337%:%
%:%1046=338%:%
%:%1047=338%:%
%:%1048=339%:%
%:%1049=339%:%
%:%1050=340%:%
%:%1051=340%:%
%:%1052=341%:%
%:%1053=341%:%
%:%1054=342%:%
%:%1055=342%:%
%:%1056=342%:%
%:%1057=343%:%
%:%1058=343%:%
%:%1059=344%:%
%:%1060=344%:%
%:%1061=345%:%
%:%1062=345%:%
%:%1063=346%:%
%:%1064=346%:%
%:%1065=346%:%
%:%1066=347%:%
%:%1067=347%:%
%:%1068=348%:%
%:%1069=348%:%
%:%1070=349%:%
%:%1071=349%:%
%:%1072=349%:%
%:%1073=349%:%
%:%1074=350%:%
%:%1075=351%:%
%:%1076=351%:%
%:%1077=352%:%
%:%1078=352%:%
%:%1079=353%:%
%:%1080=353%:%
%:%1081=354%:%
%:%1082=354%:%
%:%1083=355%:%
%:%1084=355%:%
%:%1085=355%:%
%:%1086=356%:%
%:%1087=356%:%
%:%1088=357%:%
%:%1089=357%:%
%:%1090=358%:%
%:%1091=358%:%
%:%1092=359%:%
%:%1093=359%:%
%:%1094=360%:%
%:%1095=360%:%
%:%1096=360%:%
%:%1097=361%:%
%:%1098=361%:%
%:%1099=361%:%
%:%1100=362%:%
%:%1101=362%:%
%:%1102=362%:%
%:%1103=363%:%
%:%1104=363%:%
%:%1105=364%:%
%:%1106=364%:%
%:%1107=365%:%
%:%1108=365%:%
%:%1109=366%:%
%:%1110=366%:%
%:%1111=366%:%
%:%1112=367%:%
%:%1113=367%:%
%:%1114=368%:%
%:%1115=368%:%
%:%1116=369%:%
%:%1117=369%:%
%:%1118=370%:%
%:%1119=370%:%
%:%1120=370%:%
%:%1121=371%:%
%:%1122=371%:%
%:%1123=372%:%
%:%1124=372%:%
%:%1125=373%:%
%:%1126=373%:%
%:%1127=373%:%
%:%1128=374%:%
%:%1129=374%:%
%:%1130=375%:%
%:%1131=375%:%
%:%1132=376%:%
%:%1133=376%:%
%:%1134=377%:%
%:%1135=377%:%
%:%1136=377%:%
%:%1137=377%:%
%:%1138=378%:%
%:%1144=378%:%
%:%1147=379%:%
%:%1148=380%:%
%:%1149=380%:%
%:%1150=381%:%
%:%1157=382%:%
%:%1158=382%:%
%:%1159=383%:%
%:%1160=383%:%
%:%1161=384%:%
%:%1162=384%:%
%:%1163=384%:%
%:%1164=385%:%
%:%1165=385%:%
%:%1166=385%:%
%:%1167=386%:%
%:%1168=386%:%
%:%1169=387%:%
%:%1170=387%:%
%:%1171=388%:%
%:%1172=388%:%
%:%1173=389%:%
%:%1174=389%:%
%:%1175=389%:%
%:%1176=389%:%
%:%1177=390%:%
%:%1183=390%:%
%:%1186=391%:%
%:%1187=392%:%
%:%1188=392%:%
%:%1189=393%:%
%:%1196=394%:%
%:%1197=394%:%
%:%1198=395%:%
%:%1199=395%:%
%:%1200=396%:%
%:%1201=396%:%
%:%1202=397%:%
%:%1203=397%:%
%:%1204=398%:%
%:%1205=398%:%
%:%1206=399%:%
%:%1207=399%:%
%:%1208=400%:%
%:%1209=400%:%
%:%1210=401%:%
%:%1211=401%:%
%:%1212=401%:%
%:%1213=402%:%
%:%1214=403%:%
%:%1215=404%:%
%:%1216=404%:%
%:%1217=405%:%
%:%1218=405%:%
%:%1219=405%:%
%:%1220=406%:%
%:%1221=406%:%
%:%1222=406%:%
%:%1223=406%:%
%:%1224=407%:%
%:%1225=407%:%
%:%1226=408%:%
%:%1227=408%:%
%:%1228=409%:%
%:%1229=409%:%
%:%1230=410%:%
%:%1231=410%:%
%:%1232=410%:%
%:%1233=411%:%
%:%1234=411%:%
%:%1235=411%:%
%:%1236=412%:%
%:%1237=412%:%
%:%1238=413%:%
%:%1239=413%:%
%:%1240=413%:%
%:%1241=414%:%
%:%1242=414%:%
%:%1243=414%:%
%:%1244=415%:%
%:%1245=415%:%
%:%1246=416%:%
%:%1247=416%:%
%:%1248=416%:%
%:%1249=417%:%
%:%1250=417%:%
%:%1251=417%:%
%:%1252=417%:%
%:%1253=417%:%
%:%1254=418%:%
%:%1255=418%:%
%:%1256=418%:%
%:%1257=418%:%
%:%1258=419%:%
%:%1259=419%:%
%:%1260=420%:%
%:%1261=420%:%
%:%1262=420%:%
%:%1263=421%:%
%:%1264=421%:%
%:%1265=421%:%
%:%1266=422%:%
%:%1267=422%:%
%:%1268=422%:%
%:%1269=423%:%
%:%1270=423%:%
%:%1271=424%:%
%:%1272=424%:%
%:%1273=424%:%
%:%1274=425%:%
%:%1275=425%:%
%:%1276=425%:%
%:%1277=426%:%
%:%1278=426%:%
%:%1279=427%:%
%:%1280=427%:%
%:%1281=427%:%
%:%1282=428%:%
%:%1283=428%:%
%:%1284=428%:%
%:%1285=428%:%
%:%1286=428%:%
%:%1287=429%:%
%:%1288=429%:%
%:%1289=429%:%
%:%1290=429%:%
%:%1291=430%:%
%:%1292=430%:%
%:%1293=430%:%
%:%1294=430%:%
%:%1295=431%:%
%:%1296=431%:%
%:%1297=431%:%
%:%1298=431%:%
%:%1299=431%:%
%:%1300=432%:%
%:%1301=432%:%
%:%1302=433%:%
%:%1303=433%:%
%:%1304=433%:%
%:%1305=434%:%
%:%1306=434%:%
%:%1307=435%:%
%:%1308=435%:%
%:%1309=436%:%
%:%1310=436%:%
%:%1311=436%:%
%:%1312=437%:%
%:%1313=437%:%
%:%1314=437%:%
%:%1315=437%:%
%:%1316=437%:%
%:%1317=438%:%
%:%1318=438%:%
%:%1319=438%:%
%:%1320=438%:%
%:%1321=439%:%
%:%1322=439%:%
%:%1323=439%:%
%:%1324=439%:%
%:%1325=440%:%
%:%1326=440%:%
%:%1327=441%:%
%:%1328=441%:%
%:%1329=441%:%
%:%1330=442%:%
%:%1331=442%:%
%:%1332=442%:%
%:%1333=443%:%
%:%1334=443%:%
%:%1335=444%:%
%:%1336=444%:%
%:%1337=444%:%
%:%1338=445%:%
%:%1339=445%:%
%:%1340=446%:%
%:%1341=446%:%
%:%1342=446%:%
%:%1343=446%:%
%:%1344=446%:%
%:%1345=447%:%
%:%1346=447%:%
%:%1347=447%:%
%:%1348=447%:%
%:%1349=448%:%
%:%1350=448%:%
%:%1351=449%:%
%:%1352=449%:%
%:%1353=450%:%
%:%1354=450%:%
%:%1355=450%:%
%:%1356=451%:%
%:%1357=451%:%
%:%1358=451%:%
%:%1359=452%:%
%:%1365=452%:%
%:%1368=453%:%
%:%1369=454%:%
%:%1370=454%:%
%:%1377=455%:%