### 2024-11-01 code/output/document/Choice_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Choice{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Axiom of Choice in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Choice{\isacharunderscore}{\kern0pt}Axiom\isanewline
\ \ \isakeyword{imports}\ Powerset{\isacharunderscore}{\kern0pt}Axiom\ Pairing{\isacharunderscore}{\kern0pt}Axiom\ Union{\isacharunderscore}{\kern0pt}Axiom\ Extensionality{\isacharunderscore}{\kern0pt}Axiom\ \isanewline
\ \ \ \ \ \ \ \ \ \ Foundation{\isacharunderscore}{\kern0pt}Axiom\ Powerset{\isacharunderscore}{\kern0pt}Axiom\ Separation{\isacharunderscore}{\kern0pt}Axiom\ \isanewline
\ \ \ \ \ \ \ \ \ \ Replacement{\isacharunderscore}{\kern0pt}Axiom\ Interface\ Infinity{\isacharunderscore}{\kern0pt}Axiom\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ induced{\isacharunderscore}{\kern0pt}surj\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}range{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isasymtimes}{\isacharbraceleft}{\kern0pt}e{\isacharbraceright}{\kern0pt}\ {\isasymunion}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}induced{\isacharunderscore}{\kern0pt}surj{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}restrict\ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}prod\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ range{\isacharunderscore}{\kern0pt}restrict{\isacharunderscore}{\kern0pt}vimage{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ a{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}restrictI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ y\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}x\ {\isasymin}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}restrict\ domainI{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}function{\isacharparenleft}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x\ {\isasymin}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y{\isacharequal}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\ {\isacartoucheopen}x{\isasymin}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isasymrangle}\ {\isasymin}\ f{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ assms\ {\isacartoucheopen}x\ {\isasymin}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacartoucheclose}\ \isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y{\isasymin}a{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}vimage{\isacharbrackleft}{\kern0pt}of\ f\ a{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}type{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isacharcolon}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isacharbraceleft}{\kern0pt}e{\isacharbraceright}{\kern0pt}\ {\isasymunion}\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\isanewline
\ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a\ {\isasymLongrightarrow}\ induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}range{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}e{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasyminter}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}restrict\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}range{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}prod\ function{\isacharunderscore}{\kern0pt}vimage{\isacharunderscore}{\kern0pt}Diff\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasyminter}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isacharbraceleft}{\kern0pt}e{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\ relation{\isacharunderscore}{\kern0pt}def\ range{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharcolon}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ range{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}Pi\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharcolon}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ range{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}Pi{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ f\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ function{\isacharunderscore}{\kern0pt}restrictI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ a{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ range{\isacharunderscore}{\kern0pt}restrict{\isacharunderscore}{\kern0pt}vimage\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isacharcolon}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isacharbraceleft}{\kern0pt}e{\isacharbraceright}{\kern0pt}\ {\isasymunion}\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ fun{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}function\ fun{\isacharunderscore}{\kern0pt}disjoint{\isacharunderscore}{\kern0pt}Un\ fun{\isacharunderscore}{\kern0pt}weaken{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}restrict\ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}prod\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isacharcolon}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isacharbraceleft}{\kern0pt}e{\isacharbraceright}{\kern0pt}\ {\isasymunion}\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ restrict\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x\ {\isasymin}\ f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isacartoucheclose}\ \isakeyword{and}\ {\isadigit{1}}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymnotin}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ fun{\isacharunderscore}{\kern0pt}disjoint{\isacharunderscore}{\kern0pt}apply{\isadigit{2}}{\isacharbrackleft}{\kern0pt}of\ x\ {\isacharquery}{\kern0pt}f{\isadigit{1}}\ {\isacharquery}{\kern0pt}f{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}e{\isasymin}a{\isachardoublequoteclose}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymalpha}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ a\ {\isasymLongrightarrow}\ {\isasymexists}x{\isasymin}{\isasymalpha}{\isachardot}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ CollectI\ ballI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isasymalpha}\ {\isasymrightarrow}\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ f\ a\ e{\isacharbrackright}{\kern0pt}\ cons{\isacharunderscore}{\kern0pt}eq\ cons{\isacharunderscore}{\kern0pt}absorb\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ y\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ a{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}x{\isasymin}{\isasymalpha}{\isachardot}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ {\isachardoublequoteopen}x{\isasymin}{\isasymalpha}{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}y{\isasymin}a{\isacartoucheclose}\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}f{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ vimage{\isacharunderscore}{\kern0pt}iff\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharbrackleft}{\kern0pt}of\ f\ x{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}f\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ y{\isacartoucheclose}\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ a{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}{\isasymalpha}{\isacartoucheclose}\ \isacommand{show}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymexists}x{\isasymin}{\isasymalpha}{\isachardot}{\kern0pt}\ induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ a{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{context}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\ \isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ upair{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isasymtau}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}xo{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}yo{\isasymin}M{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}xo{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}yo{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ upair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}xo{\isacharcomma}{\kern0pt}yo{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ z\ {\isacharequal}{\kern0pt}\ upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}M{\isacharsemicolon}{\kern0pt}\ y{\isasymin}M\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}o{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}y{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}o{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}upair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}z{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ s{\isasymin}nat{\isacharsemicolon}{\kern0pt}x{\isasymin}nat{\isacharsemicolon}{\kern0pt}y{\isasymin}nat{\isacharsemicolon}{\kern0pt}o{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}s{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}o{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}o{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}one{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ opair{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}upxx{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}upxy{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}upxx{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}upxy{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymand}\ is{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}upxx{\isacharcomma}{\kern0pt}upxy{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ z\ {\isacharequal}{\kern0pt}\ opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}abs\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}M{\isacharsemicolon}{\kern0pt}\ y{\isasymin}M\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}o{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}y{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}o{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}o{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}z{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ s{\isasymin}nat{\isacharsemicolon}{\kern0pt}x{\isasymin}nat{\isacharsemicolon}{\kern0pt}y{\isasymin}nat{\isacharsemicolon}{\kern0pt}o{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}s{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}o{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}o{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}one{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ sats{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}Upair\ \ generic\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymlangle}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isasymrangle}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ Pair{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}upair{\isacharunderscore}{\kern0pt}name\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}RepFun{\isacharunderscore}{\kern0pt}one{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isasymlangle}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isachardot}{\kern0pt}\ x{\isasymin}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x{\isasymin}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}A\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}Q\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlambda}{\isasymlangle}x{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isachardot}{\kern0pt}\ p\ {\isacharequal}{\kern0pt}\ one{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}one\ {\isasymin}\ P{\isasyminter}G{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ generic\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}{\isasymlangle}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ a{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}t\ {\isasymin}\ {\isacharquery}{\kern0pt}A\ {\isasymtimes}\ P\ {\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isasymlangle}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}\ \ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}t\ {\isasymin}\ {\isacharquery}{\kern0pt}A\ {\isasymtimes}\ P\ {\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t\ {\isasymin}\ {\isacharquery}{\kern0pt}A\ {\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isasyminter}G\ {\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}alt\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ t\ {\isasymin}\ {\isacharquery}{\kern0pt}A\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}one{\isasymin}P{\isasyminter}G{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{$M[G]$ is a transitive model of ZF%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{interpretation}\isamarkupfalse%
\ mgzf{\isacharcolon}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}MG\ generic\ pairing{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ Union{\isacharunderscore}{\kern0pt}MG\ \isanewline
\ \ \ \ extensionality{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ power{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ foundation{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \ \isanewline
\ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ separation{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ infinity{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ unfold{\isacharunderscore}{\kern0pt}locales\ simp{\isacharunderscore}{\kern0pt}all%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}s{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}chx{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}sx{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}chx{\isacharparenright}{\kern0pt}\ {\isasymand}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}s{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}sx{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}chx{\isacharcomma}{\kern0pt}sx{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}s{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}o{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}s{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ s{\isasymin}nat{\isacharsemicolon}{\kern0pt}x{\isasymin}nat{\isacharsemicolon}{\kern0pt}y{\isasymin}nat{\isacharsemicolon}{\kern0pt}o{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}s{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}o{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}o{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}one{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}y{\isacharless}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}o{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ sats{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm\ sats{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}s{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}s{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ y\ {\isacharequal}{\kern0pt}\ opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}s{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}def\ \ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ check{\isacharunderscore}{\kern0pt}abs\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}abs\ apply{\isacharunderscore}{\kern0pt}abs\ apply{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ repl{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isasymin}M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharbraceright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}\ {\isadigit{4}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ rcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ trans{\isacharunderscore}{\kern0pt}closure{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Replace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ finite{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ restriction{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}case{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ quasinat{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Memrel{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ singleton{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ fm{\isacharunderscore}{\kern0pt}defs\ iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}A\ {\isasymLongrightarrow}\ opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ opair{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}closed\ apply{\isacharunderscore}{\kern0pt}closed\ transitivity\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}of\ f{\isacharbrackright}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm\isanewline
\ \ \ \ \ \ \ \ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ Repl{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}opname{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{theorem}\isamarkupfalse%
\ choice{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}choice{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}choice{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtau}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymtau}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtau}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ s\ {\isasymalpha}\ \isakeyword{where}\ {\isachardoublequoteopen}s{\isasymin}surj{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}s{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymalpha}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ choice{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}abs\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymalpha}{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}MG\ generic\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ subsetD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}g\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}s{\isacharbackquote}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymbeta}{\isasymin}{\isasymalpha}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}g\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}s{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isasymalpha}{\isasymin}M{\isacartoucheclose}\ repl{\isacharunderscore}{\kern0pt}opname{\isacharunderscore}{\kern0pt}check\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}dot{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}{\isasymlangle}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}s{\isacharbackquote}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isachardot}{\kern0pt}\ {\isasymbeta}{\isasymin}{\isasymalpha}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}dot\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}g\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ singletonM\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{define}\isamarkupfalse%
\ f\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}f\ {\isasymequiv}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}dot{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}g{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}dot\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}g{\isasymtimes}{\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}dot{\isasymin}M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ intro{\isacharcolon}{\kern0pt}GenExtI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}s{\isacharbackquote}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isasymbeta}{\isasymin}{\isasymalpha}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}RepFun{\isacharunderscore}{\kern0pt}one\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isasymbeta}{\isacharcomma}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}s{\isacharbackquote}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymbeta}{\isasymin}{\isasymalpha}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}opair{\isacharunderscore}{\kern0pt}name\ valcheck\ generic\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isasymbeta}{\isacharcomma}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}s{\isacharbackquote}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymbeta}{\isasymin}{\isasymalpha}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymalpha}{\isachardoublequoteclose}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ a\ {\isasymLongrightarrow}\ {\isasymexists}x{\isasymin}{\isasymalpha}{\isachardot}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isakeyword{for}\ y\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}y\ {\isasymin}\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymsigma}\ \isakeyword{where}\ \ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharbrackleft}{\kern0pt}of\ y\ {\isacharunderscore}{\kern0pt}\ {\isasymtau}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}s{\isasymin}surj{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymbeta}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbeta}{\isasymin}{\isasymalpha}{\isachardoublequoteclose}\ {\isachardoublequoteopen}s{\isacharbackquote}{\kern0pt}{\isasymbeta}\ {\isacharequal}{\kern0pt}\ {\isasymsigma}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}s{\isacharbackquote}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}f\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isasymbeta}{\isacharcomma}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}s{\isacharbackquote}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymbeta}{\isasymin}{\isasymalpha}{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymbeta}{\isasymin}{\isasymalpha}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymbeta}{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}f{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}{\isasymbeta}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymbeta}{\isasymin}{\isasymalpha}{\isacartoucheclose}\ \isacommand{show}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymexists}{\isasymbeta}{\isasymin}{\isasymalpha}{\isachardot}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ {\isasymbeta}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}{\isasymalpha}{\isasymin}{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}f{\isacharprime}{\kern0pt}{\isasymin}{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isasymand}\ f{\isacharprime}{\kern0pt}\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}a{\isacharequal}{\kern0pt}{\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}{\isasymin}surj{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ e\ \isakeyword{where}\ {\isachardoublequoteopen}e{\isasymin}a{\isachardoublequoteclose}\ {\isachardoublequoteopen}e{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\ \isakeyword{and}\ {\isadigit{2}}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}f{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}e{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}induced{\isacharunderscore}{\kern0pt}surj{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ induced{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ {\isacartoucheopen}{\isasymalpha}{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ mgzf{\isachardot}{\kern0pt}choice{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}abs\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\ \ \isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Choice_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%32=6%:%
%:%37=6%:%
%:%40=7%:%
%:%41=8%:%
%:%42=8%:%
%:%43=9%:%
%:%44=10%:%
%:%45=11%:%
%:%46=12%:%
%:%47=12%:%
%:%50=13%:%
%:%54=13%:%
%:%55=13%:%
%:%56=13%:%
%:%57=13%:%
%:%62=13%:%
%:%65=14%:%
%:%66=15%:%
%:%67=15%:%
%:%68=16%:%
%:%69=17%:%
%:%76=18%:%
%:%77=18%:%
%:%78=19%:%
%:%79=19%:%
%:%80=20%:%
%:%81=20%:%
%:%82=21%:%
%:%83=21%:%
%:%84=21%:%
%:%85=22%:%
%:%86=22%:%
%:%87=23%:%
%:%88=23%:%
%:%89=24%:%
%:%90=24%:%
%:%91=25%:%
%:%92=25%:%
%:%93=26%:%
%:%94=26%:%
%:%95=26%:%
%:%96=27%:%
%:%97=27%:%
%:%98=28%:%
%:%99=28%:%
%:%100=29%:%
%:%101=29%:%
%:%102=30%:%
%:%103=30%:%
%:%104=31%:%
%:%105=31%:%
%:%106=31%:%
%:%107=32%:%
%:%108=32%:%
%:%109=32%:%
%:%110=33%:%
%:%111=33%:%
%:%112=34%:%
%:%113=34%:%
%:%114=35%:%
%:%115=35%:%
%:%116=36%:%
%:%117=36%:%
%:%118=36%:%
%:%119=37%:%
%:%120=37%:%
%:%121=37%:%
%:%122=38%:%
%:%123=38%:%
%:%124=39%:%
%:%125=39%:%
%:%126=39%:%
%:%127=40%:%
%:%128=40%:%
%:%129=41%:%
%:%130=41%:%
%:%131=42%:%
%:%132=42%:%
%:%133=43%:%
%:%134=43%:%
%:%135=44%:%
%:%136=44%:%
%:%137=44%:%
%:%138=45%:%
%:%144=45%:%
%:%147=46%:%
%:%148=47%:%
%:%149=47%:%
%:%150=48%:%
%:%151=49%:%
%:%152=50%:%
%:%153=51%:%
%:%154=52%:%
%:%155=53%:%
%:%162=54%:%
%:%163=54%:%
%:%164=55%:%
%:%165=55%:%
%:%166=56%:%
%:%167=56%:%
%:%168=57%:%
%:%169=57%:%
%:%170=57%:%
%:%171=58%:%
%:%172=58%:%
%:%173=58%:%
%:%174=59%:%
%:%175=59%:%
%:%176=60%:%
%:%177=60%:%
%:%178=60%:%
%:%179=61%:%
%:%180=61%:%
%:%181=62%:%
%:%182=62%:%
%:%183=63%:%
%:%184=63%:%
%:%185=64%:%
%:%186=64%:%
%:%187=65%:%
%:%188=65%:%
%:%189=66%:%
%:%190=66%:%
%:%191=66%:%
%:%192=67%:%
%:%193=67%:%
%:%194=67%:%
%:%195=68%:%
%:%196=68%:%
%:%197=69%:%
%:%198=69%:%
%:%199=69%:%
%:%200=70%:%
%:%201=70%:%
%:%202=70%:%
%:%203=71%:%
%:%204=71%:%
%:%205=72%:%
%:%206=72%:%
%:%207=72%:%
%:%208=73%:%
%:%209=73%:%
%:%210=73%:%
%:%211=74%:%
%:%212=74%:%
%:%213=75%:%
%:%214=75%:%
%:%215=75%:%
%:%216=76%:%
%:%217=76%:%
%:%218=77%:%
%:%219=77%:%
%:%220=78%:%
%:%221=78%:%
%:%222=78%:%
%:%223=78%:%
%:%224=79%:%
%:%225=79%:%
%:%226=80%:%
%:%227=80%:%
%:%228=81%:%
%:%229=81%:%
%:%230=81%:%
%:%231=82%:%
%:%232=82%:%
%:%233=83%:%
%:%234=83%:%
%:%235=84%:%
%:%236=84%:%
%:%237=85%:%
%:%238=85%:%
%:%239=86%:%
%:%240=86%:%
%:%241=87%:%
%:%242=87%:%
%:%243=88%:%
%:%244=88%:%
%:%245=88%:%
%:%246=89%:%
%:%247=89%:%
%:%248=89%:%
%:%249=90%:%
%:%250=90%:%
%:%251=91%:%
%:%252=91%:%
%:%253=92%:%
%:%254=92%:%
%:%255=93%:%
%:%256=93%:%
%:%257=94%:%
%:%258=94%:%
%:%259=94%:%
%:%260=94%:%
%:%261=95%:%
%:%267=95%:%
%:%270=96%:%
%:%271=97%:%
%:%272=97%:%
%:%273=98%:%
%:%274=99%:%
%:%275=100%:%
%:%276=101%:%
%:%279=102%:%
%:%283=102%:%
%:%284=102%:%
%:%285=103%:%
%:%286=103%:%
%:%287=104%:%
%:%288=104%:%
%:%289=105%:%
%:%290=105%:%
%:%291=106%:%
%:%292=106%:%
%:%293=106%:%
%:%294=107%:%
%:%295=107%:%
%:%296=108%:%
%:%297=108%:%
%:%298=109%:%
%:%299=109%:%
%:%300=110%:%
%:%301=110%:%
%:%302=111%:%
%:%303=111%:%
%:%304=112%:%
%:%305=112%:%
%:%306=113%:%
%:%307=113%:%
%:%308=113%:%
%:%309=114%:%
%:%310=114%:%
%:%311=115%:%
%:%312=115%:%
%:%313=116%:%
%:%314=116%:%
%:%315=116%:%
%:%316=117%:%
%:%317=117%:%
%:%318=118%:%
%:%319=118%:%
%:%320=119%:%
%:%321=119%:%
%:%322=119%:%
%:%323=120%:%
%:%324=120%:%
%:%325=120%:%
%:%326=121%:%
%:%327=121%:%
%:%328=122%:%
%:%334=122%:%
%:%337=123%:%
%:%338=124%:%
%:%339=124%:%
%:%340=125%:%
%:%341=126%:%
%:%342=127%:%
%:%343=127%:%
%:%344=128%:%
%:%345=129%:%
%:%346=130%:%
%:%347=131%:%
%:%348=131%:%
%:%349=132%:%
%:%350=133%:%
%:%351=134%:%
%:%352=135%:%
%:%353=136%:%
%:%354=136%:%
%:%355=137%:%
%:%356=138%:%
%:%359=139%:%
%:%363=139%:%
%:%364=139%:%
%:%365=139%:%
%:%366=139%:%
%:%371=139%:%
%:%374=140%:%
%:%375=141%:%
%:%376=141%:%
%:%377=142%:%
%:%380=143%:%
%:%384=143%:%
%:%385=143%:%
%:%386=143%:%
%:%387=143%:%
%:%392=143%:%
%:%395=144%:%
%:%396=145%:%
%:%397=145%:%
%:%398=146%:%
%:%399=147%:%
%:%400=148%:%
%:%401=149%:%
%:%402=150%:%
%:%403=150%:%
%:%404=151%:%
%:%407=152%:%
%:%411=152%:%
%:%412=152%:%
%:%413=152%:%
%:%418=152%:%
%:%421=153%:%
%:%422=154%:%
%:%423=154%:%
%:%424=155%:%
%:%425=156%:%
%:%426=157%:%
%:%429=158%:%
%:%433=158%:%
%:%434=158%:%
%:%435=158%:%
%:%436=158%:%
%:%441=158%:%
%:%444=159%:%
%:%445=160%:%
%:%446=160%:%
%:%447=161%:%
%:%448=162%:%
%:%449=163%:%
%:%450=164%:%
%:%451=164%:%
%:%452=165%:%
%:%453=166%:%
%:%454=167%:%
%:%455=168%:%
%:%456=169%:%
%:%457=169%:%
%:%458=170%:%
%:%459=171%:%
%:%462=172%:%
%:%466=172%:%
%:%467=172%:%
%:%468=172%:%
%:%469=172%:%
%:%474=172%:%
%:%477=173%:%
%:%478=174%:%
%:%479=174%:%
%:%480=175%:%
%:%483=176%:%
%:%487=176%:%
%:%488=176%:%
%:%489=176%:%
%:%490=176%:%
%:%495=176%:%
%:%498=177%:%
%:%499=178%:%
%:%500=178%:%
%:%501=179%:%
%:%502=180%:%
%:%503=181%:%
%:%504=182%:%
%:%505=183%:%
%:%506=183%:%
%:%507=184%:%
%:%510=185%:%
%:%514=185%:%
%:%515=185%:%
%:%516=185%:%
%:%521=185%:%
%:%524=186%:%
%:%525=187%:%
%:%526=187%:%
%:%527=188%:%
%:%528=189%:%
%:%529=190%:%
%:%532=191%:%
%:%536=191%:%
%:%537=191%:%
%:%538=191%:%
%:%539=191%:%
%:%544=191%:%
%:%547=192%:%
%:%548=193%:%
%:%549=193%:%
%:%552=194%:%
%:%556=194%:%
%:%557=194%:%
%:%558=194%:%
%:%559=194%:%
%:%564=194%:%
%:%567=195%:%
%:%568=196%:%
%:%569=196%:%
%:%572=197%:%
%:%576=197%:%
%:%577=197%:%
%:%578=197%:%
%:%579=197%:%
%:%584=197%:%
%:%587=198%:%
%:%588=199%:%
%:%589=199%:%
%:%596=200%:%
%:%597=200%:%
%:%598=201%:%
%:%599=201%:%
%:%600=202%:%
%:%601=202%:%
%:%602=203%:%
%:%603=203%:%
%:%604=203%:%
%:%605=203%:%
%:%606=204%:%
%:%607=204%:%
%:%608=205%:%
%:%609=205%:%
%:%610=205%:%
%:%611=206%:%
%:%612=206%:%
%:%613=207%:%
%:%614=207%:%
%:%615=208%:%
%:%616=208%:%
%:%617=209%:%
%:%618=209%:%
%:%619=210%:%
%:%620=210%:%
%:%621=211%:%
%:%622=211%:%
%:%623=211%:%
%:%624=212%:%
%:%625=212%:%
%:%626=213%:%
%:%627=213%:%
%:%628=214%:%
%:%629=214%:%
%:%630=214%:%
%:%631=215%:%
%:%632=215%:%
%:%633=216%:%
%:%634=216%:%
%:%635=217%:%
%:%636=217%:%
%:%637=218%:%
%:%638=218%:%
%:%639=218%:%
%:%640=218%:%
%:%641=219%:%
%:%656=221%:%
%:%666=223%:%
%:%667=223%:%
%:%670=224%:%
%:%674=224%:%
%:%675=224%:%
%:%676=225%:%
%:%677=226%:%
%:%678=227%:%
%:%679=227%:%
%:%684=227%:%
%:%687=228%:%
%:%688=229%:%
%:%689=230%:%
%:%690=230%:%
%:%691=231%:%
%:%692=232%:%
%:%693=233%:%
%:%694=234%:%
%:%695=235%:%
%:%696=235%:%
%:%697=236%:%
%:%698=237%:%
%:%699=238%:%
%:%700=239%:%
%:%701=240%:%
%:%702=240%:%
%:%703=241%:%
%:%706=242%:%
%:%710=242%:%
%:%711=242%:%
%:%712=242%:%
%:%717=242%:%
%:%720=243%:%
%:%721=244%:%
%:%722=244%:%
%:%723=245%:%
%:%724=246%:%
%:%725=247%:%
%:%726=248%:%
%:%729=249%:%
%:%733=249%:%
%:%734=249%:%
%:%735=250%:%
%:%736=250%:%
%:%737=250%:%
%:%742=250%:%
%:%745=251%:%
%:%746=252%:%
%:%747=253%:%
%:%748=253%:%
%:%749=254%:%
%:%750=255%:%
%:%753=256%:%
%:%757=256%:%
%:%758=256%:%
%:%759=257%:%
%:%760=257%:%
%:%761=257%:%
%:%766=257%:%
%:%769=258%:%
%:%770=259%:%
%:%771=259%:%
%:%772=260%:%
%:%773=261%:%
%:%774=262%:%
%:%775=263%:%
%:%782=264%:%
%:%783=264%:%
%:%784=265%:%
%:%785=265%:%
%:%786=266%:%
%:%787=266%:%
%:%788=267%:%
%:%789=268%:%
%:%790=269%:%
%:%791=270%:%
%:%792=271%:%
%:%793=271%:%
%:%794=272%:%
%:%795=272%:%
%:%796=273%:%
%:%797=273%:%
%:%798=274%:%
%:%799=274%:%
%:%800=275%:%
%:%801=275%:%
%:%802=276%:%
%:%803=276%:%
%:%804=277%:%
%:%805=277%:%
%:%806=277%:%
%:%807=278%:%
%:%808=279%:%
%:%809=280%:%
%:%810=281%:%
%:%811=281%:%
%:%812=282%:%
%:%818=282%:%
%:%821=283%:%
%:%822=284%:%
%:%823=285%:%
%:%824=286%:%
%:%825=286%:%
%:%826=287%:%
%:%827=288%:%
%:%834=289%:%
%:%835=289%:%
%:%836=290%:%
%:%837=290%:%
%:%838=291%:%
%:%839=291%:%
%:%840=292%:%
%:%841=292%:%
%:%842=293%:%
%:%843=293%:%
%:%844=294%:%
%:%845=294%:%
%:%846=295%:%
%:%847=295%:%
%:%848=295%:%
%:%849=296%:%
%:%850=296%:%
%:%851=297%:%
%:%852=297%:%
%:%853=298%:%
%:%854=298%:%
%:%855=298%:%
%:%856=299%:%
%:%857=299%:%
%:%858=300%:%
%:%859=300%:%
%:%860=301%:%
%:%861=301%:%
%:%862=301%:%
%:%863=302%:%
%:%864=302%:%
%:%865=303%:%
%:%866=303%:%
%:%867=304%:%
%:%868=304%:%
%:%869=304%:%
%:%870=305%:%
%:%871=305%:%
%:%872=306%:%
%:%873=306%:%
%:%874=307%:%
%:%875=307%:%
%:%876=307%:%
%:%877=307%:%
%:%878=308%:%
%:%879=308%:%
%:%880=309%:%
%:%881=309%:%
%:%882=309%:%
%:%883=310%:%
%:%884=310%:%
%:%885=310%:%
%:%886=310%:%
%:%887=310%:%
%:%888=311%:%
%:%889=311%:%
%:%890=312%:%
%:%891=313%:%
%:%892=313%:%
%:%893=314%:%
%:%894=314%:%
%:%895=315%:%
%:%896=315%:%
%:%897=315%:%
%:%898=316%:%
%:%899=316%:%
%:%900=317%:%
%:%901=317%:%
%:%902=318%:%
%:%903=318%:%
%:%904=318%:%
%:%905=319%:%
%:%906=319%:%
%:%907=320%:%
%:%908=320%:%
%:%909=320%:%
%:%910=320%:%
%:%911=321%:%
%:%912=321%:%
%:%913=322%:%
%:%914=322%:%
%:%915=323%:%
%:%916=323%:%
%:%917=323%:%
%:%918=324%:%
%:%919=324%:%
%:%920=325%:%
%:%921=325%:%
%:%922=325%:%
%:%923=326%:%
%:%924=326%:%
%:%925=327%:%
%:%926=327%:%
%:%927=328%:%
%:%928=328%:%
%:%929=328%:%
%:%930=329%:%
%:%931=329%:%
%:%932=330%:%
%:%933=330%:%
%:%934=331%:%
%:%935=331%:%
%:%936=332%:%
%:%937=332%:%
%:%938=333%:%
%:%939=334%:%
%:%940=334%:%
%:%941=335%:%
%:%942=335%:%
%:%943=336%:%
%:%944=336%:%
%:%945=336%:%
%:%946=337%:%
%:%947=337%:%
%:%948=338%:%
%:%949=338%:%
%:%950=339%:%
%:%951=339%:%
%:%952=339%:%
%:%953=340%:%
%:%954=340%:%
%:%955=341%:%
%:%956=341%:%
%:%957=342%:%
%:%958=342%:%
%:%959=343%:%
%:%960=343%:%
%:%961=344%:%
%:%962=344%:%
%:%963=345%:%
%:%964=345%:%
%:%965=346%:%
%:%966=346%:%
%:%967=347%:%
%:%968=347%:%
%:%969=348%:%
%:%970=348%:%
%:%971=348%:%
%:%972=349%:%
%:%973=349%:%
%:%974=349%:%
%:%975=350%:%
%:%976=351%:%
%:%977=351%:%
%:%978=352%:%
%:%979=352%:%
%:%980=353%:%
%:%981=353%:%
%:%982=354%:%
%:%983=354%:%
%:%984=355%:%
%:%985=355%:%
%:%986=356%:%
%:%987=356%:%
%:%988=357%:%
%:%989=357%:%
%:%990=358%:%
%:%991=358%:%
%:%992=359%:%
%:%993=359%:%
%:%994=359%:%
%:%995=360%:%
%:%996=360%:%
%:%997=361%:%
%:%998=361%:%
%:%999=361%:%
%:%1000=361%:%
%:%1001=362%:%
%:%1002=362%:%
%:%1003=363%:%
%:%1004=363%:%
%:%1005=364%:%
%:%1006=364%:%
%:%1007=365%:%
%:%1008=365%:%
%:%1009=366%:%
%:%1010=366%:%
%:%1011=366%:%
%:%1012=367%:%
%:%1013=367%:%
%:%1014=368%:%
%:%1015=368%:%
%:%1016=369%:%
%:%1017=369%:%
%:%1018=369%:%
%:%1019=370%:%
%:%1020=370%:%
%:%1021=370%:%
%:%1022=371%:%
%:%1023=371%:%
%:%1024=372%:%
%:%1025=372%:%
%:%1026=373%:%
%:%1027=373%:%
%:%1028=374%:%
%:%1029=374%:%
%:%1030=374%:%
%:%1031=375%:%
%:%1032=376%:%
%:%1033=376%:%
%:%1034=376%:%
%:%1035=376%:%
%:%1036=377%:%
%:%1037=377%:%
%:%1038=378%:%
%:%1039=378%:%
%:%1040=379%:%
%:%1041=379%:%
%:%1042=380%:%
%:%1043=380%:%
%:%1044=380%:%
%:%1045=380%:%
%:%1046=381%:%
%:%1052=381%:%
%:%1055=382%:%
%:%1056=383%:%
%:%1057=383%:%
%:%1060=384%:%
%:%1065=385%:%