### 2024-11-01 code/output/document/Delta0.tex

%
\begin{isabellebody}%
\setisabellecontext{Delta{\isadigit{0}}}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Delta{\isadigit{0}}\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ Utilities{\isacharunderscore}{\kern0pt}M\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{definition}\isamarkupfalse%
\ BExists{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\isacommand{definition}\isamarkupfalse%
\ BExists\ \isakeyword{where}\ {\isachardoublequoteopen}BExists{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ BForall\ \isakeyword{where}\ {\isachardoublequoteopen}BForall{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Neg{\isacharparenleft}{\kern0pt}BExists{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ X\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}\ Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isachargreater}{\kern0pt}\ {\isasymin}\ X\ {\isasymtimes}\ X\ {\isacharbraceright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}\ {\isasympsi}\ {\isasymin}\ formula{\isachardot}{\kern0pt}\ {\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}{\isasymphi}\ {\isasymin}\ X{\isachardot}{\kern0pt}\ n\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\ {\isasympsi}\ {\isacharequal}{\kern0pt}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ formula{\isachardot}{\kern0pt}\ {\isasymexists}\ n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}\ m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isacharequal}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isasymphi}\ {\isacharequal}{\kern0pt}\ Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymDelta}{\isadigit{0}}\ {\isasymequiv}\ {\isasymUnion}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymDelta}{\isadigit{0}}\ {\isasymsubseteq}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}n{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}n{\isachardot}{\kern0pt}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ formula{\isachardoublequoteclose}\ \isakeyword{in}\ nat{\isacharunderscore}{\kern0pt}induct{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}increasing\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymle}\ m{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}m{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ n\ {\isasymle}\ m\ {\isasymlongrightarrow}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}m{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}m{\isachardot}{\kern0pt}\ {\isasymforall}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ n\ {\isasymle}\ m\ {\isasymlongrightarrow}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}m{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ nat{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k\ n{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}n\ {\isasymle}\ k{\isachardoublequoteclose}\ \isakeyword{and}\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}n\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}k{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ disjE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ subst\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Member{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Equal{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Nand{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ {\isasympsi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ n\ \isakeyword{where}\ nH{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ mH{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasympsi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}m{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ k\ \isakeyword{where}\ {\isachardoublequoteopen}k\ {\isasymequiv}\ n\ {\isasymunion}\ m{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ knat{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}type\ nH\ mH\ k{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mnle{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymle}\ k\ {\isasymand}\ n\ {\isasymle}\ k{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ k{\isacharunderscore}{\kern0pt}def\ max{\isacharunderscore}{\kern0pt}le{\isadigit{1}}\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\ nH\ mH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}k{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isasympsi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}k{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mnle\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}increasing\ knat\ mH\ nH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}k{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ knat\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Neg{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Neg{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ And{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ {\isasympsi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ And{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Or{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ {\isasympsi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Or{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Or{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Implies{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ {\isasympsi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Implies{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Implies{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ Iff{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ {\isasympsi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Iff{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Iff{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ BExists{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}BExists{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ mH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}m{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}BExists{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ BExists{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymphi}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ mH\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ BExists{\isacharunderscore}{\kern0pt}formula\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}BExists{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ BForall{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}BForall{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ BForall{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ B\ env\ {\isasymphi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A\ {\isasymsubseteq}\ B{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}B{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}n{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ {\isasymforall}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}B{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}n{\isachardot}{\kern0pt}\ {\isasymforall}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}B{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ nat{\isacharunderscore}{\kern0pt}induct{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}{\isadigit{0}}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}B{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ {\isasymphi}\ n\ m\ env{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ {\isasymphi}\ n\ m\ env{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymphi}\ {\isacharequal}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}{\isasymphi}{\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymforall}env{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \isacommand{define}\isamarkupfalse%
\ X\ \isakeyword{where}\ {\isachardoublequoteopen}X\ {\isasymequiv}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Xsubset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}X\ {\isasymsubseteq}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ X{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ ih{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}env\ {\isasymphi}{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ X\ {\isasymLongrightarrow}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}\ {\isasymLongrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ X{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ ih{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}env\ {\isasymphi}{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ X\ {\isasymLongrightarrow}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}\ {\isasymLongrightarrow}\ A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ X{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Hnand{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}{\isasymphi}\ {\isasympsi}\ env{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ X\ {\isasymLongrightarrow}\ {\isasympsi}\ {\isasymin}\ X\ {\isasymLongrightarrow}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ {\isasymphi}\ {\isasympsi}\ env\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ X{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}\ {\isasymin}\ X{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ max{\isacharunderscore}{\kern0pt}le{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ Xsubset\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ iff{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ih{\isadigit{1}}\ ih{\isadigit{2}}\ assms{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ Xsubset\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ iff{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasympsi}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasympsi}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ih{\isadigit{1}}\ ih{\isadigit{2}}\ assms{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ envinBlist\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ assms\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ iff{\isadigit{1}}\ iff{\isadigit{2}}\ assms{\isadigit{2}}\ Xsubset\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Hbex\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}m\ {\isasymphi}\ env{\isachardot}{\kern0pt}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ {\isasymphi}\ {\isasymin}\ X\ {\isasymLongrightarrow}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ m\ {\isasymphi}\ env\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ X{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms{\isadigit{2}}\ X{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ phiformula{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ envinBlist\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ assms\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ eq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}pred{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ Un{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ phiformula\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ eq{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ Ord{\isacharunderscore}{\kern0pt}Un{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ phiformula\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ arityle{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ mle\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ eq{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Ord{\isacharunderscore}{\kern0pt}Un{\isacharcomma}{\kern0pt}\ rule\ Ord{\isacharunderscore}{\kern0pt}Un{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ phiformula\ assms{\isadigit{2}}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isasymin}\ nth{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ B{\isachardot}{\kern0pt}\ x\ {\isasymin}\ nth{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ sats{\isacharparenleft}{\kern0pt}B{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ ih{\isadigit{1}}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}arityle{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ih{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}arityle{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ mem{\isacharunderscore}{\kern0pt}irrefl{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x\ m{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}m{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Transset{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ nth{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ mle{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ envinBlist\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ I{\isadigit{1}}\ I{\isadigit{2}}\ I{\isadigit{3}}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}{\isasymphi}{\isasymin}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymforall}env{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}X\ \isakeyword{in}\ ssubst{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}X{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ {\isasymphi}\ env{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ X{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ih{\isadigit{1}}\ ih{\isadigit{2}}\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjE{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}clarify{\isacharcomma}{\kern0pt}\ rule\ Hnand{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}clarify{\isacharcomma}{\kern0pt}\ rule\ Hbex{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}\ {\isasymlongleftrightarrow}\ B{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ m\ n\ f\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ value{\isacharunderscore}{\kern0pt}type{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}k\ f\ m\ n{\isachardot}{\kern0pt}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ f\ {\isasymin}\ m\ {\isasymrightarrow}\ n\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k\ f\ m\ n{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isasymin}\ m{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ k\ f\ m\ n\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}k\ {\isasymin}\ m{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isachardoublequoteclose}\ \ \ \ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}k\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}rangeI\ assms{\isadigit{2}}\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}k\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}k\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltI\ assms{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ k\ f\ m\ n\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}k\ {\isasymnotin}\ m{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}k\ {\isasymnotin}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}k\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}i{\isachardot}{\kern0pt}\ i\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ \ {\isasymforall}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymforall}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymforall}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isachardot}{\kern0pt}\ {\isasymforall}{\isasymphi}\ {\isasymin}\ formula{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isasymlongrightarrow}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}i{\isachardot}{\kern0pt}\ {\isasymforall}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymforall}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymforall}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isachardot}{\kern0pt}\ {\isasymforall}{\isasymphi}\ {\isasymin}\ formula{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isasymlongrightarrow}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ nat{\isacharunderscore}{\kern0pt}induct{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ i\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}m{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymforall}f{\isasymin}m\ {\isasymrightarrow}\ n{\isachardot}{\kern0pt}\ {\isasymforall}{\isasymphi}{\isasymin}formula{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}{\isadigit{0}}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isasymlongrightarrow}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ m\ {\isacharbackquote}{\kern0pt}\ n\ {\isacharbackquote}{\kern0pt}\ f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}{\isadigit{0}}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ ballI\ {\isacharcomma}{\kern0pt}\ rule\ impI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ m\ n\ f\ {\isasymphi}\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}{\isadigit{0}}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ m\ {\isacharbackquote}{\kern0pt}\ n\ {\isacharbackquote}{\kern0pt}\ f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}{\isadigit{0}}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ value{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ i\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}m{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymforall}f{\isasymin}m\ {\isasymrightarrow}\ n{\isachardot}{\kern0pt}\ {\isasymforall}{\isasymphi}{\isasymin}formula{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isasymlongrightarrow}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ m\ {\isacharbackquote}{\kern0pt}\ n\ {\isacharbackquote}{\kern0pt}\ f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ ih\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}m\ n\ f\ {\isasymphi}{\isachardot}{\kern0pt}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ f\ {\isasymin}\ m\ {\isasymrightarrow}\ n\ {\isasymLongrightarrow}\ {\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isasymLongrightarrow}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ m\ n\ f\ {\isasymphi}{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isasymforall}m{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymforall}f{\isasymin}m\ {\isasymrightarrow}\ n{\isachardot}{\kern0pt}\ {\isasymforall}{\isasymphi}{\isasymin}formula{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isasymlongrightarrow}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ m\ {\isacharbackquote}{\kern0pt}\ n\ {\isacharbackquote}{\kern0pt}\ f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ ballI\ {\isacharcomma}{\kern0pt}\ rule\ impI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ m\ n\ f\ {\isasymphi}\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isasymin}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}y{\isasymin}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isacharequal}{\kern0pt}\ Nand{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}j{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}{\isasympsi}{\isasymin}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ j\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\ {\isasymphi}\ {\isacharequal}{\kern0pt}\ BExists{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ m\ {\isacharbackquote}{\kern0pt}\ n\ {\isacharbackquote}{\kern0pt}\ f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjE{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms{\isadigit{2}}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjE{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x\ y{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}\ {\isasymand}\ y\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x\ y{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymand}\ ren{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i\ {\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ ih{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x\ y{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ max{\isacharunderscore}{\kern0pt}le{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ih{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x\ y{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ j\ {\isasympsi}{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasympsi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ j\ {\isasympsi}{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}j\ {\isasymle}\ m{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ren{\isacharunderscore}{\kern0pt}tc{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ j\ {\isasympsi}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}j{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ j\ {\isasympsi}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}j\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ sum{\isacharunderscore}{\kern0pt}idS{\isacharbrackleft}{\kern0pt}\isakeyword{where}\ q{\isacharequal}{\kern0pt}n{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltD{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ Exists{\isacharunderscore}{\kern0pt}def\ Neg{\isacharunderscore}{\kern0pt}def\ And{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ f{\isacharparenright}{\kern0pt}\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sum{\isacharunderscore}{\kern0pt}id{\isadigit{0}}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ih{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ j\ {\isasympsi}{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ j\ {\isasympsi}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}pred{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ m{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ value{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ j\ {\isasympsi}{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}{\isadigit{1}}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}pred{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\ assms{\isadigit{2}}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}i\ m\ n\ f\ {\isasymphi}{\isachardot}{\kern0pt}\ i\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ f\ {\isasymin}\ m\ {\isasymrightarrow}\ n\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ \ m\ {\isasymLongrightarrow}\ {\isasymphi}\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ {\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}f\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}i\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ i\ \isakeyword{where}\ iH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}from{\isacharcircum}{\kern0pt}i{\isacharparenleft}{\kern0pt}{\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}i\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ main{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ iH\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}subset\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}BExists{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ A\ {\isasymphi}\ env\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ BExists{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isasymin}\ nth{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ BExists{\isacharunderscore}{\kern0pt}def\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}BForall{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ {\isasymphi}\ env\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ BForall{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isasymin}\ nth{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ BForall{\isacharunderscore}{\kern0pt}def\ BExists{\isacharunderscore}{\kern0pt}def\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}BExists\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}BExists{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymunion}\ pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ BExists{\isacharunderscore}{\kern0pt}def\ BExists{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Delta0.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%19=4%:%
%:%22=5%:%
%:%23=6%:%
%:%24=6%:%
%:%25=7%:%
%:%26=7%:%
%:%27=8%:%
%:%28=8%:%
%:%29=9%:%
%:%30=10%:%
%:%31=10%:%
%:%32=11%:%
%:%33=11%:%
%:%34=12%:%
%:%35=13%:%
%:%36=13%:%
%:%37=14%:%
%:%38=15%:%
%:%39=15%:%
%:%46=16%:%
%:%47=16%:%
%:%48=17%:%
%:%49=17%:%
%:%50=18%:%
%:%51=18%:%
%:%52=19%:%
%:%53=19%:%
%:%54=20%:%
%:%55=20%:%
%:%56=21%:%
%:%57=21%:%
%:%58=21%:%
%:%59=22%:%
%:%60=22%:%
%:%61=23%:%
%:%62=23%:%
%:%63=24%:%
%:%69=24%:%
%:%72=25%:%
%:%73=26%:%
%:%74=26%:%
%:%75=27%:%
%:%76=28%:%
%:%77=29%:%
%:%84=30%:%
%:%85=30%:%
%:%86=31%:%
%:%87=31%:%
%:%88=32%:%
%:%89=32%:%
%:%90=33%:%
%:%91=33%:%
%:%92=34%:%
%:%93=34%:%
%:%94=35%:%
%:%95=35%:%
%:%96=36%:%
%:%97=36%:%
%:%98=37%:%
%:%99=37%:%
%:%100=38%:%
%:%101=38%:%
%:%102=39%:%
%:%103=39%:%
%:%104=40%:%
%:%105=40%:%
%:%106=41%:%
%:%107=41%:%
%:%108=41%:%
%:%109=42%:%
%:%110=42%:%
%:%111=43%:%
%:%112=43%:%
%:%113=44%:%
%:%119=44%:%
%:%122=45%:%
%:%123=46%:%
%:%124=46%:%
%:%125=47%:%
%:%126=48%:%
%:%127=49%:%
%:%130=50%:%
%:%134=50%:%
%:%135=50%:%
%:%136=51%:%
%:%137=51%:%
%:%138=52%:%
%:%139=52%:%
%:%140=53%:%
%:%141=53%:%
%:%142=54%:%
%:%143=54%:%
%:%144=55%:%
%:%145=55%:%
%:%150=55%:%
%:%153=56%:%
%:%154=57%:%
%:%155=57%:%
%:%156=58%:%
%:%157=59%:%
%:%158=60%:%
%:%161=61%:%
%:%165=61%:%
%:%166=61%:%
%:%167=62%:%
%:%168=62%:%
%:%169=63%:%
%:%170=63%:%
%:%171=64%:%
%:%172=64%:%
%:%173=65%:%
%:%174=65%:%
%:%175=66%:%
%:%176=66%:%
%:%181=66%:%
%:%184=67%:%
%:%185=68%:%
%:%186=68%:%
%:%187=69%:%
%:%188=70%:%
%:%189=71%:%
%:%196=72%:%
%:%197=72%:%
%:%198=73%:%
%:%199=73%:%
%:%200=74%:%
%:%201=74%:%
%:%202=75%:%
%:%203=75%:%
%:%204=76%:%
%:%205=76%:%
%:%206=77%:%
%:%207=78%:%
%:%208=78%:%
%:%209=79%:%
%:%210=79%:%
%:%211=80%:%
%:%212=80%:%
%:%213=81%:%
%:%214=81%:%
%:%215=82%:%
%:%216=83%:%
%:%217=83%:%
%:%218=84%:%
%:%219=85%:%
%:%220=85%:%
%:%221=85%:%
%:%222=85%:%
%:%223=86%:%
%:%224=86%:%
%:%225=86%:%
%:%226=86%:%
%:%227=87%:%
%:%228=88%:%
%:%229=88%:%
%:%230=89%:%
%:%231=89%:%
%:%232=90%:%
%:%233=90%:%
%:%234=91%:%
%:%235=91%:%
%:%236=91%:%
%:%237=92%:%
%:%238=92%:%
%:%239=93%:%
%:%240=93%:%
%:%241=94%:%
%:%242=94%:%
%:%243=95%:%
%:%244=95%:%
%:%245=95%:%
%:%246=96%:%
%:%247=96%:%
%:%248=97%:%
%:%249=97%:%
%:%250=98%:%
%:%251=98%:%
%:%252=99%:%
%:%258=99%:%
%:%261=100%:%
%:%262=101%:%
%:%263=101%:%
%:%264=102%:%
%:%265=103%:%
%:%266=104%:%
%:%269=105%:%
%:%273=105%:%
%:%274=105%:%
%:%275=106%:%
%:%276=106%:%
%:%277=107%:%
%:%278=107%:%
%:%283=107%:%
%:%286=108%:%
%:%287=109%:%
%:%288=109%:%
%:%289=110%:%
%:%290=111%:%
%:%291=112%:%
%:%294=113%:%
%:%298=113%:%
%:%299=113%:%
%:%300=114%:%
%:%301=114%:%
%:%302=115%:%
%:%303=115%:%
%:%308=115%:%
%:%311=116%:%
%:%312=117%:%
%:%313=117%:%
%:%314=118%:%
%:%315=119%:%
%:%316=120%:%
%:%319=121%:%
%:%323=121%:%
%:%324=121%:%
%:%325=122%:%
%:%326=122%:%
%:%327=123%:%
%:%328=123%:%
%:%333=123%:%
%:%336=124%:%
%:%337=125%:%
%:%338=125%:%
%:%339=126%:%
%:%340=127%:%
%:%341=128%:%
%:%344=129%:%
%:%348=129%:%
%:%349=129%:%
%:%350=130%:%
%:%351=130%:%
%:%352=131%:%
%:%353=131%:%
%:%358=131%:%
%:%361=132%:%
%:%362=133%:%
%:%363=133%:%
%:%364=134%:%
%:%365=135%:%
%:%366=136%:%
%:%369=137%:%
%:%373=137%:%
%:%374=137%:%
%:%375=138%:%
%:%376=138%:%
%:%377=139%:%
%:%378=139%:%
%:%383=139%:%
%:%386=140%:%
%:%387=141%:%
%:%388=141%:%
%:%389=142%:%
%:%390=143%:%
%:%391=144%:%
%:%398=145%:%
%:%399=145%:%
%:%400=146%:%
%:%401=146%:%
%:%402=146%:%
%:%403=146%:%
%:%404=146%:%
%:%405=147%:%
%:%406=147%:%
%:%407=147%:%
%:%408=148%:%
%:%409=148%:%
%:%410=149%:%
%:%411=149%:%
%:%412=150%:%
%:%413=150%:%
%:%414=151%:%
%:%415=151%:%
%:%416=152%:%
%:%417=152%:%
%:%418=153%:%
%:%419=153%:%
%:%420=154%:%
%:%421=154%:%
%:%422=155%:%
%:%423=155%:%
%:%424=156%:%
%:%425=156%:%
%:%426=157%:%
%:%427=157%:%
%:%428=158%:%
%:%429=158%:%
%:%430=158%:%
%:%431=159%:%
%:%432=159%:%
%:%433=160%:%
%:%434=160%:%
%:%435=161%:%
%:%436=161%:%
%:%437=162%:%
%:%443=162%:%
%:%446=163%:%
%:%447=164%:%
%:%448=164%:%
%:%449=165%:%
%:%450=166%:%
%:%451=167%:%
%:%454=168%:%
%:%458=168%:%
%:%459=168%:%
%:%460=169%:%
%:%461=169%:%
%:%466=169%:%
%:%469=170%:%
%:%470=171%:%
%:%471=171%:%
%:%472=172%:%
%:%473=173%:%
%:%474=174%:%
%:%477=175%:%
%:%481=175%:%
%:%482=175%:%
%:%483=176%:%
%:%484=176%:%
%:%485=177%:%
%:%486=177%:%
%:%491=177%:%
%:%494=178%:%
%:%495=179%:%
%:%496=179%:%
%:%497=180%:%
%:%498=181%:%
%:%499=182%:%
%:%506=183%:%
%:%507=183%:%
%:%508=184%:%
%:%509=184%:%
%:%510=185%:%
%:%511=185%:%
%:%512=186%:%
%:%513=186%:%
%:%514=187%:%
%:%515=187%:%
%:%516=188%:%
%:%517=188%:%
%:%518=189%:%
%:%519=189%:%
%:%520=190%:%
%:%521=190%:%
%:%522=191%:%
%:%523=191%:%
%:%524=192%:%
%:%525=192%:%
%:%526=193%:%
%:%527=193%:%
%:%528=194%:%
%:%529=194%:%
%:%530=195%:%
%:%531=195%:%
%:%532=196%:%
%:%533=196%:%
%:%534=197%:%
%:%535=198%:%
%:%536=198%:%
%:%537=199%:%
%:%538=200%:%
%:%539=200%:%
%:%540=201%:%
%:%541=201%:%
%:%542=202%:%
%:%543=202%:%
%:%544=203%:%
%:%545=203%:%
%:%546=204%:%
%:%547=205%:%
%:%548=205%:%
%:%549=205%:%
%:%550=205%:%
%:%551=206%:%
%:%552=206%:%
%:%553=206%:%
%:%554=206%:%
%:%555=207%:%
%:%556=208%:%
%:%557=208%:%
%:%558=209%:%
%:%559=209%:%
%:%560=210%:%
%:%561=210%:%
%:%562=211%:%
%:%563=211%:%
%:%564=212%:%
%:%565=212%:%
%:%566=213%:%
%:%567=213%:%
%:%568=214%:%
%:%569=214%:%
%:%570=215%:%
%:%571=215%:%
%:%572=216%:%
%:%573=216%:%
%:%574=217%:%
%:%575=217%:%
%:%576=218%:%
%:%577=218%:%
%:%578=218%:%
%:%579=219%:%
%:%580=219%:%
%:%581=219%:%
%:%582=220%:%
%:%583=221%:%
%:%584=221%:%
%:%585=222%:%
%:%586=222%:%
%:%587=223%:%
%:%588=223%:%
%:%589=224%:%
%:%590=224%:%
%:%591=225%:%
%:%592=225%:%
%:%593=226%:%
%:%594=226%:%
%:%595=227%:%
%:%596=227%:%
%:%597=227%:%
%:%598=228%:%
%:%599=228%:%
%:%600=228%:%
%:%601=229%:%
%:%602=230%:%
%:%603=230%:%
%:%604=231%:%
%:%605=231%:%
%:%606=232%:%
%:%607=232%:%
%:%608=233%:%
%:%609=233%:%
%:%610=234%:%
%:%611=235%:%
%:%612=235%:%
%:%613=235%:%
%:%614=236%:%
%:%615=236%:%
%:%616=237%:%
%:%617=237%:%
%:%618=238%:%
%:%619=238%:%
%:%620=239%:%
%:%621=240%:%
%:%622=240%:%
%:%623=241%:%
%:%624=241%:%
%:%625=242%:%
%:%626=242%:%
%:%627=243%:%
%:%628=243%:%
%:%629=244%:%
%:%630=245%:%
%:%631=245%:%
%:%632=246%:%
%:%633=246%:%
%:%634=247%:%
%:%635=247%:%
%:%636=248%:%
%:%637=248%:%
%:%638=249%:%
%:%639=249%:%
%:%640=249%:%
%:%641=250%:%
%:%642=250%:%
%:%643=251%:%
%:%644=251%:%
%:%645=252%:%
%:%646=253%:%
%:%647=253%:%
%:%648=254%:%
%:%649=254%:%
%:%650=255%:%
%:%651=255%:%
%:%652=256%:%
%:%653=256%:%
%:%654=257%:%
%:%655=258%:%
%:%656=258%:%
%:%657=259%:%
%:%658=259%:%
%:%659=260%:%
%:%660=260%:%
%:%661=261%:%
%:%662=261%:%
%:%663=262%:%
%:%664=262%:%
%:%665=263%:%
%:%666=263%:%
%:%667=264%:%
%:%668=264%:%
%:%669=265%:%
%:%670=265%:%
%:%671=266%:%
%:%672=266%:%
%:%673=267%:%
%:%674=267%:%
%:%675=268%:%
%:%676=268%:%
%:%677=269%:%
%:%678=269%:%
%:%679=270%:%
%:%680=270%:%
%:%681=270%:%
%:%682=271%:%
%:%683=271%:%
%:%684=272%:%
%:%685=273%:%
%:%686=273%:%
%:%687=274%:%
%:%688=274%:%
%:%689=275%:%
%:%690=275%:%
%:%691=276%:%
%:%692=276%:%
%:%693=277%:%
%:%694=277%:%
%:%695=278%:%
%:%696=278%:%
%:%697=279%:%
%:%698=280%:%
%:%699=280%:%
%:%700=281%:%
%:%701=281%:%
%:%702=282%:%
%:%703=282%:%
%:%704=283%:%
%:%705=283%:%
%:%706=284%:%
%:%707=284%:%
%:%708=285%:%
%:%709=285%:%
%:%710=286%:%
%:%711=286%:%
%:%712=287%:%
%:%713=287%:%
%:%714=288%:%
%:%715=288%:%
%:%716=289%:%
%:%717=289%:%
%:%718=290%:%
%:%719=290%:%
%:%720=291%:%
%:%721=291%:%
%:%722=292%:%
%:%723=292%:%
%:%724=293%:%
%:%725=293%:%
%:%726=294%:%
%:%727=294%:%
%:%728=295%:%
%:%729=295%:%
%:%730=296%:%
%:%731=296%:%
%:%732=297%:%
%:%733=297%:%
%:%734=298%:%
%:%735=298%:%
%:%736=299%:%
%:%737=299%:%
%:%738=300%:%
%:%739=300%:%
%:%740=301%:%
%:%741=301%:%
%:%742=301%:%
%:%743=302%:%
%:%744=302%:%
%:%745=303%:%
%:%746=303%:%
%:%747=304%:%
%:%748=304%:%
%:%749=305%:%
%:%750=305%:%
%:%751=306%:%
%:%752=307%:%
%:%753=307%:%
%:%754=308%:%
%:%755=308%:%
%:%756=309%:%
%:%757=309%:%
%:%758=310%:%
%:%759=310%:%
%:%760=311%:%
%:%761=311%:%
%:%762=312%:%
%:%763=312%:%
%:%764=313%:%
%:%765=313%:%
%:%766=314%:%
%:%767=314%:%
%:%768=315%:%
%:%769=315%:%
%:%770=316%:%
%:%771=316%:%
%:%772=317%:%
%:%773=317%:%
%:%774=318%:%
%:%775=318%:%
%:%776=319%:%
%:%777=319%:%
%:%778=320%:%
%:%779=320%:%
%:%780=321%:%
%:%781=322%:%
%:%782=322%:%
%:%783=322%:%
%:%784=323%:%
%:%785=323%:%
%:%786=324%:%
%:%787=324%:%
%:%788=325%:%
%:%789=325%:%
%:%790=326%:%
%:%796=326%:%
%:%799=327%:%
%:%800=328%:%
%:%801=328%:%
%:%802=329%:%
%:%803=330%:%
%:%804=331%:%
%:%811=332%:%
%:%812=332%:%
%:%813=333%:%
%:%814=334%:%
%:%815=334%:%
%:%816=335%:%
%:%817=335%:%
%:%818=336%:%
%:%819=336%:%
%:%820=336%:%
%:%821=337%:%
%:%822=337%:%
%:%823=337%:%
%:%824=337%:%
%:%825=337%:%
%:%826=338%:%
%:%827=338%:%
%:%828=338%:%
%:%829=338%:%
%:%830=338%:%
%:%831=339%:%
%:%832=339%:%
%:%833=339%:%
%:%834=339%:%
%:%835=339%:%
%:%836=340%:%
%:%837=340%:%
%:%838=340%:%
%:%839=340%:%
%:%840=340%:%
%:%841=341%:%
%:%842=341%:%
%:%843=342%:%
%:%844=342%:%
%:%845=342%:%
%:%846=343%:%
%:%847=343%:%
%:%848=343%:%
%:%849=343%:%
%:%850=343%:%
%:%851=344%:%
%:%852=344%:%
%:%853=344%:%
%:%854=344%:%
%:%855=344%:%
%:%856=345%:%
%:%857=345%:%
%:%858=345%:%
%:%859=345%:%
%:%860=346%:%
%:%861=346%:%
%:%862=347%:%
%:%863=348%:%
%:%864=348%:%
%:%865=349%:%
%:%866=349%:%
%:%867=350%:%
%:%868=350%:%
%:%869=351%:%
%:%870=351%:%
%:%871=352%:%
%:%872=352%:%
%:%873=353%:%
%:%874=353%:%
%:%875=354%:%
%:%876=354%:%
%:%877=355%:%
%:%878=355%:%
%:%879=356%:%
%:%880=356%:%
%:%881=357%:%
%:%882=357%:%
%:%883=358%:%
%:%884=358%:%
%:%885=359%:%
%:%886=359%:%
%:%887=360%:%
%:%888=360%:%
%:%889=361%:%
%:%890=361%:%
%:%891=362%:%
%:%892=362%:%
%:%893=363%:%
%:%894=363%:%
%:%895=364%:%
%:%896=364%:%
%:%897=365%:%
%:%898=365%:%
%:%899=366%:%
%:%900=367%:%
%:%901=367%:%
%:%902=368%:%
%:%903=368%:%
%:%904=369%:%
%:%905=369%:%
%:%906=370%:%
%:%907=370%:%
%:%908=371%:%
%:%909=371%:%
%:%910=372%:%
%:%911=372%:%
%:%912=373%:%
%:%913=373%:%
%:%914=374%:%
%:%915=375%:%
%:%916=375%:%
%:%917=376%:%
%:%918=376%:%
%:%919=377%:%
%:%920=377%:%
%:%921=378%:%
%:%922=378%:%
%:%923=379%:%
%:%924=380%:%
%:%925=380%:%
%:%926=381%:%
%:%927=382%:%
%:%928=382%:%
%:%929=383%:%
%:%930=383%:%
%:%931=384%:%
%:%932=385%:%
%:%933=385%:%
%:%934=386%:%
%:%935=386%:%
%:%936=387%:%
%:%937=387%:%
%:%938=388%:%
%:%939=388%:%
%:%940=389%:%
%:%941=389%:%
%:%942=390%:%
%:%943=390%:%
%:%944=391%:%
%:%945=391%:%
%:%946=392%:%
%:%947=392%:%
%:%948=393%:%
%:%949=393%:%
%:%950=394%:%
%:%951=394%:%
%:%952=395%:%
%:%953=395%:%
%:%954=396%:%
%:%955=396%:%
%:%956=397%:%
%:%957=397%:%
%:%958=398%:%
%:%959=398%:%
%:%960=399%:%
%:%961=399%:%
%:%962=400%:%
%:%963=400%:%
%:%964=401%:%
%:%965=401%:%
%:%966=402%:%
%:%967=402%:%
%:%968=403%:%
%:%969=403%:%
%:%970=404%:%
%:%971=404%:%
%:%972=405%:%
%:%973=405%:%
%:%974=406%:%
%:%975=406%:%
%:%976=407%:%
%:%977=407%:%
%:%978=408%:%
%:%979=408%:%
%:%980=409%:%
%:%981=409%:%
%:%982=410%:%
%:%983=410%:%
%:%984=411%:%
%:%985=411%:%
%:%986=412%:%
%:%987=412%:%
%:%988=413%:%
%:%989=413%:%
%:%990=414%:%
%:%991=414%:%
%:%992=415%:%
%:%993=415%:%
%:%994=416%:%
%:%995=416%:%
%:%996=417%:%
%:%997=417%:%
%:%998=418%:%
%:%999=418%:%
%:%1000=419%:%
%:%1001=419%:%
%:%1002=420%:%
%:%1003=420%:%
%:%1004=421%:%
%:%1005=421%:%
%:%1006=422%:%
%:%1007=422%:%
%:%1008=423%:%
%:%1009=423%:%
%:%1010=424%:%
%:%1011=424%:%
%:%1012=425%:%
%:%1013=425%:%
%:%1014=426%:%
%:%1015=426%:%
%:%1016=427%:%
%:%1017=427%:%
%:%1018=428%:%
%:%1019=428%:%
%:%1020=429%:%
%:%1021=429%:%
%:%1022=430%:%
%:%1023=430%:%
%:%1024=431%:%
%:%1025=431%:%
%:%1026=432%:%
%:%1027=432%:%
%:%1028=433%:%
%:%1029=433%:%
%:%1030=434%:%
%:%1031=434%:%
%:%1032=435%:%
%:%1033=435%:%
%:%1034=436%:%
%:%1035=436%:%
%:%1036=437%:%
%:%1037=437%:%
%:%1038=438%:%
%:%1039=438%:%
%:%1040=439%:%
%:%1041=439%:%
%:%1042=440%:%
%:%1043=440%:%
%:%1044=441%:%
%:%1045=441%:%
%:%1046=442%:%
%:%1047=442%:%
%:%1048=443%:%
%:%1049=443%:%
%:%1050=444%:%
%:%1051=444%:%
%:%1052=445%:%
%:%1053=445%:%
%:%1054=446%:%
%:%1055=446%:%
%:%1056=447%:%
%:%1057=447%:%
%:%1058=448%:%
%:%1059=448%:%
%:%1060=449%:%
%:%1061=449%:%
%:%1062=450%:%
%:%1063=450%:%
%:%1064=451%:%
%:%1065=451%:%
%:%1066=452%:%
%:%1067=452%:%
%:%1068=453%:%
%:%1069=453%:%
%:%1070=454%:%
%:%1071=454%:%
%:%1072=455%:%
%:%1073=455%:%
%:%1074=456%:%
%:%1075=456%:%
%:%1076=457%:%
%:%1077=457%:%
%:%1078=458%:%
%:%1079=458%:%
%:%1080=459%:%
%:%1081=459%:%
%:%1082=460%:%
%:%1083=460%:%
%:%1084=461%:%
%:%1085=461%:%
%:%1086=462%:%
%:%1087=462%:%
%:%1088=463%:%
%:%1089=463%:%
%:%1090=464%:%
%:%1091=464%:%
%:%1092=465%:%
%:%1093=465%:%
%:%1094=466%:%
%:%1095=466%:%
%:%1096=467%:%
%:%1097=467%:%
%:%1098=468%:%
%:%1099=468%:%
%:%1100=469%:%
%:%1101=469%:%
%:%1102=470%:%
%:%1103=470%:%
%:%1104=471%:%
%:%1105=471%:%
%:%1106=472%:%
%:%1107=472%:%
%:%1108=473%:%
%:%1109=473%:%
%:%1110=474%:%
%:%1111=474%:%
%:%1112=475%:%
%:%1113=475%:%
%:%1114=476%:%
%:%1115=476%:%
%:%1116=477%:%
%:%1117=477%:%
%:%1118=478%:%
%:%1119=478%:%
%:%1120=479%:%
%:%1121=479%:%
%:%1122=480%:%
%:%1123=480%:%
%:%1124=481%:%
%:%1125=481%:%
%:%1126=482%:%
%:%1127=482%:%
%:%1128=482%:%
%:%1129=483%:%
%:%1130=483%:%
%:%1131=484%:%
%:%1132=485%:%
%:%1133=485%:%
%:%1134=486%:%
%:%1135=486%:%
%:%1136=487%:%
%:%1137=487%:%
%:%1138=488%:%
%:%1139=488%:%
%:%1140=489%:%
%:%1141=489%:%
%:%1142=489%:%
%:%1143=489%:%
%:%1144=490%:%
%:%1145=491%:%
%:%1146=491%:%
%:%1147=492%:%
%:%1148=492%:%
%:%1149=493%:%
%:%1150=493%:%
%:%1151=494%:%
%:%1152=494%:%
%:%1153=495%:%
%:%1154=495%:%
%:%1155=496%:%
%:%1156=496%:%
%:%1157=497%:%
%:%1158=497%:%
%:%1159=498%:%
%:%1160=498%:%
%:%1161=499%:%
%:%1167=499%:%
%:%1170=500%:%
%:%1171=501%:%
%:%1172=501%:%
%:%1173=502%:%
%:%1174=503%:%
%:%1175=504%:%
%:%1178=505%:%
%:%1182=505%:%
%:%1183=505%:%
%:%1184=506%:%
%:%1185=506%:%
%:%1186=507%:%
%:%1187=507%:%
%:%1192=507%:%
%:%1195=508%:%
%:%1196=509%:%
%:%1197=509%:%
%:%1198=510%:%
%:%1199=511%:%
%:%1200=512%:%
%:%1203=513%:%
%:%1207=513%:%
%:%1208=513%:%
%:%1209=514%:%
%:%1210=514%:%
%:%1211=515%:%
%:%1212=515%:%
%:%1217=515%:%
%:%1220=516%:%
%:%1221=517%:%
%:%1222=517%:%
%:%1223=518%:%
%:%1224=519%:%
%:%1225=520%:%
%:%1228=521%:%
%:%1229=522%:%
%:%1233=522%:%
%:%1234=522%:%
%:%1235=523%:%
%:%1236=523%:%
%:%1237=524%:%
%:%1238=524%:%
%:%1239=525%:%
%:%1240=525%:%
%:%1241=526%:%
%:%1242=526%:%
%:%1243=527%:%
%:%1244=527%:%
%:%1245=528%:%
%:%1246=528%:%
%:%1247=529%:%
%:%1248=529%:%
%:%1249=530%:%
%:%1250=530%:%
%:%1251=531%:%
%:%1252=531%:%
%:%1253=532%:%
%:%1259=532%:%
%:%1264=533%:%
%:%1269=534%:%