### 2024-11-01 code/output/document/Extensionality_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Extensionality{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Axiom of Extensionality in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Extensionality{\isacharunderscore}{\kern0pt}Axiom\isanewline
\isakeyword{imports}\isanewline
\ \ Names\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data\isanewline
\isakeyword{begin}\isanewline
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ extensionality{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}extensionality{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ y\ z\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ asms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}w{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ w\ {\isasymin}\ x\ {\isasymlongleftrightarrow}\ w\ {\isasymin}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}z{\isasymin}x\ {\isasymlongleftrightarrow}\ z{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymand}\ z{\isasymin}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ z{\isasymin}y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ asms\ transitivity{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}z{\isasymin}x\ {\isasymlongleftrightarrow}\ z{\isasymin}y{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isasymforall}y{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}z{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ z\ {\isasymin}\ x\ {\isasymlongleftrightarrow}\ z\ {\isasymin}\ y{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ extensionality{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \isanewline
\isacommand{end}\isamarkupfalse%
\ \ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Extensionality_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%36=5%:%
%:%39=6%:%
%:%40=7%:%
%:%41=7%:%
%:%42=8%:%
%:%43=9%:%
%:%44=10%:%
%:%45=10%:%
%:%52=11%:%
%:%53=11%:%
%:%54=12%:%
%:%55=12%:%
%:%56=13%:%
%:%57=13%:%
%:%58=14%:%
%:%59=14%:%
%:%60=15%:%
%:%61=16%:%
%:%62=16%:%
%:%63=16%:%
%:%64=17%:%
%:%65=18%:%
%:%66=18%:%
%:%67=18%:%
%:%68=19%:%
%:%69=19%:%
%:%70=19%:%
%:%71=20%:%
%:%72=21%:%
%:%73=21%:%
%:%74=21%:%
%:%75=22%:%
%:%76=22%:%
%:%77=22%:%
%:%78=23%:%
%:%79=23%:%
%:%80=24%:%
%:%81=24%:%
%:%82=25%:%
%:%83=25%:%
%:%84=25%:%
%:%85=26%:%
%:%86=27%:%
%:%87=27%:%
%:%88=28%:%
%:%89=28%:%
%:%90=28%:%
%:%91=28%:%
%:%92=28%:%
%:%93=29%:%
%:%99=29%:%
%:%102=30%:%
%:%103=31%:%
%:%104=31%:%
%:%111=32%:%