### 2024-11-01 code/output/document/Fn_Perm_Automorphism.tex

%
\begin{isabellebody}%
\setisabellecontext{Fn{\isacharunderscore}{\kern0pt}Perm{\isacharunderscore}{\kern0pt}Automorphism}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}Perm{\isacharunderscore}{\kern0pt}Automorphism\ \isanewline
\ \ \isakeyword{imports}\ Fn{\isacharunderscore}{\kern0pt}Perm{\isacharunderscore}{\kern0pt}Definition\ \isanewline
\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ \isakeyword{begin}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}bij\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ converse{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ composition{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ g\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}f\ O\ g\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ comp{\isacharunderscore}{\kern0pt}closed\ comp{\isacharunderscore}{\kern0pt}bij\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}subset\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ p\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}m{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}l{\isasymin}{\isadigit{2}}{\isachardot}{\kern0pt}\ {\isasymlangle}{\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}\ {\isasymin}\ p\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ vin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ l\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isasymin}\ {\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ H\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ p\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\isanewline
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ allI{\isacharcomma}{\kern0pt}\ rule\ allI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharcomma}{\kern0pt}\ rule\ allI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ y\ y{\isacharprime}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ l\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l{\isacharprime}{\kern0pt}\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}\ m{\isacharprime}{\kern0pt}\ l{\isacharprime}{\kern0pt}\ \isakeyword{where}\ H{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ eq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}\ {\isasymand}\ m\ {\isacharequal}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ H{\isacharprime}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ eq{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ H{\isacharprime}{\kern0pt}\ eq\ inj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}l\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\ H\ H{\isacharprime}{\kern0pt}\ eq\ eq{\isacharprime}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ H{\isacharprime}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ p\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ u\ \isakeyword{where}\ uH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ u{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ u{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ \isakeyword{where}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ u{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n\ m{\isachardot}{\kern0pt}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ relation{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ l\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ p\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isacharquery}{\kern0pt}A{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ u\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ u{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ u{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ u{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ u{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ u{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ inj{\isacharunderscore}{\kern0pt}def\ H\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}finite{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isacharquery}{\kern0pt}B{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}finite{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ g\ N\ \isakeyword{where}\ gNH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}g\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ N{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}N\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \ \ \isacommand{define}\isamarkupfalse%
\ h\ \isakeyword{where}\ {\isachardoublequoteopen}h\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ hinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}h\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \isacommand{define}\isamarkupfalse%
\ {\isasymphi}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{8}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{9}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \isacommand{define}\isamarkupfalse%
\ X\ \isakeyword{where}\ {\isachardoublequoteopen}X\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ v\ {\isasymin}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ N{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ f{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ XinM{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}X\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ X{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}notation{\isacharcomma}{\kern0pt}\ rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ gNH\ assms\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\ gNH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}X\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ v\ {\isasymin}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ N{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymexists}gnm\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}nm\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}fnm\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}fn\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}n\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nm\ {\isasymand}\ nm\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ fn\ {\isasymand}\ fnm\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}fn{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymand}\ gnm\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}nm\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}fnm{\isacharcomma}{\kern0pt}\ gnm{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ X{\isacharunderscore}{\kern0pt}def\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ v{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}v\ {\isasymin}\ M\ {\isasymand}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ M\ {\isasymand}\ f\ {\isasymin}\ M\ {\isasymand}\ g\ {\isasymin}\ M\ {\isasymand}\ N\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ succ{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MI\ gNH\ domain{\isacharunderscore}{\kern0pt}closed\ assms\ transM\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ succ{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MI\ gNH\ domain{\isacharunderscore}{\kern0pt}closed\ assms\ transM\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ v\ {\isasymin}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ N{\isachardot}{\kern0pt}\ {\isasymexists}n\ m{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n\ m\ l{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n\ m\ l{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n\ m\ l{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n\ m\ l{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n\ m\ l{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n\ m\ l{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}m{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n\ m\ l{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}n{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ succ{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MI\ gNH\ domain{\isacharunderscore}{\kern0pt}closed\ assms\ transM\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ h{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}h\ {\isasymsubseteq}\ {\isacharbraceleft}{\kern0pt}v\ {\isasymin}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ N\ {\isachardot}{\kern0pt}\ {\isasymexists}n\ m{\isachardot}{\kern0pt}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ h{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ \isakeyword{where}\ nmH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ h{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ fnin{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nmH\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ N{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ gNH\ nmH\ inj{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}v\ {\isasymin}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ N\ {\isachardot}{\kern0pt}\ {\isasymexists}n\ m{\isachardot}{\kern0pt}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nmH\ fnin\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}v\ {\isasymin}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ N\ {\isachardot}{\kern0pt}\ {\isasymexists}n\ m{\isachardot}{\kern0pt}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isasymrangle}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ h{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ h{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\ \isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}h\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ XinM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ htype{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}h\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ N{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ h{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ allI{\isacharcomma}{\kern0pt}\ rule\ allI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharcomma}{\kern0pt}\ rule\ allI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ y\ y{\isacharprime}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ h{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ h{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ h{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}\ m{\isacharprime}{\kern0pt}\ \isakeyword{where}\ H{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ h{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ H{\isacharprime}{\kern0pt}\ inj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ H{\isacharprime}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ domain{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ h{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}h{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ N{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ abH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ h{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isasymin}\ N{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ gNH\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ N{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ abH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}x{\isasymin}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymforall}y{\isasymin}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymlongrightarrow}\ x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ y\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}h\ {\isacharbackquote}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ vx\ \isakeyword{where}\ vxH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ vx{\isachargreater}{\kern0pt}\ {\isasymin}\ h{\isachardoublequoteclose}\ {\isachardoublequoteopen}vx\ {\isasymin}\ N{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ htype\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ vy\ \isakeyword{where}\ vyH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ vy{\isachargreater}{\kern0pt}\ {\isasymin}\ h{\isachardoublequoteclose}\ {\isachardoublequoteopen}vy\ {\isasymin}\ N{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ htype\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ hxeq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}h{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ vx{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ vxH\ htype\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ hyeq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}h{\isacharbackquote}{\kern0pt}y\ {\isacharequal}{\kern0pt}\ vy{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ vyH\ htype\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ \isakeyword{where}\ nmH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}vx\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ vxH\ \isacommand{unfolding}\isamarkupfalse%
\ h{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}\ m{\isacharprime}{\kern0pt}\ \isakeyword{where}\ n{\isacharprime}{\kern0pt}m{\isacharprime}{\kern0pt}H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}vy\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ vyH\ \isacommand{unfolding}\isamarkupfalse%
\ h{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ inj{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ gNH\ assms{\isadigit{1}}\ nmH\ n{\isacharprime}{\kern0pt}m{\isacharprime}{\kern0pt}H\ hxeq\ hyeq\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ nmH\ n{\isacharprime}{\kern0pt}m{\isacharprime}{\kern0pt}H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}finite{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}N\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}h\ \isakeyword{in}\ not{\isacharunderscore}{\kern0pt}emptyI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ hinM\ htype\ gNH\ transM\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isadigit{2}}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isacharquery}{\kern0pt}C{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ u\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}u{\isacharcomma}{\kern0pt}\ v{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isacommand{thm}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}permE\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}m{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}l{\isasymin}{\isadigit{2}}{\isachardot}{\kern0pt}\ {\isasymlangle}{\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}u{\isacharcomma}{\kern0pt}\ v{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}subset\ function{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm\ {\isacartoucheopen}{\isacharquery}{\kern0pt}A{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}B{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}C{\isacartoucheclose}\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}comp\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ f{\isacharprime}{\kern0pt}\ p\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}\ O\ f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l{\isacharprime}{\kern0pt}\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isasymexists}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isasymand}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{4}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{5}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{6}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}\ O\ f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{7}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}\ O\ f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ comp{\isacharunderscore}{\kern0pt}bij{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ inj{\isacharunderscore}{\kern0pt}def\ comp{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}\ O\ f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ I{\isadigit{1}}\ I{\isadigit{2}}\ I{\isadigit{3}}\ I{\isadigit{4}}\ I{\isadigit{5}}\ I{\isadigit{6}}\ I{\isadigit{7}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ p\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p\ {\isacharequal}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ p\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ Fn\ {\isasymrightarrow}\ Fn{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Fn{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}id\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ p\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}p\ {\isasymsubseteq}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isadigit{2}}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}id\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}Fn{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}id{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ domain{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ impI{\isacharcomma}{\kern0pt}\ subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}id{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ sym{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ rule\ idI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ id{\isacharunderscore}{\kern0pt}type\ Pi{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ id{\isacharunderscore}{\kern0pt}bij{\isacharcomma}{\kern0pt}\ rule\ id{\isacharunderscore}{\kern0pt}closed{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}comp\ {\isacharcolon}{\kern0pt}\ \ \isanewline
\ \ \isakeyword{fixes}\ f\ f{\isacharprime}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ O\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f\ O\ f{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ O\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Fn{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ comp{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ domain{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ subst\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ comp{\isacharunderscore}{\kern0pt}bij{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ comp{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ subst\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ simp\ {\isacharcomma}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ comp{\isacharunderscore}{\kern0pt}bij{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ comp{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ domain{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ subst\ domain{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ assms\ Pi{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}bij\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}Fn{\isacharcomma}{\kern0pt}\ Fn{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ g{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ fg{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}bijective{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}comp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ right{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharcomma}{\kern0pt}\ rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}id\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}id\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}comp{\isacharcomma}{\kern0pt}\ rule\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ left{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharcomma}{\kern0pt}\ rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}id\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}id\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}converse\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}Fn{\isacharcomma}{\kern0pt}\ Fn{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}bij\ assms\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ surjE{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}p{\isachardot}{\kern0pt}\ p\ {\isasymin}\ Fn\ {\isasymLongrightarrow}\ {\isasymexists}p{\isacharprime}{\kern0pt}\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ eq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}p{\isachardot}{\kern0pt}\ p\ {\isasymin}\ Fn\ {\isasymLongrightarrow}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ right{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ assms\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}id{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}bij\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ converse{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ assms\ Pi{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isacharprime}{\kern0pt}\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms\ assms\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ surjE\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{4}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isacharprime}{\kern0pt}\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}p{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply\ {\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}comp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ right{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}id{\isacharcomma}{\kern0pt}\ subst\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ rule\ idI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}bij\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{5}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isacharprime}{\kern0pt}\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}p{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ assms\ Pi{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{6}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ assms\ Pi{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}perms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ converse{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ I{\isadigit{1}}\ I{\isadigit{2}}\ I{\isadigit{3}}\ I{\isadigit{4}}\ I{\isadigit{5}}\ I{\isadigit{6}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}order\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ p\ p{\isacharprime}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}p{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasymsubseteq}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}\ {\isasymsubseteq}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isacharprime}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ l\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isasymin}\ {\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}p{\isacharprime}{\kern0pt}\ {\isasymsubseteq}\ p{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}order{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ p\ p{\isacharprime}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}p{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ injE{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}\ n\ m{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}m\ {\isasymLongrightarrow}\ n\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ subsetH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}\ {\isasymsubseteq}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasymsubseteq}\ p{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ p{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasymsubseteq}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymtimes}\ nat{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ vin\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ m\ l\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isasymin}\ {\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ vin\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ vin\ H\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ subsetH\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ assms\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}m{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}l{\isacharprime}{\kern0pt}\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ injE\ H\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ p{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ f\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial{\isachardot}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}Fn{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharcomma}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial{\isachardot}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}bij{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}order{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}order{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Fn_Perm_Automorphism.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%19=4%:%
%:%22=5%:%
%:%23=5%:%
%:%24=6%:%
%:%25=7%:%
%:%26=7%:%
%:%27=8%:%
%:%28=9%:%
%:%29=10%:%
%:%32=11%:%
%:%36=11%:%
%:%37=11%:%
%:%38=12%:%
%:%39=12%:%
%:%40=13%:%
%:%41=13%:%
%:%46=13%:%
%:%49=14%:%
%:%50=15%:%
%:%51=15%:%
%:%52=16%:%
%:%53=17%:%
%:%54=18%:%
%:%57=19%:%
%:%61=19%:%
%:%62=19%:%
%:%63=20%:%
%:%64=20%:%
%:%65=21%:%
%:%66=21%:%
%:%71=21%:%
%:%74=22%:%
%:%75=23%:%
%:%76=23%:%
%:%77=24%:%
%:%78=25%:%
%:%79=26%:%
%:%86=27%:%
%:%87=27%:%
%:%88=28%:%
%:%89=28%:%
%:%90=28%:%
%:%91=29%:%
%:%92=29%:%
%:%93=30%:%
%:%94=30%:%
%:%95=31%:%
%:%96=31%:%
%:%97=32%:%
%:%98=32%:%
%:%99=33%:%
%:%100=33%:%
%:%101=33%:%
%:%102=33%:%
%:%103=34%:%
%:%104=34%:%
%:%105=35%:%
%:%106=35%:%
%:%107=36%:%
%:%108=36%:%
%:%109=37%:%
%:%110=37%:%
%:%111=38%:%
%:%112=38%:%
%:%113=38%:%
%:%114=39%:%
%:%115=39%:%
%:%116=40%:%
%:%117=40%:%
%:%118=41%:%
%:%124=41%:%
%:%127=42%:%
%:%128=43%:%
%:%129=43%:%
%:%130=44%:%
%:%131=45%:%
%:%132=46%:%
%:%135=47%:%
%:%139=47%:%
%:%140=47%:%
%:%141=48%:%
%:%142=48%:%
%:%143=49%:%
%:%144=49%:%
%:%145=50%:%
%:%146=50%:%
%:%147=51%:%
%:%148=51%:%
%:%149=52%:%
%:%150=52%:%
%:%151=53%:%
%:%152=53%:%
%:%153=54%:%
%:%154=54%:%
%:%155=55%:%
%:%156=55%:%
%:%157=55%:%
%:%158=55%:%
%:%159=56%:%
%:%160=57%:%
%:%161=57%:%
%:%162=58%:%
%:%163=58%:%
%:%164=59%:%
%:%165=59%:%
%:%166=60%:%
%:%167=60%:%
%:%168=61%:%
%:%169=61%:%
%:%170=61%:%
%:%171=61%:%
%:%172=62%:%
%:%173=63%:%
%:%174=63%:%
%:%175=63%:%
%:%176=63%:%
%:%177=64%:%
%:%178=65%:%
%:%179=65%:%
%:%180=66%:%
%:%181=66%:%
%:%182=67%:%
%:%183=67%:%
%:%184=68%:%
%:%185=68%:%
%:%186=68%:%
%:%187=68%:%
%:%188=68%:%
%:%189=69%:%
%:%190=70%:%
%:%191=70%:%
%:%192=70%:%
%:%193=70%:%
%:%194=71%:%
%:%195=71%:%
%:%196=71%:%
%:%197=71%:%
%:%198=71%:%
%:%199=72%:%
%:%200=72%:%
%:%201=72%:%
%:%202=72%:%
%:%203=72%:%
%:%204=73%:%
%:%210=73%:%
%:%213=74%:%
%:%214=75%:%
%:%215=75%:%
%:%216=76%:%
%:%217=77%:%
%:%218=78%:%
%:%225=79%:%
%:%226=79%:%
%:%227=80%:%
%:%228=80%:%
%:%229=80%:%
%:%230=81%:%
%:%231=81%:%
%:%232=81%:%
%:%233=81%:%
%:%234=82%:%
%:%235=82%:%
%:%236=82%:%
%:%237=83%:%
%:%238=83%:%
%:%239=84%:%
%:%240=84%:%
%:%241=85%:%
%:%242=85%:%
%:%243=86%:%
%:%244=86%:%
%:%245=86%:%
%:%246=86%:%
%:%247=87%:%
%:%248=87%:%
%:%249=87%:%
%:%250=87%:%
%:%251=88%:%
%:%252=88%:%
%:%253=89%:%
%:%254=89%:%
%:%255=89%:%
%:%256=90%:%
%:%257=90%:%
%:%258=90%:%
%:%259=91%:%
%:%260=91%:%
%:%261=92%:%
%:%262=92%:%
%:%263=93%:%
%:%264=93%:%
%:%265=94%:%
%:%266=94%:%
%:%267=95%:%
%:%268=95%:%
%:%269=95%:%
%:%270=95%:%
%:%271=96%:%
%:%272=96%:%
%:%273=96%:%
%:%274=96%:%
%:%275=97%:%
%:%276=97%:%
%:%277=97%:%
%:%278=98%:%
%:%279=98%:%
%:%280=99%:%
%:%281=99%:%
%:%282=100%:%
%:%283=100%:%
%:%284=100%:%
%:%285=100%:%
%:%286=100%:%
%:%287=101%:%
%:%293=101%:%
%:%296=102%:%
%:%297=103%:%
%:%298=103%:%
%:%299=104%:%
%:%300=105%:%
%:%301=106%:%
%:%308=107%:%
%:%309=107%:%
%:%310=108%:%
%:%311=109%:%
%:%312=109%:%
%:%313=110%:%
%:%314=110%:%
%:%315=111%:%
%:%316=111%:%
%:%317=111%:%
%:%318=112%:%
%:%319=112%:%
%:%320=112%:%
%:%321=112%:%
%:%322=113%:%
%:%323=113%:%
%:%324=113%:%
%:%325=114%:%
%:%326=114%:%
%:%327=115%:%
%:%328=115%:%
%:%329=116%:%
%:%330=116%:%
%:%331=117%:%
%:%332=117%:%
%:%333=117%:%
%:%334=117%:%
%:%335=118%:%
%:%336=118%:%
%:%337=118%:%
%:%338=118%:%
%:%339=119%:%
%:%340=119%:%
%:%341=120%:%
%:%342=120%:%
%:%343=121%:%
%:%344=121%:%
%:%345=122%:%
%:%346=122%:%
%:%347=123%:%
%:%348=123%:%
%:%349=123%:%
%:%350=123%:%
%:%351=123%:%
%:%352=124%:%
%:%353=124%:%
%:%354=125%:%
%:%355=126%:%
%:%356=126%:%
%:%357=127%:%
%:%358=127%:%
%:%359=128%:%
%:%360=128%:%
%:%361=128%:%
%:%362=128%:%
%:%363=129%:%
%:%364=129%:%
%:%365=129%:%
%:%366=129%:%
%:%367=129%:%
%:%368=130%:%
%:%369=131%:%
%:%370=131%:%
%:%371=132%:%
%:%372=133%:%
%:%373=133%:%
%:%374=134%:%
%:%375=134%:%
%:%376=135%:%
%:%377=136%:%
%:%378=136%:%
%:%384=142%:%
%:%385=143%:%
%:%386=143%:%
%:%387=144%:%
%:%388=145%:%
%:%389=145%:%
%:%390=146%:%
%:%391=146%:%
%:%392=147%:%
%:%393=147%:%
%:%394=148%:%
%:%395=148%:%
%:%396=149%:%
%:%397=149%:%
%:%398=150%:%
%:%399=150%:%
%:%400=151%:%
%:%401=151%:%
%:%402=152%:%
%:%403=152%:%
%:%404=153%:%
%:%405=153%:%
%:%406=154%:%
%:%407=154%:%
%:%408=155%:%
%:%409=155%:%
%:%410=156%:%
%:%411=157%:%
%:%412=157%:%
%:%414=159%:%
%:%415=160%:%
%:%416=161%:%
%:%417=161%:%
%:%418=162%:%
%:%419=162%:%
%:%420=163%:%
%:%421=163%:%
%:%422=164%:%
%:%423=164%:%
%:%424=165%:%
%:%425=165%:%
%:%426=166%:%
%:%427=166%:%
%:%428=167%:%
%:%429=167%:%
%:%430=168%:%
%:%431=168%:%
%:%432=169%:%
%:%433=169%:%
%:%434=169%:%
%:%435=170%:%
%:%436=170%:%
%:%437=171%:%
%:%438=171%:%
%:%439=172%:%
%:%440=172%:%
%:%441=173%:%
%:%442=173%:%
%:%443=174%:%
%:%444=174%:%
%:%445=175%:%
%:%446=175%:%
%:%447=176%:%
%:%448=176%:%
%:%449=177%:%
%:%450=177%:%
%:%451=178%:%
%:%452=178%:%
%:%453=179%:%
%:%454=179%:%
%:%455=180%:%
%:%456=180%:%
%:%457=181%:%
%:%458=181%:%
%:%459=182%:%
%:%460=182%:%
%:%461=182%:%
%:%462=183%:%
%:%463=183%:%
%:%464=184%:%
%:%465=184%:%
%:%466=185%:%
%:%467=185%:%
%:%468=186%:%
%:%469=186%:%
%:%470=186%:%
%:%471=187%:%
%:%472=187%:%
%:%473=187%:%
%:%474=188%:%
%:%475=188%:%
%:%476=189%:%
%:%477=189%:%
%:%478=190%:%
%:%479=190%:%
%:%480=191%:%
%:%481=191%:%
%:%482=192%:%
%:%483=192%:%
%:%484=193%:%
%:%485=193%:%
%:%486=194%:%
%:%487=194%:%
%:%488=195%:%
%:%489=195%:%
%:%490=196%:%
%:%491=196%:%
%:%492=196%:%
%:%493=197%:%
%:%494=197%:%
%:%495=198%:%
%:%496=198%:%
%:%497=199%:%
%:%498=199%:%
%:%499=200%:%
%:%500=200%:%
%:%501=200%:%
%:%502=201%:%
%:%503=201%:%
%:%504=202%:%
%:%505=202%:%
%:%506=203%:%
%:%507=203%:%
%:%508=204%:%
%:%509=204%:%
%:%510=205%:%
%:%511=205%:%
%:%512=206%:%
%:%513=206%:%
%:%514=207%:%
%:%515=207%:%
%:%516=208%:%
%:%517=208%:%
%:%518=209%:%
%:%519=209%:%
%:%520=210%:%
%:%521=210%:%
%:%522=210%:%
%:%523=210%:%
%:%524=210%:%
%:%525=211%:%
%:%526=211%:%
%:%527=212%:%
%:%528=213%:%
%:%529=213%:%
%:%530=214%:%
%:%531=214%:%
%:%532=215%:%
%:%533=215%:%
%:%534=216%:%
%:%535=216%:%
%:%536=217%:%
%:%537=217%:%
%:%538=218%:%
%:%539=218%:%
%:%540=219%:%
%:%541=219%:%
%:%542=220%:%
%:%543=220%:%
%:%544=221%:%
%:%545=221%:%
%:%546=222%:%
%:%547=222%:%
%:%548=223%:%
%:%549=223%:%
%:%550=224%:%
%:%551=224%:%
%:%552=225%:%
%:%553=225%:%
%:%554=226%:%
%:%555=226%:%
%:%556=227%:%
%:%557=227%:%
%:%558=228%:%
%:%559=228%:%
%:%560=229%:%
%:%561=229%:%
%:%562=230%:%
%:%563=230%:%
%:%564=231%:%
%:%565=231%:%
%:%566=232%:%
%:%567=232%:%
%:%568=233%:%
%:%569=233%:%
%:%570=234%:%
%:%571=234%:%
%:%572=235%:%
%:%573=235%:%
%:%574=236%:%
%:%575=237%:%
%:%576=237%:%
%:%577=238%:%
%:%578=238%:%
%:%579=239%:%
%:%580=239%:%
%:%581=240%:%
%:%582=240%:%
%:%583=240%:%
%:%584=240%:%
%:%585=240%:%
%:%586=241%:%
%:%587=241%:%
%:%588=241%:%
%:%589=241%:%
%:%590=241%:%
%:%591=242%:%
%:%592=242%:%
%:%593=243%:%
%:%594=243%:%
%:%595=244%:%
%:%596=244%:%
%:%597=245%:%
%:%598=245%:%
%:%599=246%:%
%:%600=246%:%
%:%601=247%:%
%:%602=247%:%
%:%603=248%:%
%:%604=248%:%
%:%605=248%:%
%:%606=249%:%
%:%607=249%:%
%:%608=250%:%
%:%609=250%:%
%:%610=251%:%
%:%611=251%:%
%:%612=252%:%
%:%613=252%:%
%:%614=253%:%
%:%615=253%:%
%:%616=253%:%
%:%617=253%:%
%:%618=254%:%
%:%619=254%:%
%:%620=255%:%
%:%621=255%:%
%:%622=256%:%
%:%623=256%:%
%:%624=257%:%
%:%625=257%:%
%:%626=257%:%
%:%627=258%:%
%:%628=258%:%
%:%629=258%:%
%:%630=259%:%
%:%631=259%:%
%:%632=260%:%
%:%633=260%:%
%:%634=261%:%
%:%635=261%:%
%:%636=262%:%
%:%637=262%:%
%:%638=263%:%
%:%639=263%:%
%:%640=263%:%
%:%641=264%:%
%:%642=264%:%
%:%643=265%:%
%:%644=265%:%
%:%645=266%:%
%:%646=266%:%
%:%647=267%:%
%:%648=267%:%
%:%649=267%:%
%:%650=267%:%
%:%651=267%:%
%:%652=268%:%
%:%653=268%:%
%:%654=269%:%
%:%655=269%:%
%:%656=270%:%
%:%657=271%:%
%:%658=271%:%
%:%659=272%:%
%:%660=272%:%
%:%661=273%:%
%:%662=273%:%
%:%663=274%:%
%:%664=274%:%
%:%665=275%:%
%:%666=275%:%
%:%667=275%:%
%:%668=275%:%
%:%669=276%:%
%:%670=276%:%
%:%671=276%:%
%:%672=276%:%
%:%673=277%:%
%:%674=278%:%
%:%675=278%:%
%:%676=279%:%
%:%677=279%:%
%:%678=280%:%
%:%679=280%:%
%:%680=281%:%
%:%681=281%:%
%:%682=282%:%
%:%683=282%:%
%:%684=283%:%
%:%685=283%:%
%:%686=284%:%
%:%687=284%:%
%:%688=285%:%
%:%689=285%:%
%:%690=286%:%
%:%691=287%:%
%:%692=287%:%
%:%693=288%:%
%:%694=288%:%
%:%695=288%:%
%:%696=289%:%
%:%697=289%:%
%:%698=290%:%
%:%699=290%:%
%:%700=291%:%
%:%701=291%:%
%:%702=292%:%
%:%703=292%:%
%:%704=293%:%
%:%705=293%:%
%:%706=293%:%
%:%707=294%:%
%:%708=294%:%
%:%709=295%:%
%:%710=295%:%
%:%711=296%:%
%:%712=296%:%
%:%713=297%:%
%:%714=298%:%
%:%715=298%:%
%:%716=299%:%
%:%717=299%:%
%:%718=300%:%
%:%719=300%:%
%:%720=301%:%
%:%721=301%:%
%:%722=302%:%
%:%723=302%:%
%:%724=302%:%
%:%725=302%:%
%:%726=302%:%
%:%727=303%:%
%:%728=303%:%
%:%729=304%:%
%:%730=305%:%
%:%731=305%:%
%:%732=305%:%
%:%733=306%:%
%:%734=306%:%
%:%735=307%:%
%:%736=307%:%
%:%737=308%:%
%:%738=308%:%
%:%739=309%:%
%:%740=309%:%
%:%741=310%:%
%:%742=310%:%
%:%743=311%:%
%:%744=311%:%
%:%745=312%:%
%:%746=312%:%
%:%747=313%:%
%:%748=314%:%
%:%749=314%:%
%:%750=315%:%
%:%751=315%:%
%:%752=316%:%
%:%753=316%:%
%:%754=316%:%
%:%755=317%:%
%:%756=317%:%
%:%757=317%:%
%:%758=317%:%
%:%759=317%:%
%:%760=318%:%
%:%761=318%:%
%:%762=318%:%
%:%763=319%:%
%:%764=319%:%
%:%765=320%:%
%:%766=320%:%
%:%767=321%:%
%:%768=321%:%
%:%769=322%:%
%:%770=322%:%
%:%771=322%:%
%:%772=322%:%
%:%773=323%:%
%:%774=323%:%
%:%775=324%:%
%:%776=325%:%
%:%777=325%:%
%:%778=325%:%
%:%779=326%:%
%:%780=326%:%
%:%781=327%:%
%:%782=327%:%
%:%783=328%:%
%:%784=328%:%
%:%785=329%:%
%:%791=329%:%
%:%794=330%:%
%:%795=331%:%
%:%796=331%:%
%:%797=332%:%
%:%798=333%:%
%:%799=334%:%
%:%806=335%:%
%:%807=335%:%
%:%808=336%:%
%:%809=336%:%
%:%810=337%:%
%:%811=337%:%
%:%812=338%:%
%:%813=338%:%
%:%814=339%:%
%:%815=339%:%
%:%816=340%:%
%:%817=340%:%
%:%818=341%:%
%:%819=341%:%
%:%820=342%:%
%:%821=342%:%
%:%822=343%:%
%:%823=343%:%
%:%824=344%:%
%:%825=344%:%
%:%826=345%:%
%:%827=345%:%
%:%828=346%:%
%:%829=346%:%
%:%830=347%:%
%:%831=347%:%
%:%832=348%:%
%:%833=348%:%
%:%834=349%:%
%:%835=349%:%
%:%836=350%:%
%:%837=350%:%
%:%838=351%:%
%:%839=351%:%
%:%840=351%:%
%:%841=352%:%
%:%842=352%:%
%:%843=353%:%
%:%844=353%:%
%:%845=354%:%
%:%846=354%:%
%:%847=355%:%
%:%848=355%:%
%:%849=356%:%
%:%850=356%:%
%:%851=357%:%
%:%852=357%:%
%:%853=358%:%
%:%854=358%:%
%:%855=358%:%
%:%856=359%:%
%:%857=359%:%
%:%858=360%:%
%:%859=360%:%
%:%860=361%:%
%:%861=361%:%
%:%862=362%:%
%:%863=362%:%
%:%864=363%:%
%:%865=363%:%
%:%866=364%:%
%:%867=364%:%
%:%868=365%:%
%:%869=365%:%
%:%870=366%:%
%:%871=366%:%
%:%872=367%:%
%:%873=367%:%
%:%874=368%:%
%:%875=368%:%
%:%876=369%:%
%:%877=369%:%
%:%878=370%:%
%:%879=370%:%
%:%880=371%:%
%:%881=371%:%
%:%882=372%:%
%:%883=372%:%
%:%884=373%:%
%:%885=373%:%
%:%886=374%:%
%:%887=374%:%
%:%888=374%:%
%:%889=374%:%
%:%890=375%:%
%:%896=375%:%
%:%899=376%:%
%:%900=377%:%
%:%901=377%:%
%:%902=378%:%
%:%903=379%:%
%:%904=380%:%
%:%907=381%:%
%:%911=381%:%
%:%912=381%:%
%:%913=382%:%
%:%914=382%:%
%:%915=383%:%
%:%916=383%:%
%:%917=384%:%
%:%918=384%:%
%:%923=384%:%
%:%926=385%:%
%:%927=386%:%
%:%928=386%:%
%:%929=387%:%
%:%930=388%:%
%:%931=389%:%
%:%934=390%:%
%:%935=391%:%
%:%939=391%:%
%:%940=391%:%
%:%941=392%:%
%:%942=392%:%
%:%947=392%:%
%:%950=393%:%
%:%951=394%:%
%:%952=394%:%
%:%953=395%:%
%:%954=396%:%
%:%955=397%:%
%:%958=398%:%
%:%962=398%:%
%:%963=398%:%
%:%964=399%:%
%:%965=399%:%
%:%966=400%:%
%:%967=400%:%
%:%972=400%:%
%:%975=401%:%
%:%976=402%:%
%:%977=402%:%
%:%978=403%:%
%:%979=404%:%
%:%980=405%:%
%:%983=406%:%
%:%987=406%:%
%:%988=406%:%
%:%989=407%:%
%:%990=407%:%
%:%995=407%:%
%:%998=408%:%
%:%999=409%:%
%:%1000=409%:%
%:%1001=410%:%
%:%1002=411%:%
%:%1003=412%:%
%:%1006=413%:%
%:%1010=413%:%
%:%1011=413%:%
%:%1012=414%:%
%:%1013=414%:%
%:%1014=415%:%
%:%1015=415%:%
%:%1016=416%:%
%:%1017=416%:%
%:%1022=416%:%
%:%1025=417%:%
%:%1026=418%:%
%:%1027=418%:%
%:%1028=419%:%
%:%1031=420%:%
%:%1035=420%:%
%:%1036=420%:%
%:%1037=421%:%
%:%1038=421%:%
%:%1039=422%:%
%:%1040=422%:%
%:%1041=423%:%
%:%1042=423%:%
%:%1043=424%:%
%:%1044=424%:%
%:%1045=425%:%
%:%1046=425%:%
%:%1047=426%:%
%:%1048=426%:%
%:%1049=427%:%
%:%1050=427%:%
%:%1051=428%:%
%:%1052=428%:%
%:%1053=429%:%
%:%1054=429%:%
%:%1055=430%:%
%:%1056=430%:%
%:%1057=431%:%
%:%1058=431%:%
%:%1059=432%:%
%:%1060=432%:%
%:%1061=433%:%
%:%1062=433%:%
%:%1067=433%:%
%:%1070=434%:%
%:%1071=435%:%
%:%1072=435%:%
%:%1073=436%:%
%:%1074=437%:%
%:%1075=438%:%
%:%1078=439%:%
%:%1082=439%:%
%:%1083=439%:%
%:%1084=440%:%
%:%1085=440%:%
%:%1086=441%:%
%:%1087=441%:%
%:%1088=442%:%
%:%1089=442%:%
%:%1090=443%:%
%:%1091=443%:%
%:%1092=444%:%
%:%1093=444%:%
%:%1094=445%:%
%:%1095=445%:%
%:%1096=446%:%
%:%1097=446%:%
%:%1098=447%:%
%:%1099=447%:%
%:%1100=448%:%
%:%1101=448%:%
%:%1102=449%:%
%:%1103=449%:%
%:%1104=450%:%
%:%1105=450%:%
%:%1106=451%:%
%:%1107=451%:%
%:%1108=452%:%
%:%1109=452%:%
%:%1110=453%:%
%:%1111=453%:%
%:%1112=454%:%
%:%1113=454%:%
%:%1114=455%:%
%:%1115=455%:%
%:%1116=456%:%
%:%1117=456%:%
%:%1118=457%:%
%:%1119=457%:%
%:%1120=458%:%
%:%1121=458%:%
%:%1122=459%:%
%:%1123=459%:%
%:%1124=460%:%
%:%1125=460%:%
%:%1126=461%:%
%:%1127=461%:%
%:%1128=462%:%
%:%1129=462%:%
%:%1130=463%:%
%:%1131=463%:%
%:%1136=463%:%
%:%1139=464%:%
%:%1140=465%:%
%:%1141=465%:%
%:%1142=466%:%
%:%1143=467%:%
%:%1144=468%:%
%:%1147=469%:%
%:%1151=469%:%
%:%1152=469%:%
%:%1153=470%:%
%:%1154=470%:%
%:%1155=471%:%
%:%1156=471%:%
%:%1157=472%:%
%:%1158=472%:%
%:%1159=473%:%
%:%1160=473%:%
%:%1161=474%:%
%:%1162=474%:%
%:%1163=475%:%
%:%1164=475%:%
%:%1165=476%:%
%:%1166=476%:%
%:%1167=477%:%
%:%1168=477%:%
%:%1169=478%:%
%:%1170=478%:%
%:%1171=479%:%
%:%1172=479%:%
%:%1173=480%:%
%:%1174=480%:%
%:%1175=481%:%
%:%1176=481%:%
%:%1177=482%:%
%:%1183=482%:%
%:%1186=483%:%
%:%1187=484%:%
%:%1188=484%:%
%:%1189=485%:%
%:%1190=486%:%
%:%1191=487%:%
%:%1198=488%:%
%:%1199=488%:%
%:%1200=489%:%
%:%1201=489%:%
%:%1202=490%:%
%:%1203=491%:%
%:%1204=491%:%
%:%1205=492%:%
%:%1206=492%:%
%:%1207=493%:%
%:%1208=493%:%
%:%1209=494%:%
%:%1210=494%:%
%:%1211=494%:%
%:%1212=494%:%
%:%1213=494%:%
%:%1214=495%:%
%:%1215=496%:%
%:%1216=496%:%
%:%1217=497%:%
%:%1218=497%:%
%:%1219=498%:%
%:%1220=498%:%
%:%1221=499%:%
%:%1222=499%:%
%:%1223=500%:%
%:%1224=500%:%
%:%1225=501%:%
%:%1226=501%:%
%:%1227=502%:%
%:%1228=502%:%
%:%1229=503%:%
%:%1230=503%:%
%:%1231=504%:%
%:%1232=504%:%
%:%1233=505%:%
%:%1234=505%:%
%:%1235=506%:%
%:%1236=506%:%
%:%1237=507%:%
%:%1238=507%:%
%:%1239=508%:%
%:%1240=508%:%
%:%1241=509%:%
%:%1242=509%:%
%:%1243=510%:%
%:%1244=510%:%
%:%1245=511%:%
%:%1246=512%:%
%:%1247=512%:%
%:%1248=513%:%
%:%1249=513%:%
%:%1250=514%:%
%:%1251=514%:%
%:%1252=515%:%
%:%1253=515%:%
%:%1254=516%:%
%:%1255=516%:%
%:%1256=517%:%
%:%1257=517%:%
%:%1258=518%:%
%:%1259=518%:%
%:%1260=519%:%
%:%1261=519%:%
%:%1262=520%:%
%:%1263=520%:%
%:%1264=521%:%
%:%1265=521%:%
%:%1266=522%:%
%:%1267=522%:%
%:%1268=523%:%
%:%1269=523%:%
%:%1270=524%:%
%:%1271=524%:%
%:%1272=525%:%
%:%1273=525%:%
%:%1274=526%:%
%:%1275=526%:%
%:%1276=527%:%
%:%1277=527%:%
%:%1278=528%:%
%:%1279=528%:%
%:%1280=529%:%
%:%1281=529%:%
%:%1282=530%:%
%:%1283=530%:%
%:%1284=531%:%
%:%1285=531%:%
%:%1286=532%:%
%:%1287=532%:%
%:%1288=533%:%
%:%1289=533%:%
%:%1290=534%:%
%:%1291=534%:%
%:%1292=535%:%
%:%1293=535%:%
%:%1294=536%:%
%:%1295=536%:%
%:%1296=537%:%
%:%1297=537%:%
%:%1298=538%:%
%:%1299=538%:%
%:%1300=539%:%
%:%1301=539%:%
%:%1302=540%:%
%:%1303=540%:%
%:%1304=541%:%
%:%1310=541%:%
%:%1313=542%:%
%:%1314=543%:%
%:%1315=543%:%
%:%1316=544%:%
%:%1317=545%:%
%:%1318=546%:%
%:%1325=547%:%
%:%1326=547%:%
%:%1327=548%:%
%:%1328=548%:%
%:%1329=548%:%
%:%1330=548%:%
%:%1331=549%:%
%:%1332=550%:%
%:%1333=550%:%
%:%1334=551%:%
%:%1335=551%:%
%:%1336=552%:%
%:%1337=552%:%
%:%1338=553%:%
%:%1339=553%:%
%:%1340=554%:%
%:%1341=554%:%
%:%1342=554%:%
%:%1343=554%:%
%:%1344=554%:%
%:%1345=555%:%
%:%1346=555%:%
%:%1347=555%:%
%:%1348=556%:%
%:%1349=556%:%
%:%1350=557%:%
%:%1351=557%:%
%:%1352=558%:%
%:%1353=558%:%
%:%1354=559%:%
%:%1355=559%:%
%:%1356=559%:%
%:%1357=559%:%
%:%1358=560%:%
%:%1359=560%:%
%:%1360=560%:%
%:%1361=560%:%
%:%1362=560%:%
%:%1363=561%:%
%:%1364=561%:%
%:%1365=561%:%
%:%1366=562%:%
%:%1367=562%:%
%:%1368=563%:%
%:%1369=563%:%
%:%1370=564%:%
%:%1371=564%:%
%:%1372=564%:%
%:%1373=565%:%
%:%1374=565%:%
%:%1375=566%:%
%:%1376=566%:%
%:%1377=567%:%
%:%1378=567%:%
%:%1379=568%:%
%:%1380=568%:%
%:%1381=568%:%
%:%1382=568%:%
%:%1383=568%:%
%:%1384=569%:%
%:%1385=569%:%
%:%1386=570%:%
%:%1387=570%:%
%:%1388=570%:%
%:%1389=571%:%
%:%1390=571%:%
%:%1391=572%:%
%:%1392=572%:%
%:%1393=573%:%
%:%1394=573%:%
%:%1395=574%:%
%:%1396=574%:%
%:%1397=575%:%
%:%1398=575%:%
%:%1399=576%:%
%:%1400=576%:%
%:%1401=577%:%
%:%1407=577%:%
%:%1410=578%:%
%:%1411=579%:%
%:%1412=579%:%
%:%1413=580%:%
%:%1414=581%:%
%:%1415=582%:%
%:%1422=583%:%
%:%1423=583%:%
%:%1424=584%:%
%:%1425=584%:%
%:%1426=584%:%
%:%1427=584%:%
%:%1428=585%:%
%:%1429=585%:%
%:%1430=585%:%
%:%1431=586%:%
%:%1432=586%:%
%:%1433=587%:%
%:%1434=587%:%
%:%1435=588%:%
%:%1436=589%:%
%:%1437=589%:%
%:%1438=590%:%
%:%1439=590%:%
%:%1440=591%:%
%:%1441=591%:%
%:%1442=592%:%
%:%1443=592%:%
%:%1444=593%:%
%:%1445=593%:%
%:%1446=594%:%
%:%1447=594%:%
%:%1448=594%:%
%:%1449=595%:%
%:%1450=595%:%
%:%1451=595%:%
%:%1452=596%:%
%:%1453=596%:%
%:%1454=597%:%
%:%1455=597%:%
%:%1456=598%:%
%:%1457=598%:%
%:%1458=598%:%
%:%1459=599%:%
%:%1460=599%:%
%:%1461=600%:%
%:%1462=600%:%
%:%1463=601%:%
%:%1464=601%:%
%:%1465=601%:%
%:%1466=602%:%
%:%1467=602%:%
%:%1468=603%:%
%:%1469=603%:%
%:%1470=604%:%
%:%1471=604%:%
%:%1472=604%:%
%:%1473=605%:%
%:%1474=605%:%
%:%1475=606%:%
%:%1476=606%:%
%:%1477=607%:%
%:%1478=607%:%
%:%1479=608%:%
%:%1480=608%:%
%:%1481=608%:%
%:%1482=609%:%
%:%1483=609%:%
%:%1484=610%:%
%:%1485=610%:%
%:%1486=611%:%
%:%1487=611%:%
%:%1488=612%:%
%:%1489=612%:%
%:%1490=612%:%
%:%1491=613%:%
%:%1492=613%:%
%:%1493=614%:%
%:%1494=614%:%
%:%1495=615%:%
%:%1496=615%:%
%:%1497=615%:%
%:%1498=616%:%
%:%1499=616%:%
%:%1500=617%:%
%:%1501=617%:%
%:%1502=618%:%
%:%1503=618%:%
%:%1504=618%:%
%:%1505=619%:%
%:%1506=619%:%
%:%1507=620%:%
%:%1508=620%:%
%:%1509=621%:%
%:%1510=621%:%
%:%1511=622%:%
%:%1512=622%:%
%:%1513=622%:%
%:%1514=623%:%
%:%1515=623%:%
%:%1516=624%:%
%:%1517=624%:%
%:%1518=625%:%
%:%1519=625%:%
%:%1520=625%:%
%:%1521=626%:%
%:%1522=626%:%
%:%1523=626%:%
%:%1524=627%:%
%:%1525=627%:%
%:%1526=628%:%
%:%1527=628%:%
%:%1528=628%:%
%:%1529=629%:%
%:%1530=629%:%
%:%1531=630%:%
%:%1532=630%:%
%:%1533=631%:%
%:%1534=631%:%
%:%1535=632%:%
%:%1541=632%:%
%:%1544=633%:%
%:%1545=634%:%
%:%1546=634%:%
%:%1547=635%:%
%:%1548=636%:%
%:%1549=637%:%
%:%1552=638%:%
%:%1553=639%:%
%:%1557=639%:%
%:%1558=639%:%
%:%1559=640%:%
%:%1560=640%:%
%:%1561=641%:%
%:%1562=641%:%
%:%1563=642%:%
%:%1564=642%:%
%:%1565=643%:%
%:%1566=643%:%
%:%1567=644%:%
%:%1568=644%:%
%:%1569=645%:%
%:%1570=645%:%
%:%1571=646%:%
%:%1572=646%:%
%:%1573=647%:%
%:%1574=647%:%
%:%1575=648%:%
%:%1576=648%:%
%:%1577=649%:%
%:%1578=649%:%
%:%1583=649%:%
%:%1586=650%:%
%:%1587=651%:%
%:%1588=652%:%
%:%1589=652%:%
%:%1596=653%:%