### 2024-11-01 code/output/document/Forcing_Data.tex

%
\begin{isabellebody}%
\setisabellecontext{Forcing{\isacharunderscore}{\kern0pt}Data}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Transitive set models of ZF%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
This theory defines the locale \isa{M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans} for
transitive models of ZF, and the associated \isa{forcing{\isacharunderscore}{\kern0pt}data}
 that adds a forcing notion%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Forcing{\isacharunderscore}{\kern0pt}Data\isanewline
\ \ \isakeyword{imports}\ \ \isanewline
\ \ \ \ Forcing{\isacharunderscore}{\kern0pt}Notions\ \isanewline
\ \ \ \ Interface\isanewline
\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \ y{\isasymin}x\ {\isasymLongrightarrow}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ y\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ Transset{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \ \isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF\ {\isacharequal}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ M\ \isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ upair{\isacharunderscore}{\kern0pt}ax{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}upair{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ Union{\isacharunderscore}{\kern0pt}ax{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ power{\isacharunderscore}{\kern0pt}ax{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}power{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ extensionality{\isacharcolon}{\kern0pt}\ \ \ {\isachardoublequoteopen}extensionality{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ foundation{\isacharunderscore}{\kern0pt}ax{\isacharcolon}{\kern0pt}\ \ \ \ {\isachardoublequoteopen}foundation{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ infinity{\isacharunderscore}{\kern0pt}ax{\isacharcolon}{\kern0pt}\ \ \ \ \ \ {\isachardoublequoteopen}infinity{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ separation{\isacharunderscore}{\kern0pt}ax{\isacharcolon}{\kern0pt}\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isakeyword{and}\ replacement{\isacharunderscore}{\kern0pt}ax{\isacharcolon}{\kern0pt}\ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isacharequal}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}ZF\ {\isacharplus}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ enum\isanewline
\ \ \isakeyword{assumes}\ M{\isacharunderscore}{\kern0pt}countable{\isacharcolon}{\kern0pt}\ \ \ \ \ \ {\isachardoublequoteopen}enum{\isasymin}bij{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ trans{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isakeyword{begin}\isanewline
\isacommand{interpretation}\isamarkupfalse%
\ intf{\isacharcolon}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans\ {\isachardoublequoteopen}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans{\isachardot}{\kern0pt}intro\isanewline
\ \ \ \ trans{\isacharunderscore}{\kern0pt}M\ upair{\isacharunderscore}{\kern0pt}ax\ Union{\isacharunderscore}{\kern0pt}ax\ power{\isacharunderscore}{\kern0pt}ax\ extensionality\isanewline
\ \ \ \ foundation{\isacharunderscore}{\kern0pt}ax\ infinity{\isacharunderscore}{\kern0pt}ax\ separation{\isacharunderscore}{\kern0pt}ax{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ replacement{\isacharunderscore}{\kern0pt}ax{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ transitivity\ {\isacharequal}{\kern0pt}\ Transset{\isacharunderscore}{\kern0pt}intf{\isacharbrackleft}{\kern0pt}OF\ trans{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ tuples{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A{\isasymin}M\ {\isasymLongrightarrow}\ B{\isasymin}M\ {\isasymLongrightarrow}\ {\isasymlangle}A{\isacharcomma}{\kern0pt}B{\isasymrangle}{\isasymin}M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}nat\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isasymin}nat\ {\isasymLongrightarrow}\ n{\isasymin}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transitivity\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ mtriv{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}trivial{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}mtriv{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ mtrans{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}trans{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}mtrans{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ mbasic{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}basic{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}mbasic{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ mtrancl{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}trancl{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}mtrancl{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ mdatatypes{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}datatypes{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}mdatatypes{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ meclose{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}meclose{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ meclose{\isacharunderscore}{\kern0pt}pow{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}pow{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ intf{\isachardot}{\kern0pt}meclose{\isacharunderscore}{\kern0pt}pow{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
\isanewline
\isanewline
\isacommand{sublocale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isasymsubseteq}\ M{\isacharunderscore}{\kern0pt}trivial\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}rule\ mtriv{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{sublocale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isasymsubseteq}\ M{\isacharunderscore}{\kern0pt}trans\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}rule\ mtrans{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{sublocale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isasymsubseteq}\ M{\isacharunderscore}{\kern0pt}basic\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}rule\ mbasic{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{sublocale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isasymsubseteq}\ M{\isacharunderscore}{\kern0pt}trancl\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}rule\ mtrancl{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{sublocale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isasymsubseteq}\ M{\isacharunderscore}{\kern0pt}datatypes\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}rule\ mdatatypes{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{sublocale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isasymsubseteq}\ M{\isacharunderscore}{\kern0pt}eclose\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}rule\ meclose{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{sublocale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isasymsubseteq}\ M{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}pow\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}rule\ meclose{\isacharunderscore}{\kern0pt}pow{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\isanewline
\isakeyword{begin}%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{\isa{Collects} in $M$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{lemma}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}{\isadigit{0}}p\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ Qfm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Q{\isacharunderscore}{\kern0pt}fm\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qarty\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qsats\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qabs\ \ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}Q{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}A\ {\isasymLongrightarrow}\ z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ transitivity{\isacharbrackleft}{\kern0pt}of\ z\ A{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}Q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Qabs\ Collect{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Q{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}ax\ Qsats\ Qarty\ Qfm\isanewline
\ \ \ \ \ \ separation{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}y{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}y{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}closed\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}{\isadigit{2}}p\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ Qfm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Q{\isacharunderscore}{\kern0pt}fm\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qarty\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{3}}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ params{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qsats\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qabs\ \ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}A\ {\isasymLongrightarrow}\ z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ transitivity{\isacharbrackleft}{\kern0pt}of\ z\ A{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Qabs\ Collect{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}ax\ Qsats\ Qarty\ Qfm\ params{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ separation{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}closed\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}{\isadigit{4}}p\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ Qfm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Q{\isacharunderscore}{\kern0pt}fm\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qarty\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{5}}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ params{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a{\isadigit{1}}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isadigit{2}}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isadigit{3}}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isadigit{4}}{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qsats\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ Qabs\ \ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}A\ {\isasymLongrightarrow}\ z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ transitivity{\isacharbrackleft}{\kern0pt}of\ z\ A{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Qabs\ Collect{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}ax\ Qsats\ Qarty\ Qfm\ params{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ separation{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}closed\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Repl{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ f{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}f{\isacharunderscore}{\kern0pt}fm\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ f{\isacharunderscore}{\kern0pt}ar{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ fsats{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x\ y{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymLongrightarrow}\ y{\isasymin}M\ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ fabs{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isasymAnd}x\ y{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymLongrightarrow}\ y{\isasymin}M\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ y\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ fclosed{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x{\isasymin}A\ {\isasymLongrightarrow}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \ \isakeyword{and}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharbraceright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ replacement{\isacharunderscore}{\kern0pt}ax\ f{\isacharunderscore}{\kern0pt}fm\ f{\isacharunderscore}{\kern0pt}ar\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fsats\ fabs\ \isanewline
\ \ \ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ y\ {\isachardot}{\kern0pt}\ x{\isasymin}A\ {\isacharcomma}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ fclosed\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ y\ {\isachardot}{\kern0pt}\ x{\isasymin}A\ {\isacharcomma}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{A forcing locale and generic filters%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{locale}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data\ {\isacharequal}{\kern0pt}\ forcing{\isacharunderscore}{\kern0pt}notion\ {\isacharplus}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}ctm\ {\isacharplus}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}P\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}leq\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ transD\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ M\ {\isasymLongrightarrow}\ y\ {\isasymsubseteq}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}unfold\ Transset{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ blast{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}M\ {\isacharequal}{\kern0pt}\ transD{\isacharbrackleft}{\kern0pt}OF\ trans{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ M{\isacharunderscore}{\kern0pt}generic\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}D{\isasymin}M{\isachardot}{\kern0pt}\ D{\isasymsubseteq}P\ {\isasymand}\ dense{\isacharparenleft}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isasymlongrightarrow}D{\isasyminter}G{\isasymnoteq}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}genericD\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x{\isasymin}G\ {\isasymLongrightarrow}\ x{\isasymin}P{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ dest{\isacharcolon}{\kern0pt}filterD{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}leqD\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p{\isasymin}G\ {\isasymLongrightarrow}\ q{\isasymin}P\ {\isasymLongrightarrow}\ p{\isasympreceq}q\ {\isasymLongrightarrow}\ q{\isasymin}G{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ dest{\isacharcolon}{\kern0pt}filter{\isacharunderscore}{\kern0pt}leqD{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}compatD\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p{\isasymin}G\ {\isasymLongrightarrow}\ r{\isasymin}G\ {\isasymLongrightarrow}\ {\isasymexists}q{\isasymin}G{\isachardot}{\kern0pt}\ q{\isasympreceq}p\ {\isasymand}\ q{\isasympreceq}r{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ dest{\isacharcolon}{\kern0pt}low{\isacharunderscore}{\kern0pt}bound{\isacharunderscore}{\kern0pt}filter{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}denseD\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ dense{\isacharparenleft}{\kern0pt}D{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ D{\isasymsubseteq}P\ {\isasymLongrightarrow}\ D{\isasymin}M\ {\isasymLongrightarrow}\ {\isasymexists}q{\isasymin}G{\isachardot}{\kern0pt}\ q{\isasymin}D{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}nonempty{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ G{\isasymnoteq}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isasymsubseteq}P{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}dense\ {\isacartoucheopen}P{\isasymsubseteq}P{\isacartoucheclose}\ \isacommand{show}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}G\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ {\isachardoublequoteopen}one\ {\isasymin}\ G{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}G{\isasymsubseteq}P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isakeyword{and}\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}increasing{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isakeyword{and}\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}G{\isasymsubseteq}P{\isacartoucheclose}\ \isakeyword{and}\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}nonempty\ \isakeyword{and}\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ \isakeyword{and}\ one{\isacharunderscore}{\kern0pt}max\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ increasing{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ G\ {\isasymsubseteq}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{declare}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}trans\ {\isacharbrackleft}{\kern0pt}trans{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}existence{\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}p{\isasymin}P\ {\isasymLongrightarrow}\ {\isasymexists}G{\isachardot}{\kern0pt}\ p{\isasymin}G\ {\isasymand}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}D{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}if\ {\isacharparenleft}{\kern0pt}enum{\isacharbackquote}{\kern0pt}n{\isasymsubseteq}P\ {\isasymand}\ dense{\isacharparenleft}{\kern0pt}enum{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \ then\ enum{\isacharbackquote}{\kern0pt}n\ else\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}D{\isacharbackquote}{\kern0pt}n\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}D{\isacharcolon}{\kern0pt}nat{\isasymrightarrow}Pow{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ lam{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{4}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ dense{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}intro\ ballI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ dense{\isacharparenleft}{\kern0pt}if\ enum{\isacharbackquote}{\kern0pt}n\ {\isasymsubseteq}\ P\ {\isasymand}\ dense{\isacharparenleft}{\kern0pt}enum{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}\ then\ enum{\isacharbackquote}{\kern0pt}n\ else\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ \ {\isacharparenleft}{\kern0pt}{\isasymnot}{\isacharparenleft}{\kern0pt}enum{\isacharbackquote}{\kern0pt}n\ {\isasymsubseteq}\ P\ {\isasymand}\ dense{\isacharparenleft}{\kern0pt}enum{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ dense{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ split{\isacharunderscore}{\kern0pt}if\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}dense\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}D{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ \isakeyword{and}\ Eq{\isadigit{4}}\ \isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ cg{\isacharcolon}{\kern0pt}\ countable{\isacharunderscore}{\kern0pt}generic\ P\ leq\ one\ {\isacharquery}{\kern0pt}D\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}unfold{\isacharunderscore}{\kern0pt}locales{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ G\ \isakeyword{where}\ Eq{\isadigit{6}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}G\ {\isasymand}\ filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasyminter}G{\isasymnoteq}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ cg{\isachardot}{\kern0pt}countable{\isacharunderscore}{\kern0pt}rasiowa{\isacharunderscore}{\kern0pt}sikorski{\isacharbrackleft}{\kern0pt}\isakeyword{where}\ M{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \ P{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ M{\isacharunderscore}{\kern0pt}countable{\isacharbrackleft}{\kern0pt}THEN\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun{\isacharbrackright}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}countable{\isacharbrackleft}{\kern0pt}THEN\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj{\isacharcomma}{\kern0pt}\ THEN\ surj{\isacharunderscore}{\kern0pt}range{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ cg{\isachardot}{\kern0pt}D{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{7}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}D{\isasymin}M{\isachardot}{\kern0pt}\ D{\isasymsubseteq}P\ {\isasymand}\ dense{\isacharparenleft}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isasymlongrightarrow}D{\isasyminter}G{\isasymnoteq}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ ballI\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ D\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}D{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{and}\ Eq{\isadigit{9}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}D\ {\isasymsubseteq}\ P\ {\isasymand}\ dense{\isacharparenleft}{\kern0pt}D{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}y{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}x{\isasymin}nat{\isachardot}{\kern0pt}\ enum{\isacharbackquote}{\kern0pt}x{\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}countable\ \isakeyword{and}\ \ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}D{\isasymin}M{\isacartoucheclose}\ \isacommand{obtain}\isamarkupfalse%
\ n\ \isakeyword{where}\ Eq{\isadigit{1}}{\isadigit{0}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isasymin}nat\ {\isasymand}\ enum{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ D{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ Eq{\isadigit{9}}\ \isakeyword{and}\ if{\isacharunderscore}{\kern0pt}P\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}D{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ D{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ Eq{\isadigit{6}}\ \isakeyword{and}\ Eq{\isadigit{1}}{\isadigit{0}}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}D{\isasyminter}G{\isasymnoteq}{\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ Eq{\isadigit{6}}\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}d{\isasymin}A\ {\isasymLongrightarrow}\ d{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ d\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}d{\isasymin}A\ {\isasymLongrightarrow}\ {\isasymlangle}d{\isacharcomma}{\kern0pt}\ t{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}t{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ t\ d\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ transitivity\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}p{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}r{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}q{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}r{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ A{\isasymin}nat{\isacharsemicolon}{\kern0pt}r{\isasymin}nat{\isacharsemicolon}{\kern0pt}p{\isasymin}nat{\isacharsemicolon}{\kern0pt}q{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}nat{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}p{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Forcing_Data.thy%:%
%:%11=1%:%
%:%23=2%:%
%:%24=3%:%
%:%25=4%:%
%:%33=5%:%
%:%34=5%:%
%:%35=6%:%
%:%36=7%:%
%:%37=8%:%
%:%38=9%:%
%:%39=10%:%
%:%46=10%:%
%:%47=11%:%
%:%48=12%:%
%:%49=12%:%
%:%50=13%:%
%:%53=14%:%
%:%57=14%:%
%:%58=14%:%
%:%63=14%:%
%:%66=15%:%
%:%67=16%:%
%:%68=17%:%
%:%69=17%:%
%:%70=18%:%
%:%71=19%:%
%:%72=20%:%
%:%73=21%:%
%:%74=22%:%
%:%75=23%:%
%:%76=24%:%
%:%77=25%:%
%:%78=26%:%
%:%79=27%:%
%:%80=28%:%
%:%81=29%:%
%:%82=30%:%
%:%83=31%:%
%:%84=31%:%
%:%85=32%:%
%:%86=33%:%
%:%87=34%:%
%:%88=35%:%
%:%89=36%:%
%:%90=37%:%
%:%91=37%:%
%:%94=38%:%
%:%98=38%:%
%:%99=38%:%
%:%100=39%:%
%:%101=40%:%
%:%102=41%:%
%:%103=42%:%
%:%104=42%:%
%:%109=42%:%
%:%112=43%:%
%:%113=44%:%
%:%114=45%:%
%:%115=45%:%
%:%116=46%:%
%:%117=47%:%
%:%118=47%:%
%:%121=48%:%
%:%125=48%:%
%:%126=48%:%
%:%131=48%:%
%:%134=49%:%
%:%135=50%:%
%:%136=50%:%
%:%139=51%:%
%:%143=51%:%
%:%144=51%:%
%:%149=51%:%
%:%152=52%:%
%:%153=53%:%
%:%154=53%:%
%:%157=54%:%
%:%161=54%:%
%:%162=54%:%
%:%167=54%:%
%:%170=55%:%
%:%171=56%:%
%:%172=56%:%
%:%175=57%:%
%:%179=57%:%
%:%180=57%:%
%:%181=57%:%
%:%186=57%:%
%:%189=58%:%
%:%190=59%:%
%:%191=59%:%
%:%194=60%:%
%:%198=60%:%
%:%199=60%:%
%:%204=60%:%
%:%207=61%:%
%:%208=62%:%
%:%209=62%:%
%:%212=63%:%
%:%216=63%:%
%:%217=63%:%
%:%222=63%:%
%:%225=64%:%
%:%226=65%:%
%:%227=65%:%
%:%230=66%:%
%:%234=66%:%
%:%235=66%:%
%:%240=66%:%
%:%243=67%:%
%:%244=68%:%
%:%245=68%:%
%:%248=69%:%
%:%252=69%:%
%:%253=69%:%
%:%258=69%:%
%:%261=70%:%
%:%262=71%:%
%:%263=71%:%
%:%266=72%:%
%:%270=72%:%
%:%271=72%:%
%:%276=72%:%
%:%279=73%:%
%:%280=74%:%
%:%281=74%:%
%:%284=75%:%
%:%288=75%:%
%:%289=75%:%
%:%294=75%:%
%:%297=76%:%
%:%298=77%:%
%:%299=77%:%
%:%302=78%:%
%:%306=78%:%
%:%307=78%:%
%:%312=78%:%
%:%315=79%:%
%:%316=80%:%
%:%317=81%:%
%:%318=82%:%
%:%319=82%:%
%:%320=83%:%
%:%321=84%:%
%:%322=85%:%
%:%323=85%:%
%:%326=86%:%
%:%330=86%:%
%:%331=86%:%
%:%336=86%:%
%:%339=87%:%
%:%340=88%:%
%:%341=88%:%
%:%344=89%:%
%:%348=89%:%
%:%349=89%:%
%:%354=89%:%
%:%357=90%:%
%:%358=91%:%
%:%359=91%:%
%:%362=92%:%
%:%366=92%:%
%:%367=92%:%
%:%372=92%:%
%:%375=93%:%
%:%376=94%:%
%:%377=94%:%
%:%380=95%:%
%:%384=95%:%
%:%385=95%:%
%:%390=95%:%
%:%393=96%:%
%:%394=97%:%
%:%395=97%:%
%:%398=98%:%
%:%402=98%:%
%:%403=98%:%
%:%408=98%:%
%:%411=99%:%
%:%412=100%:%
%:%413=100%:%
%:%416=101%:%
%:%420=101%:%
%:%421=101%:%
%:%426=101%:%
%:%429=102%:%
%:%430=103%:%
%:%431=103%:%
%:%434=104%:%
%:%438=104%:%
%:%439=104%:%
%:%444=104%:%
%:%447=105%:%
%:%448=106%:%
%:%449=107%:%
%:%450=108%:%
%:%451=108%:%
%:%452=109%:%
%:%459=111%:%
%:%469=112%:%
%:%470=112%:%
%:%471=113%:%
%:%472=114%:%
%:%473=115%:%
%:%474=116%:%
%:%475=117%:%
%:%476=118%:%
%:%477=119%:%
%:%478=120%:%
%:%485=121%:%
%:%486=121%:%
%:%487=122%:%
%:%488=122%:%
%:%489=123%:%
%:%490=123%:%
%:%491=123%:%
%:%492=124%:%
%:%493=124%:%
%:%494=125%:%
%:%495=125%:%
%:%496=126%:%
%:%497=126%:%
%:%498=126%:%
%:%499=127%:%
%:%500=127%:%
%:%501=128%:%
%:%502=128%:%
%:%503=129%:%
%:%504=130%:%
%:%505=130%:%
%:%506=131%:%
%:%507=131%:%
%:%508=132%:%
%:%509=132%:%
%:%510=133%:%
%:%511=133%:%
%:%512=133%:%
%:%513=134%:%
%:%514=134%:%
%:%515=135%:%
%:%516=135%:%
%:%517=135%:%
%:%518=135%:%
%:%519=136%:%
%:%525=136%:%
%:%528=137%:%
%:%529=138%:%
%:%530=138%:%
%:%531=139%:%
%:%532=140%:%
%:%533=141%:%
%:%534=142%:%
%:%535=143%:%
%:%536=144%:%
%:%537=145%:%
%:%538=146%:%
%:%539=147%:%
%:%546=148%:%
%:%547=148%:%
%:%548=149%:%
%:%549=149%:%
%:%550=150%:%
%:%551=150%:%
%:%552=150%:%
%:%553=151%:%
%:%554=151%:%
%:%555=152%:%
%:%556=152%:%
%:%557=153%:%
%:%558=153%:%
%:%559=153%:%
%:%560=154%:%
%:%561=154%:%
%:%562=155%:%
%:%563=155%:%
%:%564=156%:%
%:%565=157%:%
%:%566=157%:%
%:%567=158%:%
%:%568=158%:%
%:%569=159%:%
%:%570=159%:%
%:%571=160%:%
%:%572=160%:%
%:%573=160%:%
%:%574=161%:%
%:%575=161%:%
%:%576=162%:%
%:%577=162%:%
%:%578=162%:%
%:%579=162%:%
%:%580=163%:%
%:%586=163%:%
%:%589=164%:%
%:%590=165%:%
%:%591=165%:%
%:%592=166%:%
%:%593=167%:%
%:%594=168%:%
%:%595=169%:%
%:%596=170%:%
%:%597=171%:%
%:%598=172%:%
%:%599=173%:%
%:%600=174%:%
%:%607=175%:%
%:%608=175%:%
%:%609=176%:%
%:%610=176%:%
%:%611=177%:%
%:%612=177%:%
%:%613=177%:%
%:%614=178%:%
%:%615=178%:%
%:%616=179%:%
%:%617=179%:%
%:%618=180%:%
%:%619=180%:%
%:%620=181%:%
%:%621=181%:%
%:%622=182%:%
%:%623=182%:%
%:%624=183%:%
%:%625=183%:%
%:%626=184%:%
%:%627=185%:%
%:%628=186%:%
%:%629=186%:%
%:%630=187%:%
%:%631=187%:%
%:%632=188%:%
%:%633=188%:%
%:%634=189%:%
%:%635=189%:%
%:%636=189%:%
%:%637=190%:%
%:%638=190%:%
%:%639=191%:%
%:%640=191%:%
%:%641=191%:%
%:%642=191%:%
%:%643=192%:%
%:%649=192%:%
%:%652=193%:%
%:%653=194%:%
%:%654=194%:%
%:%655=195%:%
%:%656=196%:%
%:%657=197%:%
%:%658=198%:%
%:%659=199%:%
%:%660=200%:%
%:%661=201%:%
%:%662=202%:%
%:%669=203%:%
%:%670=203%:%
%:%671=204%:%
%:%672=204%:%
%:%673=205%:%
%:%674=205%:%
%:%675=205%:%
%:%676=206%:%
%:%677=206%:%
%:%678=207%:%
%:%679=207%:%
%:%680=208%:%
%:%681=208%:%
%:%682=209%:%
%:%683=210%:%
%:%684=210%:%
%:%685=211%:%
%:%686=211%:%
%:%687=212%:%
%:%688=212%:%
%:%689=213%:%
%:%690=213%:%
%:%691=213%:%
%:%692=214%:%
%:%693=214%:%
%:%694=215%:%
%:%695=215%:%
%:%696=216%:%
%:%697=216%:%
%:%698=217%:%
%:%699=217%:%
%:%700=217%:%
%:%701=217%:%
%:%702=218%:%
%:%708=218%:%
%:%711=219%:%
%:%712=220%:%
%:%720=222%:%
%:%730=223%:%
%:%731=223%:%
%:%732=224%:%
%:%733=225%:%
%:%734=226%:%
%:%735=227%:%
%:%736=228%:%
%:%737=229%:%
%:%738=229%:%
%:%741=230%:%
%:%745=230%:%
%:%746=230%:%
%:%751=230%:%
%:%754=231%:%
%:%755=232%:%
%:%756=233%:%
%:%757=233%:%
%:%758=234%:%
%:%759=235%:%
%:%760=235%:%
%:%761=236%:%
%:%762=237%:%
%:%763=238%:%
%:%764=239%:%
%:%765=239%:%
%:%768=240%:%
%:%772=240%:%
%:%773=240%:%
%:%774=240%:%
%:%779=240%:%
%:%782=241%:%
%:%783=242%:%
%:%784=242%:%
%:%787=243%:%
%:%791=243%:%
%:%792=243%:%
%:%793=243%:%
%:%798=243%:%
%:%801=244%:%
%:%802=245%:%
%:%803=245%:%
%:%806=246%:%
%:%810=246%:%
%:%811=246%:%
%:%812=246%:%
%:%817=246%:%
%:%820=247%:%
%:%821=248%:%
%:%822=248%:%
%:%825=249%:%
%:%829=249%:%
%:%830=249%:%
%:%831=249%:%
%:%836=249%:%
%:%839=250%:%
%:%840=251%:%
%:%841=251%:%
%:%848=252%:%
%:%849=252%:%
%:%850=253%:%
%:%851=253%:%
%:%852=253%:%
%:%853=254%:%
%:%854=254%:%
%:%855=255%:%
%:%856=256%:%
%:%857=256%:%
%:%858=256%:%
%:%859=257%:%
%:%860=258%:%
%:%861=258%:%
%:%862=258%:%
%:%863=259%:%
%:%869=259%:%
%:%872=260%:%
%:%873=261%:%
%:%874=261%:%
%:%875=262%:%
%:%876=263%:%
%:%883=264%:%
%:%884=264%:%
%:%885=265%:%
%:%886=265%:%
%:%887=265%:%
%:%888=266%:%
%:%889=266%:%
%:%890=266%:%
%:%891=267%:%
%:%892=267%:%
%:%893=267%:%
%:%894=268%:%
%:%895=268%:%
%:%896=268%:%
%:%897=269%:%
%:%898=269%:%
%:%899=270%:%
%:%900=270%:%
%:%901=271%:%
%:%902=271%:%
%:%903=272%:%
%:%904=272%:%
%:%905=272%:%
%:%906=273%:%
%:%912=273%:%
%:%915=274%:%
%:%916=275%:%
%:%917=275%:%
%:%920=276%:%
%:%924=276%:%
%:%925=276%:%
%:%926=276%:%
%:%931=276%:%
%:%934=277%:%
%:%935=278%:%
%:%936=278%:%
%:%937=279%:%
%:%938=280%:%
%:%939=280%:%
%:%940=281%:%
%:%947=282%:%
%:%948=282%:%
%:%949=283%:%
%:%950=283%:%
%:%951=284%:%
%:%952=284%:%
%:%953=285%:%
%:%954=285%:%
%:%955=286%:%
%:%956=286%:%
%:%957=287%:%
%:%958=287%:%
%:%959=288%:%
%:%960=288%:%
%:%961=289%:%
%:%962=289%:%
%:%963=289%:%
%:%964=290%:%
%:%965=290%:%
%:%966=291%:%
%:%967=291%:%
%:%968=292%:%
%:%969=292%:%
%:%970=293%:%
%:%971=293%:%
%:%972=294%:%
%:%973=294%:%
%:%974=295%:%
%:%975=295%:%
%:%976=296%:%
%:%977=296%:%
%:%978=297%:%
%:%979=297%:%
%:%980=298%:%
%:%981=298%:%
%:%982=299%:%
%:%983=299%:%
%:%984=299%:%
%:%985=300%:%
%:%986=300%:%
%:%987=301%:%
%:%988=301%:%
%:%989=302%:%
%:%990=302%:%
%:%991=302%:%
%:%992=303%:%
%:%993=303%:%
%:%994=304%:%
%:%995=304%:%
%:%996=305%:%
%:%997=305%:%
%:%998=306%:%
%:%999=306%:%
%:%1000=307%:%
%:%1001=307%:%
%:%1002=308%:%
%:%1003=308%:%
%:%1004=309%:%
%:%1005=309%:%
%:%1006=310%:%
%:%1007=311%:%
%:%1008=311%:%
%:%1009=311%:%
%:%1010=312%:%
%:%1011=312%:%
%:%1012=313%:%
%:%1013=313%:%
%:%1014=314%:%
%:%1015=314%:%
%:%1016=315%:%
%:%1017=315%:%
%:%1018=316%:%
%:%1019=316%:%
%:%1020=317%:%
%:%1021=317%:%
%:%1022=318%:%
%:%1023=318%:%
%:%1024=318%:%
%:%1025=318%:%
%:%1026=319%:%
%:%1027=319%:%
%:%1028=319%:%
%:%1029=320%:%
%:%1030=320%:%
%:%1031=321%:%
%:%1032=321%:%
%:%1033=322%:%
%:%1034=322%:%
%:%1035=322%:%
%:%1036=323%:%
%:%1037=323%:%
%:%1038=324%:%
%:%1039=324%:%
%:%1040=324%:%
%:%1041=325%:%
%:%1042=325%:%
%:%1043=326%:%
%:%1044=326%:%
%:%1045=327%:%
%:%1046=327%:%
%:%1047=327%:%
%:%1048=327%:%
%:%1049=328%:%
%:%1055=328%:%
%:%1058=329%:%
%:%1059=330%:%
%:%1060=331%:%
%:%1061=331%:%
%:%1062=332%:%
%:%1063=333%:%
%:%1064=334%:%
%:%1065=335%:%
%:%1072=336%:%
%:%1073=336%:%
%:%1074=337%:%
%:%1075=337%:%
%:%1076=338%:%
%:%1077=338%:%
%:%1078=338%:%
%:%1079=339%:%
%:%1080=339%:%
%:%1081=339%:%
%:%1082=340%:%
%:%1083=340%:%
%:%1084=341%:%
%:%1085=341%:%
%:%1086=341%:%
%:%1087=342%:%
%:%1088=342%:%
%:%1089=343%:%
%:%1090=343%:%
%:%1091=344%:%
%:%1092=344%:%
%:%1093=345%:%
%:%1094=345%:%
%:%1095=345%:%
%:%1096=346%:%
%:%1102=346%:%
%:%1105=347%:%
%:%1106=348%:%
%:%1107=348%:%
%:%1108=349%:%
%:%1109=350%:%
%:%1112=353%:%
%:%1113=354%:%
%:%1114=355%:%
%:%1115=355%:%
%:%1116=356%:%
%:%1119=357%:%
%:%1123=357%:%
%:%1124=357%:%
%:%1125=357%:%
%:%1130=357%:%
%:%1133=358%:%
%:%1134=359%:%
%:%1135=359%:%
%:%1136=360%:%
%:%1137=361%:%
%:%1138=362%:%
%:%1139=363%:%
%:%1140=364%:%
%:%1143=365%:%
%:%1147=365%:%
%:%1148=365%:%
%:%1149=365%:%
%:%1150=365%:%
%:%1155=365%:%
%:%1158=366%:%
%:%1159=367%:%
%:%1160=367%:%
%:%1163=368%:%
%:%1168=369%:%