### 2024-11-01 code/output/document/Forcing_Main.tex

%
\begin{isabellebody}%
\setisabellecontext{Forcing{\isacharunderscore}{\kern0pt}Main}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The main theorem%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Forcing{\isacharunderscore}{\kern0pt}Main\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ Internal{\isacharunderscore}{\kern0pt}ZFC{\isacharunderscore}{\kern0pt}Axioms\isanewline
\ \ Choice{\isacharunderscore}{\kern0pt}Axiom\isanewline
\ \ Ordinals{\isacharunderscore}{\kern0pt}In{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ Succession{\isacharunderscore}{\kern0pt}Poset\isanewline
\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{The generic extension is countable%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ minimum\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}B{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ THE\ b{\isachardot}{\kern0pt}\ b{\isasymin}B\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isasymin}B{\isachardot}{\kern0pt}\ y\ {\isasymnoteq}\ b\ {\isasymlongrightarrow}\ {\isasymlangle}b{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymin}\ r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}min{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}well{\isacharunderscore}{\kern0pt}ord{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}B\ {\isasymsubseteq}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}B\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}B{\isacharparenright}{\kern0pt}\ {\isasymin}\ B{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}well{\isacharunderscore}{\kern0pt}ord{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}wf{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}wf{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}well{\isacharunderscore}{\kern0pt}ord{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}B{\isasymsubseteq}A{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}wf{\isacharbrackleft}{\kern0pt}B{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Sigma{\isacharunderscore}{\kern0pt}mono\ Int{\isacharunderscore}{\kern0pt}mono\ wf{\isacharunderscore}{\kern0pt}subset\ \isacommand{unfolding}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}\ x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ B\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}z{\isasymin}B{\isachardot}{\kern0pt}\ {\isasymforall}y{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ z{\isasymrangle}\ {\isasymin}\ r{\isasyminter}B{\isasymtimes}B\ {\isasymlongrightarrow}\ y\ {\isasymnotin}\ B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}minimal\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}B{\isasymnoteq}{\isadigit{0}}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ z\ \isakeyword{where}\isanewline
\ \ \ \ B{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z{\isasymin}B\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}z{\isasymrangle}{\isasymin}r{\isasyminter}B{\isasymtimes}B\ {\isasymlongrightarrow}\ y{\isasymnotin}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}B\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isasymin}B{\isachardot}{\kern0pt}\ y\ {\isasymnoteq}\ z\ {\isasymlongrightarrow}\ {\isasymlangle}z{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymin}\ r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}y{\isasymin}B{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymnoteq}z{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}well{\isacharunderscore}{\kern0pt}ord{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ B\ {\isacartoucheopen}B{\isasymsubseteq}A{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}z{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}r{\isacharbar}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}z{\isasymrangle}{\isasymin}r{\isacharbar}{\kern0pt}y{\isacharequal}{\kern0pt}z{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}def\ tot{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}def\ linear{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ B\ {\isacartoucheopen}y{\isasymin}B{\isacartoucheclose}\ {\isacartoucheopen}y{\isasymnoteq}z{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}z{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}r{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharsemicolon}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ B\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ z{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}v{\isasymin}B\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isasymin}B{\isachardot}{\kern0pt}\ y\ {\isasymnoteq}\ v\ {\isasymlongrightarrow}\ {\isasymlangle}v{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymin}\ r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ v\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ B\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}z{\isasymin}B\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isasymin}B{\isachardot}{\kern0pt}\ y\ {\isasymnoteq}\ z\ {\isasymlongrightarrow}\ {\isasymlangle}z{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymin}\ r{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ minimum{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ the{\isacharunderscore}{\kern0pt}equality{\isadigit{2}}{\isacharbrackleft}{\kern0pt}OF\ ex{\isadigit{1}}I{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isasymlambda}x\ {\isachardot}{\kern0pt}x{\isasymin}B\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isasymin}B{\isachardot}{\kern0pt}\ y\ {\isasymnoteq}\ x\ {\isasymlongrightarrow}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymin}\ r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ z{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}lepoll{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}well{\isacharunderscore}{\kern0pt}ord{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}h\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}B\ {\isasymlesssim}\ A{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}f{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}b{\isasymin}B{\isachardot}{\kern0pt}\ minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isasymin}A{\isachardot}{\kern0pt}\ h{\isacharbackquote}{\kern0pt}a{\isacharequal}{\kern0pt}b{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isasymin}\ B\ {\isasymLongrightarrow}\ minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a\ {\isasymin}\ A\ {\isachardot}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ a\ {\isacharequal}{\kern0pt}\ b{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}a{\isasymin}A{\isachardot}{\kern0pt}\ h{\isacharbackquote}{\kern0pt}a{\isacharequal}{\kern0pt}b{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ b\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}b{\isasymin}B{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}h\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}a{\isasymin}A{\isachardot}{\kern0pt}\ h{\isacharbackquote}{\kern0pt}a{\isacharequal}{\kern0pt}b{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}a{\isasymin}A{\isachardot}{\kern0pt}\ h{\isacharbackquote}{\kern0pt}a{\isacharequal}{\kern0pt}b{\isacharbraceright}{\kern0pt}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isasymin}A{\isachardot}{\kern0pt}\ h{\isacharbackquote}{\kern0pt}a{\isacharequal}{\kern0pt}b{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}a{\isasymin}A{\isachardot}{\kern0pt}\ h{\isacharbackquote}{\kern0pt}a{\isacharequal}{\kern0pt}b{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}min\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f\ {\isacharcolon}{\kern0pt}\ B\ {\isasymrightarrow}\ A{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ lam{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ B\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}A{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ w\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymLongrightarrow}\ w\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}w{\isasymin}B{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}B{\isachardoublequoteclose}\ \isakeyword{for}\ w\ x\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ calculation{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}OF\ that{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ calculation{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}OF\ that{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isacharequal}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a\ {\isasymin}\ A\ {\isachardot}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ a\ {\isacharequal}{\kern0pt}\ w{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a\ {\isasymin}\ A\ {\isachardot}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ a\ {\isacharequal}{\kern0pt}\ x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \ \isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ w\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ that\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a\ {\isasymin}\ A\ {\isachardot}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ a\ {\isacharequal}{\kern0pt}\ w{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ minimum{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a\ {\isasymin}\ A\ {\isachardot}{\kern0pt}\ h\ {\isacharbackquote}{\kern0pt}\ a\ {\isacharequal}{\kern0pt}\ x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{4}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}w{\isacharequal}{\kern0pt}x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ lepoll{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ forcing{\isacharunderscore}{\kern0pt}data{\isacharparenright}{\kern0pt}\ surj{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymexists}f{\isachardot}{\kern0pt}\ f\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}f{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}n{\isasymin}nat{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}enum{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ enum\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtD{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ of\ {\isacharunderscore}{\kern0pt}\ G{\isacharbrackright}{\kern0pt}\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun{\isacharbrackleft}{\kern0pt}OF\ M{\isacharunderscore}{\kern0pt}countable{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isacharcolon}{\kern0pt}\ nat\ {\isasymrightarrow}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ lam{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ nat\ {\isachardoublequoteopen}{\isasymlambda}n{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}enum{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ GenExtD{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ G{\isacharbrackright}{\kern0pt}\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj{\isacharbrackleft}{\kern0pt}OF\ M{\isacharunderscore}{\kern0pt}countable{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ G{\isacharunderscore}{\kern0pt}generic{\isacharparenright}{\kern0pt}\ MG{\isacharunderscore}{\kern0pt}eqpoll{\isacharunderscore}{\kern0pt}nat{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymapprox}\ nat{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ MG{\isacharcolon}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}MG\ generic\ pairing{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \isanewline
\ \ \ \ \ \ Union{\isacharunderscore}{\kern0pt}MG\ \ extensionality{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ power{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \ \ \ \ foundation{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ separation{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\ infinity{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ unfold{\isacharunderscore}{\kern0pt}locales\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ f\ \isakeyword{where}\ {\isachardoublequoteopen}f\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymlesssim}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}surj{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}lepoll\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}Memrel{\isacharbrackleft}{\kern0pt}of\ nat{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nat\ {\isasymlesssim}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ MG{\isachardot}{\kern0pt}nat{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}M\ subset{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}lepoll\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ eqpollI\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{The main result%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{theorem}\isamarkupfalse%
\ extensions{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}ctms{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M\ {\isasymapprox}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M\ {\isasymTurnstile}\ ZF{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymexists}N{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ M\ {\isasymsubseteq}\ N\ {\isasymand}\ N\ {\isasymapprox}\ nat\ {\isasymand}\ Transset{\isacharparenleft}{\kern0pt}N{\isacharparenright}{\kern0pt}\ {\isasymand}\ N\ {\isasymTurnstile}\ ZF\ {\isasymand}\ M{\isasymnoteq}N\ {\isasymand}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}{\isasymalpha}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymalpha}\ {\isasymin}\ M\ {\isasymlongleftrightarrow}\ {\isasymalpha}\ {\isasymin}\ N{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isasymTurnstile}\ AC\ {\isasymlongrightarrow}\ N\ {\isasymTurnstile}\ ZFC{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}M\ {\isasymapprox}\ nat{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ enum\ \isakeyword{where}\ {\isachardoublequoteopen}enum\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ eqpoll{\isacharunderscore}{\kern0pt}sym\ \isacommand{unfolding}\isamarkupfalse%
\ eqpoll{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ M\ enum\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}satT\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ intro{\isacharunderscore}{\kern0pt}locales\ {\isacharparenleft}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all\ add{\isacharcolon}{\kern0pt}M{\isacharunderscore}{\kern0pt}ctm{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ ctm{\isacharunderscore}{\kern0pt}separative\ {\isachardoublequoteopen}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ seqle\ {\isadigit{0}}\ M\ enum\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}unfold{\isacharunderscore}{\kern0pt}locales{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ f\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}q\ {\isasympreceq}s\ f\ {\isasymand}\ {\isacharquery}{\kern0pt}r\ {\isasympreceq}s\ f\ {\isasymand}\ {\isacharquery}{\kern0pt}q\ {\isasymbottom}s\ {\isacharquery}{\kern0pt}r{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ upd{\isacharunderscore}{\kern0pt}leI\ seqspace{\isacharunderscore}{\kern0pt}separative\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}q\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}r\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ f\ {\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}q{\isasymin}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardot}{\kern0pt}\ \ {\isasymexists}r{\isasymin}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardot}{\kern0pt}\ q\ {\isasympreceq}s\ f\ {\isasymand}\ r\ {\isasympreceq}s\ f\ {\isasymand}\ q\ {\isasymbottom}s\ r{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ bexI{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\ %
\isamarkupcmt{why the heck auto-tools don't solve this?%
}\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}M\ seqspace{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}seqle\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ seqle{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ cohen{\isacharunderscore}{\kern0pt}extension{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}proper\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ G\ \isakeyword{where}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M\ {\isasymnoteq}\ GenExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}M{\isasymnoteq}{\isacharquery}{\kern0pt}N{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\ {\isachardoublequoteopen}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ seqle\ {\isadigit{0}}\ {\isacharunderscore}{\kern0pt}\ enum\ G\ \isacommand{by}\isamarkupfalse%
\ unfold{\isacharunderscore}{\kern0pt}locales\isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ MG{\isacharcolon}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}ZF\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}N{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic\ pairing{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \isanewline
\ \ \ \ \ \ Union{\isacharunderscore}{\kern0pt}MG\ \ extensionality{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ power{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \ \ \ \ foundation{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ separation{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\ infinity{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ unfold{\isacharunderscore}{\kern0pt}locales\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}N\ {\isasymTurnstile}\ ZF{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}satT{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}N{\isacharbrackright}{\kern0pt}\ MG{\isachardot}{\kern0pt}M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}axioms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isasymTurnstile}\ AC\ {\isasymLongrightarrow}\ {\isacharquery}{\kern0pt}N\ {\isasymTurnstile}\ ZFC{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ AC{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}choice{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}auto\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}choice{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharquery}{\kern0pt}N{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ choice{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}N\ {\isasymTurnstile}\ ZF{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}N\ {\isasymTurnstile}\ ZFC{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}auto\ sats{\isacharunderscore}{\kern0pt}ZFC{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}AC\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}M\ {\isasymnoteq}\ {\isacharquery}{\kern0pt}N{\isacartoucheclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}N{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}MG\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M\ {\isasymsubseteq}\ {\isacharquery}{\kern0pt}N{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}MG{\isacharbrackleft}{\kern0pt}OF\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ generic\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}MG{\isacharunderscore}{\kern0pt}iff\ MG{\isacharunderscore}{\kern0pt}eqpoll{\isacharunderscore}{\kern0pt}nat\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}N{\isachardoublequoteclose}\ \isakeyword{in}\ exI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Forcing_Main.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%32=6%:%
%:%33=7%:%
%:%34=8%:%
%:%35=9%:%
%:%49=11%:%
%:%59=19%:%
%:%60=19%:%
%:%61=20%:%
%:%62=21%:%
%:%63=22%:%
%:%64=23%:%
%:%65=23%:%
%:%66=24%:%
%:%67=25%:%
%:%68=26%:%
%:%69=27%:%
%:%76=28%:%
%:%77=28%:%
%:%78=29%:%
%:%79=29%:%
%:%80=30%:%
%:%81=30%:%
%:%82=31%:%
%:%83=31%:%
%:%84=31%:%
%:%85=32%:%
%:%86=32%:%
%:%87=33%:%
%:%88=33%:%
%:%89=34%:%
%:%90=34%:%
%:%91=34%:%
%:%92=34%:%
%:%93=35%:%
%:%94=35%:%
%:%95=36%:%
%:%96=36%:%
%:%97=37%:%
%:%98=37%:%
%:%99=37%:%
%:%100=38%:%
%:%101=38%:%
%:%102=39%:%
%:%103=39%:%
%:%104=40%:%
%:%105=40%:%
%:%106=41%:%
%:%107=42%:%
%:%108=42%:%
%:%109=43%:%
%:%110=43%:%
%:%111=44%:%
%:%112=44%:%
%:%113=45%:%
%:%114=45%:%
%:%115=46%:%
%:%116=46%:%
%:%117=47%:%
%:%118=47%:%
%:%119=48%:%
%:%120=48%:%
%:%121=49%:%
%:%122=49%:%
%:%123=50%:%
%:%124=50%:%
%:%125=51%:%
%:%126=51%:%
%:%127=51%:%
%:%128=52%:%
%:%129=52%:%
%:%130=53%:%
%:%131=53%:%
%:%132=54%:%
%:%133=54%:%
%:%134=55%:%
%:%135=55%:%
%:%136=56%:%
%:%137=56%:%
%:%138=57%:%
%:%139=57%:%
%:%140=57%:%
%:%141=58%:%
%:%142=58%:%
%:%143=59%:%
%:%144=59%:%
%:%145=60%:%
%:%146=60%:%
%:%147=60%:%
%:%148=61%:%
%:%149=61%:%
%:%150=62%:%
%:%151=62%:%
%:%152=63%:%
%:%153=63%:%
%:%154=64%:%
%:%155=64%:%
%:%156=65%:%
%:%157=65%:%
%:%158=66%:%
%:%164=66%:%
%:%167=67%:%
%:%168=68%:%
%:%169=68%:%
%:%170=69%:%
%:%171=70%:%
%:%178=71%:%
%:%179=71%:%
%:%180=72%:%
%:%181=72%:%
%:%182=73%:%
%:%183=73%:%
%:%184=74%:%
%:%185=74%:%
%:%186=75%:%
%:%187=75%:%
%:%188=76%:%
%:%189=76%:%
%:%190=77%:%
%:%191=77%:%
%:%192=78%:%
%:%193=78%:%
%:%194=79%:%
%:%195=79%:%
%:%196=79%:%
%:%197=80%:%
%:%198=80%:%
%:%199=81%:%
%:%200=81%:%
%:%201=82%:%
%:%202=82%:%
%:%203=83%:%
%:%204=83%:%
%:%205=84%:%
%:%206=84%:%
%:%207=85%:%
%:%208=85%:%
%:%209=85%:%
%:%210=86%:%
%:%211=86%:%
%:%212=87%:%
%:%213=87%:%
%:%214=87%:%
%:%215=88%:%
%:%216=88%:%
%:%217=89%:%
%:%218=89%:%
%:%219=89%:%
%:%220=90%:%
%:%221=90%:%
%:%222=91%:%
%:%223=91%:%
%:%224=92%:%
%:%225=92%:%
%:%226=93%:%
%:%227=93%:%
%:%228=94%:%
%:%229=94%:%
%:%230=95%:%
%:%231=96%:%
%:%232=96%:%
%:%233=97%:%
%:%234=97%:%
%:%235=98%:%
%:%236=98%:%
%:%237=99%:%
%:%238=99%:%
%:%239=99%:%
%:%240=100%:%
%:%241=100%:%
%:%242=101%:%
%:%243=101%:%
%:%244=102%:%
%:%245=102%:%
%:%246=102%:%
%:%247=103%:%
%:%248=103%:%
%:%249=103%:%
%:%250=104%:%
%:%251=104%:%
%:%252=105%:%
%:%253=105%:%
%:%254=106%:%
%:%255=106%:%
%:%256=107%:%
%:%257=107%:%
%:%258=107%:%
%:%259=108%:%
%:%265=108%:%
%:%268=109%:%
%:%269=110%:%
%:%270=110%:%
%:%271=111%:%
%:%278=112%:%
%:%279=112%:%
%:%280=113%:%
%:%281=113%:%
%:%282=114%:%
%:%283=114%:%
%:%284=115%:%
%:%285=115%:%
%:%286=115%:%
%:%287=116%:%
%:%288=116%:%
%:%289=117%:%
%:%290=117%:%
%:%291=118%:%
%:%292=118%:%
%:%293=118%:%
%:%294=119%:%
%:%295=119%:%
%:%296=120%:%
%:%297=120%:%
%:%298=121%:%
%:%299=121%:%
%:%300=122%:%
%:%301=122%:%
%:%302=122%:%
%:%303=123%:%
%:%304=123%:%
%:%305=124%:%
%:%306=124%:%
%:%307=125%:%
%:%308=125%:%
%:%309=125%:%
%:%310=126%:%
%:%316=126%:%
%:%319=127%:%
%:%320=128%:%
%:%321=128%:%
%:%328=129%:%
%:%329=129%:%
%:%330=130%:%
%:%331=130%:%
%:%332=131%:%
%:%333=131%:%
%:%334=132%:%
%:%335=133%:%
%:%336=134%:%
%:%337=135%:%
%:%338=135%:%
%:%339=136%:%
%:%340=136%:%
%:%341=137%:%
%:%342=137%:%
%:%343=137%:%
%:%344=138%:%
%:%345=138%:%
%:%346=139%:%
%:%347=139%:%
%:%348=140%:%
%:%349=140%:%
%:%350=141%:%
%:%351=141%:%
%:%352=142%:%
%:%353=142%:%
%:%354=143%:%
%:%355=143%:%
%:%356=144%:%
%:%357=144%:%
%:%358=144%:%
%:%359=145%:%
%:%360=145%:%
%:%361=146%:%
%:%362=146%:%
%:%363=146%:%
%:%364=147%:%
%:%365=147%:%
%:%366=148%:%
%:%381=150%:%
%:%391=152%:%
%:%392=152%:%
%:%393=153%:%
%:%394=154%:%
%:%395=155%:%
%:%396=156%:%
%:%399=159%:%
%:%406=160%:%
%:%407=160%:%
%:%408=161%:%
%:%409=161%:%
%:%410=162%:%
%:%411=162%:%
%:%412=163%:%
%:%413=163%:%
%:%414=163%:%
%:%415=163%:%
%:%416=164%:%
%:%417=164%:%
%:%418=165%:%
%:%419=165%:%
%:%420=166%:%
%:%421=166%:%
%:%422=167%:%
%:%423=167%:%
%:%424=168%:%
%:%425=168%:%
%:%426=169%:%
%:%427=169%:%
%:%428=170%:%
%:%429=170%:%
%:%430=171%:%
%:%431=171%:%
%:%432=172%:%
%:%433=172%:%
%:%434=173%:%
%:%435=173%:%
%:%436=174%:%
%:%437=174%:%
%:%438=175%:%
%:%439=175%:%
%:%440=175%:%
%:%441=176%:%
%:%442=176%:%
%:%443=176%:%
%:%444=177%:%
%:%445=177%:%
%:%446=178%:%
%:%447=178%:%
%:%448=178%:%
%:%449=179%:%
%:%450=179%:%
%:%451=180%:%
%:%452=180%:%
%:%453=181%:%
%:%454=181%:%
%:%455=181%:%
%:%456=181%:%
%:%457=182%:%
%:%458=182%:%
%:%459=183%:%
%:%460=183%:%
%:%461=183%:%
%:%462=183%:%
%:%463=184%:%
%:%464=184%:%
%:%465=185%:%
%:%466=185%:%
%:%467=185%:%
%:%468=185%:%
%:%469=186%:%
%:%470=186%:%
%:%471=187%:%
%:%472=187%:%
%:%473=188%:%
%:%474=188%:%
%:%475=189%:%
%:%476=190%:%
%:%477=190%:%
%:%478=191%:%
%:%479=191%:%
%:%480=192%:%
%:%481=192%:%
%:%482=192%:%
%:%483=193%:%
%:%484=193%:%
%:%485=194%:%
%:%486=194%:%
%:%487=195%:%
%:%488=196%:%
%:%489=197%:%
%:%490=198%:%
%:%491=198%:%
%:%492=199%:%
%:%493=199%:%
%:%494=200%:%
%:%495=200%:%
%:%496=200%:%
%:%497=201%:%
%:%498=201%:%
%:%499=202%:%
%:%500=202%:%
%:%501=203%:%
%:%502=203%:%
%:%503=204%:%
%:%504=204%:%
%:%505=205%:%
%:%506=205%:%
%:%507=206%:%
%:%508=206%:%
%:%509=207%:%
%:%510=207%:%
%:%511=207%:%
%:%512=207%:%
%:%513=208%:%
%:%514=208%:%
%:%515=209%:%
%:%516=209%:%
%:%517=209%:%
%:%518=209%:%
%:%519=210%:%
%:%520=210%:%
%:%521=211%:%
%:%522=211%:%
%:%523=212%:%
%:%524=212%:%
%:%525=213%:%
%:%526=213%:%
%:%527=213%:%
%:%528=214%:%
%:%529=214%:%
%:%530=215%:%
%:%531=215%:%
%:%532=216%:%
%:%533=216%:%
%:%534=217%:%
%:%535=217%:%
%:%536=218%:%
%:%537=218%:%
%:%538=218%:%
%:%539=218%:%
%:%540=219%:%
%:%541=219%:%
%:%542=220%:%
%:%543=220%:%
%:%544=220%:%
%:%545=220%:%
%:%546=221%:%
%:%547=221%:%
%:%548=222%:%
%:%549=222%:%
%:%550=223%:%
%:%551=223%:%
%:%552=224%:%
%:%553=224%:%
%:%554=225%:%
%:%560=225%:%
%:%565=226%:%
%:%570=227%:%