### 2024-11-01 code/output/document/Forcing_Notions.tex

%
\begin{isabellebody}%
\setisabellecontext{Forcing{\isacharunderscore}{\kern0pt}Notions}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Forcing notions%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
This theory defines a locale for forcing notions, that is,
 preorders with a distinguished maximum element.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Forcing{\isacharunderscore}{\kern0pt}Notions\isanewline
\ \ \isakeyword{imports}\ {\isachardoublequoteopen}ZF{\isacharminus}{\kern0pt}Constructible{\isachardot}{\kern0pt}Relative{\isachardoublequoteclose}\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Basic concepts%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
We say that two elements $p,q$ are
  \emph{compatible} if they have a lower bound in $P$%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}d{\isasymin}A\ {\isachardot}{\kern0pt}\ {\isasymlangle}d{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}r\ {\isasymand}\ {\isasymlangle}d{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}r{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}in\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}d{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ d{\isasymin}A\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}dp{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}d{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}dp{\isacharparenright}{\kern0pt}\ {\isasymand}\ dp{\isasymin}r\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}dq{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}d{\isacharcomma}{\kern0pt}q{\isacharcomma}{\kern0pt}dq{\isacharparenright}{\kern0pt}\ {\isasymand}\ dq{\isasymin}r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}inI\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ d{\isasymin}A\ {\isacharsemicolon}{\kern0pt}\ {\isasymlangle}d{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}r\ {\isacharsemicolon}{\kern0pt}\ {\isasymlangle}d{\isacharcomma}{\kern0pt}g{\isasymrangle}{\isasymin}r\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ refl{\isacharunderscore}{\kern0pt}compat{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ refl{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ {\isasymlangle}p{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ r\ {\isacharbar}{\kern0pt}\ p{\isacharequal}{\kern0pt}q\ {\isacharbar}{\kern0pt}\ {\isasymlangle}q{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ r\ {\isacharsemicolon}{\kern0pt}\ p{\isasymin}A\ {\isacharsemicolon}{\kern0pt}\ q{\isasymin}A{\isasymrbrakk}\ {\isasymLongrightarrow}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ refl{\isacharunderscore}{\kern0pt}def\ compat{\isacharunderscore}{\kern0pt}inI{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ chain{\isacharunderscore}{\kern0pt}compat{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}refl{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ linear{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \ {\isacharparenleft}{\kern0pt}{\isasymforall}p{\isasymin}A{\isachardot}{\kern0pt}{\isasymforall}q{\isasymin}A{\isachardot}{\kern0pt}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ refl{\isacharunderscore}{\kern0pt}compat\ linear{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ subset{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}image{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharcolon}{\kern0pt}N{\isasymrightarrow}P\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}N{\isasymsubseteq}P{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ image{\isacharunderscore}{\kern0pt}fun\ apply{\isacharunderscore}{\kern0pt}funtype{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ refl{\isacharunderscore}{\kern0pt}monot{\isacharunderscore}{\kern0pt}domain{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}refl{\isacharparenleft}{\kern0pt}B{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ A{\isasymsubseteq}B\ {\isasymLongrightarrow}\ refl{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ refl{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ antichain\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}antichain{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ A{\isasymsubseteq}P\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}p{\isasymin}A{\isachardot}{\kern0pt}{\isasymforall}q{\isasymin}A{\isachardot}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymnot}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ ccc\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ccc{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymforall}A{\isachardot}{\kern0pt}\ antichain{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharbar}{\kern0pt}A{\isacharbar}{\kern0pt}\ {\isasymle}\ nat{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}notion\ {\isacharequal}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ P\ leq\ one\isanewline
\ \ \isakeyword{assumes}\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}one\ {\isasymin}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ leq{\isacharunderscore}{\kern0pt}preord{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ {\isachardoublequoteopen}preorder{\isacharunderscore}{\kern0pt}on{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ one{\isacharunderscore}{\kern0pt}max{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymlangle}p{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isasymin}leq{\isachardoublequoteclose}\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ Leq\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}\isakeyword{infixl}\ {\isachardoublequoteopen}{\isasympreceq}{\isachardoublequoteclose}\ {\isadigit{5}}{\isadigit{0}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}x\ {\isasympreceq}\ y\ {\isasymequiv}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}leq{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ refl{\isacharunderscore}{\kern0pt}leq{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}r{\isasymin}P\ {\isasymLongrightarrow}\ r{\isasympreceq}r{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}preord\ \isacommand{unfolding}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ refl{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
A set $D$ is \emph{dense} if every element $p\in P$ has a lower 
bound in $D$.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ dense\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}D{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymforall}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymexists}d{\isasymin}D\ {\isachardot}{\kern0pt}\ d{\isasympreceq}p{\isachardoublequoteclose}%
\begin{isamarkuptext}%
There is also a weaker definition which asks for 
a lower bound in $D$ only for the elements below some fixed 
element $q$.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ dense{\isacharunderscore}{\kern0pt}below\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymforall}p{\isasymin}P{\isachardot}{\kern0pt}\ p{\isasympreceq}q\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}d{\isasymin}D{\isachardot}{\kern0pt}\ d{\isasymin}P\ {\isasymand}\ d{\isasympreceq}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}dense{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ leq{\isacharunderscore}{\kern0pt}preord{\isacharcomma}{\kern0pt}\ auto\ simp\ add{\isacharcolon}{\kern0pt}\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ refl{\isacharunderscore}{\kern0pt}def\ dense{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ increasing\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}increasing{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymforall}x{\isasymin}F{\isachardot}{\kern0pt}\ {\isasymforall}\ p\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ x{\isasympreceq}p\ {\isasymlongrightarrow}\ p{\isasymin}F{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ compat\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}compat{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}transD{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}a{\isasympreceq}b\ {\isasymLongrightarrow}\ b{\isasympreceq}c\ {\isasymLongrightarrow}\ a\ {\isasymin}\ P{\isasymLongrightarrow}\ b\ {\isasymin}\ P{\isasymLongrightarrow}\ c\ {\isasymin}\ P{\isasymLongrightarrow}\ a{\isasympreceq}c{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}preord\ trans{\isacharunderscore}{\kern0pt}onD\ \isacommand{unfolding}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}transD{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}A{\isasymsubseteq}P\ {\isasymLongrightarrow}\ a{\isasympreceq}b\ {\isasymLongrightarrow}\ b{\isasympreceq}c\ {\isasymLongrightarrow}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ b\ {\isasymin}\ P{\isasymLongrightarrow}\ c\ {\isasymin}\ P{\isasymLongrightarrow}\ a{\isasympreceq}c{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}preord\ trans{\isacharunderscore}{\kern0pt}onD\ subsetD\ \isacommand{unfolding}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}reflI{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}P\ {\isasymLongrightarrow}\ p{\isasympreceq}p{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}preord\ \isacommand{unfolding}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ refl{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ compatD{\isacharbrackleft}{\kern0pt}dest{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}compat{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}d{\isasymin}P{\isachardot}{\kern0pt}\ d{\isasympreceq}p\ {\isasymand}\ d{\isasympreceq}q{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}def\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ Incompatible\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}\isakeyword{infixl}\ {\isachardoublequoteopen}{\isasymbottom}{\isachardoublequoteclose}\ {\isadigit{5}}{\isadigit{0}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}p\ {\isasymbottom}\ q\ {\isasymequiv}\ {\isasymnot}\ compat{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ compatI{\isacharbrackleft}{\kern0pt}intro{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}d{\isasymin}P\ {\isasymLongrightarrow}\ d{\isasympreceq}p\ {\isasymLongrightarrow}\ d{\isasympreceq}q\ {\isasymLongrightarrow}\ compat{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}def\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ denseD\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}D{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p{\isasymin}P\ {\isasymLongrightarrow}\ \ {\isasymexists}d{\isasymin}D{\isachardot}{\kern0pt}\ d{\isasympreceq}\ p{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ denseI\ {\isacharbrackleft}{\kern0pt}intro{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ {\isasymAnd}p{\isachardot}{\kern0pt}\ p{\isasymin}P\ {\isasymLongrightarrow}\ {\isasymexists}d{\isasymin}D{\isachardot}{\kern0pt}\ d{\isasympreceq}\ p\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ dense{\isacharparenleft}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}belowD\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}d{\isasymin}D{\isachardot}{\kern0pt}\ d{\isasymin}P\ {\isasymand}\ d{\isasympreceq}q{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}belowI\ {\isacharbrackleft}{\kern0pt}intro{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}q{\isachardot}{\kern0pt}\ q{\isasymin}P\ {\isasymLongrightarrow}\ q{\isasympreceq}p\ {\isasymLongrightarrow}\ {\isasymexists}d{\isasymin}D{\isachardot}{\kern0pt}\ d{\isasymin}P\ {\isasymand}\ d{\isasympreceq}q{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}cong{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}P\ {\isasymLongrightarrow}\ D\ {\isacharequal}{\kern0pt}\ D{\isacharprime}{\kern0pt}\ {\isasymLongrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}cong{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}P\ {\isasymLongrightarrow}\ {\isasymlbrakk}{\isasymAnd}x{\isachardot}{\kern0pt}\ x{\isasymin}P\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymrbrakk}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ Q{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}mono{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}P\ {\isasymLongrightarrow}\ D\ {\isasymsubseteq}\ D{\isacharprime}{\kern0pt}\ {\isasymLongrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}under{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ leq{\isacharunderscore}{\kern0pt}transD\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ideal{\isacharunderscore}{\kern0pt}dense{\isacharunderscore}{\kern0pt}below{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}q{\isachardot}{\kern0pt}\ q{\isasymin}P\ {\isasymLongrightarrow}\ q{\isasympreceq}p\ {\isasymLongrightarrow}\ q{\isasymin}D{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ leq{\isacharunderscore}{\kern0pt}reflI\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}dense{\isacharunderscore}{\kern0pt}below{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}D{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ leq{\isacharunderscore}{\kern0pt}transD\ leq{\isacharunderscore}{\kern0pt}reflI\ \ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\ \ \ \ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ antichain\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}antichain{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ A{\isasymsubseteq}P\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}p{\isasymin}A{\isachardot}{\kern0pt}{\isasymforall}q{\isasymin}A{\isachardot}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymnot}compat{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
A filter is an increasing set $G$ with all its elements 
being compatible in $G$.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ filter\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ G{\isasymsubseteq}P\ {\isasymand}\ increasing{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}p{\isasymin}G{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}G{\isachardot}{\kern0pt}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ filterD\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ G\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add\ {\isacharcolon}{\kern0pt}\ subsetD\ filter{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}leqD\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ G\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P\ {\isasymLongrightarrow}\ x{\isasympreceq}y\ {\isasymLongrightarrow}\ y\ {\isasymin}\ G{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ filter{\isacharunderscore}{\kern0pt}def\ increasing{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}compat{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p{\isasymin}G\ {\isasymLongrightarrow}\ q{\isasymin}G\ {\isasymLongrightarrow}\ compat{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}def\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ compat{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ low{\isacharunderscore}{\kern0pt}bound{\isacharunderscore}{\kern0pt}filter{\isacharcolon}{\kern0pt}\ %
\isamarkupcmt{says the compatibility is attained inside G%
}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}q{\isasymin}G{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}r{\isasymin}G{\isachardot}{\kern0pt}\ r{\isasympreceq}p\ {\isasymand}\ r{\isasympreceq}q{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
We finally introduce the upward closure of a set
and prove that the closure of $A$ is a filter if its elements are
compatible in $A$.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ \ \isanewline
\ \ upclosure\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}p{\isasymin}P{\isachardot}{\kern0pt}{\isasymexists}a{\isasymin}A{\isachardot}{\kern0pt}\ a{\isasympreceq}p{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ upclosureI\ {\isacharbrackleft}{\kern0pt}intro{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}P\ {\isasymLongrightarrow}\ a{\isasymin}A\ {\isasymLongrightarrow}\ a{\isasympreceq}p\ {\isasymLongrightarrow}\ p{\isasymin}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}upclosure{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ upclosureE\ {\isacharbrackleft}{\kern0pt}elim{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}p{\isasymin}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x\ a{\isachardot}{\kern0pt}\ x{\isasymin}P\ {\isasymLongrightarrow}\ a{\isasymin}A\ {\isasymLongrightarrow}\ a{\isasympreceq}x\ {\isasymLongrightarrow}\ R{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ R{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}upclosure{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ upclosureD\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}p{\isasymin}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}a{\isasymin}A{\isachardot}{\kern0pt}{\isacharparenleft}{\kern0pt}a{\isasympreceq}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ p{\isasymin}P{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}upclosure{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ upclosure{\isacharunderscore}{\kern0pt}increasing\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A{\isasymsubseteq}P{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}increasing{\isacharparenleft}{\kern0pt}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ increasing{\isacharunderscore}{\kern0pt}def\ upclosure{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}transD{\isacharprime}{\kern0pt}{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}A{\isasymsubseteq}P{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ upclosure{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isasymsubseteq}\ P\ {\isasymLongrightarrow}\ upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ subsetI\ upclosure{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ A{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}upclosure{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isasymsubseteq}\ P\ {\isasymLongrightarrow}\ A{\isasymsubseteq}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ subsetI\ leq{\isacharunderscore}{\kern0pt}preord\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ upclosure{\isacharunderscore}{\kern0pt}def\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ refl{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ elem{\isacharunderscore}{\kern0pt}upclosure{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A{\isasymsubseteq}P\ {\isasymLongrightarrow}\ x{\isasymin}A\ \ {\isasymLongrightarrow}\ x{\isasymin}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ dest{\isacharcolon}{\kern0pt}A{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}upclosure{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ closure{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}filter{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A{\isasymsubseteq}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}p{\isasymin}A{\isachardot}{\kern0pt}{\isasymforall}q{\isasymin}A{\isachardot}{\kern0pt}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}def\isanewline
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}increasing{\isacharparenleft}{\kern0pt}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ upclosure{\isacharunderscore}{\kern0pt}increasing\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}UA{\isacharequal}{\kern0pt}{\isachardoublequoteopen}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}upclosure{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}p{\isasymin}{\isacharquery}{\kern0pt}UA{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}{\isacharquery}{\kern0pt}UA{\isachardoublequoteclose}\ \isakeyword{for}\ p\ q\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ that\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}a{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasympreceq}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasympreceq}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ upclosureD{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}p{\isasymin}{\isacharquery}{\kern0pt}UA{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ upclosureD{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}q{\isasymin}{\isacharquery}{\kern0pt}UA{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ d\ \isakeyword{where}\ {\isachardoublequoteopen}d{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}d{\isasympreceq}a{\isachardoublequoteclose}\ {\isachardoublequoteopen}d{\isasympreceq}b{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}d{\isasympreceq}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}d{\isasympreceq}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}d{\isasymin}{\isacharquery}{\kern0pt}UA{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ A{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}upclosure{\isacharbrackleft}{\kern0pt}THEN\ subsetD{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}A{\isasymsubseteq}P{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ leq{\isacharunderscore}{\kern0pt}transD{\isacharprime}{\kern0pt}{\isacharbrackleft}{\kern0pt}of\ A\ d\ a{\isacharbrackright}{\kern0pt}\ leq{\isacharunderscore}{\kern0pt}transD{\isacharprime}{\kern0pt}{\isacharbrackleft}{\kern0pt}of\ A\ d\ b{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ aux{\isacharunderscore}{\kern0pt}RS{\isadigit{1}}{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}f\ {\isasymin}\ N\ {\isasymrightarrow}\ P\ {\isasymLongrightarrow}\ n{\isasymin}N\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}n\ {\isasymin}\ upclosure{\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}N{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}upclosure{\isacharbrackleft}{\kern0pt}OF\ subset{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}image{\isacharbrackright}{\kern0pt}\ image{\isacharunderscore}{\kern0pt}fun\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ blast{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ decr{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}decr{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat\ {\isasymrightarrow}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}preorder{\isacharunderscore}{\kern0pt}on{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ \ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ n{\isasymrangle}\ {\isasymin}\ leq{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}m{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}n{\isasymin}nat\ {\isasymLongrightarrow}\ n{\isasymle}m\ {\isasymLongrightarrow}\ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ m{\isacharcomma}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ n{\isasymrangle}\ {\isasymin}\ leq{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}m{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}induct\ m{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{0}}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ assms\ leq{\isacharunderscore}{\kern0pt}reflI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}succ\ x{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasympreceq}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}lt{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}n{\isacharless}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}n{\isacharequal}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ succ\ le{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ lt\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ leI\ succ\ leq{\isacharunderscore}{\kern0pt}transD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ eq\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}reflI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ decr{\isacharunderscore}{\kern0pt}seq{\isacharunderscore}{\kern0pt}linear{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}refl{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat\ {\isasymrightarrow}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ \ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ n{\isasymrangle}\ {\isasymin}\ leq{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}trans{\isacharbrackleft}{\kern0pt}P{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}linear{\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ nat{\isacharcomma}{\kern0pt}\ leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}preorder{\isacharunderscore}{\kern0pt}on{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n\ m\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}m\ {\isasympreceq}\ f{\isacharbackquote}{\kern0pt}n\ {\isasymor}\ f{\isacharbackquote}{\kern0pt}n\ {\isasympreceq}\ f{\isacharbackquote}{\kern0pt}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}m{\isasymle}n{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}m{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ decr{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}decr{\isacharbrackleft}{\kern0pt}of\ f\ n\ m{\isacharbrackright}{\kern0pt}\ assms\ leI\ {\isacartoucheopen}preorder{\isacharunderscore}{\kern0pt}on{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}m{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ decr{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}decr{\isacharbrackleft}{\kern0pt}of\ f\ m\ n{\isacharbrackright}{\kern0pt}\ assms\ leI\ not{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}lt\ {\isacartoucheopen}preorder{\isacharunderscore}{\kern0pt}on{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ linear{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ ball{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}simp\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Towards Rasiowa-Sikorski Lemma (RSL)%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{locale}\isamarkupfalse%
\ countable{\isacharunderscore}{\kern0pt}generic\ {\isacharequal}{\kern0pt}\ forcing{\isacharunderscore}{\kern0pt}notion\ {\isacharplus}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ {\isasymD}\isanewline
\ \ \isakeyword{assumes}\ countable{\isacharunderscore}{\kern0pt}subs{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}P{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isasymD}\ {\isasymin}\ nat{\isasymrightarrow}Pow{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ \ \ \ \ seq{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}denses{\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ dense{\isacharparenleft}{\kern0pt}{\isasymD}{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ D{\isacharunderscore}{\kern0pt}generic\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}D{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymD}{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasyminter}G{\isasymnoteq}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The next lemma identifies a sufficient condition for obtaining
RSL.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ RS{\isacharunderscore}{\kern0pt}sequence{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}rasiowa{\isacharunderscore}{\kern0pt}sikorski{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isacharcolon}{\kern0pt}\ nat{\isasymrightarrow}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isacharbackquote}{\kern0pt}\ {\isadigit{0}}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}n{\isachardot}{\kern0pt}\ n{\isasymin}nat\ {\isasymLongrightarrow}\ f\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasympreceq}\ f\ {\isacharbackquote}{\kern0pt}\ n\ {\isasymand}\ f\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymexists}G{\isachardot}{\kern0pt}\ p{\isasymin}G\ {\isasymand}\ D{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}nat\ \ {\isasymsubseteq}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}image{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}refl{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ leq{\isacharparenright}{\kern0pt}\ {\isasymand}\ trans{\isacharbrackleft}{\kern0pt}P{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}preord\ \isacommand{unfolding}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ intro{\isacharcolon}{\kern0pt}refl{\isacharunderscore}{\kern0pt}monot{\isacharunderscore}{\kern0pt}domain{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ \ f\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasympreceq}\ f\ {\isacharbackquote}{\kern0pt}\ n{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}linear{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ leq{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}preord\ \isakeyword{and}\ decr{\isacharunderscore}{\kern0pt}seq{\isacharunderscore}{\kern0pt}linear\ \isacommand{unfolding}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}p{\isasymin}f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}nat{\isachardot}{\kern0pt}{\isasymforall}q{\isasymin}f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}nat{\isachardot}{\kern0pt}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}nat{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ chain{\isacharunderscore}{\kern0pt}compat\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}upclosure{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ closure{\isacharunderscore}{\kern0pt}compat{\isacharunderscore}{\kern0pt}filter\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ n\ {\isasyminter}\ {\isacharquery}{\kern0pt}G\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharquery}{\kern0pt}G\ {\isasymand}\ f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ aux{\isacharunderscore}{\kern0pt}RS{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymD}\ {\isacharbackquote}{\kern0pt}\ n\ {\isasyminter}\ {\isacharquery}{\kern0pt}G\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ {\isacharquery}{\kern0pt}G{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ aux{\isacharunderscore}{\kern0pt}RS{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ D{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
%
\begin{isamarkuptext}%
Now, the following recursive definition will fulfill the 
requirements of lemma \isa{RS{\isacharunderscore}{\kern0pt}sequence{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}rasiowa{\isacharunderscore}{\kern0pt}sikorski}%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{consts}\isamarkupfalse%
\ RS{\isacharunderscore}{\kern0pt}seq\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\isanewline
\isacommand{primrec}\isamarkupfalse%
\isanewline
\ \ {\isachardoublequoteopen}RS{\isacharunderscore}{\kern0pt}seq{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}enum{\isacharcomma}{\kern0pt}{\isasymD}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}RS{\isacharunderscore}{\kern0pt}seq{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}enum{\isacharcomma}{\kern0pt}{\isasymD}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \isanewline
\ \ \ \ enum{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymmu}\ m{\isachardot}{\kern0pt}\ {\isasymlangle}enum{\isacharbackquote}{\kern0pt}m{\isacharcomma}{\kern0pt}\ RS{\isacharunderscore}{\kern0pt}seq{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}enum{\isacharcomma}{\kern0pt}{\isasymD}{\isacharparenright}{\kern0pt}{\isasymrangle}\ {\isasymin}\ leq\ {\isasymand}\ enum{\isacharbackquote}{\kern0pt}m\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ countable{\isacharunderscore}{\kern0pt}generic\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ preimage{\isacharunderscore}{\kern0pt}rangeD{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f{\isasymin}Pi{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}a{\isasymin}A{\isachardot}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ b{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ apply{\isacharunderscore}{\kern0pt}equality{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ assms{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ of\ {\isacharunderscore}{\kern0pt}\ b{\isacharbrackright}{\kern0pt}\ domain{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ assms{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ countable{\isacharunderscore}{\kern0pt}RS{\isacharunderscore}{\kern0pt}sequence{\isacharunderscore}{\kern0pt}aux{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ p\ enum\isanewline
\ \ \isakeyword{defines}\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ RS{\isacharunderscore}{\kern0pt}seq{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}enum{\isacharcomma}{\kern0pt}{\isasymD}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ \ \ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}k{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ enum{\isacharbackquote}{\kern0pt}m{\isasympreceq}\ q\ {\isasymand}\ enum{\isacharbackquote}{\kern0pt}m\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ k{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}P\ {\isasymsubseteq}\ range{\isacharparenleft}{\kern0pt}enum{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isacharcolon}{\kern0pt}nat{\isasymrightarrow}M{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}x\ k{\isachardot}{\kern0pt}\ x{\isasymin}P\ {\isasymLongrightarrow}\ k{\isasymin}nat\ {\isasymLongrightarrow}\ \ {\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}\ x\ {\isasymand}\ q\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ k{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ P\ {\isasymand}\ f{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasympreceq}\ f{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymand}\ f{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{0}}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ {\isacartoucheopen}P\ {\isasymsubseteq}\ range{\isacharparenleft}{\kern0pt}enum{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isacharbackquote}{\kern0pt}m\ {\isacharequal}{\kern0pt}\ q{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ preimage{\isacharunderscore}{\kern0pt}rangeD{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}enum{\isacharcolon}{\kern0pt}nat{\isasymrightarrow}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymD}{\isacharbackquote}{\kern0pt}{\isadigit{0}}\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}funtype{\isacharbrackleft}{\kern0pt}OF\ countable{\isacharunderscore}{\kern0pt}subs{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}P{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ LeastI{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ m{\isacharbrackright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ Q{\isacharunderscore}{\kern0pt}def\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}succ\ n{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}\ f{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ {\isacartoucheopen}P\ {\isasymsubseteq}\ range{\isacharparenleft}{\kern0pt}enum{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isacharbackquote}{\kern0pt}m{\isasympreceq}\ f{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isacharbackquote}{\kern0pt}m\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ preimage{\isacharunderscore}{\kern0pt}rangeD{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}enum{\isacharcolon}{\kern0pt}nat{\isasymrightarrow}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ succ\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymD}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}funtype{\isacharbrackleft}{\kern0pt}OF\ countable{\isacharunderscore}{\kern0pt}subs{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}P{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ LeastI{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ m{\isacharbrackright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ Q{\isacharunderscore}{\kern0pt}def\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ countable{\isacharunderscore}{\kern0pt}RS{\isacharunderscore}{\kern0pt}sequence{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ p\ enum\isanewline
\ \ \isakeyword{defines}\ {\isachardoublequoteopen}f\ {\isasymequiv}\ {\isasymlambda}n{\isasymin}nat{\isachardot}{\kern0pt}\ RS{\isacharunderscore}{\kern0pt}seq{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}enum{\isacharcomma}{\kern0pt}{\isasymD}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ \ \ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}k{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ enum{\isacharbackquote}{\kern0pt}m{\isasympreceq}\ q\ {\isasymand}\ enum{\isacharbackquote}{\kern0pt}m\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ k{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}P\ {\isasymsubseteq}\ range{\isacharparenleft}{\kern0pt}enum{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isacharcolon}{\kern0pt}nat{\isasymrightarrow}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}{\isadigit{0}}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasympreceq}\ f{\isacharbackquote}{\kern0pt}n\ {\isasymand}\ f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}{\isadigit{0}}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ k\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}k{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}\ x\ {\isasymand}\ q\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ k{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}denses\ apply{\isacharunderscore}{\kern0pt}funtype{\isacharbrackleft}{\kern0pt}OF\ countable{\isacharunderscore}{\kern0pt}subs{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}P{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasympreceq}\ f{\isacharbackquote}{\kern0pt}n\ {\isasymand}\ f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymD}\ {\isacharbackquote}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ countable{\isacharunderscore}{\kern0pt}RS{\isacharunderscore}{\kern0pt}sequence{\isacharunderscore}{\kern0pt}aux\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ RS{\isacharunderscore}{\kern0pt}seq{\isacharunderscore}{\kern0pt}type{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}P\ {\isasymsubseteq}\ range{\isacharparenleft}{\kern0pt}enum{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isacharcolon}{\kern0pt}nat{\isasymrightarrow}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}RS{\isacharunderscore}{\kern0pt}seq{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}enum{\isacharcomma}{\kern0pt}{\isasymD}{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ countable{\isacharunderscore}{\kern0pt}RS{\isacharunderscore}{\kern0pt}sequence{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct{\isacharsemicolon}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ RS{\isacharunderscore}{\kern0pt}seq{\isacharunderscore}{\kern0pt}funtype{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}P\ {\isasymsubseteq}\ range{\isacharparenleft}{\kern0pt}enum{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isacharcolon}{\kern0pt}nat{\isasymrightarrow}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymlambda}n{\isasymin}nat{\isachardot}{\kern0pt}\ RS{\isacharunderscore}{\kern0pt}seq{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}enum{\isacharcomma}{\kern0pt}{\isasymD}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcolon}{\kern0pt}\ nat\ {\isasymrightarrow}\ P{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ lam{\isacharunderscore}{\kern0pt}type\ RS{\isacharunderscore}{\kern0pt}seq{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ countable{\isacharunderscore}{\kern0pt}rasiowa{\isacharunderscore}{\kern0pt}sikorski\ {\isacharequal}{\kern0pt}\ \isanewline
\ \ RS{\isacharunderscore}{\kern0pt}sequence{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}rasiowa{\isacharunderscore}{\kern0pt}sikorski{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ RS{\isacharunderscore}{\kern0pt}seq{\isacharunderscore}{\kern0pt}funtype\ countable{\isacharunderscore}{\kern0pt}RS{\isacharunderscore}{\kern0pt}sequence{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Forcing_Notions.thy%:%
%:%11=1%:%
%:%23=2%:%
%:%24=3%:%
%:%32=5%:%
%:%33=5%:%
%:%34=6%:%
%:%35=7%:%
%:%49=9%:%
%:%61=10%:%
%:%62=11%:%
%:%64=12%:%
%:%65=12%:%
%:%66=13%:%
%:%67=14%:%
%:%68=15%:%
%:%69=15%:%
%:%70=16%:%
%:%71=17%:%
%:%72=18%:%
%:%73=19%:%
%:%74=20%:%
%:%75=20%:%
%:%76=21%:%
%:%79=22%:%
%:%83=22%:%
%:%84=22%:%
%:%89=22%:%
%:%92=23%:%
%:%93=24%:%
%:%94=24%:%
%:%95=25%:%
%:%98=26%:%
%:%102=26%:%
%:%103=26%:%
%:%108=26%:%
%:%111=27%:%
%:%112=28%:%
%:%113=28%:%
%:%114=29%:%
%:%117=30%:%
%:%121=30%:%
%:%122=30%:%
%:%127=30%:%
%:%130=31%:%
%:%131=32%:%
%:%132=32%:%
%:%135=33%:%
%:%139=33%:%
%:%140=33%:%
%:%145=33%:%
%:%148=34%:%
%:%149=35%:%
%:%150=35%:%
%:%153=36%:%
%:%157=36%:%
%:%158=36%:%
%:%159=36%:%
%:%164=36%:%
%:%167=37%:%
%:%168=38%:%
%:%169=38%:%
%:%170=39%:%
%:%171=40%:%
%:%172=41%:%
%:%173=42%:%
%:%174=42%:%
%:%175=43%:%
%:%176=44%:%
%:%177=45%:%
%:%178=46%:%
%:%179=46%:%
%:%180=47%:%
%:%181=48%:%
%:%182=49%:%
%:%183=50%:%
%:%184=51%:%
%:%185=52%:%
%:%186=53%:%
%:%187=53%:%
%:%188=54%:%
%:%189=55%:%
%:%190=56%:%
%:%191=56%:%
%:%192=57%:%
%:%195=58%:%
%:%199=58%:%
%:%200=58%:%
%:%201=58%:%
%:%202=58%:%
%:%211=60%:%
%:%212=61%:%
%:%214=62%:%
%:%215=62%:%
%:%216=63%:%
%:%217=64%:%
%:%219=66%:%
%:%220=67%:%
%:%221=68%:%
%:%223=69%:%
%:%224=69%:%
%:%225=70%:%
%:%226=71%:%
%:%227=72%:%
%:%228=73%:%
%:%229=73%:%
%:%232=74%:%
%:%236=74%:%
%:%237=74%:%
%:%242=74%:%
%:%245=75%:%
%:%246=76%:%
%:%247=76%:%
%:%248=77%:%
%:%249=78%:%
%:%250=79%:%
%:%251=80%:%
%:%252=80%:%
%:%253=81%:%
%:%254=82%:%
%:%255=83%:%
%:%256=84%:%
%:%257=84%:%
%:%260=85%:%
%:%264=85%:%
%:%265=85%:%
%:%266=85%:%
%:%267=85%:%
%:%272=85%:%
%:%275=86%:%
%:%276=87%:%
%:%277=87%:%
%:%280=88%:%
%:%284=88%:%
%:%285=88%:%
%:%286=88%:%
%:%287=88%:%
%:%292=88%:%
%:%295=89%:%
%:%296=90%:%
%:%297=91%:%
%:%298=91%:%
%:%301=92%:%
%:%305=92%:%
%:%306=92%:%
%:%307=92%:%
%:%308=92%:%
%:%313=92%:%
%:%316=93%:%
%:%317=94%:%
%:%318=94%:%
%:%321=95%:%
%:%325=95%:%
%:%326=95%:%
%:%332=95%:%
%:%335=96%:%
%:%336=97%:%
%:%337=97%:%
%:%338=98%:%
%:%339=99%:%
%:%340=100%:%
%:%341=100%:%
%:%344=101%:%
%:%348=101%:%
%:%349=101%:%
%:%350=101%:%
%:%355=101%:%
%:%358=102%:%
%:%359=103%:%
%:%360=103%:%
%:%363=104%:%
%:%367=104%:%
%:%368=104%:%
%:%369=104%:%
%:%374=104%:%
%:%377=105%:%
%:%378=106%:%
%:%379=106%:%
%:%382=107%:%
%:%386=107%:%
%:%387=107%:%
%:%388=107%:%
%:%393=107%:%
%:%396=108%:%
%:%397=109%:%
%:%398=109%:%
%:%399=110%:%
%:%400=111%:%
%:%403=112%:%
%:%407=112%:%
%:%408=112%:%
%:%409=112%:%
%:%410=112%:%
%:%415=112%:%
%:%418=113%:%
%:%418=114%:%
%:%419=115%:%
%:%420=116%:%
%:%421=116%:%
%:%422=117%:%
%:%423=118%:%
%:%426=119%:%
%:%430=119%:%
%:%431=119%:%
%:%432=119%:%
%:%433=119%:%
%:%438=119%:%
%:%441=120%:%
%:%442=121%:%
%:%443=121%:%
%:%446=122%:%
%:%450=122%:%
%:%451=122%:%
%:%456=122%:%
%:%459=123%:%
%:%460=124%:%
%:%461=124%:%
%:%462=125%:%
%:%465=126%:%
%:%469=126%:%
%:%470=126%:%
%:%475=126%:%
%:%478=127%:%
%:%479=128%:%
%:%480=128%:%
%:%483=129%:%
%:%487=129%:%
%:%488=129%:%
%:%493=129%:%
%:%496=130%:%
%:%497=131%:%
%:%498=131%:%
%:%499=132%:%
%:%500=133%:%
%:%503=134%:%
%:%507=134%:%
%:%508=134%:%
%:%509=134%:%
%:%514=134%:%
%:%517=135%:%
%:%518=136%:%
%:%519=136%:%
%:%520=137%:%
%:%521=138%:%
%:%524=139%:%
%:%528=139%:%
%:%529=139%:%
%:%530=139%:%
%:%535=139%:%
%:%538=140%:%
%:%539=141%:%
%:%540=141%:%
%:%541=142%:%
%:%542=143%:%
%:%545=144%:%
%:%549=144%:%
%:%550=144%:%
%:%551=144%:%
%:%556=144%:%
%:%559=145%:%
%:%560=146%:%
%:%560=161%:%
%:%561=162%:%
%:%562=163%:%
%:%563=163%:%
%:%564=164%:%
%:%565=165%:%
%:%567=167%:%
%:%568=168%:%
%:%570=169%:%
%:%571=169%:%
%:%572=170%:%
%:%573=171%:%
%:%574=172%:%
%:%575=173%:%
%:%576=173%:%
%:%579=174%:%
%:%583=174%:%
%:%584=174%:%
%:%589=174%:%
%:%592=175%:%
%:%593=176%:%
%:%594=176%:%
%:%597=177%:%
%:%601=177%:%
%:%602=177%:%
%:%607=177%:%
%:%610=178%:%
%:%611=179%:%
%:%612=179%:%
%:%615=180%:%
%:%619=180%:%
%:%620=180%:%
%:%621=180%:%
%:%626=180%:%
%:%629=181%:%
%:%630=182%:%
%:%631=182%:%
%:%632=182%:%
%:%633=182%:%
%:%634=183%:%
%:%635=184%:%
%:%638=185%:%
%:%642=185%:%
%:%643=185%:%
%:%644=186%:%
%:%645=186%:%
%:%646=186%:%
%:%655=188%:%
%:%656=189%:%
%:%657=190%:%
%:%659=191%:%
%:%660=191%:%
%:%661=192%:%
%:%662=193%:%
%:%663=194%:%
%:%664=195%:%
%:%665=195%:%
%:%668=196%:%
%:%672=196%:%
%:%673=196%:%
%:%678=196%:%
%:%681=197%:%
%:%682=198%:%
%:%683=198%:%
%:%684=199%:%
%:%687=200%:%
%:%691=200%:%
%:%692=200%:%
%:%697=200%:%
%:%700=201%:%
%:%701=202%:%
%:%702=202%:%
%:%703=203%:%
%:%706=204%:%
%:%710=204%:%
%:%711=204%:%
%:%716=204%:%
%:%719=205%:%
%:%720=206%:%
%:%721=206%:%
%:%722=207%:%
%:%723=208%:%
%:%726=209%:%
%:%730=209%:%
%:%731=209%:%
%:%732=210%:%
%:%733=210%:%
%:%734=210%:%
%:%739=210%:%
%:%742=211%:%
%:%743=212%:%
%:%744=212%:%
%:%747=213%:%
%:%751=213%:%
%:%752=213%:%
%:%753=213%:%
%:%758=213%:%
%:%761=214%:%
%:%762=215%:%
%:%763=215%:%
%:%766=216%:%
%:%770=216%:%
%:%771=216%:%
%:%772=217%:%
%:%773=217%:%
%:%774=217%:%
%:%779=217%:%
%:%782=218%:%
%:%783=219%:%
%:%784=219%:%
%:%787=220%:%
%:%791=220%:%
%:%792=220%:%
%:%797=220%:%
%:%800=221%:%
%:%801=222%:%
%:%802=222%:%
%:%803=223%:%
%:%804=224%:%
%:%807=225%:%
%:%811=225%:%
%:%812=225%:%
%:%813=226%:%
%:%814=226%:%
%:%815=227%:%
%:%816=227%:%
%:%817=228%:%
%:%818=228%:%
%:%819=228%:%
%:%820=229%:%
%:%821=229%:%
%:%822=230%:%
%:%823=230%:%
%:%824=231%:%
%:%825=231%:%
%:%826=232%:%
%:%827=232%:%
%:%828=233%:%
%:%829=233%:%
%:%830=234%:%
%:%831=234%:%
%:%832=235%:%
%:%833=235%:%
%:%834=235%:%
%:%835=236%:%
%:%836=236%:%
%:%837=237%:%
%:%838=237%:%
%:%839=238%:%
%:%840=238%:%
%:%841=238%:%
%:%842=239%:%
%:%843=239%:%
%:%844=240%:%
%:%845=240%:%
%:%846=241%:%
%:%847=241%:%
%:%848=242%:%
%:%849=242%:%
%:%850=243%:%
%:%851=243%:%
%:%852=244%:%
%:%853=244%:%
%:%854=244%:%
%:%855=244%:%
%:%856=245%:%
%:%857=245%:%
%:%858=246%:%
%:%864=246%:%
%:%867=247%:%
%:%868=248%:%
%:%869=248%:%
%:%872=249%:%
%:%876=249%:%
%:%877=249%:%
%:%878=250%:%
%:%879=250%:%
%:%884=250%:%
%:%887=251%:%
%:%888=252%:%
%:%889=252%:%
%:%890=253%:%
%:%891=254%:%
%:%892=255%:%
%:%893=256%:%
%:%896=257%:%
%:%900=257%:%
%:%901=257%:%
%:%902=258%:%
%:%903=258%:%
%:%904=259%:%
%:%905=259%:%
%:%906=260%:%
%:%907=260%:%
%:%908=260%:%
%:%909=260%:%
%:%910=260%:%
%:%911=261%:%
%:%912=261%:%
%:%913=262%:%
%:%914=262%:%
%:%915=263%:%
%:%916=263%:%
%:%917=264%:%
%:%918=264%:%
%:%919=265%:%
%:%920=265%:%
%:%921=265%:%
%:%922=266%:%
%:%923=266%:%
%:%924=267%:%
%:%925=267%:%
%:%926=267%:%
%:%927=268%:%
%:%928=268%:%
%:%929=269%:%
%:%930=269%:%
%:%931=270%:%
%:%932=270%:%
%:%933=271%:%
%:%934=271%:%
%:%935=272%:%
%:%936=272%:%
%:%937=272%:%
%:%938=272%:%
%:%939=272%:%
%:%940=273%:%
%:%941=273%:%
%:%942=274%:%
%:%943=274%:%
%:%944=275%:%
%:%945=275%:%
%:%946=275%:%
%:%947=275%:%
%:%948=275%:%
%:%949=276%:%
%:%950=276%:%
%:%951=277%:%
%:%957=277%:%
%:%960=278%:%
%:%961=279%:%
%:%962=279%:%
%:%963=280%:%
%:%964=281%:%
%:%965=282%:%
%:%966=283%:%
%:%973=284%:%
%:%974=284%:%
%:%975=285%:%
%:%976=285%:%
%:%977=286%:%
%:%978=286%:%
%:%979=286%:%
%:%980=286%:%
%:%981=287%:%
%:%982=287%:%
%:%983=288%:%
%:%984=288%:%
%:%985=289%:%
%:%986=289%:%
%:%987=290%:%
%:%988=290%:%
%:%989=291%:%
%:%990=291%:%
%:%991=292%:%
%:%992=292%:%
%:%993=293%:%
%:%994=293%:%
%:%995=294%:%
%:%996=294%:%
%:%997=295%:%
%:%998=295%:%
%:%999=296%:%
%:%1000=296%:%
%:%1001=296%:%
%:%1002=297%:%
%:%1003=297%:%
%:%1004=298%:%
%:%1005=298%:%
%:%1006=299%:%
%:%1007=299%:%
%:%1008=300%:%
%:%1009=300%:%
%:%1010=301%:%
%:%1011=301%:%
%:%1012=301%:%
%:%1013=302%:%
%:%1014=302%:%
%:%1015=303%:%
%:%1016=303%:%
%:%1017=304%:%
%:%1018=304%:%
%:%1019=305%:%
%:%1020=305%:%
%:%1021=306%:%
%:%1022=306%:%
%:%1023=306%:%
%:%1024=306%:%
%:%1025=307%:%
%:%1031=307%:%
%:%1034=308%:%
%:%1035=309%:%
%:%1043=311%:%
%:%1053=312%:%
%:%1054=312%:%
%:%1055=313%:%
%:%1056=314%:%
%:%1057=315%:%
%:%1058=316%:%
%:%1059=317%:%
%:%1060=318%:%
%:%1061=319%:%
%:%1062=319%:%
%:%1063=320%:%
%:%1064=321%:%
%:%1066=323%:%
%:%1067=324%:%
%:%1069=325%:%
%:%1070=325%:%
%:%1071=326%:%
%:%1072=327%:%
%:%1073=328%:%
%:%1074=329%:%
%:%1075=330%:%
%:%1082=331%:%
%:%1083=331%:%
%:%1084=332%:%
%:%1085=332%:%
%:%1086=333%:%
%:%1087=333%:%
%:%1088=333%:%
%:%1089=334%:%
%:%1090=334%:%
%:%1091=335%:%
%:%1092=335%:%
%:%1093=336%:%
%:%1094=336%:%
%:%1095=336%:%
%:%1096=337%:%
%:%1097=337%:%
%:%1098=338%:%
%:%1099=338%:%
%:%1100=338%:%
%:%1101=338%:%
%:%1102=339%:%
%:%1103=339%:%
%:%1104=339%:%
%:%1105=340%:%
%:%1106=340%:%
%:%1107=340%:%
%:%1108=341%:%
%:%1109=341%:%
%:%1110=341%:%
%:%1111=342%:%
%:%1112=342%:%
%:%1113=343%:%
%:%1114=343%:%
%:%1115=343%:%
%:%1116=343%:%
%:%1117=344%:%
%:%1118=344%:%
%:%1119=344%:%
%:%1120=345%:%
%:%1121=345%:%
%:%1122=346%:%
%:%1123=346%:%
%:%1124=346%:%
%:%1125=347%:%
%:%1126=347%:%
%:%1127=348%:%
%:%1128=348%:%
%:%1129=349%:%
%:%1130=349%:%
%:%1131=349%:%
%:%1132=350%:%
%:%1133=350%:%
%:%1134=351%:%
%:%1135=351%:%
%:%1136=352%:%
%:%1137=352%:%
%:%1138=353%:%
%:%1139=353%:%
%:%1140=354%:%
%:%1141=354%:%
%:%1142=355%:%
%:%1143=355%:%
%:%1144=356%:%
%:%1145=356%:%
%:%1146=357%:%
%:%1147=357%:%
%:%1148=357%:%
%:%1149=358%:%
%:%1150=358%:%
%:%1151=359%:%
%:%1152=359%:%
%:%1153=359%:%
%:%1154=360%:%
%:%1155=360%:%
%:%1156=361%:%
%:%1157=361%:%
%:%1158=361%:%
%:%1159=362%:%
%:%1160=362%:%
%:%1161=363%:%
%:%1162=363%:%
%:%1163=363%:%
%:%1164=364%:%
%:%1165=364%:%
%:%1166=365%:%
%:%1167=365%:%
%:%1168=365%:%
%:%1169=365%:%
%:%1170=366%:%
%:%1176=366%:%
%:%1179=367%:%
%:%1180=368%:%
%:%1183=370%:%
%:%1184=371%:%
%:%1186=373%:%
%:%1187=373%:%
%:%1188=374%:%
%:%1189=374%:%
%:%1190=375%:%
%:%1191=376%:%
%:%1192=377%:%
%:%1193=378%:%
%:%1194=379%:%
%:%1195=379%:%
%:%1196=380%:%
%:%1197=381%:%
%:%1198=382%:%
%:%1199=382%:%
%:%1200=383%:%
%:%1201=384%:%
%:%1204=385%:%
%:%1208=385%:%
%:%1209=385%:%
%:%1210=385%:%
%:%1215=385%:%
%:%1218=386%:%
%:%1219=387%:%
%:%1220=387%:%
%:%1221=388%:%
%:%1222=389%:%
%:%1223=390%:%
%:%1224=391%:%
%:%1225=392%:%
%:%1226=393%:%
%:%1227=394%:%
%:%1230=395%:%
%:%1234=395%:%
%:%1235=395%:%
%:%1236=396%:%
%:%1237=396%:%
%:%1238=397%:%
%:%1239=397%:%
%:%1240=398%:%
%:%1241=398%:%
%:%1242=399%:%
%:%1243=399%:%
%:%1244=399%:%
%:%1245=400%:%
%:%1246=400%:%
%:%1247=400%:%
%:%1248=401%:%
%:%1249=401%:%
%:%1250=402%:%
%:%1251=402%:%
%:%1252=402%:%
%:%1253=403%:%
%:%1254=403%:%
%:%1255=404%:%
%:%1256=404%:%
%:%1257=405%:%
%:%1258=405%:%
%:%1259=405%:%
%:%1260=406%:%
%:%1261=406%:%
%:%1262=406%:%
%:%1263=407%:%
%:%1264=407%:%
%:%1265=408%:%
%:%1266=408%:%
%:%1267=409%:%
%:%1268=409%:%
%:%1269=409%:%
%:%1270=409%:%
%:%1271=410%:%
%:%1272=410%:%
%:%1273=411%:%
%:%1274=411%:%
%:%1275=412%:%
%:%1276=412%:%
%:%1277=413%:%
%:%1278=413%:%
%:%1279=413%:%
%:%1280=414%:%
%:%1281=414%:%
%:%1282=414%:%
%:%1283=415%:%
%:%1284=415%:%
%:%1285=416%:%
%:%1286=416%:%
%:%1287=416%:%
%:%1288=417%:%
%:%1289=417%:%
%:%1290=417%:%
%:%1291=418%:%
%:%1292=418%:%
%:%1293=418%:%
%:%1294=419%:%
%:%1295=419%:%
%:%1296=420%:%
%:%1297=420%:%
%:%1298=420%:%
%:%1299=421%:%
%:%1300=421%:%
%:%1301=422%:%
%:%1302=422%:%
%:%1303=423%:%
%:%1304=423%:%
%:%1305=423%:%
%:%1306=423%:%
%:%1307=424%:%
%:%1313=424%:%
%:%1316=425%:%
%:%1317=426%:%
%:%1318=426%:%
%:%1319=427%:%
%:%1320=428%:%
%:%1321=429%:%
%:%1322=430%:%
%:%1323=431%:%
%:%1324=432%:%
%:%1331=433%:%
%:%1332=433%:%
%:%1333=434%:%
%:%1334=434%:%
%:%1335=435%:%
%:%1336=435%:%
%:%1337=435%:%
%:%1338=436%:%
%:%1339=436%:%
%:%1340=437%:%
%:%1341=437%:%
%:%1342=438%:%
%:%1343=438%:%
%:%1344=439%:%
%:%1345=439%:%
%:%1346=440%:%
%:%1347=440%:%
%:%1348=441%:%
%:%1349=441%:%
%:%1350=442%:%
%:%1351=442%:%
%:%1352=442%:%
%:%1353=443%:%
%:%1354=443%:%
%:%1355=444%:%
%:%1356=444%:%
%:%1357=445%:%
%:%1358=445%:%
%:%1359=446%:%
%:%1360=446%:%
%:%1361=446%:%
%:%1362=446%:%
%:%1363=447%:%
%:%1369=447%:%
%:%1372=448%:%
%:%1373=449%:%
%:%1374=449%:%
%:%1375=450%:%
%:%1376=451%:%
%:%1379=452%:%
%:%1383=452%:%
%:%1384=452%:%
%:%1385=453%:%
%:%1386=453%:%
%:%1391=453%:%
%:%1394=454%:%
%:%1395=455%:%
%:%1396=455%:%
%:%1397=456%:%
%:%1398=457%:%
%:%1401=458%:%
%:%1405=458%:%
%:%1406=458%:%
%:%1407=458%:%
%:%1412=458%:%
%:%1415=459%:%
%:%1416=460%:%
%:%1417=460%:%
%:%1418=461%:%
%:%1419=462%:%
%:%1420=462%:%
%:%1423=463%:%
%:%1428=464%:%