### 2024-11-01 code/output/document/Foundation_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Foundation{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Axiom of Foundation in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Foundation{\isacharunderscore}{\kern0pt}Axiom\isanewline
\isakeyword{imports}\isanewline
\ \ Names\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data\isanewline
\isakeyword{begin}\isanewline
\ \ \isanewline
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ foundation{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}foundation{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ foundation{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ rallI{\isacharcomma}{\kern0pt}\ cut{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ foundation{\isacharcomma}{\kern0pt}\ auto\ intro{\isacharcolon}{\kern0pt}\ transitivity{\isacharunderscore}{\kern0pt}MG{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}foundation{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\ \ \ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymexists}y{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y{\isasymin}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y{\isasymin}x{\isasyminter}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ {\isachardoublequoteopen}y{\isasymin}x{\isasyminter}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}z{\isasymin}y{\isachardot}{\kern0pt}\ z\ {\isasymnotin}\ x{\isasyminter}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ foundation{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}x{\isasyminter}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ x\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}z{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ z\ {\isasymnotin}\ x\ {\isasymor}\ z\ {\isasymnotin}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ foundation{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isacommand{end}\isamarkupfalse%
\ \ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Foundation_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%36=5%:%
%:%39=6%:%
%:%40=7%:%
%:%41=7%:%
%:%42=8%:%
%:%43=9%:%
%:%44=10%:%
%:%45=11%:%
%:%46=11%:%
%:%49=12%:%
%:%53=12%:%
%:%54=12%:%
%:%55=13%:%
%:%56=13%:%
%:%61=13%:%
%:%64=14%:%
%:%65=16%:%
%:%66=17%:%
%:%67=17%:%
%:%74=18%:%
%:%75=18%:%
%:%76=19%:%
%:%77=19%:%
%:%78=20%:%
%:%79=20%:%
%:%80=21%:%
%:%81=21%:%
%:%82=22%:%
%:%83=22%:%
%:%84=23%:%
%:%85=23%:%
%:%86=23%:%
%:%87=24%:%
%:%88=24%:%
%:%89=25%:%
%:%90=25%:%
%:%91=26%:%
%:%92=26%:%
%:%93=26%:%
%:%94=27%:%
%:%95=27%:%
%:%96=28%:%
%:%97=28%:%
%:%98=28%:%
%:%99=29%:%
%:%100=29%:%
%:%101=30%:%
%:%102=30%:%
%:%103=30%:%
%:%104=31%:%
%:%105=31%:%
%:%106=31%:%
%:%107=32%:%
%:%113=32%:%
%:%116=33%:%
%:%117=34%:%
%:%118=34%:%
%:%125=35%:%