### 2024-11-01 code/output/document/FrecR.tex

%
\begin{isabellebody}%
\setisabellecontext{FrecR}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Well-founded relation on names%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ FrecR\ \isakeyword{imports}\ Names\ Synthetic{\isacharunderscore}{\kern0pt}Definition\ \isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}rules{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ nth{\isacharunderscore}{\kern0pt}ConsI\ FOL{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ function{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\isanewline
\ \ fun{\isacharunderscore}{\kern0pt}plus{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ omega{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff%
\begin{isamarkuptext}%
\isa{frecR} is the well-founded relation on names that allows
us to define forcing for atomic formulas.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}hcomp\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isasymRightarrow}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isasymRightarrow}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}hcomp{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}f{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}g{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}w{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}z{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}g{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}f{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}trivial{\isacharparenright}{\kern0pt}\ hcomp{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ is{\isacharunderscore}{\kern0pt}f{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}a\ z{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}f{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ z\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ is{\isacharunderscore}{\kern0pt}g{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}a\ z{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}g{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ z\ {\isacharequal}{\kern0pt}\ g{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ g{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}a{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}g{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}hcomp{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}f{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}g{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}w{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ w\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}g{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}hcomp{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ hcomp{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}i{\isasymRightarrow}i{\isacharcomma}{\kern0pt}i{\isasymRightarrow}i{\isasymRightarrow}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}hcomp{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}pf{\isacharcomma}{\kern0pt}pg{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}w{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pg{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}pf{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}hcomp{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ f{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}a\ b\ z{\isachardot}{\kern0pt}\ a{\isasymin}nat\ {\isasymLongrightarrow}\ b{\isasymin}nat\ {\isasymLongrightarrow}\ z{\isasymin}M\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}f{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}b{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}pf{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\isanewline
\ \ \ \ g{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}a\ b\ z{\isachardot}{\kern0pt}\ a{\isasymin}nat\ {\isasymLongrightarrow}\ b{\isasymin}nat\ {\isasymLongrightarrow}\ z{\isasymin}M\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}g{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}b{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}pg{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\isanewline
\ \ \ \ {\isachardoublequoteopen}a{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}w{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}hcomp{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}pf{\isacharcomma}{\kern0pt}pg{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}hcomp{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}f{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}g{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}w{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ pf{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}w{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}w{\isasymin}nat{\isachardoublequoteclose}\ \isakeyword{for}\ x\ w\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}of\ {\isadigit{0}}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ x{\isacharbrackright}{\kern0pt}\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ pg{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}g{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}nat{\isachardoublequoteclose}\ \isakeyword{for}\ x\ a\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isadigit{0}}\ x{\isacharbrackright}{\kern0pt}\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}hcomp{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ftype\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ftype\ {\isasymequiv}\ fst{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ name{\isadigit{1}}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ fst{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ name{\isadigit{2}}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ fst{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ cond{\isacharunderscore}{\kern0pt}of\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ snd{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ components{\isacharunderscore}{\kern0pt}simp{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}ftype{\isacharparenleft}{\kern0pt}{\isasymlangle}f{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharcomma}{\kern0pt}n{\isadigit{2}}{\isacharcomma}{\kern0pt}c{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ f{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}{\isasymlangle}f{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharcomma}{\kern0pt}n{\isadigit{2}}{\isacharcomma}{\kern0pt}c{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}{\isasymlangle}f{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharcomma}{\kern0pt}n{\isadigit{2}}{\isacharcomma}{\kern0pt}c{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}{\isasymlangle}f{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharcomma}{\kern0pt}n{\isadigit{2}}{\isacharcomma}{\kern0pt}c{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ftype{\isacharunderscore}{\kern0pt}def\ name{\isadigit{1}}{\isacharunderscore}{\kern0pt}def\ name{\isadigit{2}}{\isacharunderscore}{\kern0pt}def\ cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}n\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}eclose{\isacharunderscore}{\kern0pt}n{\isacharparenleft}{\kern0pt}name{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}name{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ecloseN\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ecloseN{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ eclose{\isacharunderscore}{\kern0pt}n{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ eclose{\isacharunderscore}{\kern0pt}n{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ components{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}n{\isadigit{1}}\ {\isasymin}\ ecloseN{\isacharparenleft}{\kern0pt}{\isasymlangle}f{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharcomma}{\kern0pt}n{\isadigit{2}}{\isacharcomma}{\kern0pt}c{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}n{\isadigit{2}}\ {\isasymin}\ ecloseN{\isacharparenleft}{\kern0pt}{\isasymlangle}f{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharcomma}{\kern0pt}n{\isadigit{2}}{\isacharcomma}{\kern0pt}c{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ components{\isacharunderscore}{\kern0pt}simp\ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ names{\isacharunderscore}{\kern0pt}simp\ {\isacharequal}{\kern0pt}\ components{\isacharunderscore}{\kern0pt}simp{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ components{\isacharunderscore}{\kern0pt}simp{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ecloseNI{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}n{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymor}\ x{\isasymin}eclose{\isacharparenleft}{\kern0pt}n{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}x\ {\isasymin}\ ecloseN{\isacharparenleft}{\kern0pt}{\isasymlangle}f{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharcomma}{\kern0pt}n{\isadigit{2}}{\isacharcomma}{\kern0pt}c{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ eclose{\isacharunderscore}{\kern0pt}sing\ names{\isacharunderscore}{\kern0pt}simp\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ ecloseNI\ {\isacharequal}{\kern0pt}\ ecloseNI{\isadigit{1}}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}u\ {\isasymin}\ ecloseN{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ ecloseN{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ ecloseN{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}u\ {\isasymin}\ ecloseN{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}u{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}u{\isasymin}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}def\ \ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ cases\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}def\ \ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}singE{\isacharbrackleft}{\kern0pt}OF\ a{\isacharbrackright}{\kern0pt}\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}of\ u\ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}singE{\isacharbrackleft}{\kern0pt}OF\ b{\isacharbrackright}{\kern0pt}\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}of\ u\ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}fst\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}i{\isasymRightarrow}o{\isacharparenright}{\kern0pt}{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}fst{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharparenleft}{\kern0pt}{\isasymexists}z{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}t{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}z{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}w{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}w{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ empty{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ fst{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}fst{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Or{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Neg{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ \isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ fst{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}fst{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ fst{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ is{\isacharunderscore}{\kern0pt}ftype\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}i{\isasymRightarrow}o{\isacharparenright}{\kern0pt}{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}ftype\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}fst{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ftype{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ftype{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\ fst{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ \isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ ftype{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ftype{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}sats{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ aa{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}b{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ bb{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}aa{\isacharcomma}{\kern0pt}bb{\isacharparenright}{\kern0pt}\ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}snd\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}i{\isasymRightarrow}o{\isacharparenright}{\kern0pt}{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharparenleft}{\kern0pt}{\isasymexists}z{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}t{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}z{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}w{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}w{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ empty{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ snd{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}snd{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Or{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Neg{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ \isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ snd{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ snd{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}i{\isasymRightarrow}o{\isacharparenright}{\kern0pt}{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}hcomp{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}fst{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ \isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm\ sats{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}hcomp{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ A\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}fst{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ fst{\isacharunderscore}{\kern0pt}fm\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ aa{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}b{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ bb{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}aa{\isacharcomma}{\kern0pt}bb{\isacharparenright}{\kern0pt}\ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}sats{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}i{\isasymRightarrow}o{\isacharparenright}{\kern0pt}{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}hcomp{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}snd{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ \isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}hcomp{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ A\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ snd{\isacharunderscore}{\kern0pt}fm\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}i{\isasymRightarrow}o{\isacharparenright}{\kern0pt}{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{3}}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}hcomp{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}fst{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{3}}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ \isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}fm\ sats{\isacharunderscore}{\kern0pt}snd{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm\isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}hcomp{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ A\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}fst{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ fst{\isacharunderscore}{\kern0pt}fm\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ aa{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}b{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ bb{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}aa{\isacharcomma}{\kern0pt}bb{\isacharparenright}{\kern0pt}\ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}sats{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}i{\isasymRightarrow}o{\isacharparenright}{\kern0pt}{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{4}}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}hcomp{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{4}}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}t{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ \isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm\ sats{\isacharunderscore}{\kern0pt}snd{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm\isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}hcomp{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ A\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ snd{\isacharunderscore}{\kern0pt}fm\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ aa{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}b{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ bb{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}aa{\isacharcomma}{\kern0pt}bb{\isacharparenright}{\kern0pt}\ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}sats{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ components{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}a{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}ftype{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ ftype{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ fst{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ snd{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}components{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}fm\ sats{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm\ sats{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm\ sats{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm\isanewline
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ components{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ {\isacharequal}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}ftype{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\isanewline
\ \ is{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\isanewline
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ components{\isacharunderscore}{\kern0pt}defs\ {\isacharequal}{\kern0pt}\ fst{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ ftype{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ snd{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ snd{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ hcomp{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ cond{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}name{\isacharcomma}{\kern0pt}en{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ \ \ \ \ {\isasymexists}n{\isadigit{1}}{\isacharbrackleft}{\kern0pt}N{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}{\isasymexists}s{\isadigit{1}}{\isacharbrackleft}{\kern0pt}N{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}t{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}singleton{\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}n{\isadigit{1}}{\isacharcomma}{\kern0pt}s{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}s{\isadigit{1}}{\isacharcomma}{\kern0pt}en{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}t{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}t{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}ecloseN\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}ecloseN{\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}en{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}en{\isadigit{1}}{\isacharbrackleft}{\kern0pt}N{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}{\isasymexists}en{\isadigit{2}}{\isacharbrackleft}{\kern0pt}N{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharcomma}{\kern0pt}en{\isadigit{1}}{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharcomma}{\kern0pt}en{\isadigit{2}}{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ union{\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}en{\isadigit{1}}{\isacharcomma}{\kern0pt}en{\isadigit{2}}{\isacharcomma}{\kern0pt}en{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ ecloseN{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ecloseN{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}en{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}t{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}t{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}en{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ en\ {\isasymin}\ nat\ {\isacharsemicolon}{\kern0pt}\ t\ {\isasymin}\ nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ ecloseN{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}en{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ecloseN{\isacharunderscore}{\kern0pt}fm\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ en\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ t\ {\isasymin}\ nat\ {\isacharsemicolon}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ ecloseN{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}en{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}ecloseN{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}en{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}t{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ecloseN{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}ecloseN{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ \ nth{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ nth{\isacharunderscore}{\kern0pt}ConsI\ sats{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm\ sats{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm\ \isanewline
\ \ \ \ is{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ frecR\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymequiv}\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ \isanewline
\ \ \ \ \ \ {\isasymand}\ {\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymor}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ {\isasymor}\ {\isacharparenleft}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isadigit{1}}\ {\isasymand}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}ftypeD\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isacharparenleft}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecRI{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}s\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymor}\ s\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ n{\isadigit{2}}{\isacharcomma}{\kern0pt}\ q{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecRI{\isadigit{1}}{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}s\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ n{\isadigit{2}}{\isacharcomma}{\kern0pt}\ q{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecRI{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}s\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymor}\ s\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ n{\isadigit{2}}{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ n{\isadigit{2}}{\isacharcomma}{\kern0pt}\ q{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecRI{\isadigit{2}}{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}s\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ n{\isadigit{2}}{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ n{\isadigit{2}}{\isacharcomma}{\kern0pt}\ q{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecRI{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlangle}s{\isacharcomma}{\kern0pt}\ r{\isasymrangle}\ {\isasymin}\ n{\isadigit{2}}\ {\isasymLongrightarrow}\ frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ n{\isadigit{2}}{\isacharcomma}{\kern0pt}\ q{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecRI{\isadigit{3}}{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}s\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}n{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ n{\isadigit{1}}{\isacharcomma}{\kern0pt}\ n{\isadigit{2}}{\isacharcomma}{\kern0pt}\ q{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ \isanewline
\ \ \ \ \ \ {\isasymand}\ {\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymor}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ {\isasymor}\ {\isacharparenleft}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isadigit{1}}\ {\isasymand}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}D{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymor}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}D{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymLongrightarrow}\ ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isadigit{1}}\ {\isasymand}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}DI\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isacharcomma}{\kern0pt}c{\isacharcomma}{\kern0pt}d{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isacharcomma}{\kern0pt}c{\isacharcomma}{\kern0pt}d{\isasymrangle}{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}frecR\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}frecR{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}\ ftx{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}\ n{\isadigit{1}}x{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}n{\isadigit{2}}x{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}fty{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}n{\isadigit{1}}y{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}n{\isadigit{2}}y{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}dn{\isadigit{1}}{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}dn{\isadigit{2}}{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\isanewline
\ \ is{\isacharunderscore}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}ftx{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}n{\isadigit{1}}x{\isacharparenright}{\kern0pt}{\isasymand}\ is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}n{\isadigit{2}}x{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ is{\isacharunderscore}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}fty{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}n{\isadigit{1}}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}n{\isadigit{2}}y{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isasymand}\ is{\isacharunderscore}{\kern0pt}domain{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}n{\isadigit{1}}y{\isacharcomma}{\kern0pt}dn{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}domain{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}n{\isadigit{2}}y{\isacharcomma}{\kern0pt}dn{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}\ \ {\isacharparenleft}{\kern0pt}number{\isadigit{1}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}ftx{\isacharparenright}{\kern0pt}\ {\isasymand}\ empty{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}fty{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}n{\isadigit{1}}x\ {\isasymin}\ dn{\isadigit{1}}\ {\isasymor}\ n{\isadigit{1}}x\ {\isasymin}\ dn{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}n{\isadigit{2}}x\ {\isacharequal}{\kern0pt}\ n{\isadigit{1}}y\ {\isasymor}\ n{\isadigit{2}}x\ {\isacharequal}{\kern0pt}\ n{\isadigit{2}}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ {\isasymor}\ {\isacharparenleft}{\kern0pt}empty{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}ftx{\isacharparenright}{\kern0pt}\ {\isasymand}\ number{\isadigit{1}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}fty{\isacharparenright}{\kern0pt}\ {\isasymand}\ n{\isadigit{1}}x\ {\isacharequal}{\kern0pt}\ n{\isadigit{1}}y\ {\isasymand}\ n{\isadigit{2}}x\ {\isasymin}\ dn{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}frecR{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}i{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ b{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}frecR{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fr{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ \ is{\isacharunderscore}{\kern0pt}frecR{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}Collect{\isacharunderscore}{\kern0pt}def\ \ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules{\isacharprime}{\kern0pt}\ cartprod{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ \ components{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\isanewline
\ \ \ \ \ \ \ \ {\isacharbar}{\kern0pt}\ simp\ del{\isacharcolon}{\kern0pt}sats{\isacharunderscore}{\kern0pt}cartprod{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}frecR{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ sats{\isacharunderscore}{\kern0pt}frecR{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ eq{\isacharunderscore}{\kern0pt}ftypep{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}frecrR{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymnot}\ frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ frecR{\isacharunderscore}{\kern0pt}ftypeD\ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ rank{\isacharunderscore}{\kern0pt}names\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ max{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rank{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}types\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ max{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ Ord{\isacharunderscore}{\kern0pt}Un\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ mtype{\isacharunderscore}{\kern0pt}form\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}mtype{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ if\ rank{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ then\ {\isadigit{0}}\ else\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ type{\isacharunderscore}{\kern0pt}form\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ if\ ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ mtype{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}tc\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isadigit{3}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ mtype{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}rnk{\isacharunderscore}{\kern0pt}names\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymle}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ c\ d\ \ \isakeyword{where}\isanewline
\ \ \ \ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isacharequal}{\kern0pt}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}c\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ {\isacharequal}{\kern0pt}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}a\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isasymunion}domain{\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}b{\isacharequal}{\kern0pt}c\ {\isasymor}\ b\ {\isacharequal}{\kern0pt}\ d{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isacharparenleft}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ c\ {\isasymand}\ b\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}b\ {\isacharequal}{\kern0pt}\ c\ {\isasymor}\ b\ {\isacharequal}{\kern0pt}\ d{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\isanewline
\ \ \ \ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}b\ {\isacharequal}{\kern0pt}\ c\ {\isasymor}\ b\ {\isacharequal}{\kern0pt}\ d{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}o{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}b\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}\ {\isasymand}\ a\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ m\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}lt\ \ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}rank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ H\ m\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ max{\isacharunderscore}{\kern0pt}cong\ max{\isacharunderscore}{\kern0pt}cong{\isadigit{2}}\ leI\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ n\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}lt\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}rank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ H\ n\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ max{\isacharunderscore}{\kern0pt}cong{\isadigit{2}}\ max{\isacharunderscore}{\kern0pt}cong\ max{\isacharunderscore}{\kern0pt}commutes{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ leI\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ o\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}b\ {\isacharless}{\kern0pt}\ {\isacharquery}{\kern0pt}d{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}lt\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ H\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ max{\isacharunderscore}{\kern0pt}commutes\ max{\isacharunderscore}{\kern0pt}cong{\isadigit{2}}{\isacharbrackleft}{\kern0pt}OF\ leI{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}b\ {\isacharless}{\kern0pt}\ {\isacharquery}{\kern0pt}d{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ of\ {\isacharquery}{\kern0pt}a{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ {\isasymGamma}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}{\isasymGamma}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{3}}\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isasymGamma}{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}{\isasymGamma}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ {\isasymGamma}{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isasymGamma}{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymGamma}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ {\isasymGamma}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ F{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ {\isadigit{3}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ {\isadigit{3}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymle}\ rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}x\ {\isasymle}\ {\isacharquery}{\kern0pt}y{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ frecR{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}rnk{\isacharunderscore}{\kern0pt}names\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ max{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ leE{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}x{\isasymle}{\isacharquery}{\kern0pt}y{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymGamma}{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ oadd{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono{\isadigit{2}}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}x\ {\isacharless}{\kern0pt}\ {\isacharquery}{\kern0pt}y{\isacartoucheclose}\ F\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{2}}\isanewline
\ \ \ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymand}\ ftype{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ frecR{\isacharunderscore}{\kern0pt}ftypeD{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymor}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtau}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymsigma}{\isacharprime}{\kern0pt}\ {\isasymor}\ {\isacharquery}{\kern0pt}{\isasymtau}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtau}{\isacharprime}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymsigma}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ E{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ H\isanewline
\ \ \ \ \ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ \ {\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}lt\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ mtype{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}D{\isadigit{2}}{\isacharbrackleft}{\kern0pt}OF\ E\ a{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ E\ assms\ Ord{\isacharunderscore}{\kern0pt}rank\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ mtype{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}D{\isadigit{2}}{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ b{\isacharbrackright}{\kern0pt}\ max{\isacharunderscore}{\kern0pt}commutes\ E\ assms\ Ord{\isacharunderscore}{\kern0pt}rank\ disj{\isacharunderscore}{\kern0pt}commute\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ mtype{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacartoucheclose}\ {\isacartoucheopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymGamma}{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymsigma}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymsigma}{\isacharprime}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtau}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ frecR{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ eclose{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}lt\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ rank{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ max{\isacharunderscore}{\kern0pt}D{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ mtype{\isacharunderscore}{\kern0pt}form{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ rank{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacartoucheclose}\ {\isacartoucheopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}type{\isacharunderscore}{\kern0pt}form{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymGamma}{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ frecrel\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}frecrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Rrel{\isacharparenleft}{\kern0pt}frecR{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecrelI\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}frecrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ frecrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecrelD\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ frecrel{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isasymtimes}A{\isadigit{2}}{\isasymtimes}A{\isadigit{3}}{\isasymtimes}A{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isadigit{1}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isadigit{3}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isadigit{3}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isadigit{4}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isadigit{4}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ frecrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ ftype{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}frecrel\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}wf{\isacharparenleft}{\kern0pt}frecrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}frecrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ measure{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymGamma}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ frecrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ measure{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymGamma}{\isacharunderscore}{\kern0pt}mono\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}subset\ wf{\isacharunderscore}{\kern0pt}measure\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ core{\isacharunderscore}{\kern0pt}induction{\isacharunderscore}{\kern0pt}aux{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ A{\isadigit{1}}\ A{\isadigit{2}}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}{\isasymtau}\ {\isasymtheta}\ p{\isachardot}{\kern0pt}\ \ p\ {\isasymin}\ A{\isadigit{2}}\ {\isasymLongrightarrow}\ {\isasymlbrakk}{\isasymAnd}q\ {\isasymsigma}{\isachardot}{\kern0pt}\ {\isasymlbrakk}\ q{\isasymin}A{\isadigit{2}}\ {\isacharsemicolon}{\kern0pt}\ {\isasymsigma}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isasymrbrakk}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isasymrbrakk}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}{\isasymtau}\ {\isasymtheta}\ p{\isachardot}{\kern0pt}\ \ p\ {\isasymin}\ A{\isadigit{2}}\ {\isasymLongrightarrow}\ {\isasymlbrakk}{\isasymAnd}q\ {\isasymsigma}{\isachardot}{\kern0pt}\ {\isasymlbrakk}\ q{\isasymin}A{\isadigit{2}}\ {\isacharsemicolon}{\kern0pt}\ {\isasymsigma}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isasymrbrakk}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isasymrbrakk}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}a{\isasymin}{\isadigit{2}}{\isasymtimes}A{\isadigit{1}}{\isasymtimes}A{\isadigit{1}}{\isasymtimes}A{\isadigit{2}}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ a\ rule{\isacharcolon}{\kern0pt}wf{\isacharunderscore}{\kern0pt}induct{\isacharbrackleft}{\kern0pt}OF\ wf{\isacharunderscore}{\kern0pt}frecrel{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isadigit{2}}{\isasymtimes}A{\isadigit{1}}{\isasymtimes}A{\isadigit{1}}{\isasymtimes}A{\isadigit{2}}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}\ x{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymtau}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymtheta}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}D\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isadigit{2}}{\isasymtimes}A{\isadigit{1}}{\isasymtimes}A{\isadigit{1}}{\isasymtimes}A{\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharquery}{\kern0pt}D{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}A{\isadigit{2}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}{\isacharquery}{\kern0pt}D{\isacartoucheclose}\isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}{\isadigit{0}}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}mem{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}{\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ cases\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ eq\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isasymsigma}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isasymsigma}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}q{\isasymin}A{\isadigit{2}}{\isachardoublequoteclose}\ \isakeyword{for}\ q\ {\isasymsigma}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}A{\isadigit{1}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}eclose{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}eclose{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ \ {\isacartoucheopen}Transset{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ that{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}eclose{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ eclose{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}A{\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}subset{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}subset{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ Transset{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}arg{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}Transset{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtheta}\ {\isasymin}\ A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isasymsigma}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}T{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isasymsigma}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}U{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ frecRI{\isadigit{1}}{\isacharprime}{\kern0pt}{\isacharbrackleft}{\kern0pt}OF\ that{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ frecR{\isacharunderscore}{\kern0pt}DI\ \ {\isacartoucheopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ frecRI{\isadigit{2}}{\isacharprime}{\kern0pt}{\isacharbrackleft}{\kern0pt}OF\ that{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}{\isacharquery}{\kern0pt}D{\isacartoucheclose}\ {\isacartoucheopen}{\isasymsigma}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}q{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isacharquery}{\kern0pt}T{\isacharcomma}{\kern0pt}x{\isasymrangle}{\isasymin}\ frecrel{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isacharquery}{\kern0pt}U{\isacharcomma}{\kern0pt}x{\isasymrangle}{\isasymin}\ frecrel{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ frecrelI{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}T\ {\isacharquery}{\kern0pt}D\ x{\isacharbrackright}{\kern0pt}\ \ frecrelI{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}U\ {\isacharquery}{\kern0pt}D\ x{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymsigma}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isasymsigma}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymsigma}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymlangle}{\isacharquery}{\kern0pt}U{\isacharcomma}{\kern0pt}x{\isasymrangle}{\isasymin}\ frecrel{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isasymsigma}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ A\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}\ {\isacartoucheopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isacartoucheclose}\ {\isacartoucheopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ mem\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}\ \ {\isasymsigma}{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}q{\isasymin}A{\isadigit{2}}{\isachardoublequoteclose}\ \isakeyword{for}\ q\ {\isasymsigma}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isadigit{1}}\ assms\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}A{\isadigit{1}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}A{\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}eclose{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}eclose{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ \ {\isacartoucheopen}Transset{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ that{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}A{\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}subset{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ Transset{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}arg{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}Transset{\isacharparenleft}{\kern0pt}A{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtheta}\ {\isasymin}\ A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}\ {\isasymsigma}{\isacharcomma}{\kern0pt}\ q{\isasymrangle}{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}frecR{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}T{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ frecRI{\isadigit{3}}{\isacharprime}{\kern0pt}{\isacharbrackleft}{\kern0pt}OF\ that{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ frecR{\isacharunderscore}{\kern0pt}DI\ \ {\isacartoucheopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacartoucheclose}\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}{\isacharquery}{\kern0pt}D{\isacartoucheclose}\ {\isacartoucheopen}{\isasymsigma}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}q{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isacharquery}{\kern0pt}T{\isacharcomma}{\kern0pt}x{\isasymrangle}{\isasymin}\ frecrel{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}T{\isasymin}{\isacharquery}{\kern0pt}D{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ frecrelI{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}T\ {\isacharquery}{\kern0pt}D\ x{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymsigma}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtau}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isasymin}A{\isadigit{1}}{\isacartoucheclose}\ {\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force\ simp\ add{\isacharcolon}{\kern0pt}components{\isacharunderscore}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isacartoucheopen}ftype{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacartoucheclose}\ {\isacartoucheopen}cond{\isacharunderscore}{\kern0pt}of{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}A{\isadigit{2}}{\isacartoucheclose}\ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}frecrel\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}frecrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}z{\isasymin}A{\isasymtimes}A{\isachardot}{\kern0pt}\ {\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymand}\ frecR{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ frecrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ frecrel{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}snd{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}frecrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}z\ {\isasymin}\ A{\isasymtimes}A\ {\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ ftype{\isacharparenleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ ftype{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{1}}{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isasymor}\ {\isacharparenleft}{\kern0pt}ftype{\isacharparenleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ ftype{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymand}\ \ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ name{\isadigit{1}}{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ name{\isadigit{2}}{\isacharparenleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}name{\isadigit{2}}{\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}frecrel\ frecR{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ equalityI\ subsetI\ CollectI{\isacharsemicolon}{\kern0pt}\ elim\ CollectE{\isacharsemicolon}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/FrecR.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%33=2%:%
%:%36=3%:%
%:%37=4%:%
%:%38=4%:%
%:%39=5%:%
%:%41=7%:%
%:%42=8%:%
%:%44=11%:%
%:%45=11%:%
%:%46=12%:%
%:%47=13%:%
%:%48=14%:%
%:%49=15%:%
%:%50=15%:%
%:%51=16%:%
%:%52=17%:%
%:%53=18%:%
%:%54=19%:%
%:%55=20%:%
%:%56=21%:%
%:%57=22%:%
%:%60=23%:%
%:%64=23%:%
%:%65=23%:%
%:%66=23%:%
%:%67=23%:%
%:%72=23%:%
%:%75=24%:%
%:%76=25%:%
%:%77=25%:%
%:%78=26%:%
%:%79=27%:%
%:%80=28%:%
%:%81=29%:%
%:%82=29%:%
%:%83=30%:%
%:%84=31%:%
%:%85=32%:%
%:%86=33%:%
%:%87=34%:%
%:%88=35%:%
%:%89=36%:%
%:%90=37%:%
%:%91=38%:%
%:%92=39%:%
%:%99=40%:%
%:%100=40%:%
%:%101=41%:%
%:%102=41%:%
%:%103=42%:%
%:%104=42%:%
%:%105=42%:%
%:%106=43%:%
%:%107=43%:%
%:%108=44%:%
%:%109=44%:%
%:%110=45%:%
%:%111=45%:%
%:%112=45%:%
%:%113=46%:%
%:%114=46%:%
%:%115=47%:%
%:%116=47%:%
%:%117=47%:%
%:%118=47%:%
%:%119=47%:%
%:%120=48%:%
%:%126=48%:%
%:%129=49%:%
%:%130=50%:%
%:%131=51%:%
%:%132=52%:%
%:%133=52%:%
%:%134=53%:%
%:%135=54%:%
%:%136=55%:%
%:%137=56%:%
%:%138=56%:%
%:%139=57%:%
%:%140=58%:%
%:%141=59%:%
%:%142=60%:%
%:%143=60%:%
%:%144=61%:%
%:%145=62%:%
%:%146=63%:%
%:%147=64%:%
%:%148=64%:%
%:%149=65%:%
%:%150=66%:%
%:%151=67%:%
%:%152=68%:%
%:%153=68%:%
%:%154=69%:%
%:%155=70%:%
%:%156=71%:%
%:%157=72%:%
%:%160=73%:%
%:%164=73%:%
%:%165=73%:%
%:%166=74%:%
%:%167=74%:%
%:%172=74%:%
%:%175=75%:%
%:%176=76%:%
%:%177=76%:%
%:%178=77%:%
%:%179=78%:%
%:%180=79%:%
%:%181=79%:%
%:%182=80%:%
%:%183=81%:%
%:%184=82%:%
%:%185=83%:%
%:%186=83%:%
%:%187=84%:%
%:%188=85%:%
%:%191=86%:%
%:%195=86%:%
%:%196=86%:%
%:%197=87%:%
%:%198=87%:%
%:%199=87%:%
%:%204=87%:%
%:%207=88%:%
%:%208=89%:%
%:%209=89%:%
%:%210=90%:%
%:%211=91%:%
%:%212=91%:%
%:%213=92%:%
%:%214=93%:%
%:%217=94%:%
%:%221=94%:%
%:%222=94%:%
%:%223=95%:%
%:%224=95%:%
%:%225=96%:%
%:%226=96%:%
%:%231=96%:%
%:%234=97%:%
%:%235=98%:%
%:%236=98%:%
%:%237=99%:%
%:%238=100%:%
%:%239=100%:%
%:%240=101%:%
%:%241=102%:%
%:%248=103%:%
%:%249=103%:%
%:%250=104%:%
%:%251=104%:%
%:%252=105%:%
%:%253=105%:%
%:%254=106%:%
%:%255=106%:%
%:%256=106%:%
%:%257=107%:%
%:%258=107%:%
%:%259=108%:%
%:%260=108%:%
%:%261=109%:%
%:%262=109%:%
%:%263=110%:%
%:%264=110%:%
%:%265=111%:%
%:%266=111%:%
%:%267=112%:%
%:%268=112%:%
%:%269=113%:%
%:%270=113%:%
%:%271=114%:%
%:%272=114%:%
%:%273=114%:%
%:%274=115%:%
%:%275=115%:%
%:%276=116%:%
%:%277=116%:%
%:%278=117%:%
%:%279=117%:%
%:%280=118%:%
%:%281=118%:%
%:%282=119%:%
%:%283=119%:%
%:%284=120%:%
%:%285=120%:%
%:%286=120%:%
%:%287=121%:%
%:%288=121%:%
%:%289=122%:%
%:%295=122%:%
%:%298=123%:%
%:%299=124%:%
%:%300=125%:%
%:%301=126%:%
%:%302=127%:%
%:%303=127%:%
%:%304=128%:%
%:%305=129%:%
%:%306=130%:%
%:%307=131%:%
%:%308=132%:%
%:%309=132%:%
%:%310=133%:%
%:%311=134%:%
%:%312=135%:%
%:%313=136%:%
%:%314=137%:%
%:%315=137%:%
%:%316=138%:%
%:%318=140%:%
%:%321=141%:%
%:%325=141%:%
%:%326=141%:%
%:%331=141%:%
%:%334=142%:%
%:%335=143%:%
%:%336=143%:%
%:%337=144%:%
%:%338=145%:%
%:%339=146%:%
%:%340=147%:%
%:%341=147%:%
%:%342=148%:%
%:%343=149%:%
%:%344=150%:%
%:%345=151%:%
%:%346=151%:%
%:%347=152%:%
%:%349=154%:%
%:%352=155%:%
%:%356=155%:%
%:%357=155%:%
%:%358=156%:%
%:%359=156%:%
%:%364=156%:%
%:%367=157%:%
%:%368=158%:%
%:%369=158%:%
%:%370=159%:%
%:%371=160%:%
%:%372=161%:%
%:%373=162%:%
%:%376=163%:%
%:%380=163%:%
%:%381=163%:%
%:%382=164%:%
%:%383=164%:%
%:%388=164%:%
%:%391=165%:%
%:%392=166%:%
%:%393=166%:%
%:%394=167%:%
%:%395=168%:%
%:%396=169%:%
%:%397=170%:%
%:%398=171%:%
%:%399=171%:%
%:%400=172%:%
%:%401=173%:%
%:%402=174%:%
%:%403=175%:%
%:%404=176%:%
%:%405=176%:%
%:%406=177%:%
%:%408=179%:%
%:%411=180%:%
%:%415=180%:%
%:%416=180%:%
%:%421=180%:%
%:%424=181%:%
%:%425=182%:%
%:%426=182%:%
%:%427=183%:%
%:%428=184%:%
%:%429=185%:%
%:%430=186%:%
%:%431=186%:%
%:%432=187%:%
%:%433=188%:%
%:%434=189%:%
%:%435=190%:%
%:%436=190%:%
%:%437=191%:%
%:%439=193%:%
%:%442=194%:%
%:%446=194%:%
%:%447=194%:%
%:%448=194%:%
%:%449=195%:%
%:%450=195%:%
%:%455=195%:%
%:%458=196%:%
%:%459=197%:%
%:%460=197%:%
%:%461=198%:%
%:%462=199%:%
%:%463=200%:%
%:%464=201%:%
%:%467=202%:%
%:%471=202%:%
%:%472=202%:%
%:%473=203%:%
%:%474=203%:%
%:%479=203%:%
%:%482=204%:%
%:%483=205%:%
%:%484=205%:%
%:%485=206%:%
%:%486=207%:%
%:%487=208%:%
%:%488=209%:%
%:%489=209%:%
%:%490=210%:%
%:%491=211%:%
%:%492=212%:%
%:%493=213%:%
%:%494=213%:%
%:%495=214%:%
%:%497=216%:%
%:%500=217%:%
%:%504=217%:%
%:%505=217%:%
%:%506=217%:%
%:%507=218%:%
%:%508=218%:%
%:%513=218%:%
%:%516=219%:%
%:%517=220%:%
%:%518=220%:%
%:%519=221%:%
%:%520=222%:%
%:%521=223%:%
%:%522=224%:%
%:%523=224%:%
%:%524=225%:%
%:%525=226%:%
%:%526=227%:%
%:%527=228%:%
%:%528=228%:%
%:%529=229%:%
%:%531=231%:%
%:%534=232%:%
%:%538=232%:%
%:%539=232%:%
%:%540=232%:%
%:%541=233%:%
%:%542=233%:%
%:%547=233%:%
%:%550=234%:%
%:%551=235%:%
%:%552=235%:%
%:%553=236%:%
%:%554=237%:%
%:%555=238%:%
%:%556=239%:%
%:%559=240%:%
%:%563=240%:%
%:%564=240%:%
%:%565=241%:%
%:%566=241%:%
%:%571=241%:%
%:%574=242%:%
%:%575=243%:%
%:%576=243%:%
%:%577=244%:%
%:%578=245%:%
%:%579=246%:%
%:%580=247%:%
%:%581=247%:%
%:%582=248%:%
%:%583=249%:%
%:%584=250%:%
%:%585=251%:%
%:%586=251%:%
%:%587=252%:%
%:%589=254%:%
%:%592=255%:%
%:%596=255%:%
%:%597=255%:%
%:%598=255%:%
%:%599=256%:%
%:%600=256%:%
%:%605=256%:%
%:%608=257%:%
%:%609=258%:%
%:%610=258%:%
%:%611=259%:%
%:%612=260%:%
%:%613=261%:%
%:%614=262%:%
%:%617=263%:%
%:%621=263%:%
%:%622=263%:%
%:%623=264%:%
%:%624=264%:%
%:%629=264%:%
%:%632=265%:%
%:%633=266%:%
%:%634=266%:%
%:%635=267%:%
%:%636=268%:%
%:%637=269%:%
%:%638=270%:%
%:%639=271%:%
%:%640=272%:%
%:%643=273%:%
%:%647=273%:%
%:%648=273%:%
%:%649=274%:%
%:%650=274%:%
%:%651=275%:%
%:%652=276%:%
%:%653=276%:%
%:%658=276%:%
%:%661=277%:%
%:%662=278%:%
%:%663=278%:%
%:%664=279%:%
%:%665=280%:%
%:%666=280%:%
%:%667=281%:%
%:%668=282%:%
%:%669=283%:%
%:%670=283%:%
%:%671=284%:%
%:%672=285%:%
%:%673=286%:%
%:%674=287%:%
%:%675=287%:%
%:%676=288%:%
%:%677=289%:%
%:%678=290%:%
%:%679=291%:%
%:%680=292%:%
%:%681=292%:%
%:%682=293%:%
%:%683=294%:%
%:%684=295%:%
%:%685=296%:%
%:%686=297%:%
%:%687=297%:%
%:%688=298%:%
%:%689=299%:%
%:%690=300%:%
%:%691=301%:%
%:%692=302%:%
%:%693=302%:%
%:%694=303%:%
%:%695=304%:%
%:%697=306%:%
%:%698=307%:%
%:%699=308%:%
%:%700=308%:%
%:%701=309%:%
%:%702=310%:%
%:%703=311%:%
%:%704=312%:%
%:%705=312%:%
%:%706=313%:%
%:%709=314%:%
%:%713=314%:%
%:%714=314%:%
%:%715=314%:%
%:%720=314%:%
%:%723=315%:%
%:%724=316%:%
%:%725=316%:%
%:%726=317%:%
%:%727=318%:%
%:%730=319%:%
%:%734=319%:%
%:%735=319%:%
%:%736=320%:%
%:%737=320%:%
%:%738=321%:%
%:%739=322%:%
%:%740=322%:%
%:%745=322%:%
%:%748=323%:%
%:%749=324%:%
%:%750=325%:%
%:%751=325%:%
%:%752=326%:%
%:%753=327%:%
%:%756=330%:%
%:%757=331%:%
%:%758=332%:%
%:%759=332%:%
%:%760=333%:%
%:%761=334%:%
%:%764=335%:%
%:%768=335%:%
%:%769=335%:%
%:%770=335%:%
%:%771=335%:%
%:%776=335%:%
%:%779=336%:%
%:%780=337%:%
%:%781=337%:%
%:%784=338%:%
%:%788=338%:%
%:%789=338%:%
%:%790=338%:%
%:%795=338%:%
%:%798=339%:%
%:%799=340%:%
%:%800=340%:%
%:%803=341%:%
%:%807=341%:%
%:%808=341%:%
%:%809=341%:%
%:%814=341%:%
%:%817=342%:%
%:%818=343%:%
%:%819=343%:%
%:%822=344%:%
%:%826=344%:%
%:%827=344%:%
%:%828=344%:%
%:%833=344%:%
%:%836=345%:%
%:%837=346%:%
%:%838=346%:%
%:%841=347%:%
%:%845=347%:%
%:%846=347%:%
%:%847=347%:%
%:%852=347%:%
%:%855=348%:%
%:%856=349%:%
%:%857=350%:%
%:%858=350%:%
%:%861=351%:%
%:%865=351%:%
%:%866=351%:%
%:%867=351%:%
%:%872=351%:%
%:%875=352%:%
%:%876=353%:%
%:%877=353%:%
%:%880=354%:%
%:%884=354%:%
%:%885=354%:%
%:%886=354%:%
%:%891=354%:%
%:%894=355%:%
%:%895=356%:%
%:%896=356%:%
%:%897=357%:%
%:%900=360%:%
%:%903=361%:%
%:%907=361%:%
%:%908=361%:%
%:%914=361%:%
%:%917=362%:%
%:%918=363%:%
%:%919=363%:%
%:%920=364%:%
%:%921=365%:%
%:%924=366%:%
%:%928=366%:%
%:%929=366%:%
%:%930=367%:%
%:%931=367%:%
%:%936=367%:%
%:%939=368%:%
%:%940=369%:%
%:%941=369%:%
%:%942=370%:%
%:%943=371%:%
%:%946=372%:%
%:%950=372%:%
%:%951=372%:%
%:%952=373%:%
%:%953=373%:%
%:%958=373%:%
%:%961=374%:%
%:%962=375%:%
%:%963=375%:%
%:%964=376%:%
%:%965=377%:%
%:%968=378%:%
%:%972=378%:%
%:%973=378%:%
%:%974=378%:%
%:%975=378%:%
%:%980=378%:%
%:%983=379%:%
%:%984=383%:%
%:%985=384%:%
%:%986=384%:%
%:%987=385%:%
%:%988=386%:%
%:%993=391%:%
%:%994=392%:%
%:%995=393%:%
%:%996=393%:%
%:%997=394%:%
%:%998=395%:%
%:%999=396%:%
%:%1000=397%:%
%:%1003=398%:%
%:%1007=398%:%
%:%1008=398%:%
%:%1009=399%:%
%:%1010=399%:%
%:%1011=400%:%
%:%1016=400%:%
%:%1021=401%:%
%:%1026=402%:%
%:%1027=402%:%
%:%1032=402%:%
%:%1035=403%:%
%:%1036=404%:%
%:%1037=405%:%
%:%1038=405%:%
%:%1039=406%:%
%:%1040=407%:%
%:%1043=408%:%
%:%1047=408%:%
%:%1048=408%:%
%:%1049=408%:%
%:%1054=408%:%
%:%1057=409%:%
%:%1058=410%:%
%:%1059=411%:%
%:%1060=411%:%
%:%1061=412%:%
%:%1062=413%:%
%:%1063=414%:%
%:%1064=415%:%
%:%1065=415%:%
%:%1066=416%:%
%:%1069=417%:%
%:%1073=417%:%
%:%1074=417%:%
%:%1075=417%:%
%:%1076=417%:%
%:%1081=417%:%
%:%1084=418%:%
%:%1085=419%:%
%:%1086=419%:%
%:%1087=420%:%
%:%1088=421%:%
%:%1089=422%:%
%:%1090=423%:%
%:%1091=423%:%
%:%1092=424%:%
%:%1093=425%:%
%:%1094=426%:%
%:%1095=427%:%
%:%1096=427%:%
%:%1097=428%:%
%:%1100=429%:%
%:%1104=429%:%
%:%1105=429%:%
%:%1106=429%:%
%:%1111=429%:%
%:%1114=430%:%
%:%1115=431%:%
%:%1116=431%:%
%:%1117=432%:%
%:%1118=433%:%
%:%1125=434%:%
%:%1126=434%:%
%:%1127=435%:%
%:%1128=435%:%
%:%1129=436%:%
%:%1130=437%:%
%:%1131=438%:%
%:%1132=439%:%
%:%1133=439%:%
%:%1134=439%:%
%:%1135=439%:%
%:%1136=440%:%
%:%1137=440%:%
%:%1138=441%:%
%:%1139=441%:%
%:%1140=442%:%
%:%1141=443%:%
%:%1142=444%:%
%:%1143=445%:%
%:%1144=445%:%
%:%1145=446%:%
%:%1146=446%:%
%:%1147=446%:%
%:%1148=446%:%
%:%1149=447%:%
%:%1150=447%:%
%:%1151=448%:%
%:%1152=448%:%
%:%1153=449%:%
%:%1154=449%:%
%:%1155=450%:%
%:%1156=450%:%
%:%1157=450%:%
%:%1158=451%:%
%:%1159=451%:%
%:%1160=452%:%
%:%1161=452%:%
%:%1162=452%:%
%:%1163=452%:%
%:%1164=452%:%
%:%1165=453%:%
%:%1166=453%:%
%:%1167=454%:%
%:%1168=454%:%
%:%1169=455%:%
%:%1170=455%:%
%:%1171=456%:%
%:%1172=456%:%
%:%1173=457%:%
%:%1174=457%:%
%:%1175=457%:%
%:%1176=458%:%
%:%1177=458%:%
%:%1178=459%:%
%:%1179=459%:%
%:%1180=459%:%
%:%1181=460%:%
%:%1182=460%:%
%:%1183=460%:%
%:%1184=461%:%
%:%1185=461%:%
%:%1186=462%:%
%:%1187=462%:%
%:%1188=463%:%
%:%1189=463%:%
%:%1190=464%:%
%:%1191=464%:%
%:%1192=465%:%
%:%1193=465%:%
%:%1194=465%:%
%:%1195=466%:%
%:%1196=466%:%
%:%1197=467%:%
%:%1198=467%:%
%:%1199=467%:%
%:%1200=468%:%
%:%1201=468%:%
%:%1202=468%:%
%:%1203=469%:%
%:%1204=469%:%
%:%1205=470%:%
%:%1211=470%:%
%:%1214=471%:%
%:%1215=472%:%
%:%1216=473%:%
%:%1217=473%:%
%:%1218=474%:%
%:%1219=475%:%
%:%1220=476%:%
%:%1221=477%:%
%:%1222=477%:%
%:%1223=478%:%
%:%1226=479%:%
%:%1230=479%:%
%:%1231=479%:%
%:%1232=479%:%
%:%1237=479%:%
%:%1240=480%:%
%:%1241=481%:%
%:%1242=482%:%
%:%1243=482%:%
%:%1244=483%:%
%:%1245=484%:%
%:%1252=485%:%
%:%1253=485%:%
%:%1254=486%:%
%:%1255=486%:%
%:%1256=487%:%
%:%1257=487%:%
%:%1258=487%:%
%:%1259=488%:%
%:%1260=488%:%
%:%1261=489%:%
%:%1262=489%:%
%:%1263=490%:%
%:%1264=490%:%
%:%1265=490%:%
%:%1266=491%:%
%:%1267=491%:%
%:%1268=492%:%
%:%1269=492%:%
%:%1270=492%:%
%:%1271=492%:%
%:%1272=492%:%
%:%1273=493%:%
%:%1274=493%:%
%:%1275=494%:%
%:%1276=494%:%
%:%1277=495%:%
%:%1278=495%:%
%:%1279=496%:%
%:%1280=496%:%
%:%1281=497%:%
%:%1282=497%:%
%:%1283=498%:%
%:%1284=498%:%
%:%1285=499%:%
%:%1286=499%:%
%:%1287=499%:%
%:%1288=499%:%
%:%1289=499%:%
%:%1290=500%:%
%:%1291=500%:%
%:%1292=501%:%
%:%1293=501%:%
%:%1294=502%:%
%:%1295=502%:%
%:%1296=503%:%
%:%1297=503%:%
%:%1298=503%:%
%:%1299=504%:%
%:%1300=504%:%
%:%1301=504%:%
%:%1302=504%:%
%:%1303=505%:%
%:%1304=505%:%
%:%1305=506%:%
%:%1306=506%:%
%:%1307=507%:%
%:%1308=507%:%
%:%1309=508%:%
%:%1310=508%:%
%:%1311=508%:%
%:%1312=509%:%
%:%1313=509%:%
%:%1314=510%:%
%:%1315=510%:%
%:%1316=511%:%
%:%1317=512%:%
%:%1318=513%:%
%:%1319=513%:%
%:%1320=513%:%
%:%1321=513%:%
%:%1322=514%:%
%:%1323=514%:%
%:%1324=515%:%
%:%1325=515%:%
%:%1326=515%:%
%:%1327=516%:%
%:%1328=516%:%
%:%1329=517%:%
%:%1330=517%:%
%:%1331=518%:%
%:%1332=518%:%
%:%1333=518%:%
%:%1334=519%:%
%:%1335=519%:%
%:%1336=520%:%
%:%1337=520%:%
%:%1338=520%:%
%:%1339=521%:%
%:%1340=521%:%
%:%1341=522%:%
%:%1342=522%:%
%:%1343=523%:%
%:%1344=523%:%
%:%1345=523%:%
%:%1346=523%:%
%:%1347=524%:%
%:%1348=524%:%
%:%1349=525%:%
%:%1350=525%:%
%:%1351=526%:%
%:%1352=526%:%
%:%1353=527%:%
%:%1354=527%:%
%:%1355=528%:%
%:%1356=528%:%
%:%1357=528%:%
%:%1358=529%:%
%:%1359=529%:%
%:%1360=529%:%
%:%1361=530%:%
%:%1362=530%:%
%:%1363=531%:%
%:%1364=531%:%
%:%1365=532%:%
%:%1366=532%:%
%:%1367=532%:%
%:%1368=532%:%
%:%1369=533%:%
%:%1370=533%:%
%:%1371=534%:%
%:%1372=534%:%
%:%1373=535%:%
%:%1374=535%:%
%:%1375=535%:%
%:%1376=536%:%
%:%1377=536%:%
%:%1378=537%:%
%:%1379=537%:%
%:%1380=538%:%
%:%1381=538%:%
%:%1382=539%:%
%:%1383=539%:%
%:%1384=540%:%
%:%1385=541%:%
%:%1386=542%:%
%:%1387=542%:%
%:%1388=542%:%
%:%1389=542%:%
%:%1390=543%:%
%:%1391=543%:%
%:%1392=544%:%
%:%1393=544%:%
%:%1394=545%:%
%:%1395=545%:%
%:%1396=545%:%
%:%1397=546%:%
%:%1398=546%:%
%:%1399=547%:%
%:%1400=547%:%
%:%1401=548%:%
%:%1402=548%:%
%:%1403=548%:%
%:%1404=548%:%
%:%1405=549%:%
%:%1406=549%:%
%:%1407=550%:%
%:%1408=550%:%
%:%1409=551%:%
%:%1410=551%:%
%:%1411=551%:%
%:%1412=551%:%
%:%1413=552%:%
%:%1414=552%:%
%:%1415=553%:%
%:%1416=553%:%
%:%1417=554%:%
%:%1418=554%:%
%:%1419=554%:%
%:%1420=555%:%
%:%1421=555%:%
%:%1422=556%:%
%:%1423=556%:%
%:%1424=557%:%
%:%1430=557%:%
%:%1433=558%:%
%:%1434=559%:%
%:%1435=559%:%
%:%1436=560%:%
%:%1437=561%:%
%:%1438=562%:%
%:%1439=563%:%
%:%1440=563%:%
%:%1441=564%:%
%:%1442=565%:%
%:%1445=566%:%
%:%1449=566%:%
%:%1450=566%:%
%:%1451=566%:%
%:%1452=566%:%
%:%1457=566%:%
%:%1460=567%:%
%:%1461=568%:%
%:%1462=568%:%
%:%1463=569%:%
%:%1464=570%:%
%:%1465=571%:%
%:%1466=572%:%
%:%1467=573%:%
%:%1470=574%:%
%:%1474=574%:%
%:%1475=574%:%
%:%1476=574%:%
%:%1477=574%:%
%:%1482=574%:%
%:%1485=575%:%
%:%1486=576%:%
%:%1487=576%:%
%:%1488=577%:%
%:%1495=578%:%
%:%1496=578%:%
%:%1497=579%:%
%:%1498=579%:%
%:%1499=580%:%
%:%1500=580%:%
%:%1501=581%:%
%:%1502=581%:%
%:%1503=581%:%
%:%1504=582%:%
%:%1505=582%:%
%:%1506=582%:%
%:%1507=582%:%
%:%1508=582%:%
%:%1509=583%:%
%:%1515=583%:%
%:%1518=584%:%
%:%1519=585%:%
%:%1520=585%:%
%:%1521=586%:%
%:%1522=587%:%
%:%1523=588%:%
%:%1524=589%:%
%:%1525=590%:%
%:%1526=591%:%
%:%1533=592%:%
%:%1534=592%:%
%:%1535=593%:%
%:%1536=593%:%
%:%1537=594%:%
%:%1538=594%:%
%:%1539=595%:%
%:%1540=595%:%
%:%1541=596%:%
%:%1542=596%:%
%:%1543=597%:%
%:%1544=597%:%
%:%1545=598%:%
%:%1546=598%:%
%:%1547=599%:%
%:%1548=599%:%
%:%1549=600%:%
%:%1550=600%:%
%:%1551=601%:%
%:%1552=601%:%
%:%1553=602%:%
%:%1554=602%:%
%:%1555=603%:%
%:%1556=603%:%
%:%1557=604%:%
%:%1558=604%:%
%:%1559=605%:%
%:%1560=605%:%
%:%1561=606%:%
%:%1562=606%:%
%:%1563=607%:%
%:%1564=607%:%
%:%1565=608%:%
%:%1566=608%:%
%:%1567=609%:%
%:%1568=609%:%
%:%1569=610%:%
%:%1570=610%:%
%:%1571=611%:%
%:%1572=611%:%
%:%1573=612%:%
%:%1574=612%:%
%:%1575=613%:%
%:%1576=613%:%
%:%1577=613%:%
%:%1578=614%:%
%:%1579=614%:%
%:%1580=615%:%
%:%1581=615%:%
%:%1582=616%:%
%:%1583=616%:%
%:%1584=616%:%
%:%1585=617%:%
%:%1586=617%:%
%:%1587=618%:%
%:%1588=618%:%
%:%1589=619%:%
%:%1590=619%:%
%:%1591=620%:%
%:%1592=621%:%
%:%1593=621%:%
%:%1594=622%:%
%:%1595=622%:%
%:%1596=623%:%
%:%1597=623%:%
%:%1598=624%:%
%:%1599=625%:%
%:%1600=625%:%
%:%1601=626%:%
%:%1602=627%:%
%:%1603=627%:%
%:%1604=628%:%
%:%1605=628%:%
%:%1606=629%:%
%:%1607=629%:%
%:%1608=630%:%
%:%1609=630%:%
%:%1610=630%:%
%:%1611=631%:%
%:%1612=631%:%
%:%1613=632%:%
%:%1614=632%:%
%:%1615=632%:%
%:%1616=632%:%
%:%1617=633%:%
%:%1618=633%:%
%:%1619=633%:%
%:%1620=634%:%
%:%1621=634%:%
%:%1622=634%:%
%:%1623=634%:%
%:%1624=635%:%
%:%1625=635%:%
%:%1626=636%:%
%:%1627=636%:%
%:%1628=636%:%
%:%1629=636%:%
%:%1630=637%:%
%:%1631=637%:%
%:%1632=638%:%
%:%1633=638%:%
%:%1634=638%:%
%:%1635=638%:%
%:%1636=638%:%
%:%1637=639%:%
%:%1638=639%:%
%:%1639=640%:%
%:%1640=640%:%
%:%1641=641%:%
%:%1642=641%:%
%:%1643=642%:%
%:%1644=642%:%
%:%1645=643%:%
%:%1646=643%:%
%:%1647=644%:%
%:%1648=644%:%
%:%1649=645%:%
%:%1650=645%:%
%:%1651=645%:%
%:%1652=646%:%
%:%1653=646%:%
%:%1654=647%:%
%:%1655=647%:%
%:%1656=648%:%
%:%1657=648%:%
%:%1658=648%:%
%:%1659=649%:%
%:%1660=649%:%
%:%1661=650%:%
%:%1662=650%:%
%:%1663=651%:%
%:%1664=651%:%
%:%1665=652%:%
%:%1666=652%:%
%:%1667=653%:%
%:%1668=653%:%
%:%1669=654%:%
%:%1670=654%:%
%:%1671=655%:%
%:%1672=655%:%
%:%1673=656%:%
%:%1674=656%:%
%:%1675=657%:%
%:%1676=657%:%
%:%1677=658%:%
%:%1678=658%:%
%:%1679=659%:%
%:%1680=659%:%
%:%1681=659%:%
%:%1682=660%:%
%:%1683=660%:%
%:%1684=661%:%
%:%1685=661%:%
%:%1686=661%:%
%:%1687=662%:%
%:%1688=662%:%
%:%1689=663%:%
%:%1690=663%:%
%:%1691=663%:%
%:%1692=663%:%
%:%1693=663%:%
%:%1694=664%:%
%:%1695=664%:%
%:%1696=665%:%
%:%1702=665%:%
%:%1705=666%:%
%:%1706=667%:%
%:%1707=667%:%
%:%1710=668%:%
%:%1714=668%:%
%:%1715=668%:%
%:%1721=668%:%
%:%1724=669%:%
%:%1725=670%:%
%:%1726=670%:%
%:%1727=671%:%
%:%1732=676%:%
%:%1735=677%:%
%:%1739=677%:%
%:%1740=677%:%
%:%1741=678%:%
%:%1742=678%:%
%:%1747=678%:%
%:%1752=679%:%
%:%1757=680%:%