### 2024-11-01 code/output/document/HS_Definition.tex

%
\begin{isabellebody}%
\setisabellecontext{HS{\isacharunderscore}{\kern0pt}Definition}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}Definition\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\ \isanewline
\ \ \ \ Automorphism{\isacharunderscore}{\kern0pt}Theorems\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ G\ {\isasymsubseteq}\ {\isacharbraceleft}{\kern0pt}\ {\isasympi}\ {\isasymin}\ P\ {\isasymrightarrow}\ P{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ \isanewline
\ \ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}{\isasympi}\ {\isasymin}\ G{\isachardot}{\kern0pt}\ {\isasymforall}{\isasymtau}\ {\isasymin}\ G{\isachardot}{\kern0pt}\ {\isasympi}\ O\ {\isasymtau}\ {\isasymin}\ G{\isacharparenright}{\kern0pt}\ \isanewline
\ \ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}{\isasympi}\ {\isasymin}\ G{\isachardot}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups\ \isakeyword{where}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ H\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\ {\isacharequal}{\kern0pt}\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}partial\ {\isacharplus}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymG}\ {\isasymF}\ \isanewline
\ \ \isakeyword{assumes}\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymG}\ {\isasymin}\ M{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{and}\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharparenleft}{\kern0pt}{\isasymG}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \ \isanewline
\ \ \isakeyword{and}\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymF}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{and}\ {\isasymF}{\isacharunderscore}{\kern0pt}subset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymF}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharparenleft}{\kern0pt}{\isasymG}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ {\isasymF}{\isacharunderscore}{\kern0pt}nonempty\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymF}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ {\isasymF}{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}intersection\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}A\ {\isasymin}\ {\isasymF}{\isachardot}{\kern0pt}\ {\isasymforall}B\ {\isasymin}\ {\isasymF}{\isachardot}{\kern0pt}\ A\ {\isasyminter}\ B\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ {\isasymF}{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}supergroup\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}A\ {\isasymin}\ {\isasymF}{\isachardot}{\kern0pt}\ {\isasymforall}B\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharparenleft}{\kern0pt}{\isasymG}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ A\ {\isasymsubseteq}\ B\ {\isasymlongrightarrow}\ B\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ {\isasymF}{\isacharunderscore}{\kern0pt}normal\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}H\ {\isasymin}\ {\isasymF}{\isachardot}{\kern0pt}\ {\isasymforall}{\isasympi}\ {\isasymin}\ {\isasymG}{\isachardot}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isasympi}\ O\ {\isasymtau}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymtau}\ {\isasymin}\ H\ {\isacharbraceright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ sym\ \isakeyword{where}\ {\isachardoublequoteopen}sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isasympi}\ {\isasymin}\ {\isasymG}{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ symmetric\ \isakeyword{where}\ {\isachardoublequoteopen}symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ HHS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isakeyword{where}\ {\isachardoublequoteopen}HHS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ X{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ X\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set\ \isakeyword{where}\ {\isachardoublequoteopen}HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ transrec{\isadigit{2}}{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ HHS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ HS\ \isakeyword{where}\ {\isachardoublequoteopen}HS\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ {\isasymexists}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}a{\isachardot}{\kern0pt}\ {\isasymforall}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ trans{\isacharunderscore}{\kern0pt}induct{\isadigit{3}}{\isacharunderscore}{\kern0pt}raw{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ blast\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ a\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}b{\isasymin}a{\isachardot}{\kern0pt}\ {\isasymforall}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}c\ {\isacharless}{\kern0pt}\ b{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}b{\isacharless}{\kern0pt}a{\isachardot}{\kern0pt}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ transrec{\isadigit{2}}{\isacharunderscore}{\kern0pt}Limit\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ c\ \isakeyword{where}\ cH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}c\ {\isacharless}{\kern0pt}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ ltD\ \isacommand{by}\isamarkupfalse%
\ blast\ \ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}c\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}trans\ bH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}c\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ cH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS\ {\isasymLongrightarrow}\ {\isasymexists}b{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharprime}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}transrec{\isadigit{2}}{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ HHS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}HHS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ transrec{\isadigit{2}}{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ HHS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ transrec{\isadigit{2}}{\isacharunderscore}{\kern0pt}succ{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ HHS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}subset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ trans{\isacharunderscore}{\kern0pt}induct{\isadigit{3}}{\isacharunderscore}{\kern0pt}raw{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def\ HHS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ blast\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ a\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}b{\isasymin}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ transrec{\isadigit{2}}{\isacharunderscore}{\kern0pt}Limit\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ Limit{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ lt{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ ltD\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}lim\ assms\ bH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ HS\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ HS\ {\isasymlongleftrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ domain{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}\isanewline
\ \ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ aH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ HS{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}HS{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ aH\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ p\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ HS\ {\isasymlongleftrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ domain{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ I\ \isakeyword{where}\ {\isachardoublequoteopen}I\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}{\isasymUnion}I{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Ord{\isacharunderscore}{\kern0pt}Union{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ L\ \isakeyword{where}\ LH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymUnion}I{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ L{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ex{\isacharunderscore}{\kern0pt}larger{\isacharunderscore}{\kern0pt}limit\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}a\ {\isacharless}{\kern0pt}\ L{\isachardot}{\kern0pt}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ transrec{\isadigit{2}}{\isacharunderscore}{\kern0pt}Limit\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ p\ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ LeastI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ Union{\isacharunderscore}{\kern0pt}upper{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assm\ le{\isacharunderscore}{\kern0pt}refl\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}I{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ I{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ L{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymUnion}I{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ LH\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}a\ {\isacharless}{\kern0pt}\ L{\isachardot}{\kern0pt}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ OUN{\isacharunderscore}{\kern0pt}I{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}\ {\isasymand}\ symmetric{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ pin{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ ypH\ vin\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}range\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ ypH\ vin\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}subset\ LH\ Limit{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pin\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ LH\ Limit{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}L{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ exI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ LH\ Limit{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}subgroup{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}{\isasymG}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymG}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharparenleft}{\kern0pt}{\isasymG}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ local{\isachardot}{\kern0pt}{\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}{\isasymF}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymG}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymI}\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymI}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymI}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharparenleft}{\kern0pt}{\isasymG}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isacommand{using}\isamarkupfalse%
\ {\isasymF}{\isacharunderscore}{\kern0pt}nonempty\ {\isasymF}{\isacharunderscore}{\kern0pt}subset\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymI}\ {\isasymsubseteq}\ {\isasymG}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymG}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isasymF}{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}supergroup\ {\isasymG}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}subgroup{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}{\isasymG}\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/HS_Definition.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=10%:%
%:%30=11%:%
%:%33=14%:%
%:%34=15%:%
%:%35=16%:%
%:%36=16%:%
%:%37=17%:%
%:%38=18%:%
%:%39=18%:%
%:%40=19%:%
%:%41=20%:%
%:%42=20%:%
%:%43=21%:%
%:%44=22%:%
%:%45=23%:%
%:%46=24%:%
%:%47=25%:%
%:%48=26%:%
%:%49=27%:%
%:%50=28%:%
%:%51=29%:%
%:%52=30%:%
%:%53=31%:%
%:%54=32%:%
%:%55=32%:%
%:%56=33%:%
%:%57=34%:%
%:%58=34%:%
%:%59=35%:%
%:%60=36%:%
%:%61=36%:%
%:%62=37%:%
%:%63=38%:%
%:%64=38%:%
%:%65=39%:%
%:%66=39%:%
%:%67=40%:%
%:%68=41%:%
%:%69=41%:%
%:%72=42%:%
%:%76=42%:%
%:%77=42%:%
%:%78=43%:%
%:%79=43%:%
%:%80=44%:%
%:%81=44%:%
%:%82=45%:%
%:%83=45%:%
%:%84=46%:%
%:%85=46%:%
%:%86=47%:%
%:%87=47%:%
%:%88=48%:%
%:%89=48%:%
%:%90=49%:%
%:%91=49%:%
%:%92=50%:%
%:%93=50%:%
%:%94=50%:%
%:%95=51%:%
%:%96=51%:%
%:%97=52%:%
%:%98=52%:%
%:%99=52%:%
%:%100=52%:%
%:%101=53%:%
%:%102=53%:%
%:%103=53%:%
%:%104=53%:%
%:%105=53%:%
%:%106=54%:%
%:%107=54%:%
%:%108=54%:%
%:%109=54%:%
%:%110=54%:%
%:%111=55%:%
%:%112=55%:%
%:%113=55%:%
%:%114=55%:%
%:%115=55%:%
%:%116=56%:%
%:%117=56%:%
%:%118=56%:%
%:%119=56%:%
%:%120=56%:%
%:%121=57%:%
%:%127=57%:%
%:%130=58%:%
%:%131=59%:%
%:%132=59%:%
%:%139=60%:%
%:%140=60%:%
%:%141=61%:%
%:%142=61%:%
%:%143=62%:%
%:%144=62%:%
%:%145=62%:%
%:%146=62%:%
%:%147=62%:%
%:%148=63%:%
%:%149=63%:%
%:%150=63%:%
%:%151=63%:%
%:%152=63%:%
%:%153=64%:%
%:%159=64%:%
%:%162=65%:%
%:%163=66%:%
%:%164=66%:%
%:%167=67%:%
%:%171=67%:%
%:%172=67%:%
%:%173=68%:%
%:%174=68%:%
%:%175=69%:%
%:%176=69%:%
%:%177=70%:%
%:%178=70%:%
%:%179=71%:%
%:%180=71%:%
%:%185=71%:%
%:%188=72%:%
%:%189=73%:%
%:%190=73%:%
%:%193=74%:%
%:%197=74%:%
%:%198=74%:%
%:%199=75%:%
%:%200=75%:%
%:%201=76%:%
%:%202=76%:%
%:%203=77%:%
%:%204=77%:%
%:%205=78%:%
%:%206=78%:%
%:%207=79%:%
%:%208=79%:%
%:%209=80%:%
%:%210=80%:%
%:%211=81%:%
%:%212=81%:%
%:%213=82%:%
%:%214=82%:%
%:%215=83%:%
%:%216=83%:%
%:%217=83%:%
%:%218=84%:%
%:%219=84%:%
%:%220=84%:%
%:%221=84%:%
%:%222=84%:%
%:%223=84%:%
%:%224=85%:%
%:%225=85%:%
%:%226=85%:%
%:%227=85%:%
%:%228=85%:%
%:%229=86%:%
%:%230=86%:%
%:%231=86%:%
%:%232=86%:%
%:%233=86%:%
%:%234=87%:%
%:%235=87%:%
%:%236=87%:%
%:%237=87%:%
%:%238=87%:%
%:%239=88%:%
%:%245=88%:%
%:%248=89%:%
%:%249=90%:%
%:%250=90%:%
%:%253=91%:%
%:%257=91%:%
%:%258=91%:%
%:%259=92%:%
%:%260=92%:%
%:%261=93%:%
%:%262=93%:%
%:%263=94%:%
%:%264=94%:%
%:%265=94%:%
%:%266=95%:%
%:%267=96%:%
%:%268=96%:%
%:%269=96%:%
%:%270=96%:%
%:%271=97%:%
%:%272=97%:%
%:%273=97%:%
%:%274=97%:%
%:%275=97%:%
%:%276=98%:%
%:%277=98%:%
%:%278=98%:%
%:%279=99%:%
%:%280=99%:%
%:%281=100%:%
%:%282=100%:%
%:%283=101%:%
%:%284=101%:%
%:%285=102%:%
%:%286=102%:%
%:%287=103%:%
%:%288=103%:%
%:%289=104%:%
%:%290=104%:%
%:%291=105%:%
%:%292=105%:%
%:%293=106%:%
%:%294=106%:%
%:%295=106%:%
%:%296=106%:%
%:%297=106%:%
%:%298=106%:%
%:%299=106%:%
%:%300=106%:%
%:%301=107%:%
%:%302=107%:%
%:%303=108%:%
%:%304=108%:%
%:%305=109%:%
%:%306=109%:%
%:%307=109%:%
%:%308=110%:%
%:%309=111%:%
%:%310=112%:%
%:%311=112%:%
%:%312=113%:%
%:%313=113%:%
%:%314=113%:%
%:%315=114%:%
%:%316=114%:%
%:%317=115%:%
%:%318=115%:%
%:%319=116%:%
%:%320=116%:%
%:%321=116%:%
%:%322=117%:%
%:%323=117%:%
%:%324=117%:%
%:%325=118%:%
%:%326=118%:%
%:%327=118%:%
%:%328=118%:%
%:%329=118%:%
%:%330=118%:%
%:%331=119%:%
%:%332=119%:%
%:%333=119%:%
%:%334=120%:%
%:%335=120%:%
%:%336=121%:%
%:%337=121%:%
%:%338=121%:%
%:%339=122%:%
%:%340=122%:%
%:%341=122%:%
%:%342=122%:%
%:%343=122%:%
%:%344=123%:%
%:%345=123%:%
%:%346=123%:%
%:%347=123%:%
%:%348=123%:%
%:%349=124%:%
%:%350=124%:%
%:%351=124%:%
%:%352=125%:%
%:%353=125%:%
%:%354=126%:%
%:%355=126%:%
%:%356=127%:%
%:%357=128%:%
%:%358=128%:%
%:%359=129%:%
%:%360=129%:%
%:%361=130%:%
%:%362=130%:%
%:%363=131%:%
%:%364=131%:%
%:%365=132%:%
%:%366=132%:%
%:%367=133%:%
%:%368=133%:%
%:%369=133%:%
%:%370=133%:%
%:%371=133%:%
%:%372=134%:%
%:%373=134%:%
%:%374=134%:%
%:%375=135%:%
%:%376=135%:%
%:%377=136%:%
%:%378=136%:%
%:%379=137%:%
%:%380=137%:%
%:%381=138%:%
%:%382=138%:%
%:%383=138%:%
%:%384=139%:%
%:%385=139%:%
%:%386=140%:%
%:%387=140%:%
%:%388=141%:%
%:%389=141%:%
%:%390=142%:%
%:%391=142%:%
%:%392=143%:%
%:%393=143%:%
%:%394=143%:%
%:%395=144%:%
%:%396=144%:%
%:%397=145%:%
%:%398=145%:%
%:%399=146%:%
%:%400=146%:%
%:%401=147%:%
%:%402=147%:%
%:%403=148%:%
%:%404=148%:%
%:%405=149%:%
%:%406=149%:%
%:%407=150%:%
%:%408=150%:%
%:%409=150%:%
%:%410=151%:%
%:%411=151%:%
%:%412=151%:%
%:%413=151%:%
%:%414=151%:%
%:%415=151%:%
%:%416=152%:%
%:%417=152%:%
%:%418=152%:%
%:%419=152%:%
%:%420=152%:%
%:%421=153%:%
%:%422=153%:%
%:%423=153%:%
%:%424=153%:%
%:%425=154%:%
%:%426=154%:%
%:%427=154%:%
%:%428=154%:%
%:%429=154%:%
%:%430=155%:%
%:%431=155%:%
%:%432=156%:%
%:%433=156%:%
%:%434=156%:%
%:%435=157%:%
%:%436=157%:%
%:%437=157%:%
%:%438=158%:%
%:%439=158%:%
%:%440=158%:%
%:%441=159%:%
%:%442=159%:%
%:%443=160%:%
%:%444=160%:%
%:%445=161%:%
%:%446=161%:%
%:%447=162%:%
%:%448=162%:%
%:%449=163%:%
%:%450=163%:%
%:%451=164%:%
%:%452=164%:%
%:%453=165%:%
%:%459=165%:%
%:%462=166%:%
%:%463=167%:%
%:%464=167%:%
%:%467=168%:%
%:%471=168%:%
%:%472=168%:%
%:%473=169%:%
%:%474=169%:%
%:%475=170%:%
%:%476=170%:%
%:%481=170%:%
%:%484=171%:%
%:%485=172%:%
%:%486=172%:%
%:%493=173%:%
%:%494=173%:%
%:%495=174%:%
%:%496=174%:%
%:%497=174%:%
%:%498=174%:%
%:%499=175%:%
%:%500=175%:%
%:%501=175%:%
%:%502=175%:%
%:%503=175%:%
%:%504=176%:%
%:%505=176%:%
%:%506=176%:%
%:%507=176%:%
%:%508=176%:%
%:%509=177%:%
%:%515=177%:%
%:%518=178%:%
%:%519=179%:%
%:%520=179%:%
%:%527=180%:%