### 2024-11-01 code/output/document/HS_Forces.tex

%
\begin{isabellebody}%
\setisabellecontext{HS{\isacharunderscore}{\kern0pt}Forces}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}Forces\ \isanewline
\ \ \isakeyword{imports}\ SymExt{\isacharunderscore}{\kern0pt}Definition\ Delta{\isadigit{0}}\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\ \ \isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}k{\isacharcomma}{\kern0pt}\ k{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ k\ {\isasymin}\ n{\isacharcomma}{\kern0pt}\ {\isadigit{6}}\ {\isasymle}\ k\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{consts}\isamarkupfalse%
\ forcesHS{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\isanewline
\isacommand{primrec}\isamarkupfalse%
\isanewline
\ \ {\isachardoublequoteopen}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}forces{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{4}}{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \ {\isacharequal}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}forces{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{4}}{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \ \ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ Neg{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}leq{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \ \ {\isacharequal}{\kern0pt}\ Forall{\isacharparenleft}{\kern0pt}Implies{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ ren{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isasymunion}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isasymunion}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ forcesHS\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\ \isakeyword{begin}\ \isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ ForcesHS\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}{\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isasymtturnstile}HS\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isadigit{6}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isadigit{6}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\ {\isadigit{6}}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env\ \ \ {\isasymequiv}\ \ \ M{\isacharcomma}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isasymunion}n{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isasymunion}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isadigit{6}}\ {\isasymle}\ k{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjI{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isasymnotin}\ {\isadigit{6}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ {\isadigit{6}}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}lt\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ {\isadigit{6}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}lt\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ ltI\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjE{\isacharcomma}{\kern0pt}\ assumption{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}lt\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ ltI\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}n\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\ n\ a\ b\ c\ d\ e\ f\ env\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}\ {\isasymunion}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}c\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}e\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isacharcomma}{\kern0pt}\ c{\isacharcomma}{\kern0pt}\ d{\isacharcomma}{\kern0pt}\ e{\isacharcomma}{\kern0pt}\ f{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isasymunion}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isasymunion}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ a{\isacharcomma}{\kern0pt}\ b{\isacharcomma}{\kern0pt}\ c{\isacharcomma}{\kern0pt}\ d{\isacharcomma}{\kern0pt}\ e{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ren{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isasymin}\ {\isadigit{6}}\ {\isasymor}\ k\ {\isasymin}\ n{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isasymin}\ {\isadigit{6}}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ {\isadigit{6}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymor}\ k\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymor}\ k\ {\isacharequal}{\kern0pt}\ {\isadigit{2}}\ {\isasymor}\ k\ {\isacharequal}{\kern0pt}\ {\isadigit{3}}\ {\isasymor}\ k\ {\isacharequal}{\kern0pt}\ {\isadigit{4}}\ {\isasymor}\ k\ {\isacharequal}{\kern0pt}\ {\isadigit{5}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjE{\isacharcomma}{\kern0pt}\ assumption{\isacharcomma}{\kern0pt}\ subst\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ assumption{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjE{\isacharcomma}{\kern0pt}\ assumption{\isacharcomma}{\kern0pt}\ subst\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ rule\ UnI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ assumption{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ rule\ UnI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ assumption{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}type\ assms\ Pi{\isacharunderscore}{\kern0pt}def\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isadigit{5}}\ {\isacharless}{\kern0pt}\ k{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ rule\ UnI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}type\ assms\ Pi{\isacharunderscore}{\kern0pt}def\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ nth{\isacharunderscore}{\kern0pt}append{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ nth{\isacharunderscore}{\kern0pt}append{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ if{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}P{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ not{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ assms\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ if{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}P{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ not{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ assms\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isasymle}\ {\isadigit{5}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ ltD\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isasymin}\ {\isadigit{6}}\ {\isasymunion}\ n{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}clarify{\isacharcomma}{\kern0pt}\ rule\ Implies{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ren{\isacharunderscore}{\kern0pt}tc{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{5}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ forces{\isacharunderscore}{\kern0pt}mem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ if{\isacharunderscore}{\kern0pt}P{\isacharcomma}{\kern0pt}\ subst\ arity{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}mem{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}mem{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ forces{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ if{\isacharunderscore}{\kern0pt}P{\isacharcomma}{\kern0pt}\ subst\ arity{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ arity{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp\ {\isacharcomma}{\kern0pt}rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp\ {\isacharcomma}{\kern0pt}rule\ disjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ p{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymand}\ ren{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{6}}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{6}}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}ren{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ p{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isadigit{5}}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ren{\isacharunderscore}{\kern0pt}tc{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ arity{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}forcesHS\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ {\isasymphi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{5}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ assms\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Member\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forcesHS{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forces{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}P\ {\isasymin}\ M\ {\isasymand}\ leq\ {\isasymin}\ M\ {\isasymand}\ one\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}def\ forces{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}forces{\isacharunderscore}{\kern0pt}mem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Member\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ y\ env\ p\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}p\ {\isasymtturnstile}\ Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ env\ {\isasymlongleftrightarrow}\ p\ {\isasymtturnstile}HS\ Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Member{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Equal\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forcesHS{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forces{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}P\ {\isasymin}\ M\ {\isasymand}\ leq\ {\isasymin}\ M\ {\isasymand}\ one\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}def\ forces{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}forces{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Equal\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ y\ env\ p\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}p\ {\isasymtturnstile}\ Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ env\ {\isasymlongleftrightarrow}\ p\ {\isasymtturnstile}HS\ Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Equal{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Nand\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forcesHS{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}p{\isasymin}P\ {\isasymand}\ {\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}q{\isasymin}M{\isachardot}{\kern0pt}\ q{\isasymin}P\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}leq{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}q{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymand}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}fm\ assms\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}disjI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}disjI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}disjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ notnot{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ notnot{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}nand{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Nand{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ {\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}q{\isasymin}M{\isachardot}{\kern0pt}\ q{\isasymin}P\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}leq{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}q{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}P\ {\isasymin}\ M\ {\isasymand}\ leq\ {\isasymin}\ M\ {\isasymand}\ one\ {\isasymin}\ M\ {\isasymand}\ {\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Nand{\isacharbrackleft}{\kern0pt}OF\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharminus}{\kern0pt}{\isadigit{4}}{\isacharparenright}{\kern0pt}\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Nand{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}q{\isasymin}M{\isachardot}{\kern0pt}\ q{\isasymin}P\ {\isasymand}\ q{\isasympreceq}p\ {\isasymand}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Nand{\isacharprime}{\kern0pt}\ transitivity\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}abs\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Neg{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}q{\isasymin}M{\isachardot}{\kern0pt}\ q{\isasymin}P\ {\isasymand}\ q{\isasympreceq}p\ {\isasymand}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Neg{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Nand{\isacharprime}{\kern0pt}\ transitivity\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}abs\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Forall\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forcesHS{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ p{\isasymin}P\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x{\isasymin}HS{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ rule\ ballI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharcomma}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ sats{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ arity{\isacharunderscore}{\kern0pt}type\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{9}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}leq{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ ren{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{6}}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{6}}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}forcesHS{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ sats{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}forall{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ forcesHS{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ arity{\isacharunderscore}{\kern0pt}type\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{9}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}leq{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Forall{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x{\isasymin}HS{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharcomma}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Forall{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}P\ {\isasymin}\ M\ {\isasymand}\ leq\ {\isasymin}\ M\ {\isasymand}\ one\ {\isasymin}\ M\ {\isasymand}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}\ {\isasymin}\ M\ {\isasymand}\ p\ {\isasymin}\ M\ {\isasymand}\ HS\ {\isasymsubseteq}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ transM\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Forall{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x{\isasymin}HS{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Forall{\isacharprime}{\kern0pt}\ assms\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasympreceq}p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}env{\isachardot}{\kern0pt}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p\ {\isasymtturnstile}HS\ {\isasymphi}\ env\ {\isasymLongrightarrow}\ r\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Member\ n\ m{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Member{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}r\ {\isasymtturnstile}HS\ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Member{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ assms\ assms{\isadigit{1}}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Equal\ n\ m{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Equal{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}r\ {\isasymtturnstile}HS\ {\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Equal{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ assms\ assms{\isadigit{1}}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Nand\ {\isasymphi}\ {\isasympsi}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Nand\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ leq{\isacharunderscore}{\kern0pt}transD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Forall\ {\isasymphi}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}HS{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ ForcesHS{\isacharunderscore}{\kern0pt}Forall\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ Forall\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}HS{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ pred{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\ Forall\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Forall\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}separation\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}p{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}P\ {\isasymLongrightarrow}\ z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transitivity{\isacharbrackleft}{\kern0pt}of\ z\ P{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}p{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ forcesHS{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}forcesHS{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}ForcesHS\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ fty{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ far{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ envty{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}p{\isasymin}P\ {\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}P\ {\isasymLongrightarrow}\ z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transitivity{\isacharbrackleft}{\kern0pt}of\ z\ P{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}separation{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}closed\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}aux{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env\ \ \ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}q{\isasymin}M{\isachardot}{\kern0pt}\ q{\isasymin}P\ {\isasymand}\ q{\isasympreceq}p\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}r{\isasymin}M{\isachardot}{\kern0pt}\ r{\isasymin}P\ {\isasymand}\ r{\isasympreceq}q\ {\isasymand}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ And{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ ForcesHS{\isacharunderscore}{\kern0pt}Neg\ ForcesHS{\isacharunderscore}{\kern0pt}Nand\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ only{\isacharcolon}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}dense{\isacharunderscore}{\kern0pt}below{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}r{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}aux\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ dest{\isacharcolon}{\kern0pt}transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ rename{\isacharunderscore}{\kern0pt}tac\ q{\isacharsemicolon}{\kern0pt}\ drule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}q\ \isakeyword{in}\ bspec{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}forcesHS{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}env{\isachardot}{\kern0pt}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Member\ n\ m{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iff{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Member{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transM\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Member\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}\ Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}forces{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Member\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ Member{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Member{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Member\ transM\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Equal\ n\ m{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iff{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Equal{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transM\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Equal\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}\ Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}forces{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Equal\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ Equal{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Equal{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Equal\ transM\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Nand\ {\isasymphi}\ {\isasympsi}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\ \ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ q\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}q{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ Nand\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ d\ \isakeyword{where}\ {\isachardoublequoteopen}d{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ {\isasymtturnstile}HS\ Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ {\isachardoublequoteopen}d{\isasympreceq}\ q{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}belowI\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}{\isacharparenleft}{\kern0pt}d{\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}d\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ ForcesHS{\isacharunderscore}{\kern0pt}Nand\ leq{\isacharunderscore}{\kern0pt}reflI\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ of\ d{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}Nand{\isacharunderscore}{\kern0pt}le{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isasympsi}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharbrackleft}{\kern0pt}of\ q\ {\isasymphi}\ d\ env{\isacharbrackright}{\kern0pt}\ Un{\isacharunderscore}{\kern0pt}leD{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharbrackleft}{\kern0pt}of\ q\ {\isasympsi}\ d\ env{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Nand{\isacharbrackleft}{\kern0pt}symmetric{\isacharcomma}{\kern0pt}\ OF\ {\isacharunderscore}{\kern0pt}\ Nand{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Forall\ {\isasymphi}{\isacharparenright}{\kern0pt}\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}env{\isachardot}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ local{\isachardot}{\kern0pt}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}leq{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}a{\isasymin}HS{\isachardoublequoteclose}\ \isakeyword{for}\ a\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ r\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}r{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasympreceq}p{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}r{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymtturnstile}HS\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ Forall\ {\isacartoucheopen}a{\isasymin}HS{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Forall\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}d\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ d\ {\isasymin}\ P\ {\isasymand}\ d{\isasympreceq}r{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ Forall{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ Forall{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharminus}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}a{\isasymin}HS{\isachardoublequoteclose}\ \isakeyword{for}\ a\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ H{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pred{\isacharunderscore}{\kern0pt}le\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pred{\isacharunderscore}{\kern0pt}le\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ pred{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\ Forall\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Forall\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}density{\isacharunderscore}{\kern0pt}lemma{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env\ \ \ {\isasymlongleftrightarrow}\ \ \ dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}forcesHS\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma\ leq{\isacharunderscore}{\kern0pt}reflI\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}And{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env\ \ \ {\isasymlongleftrightarrow}\ \ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}r\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}dense{\isacharunderscore}{\kern0pt}below\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}r\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}r\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}density{\isacharunderscore}{\kern0pt}lemma{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}r\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ dense{\isacharunderscore}{\kern0pt}belowI\ bexI\ conjI{\isacharcomma}{\kern0pt}\ assumption{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ q\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}\ p{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}q{\isasymin}{\isacharbraceleft}{\kern0pt}r\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}r\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}\ q{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma\ leq{\isacharunderscore}{\kern0pt}reflI\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}dense{\isacharunderscore}{\kern0pt}below\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Nand{\isacharunderscore}{\kern0pt}alt{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Nand{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ ForcesHS{\isacharunderscore}{\kern0pt}Nand\ ForcesHS{\isacharunderscore}{\kern0pt}And\ ForcesHS{\isacharunderscore}{\kern0pt}Neg\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}Neg{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ IH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ \ {\isasymlongleftrightarrow}\ \ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ iffI{\isacharcomma}{\kern0pt}\ elim\ bexE{\isacharcomma}{\kern0pt}\ rule\ ccontr{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ p\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymnot}{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ map{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ assms\ envin\ mapin\ assms{\isadigit{1}}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ IH\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ r\ \isakeyword{where}\ {\isachardoublequoteopen}r\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}G{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ {\isachardoublequoteopen}q{\isasympreceq}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}r{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}G{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharbrackleft}{\kern0pt}\isakeyword{where}\ {\isasymphi}{\isacharequal}{\kern0pt}{\isasymphi}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}False{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Neg{\isacharbrackleft}{\kern0pt}\isakeyword{where}\ {\isasymphi}{\isacharequal}{\kern0pt}{\isasymphi}{\isacharbrackright}{\kern0pt}\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ map{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ {\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapin\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}D{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}separation{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ envin\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}separation{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ envin\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}disj\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}D\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}closed\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}D\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ q\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}d{\isasymin}{\isacharbraceleft}{\kern0pt}p\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ d{\isasympreceq}\ q{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}q\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}P{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}reflI\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}P{\isacartoucheclose}\ \isakeyword{and}\ assms\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}Neg\ envin\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Neg{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymnot}\ {\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ IH\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{2}}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}And{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\isanewline
\ \ \ \ IH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ \ {\isasymlongleftrightarrow}\ \ \ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}\ \ {\isasymlongleftrightarrow}\ \ \ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasympsi}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ iffI{\isacharcomma}{\kern0pt}\ elim\ bexE{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ map{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ p\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesHS{\isacharunderscore}{\kern0pt}And{\isacharbrackleft}{\kern0pt}OF\ M{\isacharunderscore}{\kern0pt}genericD{\isacharcomma}{\kern0pt}\ of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isasymphi}\ {\isasympsi}{\isacharbrackright}{\kern0pt}\ mapin\ envin\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ map{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ q\ r\ \isakeyword{where}\ {\isachardoublequoteopen}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}G{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapin\ envin\ ForcesHS{\isacharunderscore}{\kern0pt}And\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}p{\isasympreceq}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasympreceq}r{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasympsi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma\ envin\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ And{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}p\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ ForcesHS{\isacharunderscore}{\kern0pt}And{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ M{\isacharunderscore}{\kern0pt}genericD\ envin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{9}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}{\isadigit{7}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ iterates{\isacharparenleft}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{6}}\ {\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}b{\isacharcomma}{\kern0pt}d{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymphi}\ {\isasymin}\ formula\ {\isasymLongrightarrow}\isanewline
\ \ \ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}b{\isacharcomma}{\kern0pt}d{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharcomma}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}iff\ {\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ M\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharcomma}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}le\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}m\ {\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}m\ {\isasymin}\ nat{\isacartoucheclose}\isanewline
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{0}}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}succ\ x{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iterates{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}n\ {\isacharless}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}x\ {\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}x\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isadigit{7}}\ {\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isadigit{7}}\ {\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iterates{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}d{\isachardot}{\kern0pt}\ {\isasymexists}b{\isasymin}HS{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}d\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}lemma{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q\ R{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ x\ {\isasymin}\ HS\ {\isasymlongleftrightarrow}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ HS\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ R{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ HS{\isachardot}{\kern0pt}\ R{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}rel{\isacharunderscore}{\kern0pt}pred{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}M\ x\ a{\isadigit{1}}\ a{\isadigit{2}}\ a{\isadigit{3}}\ a{\isadigit{4}}{\isachardot}{\kern0pt}\ {\isasymexists}b{\isasymin}HS{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}M{\isachardot}{\kern0pt}\ q{\isasymin}a{\isadigit{1}}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}leq{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}q{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymnot}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharcomma}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ Forall{\isacharparenleft}{\kern0pt}Implies{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}leq{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ Neg{\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenleft}{\kern0pt}forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}q{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ \isakeyword{for}\ q\ \isacommand{using}\isamarkupfalse%
\ that\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}q{\isasymin}M{\isachardot}{\kern0pt}\ q{\isasymin}P\ {\isasymand}\ R{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ R{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ R\ Q\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlbrakk}b\ {\isasymin}\ HS{\isacharsemicolon}{\kern0pt}\ {\isasymforall}q{\isasymin}M{\isachardot}{\kern0pt}\ q\ {\isasymin}\ P\ {\isasymand}\ q\ {\isasympreceq}\ d\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymrbrakk}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ {\isasymexists}c{\isasymin}HS{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q\ {\isasympreceq}\ d\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ b\ d\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ bexI{\isacharcomma}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}rel{\isacharunderscore}{\kern0pt}pred{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}d{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b{\isasymin}HS{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}d\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}d{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ d\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ leq{\isacharunderscore}{\kern0pt}abs\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymand}\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenleft}{\kern0pt}forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Exists{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iterates{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ forcesHS{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}d{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharquery}{\kern0pt}rel{\isacharunderscore}{\kern0pt}pred{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}d{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}d{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ d\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ that\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ sats{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}fm\ \ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ imp{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ notnot{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x\ y{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ d{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenleft}{\kern0pt}forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ forcesHS{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}d{\isachardot}{\kern0pt}\ {\isasymexists}b{\isasymin}HS{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}d\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}d{\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}rel{\isacharunderscore}{\kern0pt}pred{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}d{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}d{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}d{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ H{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}d{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}d{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}d{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}d{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Exists{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ forcesHS{\isacharunderscore}{\kern0pt}type\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}type\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp\ {\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}type\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ forcesHS{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}forcesHS{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}d{\isachardot}{\kern0pt}\ {\isasymexists}b{\isasymin}HS{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q\ {\isasympreceq}\ d\ {\isasymlongrightarrow}\ satisfies{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ I{\isadigit{1}}\ I{\isadigit{2}}\ I{\isadigit{3}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}env{\isachardot}{\kern0pt}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Member\ x\ y{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ GsubsetM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}G\ {\isasymsubseteq}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ subset{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ filter{\isacharunderscore}{\kern0pt}def\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envmaptype\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ map{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ enum{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}axioms\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}axioms\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Member{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ envin\ GsubsetM\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p\ {\isasymin}\ G{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}\ Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms\ envin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{4}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isachardot}{\kern0pt}SymExt{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}GenExt\ envmaptype\ Member{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ assms{\isadigit{1}}\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isachardot}{\kern0pt}Transset{\isacharunderscore}{\kern0pt}SymExt\ length{\isacharunderscore}{\kern0pt}map\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ I{\isadigit{1}}\ I{\isadigit{2}}\ I{\isadigit{3}}\ I{\isadigit{4}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Equal\ x\ y{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ enum{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}axioms\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}axioms\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ GsubsetM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}G\ {\isasymsubseteq}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ B{\isacharequal}{\kern0pt}P\ \isakeyword{in}\ subset{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ filter{\isacharunderscore}{\kern0pt}def\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envmaptype\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ map{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}forcesHS{\isacharunderscore}{\kern0pt}Equal{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ envin\ GsubsetM\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p\ {\isasymin}\ G{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}\ Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms\ envin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ I{\isadigit{4}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ {\isasymDelta}{\isadigit{0}}{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isachardot}{\kern0pt}SymExt{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}GenExt\ envmaptype\ Member{\isacharunderscore}{\kern0pt}{\isasymDelta}{\isadigit{0}}\ assms{\isadigit{1}}\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isachardot}{\kern0pt}Transset{\isacharunderscore}{\kern0pt}SymExt\ length{\isacharunderscore}{\kern0pt}map\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ I{\isadigit{1}}\ I{\isadigit{2}}\ I{\isadigit{3}}\ I{\isadigit{4}}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Nand\ {\isasymphi}\ {\isasympsi}{\isacharparenright}{\kern0pt}\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ Nand\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ map{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ SymExt{\isacharunderscore}{\kern0pt}def\ Nand\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ arityle\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymand}\ arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Nand\ max{\isacharunderscore}{\kern0pt}le{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ arity{\isacharparenleft}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Nand\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ ForcesHS{\isacharunderscore}{\kern0pt}Nand{\isacharunderscore}{\kern0pt}alt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}genericD\ assms\ envin\ Nand\ arityle\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}Neg{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Nand\ assms\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharunderscore}{\kern0pt}And{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Nand\ arityle\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{8}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapin\ Nand\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Forall\ {\isasymphi}{\isacharparenright}{\kern0pt}\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ Forall\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ map{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ SymExt{\isacharunderscore}{\kern0pt}def\ Forall\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ ihE\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}env{\isachardot}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}{\isasymlambda}a{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Forall\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ pH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}HS{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ ForcesHS{\isacharunderscore}{\kern0pt}Forall\ envin\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}{\isasymlambda}a{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ ForcesHS{\isacharunderscore}{\kern0pt}Forall\ envin\ Forall\ pH\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ ihE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}p\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapin\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymexists}x{\isacharprime}{\kern0pt}\ {\isasymin}\ HS{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ x{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ satsforall\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}D{\isadigit{1}}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}d{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}d\ {\isasymtturnstile}HS\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}D{\isadigit{2}}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}d{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymexists}b{\isasymin}HS{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}d\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{define}\isamarkupfalse%
\ D\ \isakeyword{where}\ {\isachardoublequoteopen}D\ {\isasymequiv}\ {\isacharquery}{\kern0pt}D{\isadigit{1}}\ {\isasymunion}\ {\isacharquery}{\kern0pt}D{\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ ar{\isasymphi}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ pred{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}le\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}D{\isadigit{1}}{\isasymin}M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}ForcesHS\ ar{\isasymphi}\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}D{\isadigit{2}}{\isasymin}M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ \ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharprime}{\kern0pt}\ separation{\isacharunderscore}{\kern0pt}closed\ ar{\isasymphi}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}D{\isasymin}M{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ D{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}D\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ D{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}D{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ p\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}d{\isasymin}D{\isachardot}{\kern0pt}\ d{\isasympreceq}\ p{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ D{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}reflI\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}x{\isasymin}HS{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ForcesHS{\isacharunderscore}{\kern0pt}Forall{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}p\ {\isasymin}\ P{\isacartoucheclose}\ Forall\ envin\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x{\isasymin}HS{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ forcesHS{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ False\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ {\isachardoublequoteopen}b{\isasymin}HS{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymnot}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ Forall\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ notnot{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ HS{\isacharunderscore}{\kern0pt}density{\isacharunderscore}{\kern0pt}lemma{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ envin\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}density{\isacharunderscore}{\kern0pt}lemma\ pred{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ d\ \isakeyword{where}\ {\isachardoublequoteopen}d{\isasympreceq}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}d\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}d{\isasymin}P{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ D{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}d\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}b\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ d\ \isakeyword{where}\ {\isachardoublequoteopen}d\ {\isasymin}\ D{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}d\ {\isasymin}\ G{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}d{\isasymin}{\isacharquery}{\kern0pt}D{\isadigit{1}}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}d{\isasymin}{\isacharquery}{\kern0pt}D{\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ D{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ Forall{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}d{\isasymin}G{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{2}}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ {\isachardoublequoteopen}b{\isasymin}HS{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}d\ {\isasymlongrightarrow}{\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ satsforall\ mapin\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}b{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ b{\isacharparenright}{\kern0pt}\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}b\ {\isasymin}\ HS{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}p\ {\isasymin}\ G{\isachardot}{\kern0pt}\ p\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ihE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}b\ {\isasymin}\ HS{\isacartoucheclose}\ Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Forall\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}genericD\ assms\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}d{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ {\isachardoublequoteopen}q{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}d{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasympreceq}p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ \ {\isacartoucheopen}p\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ Forall\ \ {\isacartoucheopen}b{\isasymin}HS{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ p{\isacharequal}{\kern0pt}p\ \isakeyword{in}\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}b{\isasymin}HS{\isacartoucheclose}\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ envin\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymforall}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}d\ {\isasymlongrightarrow}{\isasymnot}{\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}b{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing{\isacharunderscore}{\kern0pt}HS{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ p{\isasymin}G\ \ {\isasymlongrightarrow}\ \ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ iffI\ allI\ impI{\isacharcomma}{\kern0pt}\ elim\ conjE{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ G\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ envin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ Forall\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}{\isacharparenleft}{\kern0pt}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isasymand}\ p{\isasymin}G{\isacharparenright}{\kern0pt}{\isasymlongrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ r\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}r{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasympreceq}p{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ G\ \isakeyword{where}\ {\isachardoublequoteopen}r{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}existence\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ {\isadigit{2}}\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}leqD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ satsphi\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}s\ {\isasymin}\ G{\isachardot}{\kern0pt}\ s\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ satsphi\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ s\ \isakeyword{where}\ {\isachardoublequoteopen}s{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}s\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ satsphi\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ \ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}r{\isasymin}G{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ {\isachardoublequoteopen}q{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasympreceq}\ s{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasympreceq}\ r{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ {\isacartoucheopen}s{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}s{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ q{\isasympreceq}r\ {\isasymand}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}q\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}q\ {\isasympreceq}\ r{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ p{\isacharequal}{\kern0pt}s\ \isakeyword{in}\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}genericD\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}s{\isasymin}G{\isacartoucheclose}\ envin\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharunderscore}{\kern0pt}below{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}below{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}HS\ {\isasymphi}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}density{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ envin\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/HS_Forces.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%18=3%:%
%:%21=4%:%
%:%22=5%:%
%:%23=5%:%
%:%24=6%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=9%:%
%:%29=10%:%
%:%30=10%:%
%:%31=11%:%
%:%32=12%:%
%:%33=13%:%
%:%35=15%:%
%:%36=16%:%
%:%37=17%:%
%:%38=18%:%
%:%39=18%:%
%:%40=19%:%
%:%41=20%:%
%:%42=21%:%
%:%43=22%:%
%:%44=22%:%
%:%45=23%:%
%:%46=24%:%
%:%47=24%:%
%:%48=25%:%
%:%49=26%:%
%:%50=27%:%
%:%51=27%:%
%:%52=28%:%
%:%53=29%:%
%:%54=30%:%
%:%57=31%:%
%:%58=32%:%
%:%62=32%:%
%:%63=32%:%
%:%64=33%:%
%:%65=33%:%
%:%66=34%:%
%:%67=34%:%
%:%68=35%:%
%:%69=35%:%
%:%70=36%:%
%:%71=36%:%
%:%72=37%:%
%:%73=37%:%
%:%74=38%:%
%:%75=38%:%
%:%76=39%:%
%:%77=39%:%
%:%78=40%:%
%:%79=40%:%
%:%80=41%:%
%:%81=41%:%
%:%82=42%:%
%:%83=42%:%
%:%84=43%:%
%:%85=43%:%
%:%86=44%:%
%:%87=44%:%
%:%88=45%:%
%:%89=45%:%
%:%90=46%:%
%:%91=46%:%
%:%92=47%:%
%:%93=47%:%
%:%94=48%:%
%:%95=48%:%
%:%96=49%:%
%:%97=49%:%
%:%98=50%:%
%:%99=50%:%
%:%100=51%:%
%:%101=51%:%
%:%102=52%:%
%:%103=52%:%
%:%104=53%:%
%:%105=53%:%
%:%106=54%:%
%:%107=54%:%
%:%108=55%:%
%:%109=55%:%
%:%110=56%:%
%:%116=56%:%
%:%119=57%:%
%:%120=58%:%
%:%121=58%:%
%:%122=59%:%
%:%123=60%:%
%:%124=61%:%
%:%127=62%:%
%:%128=63%:%
%:%132=63%:%
%:%133=63%:%
%:%134=64%:%
%:%135=64%:%
%:%136=65%:%
%:%137=65%:%
%:%138=66%:%
%:%139=66%:%
%:%140=67%:%
%:%141=67%:%
%:%142=68%:%
%:%143=68%:%
%:%144=69%:%
%:%145=69%:%
%:%146=70%:%
%:%147=70%:%
%:%148=71%:%
%:%149=71%:%
%:%150=72%:%
%:%151=72%:%
%:%152=73%:%
%:%153=73%:%
%:%154=74%:%
%:%155=74%:%
%:%156=75%:%
%:%157=75%:%
%:%158=76%:%
%:%159=76%:%
%:%160=77%:%
%:%161=77%:%
%:%162=78%:%
%:%163=78%:%
%:%164=79%:%
%:%165=79%:%
%:%166=80%:%
%:%167=80%:%
%:%168=81%:%
%:%169=81%:%
%:%170=82%:%
%:%171=82%:%
%:%172=83%:%
%:%173=83%:%
%:%174=84%:%
%:%175=84%:%
%:%176=85%:%
%:%177=85%:%
%:%178=86%:%
%:%179=86%:%
%:%180=87%:%
%:%181=87%:%
%:%182=88%:%
%:%183=88%:%
%:%184=89%:%
%:%185=89%:%
%:%186=90%:%
%:%187=90%:%
%:%188=91%:%
%:%189=91%:%
%:%190=92%:%
%:%191=92%:%
%:%192=93%:%
%:%193=93%:%
%:%194=94%:%
%:%195=94%:%
%:%196=95%:%
%:%197=95%:%
%:%198=96%:%
%:%199=96%:%
%:%200=97%:%
%:%201=97%:%
%:%202=98%:%
%:%203=98%:%
%:%204=99%:%
%:%205=99%:%
%:%206=100%:%
%:%207=100%:%
%:%208=101%:%
%:%209=101%:%
%:%210=102%:%
%:%211=102%:%
%:%212=103%:%
%:%213=103%:%
%:%214=104%:%
%:%215=104%:%
%:%216=105%:%
%:%217=105%:%
%:%218=106%:%
%:%219=106%:%
%:%220=107%:%
%:%221=107%:%
%:%222=108%:%
%:%223=108%:%
%:%224=109%:%
%:%225=109%:%
%:%226=110%:%
%:%232=110%:%
%:%235=111%:%
%:%236=112%:%
%:%237=112%:%
%:%238=113%:%
%:%239=114%:%
%:%240=115%:%
%:%243=116%:%
%:%247=116%:%
%:%248=116%:%
%:%249=117%:%
%:%250=117%:%
%:%251=118%:%
%:%252=118%:%
%:%253=119%:%
%:%254=119%:%
%:%255=120%:%
%:%256=120%:%
%:%257=121%:%
%:%258=121%:%
%:%259=122%:%
%:%260=122%:%
%:%261=123%:%
%:%262=123%:%
%:%263=124%:%
%:%264=124%:%
%:%265=125%:%
%:%266=125%:%
%:%267=126%:%
%:%268=126%:%
%:%269=127%:%
%:%270=127%:%
%:%275=127%:%
%:%278=128%:%
%:%279=129%:%
%:%280=129%:%
%:%281=130%:%
%:%282=131%:%
%:%283=132%:%
%:%286=133%:%
%:%290=133%:%
%:%291=133%:%
%:%292=134%:%
%:%293=134%:%
%:%294=135%:%
%:%295=135%:%
%:%296=136%:%
%:%297=136%:%
%:%302=136%:%
%:%305=137%:%
%:%306=138%:%
%:%307=138%:%
%:%308=139%:%
%:%309=140%:%
%:%310=141%:%
%:%313=142%:%
%:%317=142%:%
%:%318=142%:%
%:%319=143%:%
%:%320=143%:%
%:%321=144%:%
%:%322=144%:%
%:%323=145%:%
%:%324=145%:%
%:%325=146%:%
%:%326=146%:%
%:%327=147%:%
%:%328=147%:%
%:%329=148%:%
%:%330=148%:%
%:%331=149%:%
%:%332=149%:%
%:%333=150%:%
%:%334=150%:%
%:%335=151%:%
%:%336=151%:%
%:%337=152%:%
%:%338=152%:%
%:%339=153%:%
%:%340=153%:%
%:%341=154%:%
%:%342=154%:%
%:%343=155%:%
%:%344=155%:%
%:%345=156%:%
%:%346=156%:%
%:%347=157%:%
%:%348=157%:%
%:%349=158%:%
%:%350=158%:%
%:%351=159%:%
%:%352=159%:%
%:%353=160%:%
%:%354=160%:%
%:%355=161%:%
%:%356=161%:%
%:%357=162%:%
%:%358=162%:%
%:%359=163%:%
%:%360=163%:%
%:%361=164%:%
%:%362=164%:%
%:%363=165%:%
%:%364=165%:%
%:%365=166%:%
%:%366=166%:%
%:%367=167%:%
%:%368=167%:%
%:%369=168%:%
%:%370=168%:%
%:%371=169%:%
%:%372=169%:%
%:%373=170%:%
%:%374=170%:%
%:%375=171%:%
%:%376=171%:%
%:%377=172%:%
%:%378=172%:%
%:%379=173%:%
%:%380=173%:%
%:%381=174%:%
%:%382=174%:%
%:%383=175%:%
%:%384=175%:%
%:%385=176%:%
%:%386=176%:%
%:%387=177%:%
%:%388=177%:%
%:%389=178%:%
%:%390=178%:%
%:%391=179%:%
%:%392=179%:%
%:%393=180%:%
%:%394=180%:%
%:%395=181%:%
%:%396=181%:%
%:%397=182%:%
%:%398=182%:%
%:%399=183%:%
%:%400=183%:%
%:%401=184%:%
%:%402=184%:%
%:%403=185%:%
%:%404=185%:%
%:%405=186%:%
%:%406=186%:%
%:%407=187%:%
%:%408=187%:%
%:%409=188%:%
%:%410=188%:%
%:%411=189%:%
%:%412=189%:%
%:%413=190%:%
%:%414=190%:%
%:%415=191%:%
%:%416=191%:%
%:%417=192%:%
%:%418=192%:%
%:%419=193%:%
%:%420=193%:%
%:%421=194%:%
%:%422=194%:%
%:%423=195%:%
%:%424=195%:%
%:%425=196%:%
%:%426=196%:%
%:%427=197%:%
%:%428=197%:%
%:%429=198%:%
%:%430=198%:%
%:%431=199%:%
%:%432=199%:%
%:%433=200%:%
%:%434=200%:%
%:%435=201%:%
%:%436=201%:%
%:%437=202%:%
%:%438=202%:%
%:%439=203%:%
%:%440=203%:%
%:%441=204%:%
%:%442=204%:%
%:%443=205%:%
%:%444=205%:%
%:%445=206%:%
%:%446=206%:%
%:%447=207%:%
%:%448=207%:%
%:%449=208%:%
%:%450=208%:%
%:%451=209%:%
%:%452=209%:%
%:%453=210%:%
%:%459=210%:%
%:%462=211%:%
%:%463=212%:%
%:%464=212%:%
%:%465=213%:%
%:%466=214%:%
%:%467=215%:%
%:%470=216%:%
%:%474=216%:%
%:%475=216%:%
%:%476=217%:%
%:%477=217%:%
%:%478=218%:%
%:%479=218%:%
%:%480=219%:%
%:%481=219%:%
%:%482=220%:%
%:%483=220%:%
%:%484=221%:%
%:%485=221%:%
%:%486=222%:%
%:%487=222%:%
%:%492=222%:%
%:%495=223%:%
%:%496=224%:%
%:%497=224%:%
%:%498=225%:%
%:%499=226%:%
%:%500=227%:%
%:%503=228%:%
%:%507=228%:%
%:%508=228%:%
%:%509=229%:%
%:%510=229%:%
%:%511=230%:%
%:%512=230%:%
%:%513=231%:%
%:%514=231%:%
%:%515=232%:%
%:%516=232%:%
%:%517=233%:%
%:%518=233%:%
%:%519=234%:%
%:%520=235%:%
%:%521=235%:%
%:%522=236%:%
%:%523=236%:%
%:%524=237%:%
%:%525=237%:%
%:%526=238%:%
%:%527=238%:%
%:%528=239%:%
%:%529=239%:%
%:%534=239%:%
%:%537=240%:%
%:%538=241%:%
%:%539=241%:%
%:%540=242%:%
%:%541=243%:%
%:%542=244%:%
%:%545=245%:%
%:%549=245%:%
%:%550=245%:%
%:%551=246%:%
%:%552=246%:%
%:%553=247%:%
%:%554=247%:%
%:%555=248%:%
%:%556=248%:%
%:%557=249%:%
%:%558=249%:%
%:%563=249%:%
%:%566=250%:%
%:%567=251%:%
%:%568=251%:%
%:%569=252%:%
%:%570=253%:%
%:%571=254%:%
%:%574=255%:%
%:%578=255%:%
%:%579=255%:%
%:%580=256%:%
%:%581=256%:%
%:%582=257%:%
%:%583=257%:%
%:%584=258%:%
%:%585=258%:%
%:%586=259%:%
%:%587=259%:%
%:%588=260%:%
%:%589=260%:%
%:%590=261%:%
%:%591=262%:%
%:%592=262%:%
%:%593=263%:%
%:%594=263%:%
%:%595=264%:%
%:%596=264%:%
%:%597=265%:%
%:%598=265%:%
%:%599=266%:%
%:%600=266%:%
%:%605=266%:%
%:%608=267%:%
%:%609=268%:%
%:%610=268%:%
%:%611=269%:%
%:%612=270%:%
%:%613=271%:%
%:%616=272%:%
%:%620=272%:%
%:%621=272%:%
%:%622=273%:%
%:%623=273%:%
%:%624=274%:%
%:%625=274%:%
%:%626=275%:%
%:%627=275%:%
%:%628=276%:%
%:%629=276%:%
%:%634=276%:%
%:%637=277%:%
%:%638=278%:%
%:%639=278%:%
%:%640=279%:%
%:%641=280%:%
%:%643=282%:%
%:%646=283%:%
%:%650=283%:%
%:%651=283%:%
%:%652=283%:%
%:%653=284%:%
%:%654=284%:%
%:%655=285%:%
%:%656=285%:%
%:%657=286%:%
%:%658=286%:%
%:%659=287%:%
%:%660=287%:%
%:%661=288%:%
%:%662=288%:%
%:%663=289%:%
%:%664=290%:%
%:%665=290%:%
%:%666=291%:%
%:%667=291%:%
%:%668=292%:%
%:%669=292%:%
%:%670=293%:%
%:%671=293%:%
%:%672=294%:%
%:%673=295%:%
%:%674=295%:%
%:%675=296%:%
%:%676=296%:%
%:%677=297%:%
%:%683=297%:%
%:%686=298%:%
%:%687=299%:%
%:%688=299%:%
%:%689=300%:%
%:%690=301%:%
%:%691=302%:%
%:%692=303%:%
%:%695=306%:%
%:%698=307%:%
%:%702=307%:%
%:%703=307%:%
%:%704=308%:%
%:%705=308%:%
%:%706=309%:%
%:%707=309%:%
%:%708=310%:%
%:%709=310%:%
%:%710=311%:%
%:%711=311%:%
%:%712=312%:%
%:%713=312%:%
%:%718=312%:%
%:%721=313%:%
%:%722=314%:%
%:%723=314%:%
%:%724=315%:%
%:%725=316%:%
%:%726=317%:%
%:%727=318%:%
%:%730=319%:%
%:%734=319%:%
%:%735=319%:%
%:%736=320%:%
%:%737=320%:%
%:%742=320%:%
%:%745=321%:%
%:%746=322%:%
%:%747=322%:%
%:%748=323%:%
%:%749=324%:%
%:%750=325%:%
%:%751=326%:%
%:%754=327%:%
%:%758=327%:%
%:%759=327%:%
%:%760=328%:%
%:%761=328%:%
%:%762=329%:%
%:%763=329%:%
%:%768=329%:%
%:%771=330%:%
%:%772=331%:%
%:%773=331%:%
%:%774=332%:%
%:%775=333%:%
%:%776=334%:%
%:%779=335%:%
%:%783=335%:%
%:%784=335%:%
%:%785=336%:%
%:%786=336%:%
%:%787=337%:%
%:%788=337%:%
%:%789=338%:%
%:%790=338%:%
%:%791=339%:%
%:%792=339%:%
%:%793=340%:%
%:%794=340%:%
%:%795=341%:%
%:%796=341%:%
%:%797=342%:%
%:%798=342%:%
%:%799=343%:%
%:%800=343%:%
%:%801=344%:%
%:%802=344%:%
%:%803=345%:%
%:%804=345%:%
%:%805=346%:%
%:%806=346%:%
%:%807=347%:%
%:%808=347%:%
%:%809=348%:%
%:%810=348%:%
%:%811=349%:%
%:%812=349%:%
%:%813=350%:%
%:%814=350%:%
%:%815=351%:%
%:%816=351%:%
%:%817=352%:%
%:%818=352%:%
%:%819=353%:%
%:%820=354%:%
%:%821=355%:%
%:%822=355%:%
%:%823=356%:%
%:%824=356%:%
%:%825=357%:%
%:%826=357%:%
%:%827=358%:%
%:%828=358%:%
%:%829=359%:%
%:%830=359%:%
%:%831=360%:%
%:%832=360%:%
%:%833=361%:%
%:%834=361%:%
%:%835=362%:%
%:%836=362%:%
%:%837=363%:%
%:%838=363%:%
%:%839=364%:%
%:%840=364%:%
%:%845=364%:%
%:%848=365%:%
%:%849=366%:%
%:%850=366%:%
%:%851=367%:%
%:%852=368%:%
%:%853=369%:%
%:%854=370%:%
%:%857=371%:%
%:%861=371%:%
%:%862=371%:%
%:%863=372%:%
%:%864=372%:%
%:%865=373%:%
%:%866=373%:%
%:%867=374%:%
%:%868=374%:%
%:%869=375%:%
%:%870=375%:%
%:%871=376%:%
%:%872=376%:%
%:%873=377%:%
%:%874=377%:%
%:%875=378%:%
%:%876=378%:%
%:%877=379%:%
%:%878=379%:%
%:%879=380%:%
%:%880=380%:%
%:%881=381%:%
%:%882=381%:%
%:%887=381%:%
%:%890=382%:%
%:%891=383%:%
%:%892=383%:%
%:%893=384%:%
%:%894=385%:%
%:%895=386%:%
%:%896=387%:%
%:%899=388%:%
%:%903=388%:%
%:%904=388%:%
%:%905=388%:%
%:%910=388%:%
%:%913=389%:%
%:%914=390%:%
%:%915=390%:%
%:%916=391%:%
%:%917=392%:%
%:%918=393%:%
%:%919=394%:%
%:%922=395%:%
%:%926=395%:%
%:%927=395%:%
%:%928=396%:%
%:%929=396%:%
%:%930=397%:%
%:%931=397%:%
%:%932=398%:%
%:%933=398%:%
%:%934=398%:%
%:%935=399%:%
%:%936=400%:%
%:%937=401%:%
%:%938=401%:%
%:%939=402%:%
%:%940=402%:%
%:%941=402%:%
%:%942=403%:%
%:%943=403%:%
%:%944=404%:%
%:%945=404%:%
%:%946=405%:%
%:%947=405%:%
%:%948=406%:%
%:%949=406%:%
%:%950=407%:%
%:%951=407%:%
%:%952=407%:%
%:%953=408%:%
%:%954=408%:%
%:%955=409%:%
%:%956=409%:%
%:%957=410%:%
%:%958=410%:%
%:%959=411%:%
%:%960=411%:%
%:%961=411%:%
%:%962=412%:%
%:%963=412%:%
%:%964=413%:%
%:%965=413%:%
%:%966=414%:%
%:%967=414%:%
%:%968=415%:%
%:%969=415%:%
%:%970=416%:%
%:%971=416%:%
%:%972=417%:%
%:%973=417%:%
%:%974=418%:%
%:%975=418%:%
%:%976=418%:%
%:%977=419%:%
%:%978=420%:%
%:%979=421%:%
%:%980=421%:%
%:%981=422%:%
%:%982=422%:%
%:%983=422%:%
%:%984=423%:%
%:%985=423%:%
%:%986=424%:%
%:%987=424%:%
%:%988=425%:%
%:%989=425%:%
%:%990=426%:%
%:%991=426%:%
%:%992=427%:%
%:%993=427%:%
%:%994=427%:%
%:%995=428%:%
%:%996=428%:%
%:%997=429%:%
%:%998=429%:%
%:%999=430%:%
%:%1000=430%:%
%:%1001=431%:%
%:%1002=431%:%
%:%1003=431%:%
%:%1004=432%:%
%:%1005=432%:%
%:%1006=433%:%
%:%1007=433%:%
%:%1008=434%:%
%:%1009=434%:%
%:%1010=435%:%
%:%1011=435%:%
%:%1012=436%:%
%:%1013=436%:%
%:%1014=437%:%
%:%1015=437%:%
%:%1016=438%:%
%:%1017=438%:%
%:%1018=439%:%
%:%1019=439%:%
%:%1020=440%:%
%:%1021=440%:%
%:%1022=441%:%
%:%1023=441%:%
%:%1024=442%:%
%:%1025=442%:%
%:%1026=443%:%
%:%1027=443%:%
%:%1028=444%:%
%:%1029=444%:%
%:%1030=445%:%
%:%1031=445%:%
%:%1032=446%:%
%:%1033=446%:%
%:%1034=446%:%
%:%1035=447%:%
%:%1036=448%:%
%:%1037=448%:%
%:%1038=449%:%
%:%1039=449%:%
%:%1040=450%:%
%:%1041=450%:%
%:%1042=450%:%
%:%1043=451%:%
%:%1044=451%:%
%:%1045=452%:%
%:%1046=452%:%
%:%1047=453%:%
%:%1048=453%:%
%:%1049=453%:%
%:%1050=454%:%
%:%1056=454%:%
%:%1059=455%:%
%:%1060=456%:%
%:%1061=457%:%
%:%1062=458%:%
%:%1063=458%:%
%:%1064=459%:%
%:%1065=460%:%
%:%1072=461%:%
%:%1073=461%:%
%:%1074=462%:%
%:%1075=462%:%
%:%1076=463%:%
%:%1077=463%:%
%:%1078=463%:%
%:%1079=464%:%
%:%1080=464%:%
%:%1081=465%:%
%:%1082=465%:%
%:%1083=466%:%
%:%1084=466%:%
%:%1085=467%:%
%:%1086=467%:%
%:%1087=468%:%
%:%1088=468%:%
%:%1089=469%:%
%:%1090=469%:%
%:%1091=470%:%
%:%1092=470%:%
%:%1093=471%:%
%:%1094=471%:%
%:%1095=472%:%
%:%1096=472%:%
%:%1097=473%:%
%:%1098=473%:%
%:%1099=474%:%
%:%1100=474%:%
%:%1101=474%:%
%:%1102=474%:%
%:%1103=475%:%
%:%1109=475%:%
%:%1112=476%:%
%:%1113=477%:%
%:%1114=477%:%
%:%1115=478%:%
%:%1116=479%:%
%:%1117=480%:%
%:%1118=481%:%
%:%1119=482%:%
%:%1120=483%:%
%:%1127=484%:%
%:%1128=484%:%
%:%1129=485%:%
%:%1130=485%:%
%:%1131=486%:%
%:%1132=486%:%
%:%1133=486%:%
%:%1134=487%:%
%:%1135=487%:%
%:%1136=488%:%
%:%1137=488%:%
%:%1138=489%:%
%:%1139=489%:%
%:%1140=490%:%
%:%1141=490%:%
%:%1142=491%:%
%:%1143=491%:%
%:%1144=492%:%
%:%1145=492%:%
%:%1146=493%:%
%:%1147=493%:%
%:%1148=494%:%
%:%1149=494%:%
%:%1150=494%:%
%:%1151=495%:%
%:%1152=495%:%
%:%1153=495%:%
%:%1154=495%:%
%:%1155=496%:%
%:%1161=496%:%
%:%1164=497%:%
%:%1165=498%:%
%:%1166=498%:%
%:%1167=499%:%
%:%1168=500%:%
%:%1169=501%:%
%:%1170=502%:%
%:%1171=503%:%
%:%1174=504%:%
%:%1178=504%:%
%:%1179=504%:%
%:%1180=504%:%
%:%1181=504%:%
%:%1186=504%:%
%:%1189=505%:%
%:%1190=506%:%
%:%1191=506%:%
%:%1192=507%:%
%:%1193=508%:%
%:%1194=509%:%
%:%1195=510%:%
%:%1198=511%:%
%:%1202=511%:%
%:%1203=511%:%
%:%1204=511%:%
%:%1205=512%:%
%:%1206=512%:%
%:%1211=512%:%
%:%1214=513%:%
%:%1215=514%:%
%:%1216=514%:%
%:%1217=515%:%
%:%1218=516%:%
%:%1219=517%:%
%:%1220=518%:%
%:%1221=519%:%
%:%1224=520%:%
%:%1228=520%:%
%:%1229=520%:%
%:%1230=521%:%
%:%1231=521%:%
%:%1232=522%:%
%:%1233=522%:%
%:%1234=523%:%
%:%1235=523%:%
%:%1236=523%:%
%:%1237=524%:%
%:%1238=524%:%
%:%1239=525%:%
%:%1240=525%:%
%:%1241=526%:%
%:%1242=526%:%
%:%1243=527%:%
%:%1244=527%:%
%:%1245=528%:%
%:%1246=528%:%
%:%1247=529%:%
%:%1248=529%:%
%:%1249=529%:%
%:%1250=530%:%
%:%1251=530%:%
%:%1252=531%:%
%:%1253=531%:%
%:%1254=532%:%
%:%1255=532%:%
%:%1256=532%:%
%:%1257=533%:%
%:%1258=533%:%
%:%1259=534%:%
%:%1260=534%:%
%:%1261=535%:%
%:%1262=535%:%
%:%1263=536%:%
%:%1264=536%:%
%:%1265=536%:%
%:%1266=537%:%
%:%1267=537%:%
%:%1268=538%:%
%:%1269=538%:%
%:%1270=539%:%
%:%1271=539%:%
%:%1272=540%:%
%:%1273=540%:%
%:%1274=541%:%
%:%1275=541%:%
%:%1276=542%:%
%:%1277=542%:%
%:%1278=543%:%
%:%1279=543%:%
%:%1280=543%:%
%:%1281=544%:%
%:%1282=544%:%
%:%1283=545%:%
%:%1284=545%:%
%:%1285=546%:%
%:%1286=546%:%
%:%1287=547%:%
%:%1288=547%:%
%:%1289=548%:%
%:%1290=548%:%
%:%1291=549%:%
%:%1292=549%:%
%:%1293=549%:%
%:%1294=550%:%
%:%1295=550%:%
%:%1296=551%:%
%:%1297=551%:%
%:%1298=552%:%
%:%1299=552%:%
%:%1300=552%:%
%:%1301=553%:%
%:%1302=553%:%
%:%1303=554%:%
%:%1304=554%:%
%:%1305=555%:%
%:%1306=555%:%
%:%1307=556%:%
%:%1308=556%:%
%:%1309=556%:%
%:%1310=557%:%
%:%1311=557%:%
%:%1312=558%:%
%:%1313=558%:%
%:%1314=559%:%
%:%1315=559%:%
%:%1316=560%:%
%:%1317=560%:%
%:%1318=561%:%
%:%1319=561%:%
%:%1320=562%:%
%:%1321=562%:%
%:%1322=563%:%
%:%1323=563%:%
%:%1324=564%:%
%:%1325=564%:%
%:%1326=565%:%
%:%1327=565%:%
%:%1328=566%:%
%:%1329=566%:%
%:%1330=567%:%
%:%1331=567%:%
%:%1332=568%:%
%:%1333=568%:%
%:%1334=568%:%
%:%1335=569%:%
%:%1336=569%:%
%:%1337=570%:%
%:%1338=570%:%
%:%1339=570%:%
%:%1340=571%:%
%:%1341=571%:%
%:%1342=571%:%
%:%1343=572%:%
%:%1344=572%:%
%:%1345=573%:%
%:%1346=573%:%
%:%1347=573%:%
%:%1348=574%:%
%:%1349=574%:%
%:%1350=575%:%
%:%1351=575%:%
%:%1352=576%:%
%:%1353=576%:%
%:%1354=576%:%
%:%1355=577%:%
%:%1356=577%:%
%:%1357=578%:%
%:%1358=578%:%
%:%1359=578%:%
%:%1360=579%:%
%:%1361=579%:%
%:%1362=580%:%
%:%1363=580%:%
%:%1364=581%:%
%:%1365=581%:%
%:%1366=581%:%
%:%1367=582%:%
%:%1368=582%:%
%:%1369=583%:%
%:%1370=583%:%
%:%1371=584%:%
%:%1372=584%:%
%:%1373=585%:%
%:%1374=585%:%
%:%1375=585%:%
%:%1376=586%:%
%:%1377=586%:%
%:%1378=587%:%
%:%1379=587%:%
%:%1380=588%:%
%:%1381=589%:%
%:%1382=589%:%
%:%1383=589%:%
%:%1386=592%:%
%:%1387=592%:%
%:%1388=593%:%
%:%1389=594%:%
%:%1390=594%:%
%:%1391=595%:%
%:%1392=595%:%
%:%1393=596%:%
%:%1394=596%:%
%:%1395=597%:%
%:%1396=597%:%
%:%1397=598%:%
%:%1398=598%:%
%:%1399=599%:%
%:%1400=599%:%
%:%1401=600%:%
%:%1402=600%:%
%:%1403=601%:%
%:%1404=601%:%
%:%1405=602%:%
%:%1406=602%:%
%:%1407=603%:%
%:%1408=603%:%
%:%1409=603%:%
%:%1410=604%:%
%:%1411=604%:%
%:%1412=605%:%
%:%1413=605%:%
%:%1414=605%:%
%:%1415=606%:%
%:%1416=606%:%
%:%1417=607%:%
%:%1418=607%:%
%:%1419=608%:%
%:%1420=608%:%
%:%1421=609%:%
%:%1422=609%:%
%:%1423=610%:%
%:%1424=611%:%
%:%1425=611%:%
%:%1426=612%:%
%:%1427=612%:%
%:%1428=613%:%
%:%1429=613%:%
%:%1430=614%:%
%:%1431=614%:%
%:%1432=615%:%
%:%1433=615%:%
%:%1434=616%:%
%:%1435=616%:%
%:%1436=617%:%
%:%1437=617%:%
%:%1438=618%:%
%:%1439=618%:%
%:%1440=619%:%
%:%1441=619%:%
%:%1442=620%:%
%:%1443=620%:%
%:%1444=621%:%
%:%1445=621%:%
%:%1446=622%:%
%:%1447=622%:%
%:%1448=623%:%
%:%1449=623%:%
%:%1450=624%:%
%:%1451=624%:%
%:%1452=625%:%
%:%1453=625%:%
%:%1454=625%:%
%:%1455=625%:%
%:%1456=626%:%
%:%1462=626%:%
%:%1465=627%:%
%:%1466=628%:%
%:%1467=628%:%
%:%1468=629%:%
%:%1469=630%:%
%:%1470=631%:%
%:%1471=632%:%
%:%1478=633%:%
%:%1479=633%:%
%:%1480=634%:%
%:%1481=634%:%
%:%1482=635%:%
%:%1483=635%:%
%:%1484=636%:%
%:%1485=636%:%
%:%1486=637%:%
%:%1487=637%:%
%:%1488=637%:%
%:%1489=638%:%
%:%1490=638%:%
%:%1491=639%:%
%:%1492=639%:%
%:%1493=640%:%
%:%1494=640%:%
%:%1495=641%:%
%:%1496=641%:%
%:%1497=642%:%
%:%1498=642%:%
%:%1499=642%:%
%:%1500=643%:%
%:%1506=643%:%
%:%1509=644%:%
%:%1510=645%:%
%:%1511=645%:%
%:%1512=646%:%
%:%1513=647%:%
%:%1514=648%:%
%:%1515=649%:%
%:%1516=650%:%
%:%1523=651%:%
%:%1524=651%:%
%:%1525=652%:%
%:%1526=652%:%
%:%1527=653%:%
%:%1528=653%:%
%:%1529=654%:%
%:%1530=654%:%
%:%1531=655%:%
%:%1532=655%:%
%:%1533=655%:%
%:%1534=656%:%
%:%1535=656%:%
%:%1536=657%:%
%:%1537=657%:%
%:%1538=658%:%
%:%1539=658%:%
%:%1540=659%:%
%:%1541=659%:%
%:%1542=660%:%
%:%1543=660%:%
%:%1544=661%:%
%:%1545=661%:%
%:%1546=661%:%
%:%1547=662%:%
%:%1548=662%:%
%:%1549=663%:%
%:%1550=663%:%
%:%1551=664%:%
%:%1552=664%:%
%:%1553=665%:%
%:%1554=665%:%
%:%1555=666%:%
%:%1556=666%:%
%:%1557=667%:%
%:%1558=667%:%
%:%1559=668%:%
%:%1560=668%:%
%:%1561=669%:%
%:%1562=669%:%
%:%1563=670%:%
%:%1564=670%:%
%:%1565=670%:%
%:%1566=671%:%
%:%1567=671%:%
%:%1568=672%:%
%:%1569=672%:%
%:%1570=673%:%
%:%1571=673%:%
%:%1572=674%:%
%:%1573=674%:%
%:%1574=674%:%
%:%1575=675%:%
%:%1581=675%:%
%:%1584=676%:%
%:%1585=677%:%
%:%1586=677%:%
%:%1587=678%:%
%:%1588=679%:%
%:%1589=680%:%
%:%1590=681%:%
%:%1591=682%:%
%:%1594=683%:%
%:%1598=683%:%
%:%1599=683%:%
%:%1600=683%:%
%:%1605=683%:%
%:%1608=684%:%
%:%1609=685%:%
%:%1610=685%:%
%:%1611=686%:%
%:%1612=687%:%
%:%1613=688%:%
%:%1614=689%:%
%:%1615=690%:%
%:%1622=691%:%
%:%1623=691%:%
%:%1624=692%:%
%:%1625=693%:%
%:%1626=693%:%
%:%1627=694%:%
%:%1628=694%:%
%:%1629=695%:%
%:%1630=696%:%
%:%1631=696%:%
%:%1632=697%:%
%:%1633=697%:%
%:%1634=698%:%
%:%1635=698%:%
%:%1636=699%:%
%:%1637=699%:%
%:%1638=700%:%
%:%1639=700%:%
%:%1640=701%:%
%:%1641=701%:%
%:%1642=702%:%
%:%1643=702%:%
%:%1644=703%:%
%:%1645=703%:%
%:%1646=704%:%
%:%1647=705%:%
%:%1648=705%:%
%:%1649=706%:%
%:%1650=706%:%
%:%1651=706%:%
%:%1652=707%:%
%:%1653=707%:%
%:%1654=708%:%
%:%1655=708%:%
%:%1656=709%:%
%:%1657=709%:%
%:%1658=710%:%
%:%1659=710%:%
%:%1660=710%:%
%:%1661=711%:%
%:%1662=711%:%
%:%1663=711%:%
%:%1664=712%:%
%:%1665=712%:%
%:%1666=713%:%
%:%1667=713%:%
%:%1668=714%:%
%:%1669=714%:%
%:%1670=714%:%
%:%1671=715%:%
%:%1672=715%:%
%:%1673=716%:%
%:%1674=716%:%
%:%1675=716%:%
%:%1676=717%:%
%:%1677=717%:%
%:%1678=718%:%
%:%1679=718%:%
%:%1680=719%:%
%:%1681=719%:%
%:%1682=719%:%
%:%1683=720%:%
%:%1684=720%:%
%:%1685=721%:%
%:%1686=721%:%
%:%1687=722%:%
%:%1688=722%:%
%:%1689=723%:%
%:%1690=723%:%
%:%1691=724%:%
%:%1692=724%:%
%:%1693=725%:%
%:%1694=725%:%
%:%1695=726%:%
%:%1696=726%:%
%:%1697=727%:%
%:%1698=727%:%
%:%1699=728%:%
%:%1700=728%:%
%:%1701=729%:%
%:%1702=730%:%
%:%1703=730%:%
%:%1704=731%:%
%:%1705=731%:%
%:%1706=732%:%
%:%1707=732%:%
%:%1708=733%:%
%:%1709=733%:%
%:%1710=733%:%
%:%1711=734%:%
%:%1712=734%:%
%:%1713=735%:%
%:%1714=735%:%
%:%1715=736%:%
%:%1716=736%:%
%:%1717=737%:%
%:%1718=737%:%
%:%1719=738%:%
%:%1720=738%:%
%:%1721=739%:%
%:%1722=739%:%
%:%1723=740%:%
%:%1724=740%:%
%:%1725=741%:%
%:%1726=741%:%
%:%1727=742%:%
%:%1728=742%:%
%:%1729=743%:%
%:%1730=743%:%
%:%1731=744%:%
%:%1732=744%:%
%:%1733=745%:%
%:%1734=745%:%
%:%1735=746%:%
%:%1736=746%:%
%:%1737=746%:%
%:%1738=747%:%
%:%1739=747%:%
%:%1740=748%:%
%:%1741=748%:%
%:%1742=749%:%
%:%1743=749%:%
%:%1744=749%:%
%:%1745=750%:%
%:%1746=750%:%
%:%1747=751%:%
%:%1748=751%:%
%:%1749=751%:%
%:%1750=752%:%
%:%1751=752%:%
%:%1752=753%:%
%:%1753=753%:%
%:%1754=754%:%
%:%1755=754%:%
%:%1756=755%:%
%:%1757=755%:%
%:%1758=756%:%
%:%1759=756%:%
%:%1760=757%:%
%:%1761=757%:%
%:%1762=758%:%
%:%1763=758%:%
%:%1764=759%:%
%:%1765=759%:%
%:%1766=760%:%
%:%1767=760%:%
%:%1768=761%:%
%:%1769=761%:%
%:%1770=761%:%
%:%1771=761%:%
%:%1772=762%:%
%:%1773=762%:%
%:%1774=763%:%
%:%1775=763%:%
%:%1776=764%:%
%:%1777=764%:%
%:%1778=765%:%
%:%1779=765%:%
%:%1780=765%:%
%:%1781=765%:%
%:%1782=766%:%
%:%1783=766%:%
%:%1784=767%:%
%:%1785=767%:%
%:%1786=768%:%
%:%1787=768%:%
%:%1788=769%:%
%:%1789=769%:%
%:%1790=770%:%
%:%1791=770%:%
%:%1792=771%:%
%:%1793=771%:%
%:%1794=772%:%
%:%1795=772%:%
%:%1796=773%:%
%:%1797=773%:%
%:%1798=774%:%
%:%1799=774%:%
%:%1800=774%:%
%:%1801=775%:%
%:%1802=775%:%
%:%1803=776%:%
%:%1804=776%:%
%:%1805=777%:%
%:%1806=777%:%
%:%1807=778%:%
%:%1808=778%:%
%:%1809=779%:%
%:%1810=779%:%
%:%1811=780%:%
%:%1812=780%:%
%:%1813=781%:%
%:%1814=781%:%
%:%1815=782%:%
%:%1816=782%:%
%:%1817=783%:%
%:%1818=783%:%
%:%1819=784%:%
%:%1820=784%:%
%:%1821=785%:%
%:%1822=785%:%
%:%1823=785%:%
%:%1824=786%:%
%:%1825=786%:%
%:%1826=787%:%
%:%1832=787%:%
%:%1835=788%:%
%:%1836=789%:%
%:%1837=789%:%
%:%1838=790%:%
%:%1839=791%:%
%:%1840=792%:%
%:%1841=793%:%
%:%1842=794%:%
%:%1843=795%:%
%:%1844=796%:%
%:%1845=797%:%
%:%1848=798%:%
%:%1853=799%:%
%:%1854=799%:%
%:%1855=800%:%
%:%1856=800%:%
%:%1857=801%:%
%:%1858=801%:%
%:%1859=802%:%
%:%1860=802%:%
%:%1861=803%:%
%:%1862=803%:%
%:%1863=804%:%
%:%1864=804%:%
%:%1865=805%:%
%:%1866=805%:%
%:%1867=806%:%
%:%1868=806%:%
%:%1869=807%:%
%:%1870=807%:%
%:%1871=808%:%
%:%1872=809%:%
%:%1873=809%:%
%:%1874=810%:%
%:%1875=810%:%
%:%1876=811%:%
%:%1877=811%:%
%:%1878=812%:%
%:%1879=812%:%
%:%1880=813%:%
%:%1881=813%:%
%:%1882=813%:%
%:%1883=814%:%
%:%1884=814%:%
%:%1885=815%:%
%:%1886=815%:%
%:%1887=816%:%
%:%1888=816%:%
%:%1889=817%:%
%:%1890=817%:%
%:%1891=818%:%
%:%1892=818%:%
%:%1893=819%:%
%:%1894=819%:%
%:%1895=820%:%
%:%1896=820%:%
%:%1897=821%:%
%:%1898=821%:%
%:%1899=822%:%
%:%1900=822%:%
%:%1901=823%:%
%:%1902=824%:%
%:%1903=824%:%
%:%1904=825%:%
%:%1905=825%:%
%:%1906=826%:%
%:%1907=826%:%
%:%1908=827%:%
%:%1909=827%:%
%:%1910=827%:%
%:%1911=828%:%
%:%1912=828%:%
%:%1913=829%:%
%:%1914=829%:%
%:%1915=829%:%
%:%1916=830%:%
%:%1917=830%:%
%:%1918=830%:%
%:%1919=831%:%
%:%1920=831%:%
%:%1921=832%:%
%:%1922=832%:%
%:%1923=833%:%
%:%1924=833%:%
%:%1925=833%:%
%:%1926=834%:%
%:%1927=834%:%
%:%1928=835%:%
%:%1929=835%:%
%:%1930=835%:%
%:%1931=836%:%
%:%1932=836%:%
%:%1933=837%:%
%:%1934=837%:%
%:%1935=838%:%
%:%1936=838%:%
%:%1937=839%:%
%:%1938=839%:%
%:%1939=840%:%
%:%1940=840%:%
%:%1941=841%:%
%:%1942=841%:%
%:%1943=842%:%
%:%1949=842%:%
%:%1952=843%:%
%:%1953=844%:%
%:%1954=844%:%
%:%1955=845%:%
%:%1956=846%:%
%:%1959=849%:%
%:%1960=850%:%
%:%1961=851%:%
%:%1962=851%:%
%:%1963=852%:%
%:%1966=853%:%
%:%1970=853%:%
%:%1971=853%:%
%:%1972=854%:%
%:%1973=854%:%
%:%1978=854%:%
%:%1981=855%:%
%:%1982=856%:%
%:%1983=856%:%
%:%1984=857%:%
%:%1986=859%:%
%:%1989=860%:%
%:%1993=860%:%
%:%1994=860%:%
%:%1995=861%:%
%:%1996=861%:%
%:%2001=861%:%
%:%2004=862%:%
%:%2005=863%:%
%:%2006=863%:%
%:%2007=864%:%
%:%2008=865%:%
%:%2011=866%:%
%:%2015=866%:%
%:%2016=866%:%
%:%2017=867%:%
%:%2018=867%:%
%:%2019=868%:%
%:%2020=868%:%
%:%2021=869%:%
%:%2022=869%:%
%:%2023=869%:%
%:%2024=869%:%
%:%2025=869%:%
%:%2026=870%:%
%:%2027=870%:%
%:%2028=871%:%
%:%2029=871%:%
%:%2030=872%:%
%:%2031=872%:%
%:%2032=872%:%
%:%2033=873%:%
%:%2034=873%:%
%:%2035=874%:%
%:%2036=874%:%
%:%2037=875%:%
%:%2038=875%:%
%:%2039=876%:%
%:%2040=876%:%
%:%2041=877%:%
%:%2042=877%:%
%:%2043=878%:%
%:%2044=878%:%
%:%2045=879%:%
%:%2046=879%:%
%:%2047=880%:%
%:%2048=880%:%
%:%2049=881%:%
%:%2050=881%:%
%:%2051=882%:%
%:%2052=882%:%
%:%2053=883%:%
%:%2054=883%:%
%:%2055=884%:%
%:%2056=884%:%
%:%2057=885%:%
%:%2063=885%:%
%:%2066=886%:%
%:%2067=887%:%
%:%2068=887%:%
%:%2069=888%:%
%:%2070=889%:%
%:%2073=890%:%
%:%2074=891%:%
%:%2078=891%:%
%:%2079=891%:%
%:%2080=892%:%
%:%2081=892%:%
%:%2082=893%:%
%:%2083=893%:%
%:%2084=894%:%
%:%2085=894%:%
%:%2086=895%:%
%:%2087=895%:%
%:%2088=896%:%
%:%2089=896%:%
%:%2090=897%:%
%:%2091=897%:%
%:%2092=898%:%
%:%2093=898%:%
%:%2094=899%:%
%:%2095=899%:%
%:%2096=900%:%
%:%2097=900%:%
%:%2098=901%:%
%:%2099=901%:%
%:%2100=902%:%
%:%2101=902%:%
%:%2102=903%:%
%:%2103=903%:%
%:%2104=904%:%
%:%2105=904%:%
%:%2106=905%:%
%:%2107=905%:%
%:%2108=906%:%
%:%2109=906%:%
%:%2110=907%:%
%:%2111=907%:%
%:%2112=908%:%
%:%2113=908%:%
%:%2114=909%:%
%:%2115=909%:%
%:%2120=909%:%
%:%2123=910%:%
%:%2124=911%:%
%:%2125=911%:%
%:%2126=912%:%
%:%2127=913%:%
%:%2128=914%:%
%:%2129=915%:%
%:%2136=916%:%
%:%2137=916%:%
%:%2138=917%:%
%:%2139=918%:%
%:%2140=918%:%
%:%2141=919%:%
%:%2142=919%:%
%:%2143=920%:%
%:%2144=920%:%
%:%2145=921%:%
%:%2146=921%:%
%:%2147=922%:%
%:%2148=922%:%
%:%2149=923%:%
%:%2150=923%:%
%:%2151=924%:%
%:%2152=924%:%
%:%2153=925%:%
%:%2154=925%:%
%:%2155=926%:%
%:%2156=926%:%
%:%2157=927%:%
%:%2158=927%:%
%:%2159=928%:%
%:%2160=928%:%
%:%2161=929%:%
%:%2162=929%:%
%:%2163=930%:%
%:%2164=931%:%
%:%2165=931%:%
%:%2166=932%:%
%:%2167=933%:%
%:%2168=933%:%
%:%2171=936%:%
%:%2172=937%:%
%:%2173=937%:%
%:%2174=937%:%
%:%2175=937%:%
%:%2176=938%:%
%:%2177=938%:%
%:%2178=939%:%
%:%2179=939%:%
%:%2180=940%:%
%:%2181=940%:%
%:%2182=941%:%
%:%2183=941%:%
%:%2184=942%:%
%:%2185=942%:%
%:%2186=943%:%
%:%2187=944%:%
%:%2188=944%:%
%:%2189=945%:%
%:%2190=945%:%
%:%2191=946%:%
%:%2192=946%:%
%:%2193=947%:%
%:%2194=947%:%
%:%2195=948%:%
%:%2196=948%:%
%:%2197=949%:%
%:%2198=949%:%
%:%2199=950%:%
%:%2200=950%:%
%:%2201=951%:%
%:%2202=951%:%
%:%2203=952%:%
%:%2204=952%:%
%:%2205=953%:%
%:%2206=953%:%
%:%2207=954%:%
%:%2208=954%:%
%:%2209=955%:%
%:%2210=955%:%
%:%2211=956%:%
%:%2212=956%:%
%:%2213=957%:%
%:%2214=957%:%
%:%2215=958%:%
%:%2216=958%:%
%:%2217=959%:%
%:%2218=959%:%
%:%2219=960%:%
%:%2220=960%:%
%:%2221=961%:%
%:%2222=961%:%
%:%2223=962%:%
%:%2224=962%:%
%:%2225=963%:%
%:%2226=963%:%
%:%2227=964%:%
%:%2228=964%:%
%:%2229=965%:%
%:%2230=965%:%
%:%2231=966%:%
%:%2232=966%:%
%:%2233=967%:%
%:%2234=967%:%
%:%2235=968%:%
%:%2236=968%:%
%:%2237=969%:%
%:%2238=969%:%
%:%2239=970%:%
%:%2240=970%:%
%:%2241=971%:%
%:%2242=971%:%
%:%2243=972%:%
%:%2244=972%:%
%:%2245=973%:%
%:%2246=973%:%
%:%2247=974%:%
%:%2248=974%:%
%:%2249=975%:%
%:%2250=975%:%
%:%2251=976%:%
%:%2252=977%:%
%:%2253=977%:%
%:%2254=978%:%
%:%2255=979%:%
%:%2256=979%:%
%:%2257=980%:%
%:%2258=980%:%
%:%2259=981%:%
%:%2260=981%:%
%:%2261=982%:%
%:%2262=982%:%
%:%2263=983%:%
%:%2264=983%:%
%:%2265=984%:%
%:%2266=984%:%
%:%2267=985%:%
%:%2268=985%:%
%:%2269=986%:%
%:%2270=986%:%
%:%2271=986%:%
%:%2272=987%:%
%:%2273=988%:%
%:%2274=988%:%
%:%2275=989%:%
%:%2276=989%:%
%:%2277=990%:%
%:%2278=990%:%
%:%2279=991%:%
%:%2280=991%:%
%:%2281=992%:%
%:%2282=992%:%
%:%2283=993%:%
%:%2284=993%:%
%:%2285=994%:%
%:%2286=994%:%
%:%2287=995%:%
%:%2288=995%:%
%:%2289=996%:%
%:%2290=996%:%
%:%2291=997%:%
%:%2292=997%:%
%:%2293=998%:%
%:%2294=998%:%
%:%2295=999%:%
%:%2296=999%:%
%:%2297=1000%:%
%:%2298=1000%:%
%:%2299=1001%:%
%:%2300=1001%:%
%:%2301=1002%:%
%:%2302=1002%:%
%:%2303=1003%:%
%:%2304=1003%:%
%:%2305=1004%:%
%:%2306=1004%:%
%:%2307=1005%:%
%:%2308=1005%:%
%:%2309=1006%:%
%:%2310=1006%:%
%:%2311=1007%:%
%:%2312=1007%:%
%:%2313=1008%:%
%:%2314=1008%:%
%:%2315=1009%:%
%:%2316=1009%:%
%:%2317=1010%:%
%:%2318=1010%:%
%:%2319=1011%:%
%:%2320=1011%:%
%:%2321=1012%:%
%:%2322=1012%:%
%:%2323=1013%:%
%:%2324=1013%:%
%:%2325=1014%:%
%:%2326=1014%:%
%:%2327=1015%:%
%:%2328=1015%:%
%:%2329=1016%:%
%:%2330=1016%:%
%:%2331=1017%:%
%:%2332=1017%:%
%:%2333=1018%:%
%:%2334=1018%:%
%:%2335=1019%:%
%:%2336=1019%:%
%:%2337=1020%:%
%:%2338=1020%:%
%:%2339=1021%:%
%:%2340=1021%:%
%:%2341=1022%:%
%:%2342=1022%:%
%:%2343=1022%:%
%:%2344=1023%:%
%:%2345=1023%:%
%:%2346=1024%:%
%:%2347=1024%:%
%:%2348=1025%:%
%:%2354=1025%:%
%:%2357=1026%:%
%:%2358=1027%:%
%:%2359=1027%:%
%:%2360=1028%:%
%:%2361=1029%:%
%:%2362=1030%:%
%:%2363=1031%:%
%:%2364=1032%:%
%:%2367=1033%:%
%:%2371=1033%:%
%:%2372=1033%:%
%:%2373=1034%:%
%:%2374=1034%:%
%:%2375=1035%:%
%:%2376=1035%:%
%:%2377=1036%:%
%:%2378=1036%:%
%:%2379=1036%:%
%:%2380=1036%:%
%:%2381=1037%:%
%:%2382=1038%:%
%:%2383=1038%:%
%:%2384=1039%:%
%:%2385=1039%:%
%:%2386=1040%:%
%:%2387=1040%:%
%:%2388=1041%:%
%:%2389=1041%:%
%:%2390=1042%:%
%:%2391=1043%:%
%:%2392=1043%:%
%:%2393=1044%:%
%:%2394=1044%:%
%:%2395=1045%:%
%:%2396=1045%:%
%:%2397=1046%:%
%:%2398=1046%:%
%:%2399=1047%:%
%:%2400=1047%:%
%:%2401=1048%:%
%:%2402=1049%:%
%:%2403=1049%:%
%:%2404=1050%:%
%:%2405=1050%:%
%:%2406=1051%:%
%:%2407=1051%:%
%:%2408=1052%:%
%:%2409=1052%:%
%:%2410=1053%:%
%:%2411=1054%:%
%:%2412=1054%:%
%:%2413=1055%:%
%:%2414=1055%:%
%:%2415=1056%:%
%:%2416=1056%:%
%:%2417=1057%:%
%:%2418=1057%:%
%:%2419=1058%:%
%:%2420=1059%:%
%:%2421=1059%:%
%:%2422=1060%:%
%:%2423=1061%:%
%:%2424=1061%:%
%:%2425=1062%:%
%:%2426=1062%:%
%:%2427=1063%:%
%:%2428=1063%:%
%:%2429=1064%:%
%:%2430=1064%:%
%:%2431=1064%:%
%:%2432=1065%:%
%:%2433=1065%:%
%:%2434=1066%:%
%:%2435=1066%:%
%:%2436=1067%:%
%:%2437=1067%:%
%:%2438=1068%:%
%:%2439=1068%:%
%:%2440=1069%:%
%:%2441=1069%:%
%:%2442=1070%:%
%:%2443=1070%:%
%:%2444=1071%:%
%:%2445=1071%:%
%:%2446=1072%:%
%:%2447=1072%:%
%:%2448=1073%:%
%:%2449=1073%:%
%:%2450=1074%:%
%:%2451=1075%:%
%:%2452=1075%:%
%:%2453=1076%:%
%:%2454=1076%:%
%:%2455=1077%:%
%:%2456=1078%:%
%:%2457=1078%:%
%:%2458=1079%:%
%:%2459=1079%:%
%:%2460=1080%:%
%:%2461=1080%:%
%:%2462=1080%:%
%:%2463=1080%:%
%:%2464=1081%:%
%:%2465=1082%:%
%:%2466=1082%:%
%:%2467=1083%:%
%:%2468=1083%:%
%:%2469=1084%:%
%:%2470=1084%:%
%:%2471=1085%:%
%:%2472=1085%:%
%:%2473=1086%:%
%:%2474=1087%:%
%:%2475=1087%:%
%:%2476=1088%:%
%:%2477=1088%:%
%:%2478=1089%:%
%:%2479=1089%:%
%:%2480=1090%:%
%:%2481=1090%:%
%:%2482=1091%:%
%:%2483=1092%:%
%:%2484=1092%:%
%:%2485=1093%:%
%:%2486=1093%:%
%:%2487=1094%:%
%:%2488=1094%:%
%:%2489=1095%:%
%:%2490=1095%:%
%:%2491=1096%:%
%:%2492=1096%:%
%:%2493=1097%:%
%:%2494=1098%:%
%:%2495=1098%:%
%:%2496=1099%:%
%:%2497=1099%:%
%:%2498=1100%:%
%:%2499=1100%:%
%:%2500=1101%:%
%:%2501=1101%:%
%:%2502=1102%:%
%:%2503=1103%:%
%:%2504=1103%:%
%:%2505=1104%:%
%:%2506=1105%:%
%:%2507=1105%:%
%:%2508=1106%:%
%:%2509=1106%:%
%:%2510=1107%:%
%:%2511=1107%:%
%:%2512=1108%:%
%:%2513=1108%:%
%:%2514=1108%:%
%:%2515=1109%:%
%:%2516=1109%:%
%:%2517=1110%:%
%:%2518=1110%:%
%:%2519=1111%:%
%:%2520=1111%:%
%:%2521=1112%:%
%:%2522=1112%:%
%:%2523=1113%:%
%:%2524=1113%:%
%:%2525=1114%:%
%:%2526=1114%:%
%:%2527=1115%:%
%:%2528=1115%:%
%:%2529=1116%:%
%:%2530=1116%:%
%:%2531=1117%:%
%:%2532=1117%:%
%:%2533=1118%:%
%:%2534=1119%:%
%:%2535=1119%:%
%:%2536=1120%:%
%:%2537=1120%:%
%:%2538=1121%:%
%:%2539=1121%:%
%:%2540=1122%:%
%:%2541=1122%:%
%:%2542=1123%:%
%:%2543=1124%:%
%:%2544=1124%:%
%:%2545=1125%:%
%:%2546=1125%:%
%:%2547=1126%:%
%:%2548=1126%:%
%:%2549=1127%:%
%:%2550=1127%:%
%:%2551=1128%:%
%:%2552=1128%:%
%:%2553=1129%:%
%:%2554=1129%:%
%:%2555=1130%:%
%:%2556=1130%:%
%:%2557=1131%:%
%:%2558=1131%:%
%:%2559=1132%:%
%:%2560=1132%:%
%:%2561=1133%:%
%:%2562=1133%:%
%:%2563=1134%:%
%:%2564=1134%:%
%:%2565=1135%:%
%:%2566=1135%:%
%:%2567=1136%:%
%:%2568=1136%:%
%:%2569=1137%:%
%:%2570=1137%:%
%:%2571=1138%:%
%:%2572=1138%:%
%:%2573=1139%:%
%:%2574=1139%:%
%:%2575=1140%:%
%:%2576=1140%:%
%:%2577=1141%:%
%:%2578=1142%:%
%:%2579=1142%:%
%:%2580=1143%:%
%:%2581=1143%:%
%:%2582=1144%:%
%:%2583=1144%:%
%:%2584=1145%:%
%:%2585=1145%:%
%:%2586=1146%:%
%:%2587=1146%:%
%:%2588=1147%:%
%:%2589=1147%:%
%:%2590=1148%:%
%:%2591=1148%:%
%:%2592=1149%:%
%:%2593=1149%:%
%:%2594=1150%:%
%:%2595=1150%:%
%:%2596=1151%:%
%:%2597=1151%:%
%:%2598=1152%:%
%:%2599=1152%:%
%:%2600=1153%:%
%:%2601=1153%:%
%:%2602=1154%:%
%:%2603=1154%:%
%:%2604=1155%:%
%:%2605=1155%:%
%:%2606=1156%:%
%:%2607=1156%:%
%:%2608=1157%:%
%:%2609=1157%:%
%:%2610=1158%:%
%:%2611=1158%:%
%:%2612=1159%:%
%:%2613=1159%:%
%:%2614=1160%:%
%:%2615=1161%:%
%:%2616=1161%:%
%:%2617=1162%:%
%:%2618=1162%:%
%:%2619=1163%:%
%:%2620=1163%:%
%:%2621=1164%:%
%:%2622=1164%:%
%:%2623=1165%:%
%:%2624=1165%:%
%:%2625=1166%:%
%:%2626=1166%:%
%:%2627=1167%:%
%:%2628=1167%:%
%:%2629=1168%:%
%:%2630=1168%:%
%:%2631=1169%:%
%:%2632=1170%:%
%:%2633=1170%:%
%:%2634=1171%:%
%:%2635=1172%:%
%:%2636=1172%:%
%:%2637=1173%:%
%:%2638=1173%:%
%:%2639=1174%:%
%:%2640=1175%:%
%:%2641=1175%:%
%:%2642=1176%:%
%:%2643=1176%:%
%:%2644=1177%:%
%:%2645=1177%:%
%:%2646=1178%:%
%:%2647=1178%:%
%:%2648=1179%:%
%:%2649=1179%:%
%:%2650=1180%:%
%:%2651=1180%:%
%:%2652=1181%:%
%:%2653=1181%:%
%:%2654=1181%:%
%:%2655=1182%:%
%:%2656=1182%:%
%:%2657=1183%:%
%:%2658=1183%:%
%:%2659=1184%:%
%:%2660=1184%:%
%:%2661=1184%:%
%:%2662=1185%:%
%:%2663=1185%:%
%:%2664=1185%:%
%:%2665=1186%:%
%:%2666=1186%:%
%:%2667=1187%:%
%:%2668=1187%:%
%:%2669=1188%:%
%:%2670=1188%:%
%:%2671=1189%:%
%:%2672=1189%:%
%:%2673=1190%:%
%:%2674=1190%:%
%:%2675=1191%:%
%:%2676=1191%:%
%:%2677=1192%:%
%:%2678=1192%:%
%:%2679=1192%:%
%:%2680=1193%:%
%:%2681=1193%:%
%:%2682=1194%:%
%:%2683=1194%:%
%:%2684=1195%:%
%:%2685=1195%:%
%:%2686=1196%:%
%:%2687=1196%:%
%:%2688=1197%:%
%:%2689=1197%:%
%:%2690=1198%:%
%:%2691=1198%:%
%:%2692=1199%:%
%:%2693=1199%:%
%:%2694=1200%:%
%:%2695=1200%:%
%:%2696=1201%:%
%:%2697=1202%:%
%:%2698=1202%:%
%:%2699=1203%:%
%:%2700=1203%:%
%:%2701=1204%:%
%:%2702=1204%:%
%:%2703=1205%:%
%:%2704=1205%:%
%:%2705=1206%:%
%:%2706=1206%:%
%:%2707=1207%:%
%:%2708=1207%:%
%:%2709=1208%:%
%:%2710=1208%:%
%:%2711=1209%:%
%:%2712=1209%:%
%:%2713=1210%:%
%:%2714=1210%:%
%:%2715=1210%:%
%:%2716=1211%:%
%:%2717=1211%:%
%:%2718=1212%:%
%:%2719=1212%:%
%:%2720=1212%:%
%:%2721=1212%:%
%:%2722=1213%:%
%:%2723=1213%:%
%:%2724=1214%:%
%:%2725=1214%:%
%:%2726=1214%:%
%:%2727=1215%:%
%:%2728=1216%:%
%:%2729=1216%:%
%:%2730=1217%:%
%:%2731=1217%:%
%:%2732=1218%:%
%:%2733=1218%:%
%:%2734=1218%:%
%:%2735=1218%:%
%:%2736=1218%:%
%:%2737=1219%:%
%:%2738=1219%:%
%:%2739=1220%:%
%:%2740=1220%:%
%:%2741=1220%:%
%:%2742=1220%:%
%:%2743=1221%:%
%:%2744=1221%:%
%:%2745=1222%:%
%:%2746=1222%:%
%:%2747=1223%:%
%:%2748=1223%:%
%:%2749=1224%:%
%:%2750=1224%:%
%:%2751=1225%:%
%:%2752=1225%:%
%:%2753=1226%:%
%:%2754=1226%:%
%:%2755=1227%:%
%:%2756=1227%:%
%:%2757=1228%:%
%:%2758=1228%:%
%:%2759=1229%:%
%:%2760=1229%:%
%:%2761=1230%:%
%:%2762=1230%:%
%:%2763=1230%:%
%:%2764=1230%:%
%:%2765=1230%:%
%:%2766=1231%:%
%:%2767=1231%:%
%:%2768=1232%:%
%:%2769=1232%:%
%:%2770=1233%:%
%:%2771=1234%:%
%:%2772=1234%:%
%:%2773=1235%:%
%:%2774=1236%:%
%:%2775=1236%:%
%:%2776=1237%:%
%:%2777=1237%:%
%:%2778=1238%:%
%:%2779=1238%:%
%:%2780=1239%:%
%:%2781=1239%:%
%:%2782=1239%:%
%:%2783=1239%:%
%:%2784=1239%:%
%:%2785=1240%:%
%:%2786=1240%:%
%:%2787=1240%:%
%:%2788=1240%:%
%:%2789=1241%:%
%:%2790=1241%:%
%:%2791=1241%:%
%:%2792=1242%:%
%:%2793=1242%:%
%:%2794=1243%:%
%:%2795=1243%:%
%:%2796=1244%:%
%:%2797=1244%:%
%:%2798=1245%:%
%:%2799=1245%:%
%:%2800=1246%:%
%:%2801=1246%:%
%:%2802=1247%:%
%:%2803=1247%:%
%:%2804=1248%:%
%:%2805=1248%:%
%:%2806=1249%:%
%:%2807=1249%:%
%:%2808=1250%:%
%:%2809=1250%:%
%:%2810=1251%:%
%:%2811=1251%:%
%:%2812=1251%:%
%:%2813=1252%:%
%:%2814=1252%:%
%:%2815=1253%:%
%:%2816=1253%:%
%:%2817=1254%:%
%:%2818=1254%:%
%:%2819=1255%:%
%:%2820=1255%:%
%:%2821=1256%:%
%:%2822=1256%:%
%:%2823=1257%:%
%:%2824=1257%:%
%:%2825=1258%:%
%:%2826=1258%:%
%:%2827=1259%:%
%:%2828=1259%:%
%:%2829=1260%:%
%:%2830=1260%:%
%:%2831=1261%:%
%:%2832=1261%:%
%:%2833=1262%:%
%:%2834=1262%:%
%:%2835=1263%:%
%:%2836=1263%:%
%:%2837=1264%:%
%:%2838=1264%:%
%:%2839=1265%:%
%:%2840=1265%:%
%:%2841=1266%:%
%:%2842=1266%:%
%:%2843=1267%:%
%:%2844=1267%:%
%:%2845=1267%:%
%:%2846=1268%:%
%:%2847=1268%:%
%:%2848=1269%:%
%:%2849=1269%:%
%:%2850=1269%:%
%:%2851=1269%:%
%:%2852=1270%:%
%:%2853=1270%:%
%:%2854=1271%:%
%:%2855=1271%:%
%:%2856=1272%:%
%:%2857=1272%:%
%:%2858=1273%:%
%:%2859=1273%:%
%:%2860=1274%:%
%:%2861=1274%:%
%:%2862=1275%:%
%:%2863=1275%:%
%:%2864=1275%:%
%:%2865=1276%:%
%:%2866=1276%:%
%:%2867=1277%:%
%:%2868=1277%:%
%:%2869=1278%:%
%:%2870=1278%:%
%:%2871=1279%:%
%:%2872=1279%:%
%:%2873=1280%:%
%:%2874=1280%:%
%:%2875=1281%:%
%:%2876=1281%:%
%:%2877=1282%:%
%:%2878=1283%:%
%:%2879=1283%:%
%:%2880=1284%:%
%:%2881=1284%:%
%:%2882=1284%:%
%:%2883=1285%:%
%:%2884=1285%:%
%:%2885=1285%:%
%:%2886=1286%:%
%:%2887=1286%:%
%:%2888=1287%:%
%:%2889=1287%:%
%:%2890=1288%:%
%:%2891=1288%:%
%:%2892=1289%:%
%:%2893=1289%:%
%:%2894=1290%:%
%:%2895=1290%:%
%:%2896=1291%:%
%:%2897=1291%:%
%:%2898=1291%:%
%:%2899=1292%:%
%:%2900=1292%:%
%:%2901=1293%:%
%:%2902=1293%:%
%:%2903=1294%:%
%:%2904=1294%:%
%:%2905=1295%:%
%:%2906=1295%:%
%:%2907=1296%:%
%:%2908=1296%:%
%:%2909=1297%:%
%:%2910=1297%:%
%:%2911=1298%:%
%:%2912=1298%:%
%:%2913=1299%:%
%:%2914=1299%:%
%:%2915=1299%:%
%:%2916=1300%:%
%:%2917=1300%:%
%:%2918=1301%:%
%:%2919=1301%:%
%:%2920=1302%:%
%:%2921=1302%:%
%:%2922=1303%:%
%:%2923=1303%:%
%:%2924=1304%:%
%:%2925=1304%:%
%:%2926=1305%:%
%:%2927=1305%:%
%:%2928=1305%:%
%:%2929=1306%:%
%:%2930=1306%:%
%:%2931=1306%:%
%:%2932=1307%:%
%:%2933=1308%:%
%:%2934=1308%:%
%:%2935=1309%:%
%:%2936=1309%:%
%:%2937=1310%:%
%:%2938=1310%:%
%:%2939=1311%:%
%:%2940=1311%:%
%:%2941=1312%:%
%:%2942=1312%:%
%:%2943=1313%:%
%:%2944=1313%:%
%:%2945=1314%:%
%:%2946=1314%:%
%:%2947=1315%:%
%:%2948=1315%:%
%:%2949=1316%:%
%:%2950=1316%:%
%:%2951=1317%:%
%:%2952=1317%:%
%:%2953=1317%:%
%:%2954=1318%:%
%:%2955=1318%:%
%:%2956=1319%:%
%:%2957=1319%:%
%:%2958=1320%:%
%:%2964=1320%:%
%:%2967=1321%:%
%:%2968=1322%:%
%:%2969=1322%:%
%:%2970=1323%:%
%:%2971=1324%:%
%:%2972=1325%:%
%:%2973=1326%:%
%:%2974=1327%:%
%:%2981=1328%:%
%:%2982=1328%:%
%:%2983=1329%:%
%:%2984=1329%:%
%:%2985=1330%:%
%:%2986=1330%:%
%:%2987=1331%:%
%:%2988=1331%:%
%:%2989=1332%:%
%:%2990=1332%:%
%:%2991=1333%:%
%:%2992=1333%:%
%:%2993=1334%:%
%:%2994=1334%:%
%:%2995=1335%:%
%:%2996=1335%:%
%:%2997=1336%:%
%:%2998=1336%:%
%:%2999=1337%:%
%:%3000=1338%:%
%:%3001=1338%:%
%:%3002=1339%:%
%:%3003=1339%:%
%:%3004=1340%:%
%:%3005=1340%:%
%:%3006=1341%:%
%:%3007=1341%:%
%:%3008=1342%:%
%:%3009=1343%:%
%:%3010=1343%:%
%:%3011=1344%:%
%:%3012=1344%:%
%:%3013=1345%:%
%:%3014=1345%:%
%:%3015=1346%:%
%:%3016=1346%:%
%:%3017=1347%:%
%:%3018=1347%:%
%:%3019=1348%:%
%:%3020=1348%:%
%:%3021=1349%:%
%:%3022=1349%:%
%:%3023=1349%:%
%:%3024=1350%:%
%:%3025=1351%:%
%:%3026=1351%:%
%:%3027=1351%:%
%:%3028=1352%:%
%:%3029=1352%:%
%:%3030=1353%:%
%:%3031=1353%:%
%:%3032=1353%:%
%:%3033=1353%:%
%:%3034=1354%:%
%:%3035=1354%:%
%:%3036=1354%:%
%:%3037=1355%:%
%:%3038=1355%:%
%:%3039=1356%:%
%:%3040=1356%:%
%:%3041=1357%:%
%:%3042=1357%:%
%:%3043=1358%:%
%:%3044=1359%:%
%:%3045=1359%:%
%:%3046=1360%:%
%:%3047=1360%:%
%:%3048=1361%:%
%:%3049=1361%:%
%:%3050=1362%:%
%:%3051=1362%:%
%:%3052=1363%:%
%:%3053=1363%:%
%:%3054=1364%:%
%:%3055=1364%:%
%:%3056=1364%:%
%:%3057=1365%:%
%:%3058=1365%:%
%:%3059=1365%:%
%:%3060=1366%:%
%:%3061=1366%:%
%:%3062=1366%:%
%:%3063=1367%:%
%:%3064=1367%:%
%:%3065=1368%:%
%:%3066=1368%:%
%:%3067=1369%:%
%:%3068=1369%:%
%:%3069=1369%:%
%:%3070=1370%:%
%:%3071=1370%:%
%:%3072=1371%:%
%:%3073=1371%:%
%:%3074=1371%:%
%:%3075=1372%:%
%:%3076=1372%:%
%:%3077=1373%:%
%:%3078=1373%:%
%:%3079=1374%:%
%:%3080=1374%:%
%:%3081=1375%:%
%:%3082=1375%:%
%:%3083=1376%:%
%:%3084=1376%:%
%:%3085=1377%:%
%:%3086=1377%:%
%:%3087=1378%:%
%:%3088=1378%:%
%:%3089=1379%:%
%:%3090=1379%:%
%:%3091=1380%:%
%:%3092=1380%:%
%:%3093=1381%:%
%:%3094=1381%:%
%:%3095=1382%:%
%:%3096=1382%:%
%:%3097=1383%:%
%:%3098=1383%:%
%:%3099=1384%:%
%:%3100=1384%:%
%:%3101=1385%:%
%:%3102=1385%:%
%:%3103=1386%:%
%:%3104=1386%:%
%:%3105=1386%:%
%:%3106=1387%:%
%:%3107=1387%:%
%:%3108=1388%:%
%:%3109=1388%:%
%:%3110=1389%:%
%:%3111=1389%:%
%:%3112=1390%:%
%:%3113=1390%:%
%:%3114=1391%:%
%:%3115=1391%:%
%:%3116=1392%:%
%:%3117=1392%:%
%:%3118=1393%:%
%:%3124=1393%:%
%:%3127=1394%:%
%:%3128=1395%:%
%:%3129=1396%:%
%:%3130=1396%:%
%:%3137=1397%:%