### 2024-11-01 code/output/document/HS_M.tex

%
\begin{isabellebody}%
\setisabellecontext{HS{\isacharunderscore}{\kern0pt}M}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\ \isanewline
\ \ \ \ HS{\isacharunderscore}{\kern0pt}Definition\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}g{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ param{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ Forall{\isacharparenleft}{\kern0pt}Implies{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ s\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ param\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}g\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ g\ s\ param\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}g\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}s\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}param\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}g{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ param{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}clarify{\isacharcomma}{\kern0pt}\ rule\ Implies{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ Exists{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ Exists{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ And{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ assms\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ g\ s\ param\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}g\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}s\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}param\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}g{\isacharcomma}{\kern0pt}\ s{\isacharcomma}{\kern0pt}\ param{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}s{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}param{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}param{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp\ {\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ env\ i\ j\ k\ g\ s\ param\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ s{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ param{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ s\ {\isasymlongrightarrow}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ param{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ y\ {\isasymin}\ s\ {\isasymlongrightarrow}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ param{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Forall{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Implies{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ imp{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isasymexists}a\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}v\ {\isasymin}\ M{\isachardot}{\kern0pt}\ a\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ param{\isachargreater}{\kern0pt}\ {\isasymand}\ \ g{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ v\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}i\ {\isasymin}\ nat\ {\isasymand}\ k\ {\isasymin}\ nat\ {\isasymand}\ g\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nth{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ param{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nth{\isacharunderscore}{\kern0pt}type\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ s\ {\isasymlongrightarrow}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ param{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ y{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}s\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transM\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nth{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ G{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}v{\isacharcomma}{\kern0pt}\ G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ v\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ x\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ x\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \ \ \ \ \ \ \isanewline
\ \ \isakeyword{fixes}\ p\ G\ x\ v\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}G\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ G{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Exists{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \ \ \ \isanewline
\ \ \isakeyword{fixes}\ p\ G\ x\ v\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}G\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ G{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ \ succ{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ env\ p\ G\ x\ v\ i\ j\ k\ l\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymG}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}l{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}v\ {\isasymin}\ {\isasymG}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}u\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ u\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ u\ {\isacharequal}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ nth{\isacharunderscore}{\kern0pt}type\ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{9}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}v\ {\isasymin}\ {\isasymG}\ {\isasymand}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ G{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ s{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ Forall{\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ s\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ G\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ x\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ p\ G\ x\ s\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}G\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}s\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ G{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ s{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}clarify{\isacharcomma}{\kern0pt}\ rule\ Iff{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ p\ G\ x\ s\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}G\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}s\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ G{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ s{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}s{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ G\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ x\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ env\ i\ j\ k\ l\ p\ G\ x\ s\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymG}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}l{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ s{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ s\ {\isacharequal}{\kern0pt}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}lemma\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}a\ b\ c{\isachardot}{\kern0pt}\ b\ {\isacharequal}{\kern0pt}\ c\ {\isasymLongrightarrow}\ a\ {\isacharequal}{\kern0pt}\ b\ {\isasymlongleftrightarrow}\ a\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ innat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat\ {\isasymand}\ j\ {\isasymin}\ nat\ {\isasymand}\ k\ {\isasymin}\ nat\ {\isasymand}\ l\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}{\isasympi}\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasympi}\ {\isasymin}\ s\ {\isasymlongleftrightarrow}\ {\isasympi}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Forall{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Iff{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ innat\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ innat\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ s\ {\isacharequal}{\kern0pt}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ {\isasymtau}{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}s\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transM\ assms\ nth{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ {\isasymtau}{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transM\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ sym{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ S\ \isakeyword{where}\ {\isachardoublequoteopen}S\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ v\ {\isasymin}\ {\isasymG}{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}v{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ SinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}S\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ S{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}notation{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}S\ {\isacharequal}{\kern0pt}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ S{\isacharunderscore}{\kern0pt}def\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ SinM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ x{\isacharprime}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}domain{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}g\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{9}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{9}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x{\isacharprime}{\kern0pt}\ g\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ Exists{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x{\isacharprime}{\kern0pt}\ g\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{9}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp\ {\isacharcomma}{\kern0pt}rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}ren\ \isakeyword{where}\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}ren{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ i{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ j{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ ren{\isacharparenleft}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{2}}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}ren{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ren{\isacharunderscore}{\kern0pt}tc{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ range{\isacharunderscore}{\kern0pt}def\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ arity{\isacharunderscore}{\kern0pt}ren{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ range{\isacharunderscore}{\kern0pt}def\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ env\ x{\isacharprime}{\kern0pt}\ g\ i\ j\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ren{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ nth{\isacharunderscore}{\kern0pt}type\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}i\ {\isasymin}\ nat\ {\isasymand}\ j\ {\isasymin}\ nat{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ k{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}k\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymor}\ k\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ i{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ j{\isasymrangle}{\isacharbraceright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}i\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}function{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ i{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ j{\isasymrangle}{\isacharbraceright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}j\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}function{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Forall{\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ v\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ g\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ empty{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x{\isacharprime}{\kern0pt}\ g\ v\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ g\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x{\isacharprime}{\kern0pt}\ g\ v\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ v{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ ltI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M\ \isakeyword{where}\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\ \ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}if\ {\isacharparenleft}{\kern0pt}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ then\ {\isadigit{1}}\ else\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ env\ i\ j\ k\ x{\isacharprime}{\kern0pt}\ g\ v\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isacharequal}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ innat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat\ {\isasymand}\ j\ {\isasymin}\ nat\ {\isasymand}\ k\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ inM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isacharprime}{\kern0pt}\ {\isasymin}\ M\ {\isasymand}\ g\ {\isasymin}\ M\ {\isasymand}\ v\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}type\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}s\ {\isasymin}\ M{\isachardot}{\kern0pt}\ s\ {\isasymin}\ v\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}i\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ j\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}s{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ s\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Forall{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Iff{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}assms\ innat\ inM{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}s\ {\isasymin}\ M{\isachardot}{\kern0pt}\ s\ {\isasymin}\ v\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ s\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ innat\ inM\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}s{\isachardot}{\kern0pt}\ s\ {\isasymin}\ v\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ s\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ clarify{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ s{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}s\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ inM\ transM\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isacharequal}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}FGppauto{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ FGppauto\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Exists{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ And{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ ltD{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ if\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}y\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isacharparenright}{\kern0pt}\ then\ {\isadigit{1}}\ else\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}h\ g\ u{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ u\ {\isasymin}\ M\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ h\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ g\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymrightarrow}\ M\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ h\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ His{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}u{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}{\isasymlangle}u{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ h\ g\ u\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}h\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymrightarrow}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ h\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymlangle}u{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}symx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}domx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pauto\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}ppauto\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \ {\isasymexists}G\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \ {\isasymexists}Gppauto\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}F\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}FGppauto\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isasymlangle}u{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ FGppauto{\isachargreater}{\kern0pt}\ {\isasymand}\ FGppauto\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}F{\isacharcomma}{\kern0pt}\ Gppauto{\isachargreater}{\kern0pt}\ {\isasymand}\ Gppauto\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}G{\isacharcomma}{\kern0pt}\ ppauto{\isachargreater}{\kern0pt}\ {\isasymand}\ ppauto\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}p{\isacharcomma}{\kern0pt}\ pauto{\isachargreater}{\kern0pt}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ domx\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ domx\ {\isasymlongrightarrow}\ g{\isacharbackquote}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ symx\ {\isacharequal}{\kern0pt}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ symx\ {\isasymin}\ {\isasymF}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymlangle}u{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ symx\ domx\ pauto\ p\ ppauto\ G\ Gppauto\ F\ FGppauto\ x{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}forcing{\isacharunderscore}{\kern0pt}data{\isachardot}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}P{\isacharunderscore}{\kern0pt}names\ \isakeyword{in}\ ssubst{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}all{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{7}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sym{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ sym{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ domain{\isacharunderscore}{\kern0pt}closed\ transM\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ h{\isacharbackquote}{\kern0pt}y\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ sym{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ assms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}u{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}{\isasymlangle}u{\isacharcomma}{\kern0pt}\ {\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isasymrangle}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}def\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ if{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HS\ \isakeyword{where}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}u{\isachardot}{\kern0pt}\ u\ {\isasymin}\ M\ {\isasymlongrightarrow}\ is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}u{\isachardot}{\kern0pt}\ u\ {\isasymin}\ M\ {\isasymlongrightarrow}\ is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ u\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ u{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ M\ {\isasymlongrightarrow}\ is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ y\ {\isasymin}\ HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}u\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \isacommand{define}\isamarkupfalse%
\ F\ \isakeyword{where}\ {\isachardoublequoteopen}F\ {\isasymequiv}\ {\isasymlambda}y\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}u{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}lemma\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}a\ b\ c{\isachardot}{\kern0pt}\ a\ {\isacharequal}{\kern0pt}\ b\ {\isasymLongrightarrow}\ a\ {\isacharequal}{\kern0pt}\ c\ {\isasymlongleftrightarrow}\ b\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ His{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}u{\isacharcomma}{\kern0pt}\ F{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}def\ F{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}lemma{\isacharcomma}{\kern0pt}\ rule\ wftrec{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ wf{\isacharunderscore}{\kern0pt}trancl{\isacharcomma}{\kern0pt}\ rule\ wf{\isacharunderscore}{\kern0pt}Memrel{\isacharcomma}{\kern0pt}\ rule\ trans{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}y\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ sym{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isadigit{1}}\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ ifT{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ sym{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ y{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}u{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}u\ \isakeyword{in}\ vimageI{\isacharcomma}{\kern0pt}\ rule\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharparenright}{\kern0pt}\ {\isasymand}\ sym{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ y{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isacommand{unfolding}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}def\ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ HS{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ HS{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ symmetric{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ u\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ i\ j\ env\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ innat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat\ {\isasymand}\ j\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}v\ {\isasymin}\ M{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ His{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ G{\isacharequal}{\kern0pt}{\isachardoublequoteopen}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isachardoublequoteclose}\ \isakeyword{in}\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ innat\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{9}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ succ{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MI\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}His{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{7}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}HS{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS\ assms\ nth{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}separation\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}A\ {\isasyminter}\ HS\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ sep\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ S\ \isakeyword{where}\ {\isachardoublequoteopen}S\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharless}{\kern0pt}{\isasymF}{\isacharcomma}{\kern0pt}\ {\isasymG}{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}auto{\isachargreater}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ SinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}S\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ S{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}notation{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ sep\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}S\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isasymin}\ HS\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ S{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymF}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ transM\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\ \isanewline
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ A\ {\isasyminter}\ HS{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ SinM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/HS_M.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=8%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=11%:%
%:%30=12%:%
%:%31=12%:%
%:%32=13%:%
%:%33=14%:%
%:%34=15%:%
%:%35=15%:%
%:%36=16%:%
%:%37=17%:%
%:%38=18%:%
%:%41=19%:%
%:%45=19%:%
%:%46=19%:%
%:%47=20%:%
%:%48=20%:%
%:%49=21%:%
%:%50=21%:%
%:%51=22%:%
%:%52=22%:%
%:%53=23%:%
%:%54=23%:%
%:%55=24%:%
%:%56=24%:%
%:%61=24%:%
%:%64=25%:%
%:%65=26%:%
%:%66=26%:%
%:%67=27%:%
%:%68=28%:%
%:%69=29%:%
%:%72=30%:%
%:%76=30%:%
%:%77=30%:%
%:%78=31%:%
%:%79=31%:%
%:%80=32%:%
%:%81=32%:%
%:%82=33%:%
%:%83=33%:%
%:%84=34%:%
%:%85=34%:%
%:%86=35%:%
%:%87=35%:%
%:%88=36%:%
%:%89=36%:%
%:%90=37%:%
%:%91=37%:%
%:%92=38%:%
%:%93=38%:%
%:%94=39%:%
%:%95=39%:%
%:%96=40%:%
%:%97=40%:%
%:%98=41%:%
%:%99=41%:%
%:%100=42%:%
%:%101=42%:%
%:%102=43%:%
%:%103=43%:%
%:%104=44%:%
%:%105=44%:%
%:%106=45%:%
%:%107=45%:%
%:%108=46%:%
%:%109=46%:%
%:%110=47%:%
%:%111=47%:%
%:%112=48%:%
%:%113=48%:%
%:%114=49%:%
%:%115=49%:%
%:%116=50%:%
%:%117=50%:%
%:%118=51%:%
%:%119=51%:%
%:%120=52%:%
%:%121=52%:%
%:%122=53%:%
%:%123=53%:%
%:%124=54%:%
%:%125=54%:%
%:%126=55%:%
%:%127=55%:%
%:%128=56%:%
%:%129=56%:%
%:%134=56%:%
%:%137=57%:%
%:%138=58%:%
%:%139=58%:%
%:%140=59%:%
%:%141=60%:%
%:%142=61%:%
%:%149=62%:%
%:%150=62%:%
%:%151=63%:%
%:%152=63%:%
%:%153=64%:%
%:%154=64%:%
%:%155=65%:%
%:%156=65%:%
%:%157=66%:%
%:%158=66%:%
%:%159=67%:%
%:%160=67%:%
%:%161=68%:%
%:%162=68%:%
%:%163=69%:%
%:%164=69%:%
%:%165=70%:%
%:%166=70%:%
%:%167=71%:%
%:%168=71%:%
%:%169=72%:%
%:%170=72%:%
%:%171=73%:%
%:%172=73%:%
%:%173=74%:%
%:%174=74%:%
%:%175=75%:%
%:%176=75%:%
%:%177=76%:%
%:%178=76%:%
%:%179=77%:%
%:%180=77%:%
%:%181=78%:%
%:%182=78%:%
%:%183=79%:%
%:%184=79%:%
%:%185=80%:%
%:%186=80%:%
%:%187=81%:%
%:%188=81%:%
%:%189=82%:%
%:%190=82%:%
%:%191=83%:%
%:%192=83%:%
%:%193=84%:%
%:%194=84%:%
%:%195=85%:%
%:%196=85%:%
%:%197=86%:%
%:%198=86%:%
%:%199=87%:%
%:%200=87%:%
%:%201=88%:%
%:%202=88%:%
%:%203=89%:%
%:%204=89%:%
%:%205=90%:%
%:%206=90%:%
%:%207=91%:%
%:%208=91%:%
%:%209=92%:%
%:%210=92%:%
%:%211=93%:%
%:%212=93%:%
%:%213=94%:%
%:%214=94%:%
%:%215=95%:%
%:%216=95%:%
%:%217=96%:%
%:%218=96%:%
%:%219=97%:%
%:%220=97%:%
%:%221=98%:%
%:%222=98%:%
%:%223=99%:%
%:%224=99%:%
%:%225=100%:%
%:%226=100%:%
%:%227=101%:%
%:%228=101%:%
%:%229=102%:%
%:%230=102%:%
%:%231=102%:%
%:%232=103%:%
%:%233=103%:%
%:%234=104%:%
%:%235=104%:%
%:%236=105%:%
%:%237=105%:%
%:%238=106%:%
%:%239=106%:%
%:%240=107%:%
%:%241=107%:%
%:%242=108%:%
%:%243=108%:%
%:%244=109%:%
%:%245=109%:%
%:%246=110%:%
%:%247=110%:%
%:%248=110%:%
%:%249=110%:%
%:%250=111%:%
%:%256=111%:%
%:%259=112%:%
%:%260=113%:%
%:%261=113%:%
%:%262=114%:%
%:%263=115%:%
%:%264=115%:%
%:%265=116%:%
%:%266=117%:%
%:%267=118%:%
%:%268=119%:%
%:%269=119%:%
%:%270=120%:%
%:%271=121%:%
%:%272=122%:%
%:%273=122%:%
%:%274=123%:%
%:%275=124%:%
%:%276=125%:%
%:%279=126%:%
%:%280=127%:%
%:%284=127%:%
%:%285=127%:%
%:%286=128%:%
%:%287=128%:%
%:%288=129%:%
%:%289=129%:%
%:%290=130%:%
%:%291=130%:%
%:%292=131%:%
%:%293=131%:%
%:%294=132%:%
%:%295=132%:%
%:%296=133%:%
%:%297=133%:%
%:%302=133%:%
%:%305=134%:%
%:%306=135%:%
%:%307=135%:%
%:%308=136%:%
%:%309=137%:%
%:%310=138%:%
%:%313=139%:%
%:%314=140%:%
%:%318=140%:%
%:%319=140%:%
%:%320=141%:%
%:%321=141%:%
%:%322=142%:%
%:%323=142%:%
%:%324=143%:%
%:%325=143%:%
%:%326=144%:%
%:%327=144%:%
%:%328=145%:%
%:%329=145%:%
%:%330=146%:%
%:%331=146%:%
%:%332=147%:%
%:%333=147%:%
%:%334=148%:%
%:%335=148%:%
%:%336=149%:%
%:%337=149%:%
%:%338=150%:%
%:%339=150%:%
%:%340=151%:%
%:%341=151%:%
%:%342=152%:%
%:%343=152%:%
%:%344=153%:%
%:%345=153%:%
%:%346=154%:%
%:%347=154%:%
%:%348=155%:%
%:%349=155%:%
%:%350=156%:%
%:%351=156%:%
%:%352=157%:%
%:%353=157%:%
%:%354=158%:%
%:%355=158%:%
%:%356=159%:%
%:%357=159%:%
%:%358=160%:%
%:%359=160%:%
%:%364=160%:%
%:%367=161%:%
%:%368=162%:%
%:%369=162%:%
%:%370=163%:%
%:%371=164%:%
%:%372=165%:%
%:%373=166%:%
%:%374=167%:%
%:%381=168%:%
%:%382=168%:%
%:%383=169%:%
%:%384=169%:%
%:%385=170%:%
%:%386=170%:%
%:%387=171%:%
%:%388=171%:%
%:%389=172%:%
%:%390=172%:%
%:%391=173%:%
%:%392=173%:%
%:%393=174%:%
%:%394=174%:%
%:%395=175%:%
%:%396=175%:%
%:%397=175%:%
%:%398=176%:%
%:%399=176%:%
%:%400=177%:%
%:%401=177%:%
%:%402=178%:%
%:%403=178%:%
%:%404=179%:%
%:%405=179%:%
%:%406=180%:%
%:%407=180%:%
%:%408=181%:%
%:%409=181%:%
%:%410=181%:%
%:%411=182%:%
%:%412=182%:%
%:%413=183%:%
%:%414=183%:%
%:%415=184%:%
%:%416=184%:%
%:%417=185%:%
%:%418=185%:%
%:%419=186%:%
%:%420=186%:%
%:%421=187%:%
%:%422=187%:%
%:%423=187%:%
%:%424=188%:%
%:%425=188%:%
%:%426=189%:%
%:%427=189%:%
%:%428=190%:%
%:%429=190%:%
%:%430=191%:%
%:%431=191%:%
%:%432=191%:%
%:%433=191%:%
%:%434=192%:%
%:%440=192%:%
%:%443=193%:%
%:%444=193%:%
%:%445=194%:%
%:%446=195%:%
%:%447=195%:%
%:%448=196%:%
%:%449=197%:%
%:%450=198%:%
%:%451=199%:%
%:%452=200%:%
%:%453=200%:%
%:%454=201%:%
%:%455=202%:%
%:%456=203%:%
%:%457=203%:%
%:%458=204%:%
%:%459=205%:%
%:%460=206%:%
%:%463=207%:%
%:%467=207%:%
%:%468=207%:%
%:%469=208%:%
%:%470=208%:%
%:%471=209%:%
%:%472=209%:%
%:%473=210%:%
%:%474=210%:%
%:%475=211%:%
%:%476=211%:%
%:%477=212%:%
%:%478=212%:%
%:%479=213%:%
%:%480=213%:%
%:%485=213%:%
%:%488=214%:%
%:%489=215%:%
%:%490=215%:%
%:%491=216%:%
%:%492=217%:%
%:%493=218%:%
%:%496=219%:%
%:%497=220%:%
%:%501=220%:%
%:%502=220%:%
%:%503=221%:%
%:%504=221%:%
%:%505=222%:%
%:%506=222%:%
%:%507=223%:%
%:%508=223%:%
%:%509=224%:%
%:%510=224%:%
%:%511=225%:%
%:%512=225%:%
%:%513=226%:%
%:%514=226%:%
%:%515=227%:%
%:%516=227%:%
%:%517=228%:%
%:%518=228%:%
%:%519=229%:%
%:%520=229%:%
%:%521=230%:%
%:%522=230%:%
%:%523=231%:%
%:%524=231%:%
%:%525=232%:%
%:%526=232%:%
%:%527=233%:%
%:%528=233%:%
%:%529=234%:%
%:%530=234%:%
%:%531=235%:%
%:%532=235%:%
%:%533=236%:%
%:%534=236%:%
%:%539=236%:%
%:%542=237%:%
%:%543=238%:%
%:%544=238%:%
%:%545=239%:%
%:%546=240%:%
%:%547=241%:%
%:%548=242%:%
%:%555=243%:%
%:%556=243%:%
%:%557=244%:%
%:%558=245%:%
%:%559=245%:%
%:%560=245%:%
%:%561=246%:%
%:%562=247%:%
%:%563=247%:%
%:%564=248%:%
%:%565=248%:%
%:%566=248%:%
%:%567=249%:%
%:%568=250%:%
%:%569=250%:%
%:%570=251%:%
%:%571=251%:%
%:%572=252%:%
%:%573=252%:%
%:%574=253%:%
%:%575=253%:%
%:%576=254%:%
%:%577=254%:%
%:%578=255%:%
%:%579=255%:%
%:%580=256%:%
%:%581=256%:%
%:%582=257%:%
%:%583=257%:%
%:%584=258%:%
%:%585=258%:%
%:%586=259%:%
%:%587=259%:%
%:%588=260%:%
%:%589=260%:%
%:%590=260%:%
%:%591=261%:%
%:%592=261%:%
%:%593=262%:%
%:%594=262%:%
%:%595=263%:%
%:%596=263%:%
%:%597=264%:%
%:%598=264%:%
%:%599=265%:%
%:%600=265%:%
%:%601=266%:%
%:%602=266%:%
%:%603=267%:%
%:%604=267%:%
%:%605=268%:%
%:%606=268%:%
%:%607=269%:%
%:%608=269%:%
%:%609=270%:%
%:%610=270%:%
%:%611=270%:%
%:%612=270%:%
%:%613=271%:%
%:%619=271%:%
%:%622=272%:%
%:%623=273%:%
%:%624=273%:%
%:%625=274%:%
%:%626=275%:%
%:%627=276%:%
%:%630=277%:%
%:%635=278%:%
%:%636=278%:%
%:%637=279%:%
%:%638=280%:%
%:%639=280%:%
%:%640=281%:%
%:%641=282%:%
%:%642=282%:%
%:%643=283%:%
%:%644=283%:%
%:%645=284%:%
%:%646=284%:%
%:%647=285%:%
%:%648=285%:%
%:%649=286%:%
%:%650=286%:%
%:%651=287%:%
%:%652=287%:%
%:%653=288%:%
%:%654=288%:%
%:%655=289%:%
%:%656=289%:%
%:%657=290%:%
%:%658=290%:%
%:%659=291%:%
%:%660=291%:%
%:%661=292%:%
%:%662=293%:%
%:%663=293%:%
%:%664=293%:%
%:%665=294%:%
%:%666=294%:%
%:%667=295%:%
%:%668=295%:%
%:%669=296%:%
%:%670=296%:%
%:%671=297%:%
%:%672=297%:%
%:%673=298%:%
%:%674=299%:%
%:%675=299%:%
%:%676=300%:%
%:%677=300%:%
%:%678=301%:%
%:%679=301%:%
%:%680=302%:%
%:%681=302%:%
%:%682=303%:%
%:%683=303%:%
%:%684=304%:%
%:%685=304%:%
%:%686=305%:%
%:%687=305%:%
%:%688=306%:%
%:%689=306%:%
%:%690=307%:%
%:%691=307%:%
%:%692=307%:%
%:%693=307%:%
%:%694=307%:%
%:%695=308%:%
%:%701=308%:%
%:%704=309%:%
%:%705=310%:%
%:%706=310%:%
%:%707=311%:%
%:%708=323%:%
%:%709=324%:%
%:%710=325%:%
%:%711=325%:%
%:%712=326%:%
%:%721=335%:%
%:%722=336%:%
%:%723=337%:%
%:%724=338%:%
%:%725=338%:%
%:%726=339%:%
%:%727=340%:%
%:%728=341%:%
%:%729=341%:%
%:%730=342%:%
%:%731=343%:%
%:%732=344%:%
%:%735=345%:%
%:%736=346%:%
%:%740=346%:%
%:%741=346%:%
%:%742=347%:%
%:%743=347%:%
%:%744=348%:%
%:%745=348%:%
%:%746=349%:%
%:%747=349%:%
%:%748=350%:%
%:%749=350%:%
%:%750=351%:%
%:%751=351%:%
%:%752=352%:%
%:%753=352%:%
%:%758=352%:%
%:%761=353%:%
%:%762=354%:%
%:%763=354%:%
%:%764=355%:%
%:%765=356%:%
%:%766=357%:%
%:%769=358%:%
%:%770=359%:%
%:%774=359%:%
%:%775=359%:%
%:%776=360%:%
%:%777=360%:%
%:%778=361%:%
%:%779=361%:%
%:%780=362%:%
%:%781=362%:%
%:%782=363%:%
%:%783=363%:%
%:%784=364%:%
%:%785=364%:%
%:%786=365%:%
%:%787=365%:%
%:%788=366%:%
%:%789=366%:%
%:%790=367%:%
%:%791=367%:%
%:%792=368%:%
%:%793=368%:%
%:%794=369%:%
%:%795=369%:%
%:%796=370%:%
%:%797=370%:%
%:%798=371%:%
%:%799=371%:%
%:%800=372%:%
%:%801=372%:%
%:%802=373%:%
%:%803=373%:%
%:%804=374%:%
%:%805=374%:%
%:%806=375%:%
%:%807=375%:%
%:%808=376%:%
%:%809=376%:%
%:%810=377%:%
%:%811=377%:%
%:%812=378%:%
%:%813=378%:%
%:%814=379%:%
%:%815=379%:%
%:%816=380%:%
%:%817=380%:%
%:%818=381%:%
%:%819=381%:%
%:%820=382%:%
%:%821=382%:%
%:%822=383%:%
%:%823=383%:%
%:%824=384%:%
%:%825=384%:%
%:%826=385%:%
%:%827=385%:%
%:%828=386%:%
%:%829=386%:%
%:%830=387%:%
%:%831=387%:%
%:%832=388%:%
%:%833=388%:%
%:%834=389%:%
%:%835=389%:%
%:%840=389%:%
%:%843=390%:%
%:%844=391%:%
%:%845=391%:%
%:%846=392%:%
%:%847=393%:%
%:%848=393%:%
%:%849=394%:%
%:%850=394%:%
%:%851=395%:%
%:%852=396%:%
%:%853=396%:%
%:%854=397%:%
%:%855=398%:%
%:%856=399%:%
%:%857=399%:%
%:%858=400%:%
%:%859=401%:%
%:%860=402%:%
%:%863=403%:%
%:%867=403%:%
%:%868=403%:%
%:%869=404%:%
%:%870=404%:%
%:%871=405%:%
%:%872=405%:%
%:%873=406%:%
%:%874=406%:%
%:%875=407%:%
%:%876=407%:%
%:%877=408%:%
%:%878=408%:%
%:%879=409%:%
%:%880=409%:%
%:%881=410%:%
%:%882=410%:%
%:%887=410%:%
%:%890=411%:%
%:%891=412%:%
%:%892=412%:%
%:%893=413%:%
%:%894=414%:%
%:%895=415%:%
%:%898=416%:%
%:%902=416%:%
%:%903=416%:%
%:%904=417%:%
%:%905=417%:%
%:%906=418%:%
%:%907=418%:%
%:%908=419%:%
%:%909=419%:%
%:%910=420%:%
%:%911=420%:%
%:%912=421%:%
%:%913=421%:%
%:%914=422%:%
%:%915=422%:%
%:%916=423%:%
%:%917=423%:%
%:%918=424%:%
%:%919=424%:%
%:%920=425%:%
%:%921=425%:%
%:%922=426%:%
%:%923=426%:%
%:%924=427%:%
%:%925=427%:%
%:%930=427%:%
%:%933=428%:%
%:%934=429%:%
%:%935=429%:%
%:%936=430%:%
%:%937=431%:%
%:%938=432%:%
%:%941=433%:%
%:%942=434%:%
%:%946=434%:%
%:%947=434%:%
%:%948=435%:%
%:%949=435%:%
%:%950=436%:%
%:%951=436%:%
%:%952=437%:%
%:%953=437%:%
%:%954=438%:%
%:%955=438%:%
%:%956=439%:%
%:%957=439%:%
%:%958=440%:%
%:%959=440%:%
%:%960=441%:%
%:%961=441%:%
%:%962=442%:%
%:%963=442%:%
%:%964=443%:%
%:%965=443%:%
%:%966=444%:%
%:%967=444%:%
%:%968=445%:%
%:%969=445%:%
%:%970=446%:%
%:%971=446%:%
%:%972=447%:%
%:%973=447%:%
%:%974=448%:%
%:%975=448%:%
%:%976=449%:%
%:%977=449%:%
%:%978=450%:%
%:%979=450%:%
%:%980=451%:%
%:%981=451%:%
%:%982=452%:%
%:%983=452%:%
%:%984=453%:%
%:%985=453%:%
%:%986=454%:%
%:%987=454%:%
%:%988=455%:%
%:%989=455%:%
%:%990=456%:%
%:%991=456%:%
%:%992=457%:%
%:%998=457%:%
%:%1001=458%:%
%:%1002=459%:%
%:%1003=459%:%
%:%1004=460%:%
%:%1005=461%:%
%:%1006=461%:%
%:%1007=462%:%
%:%1008=463%:%
%:%1009=463%:%
%:%1010=464%:%
%:%1011=465%:%
%:%1012=466%:%
%:%1013=466%:%
%:%1014=467%:%
%:%1015=468%:%
%:%1016=469%:%
%:%1019=470%:%
%:%1023=470%:%
%:%1024=470%:%
%:%1025=471%:%
%:%1026=471%:%
%:%1027=472%:%
%:%1028=472%:%
%:%1029=473%:%
%:%1030=473%:%
%:%1031=474%:%
%:%1032=474%:%
%:%1033=475%:%
%:%1034=475%:%
%:%1039=475%:%
%:%1042=476%:%
%:%1043=477%:%
%:%1044=477%:%
%:%1045=478%:%
%:%1046=479%:%
%:%1047=480%:%
%:%1050=481%:%
%:%1054=481%:%
%:%1055=481%:%
%:%1056=482%:%
%:%1057=482%:%
%:%1058=483%:%
%:%1059=483%:%
%:%1060=484%:%
%:%1061=484%:%
%:%1062=485%:%
%:%1063=485%:%
%:%1064=486%:%
%:%1065=486%:%
%:%1066=487%:%
%:%1067=487%:%
%:%1068=488%:%
%:%1069=488%:%
%:%1070=489%:%
%:%1071=489%:%
%:%1072=490%:%
%:%1073=490%:%
%:%1074=491%:%
%:%1075=491%:%
%:%1076=492%:%
%:%1077=492%:%
%:%1078=493%:%
%:%1079=493%:%
%:%1080=494%:%
%:%1081=494%:%
%:%1082=495%:%
%:%1083=495%:%
%:%1084=496%:%
%:%1085=496%:%
%:%1086=497%:%
%:%1087=497%:%
%:%1088=498%:%
%:%1089=498%:%
%:%1090=499%:%
%:%1091=499%:%
%:%1092=500%:%
%:%1093=500%:%
%:%1094=501%:%
%:%1095=501%:%
%:%1096=502%:%
%:%1097=502%:%
%:%1098=503%:%
%:%1099=503%:%
%:%1100=504%:%
%:%1101=504%:%
%:%1102=505%:%
%:%1103=505%:%
%:%1104=506%:%
%:%1105=506%:%
%:%1110=506%:%
%:%1113=507%:%
%:%1114=508%:%
%:%1115=508%:%
%:%1116=509%:%
%:%1117=510%:%
%:%1118=511%:%
%:%1119=511%:%
%:%1120=512%:%
%:%1121=513%:%
%:%1122=514%:%
%:%1125=515%:%
%:%1130=516%:%
%:%1131=516%:%
%:%1132=517%:%
%:%1133=517%:%
%:%1134=517%:%
%:%1135=517%:%
%:%1136=518%:%
%:%1137=518%:%
%:%1138=518%:%
%:%1139=518%:%
%:%1140=519%:%
%:%1141=520%:%
%:%1142=520%:%
%:%1143=521%:%
%:%1144=522%:%
%:%1145=522%:%
%:%1146=523%:%
%:%1147=523%:%
%:%1148=524%:%
%:%1149=524%:%
%:%1150=525%:%
%:%1151=525%:%
%:%1152=526%:%
%:%1153=526%:%
%:%1154=527%:%
%:%1155=527%:%
%:%1156=528%:%
%:%1157=528%:%
%:%1158=529%:%
%:%1159=529%:%
%:%1160=530%:%
%:%1161=530%:%
%:%1162=531%:%
%:%1163=532%:%
%:%1164=532%:%
%:%1165=532%:%
%:%1166=533%:%
%:%1167=533%:%
%:%1168=534%:%
%:%1169=534%:%
%:%1170=535%:%
%:%1171=535%:%
%:%1172=536%:%
%:%1173=536%:%
%:%1174=537%:%
%:%1175=537%:%
%:%1176=538%:%
%:%1177=539%:%
%:%1178=539%:%
%:%1179=539%:%
%:%1180=540%:%
%:%1181=540%:%
%:%1182=541%:%
%:%1183=541%:%
%:%1184=542%:%
%:%1185=542%:%
%:%1186=543%:%
%:%1187=543%:%
%:%1188=544%:%
%:%1189=544%:%
%:%1190=545%:%
%:%1191=545%:%
%:%1192=546%:%
%:%1193=546%:%
%:%1194=547%:%
%:%1195=547%:%
%:%1196=548%:%
%:%1197=549%:%
%:%1198=549%:%
%:%1199=549%:%
%:%1200=550%:%
%:%1201=550%:%
%:%1202=551%:%
%:%1203=551%:%
%:%1204=552%:%
%:%1205=553%:%
%:%1206=553%:%
%:%1207=553%:%
%:%1208=553%:%
%:%1209=554%:%
%:%1215=554%:%
%:%1218=555%:%
%:%1219=556%:%
%:%1220=556%:%
%:%1221=557%:%
%:%1222=558%:%
%:%1223=558%:%
%:%1224=559%:%
%:%1225=560%:%
%:%1226=561%:%
%:%1227=561%:%
%:%1228=562%:%
%:%1229=563%:%
%:%1230=564%:%
%:%1231=564%:%
%:%1232=565%:%
%:%1233=566%:%
%:%1234=567%:%
%:%1237=568%:%
%:%1241=568%:%
%:%1242=568%:%
%:%1243=569%:%
%:%1244=569%:%
%:%1245=570%:%
%:%1246=570%:%
%:%1247=571%:%
%:%1248=571%:%
%:%1249=572%:%
%:%1250=572%:%
%:%1255=572%:%
%:%1258=573%:%
%:%1259=574%:%
%:%1260=574%:%
%:%1261=575%:%
%:%1262=576%:%
%:%1263=577%:%
%:%1266=578%:%
%:%1270=578%:%
%:%1271=578%:%
%:%1272=579%:%
%:%1273=579%:%
%:%1274=580%:%
%:%1275=580%:%
%:%1276=581%:%
%:%1277=581%:%
%:%1278=582%:%
%:%1279=582%:%
%:%1280=583%:%
%:%1281=583%:%
%:%1282=584%:%
%:%1283=584%:%
%:%1284=585%:%
%:%1285=585%:%
%:%1286=586%:%
%:%1287=586%:%
%:%1288=587%:%
%:%1289=587%:%
%:%1290=588%:%
%:%1291=588%:%
%:%1292=589%:%
%:%1293=589%:%
%:%1294=590%:%
%:%1295=590%:%
%:%1296=591%:%
%:%1297=591%:%
%:%1298=592%:%
%:%1299=592%:%
%:%1300=593%:%
%:%1301=593%:%
%:%1302=594%:%
%:%1303=594%:%
%:%1304=595%:%
%:%1305=595%:%
%:%1306=596%:%
%:%1307=596%:%
%:%1308=597%:%
%:%1309=597%:%
%:%1310=598%:%
%:%1311=598%:%
%:%1312=599%:%
%:%1313=599%:%
%:%1314=600%:%
%:%1315=600%:%
%:%1316=601%:%
%:%1317=601%:%
%:%1318=602%:%
%:%1319=602%:%
%:%1320=603%:%
%:%1321=603%:%
%:%1322=604%:%
%:%1323=604%:%
%:%1324=605%:%
%:%1325=605%:%
%:%1330=605%:%
%:%1333=606%:%
%:%1334=607%:%
%:%1335=607%:%
%:%1336=608%:%
%:%1337=609%:%
%:%1338=610%:%
%:%1339=610%:%
%:%1340=611%:%
%:%1344=615%:%
%:%1351=616%:%
%:%1352=616%:%
%:%1353=617%:%
%:%1354=617%:%
%:%1355=618%:%
%:%1356=618%:%
%:%1357=619%:%
%:%1358=620%:%
%:%1359=621%:%
%:%1360=621%:%
%:%1363=624%:%
%:%1364=625%:%
%:%1365=625%:%
%:%1366=626%:%
%:%1367=626%:%
%:%1368=627%:%
%:%1369=627%:%
%:%1370=628%:%
%:%1371=628%:%
%:%1372=629%:%
%:%1373=629%:%
%:%1374=630%:%
%:%1375=630%:%
%:%1376=631%:%
%:%1377=631%:%
%:%1378=632%:%
%:%1379=632%:%
%:%1380=633%:%
%:%1381=633%:%
%:%1382=634%:%
%:%1383=634%:%
%:%1384=635%:%
%:%1385=635%:%
%:%1386=636%:%
%:%1387=636%:%
%:%1388=637%:%
%:%1389=637%:%
%:%1390=638%:%
%:%1391=638%:%
%:%1392=639%:%
%:%1393=639%:%
%:%1394=640%:%
%:%1395=640%:%
%:%1396=641%:%
%:%1397=641%:%
%:%1398=642%:%
%:%1399=642%:%
%:%1400=643%:%
%:%1401=643%:%
%:%1402=644%:%
%:%1403=644%:%
%:%1404=644%:%
%:%1405=645%:%
%:%1406=645%:%
%:%1407=646%:%
%:%1408=646%:%
%:%1409=647%:%
%:%1410=647%:%
%:%1411=648%:%
%:%1412=648%:%
%:%1413=648%:%
%:%1414=649%:%
%:%1415=649%:%
%:%1416=650%:%
%:%1417=650%:%
%:%1418=651%:%
%:%1419=651%:%
%:%1420=652%:%
%:%1421=652%:%
%:%1422=653%:%
%:%1423=653%:%
%:%1424=653%:%
%:%1425=654%:%
%:%1426=654%:%
%:%1427=655%:%
%:%1428=655%:%
%:%1429=656%:%
%:%1430=656%:%
%:%1431=657%:%
%:%1437=657%:%
%:%1440=658%:%
%:%1441=659%:%
%:%1442=659%:%
%:%1443=660%:%
%:%1444=661%:%
%:%1445=661%:%
%:%1446=662%:%
%:%1447=663%:%
%:%1448=664%:%
%:%1451=665%:%
%:%1456=666%:%
%:%1457=666%:%
%:%1458=667%:%
%:%1459=667%:%
%:%1460=668%:%
%:%1461=668%:%
%:%1462=669%:%
%:%1463=669%:%
%:%1464=670%:%
%:%1465=670%:%
%:%1466=671%:%
%:%1467=672%:%
%:%1468=672%:%
%:%1469=673%:%
%:%1470=674%:%
%:%1471=674%:%
%:%1472=674%:%
%:%1473=675%:%
%:%1474=676%:%
%:%1475=676%:%
%:%1476=677%:%
%:%1477=677%:%
%:%1478=678%:%
%:%1479=678%:%
%:%1480=679%:%
%:%1481=679%:%
%:%1482=680%:%
%:%1483=680%:%
%:%1484=681%:%
%:%1485=681%:%
%:%1486=681%:%
%:%1487=682%:%
%:%1488=682%:%
%:%1489=683%:%
%:%1490=683%:%
%:%1491=684%:%
%:%1492=684%:%
%:%1493=685%:%
%:%1494=685%:%
%:%1495=685%:%
%:%1496=686%:%
%:%1497=686%:%
%:%1498=687%:%
%:%1499=687%:%
%:%1500=688%:%
%:%1501=688%:%
%:%1502=689%:%
%:%1503=689%:%
%:%1504=690%:%
%:%1505=690%:%
%:%1506=691%:%
%:%1507=691%:%
%:%1508=692%:%
%:%1509=692%:%
%:%1510=693%:%
%:%1511=693%:%
%:%1512=693%:%
%:%1513=694%:%
%:%1514=694%:%
%:%1515=695%:%
%:%1516=695%:%
%:%1517=696%:%
%:%1518=696%:%
%:%1519=696%:%
%:%1520=697%:%
%:%1521=697%:%
%:%1522=698%:%
%:%1523=698%:%
%:%1524=699%:%
%:%1525=699%:%
%:%1526=700%:%
%:%1527=700%:%
%:%1528=700%:%
%:%1529=701%:%
%:%1530=701%:%
%:%1531=702%:%
%:%1532=702%:%
%:%1533=703%:%
%:%1534=703%:%
%:%1535=704%:%
%:%1536=704%:%
%:%1537=704%:%
%:%1538=704%:%
%:%1539=705%:%
%:%1540=705%:%
%:%1541=706%:%
%:%1542=706%:%
%:%1543=706%:%
%:%1544=706%:%
%:%1545=706%:%
%:%1546=707%:%
%:%1552=707%:%
%:%1555=708%:%
%:%1556=709%:%
%:%1557=709%:%
%:%1558=710%:%
%:%1559=711%:%
%:%1560=712%:%
%:%1563=713%:%
%:%1568=714%:%
%:%1569=714%:%
%:%1570=715%:%
%:%1571=715%:%
%:%1572=715%:%
%:%1573=715%:%
%:%1574=716%:%
%:%1575=717%:%
%:%1576=717%:%
%:%1577=718%:%
%:%1578=718%:%
%:%1579=719%:%
%:%1580=719%:%
%:%1581=720%:%
%:%1582=720%:%
%:%1583=721%:%
%:%1584=721%:%
%:%1585=722%:%
%:%1586=722%:%
%:%1587=723%:%
%:%1588=723%:%
%:%1589=724%:%
%:%1590=724%:%
%:%1591=725%:%
%:%1592=725%:%
%:%1593=726%:%
%:%1594=726%:%
%:%1595=727%:%
%:%1596=727%:%
%:%1597=728%:%
%:%1598=728%:%
%:%1599=729%:%
%:%1600=729%:%
%:%1601=730%:%
%:%1602=730%:%
%:%1603=731%:%
%:%1604=731%:%
%:%1605=732%:%
%:%1606=732%:%
%:%1607=733%:%
%:%1608=733%:%
%:%1609=734%:%
%:%1610=734%:%
%:%1611=735%:%
%:%1612=735%:%
%:%1613=736%:%
%:%1614=736%:%
%:%1615=737%:%
%:%1616=737%:%
%:%1617=738%:%
%:%1618=738%:%
%:%1619=739%:%
%:%1620=739%:%
%:%1621=739%:%
%:%1622=739%:%
%:%1623=739%:%
%:%1624=740%:%
%:%1625=740%:%
%:%1626=740%:%
%:%1627=740%:%
%:%1628=740%:%
%:%1629=741%:%
%:%1630=741%:%
%:%1631=741%:%
%:%1632=741%:%
%:%1633=742%:%
%:%1639=742%:%
%:%1642=743%:%
%:%1643=744%:%
%:%1644=744%:%
%:%1645=745%:%
%:%1646=746%:%
%:%1647=747%:%
%:%1654=748%:%
%:%1655=748%:%
%:%1656=749%:%
%:%1657=749%:%
%:%1658=750%:%
%:%1659=750%:%
%:%1660=751%:%
%:%1661=751%:%
%:%1662=752%:%
%:%1663=752%:%
%:%1664=753%:%
%:%1665=753%:%
%:%1666=754%:%
%:%1667=754%:%
%:%1668=755%:%
%:%1669=755%:%
%:%1670=756%:%
%:%1671=756%:%
%:%1672=757%:%
%:%1673=758%:%
%:%1674=758%:%
%:%1675=759%:%
%:%1676=759%:%
%:%1677=760%:%
%:%1678=760%:%
%:%1679=761%:%
%:%1680=761%:%
%:%1681=762%:%
%:%1682=762%:%
%:%1683=763%:%
%:%1684=763%:%
%:%1685=764%:%
%:%1686=765%:%
%:%1687=765%:%
%:%1688=766%:%
%:%1689=766%:%
%:%1690=767%:%
%:%1691=767%:%
%:%1692=768%:%
%:%1693=768%:%
%:%1694=769%:%
%:%1695=769%:%
%:%1696=770%:%
%:%1697=770%:%
%:%1698=771%:%
%:%1699=771%:%
%:%1700=772%:%
%:%1701=773%:%
%:%1702=773%:%
%:%1703=773%:%
%:%1704=773%:%
%:%1705=774%:%
%:%1706=775%:%
%:%1707=775%:%
%:%1708=775%:%
%:%1709=775%:%
%:%1710=775%:%
%:%1711=776%:%
%:%1717=776%:%
%:%1720=777%:%
%:%1721=778%:%
%:%1722=778%:%
%:%1729=779%:%