### 2024-11-01 code/output/document/HS_Theorems.tex

%
\begin{isabellebody}%
\setisabellecontext{HS{\isacharunderscore}{\kern0pt}Theorems}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}Theorems\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\ \isanewline
\ \ \ \ HS{\isacharunderscore}{\kern0pt}M\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\isakeyword{begin}\isanewline
\ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M\ {\isasymlongrightarrow}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M\ {\isasymlongrightarrow}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isakeyword{in}\ eps{\isacharunderscore}{\kern0pt}induct{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}y\ {\isasymin}\ x{\isachardot}{\kern0pt}\ y\ {\isasymin}\ M\ {\isasymlongrightarrow}\ check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{thm}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}{\isasympi}{\isachardot}{\kern0pt}\ {\isasympi}\ {\isasymin}\ {\isasymG}\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ check{\isacharunderscore}{\kern0pt}fixpoint{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ assms\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sym{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymG}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}fixpoint\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ \ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}symmetric{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ symmetric{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}{\isasymF}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ assm{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ vpH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ def{\isacharunderscore}{\kern0pt}check{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ vpH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ transM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ H{\isadigit{1}}\ H{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}x\ {\isasymin}\ M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ comp{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}sym\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isasympi}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymtau}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasympi}\ O\ {\isasymtau}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}sym\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isasympi}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}function\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ assms\ sym{\isacharunderscore}{\kern0pt}def\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}function{\isacharparenright}{\kern0pt}\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ {\isasymG}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymG}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroup\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharparenleft}{\kern0pt}{\isasymG}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{thm}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isacharbraceleft}{\kern0pt}\ {\isasympi}\ {\isasymin}\ P\ {\isasymrightarrow}\ P{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharparenleft}{\kern0pt}{\isasymG}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}sym{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ sym{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ comp{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}sym\ assm\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}sym\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}helper\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ {\isasymG}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ HS\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ {\isasymG}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ HS\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ HS\ {\isasymlongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ pi{\isacharunderscore}{\kern0pt}pauto\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ pi{\isacharunderscore}{\kern0pt}pauto\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ eq{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ domsubset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ qH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ q{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y{\isachardot}{\kern0pt}\ {\isasymexists}p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ {\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ q{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ eq\ qH\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ yH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yH\ assms{\isadigit{1}}\ ed{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ {\isasympi}\ O\ {\isasymtau}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymtau}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasymin}\ {\isasymF}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ symmetric{\isacharunderscore}{\kern0pt}def\ {\isasymF}{\isacharunderscore}{\kern0pt}normal\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ {\isasympi}\ O\ {\isasymtau}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymtau}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ sym{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ {\isasymsigma}\ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isasymtau}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ sigmain\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ {\isasymG}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isasymF}{\isacharunderscore}{\kern0pt}subset\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharunderscore}{\kern0pt}def\ H{\isadigit{1}}\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtau}\ \isakeyword{where}\ tauH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isacharequal}{\kern0pt}\ {\isasympi}\ O\ {\isasymtau}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}P{\isacharunderscore}{\kern0pt}names\ \isakeyword{in}\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ pi{\isacharunderscore}{\kern0pt}pauto\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}\ O\ converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ {\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ pi{\isacharunderscore}{\kern0pt}pauto\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ tauH\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ pi{\isacharunderscore}{\kern0pt}pauto\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}\ O\ {\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ {\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ comp{\isacharunderscore}{\kern0pt}assoc\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isacharparenleft}{\kern0pt}{\isasymtau}\ O\ id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ left{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharcomma}{\kern0pt}\ rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ pi{\isacharunderscore}{\kern0pt}pauto\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}\ O\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ right{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}id{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ tauH\ sym{\isacharunderscore}{\kern0pt}def\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}comp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ pi{\isacharunderscore}{\kern0pt}pauto\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ tauH\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ tauH\ sym{\isacharunderscore}{\kern0pt}def\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ tauH\ sym{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ sigmain\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}sym{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroups{\isacharparenleft}{\kern0pt}{\isasymG}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sym{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroup{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}symmetric{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ symmetric{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H{\isadigit{1}}\ H{\isadigit{2}}\ {\isasymF}{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}supergroup\ sym{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroup\ assms{\isadigit{1}}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ HS{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ HS{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ HS{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ domsubset{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ {\isasymG}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ x\ {\isasymin}\ HS\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ {\isasympi}\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ {\isasymG}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}helper\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ {\isasympi}\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ {\isasymG}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}helper{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sym{\isacharcomma}{\kern0pt}\ rule\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ \ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ left{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}bij{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}helper\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ x\ y\ {\isasympi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ pH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pair{\isacharunderscore}{\kern0pt}relI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ x\ y\ {\isasympi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}helper{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}helper{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\ converse{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}sym\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ sym{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sym{\isacharcomma}{\kern0pt}\ rule\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ sym{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ left{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}bij{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id\ sym{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}id{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/HS_Theorems.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=10%:%
%:%36=11%:%
%:%37=11%:%
%:%38=12%:%
%:%39=12%:%
%:%40=13%:%
%:%41=13%:%
%:%42=14%:%
%:%43=14%:%
%:%44=15%:%
%:%45=15%:%
%:%46=15%:%
%:%47=16%:%
%:%48=16%:%
%:%49=17%:%
%:%50=17%:%
%:%51=17%:%
%:%52=17%:%
%:%53=18%:%
%:%54=18%:%
%:%55=19%:%
%:%56=19%:%
%:%57=20%:%
%:%58=20%:%
%:%59=21%:%
%:%60=21%:%
%:%61=22%:%
%:%62=22%:%
%:%63=23%:%
%:%64=23%:%
%:%65=23%:%
%:%66=24%:%
%:%67=24%:%
%:%68=25%:%
%:%69=25%:%
%:%70=26%:%
%:%71=26%:%
%:%72=27%:%
%:%73=27%:%
%:%74=27%:%
%:%75=28%:%
%:%76=28%:%
%:%77=29%:%
%:%78=29%:%
%:%79=30%:%
%:%80=30%:%
%:%81=31%:%
%:%82=31%:%
%:%83=32%:%
%:%84=32%:%
%:%85=33%:%
%:%86=33%:%
%:%87=33%:%
%:%88=34%:%
%:%89=34%:%
%:%90=34%:%
%:%91=34%:%
%:%92=35%:%
%:%93=35%:%
%:%94=35%:%
%:%95=36%:%
%:%96=36%:%
%:%97=36%:%
%:%98=36%:%
%:%99=36%:%
%:%100=37%:%
%:%101=37%:%
%:%102=37%:%
%:%103=37%:%
%:%104=37%:%
%:%105=38%:%
%:%106=38%:%
%:%107=39%:%
%:%108=40%:%
%:%109=40%:%
%:%110=40%:%
%:%111=40%:%
%:%112=40%:%
%:%113=41%:%
%:%114=41%:%
%:%115=42%:%
%:%116=42%:%
%:%117=42%:%
%:%118=42%:%
%:%119=42%:%
%:%120=43%:%
%:%126=43%:%
%:%129=44%:%
%:%130=45%:%
%:%131=45%:%
%:%134=46%:%
%:%138=46%:%
%:%139=46%:%
%:%140=47%:%
%:%141=47%:%
%:%142=48%:%
%:%143=48%:%
%:%144=49%:%
%:%145=49%:%
%:%146=50%:%
%:%147=50%:%
%:%148=51%:%
%:%149=51%:%
%:%150=52%:%
%:%151=52%:%
%:%152=53%:%
%:%153=53%:%
%:%154=54%:%
%:%155=54%:%
%:%156=55%:%
%:%157=55%:%
%:%158=56%:%
%:%159=56%:%
%:%160=57%:%
%:%161=57%:%
%:%162=58%:%
%:%163=58%:%
%:%168=58%:%
%:%171=59%:%
%:%172=60%:%
%:%173=60%:%
%:%180=61%:%
%:%181=61%:%
%:%182=62%:%
%:%183=62%:%
%:%184=63%:%
%:%185=64%:%
%:%186=64%:%
%:%187=65%:%
%:%188=65%:%
%:%189=66%:%
%:%190=66%:%
%:%191=67%:%
%:%192=67%:%
%:%193=68%:%
%:%194=68%:%
%:%195=68%:%
%:%196=69%:%
%:%197=69%:%
%:%198=70%:%
%:%199=70%:%
%:%200=71%:%
%:%201=71%:%
%:%202=72%:%
%:%203=72%:%
%:%204=72%:%
%:%205=72%:%
%:%206=73%:%
%:%207=73%:%
%:%208=73%:%
%:%209=74%:%
%:%210=74%:%
%:%211=75%:%
%:%212=75%:%
%:%213=76%:%
%:%214=76%:%
%:%215=76%:%
%:%216=77%:%
%:%217=77%:%
%:%218=78%:%
%:%219=78%:%
%:%220=79%:%
%:%221=80%:%
%:%222=80%:%
%:%223=81%:%
%:%224=81%:%
%:%225=82%:%
%:%226=82%:%
%:%227=83%:%
%:%228=83%:%
%:%229=84%:%
%:%230=84%:%
%:%231=84%:%
%:%232=85%:%
%:%233=85%:%
%:%234=86%:%
%:%235=86%:%
%:%236=87%:%
%:%237=87%:%
%:%238=88%:%
%:%239=89%:%
%:%240=89%:%
%:%241=89%:%
%:%242=90%:%
%:%243=90%:%
%:%244=91%:%
%:%245=91%:%
%:%246=92%:%
%:%247=92%:%
%:%248=93%:%
%:%254=93%:%
%:%257=94%:%
%:%258=95%:%
%:%259=95%:%
%:%266=96%:%
%:%267=96%:%
%:%268=97%:%
%:%269=97%:%
%:%270=98%:%
%:%271=98%:%
%:%272=99%:%
%:%273=100%:%
%:%274=100%:%
%:%275=101%:%
%:%276=101%:%
%:%277=102%:%
%:%278=102%:%
%:%279=103%:%
%:%280=103%:%
%:%281=104%:%
%:%282=105%:%
%:%283=105%:%
%:%284=105%:%
%:%285=106%:%
%:%286=106%:%
%:%287=107%:%
%:%288=107%:%
%:%289=108%:%
%:%290=108%:%
%:%291=109%:%
%:%292=109%:%
%:%293=110%:%
%:%294=110%:%
%:%295=111%:%
%:%296=111%:%
%:%297=112%:%
%:%298=112%:%
%:%299=113%:%
%:%305=113%:%
%:%308=114%:%
%:%309=115%:%
%:%310=115%:%
%:%317=116%:%
%:%318=116%:%
%:%319=117%:%
%:%320=117%:%
%:%321=118%:%
%:%322=119%:%
%:%323=119%:%
%:%324=120%:%
%:%325=120%:%
%:%326=121%:%
%:%327=121%:%
%:%328=121%:%
%:%329=122%:%
%:%330=123%:%
%:%331=123%:%
%:%332=124%:%
%:%333=124%:%
%:%334=124%:%
%:%335=125%:%
%:%336=126%:%
%:%337=126%:%
%:%338=126%:%
%:%339=126%:%
%:%340=127%:%
%:%341=128%:%
%:%342=128%:%
%:%343=128%:%
%:%344=128%:%
%:%345=129%:%
%:%346=129%:%
%:%347=130%:%
%:%348=130%:%
%:%349=131%:%
%:%350=131%:%
%:%351=131%:%
%:%352=132%:%
%:%353=132%:%
%:%354=132%:%
%:%355=132%:%
%:%356=133%:%
%:%357=133%:%
%:%358=134%:%
%:%359=134%:%
%:%360=135%:%
%:%361=135%:%
%:%362=136%:%
%:%363=136%:%
%:%364=137%:%
%:%365=137%:%
%:%366=137%:%
%:%367=137%:%
%:%368=138%:%
%:%369=138%:%
%:%370=138%:%
%:%371=138%:%
%:%372=138%:%
%:%373=139%:%
%:%374=139%:%
%:%375=139%:%
%:%376=139%:%
%:%377=139%:%
%:%378=140%:%
%:%379=140%:%
%:%380=141%:%
%:%381=142%:%
%:%382=142%:%
%:%383=142%:%
%:%384=143%:%
%:%385=143%:%
%:%386=143%:%
%:%387=144%:%
%:%388=145%:%
%:%389=145%:%
%:%390=146%:%
%:%391=146%:%
%:%392=147%:%
%:%393=147%:%
%:%394=147%:%
%:%395=148%:%
%:%396=149%:%
%:%397=149%:%
%:%398=149%:%
%:%399=149%:%
%:%400=150%:%
%:%401=151%:%
%:%402=151%:%
%:%403=151%:%
%:%404=151%:%
%:%405=152%:%
%:%406=153%:%
%:%407=153%:%
%:%408=153%:%
%:%409=153%:%
%:%410=154%:%
%:%411=154%:%
%:%412=154%:%
%:%413=155%:%
%:%414=155%:%
%:%415=156%:%
%:%416=156%:%
%:%417=157%:%
%:%418=157%:%
%:%419=158%:%
%:%420=158%:%
%:%421=158%:%
%:%422=159%:%
%:%423=159%:%
%:%424=160%:%
%:%425=160%:%
%:%426=161%:%
%:%427=161%:%
%:%428=162%:%
%:%429=162%:%
%:%430=163%:%
%:%431=163%:%
%:%432=164%:%
%:%433=164%:%
%:%434=165%:%
%:%435=165%:%
%:%436=166%:%
%:%437=166%:%
%:%438=167%:%
%:%439=167%:%
%:%440=168%:%
%:%441=168%:%
%:%442=169%:%
%:%443=169%:%
%:%444=170%:%
%:%445=170%:%
%:%446=171%:%
%:%447=171%:%
%:%448=171%:%
%:%449=172%:%
%:%450=172%:%
%:%451=172%:%
%:%452=173%:%
%:%453=173%:%
%:%454=173%:%
%:%455=174%:%
%:%456=174%:%
%:%457=175%:%
%:%458=175%:%
%:%459=176%:%
%:%460=176%:%
%:%461=177%:%
%:%462=177%:%
%:%463=177%:%
%:%464=178%:%
%:%465=178%:%
%:%466=179%:%
%:%467=179%:%
%:%468=180%:%
%:%469=180%:%
%:%470=181%:%
%:%471=181%:%
%:%472=181%:%
%:%473=182%:%
%:%474=182%:%
%:%475=183%:%
%:%476=183%:%
%:%477=184%:%
%:%478=184%:%
%:%479=185%:%
%:%480=185%:%
%:%481=186%:%
%:%482=186%:%
%:%483=187%:%
%:%484=187%:%
%:%485=188%:%
%:%486=188%:%
%:%487=188%:%
%:%488=189%:%
%:%489=189%:%
%:%490=190%:%
%:%491=190%:%
%:%492=191%:%
%:%493=191%:%
%:%494=192%:%
%:%495=192%:%
%:%496=193%:%
%:%497=193%:%
%:%498=193%:%
%:%499=194%:%
%:%500=194%:%
%:%501=194%:%
%:%502=195%:%
%:%503=195%:%
%:%504=195%:%
%:%505=196%:%
%:%506=196%:%
%:%507=197%:%
%:%508=197%:%
%:%509=198%:%
%:%510=198%:%
%:%511=199%:%
%:%512=199%:%
%:%513=200%:%
%:%514=201%:%
%:%515=201%:%
%:%516=202%:%
%:%517=202%:%
%:%518=203%:%
%:%519=203%:%
%:%520=204%:%
%:%521=204%:%
%:%522=205%:%
%:%523=205%:%
%:%524=206%:%
%:%525=207%:%
%:%526=207%:%
%:%527=207%:%
%:%528=208%:%
%:%529=208%:%
%:%530=209%:%
%:%531=209%:%
%:%532=210%:%
%:%533=210%:%
%:%534=211%:%
%:%535=212%:%
%:%536=212%:%
%:%537=212%:%
%:%538=213%:%
%:%539=213%:%
%:%540=214%:%
%:%541=214%:%
%:%542=215%:%
%:%543=215%:%
%:%544=216%:%
%:%545=216%:%
%:%546=217%:%
%:%547=217%:%
%:%548=218%:%
%:%549=218%:%
%:%550=219%:%
%:%551=219%:%
%:%552=220%:%
%:%553=220%:%
%:%554=221%:%
%:%555=221%:%
%:%556=222%:%
%:%557=223%:%
%:%558=223%:%
%:%559=223%:%
%:%560=224%:%
%:%561=224%:%
%:%562=225%:%
%:%563=225%:%
%:%564=226%:%
%:%570=226%:%
%:%573=227%:%
%:%574=228%:%
%:%575=228%:%
%:%582=229%:%
%:%583=229%:%
%:%584=230%:%
%:%585=230%:%
%:%586=230%:%
%:%587=231%:%
%:%588=231%:%
%:%589=231%:%
%:%590=232%:%
%:%591=232%:%
%:%592=232%:%
%:%593=233%:%
%:%594=233%:%
%:%595=234%:%
%:%596=234%:%
%:%597=234%:%
%:%598=235%:%
%:%599=235%:%
%:%600=236%:%
%:%601=236%:%
%:%602=237%:%
%:%603=237%:%
%:%604=238%:%
%:%605=238%:%
%:%606=239%:%
%:%607=239%:%
%:%608=240%:%
%:%609=240%:%
%:%610=241%:%
%:%611=241%:%
%:%612=242%:%
%:%613=242%:%
%:%614=243%:%
%:%615=243%:%
%:%616=243%:%
%:%617=244%:%
%:%618=244%:%
%:%619=245%:%
%:%620=245%:%
%:%621=246%:%
%:%622=246%:%
%:%623=247%:%
%:%624=247%:%
%:%625=248%:%
%:%626=248%:%
%:%627=248%:%
%:%628=249%:%
%:%629=249%:%
%:%630=250%:%
%:%631=250%:%
%:%632=251%:%
%:%633=251%:%
%:%634=252%:%
%:%635=252%:%
%:%636=253%:%
%:%637=253%:%
%:%638=254%:%
%:%639=254%:%
%:%640=254%:%
%:%641=255%:%
%:%642=255%:%
%:%643=256%:%
%:%644=256%:%
%:%645=256%:%
%:%646=257%:%
%:%647=257%:%
%:%648=258%:%
%:%649=258%:%
%:%650=259%:%
%:%651=259%:%
%:%652=259%:%
%:%653=259%:%
%:%654=259%:%
%:%655=260%:%
%:%661=260%:%
%:%664=261%:%
%:%665=262%:%
%:%666=262%:%
%:%667=263%:%
%:%668=264%:%
%:%669=265%:%
%:%676=266%:%
%:%677=266%:%
%:%678=267%:%
%:%679=267%:%
%:%680=267%:%
%:%681=267%:%
%:%682=268%:%
%:%683=268%:%
%:%684=269%:%
%:%685=269%:%
%:%686=270%:%
%:%687=270%:%
%:%688=271%:%
%:%689=271%:%
%:%690=272%:%
%:%691=272%:%
%:%692=273%:%
%:%693=273%:%
%:%694=274%:%
%:%695=274%:%
%:%696=275%:%
%:%697=275%:%
%:%698=275%:%
%:%699=276%:%
%:%700=276%:%
%:%701=277%:%
%:%702=277%:%
%:%703=278%:%
%:%704=278%:%
%:%705=278%:%
%:%706=278%:%
%:%707=279%:%
%:%713=279%:%
%:%716=280%:%
%:%717=281%:%
%:%718=281%:%
%:%719=282%:%
%:%720=283%:%
%:%721=284%:%
%:%728=285%:%
%:%729=285%:%
%:%730=286%:%
%:%731=286%:%
%:%732=287%:%
%:%733=287%:%
%:%734=287%:%
%:%735=287%:%
%:%736=288%:%
%:%737=288%:%
%:%738=288%:%
%:%739=289%:%
%:%740=289%:%
%:%741=290%:%
%:%742=290%:%
%:%743=291%:%
%:%744=291%:%
%:%745=292%:%
%:%746=292%:%
%:%747=293%:%
%:%748=293%:%
%:%749=294%:%
%:%750=295%:%
%:%751=295%:%
%:%752=296%:%
%:%753=296%:%
%:%754=297%:%
%:%755=297%:%
%:%756=298%:%
%:%757=298%:%
%:%758=299%:%
%:%759=299%:%
%:%760=300%:%
%:%761=300%:%
%:%762=301%:%
%:%763=301%:%
%:%764=302%:%
%:%765=302%:%
%:%766=303%:%
%:%767=303%:%
%:%768=303%:%
%:%769=304%:%
%:%770=304%:%
%:%771=305%:%
%:%772=305%:%
%:%773=306%:%
%:%774=306%:%
%:%775=307%:%
%:%776=307%:%
%:%777=308%:%
%:%778=308%:%
%:%779=308%:%
%:%780=309%:%
%:%781=309%:%
%:%782=310%:%
%:%783=310%:%
%:%784=311%:%
%:%785=311%:%
%:%786=312%:%
%:%787=312%:%
%:%788=313%:%
%:%789=313%:%
%:%790=314%:%
%:%791=314%:%
%:%792=314%:%
%:%793=315%:%
%:%794=315%:%
%:%795=316%:%
%:%796=316%:%
%:%797=316%:%
%:%798=317%:%
%:%799=317%:%
%:%800=318%:%
%:%801=318%:%
%:%802=319%:%
%:%803=319%:%
%:%804=319%:%
%:%805=320%:%
%:%806=320%:%
%:%807=320%:%
%:%808=321%:%
%:%814=321%:%
%:%817=322%:%
%:%818=323%:%
%:%819=323%:%
%:%826=324%:%