### 2024-11-01 code/output/document/Infinity_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Infinity{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Axiom of Infinity in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Infinity{\isacharunderscore}{\kern0pt}Axiom\isanewline
\ \ \isakeyword{imports}\ Pairing{\isacharunderscore}{\kern0pt}Axiom\ Union{\isacharunderscore}{\kern0pt}Axiom\ Separation{\isacharunderscore}{\kern0pt}Axiom\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\ \isakeyword{begin}\isanewline
\isanewline
\isacommand{interpretation}\isamarkupfalse%
\ mg{\isacharunderscore}{\kern0pt}triv{\isacharcolon}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}trivial{\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ generic\ Union{\isacharunderscore}{\kern0pt}MG\ pairing{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ unfold{\isacharunderscore}{\kern0pt}locales\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ infinity{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}infinity{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ infinity{\isacharunderscore}{\kern0pt}ax\ \isacommand{obtain}\isamarkupfalse%
\ I\ \isakeyword{where}\isanewline
\ \ \ \ Eq{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}I{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ I{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}y{\isasymin}M{\isachardot}{\kern0pt}\ y\ {\isasymin}\ I\ {\isasymlongrightarrow}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ I{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ infinity{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}I{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}I{\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ valcheck\ generic\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ GenExtI{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}I{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ G{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isadigit{0}}{\isasymin}I{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}I{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}y\ {\isasymin}\ I{\isachardoublequoteclose}\ \isakeyword{for}\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}I{\isasymin}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}I{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ I\ {\isasyminter}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}y\ {\isasymin}\ I{\isachardoublequoteclose}\ \isakeyword{for}\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ Eq{\isadigit{1}}\ transitivity{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ Eq{\isadigit{1}}\ {\isacartoucheopen}I{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isadigit{0}}{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ infinity{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Infinity_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%35=4%:%
%:%38=5%:%
%:%39=6%:%
%:%40=6%:%
%:%41=7%:%
%:%42=8%:%
%:%43=8%:%
%:%46=9%:%
%:%50=9%:%
%:%51=9%:%
%:%52=10%:%
%:%53=10%:%
%:%58=10%:%
%:%61=11%:%
%:%62=12%:%
%:%63=12%:%
%:%70=13%:%
%:%71=13%:%
%:%72=14%:%
%:%73=14%:%
%:%74=14%:%
%:%75=15%:%
%:%76=16%:%
%:%77=16%:%
%:%78=16%:%
%:%79=17%:%
%:%80=17%:%
%:%81=18%:%
%:%82=18%:%
%:%83=19%:%
%:%84=19%:%
%:%85=19%:%
%:%86=20%:%
%:%87=20%:%
%:%88=21%:%
%:%89=21%:%
%:%90=22%:%
%:%91=22%:%
%:%92=22%:%
%:%93=23%:%
%:%94=23%:%
%:%95=24%:%
%:%96=24%:%
%:%97=24%:%
%:%98=24%:%
%:%99=25%:%
%:%100=25%:%
%:%101=26%:%
%:%102=26%:%
%:%103=27%:%
%:%104=27%:%
%:%105=27%:%
%:%106=28%:%
%:%107=28%:%
%:%108=29%:%
%:%109=29%:%
%:%110=30%:%
%:%111=30%:%
%:%112=30%:%
%:%113=31%:%
%:%114=31%:%
%:%115=32%:%
%:%116=32%:%
%:%117=33%:%
%:%118=33%:%
%:%119=33%:%
%:%120=34%:%
%:%126=34%:%
%:%129=35%:%
%:%130=36%:%
%:%131=36%:%
%:%138=37%:%