### 2024-11-01 code/output/document/Internal_ZFC_Axioms.tex

%
\begin{isabellebody}%
\setisabellecontext{Internal{\isacharunderscore}{\kern0pt}ZFC{\isacharunderscore}{\kern0pt}Axioms}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The ZFC axioms, internalized%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Internal{\isacharunderscore}{\kern0pt}ZFC{\isacharunderscore}{\kern0pt}Axioms\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ Forcing{\isacharunderscore}{\kern0pt}Data\isanewline
\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}zfunion{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ ZF{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}power{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}power{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}zfpow{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ power{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ powerset{\isacharunderscore}{\kern0pt}def\ subset{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}power{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ ZF{\isacharunderscore}{\kern0pt}power{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}pairing{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}upair{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}zfpair{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}pairing{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ ZF{\isacharunderscore}{\kern0pt}pairing{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}foundation{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}foundation{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}zfpow{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ foundation{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}foundation{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ ZF{\isacharunderscore}{\kern0pt}foundation{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}extensionality{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}extensionality{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}zfpow{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ extensionality{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}extensionality{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ ZF{\isacharunderscore}{\kern0pt}extensionality{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}infinity{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}infinity{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharcomma}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ infinity{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}infinity{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ ZF{\isacharunderscore}{\kern0pt}infinity{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}choice{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharcomma}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ choice{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{syntax}\isamarkupfalse%
\isanewline
\ \ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}choice{\isachardoublequoteclose}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}{\isachardoublequoteopen}AC{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\isacommand{translations}\isamarkupfalse%
\isanewline
\ \ {\isachardoublequoteopen}AC{\isachardoublequoteclose}\ {\isasymrightharpoonup}\ {\isachardoublequoteopen}CONST\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ ZFC{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}defs\ {\isacharequal}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}extensionality{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}foundation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}pairing{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ ZF{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}infinity{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}power{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ ZFC{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats\ {\isacharequal}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}extensionality{\isacharunderscore}{\kern0pt}auto\ ZF{\isacharunderscore}{\kern0pt}foundation{\isacharunderscore}{\kern0pt}auto\ ZF{\isacharunderscore}{\kern0pt}pairing{\isacharunderscore}{\kern0pt}auto\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ ZF{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}auto\ ZF{\isacharunderscore}{\kern0pt}infinity{\isacharunderscore}{\kern0pt}auto\ ZF{\isacharunderscore}{\kern0pt}power{\isacharunderscore}{\kern0pt}auto\ ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}auto\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ZF{\isacharunderscore}{\kern0pt}fin\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}fin\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}extensionality{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}foundation{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}pairing{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ ZF{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}infinity{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}power{\isacharunderscore}{\kern0pt}fm\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ZFC{\isacharunderscore}{\kern0pt}fin\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ZFC{\isacharunderscore}{\kern0pt}fin\ {\isasymequiv}\ ZF{\isacharunderscore}{\kern0pt}fin\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}choice{\isacharunderscore}{\kern0pt}fm{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ZFC{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}ZFC{\isacharunderscore}{\kern0pt}fin\ {\isasymsubseteq}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ZFC{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}def\ ZFC{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}defs\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{The Axiom of Separation, internalized%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{lemma}\isamarkupfalse%
\ iterates{\isacharunderscore}{\kern0pt}Forall{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ n\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ p\ {\isasymin}\ formula\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ Forall{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ set{\isacharcolon}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ last{\isacharunderscore}{\kern0pt}init{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}l\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}\ a{\isasymin}A{\isachardot}{\kern0pt}\ {\isasymexists}l{\isacharprime}{\kern0pt}{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ l\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}l{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rev{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}rev{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ a\ l{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}a{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}l{\isacharprime}{\kern0pt}{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rev{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharsemicolon}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}l\ {\isacharequal}{\kern0pt}\ rev{\isacharparenleft}{\kern0pt}l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rev{\isacharparenleft}{\kern0pt}l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ rev{\isacharunderscore}{\kern0pt}rev{\isacharunderscore}{\kern0pt}ident{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}l{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ take{\isacharunderscore}{\kern0pt}drop{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}l{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymAnd}\ n\ {\isachardot}{\kern0pt}\ n\ {\isacharless}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ l\ {\isacharequal}{\kern0pt}\ take{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isacharat}{\kern0pt}\ drop{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}l{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\isacommand{proof}\isamarkupfalse%
\ induct\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ Nil\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Cons\ a\ l{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ i\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}i{\isacharless}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}l{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}lt{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}{\isasymexists}k{\isasymin}nat{\isachardot}{\kern0pt}\ i\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}k{\isacharparenright}{\kern0pt}\ {\isasymand}\ k\ {\isacharless}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}l{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \ le{\isacharunderscore}{\kern0pt}natI\ nat{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}quasinat\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ rule{\isacharcolon}{\kern0pt}nat{\isacharunderscore}{\kern0pt}cases{\isacharbrackleft}{\kern0pt}of\ i{\isacharbrackright}{\kern0pt}{\isacharsemicolon}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}take{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharat}{\kern0pt}\ drop{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Cons\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharsemicolon}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ Cons\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ list{\isacharunderscore}{\kern0pt}split\ {\isacharcolon}{\kern0pt}\isanewline
\isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}rest{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rest\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isakeyword{shows}\ \ {\isachardoublequoteopen}{\isasymexists}re{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}st{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ rest\ {\isacharequal}{\kern0pt}\ re\ {\isacharat}{\kern0pt}\ st\ {\isasymand}\ length{\isacharparenleft}{\kern0pt}re{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}rest{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}n{\isasymle}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ pred{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}rest{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}rest\ {\isacharequal}{\kern0pt}\ take{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rest{\isacharparenright}{\kern0pt}\ {\isacharat}{\kern0pt}\ drop{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rest{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}re\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}st{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ take{\isacharunderscore}{\kern0pt}drop{\isacharunderscore}{\kern0pt}eq{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}rest{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ le{\isacharunderscore}{\kern0pt}natI\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}re{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}re{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}st{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ length{\isacharunderscore}{\kern0pt}take{\isacharbrackleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}format{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}rest{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ min{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ rev{\isacharunderscore}{\kern0pt}bexI{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlambda}\ re{\isachardot}{\kern0pt}\ {\isasymexists}st{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ rest\ {\isacharequal}{\kern0pt}\ re\ {\isacharat}{\kern0pt}\ st\ {\isasymand}\ length{\isacharparenleft}{\kern0pt}re{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}re{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}rest\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}nForall{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}n{\isasymin}nat\ {\isasymLongrightarrow}\ ms\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ ms\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}Forall{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}rest\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}rest{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isasymlongrightarrow}\ M{\isacharcomma}{\kern0pt}\ rest\ {\isacharat}{\kern0pt}\ ms\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ n\ arbitrary{\isacharcolon}{\kern0pt}ms\ set{\isacharcolon}{\kern0pt}nat{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{0}}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}succ\ n{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}rest{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}rest{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ P{\isacharparenleft}{\kern0pt}rest{\isacharcomma}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}t{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymforall}res{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}res{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isasymlongrightarrow}\ P{\isacharparenleft}{\kern0pt}res\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}t{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ \isakeyword{for}\ n\ P\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ last{\isacharunderscore}{\kern0pt}init{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ this{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlambda}rest\ {\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ rest\ {\isacharat}{\kern0pt}\ ms\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}rest{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}rest{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ M{\isacharcomma}{\kern0pt}\ rest\ {\isacharat}{\kern0pt}\ ms\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}t{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymforall}res{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}res{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isasymlongrightarrow}\ \ M{\isacharcomma}{\kern0pt}\ {\isacharparenleft}{\kern0pt}res\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}t{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharat}{\kern0pt}\ ms\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\ succ{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}ms{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isacharunderscore}{\kern0pt}\ M\ {\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}\ ms\ {\isacharat}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ app{\isacharunderscore}{\kern0pt}assoc\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Forall{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ incr{\isacharunderscore}{\kern0pt}bv{\isadigit{1}}{\isacharcircum}{\kern0pt}{\isadigit{2}}{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}ms{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rest{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ rest\ {\isacharat}{\kern0pt}\ ms\ {\isasymTurnstile}\ sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ rest\ {\isacharat}{\kern0pt}\ ms\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ formula{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}params{\isadigit{1}}{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isadigit{2}}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ separation{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Forall{\isacharcircum}{\kern0pt}{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ {\isasymlongleftrightarrow}\isanewline
\ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ iffI\ ballI\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ env\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ some\ rest\ \isakeyword{where}\ {\isachardoublequoteopen}some{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rest{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isacharequal}{\kern0pt}\ some\ {\isacharat}{\kern0pt}\ rest{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}some{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ list{\isacharunderscore}{\kern0pt}split{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \isacommand{using}\isamarkupfalse%
\ succpred{\isacharunderscore}{\kern0pt}leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ some\ {\isasymTurnstile}\ sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}nForall{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharquery}{\kern0pt}n{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ M\ some{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ rest\ M\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ some{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ separation{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ some\ {\isacharat}{\kern0pt}\ rest{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\ %
\isamarkupcmt{almost equal to the previous implication%
}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ asm{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ some\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}some{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}some{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}some{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ succpred{\isacharunderscore}{\kern0pt}leI{\isacharbrackright}{\kern0pt}\ succpred{\isacharunderscore}{\kern0pt}leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isakeyword{and}\ asm\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ some\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ some\ {\isasymTurnstile}\ sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ M\ some{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isacharunderscore}{\kern0pt}\ M\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ some{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ some\ {\isacharat}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}nForall{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharquery}{\kern0pt}n{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{The Axiom of Replacement, internalized%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ \isanewline
\ \ \ \ Q{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}x\ y\ z{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ y\ {\isasymin}\ A\ {\isasymLongrightarrow}\ z{\isasymin}A\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Q{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}x\ y\ z{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ y\ {\isasymin}\ A\ {\isasymLongrightarrow}\ z{\isasymin}A\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Q{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ \isanewline
\ \ \ \ asms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ B{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}B{\isacharcomma}{\kern0pt}Q{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ A{\isacharcomma}{\kern0pt}env\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}ufm{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ asms{\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ Q{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\ \ \isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize{\isacharunderscore}{\kern0pt}notc}\isamarkupfalse%
\ {\isachardoublequoteopen}univalent{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ sats{\isacharunderscore}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}q{\isadigit{1}}{\isasymin}\ formula\ {\isasymLongrightarrow}\ q{\isadigit{2}}{\isasymin}formula\ {\isasymLongrightarrow}\ i{\isasymin}nat\ {\isasymLongrightarrow}\ \isanewline
\ \ univalent{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}q{\isadigit{2}}{\isacharcomma}{\kern0pt}q{\isadigit{1}}{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ Q{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}x\ y\ z{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ y\ {\isasymin}\ A\ {\isasymLongrightarrow}\ z{\isasymin}A\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Q{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}x\ y\ z{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ y\ {\isasymin}\ A\ {\isasymLongrightarrow}\ z{\isasymin}A\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Q{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Q{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ \isanewline
\ \ \ \ asms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ B{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isacharcomma}{\kern0pt}env\ {\isasymTurnstile}\ univalent{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}Q{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}Q{\isadigit{2}}{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}B{\isacharcomma}{\kern0pt}Q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ asms\ sats{\isacharunderscore}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharbrackleft}{\kern0pt}OF\ Q{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ swap{\isacharunderscore}{\kern0pt}vars\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}swap{\isacharunderscore}{\kern0pt}vars{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ \ \ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}iterates{\isacharparenleft}{\kern0pt}{\isasymlambda}p{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{2}}\ {\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ swap{\isacharunderscore}{\kern0pt}vars{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ swap{\isacharunderscore}{\kern0pt}vars{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ swap{\isacharunderscore}{\kern0pt}vars{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}swap{\isacharunderscore}{\kern0pt}vars\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ swap{\isacharunderscore}{\kern0pt}vars{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymlongleftrightarrow}\ M{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ swap{\isacharunderscore}{\kern0pt}vars{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}iff\ {\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ M\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{1}}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{1}}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ incr{\isacharunderscore}{\kern0pt}bv{\isadigit{1}}{\isacharparenleft}{\kern0pt}swap{\isacharunderscore}{\kern0pt}vars{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{2}}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{2}}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}swap{\isacharunderscore}{\kern0pt}vars{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}Qs{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{1}}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{2}}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{1}}{\isacharunderscore}{\kern0pt}def\ univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{2}}{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}assm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{1}}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{2}}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{1}}{\isacharunderscore}{\kern0pt}def\ univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{2}}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ \isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ A\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ %
\isamarkupcmt{simplifies iterates of \isa{{\isasymlambda}x{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isadigit{0}}}%
}\isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isadigit{1}}{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ A\ z\ y{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}swap{\isacharunderscore}{\kern0pt}vars\ \ assms\ \isanewline
\ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Forall{\isacharparenleft}{\kern0pt}Implies{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ univalent{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{1}}{\isacharparenleft}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{2}}{\isacharparenleft}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ Exists{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}simps\ {\isacharequal}{\kern0pt}\ formula{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}params{\isadigit{1}}{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isadigit{2}}\ {\isacharunderscore}{\kern0pt}\ M\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbrackright}{\kern0pt}\isanewline
\ \ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ M\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ %
\isamarkupcmt{simplifies iterates of \isa{{\isasymlambda}x{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isadigit{0}}}%
}\isanewline
\ \ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ M\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}%
\isamarkupcmt{simplifies \isa{{\isasymlambda}x{\isachardot}{\kern0pt}\ incr{\isacharunderscore}{\kern0pt}bv{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isadigit{2}}}%
}\isanewline
\ \ sats{\isacharunderscore}{\kern0pt}incr{\isacharunderscore}{\kern0pt}bv{\isadigit{1}}{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ M{\isacharbrackright}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}swap{\isacharunderscore}{\kern0pt}vars\ \isakeyword{for}\ {\isasymphi}\ M\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}ms{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rest{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ rest\ {\isacharat}{\kern0pt}\ ms\ {\isasymTurnstile}\ rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ rest\ {\isacharat}{\kern0pt}\ ms\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}simps\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}def\ univalent{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{1}}{\isacharunderscore}{\kern0pt}def\ univalent{\isacharunderscore}{\kern0pt}Q{\isadigit{2}}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Forall{\isacharcircum}{\kern0pt}{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ {\isasymlongleftrightarrow}\isanewline
\ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ iffI\ ballI\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ env\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}succ{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ pred{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ some\ rest\ \isakeyword{where}\ {\isachardoublequoteopen}some{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rest{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isacharequal}{\kern0pt}\ some\ {\isacharat}{\kern0pt}\ rest{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}some{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ list{\isacharunderscore}{\kern0pt}split{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}pred{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ succpred{\isacharunderscore}{\kern0pt}leI{\isacharbrackright}{\kern0pt}\ succpred{\isacharunderscore}{\kern0pt}leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ some\ {\isasymTurnstile}\ rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}nForall{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharquery}{\kern0pt}n{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ M\ some{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ rest\ M\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ some{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ some\ {\isacharat}{\kern0pt}\ rest{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\ %
\isamarkupcmt{almost equal to the previous implication%
}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ asm{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ some\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}some{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}some{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}some{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ succpred{\isacharunderscore}{\kern0pt}leI{\isacharbrackright}{\kern0pt}\ succpred{\isacharunderscore}{\kern0pt}leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isakeyword{and}\ asm\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ some\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ some\ {\isasymTurnstile}\ rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ M\ some{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isacharunderscore}{\kern0pt}\ M\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ some{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ some\ {\isacharat}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}nForall{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}rep{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharquery}{\kern0pt}n{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ZF{\isacharunderscore}{\kern0pt}inf\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}inf\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ p\ {\isasymin}\ formula\ {\isacharbraceright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ p\ {\isasymin}\ formula\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}formula{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A{\isasymsubseteq}formula\ {\isasymand}\ B{\isasymsubseteq}formula\ {\isasymLongrightarrow}\ A{\isasymunion}B\ {\isasymsubseteq}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}inf{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}formula\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}inf\ {\isasymsubseteq}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}inf{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ZFC\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ZFC\ {\isasymequiv}\ ZF{\isacharunderscore}{\kern0pt}inf\ {\isasymunion}\ ZFC{\isacharunderscore}{\kern0pt}fin{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ZF\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ZF\ {\isasymequiv}\ ZF{\isacharunderscore}{\kern0pt}inf\ {\isasymunion}\ ZF{\isacharunderscore}{\kern0pt}fin{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ ZF{\isacharunderscore}{\kern0pt}minus{\isacharunderscore}{\kern0pt}P\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ZF{\isacharunderscore}{\kern0pt}minus{\isacharunderscore}{\kern0pt}P\ {\isasymequiv}\ ZF\ {\isacharminus}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ ZF{\isacharunderscore}{\kern0pt}power{\isacharunderscore}{\kern0pt}fm\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ZFC{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}formula{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}ZFC\ {\isasymsubseteq}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}ZFC{\isacharunderscore}{\kern0pt}def\ Un{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}formula\ ZF{\isacharunderscore}{\kern0pt}inf{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}formula\ ZFC{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptxt}%
Satisfaction of a set of sentences%
\end{isamarkuptxt}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ satT\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}{\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isasymTurnstile}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isadigit{6}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\ {\isadigit{6}}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}A\ {\isasymTurnstile}\ {\isasymPhi}\ \ {\isasymequiv}\ \ {\isasymforall}{\isasymphi}{\isasymin}{\isasymPhi}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ satTI\ {\isacharbrackleft}{\kern0pt}intro{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}{\isasymphi}{\isachardot}{\kern0pt}\ {\isasymphi}{\isasymin}{\isasymPhi}\ {\isasymLongrightarrow}\ A{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}A\ {\isasymTurnstile}\ {\isasymPhi}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ satT{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ satTD\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}{\isachardoublequoteopen}A\ {\isasymTurnstile}\ {\isasymPhi}\ {\isasymLongrightarrow}\ \ {\isasymphi}{\isasymin}{\isasymPhi}\ {\isasymLongrightarrow}\ A{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ satT{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ZFC{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}AC{\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}N\ {\isasymTurnstile}\ ZFC{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}N\ {\isasymTurnstile}\ ZF{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}N{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ AC{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ZFC{\isacharunderscore}{\kern0pt}def\ ZFC{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}satT{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}ZF{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}M\ {\isasymTurnstile}\ ZF{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}M\ {\isasymTurnstile}\ ZF{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ fin{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}upair{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}power{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}extensionality{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}foundation{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}infinity{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}def\ ZFC{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}defs\ satT{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ZFC{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}of\ M{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ {\isasymphi}\ env\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}M\ {\isasymTurnstile}\ ZF{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}formula{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}formula{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}inf{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ sats{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \ \isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ fin\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}ZF{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}ZF{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M\ {\isasymTurnstile}\ ZF{\isacharunderscore}{\kern0pt}fin{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}def\ ZFC{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}defs\ satT{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ZFC{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}of\ M{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}ZF{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}formula{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}formula{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ sats{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}M\ {\isasymTurnstile}\ ZF{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}inf{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Internal_ZFC_Axioms.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%32=6%:%
%:%37=6%:%
%:%40=7%:%
%:%41=8%:%
%:%42=8%:%
%:%43=9%:%
%:%46=10%:%
%:%50=10%:%
%:%51=10%:%
%:%52=11%:%
%:%53=11%:%
%:%58=11%:%
%:%63=12%:%
%:%68=13%:%
%:%69=13%:%
%:%74=13%:%
%:%77=14%:%
%:%78=15%:%
%:%79=15%:%
%:%80=16%:%
%:%83=17%:%
%:%87=17%:%
%:%88=17%:%
%:%89=18%:%
%:%90=18%:%
%:%95=18%:%
%:%100=19%:%
%:%105=20%:%
%:%106=20%:%
%:%111=20%:%
%:%114=21%:%
%:%115=22%:%
%:%116=22%:%
%:%117=23%:%
%:%120=24%:%
%:%124=24%:%
%:%125=24%:%
%:%126=25%:%
%:%127=25%:%
%:%132=25%:%
%:%137=26%:%
%:%142=27%:%
%:%143=27%:%
%:%148=27%:%
%:%151=28%:%
%:%152=29%:%
%:%153=29%:%
%:%154=30%:%
%:%157=31%:%
%:%161=31%:%
%:%162=31%:%
%:%163=32%:%
%:%164=32%:%
%:%169=32%:%
%:%174=33%:%
%:%179=34%:%
%:%180=34%:%
%:%185=34%:%
%:%188=35%:%
%:%189=36%:%
%:%190=36%:%
%:%191=37%:%
%:%194=38%:%
%:%198=38%:%
%:%199=38%:%
%:%200=39%:%
%:%201=39%:%
%:%206=39%:%
%:%211=40%:%
%:%216=41%:%
%:%217=41%:%
%:%222=41%:%
%:%225=42%:%
%:%226=43%:%
%:%227=43%:%
%:%228=44%:%
%:%231=45%:%
%:%235=45%:%
%:%236=45%:%
%:%237=46%:%
%:%238=46%:%
%:%243=46%:%
%:%248=47%:%
%:%253=48%:%
%:%254=48%:%
%:%259=48%:%
%:%262=49%:%
%:%263=50%:%
%:%264=50%:%
%:%265=51%:%
%:%268=52%:%
%:%272=52%:%
%:%273=52%:%
%:%274=53%:%
%:%275=53%:%
%:%280=53%:%
%:%285=54%:%
%:%290=55%:%
%:%291=55%:%
%:%296=55%:%
%:%299=56%:%
%:%300=57%:%
%:%301=57%:%
%:%302=58%:%
%:%303=59%:%
%:%304=59%:%
%:%305=60%:%
%:%306=61%:%
%:%307=62%:%
%:%308=62%:%
%:%309=63%:%
%:%310=64%:%
%:%311=65%:%
%:%312=65%:%
%:%313=66%:%
%:%314=67%:%
%:%315=68%:%
%:%316=68%:%
%:%317=69%:%
%:%318=70%:%
%:%319=71%:%
%:%320=72%:%
%:%321=73%:%
%:%322=73%:%
%:%323=74%:%
%:%324=75%:%
%:%325=76%:%
%:%326=77%:%
%:%327=77%:%
%:%330=78%:%
%:%334=78%:%
%:%335=78%:%
%:%336=78%:%
%:%350=80%:%
%:%360=81%:%
%:%361=81%:%
%:%362=82%:%
%:%365=83%:%
%:%369=83%:%
%:%370=83%:%
%:%375=83%:%
%:%378=84%:%
%:%379=85%:%
%:%380=85%:%
%:%381=86%:%
%:%382=87%:%
%:%389=88%:%
%:%390=88%:%
%:%391=89%:%
%:%392=89%:%
%:%393=90%:%
%:%394=90%:%
%:%395=91%:%
%:%396=91%:%
%:%397=92%:%
%:%398=92%:%
%:%399=93%:%
%:%400=93%:%
%:%401=94%:%
%:%402=94%:%
%:%403=95%:%
%:%404=95%:%
%:%405=96%:%
%:%406=96%:%
%:%407=97%:%
%:%408=97%:%
%:%409=97%:%
%:%410=98%:%
%:%411=98%:%
%:%412=99%:%
%:%413=99%:%
%:%414=99%:%
%:%415=100%:%
%:%421=100%:%
%:%424=101%:%
%:%425=102%:%
%:%426=102%:%
%:%427=103%:%
%:%428=104%:%
%:%431=105%:%
%:%435=105%:%
%:%436=105%:%
%:%437=106%:%
%:%438=106%:%
%:%439=107%:%
%:%440=107%:%
%:%441=108%:%
%:%442=108%:%
%:%443=108%:%
%:%444=108%:%
%:%445=109%:%
%:%446=109%:%
%:%447=110%:%
%:%448=110%:%
%:%449=111%:%
%:%450=111%:%
%:%451=111%:%
%:%452=112%:%
%:%453=112%:%
%:%454=113%:%
%:%455=113%:%
%:%456=114%:%
%:%457=114%:%
%:%458=115%:%
%:%459=115%:%
%:%460=116%:%
%:%461=116%:%
%:%462=117%:%
%:%463=117%:%
%:%464=118%:%
%:%465=118%:%
%:%466=119%:%
%:%467=119%:%
%:%468=120%:%
%:%469=120%:%
%:%470=121%:%
%:%471=121%:%
%:%472=122%:%
%:%473=122%:%
%:%474=123%:%
%:%475=123%:%
%:%476=124%:%
%:%477=124%:%
%:%478=125%:%
%:%479=125%:%
%:%480=125%:%
%:%481=125%:%
%:%482=125%:%
%:%483=126%:%
%:%484=126%:%
%:%485=127%:%
%:%491=127%:%
%:%494=128%:%
%:%495=129%:%
%:%496=129%:%
%:%497=130%:%
%:%498=131%:%
%:%505=132%:%
%:%506=132%:%
%:%507=133%:%
%:%508=133%:%
%:%509=134%:%
%:%510=134%:%
%:%511=135%:%
%:%512=135%:%
%:%513=135%:%
%:%514=136%:%
%:%515=136%:%
%:%516=137%:%
%:%517=137%:%
%:%518=138%:%
%:%519=138%:%
%:%520=138%:%
%:%521=139%:%
%:%522=139%:%
%:%523=140%:%
%:%524=140%:%
%:%525=141%:%
%:%526=141%:%
%:%527=142%:%
%:%528=142%:%
%:%529=143%:%
%:%530=143%:%
%:%531=144%:%
%:%532=144%:%
%:%533=145%:%
%:%534=145%:%
%:%535=146%:%
%:%536=146%:%
%:%537=147%:%
%:%538=148%:%
%:%539=148%:%
%:%540=149%:%
%:%546=149%:%
%:%549=150%:%
%:%550=151%:%
%:%551=151%:%
%:%552=152%:%
%:%553=153%:%
%:%554=154%:%
%:%555=155%:%
%:%557=157%:%
%:%564=158%:%
%:%565=158%:%
%:%566=159%:%
%:%567=159%:%
%:%568=160%:%
%:%569=160%:%
%:%570=161%:%
%:%571=161%:%
%:%572=161%:%
%:%573=162%:%
%:%574=162%:%
%:%575=163%:%
%:%576=163%:%
%:%577=164%:%
%:%578=164%:%
%:%579=165%:%
%:%580=166%:%
%:%581=167%:%
%:%582=167%:%
%:%583=167%:%
%:%584=168%:%
%:%585=168%:%
%:%586=169%:%
%:%587=169%:%
%:%588=170%:%
%:%589=171%:%
%:%590=171%:%
%:%591=172%:%
%:%592=172%:%
%:%593=173%:%
%:%594=173%:%
%:%595=174%:%
%:%596=174%:%
%:%597=175%:%
%:%598=175%:%
%:%599=176%:%
%:%605=176%:%
%:%608=177%:%
%:%609=178%:%
%:%610=178%:%
%:%611=179%:%
%:%612=180%:%
%:%614=182%:%
%:%615=183%:%
%:%616=184%:%
%:%617=184%:%
%:%620=185%:%
%:%624=185%:%
%:%625=185%:%
%:%630=185%:%
%:%633=186%:%
%:%634=187%:%
%:%635=187%:%
%:%636=188%:%
%:%637=189%:%
%:%638=190%:%
%:%639=191%:%
%:%640=192%:%
%:%643=193%:%
%:%647=193%:%
%:%648=193%:%
%:%649=194%:%
%:%650=194%:%
%:%651=194%:%
%:%656=194%:%
%:%659=195%:%
%:%660=196%:%
%:%661=196%:%
%:%662=197%:%
%:%663=198%:%
%:%664=199%:%
%:%665=200%:%
%:%666=200%:%
%:%669=201%:%
%:%673=201%:%
%:%674=201%:%
%:%679=201%:%
%:%682=202%:%
%:%683=203%:%
%:%684=203%:%
%:%685=204%:%
%:%686=205%:%
%:%687=206%:%
%:%688=207%:%
%:%691=210%:%
%:%698=211%:%
%:%699=211%:%
%:%700=212%:%
%:%701=212%:%
%:%702=213%:%
%:%703=213%:%
%:%704=214%:%
%:%705=214%:%
%:%706=215%:%
%:%707=215%:%
%:%708=216%:%
%:%709=216%:%
%:%710=216%:%
%:%711=217%:%
%:%712=217%:%
%:%713=217%:%
%:%714=218%:%
%:%715=218%:%
%:%716=219%:%
%:%717=219%:%
%:%718=220%:%
%:%719=221%:%
%:%720=221%:%
%:%721=221%:%
%:%722=222%:%
%:%723=222%:%
%:%724=222%:%
%:%725=223%:%
%:%726=223%:%
%:%727=224%:%
%:%728=224%:%
%:%729=224%:%
%:%730=225%:%
%:%731=225%:%
%:%732=226%:%
%:%733=226%:%
%:%734=227%:%
%:%735=227%:%
%:%736=228%:%
%:%737=228%:%
%:%738=229%:%
%:%739=229%:%
%:%740=229%:%
%:%741=230%:%
%:%742=230%:%
%:%743=231%:%
%:%744=231%:%
%:%745=232%:%
%:%746=232%:%
%:%747=232%:%
%:%748=233%:%
%:%749=233%:%
%:%750=234%:%
%:%751=234%:%
%:%752=235%:%
%:%753=235%:%
%:%754=236%:%
%:%755=236%:%
%:%756=237%:%
%:%757=238%:%
%:%758=239%:%
%:%759=239%:%
%:%760=240%:%
%:%761=240%:%
%:%762=240%:%
%:%763=240%:%
%:%764=241%:%
%:%765=241%:%
%:%766=242%:%
%:%767=242%:%
%:%768=243%:%
%:%769=244%:%
%:%770=244%:%
%:%771=245%:%
%:%772=245%:%
%:%773=246%:%
%:%774=246%:%
%:%775=247%:%
%:%776=247%:%
%:%777=248%:%
%:%778=248%:%
%:%779=249%:%
%:%780=249%:%
%:%781=249%:%
%:%782=250%:%
%:%783=250%:%
%:%784=251%:%
%:%785=251%:%
%:%786=251%:%
%:%787=252%:%
%:%788=252%:%
%:%789=252%:%
%:%790=253%:%
%:%791=253%:%
%:%792=253%:%
%:%793=254%:%
%:%794=254%:%
%:%795=255%:%
%:%796=255%:%
%:%797=256%:%
%:%798=256%:%
%:%799=257%:%
%:%800=258%:%
%:%801=259%:%
%:%802=259%:%
%:%803=260%:%
%:%804=260%:%
%:%805=261%:%
%:%806=261%:%
%:%807=262%:%
%:%808=262%:%
%:%809=263%:%
%:%810=263%:%
%:%811=264%:%
%:%812=264%:%
%:%813=265%:%
%:%814=265%:%
%:%815=266%:%
%:%830=268%:%
%:%840=269%:%
%:%841=269%:%
%:%842=270%:%
%:%843=271%:%
%:%843=273%:%
%:%844=274%:%
%:%845=275%:%
%:%846=276%:%
%:%847=277%:%
%:%848=278%:%
%:%849=279%:%
%:%850=280%:%
%:%851=281%:%
%:%854=282%:%
%:%858=282%:%
%:%859=282%:%
%:%860=283%:%
%:%861=283%:%
%:%866=283%:%
%:%871=284%:%
%:%876=285%:%
%:%877=285%:%
%:%882=285%:%
%:%885=286%:%
%:%886=287%:%
%:%887=287%:%
%:%888=288%:%
%:%891=289%:%
%:%895=289%:%
%:%896=289%:%
%:%901=289%:%
%:%904=290%:%
%:%905=291%:%
%:%906=291%:%
%:%907=292%:%
%:%908=293%:%
%:%909=294%:%
%:%910=295%:%
%:%911=296%:%
%:%912=297%:%
%:%913=298%:%
%:%914=299%:%
%:%915=300%:%
%:%918=301%:%
%:%922=301%:%
%:%923=301%:%
%:%924=301%:%
%:%925=301%:%
%:%930=301%:%
%:%933=302%:%
%:%934=303%:%
%:%935=303%:%
%:%936=304%:%
%:%937=305%:%
%:%938=306%:%
%:%939=307%:%
%:%940=308%:%
%:%941=308%:%
%:%942=309%:%
%:%945=310%:%
%:%949=310%:%
%:%950=310%:%
%:%951=310%:%
%:%956=310%:%
%:%959=311%:%
%:%960=312%:%
%:%961=312%:%
%:%962=313%:%
%:%963=314%:%
%:%966=315%:%
%:%970=315%:%
%:%971=315%:%
%:%972=316%:%
%:%973=316%:%
%:%974=316%:%
%:%979=316%:%
%:%982=317%:%
%:%983=318%:%
%:%984=318%:%
%:%985=319%:%
%:%986=320%:%
%:%987=321%:%
%:%988=322%:%
%:%989=322%:%
%:%990=323%:%
%:%991=324%:%
%:%992=325%:%
%:%993=326%:%
%:%994=326%:%
%:%995=327%:%
%:%996=328%:%
%:%999=329%:%
%:%1003=329%:%
%:%1004=329%:%
%:%1005=329%:%
%:%1006=329%:%
%:%1011=329%:%
%:%1014=330%:%
%:%1015=331%:%
%:%1016=331%:%
%:%1017=332%:%
%:%1018=333%:%
%:%1019=334%:%
%:%1020=335%:%
%:%1021=336%:%
%:%1024=337%:%
%:%1028=337%:%
%:%1029=337%:%
%:%1030=338%:%
%:%1031=338%:%
%:%1032=339%:%
%:%1033=339%:%
%:%1034=339%:%
%:%1035=340%:%
%:%1036=341%:%
%:%1037=342%:%
%:%1038=342%:%
%:%1043=342%:%
%:%1046=343%:%
%:%1047=344%:%
%:%1048=344%:%
%:%1049=345%:%
%:%1050=346%:%
%:%1053=349%:%
%:%1054=350%:%
%:%1055=351%:%
%:%1056=351%:%
%:%1059=352%:%
%:%1063=352%:%
%:%1064=352%:%
%:%1069=352%:%
%:%1072=353%:%
%:%1073=354%:%
%:%1074=354%:%
%:%1075=355%:%
%:%1076=355%:%
%:%1077=355%:%
%:%1078=356%:%
%:%1079=356%:%
%:%1080=356%:%
%:%1081=357%:%
%:%1082=358%:%
%:%1083=359%:%
%:%1084=359%:%
%:%1085=360%:%
%:%1086=361%:%
%:%1087=362%:%
%:%1088=363%:%
%:%1089=364%:%
%:%1092=365%:%
%:%1096=365%:%
%:%1097=365%:%
%:%1098=366%:%
%:%1099=366%:%
%:%1100=367%:%
%:%1101=367%:%
%:%1102=368%:%
%:%1103=368%:%
%:%1108=368%:%
%:%1111=369%:%
%:%1112=370%:%
%:%1113=370%:%
%:%1114=371%:%
%:%1115=372%:%
%:%1116=373%:%
%:%1117=374%:%
%:%1118=374%:%
%:%1121=375%:%
%:%1125=375%:%
%:%1126=375%:%
%:%1131=375%:%
%:%1134=376%:%
%:%1135=377%:%
%:%1136=377%:%
%:%1137=378%:%
%:%1138=379%:%
%:%1139=380%:%
%:%1140=381%:%
%:%1143=384%:%
%:%1150=385%:%
%:%1151=385%:%
%:%1152=386%:%
%:%1153=386%:%
%:%1154=387%:%
%:%1155=387%:%
%:%1156=388%:%
%:%1157=388%:%
%:%1158=389%:%
%:%1159=389%:%
%:%1160=389%:%
%:%1161=390%:%
%:%1162=390%:%
%:%1163=390%:%
%:%1164=391%:%
%:%1165=391%:%
%:%1166=391%:%
%:%1167=392%:%
%:%1168=392%:%
%:%1169=393%:%
%:%1170=393%:%
%:%1171=393%:%
%:%1172=394%:%
%:%1173=394%:%
%:%1174=394%:%
%:%1175=395%:%
%:%1176=395%:%
%:%1177=396%:%
%:%1178=397%:%
%:%1179=397%:%
%:%1180=397%:%
%:%1181=398%:%
%:%1182=398%:%
%:%1183=399%:%
%:%1184=399%:%
%:%1185=400%:%
%:%1186=400%:%
%:%1187=400%:%
%:%1188=401%:%
%:%1189=401%:%
%:%1190=402%:%
%:%1191=402%:%
%:%1192=402%:%
%:%1193=403%:%
%:%1194=403%:%
%:%1195=403%:%
%:%1196=404%:%
%:%1197=404%:%
%:%1198=405%:%
%:%1199=405%:%
%:%1200=406%:%
%:%1201=406%:%
%:%1202=407%:%
%:%1203=407%:%
%:%1204=408%:%
%:%1205=408%:%
%:%1206=409%:%
%:%1207=409%:%
%:%1208=410%:%
%:%1209=410%:%
%:%1210=411%:%
%:%1211=412%:%
%:%1212=413%:%
%:%1213=413%:%
%:%1214=414%:%
%:%1215=414%:%
%:%1216=414%:%
%:%1217=414%:%
%:%1218=415%:%
%:%1219=415%:%
%:%1220=416%:%
%:%1221=416%:%
%:%1222=417%:%
%:%1223=418%:%
%:%1224=418%:%
%:%1225=419%:%
%:%1226=419%:%
%:%1227=420%:%
%:%1228=420%:%
%:%1229=421%:%
%:%1230=421%:%
%:%1231=422%:%
%:%1232=422%:%
%:%1233=423%:%
%:%1234=423%:%
%:%1235=423%:%
%:%1236=424%:%
%:%1237=424%:%
%:%1238=425%:%
%:%1239=425%:%
%:%1240=425%:%
%:%1241=426%:%
%:%1242=426%:%
%:%1243=426%:%
%:%1244=427%:%
%:%1245=427%:%
%:%1246=427%:%
%:%1247=428%:%
%:%1248=428%:%
%:%1249=429%:%
%:%1250=429%:%
%:%1251=430%:%
%:%1252=430%:%
%:%1253=431%:%
%:%1254=432%:%
%:%1255=433%:%
%:%1256=433%:%
%:%1257=434%:%
%:%1258=434%:%
%:%1259=435%:%
%:%1260=435%:%
%:%1261=436%:%
%:%1262=436%:%
%:%1263=437%:%
%:%1264=437%:%
%:%1265=438%:%
%:%1266=438%:%
%:%1267=439%:%
%:%1268=439%:%
%:%1269=440%:%
%:%1275=440%:%
%:%1278=441%:%
%:%1279=442%:%
%:%1280=442%:%
%:%1281=443%:%
%:%1282=444%:%
%:%1283=445%:%
%:%1284=446%:%
%:%1285=446%:%
%:%1288=447%:%
%:%1292=447%:%
%:%1293=447%:%
%:%1298=447%:%
%:%1301=448%:%
%:%1302=449%:%
%:%1303=449%:%
%:%1306=450%:%
%:%1310=450%:%
%:%1311=450%:%
%:%1312=450%:%
%:%1317=450%:%
%:%1320=451%:%
%:%1321=452%:%
%:%1322=452%:%
%:%1323=453%:%
%:%1324=454%:%
%:%1325=455%:%
%:%1326=456%:%
%:%1327=456%:%
%:%1328=457%:%
%:%1329=458%:%
%:%1330=459%:%
%:%1331=460%:%
%:%1332=460%:%
%:%1333=461%:%
%:%1334=462%:%
%:%1335=463%:%
%:%1336=464%:%
%:%1337=464%:%
%:%1340=465%:%
%:%1344=465%:%
%:%1345=465%:%
%:%1354=467%:%
%:%1356=468%:%
%:%1357=468%:%
%:%1358=469%:%
%:%1359=470%:%
%:%1360=471%:%
%:%1361=472%:%
%:%1362=472%:%
%:%1363=473%:%
%:%1364=474%:%
%:%1367=475%:%
%:%1371=475%:%
%:%1372=475%:%
%:%1373=475%:%
%:%1374=475%:%
%:%1379=475%:%
%:%1382=476%:%
%:%1383=477%:%
%:%1384=477%:%
%:%1387=478%:%
%:%1391=478%:%
%:%1392=478%:%
%:%1393=478%:%
%:%1398=478%:%
%:%1401=479%:%
%:%1402=480%:%
%:%1403=480%:%
%:%1404=481%:%
%:%1407=482%:%
%:%1411=482%:%
%:%1412=482%:%
%:%1413=482%:%
%:%1418=482%:%
%:%1421=483%:%
%:%1422=484%:%
%:%1423=484%:%
%:%1430=485%:%
%:%1431=485%:%
%:%1432=486%:%
%:%1433=486%:%
%:%1434=487%:%
%:%1435=487%:%
%:%1436=488%:%
%:%1437=488%:%
%:%1438=489%:%
%:%1439=490%:%
%:%1440=490%:%
%:%1441=491%:%
%:%1442=491%:%
%:%1443=491%:%
%:%1444=492%:%
%:%1445=492%:%
%:%1446=493%:%
%:%1447=493%:%
%:%1448=494%:%
%:%1449=494%:%
%:%1450=495%:%
%:%1451=495%:%
%:%1452=495%:%
%:%1453=496%:%
%:%1454=496%:%
%:%1455=497%:%
%:%1456=498%:%
%:%1457=498%:%
%:%1458=498%:%
%:%1459=499%:%
%:%1460=499%:%
%:%1461=499%:%
%:%1462=500%:%
%:%1463=500%:%
%:%1464=501%:%
%:%1465=502%:%
%:%1466=502%:%
%:%1467=502%:%
%:%1468=503%:%
%:%1469=503%:%
%:%1470=504%:%
%:%1471=504%:%
%:%1472=505%:%
%:%1473=505%:%
%:%1474=506%:%
%:%1475=506%:%
%:%1476=506%:%
%:%1477=507%:%
%:%1478=507%:%
%:%1479=508%:%
%:%1480=508%:%
%:%1481=509%:%
%:%1482=509%:%
%:%1483=510%:%
%:%1484=510%:%
%:%1485=511%:%
%:%1486=511%:%
%:%1487=512%:%
%:%1488=512%:%
%:%1489=512%:%
%:%1490=513%:%
%:%1491=513%:%
%:%1492=513%:%
%:%1493=514%:%
%:%1494=514%:%
%:%1495=515%:%
%:%1496=516%:%
%:%1497=516%:%
%:%1498=516%:%
%:%1499=517%:%
%:%1500=517%:%
%:%1501=518%:%
%:%1502=518%:%
%:%1503=519%:%
%:%1504=519%:%
%:%1505=520%:%
%:%1506=520%:%
%:%1507=520%:%
%:%1508=521%:%
%:%1514=521%:%
%:%1519=522%:%
%:%1524=523%:%