### 2024-11-01 code/output/document/Internalizations.tex

%
\begin{isabellebody}%
\setisabellecontext{Internalizations}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Aids to internalize formulas%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Internalizations\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}ZF{\isacharminus}{\kern0pt}Constructible{\isachardot}{\kern0pt}DPow{\isacharunderscore}{\kern0pt}absolute{\isachardoublequoteclose}\ \isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
We found it useful to have slightly different versions of some 
results in ZF-Constructible:%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isadigit{0}}{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymin}A{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ env{\isacharsemicolon}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ {\isacharequal}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}Nand{\isacharunderscore}{\kern0pt}iff\ sats{\isacharunderscore}{\kern0pt}Forall{\isacharunderscore}{\kern0pt}iff\ sats{\isacharunderscore}{\kern0pt}Neg{\isacharunderscore}{\kern0pt}iff\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ sats{\isacharunderscore}{\kern0pt}Or{\isacharunderscore}{\kern0pt}iff\ sats{\isacharunderscore}{\kern0pt}Implies{\isacharunderscore}{\kern0pt}iff\ sats{\isacharunderscore}{\kern0pt}Iff{\isacharunderscore}{\kern0pt}iff\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}ConsI{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}nth{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharsemicolon}{\kern0pt}\ n\ {\isasymin}\ nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}rules\ {\isacharequal}{\kern0pt}\ nth{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ nth{\isacharunderscore}{\kern0pt}ConsI\ nat{\isacharunderscore}{\kern0pt}{\isadigit{0}}I\ nat{\isacharunderscore}{\kern0pt}succI\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharequal}{\kern0pt}\ nth{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ nth{\isacharunderscore}{\kern0pt}ConsI\ FOL{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ function{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ fun{\isacharunderscore}{\kern0pt}plus{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ successor{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ omega{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ Replace{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats%
\begin{isamarkuptext}%
Also a different compilation of lemmas (term\isa{sep{\isacharunderscore}{\kern0pt}rules}) used in formula
 synthesis%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemmas}\isamarkupfalse%
\ fm{\isacharunderscore}{\kern0pt}defs\ {\isacharequal}{\kern0pt}\ omega{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ limit{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ empty{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ typed{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ pair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ succ{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ cons{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ union{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ relation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ composition{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ field{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ range{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ transset{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ subset{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Replace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Internalizations.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%40=7%:%
%:%41=8%:%
%:%43=9%:%
%:%44=9%:%
%:%45=10%:%
%:%46=11%:%
%:%49=12%:%
%:%53=12%:%
%:%54=12%:%
%:%55=12%:%
%:%56=12%:%
%:%61=12%:%
%:%64=13%:%
%:%65=14%:%
%:%66=14%:%
%:%67=15%:%
%:%68=16%:%
%:%69=17%:%
%:%70=17%:%
%:%77=18%:%
%:%78=18%:%
%:%83=18%:%
%:%86=19%:%
%:%87=20%:%
%:%88=20%:%
%:%89=21%:%
%:%90=21%:%
%:%91=22%:%
%:%92=23%:%
%:%94=25%:%
%:%95=26%:%
%:%97=27%:%
%:%98=27%:%
%:%99=28%:%
%:%100=29%:%
%:%101=30%:%
%:%102=31%:%
%:%103=32%:%
%:%106=33%:%
%:%111=34%:%