### 2024-11-01 code/output/document/Least.tex

%
\begin{isabellebody}%
\setisabellecontext{Least}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The binder \isa{Least}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Least\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ Names\isanewline
\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
We have some basic results on the least ordinal satisfying
a predicate.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ Least{\isacharunderscore}{\kern0pt}Ord{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ R{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isasymand}\ R{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Least{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}lt{\isacharunderscore}{\kern0pt}Ord{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}Least{\isacharunderscore}{\kern0pt}cong{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ R{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ R{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isasymand}\ R{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ Least{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ least\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}least{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ ordinal{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}empty{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ ordinal{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isasymnot}Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ {\isasymor}\ {\isacharparenleft}{\kern0pt}Q{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ ordinal{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ b{\isasymin}i{\isasymlongrightarrow}\ {\isasymnot}Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ least{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}least{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ And{\isacharparenleft}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ Or{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}Implies{\isacharparenleft}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Neg{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ Forall{\isacharparenleft}{\kern0pt}Implies{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Neg{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}{\isachardoublequoteopen}i\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ q{\isasymin}formula\ {\isasymLongrightarrow}\ least{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ basic{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}simps\ {\isacharequal}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}transset{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ p{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}a{\isachardot}{\kern0pt}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ {\isadigit{0}}{\isasymin}A{\isasymrbrakk}\isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ least{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}closed\ p{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ \isacommand{unfolding}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}def\ least{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}basic{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}simps{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ is{\isacharunderscore}{\kern0pt}Q{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}a{\isachardot}{\kern0pt}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}Q{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ q{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isacharsemicolon}{\kern0pt}\ j\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ {\isadigit{0}}{\isasymin}A{\isasymrbrakk}\isanewline
\ \ \ {\isasymLongrightarrow}\ least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Q{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ least{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}fm\ {\isacharbrackleft}{\kern0pt}OF\ is{\isacharunderscore}{\kern0pt}Q{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcomma}{\kern0pt}\ of\ j\ {\isacharcomma}{\kern0pt}\ symmetric{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}conj{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a{\isasymin}M\ {\isasymLongrightarrow}\ least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ x{\isasymin}M\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}ctm{\isacharparenright}{\kern0pt}\ unique{\isacharunderscore}{\kern0pt}least{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a{\isasymin}M\ {\isasymLongrightarrow}\ b{\isasymin}M\ {\isasymLongrightarrow}\ least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ a{\isacharequal}{\kern0pt}b{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ erule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}a\ \isakeyword{and}\ j{\isacharequal}{\kern0pt}b\ \isakeyword{in}\ Ord{\isacharunderscore}{\kern0pt}linear{\isacharunderscore}{\kern0pt}lt{\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}drule\ ltD\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ auto\ intro{\isacharcolon}{\kern0pt}Ord{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Ord{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}trivial\isanewline
\isakeyword{begin}%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Absoluteness and closure under \isa{Least}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{lemma}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}least{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Q{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ a\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}def\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}{\isasymforall}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isasymnot}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharsemicolon}{\kern0pt}\ intro\ iffI{\isacharsemicolon}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ {\isacharparenleft}{\kern0pt}{\isasymexists}i{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isacharequal}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Least{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}ordinal{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}empty{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ Least{\isacharparenleft}{\kern0pt}Q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ Q{\isacharparenleft}{\kern0pt}Least{\isacharparenleft}{\kern0pt}Q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ i\ \isakeyword{where}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ \ {\isacartoucheopen}Q{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isacharquery}{\kern0pt}G{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ intro{\isacharcolon}{\kern0pt}LeastI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ b\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isasymnot}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}H{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ less{\isacharunderscore}{\kern0pt}LeastE{\isacharbrackleft}{\kern0pt}of\ Q\ {\isacharunderscore}{\kern0pt}\ False{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ drule{\isacharunderscore}{\kern0pt}tac\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ blast{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}ordinal{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}empty{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isasymnot}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ {\isacharquery}{\kern0pt}G\ {\isasymand}\ {\isacharquery}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymexists}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ i\ \isakeyword{where}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isasymnot}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ Q{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ b\ {\isasymin}\ a\ {\isasymlongrightarrow}\ {\isasymnot}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}b{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ b\ {\isasymin}\ a\ {\isasymlongrightarrow}\ {\isasymnot}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ {\isacartoucheopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ b\ {\isasymin}\ a\ {\isasymLongrightarrow}\ {\isasymnot}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ b\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a\ {\isasymLongrightarrow}\ {\isasymnot}\ Q{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ b\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Least{\isacharunderscore}{\kern0pt}equality\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Least{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}{\isasymmu}\ x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ LeastI{\isacharbrackleft}{\kern0pt}of\ Q{\isacharbrackright}{\kern0pt}\ Least{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}i{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Least.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%32=6%:%
%:%41=8%:%
%:%42=9%:%
%:%44=11%:%
%:%45=11%:%
%:%48=12%:%
%:%52=12%:%
%:%53=12%:%
%:%54=12%:%
%:%59=12%:%
%:%62=13%:%
%:%63=14%:%
%:%64=14%:%
%:%65=15%:%
%:%66=16%:%
%:%73=17%:%
%:%74=17%:%
%:%75=18%:%
%:%76=18%:%
%:%77=19%:%
%:%78=19%:%
%:%79=20%:%
%:%80=20%:%
%:%81=21%:%
%:%82=21%:%
%:%83=22%:%
%:%84=22%:%
%:%85=22%:%
%:%86=22%:%
%:%87=23%:%
%:%93=23%:%
%:%96=24%:%
%:%97=25%:%
%:%98=25%:%
%:%99=26%:%
%:%100=27%:%
%:%102=29%:%
%:%103=30%:%
%:%104=31%:%
%:%105=31%:%
%:%106=32%:%
%:%107=33%:%
%:%110=36%:%
%:%111=37%:%
%:%112=38%:%
%:%113=38%:%
%:%116=39%:%
%:%120=39%:%
%:%121=39%:%
%:%122=40%:%
%:%123=40%:%
%:%128=40%:%
%:%131=41%:%
%:%132=42%:%
%:%133=43%:%
%:%134=43%:%
%:%135=44%:%
%:%136=45%:%
%:%137=45%:%
%:%138=46%:%
%:%139=47%:%
%:%140=48%:%
%:%141=49%:%
%:%143=51%:%
%:%146=52%:%
%:%150=52%:%
%:%151=52%:%
%:%152=52%:%
%:%153=53%:%
%:%154=53%:%
%:%159=53%:%
%:%162=54%:%
%:%163=55%:%
%:%164=55%:%
%:%165=56%:%
%:%166=57%:%
%:%167=58%:%
%:%168=59%:%
%:%169=60%:%
%:%172=61%:%
%:%176=61%:%
%:%177=61%:%
%:%178=62%:%
%:%179=62%:%
%:%184=62%:%
%:%187=63%:%
%:%188=64%:%
%:%189=64%:%
%:%192=65%:%
%:%196=65%:%
%:%197=65%:%
%:%198=65%:%
%:%203=65%:%
%:%206=66%:%
%:%207=67%:%
%:%208=68%:%
%:%209=68%:%
%:%212=69%:%
%:%216=69%:%
%:%217=69%:%
%:%218=70%:%
%:%219=70%:%
%:%224=70%:%
%:%227=71%:%
%:%228=72%:%
%:%229=72%:%
%:%230=73%:%
%:%237=75%:%
%:%247=77%:%
%:%248=77%:%
%:%249=78%:%
%:%250=79%:%
%:%253=80%:%
%:%257=80%:%
%:%258=80%:%
%:%259=81%:%
%:%260=81%:%
%:%261=82%:%
%:%262=82%:%
%:%263=83%:%
%:%264=83%:%
%:%265=84%:%
%:%266=84%:%
%:%267=84%:%
%:%268=85%:%
%:%269=85%:%
%:%270=86%:%
%:%271=86%:%
%:%272=86%:%
%:%273=86%:%
%:%274=87%:%
%:%275=87%:%
%:%276=88%:%
%:%277=88%:%
%:%278=89%:%
%:%279=89%:%
%:%280=90%:%
%:%281=90%:%
%:%282=91%:%
%:%283=91%:%
%:%284=92%:%
%:%285=92%:%
%:%286=93%:%
%:%287=93%:%
%:%288=93%:%
%:%289=94%:%
%:%290=94%:%
%:%291=95%:%
%:%292=95%:%
%:%293=96%:%
%:%294=96%:%
%:%295=97%:%
%:%296=97%:%
%:%297=97%:%
%:%298=98%:%
%:%299=98%:%
%:%300=99%:%
%:%301=99%:%
%:%302=100%:%
%:%303=100%:%
%:%304=101%:%
%:%305=101%:%
%:%306=102%:%
%:%307=102%:%
%:%308=103%:%
%:%309=103%:%
%:%310=104%:%
%:%311=104%:%
%:%312=105%:%
%:%313=105%:%
%:%314=106%:%
%:%315=106%:%
%:%316=107%:%
%:%317=107%:%
%:%318=108%:%
%:%319=108%:%
%:%320=109%:%
%:%321=109%:%
%:%322=110%:%
%:%323=110%:%
%:%324=110%:%
%:%325=111%:%
%:%326=111%:%
%:%327=112%:%
%:%328=112%:%
%:%329=113%:%
%:%330=113%:%
%:%331=114%:%
%:%332=114%:%
%:%333=115%:%
%:%334=115%:%
%:%335=115%:%
%:%336=116%:%
%:%337=116%:%
%:%338=117%:%
%:%339=117%:%
%:%340=118%:%
%:%341=118%:%
%:%342=118%:%
%:%343=119%:%
%:%344=119%:%
%:%345=120%:%
%:%346=120%:%
%:%347=120%:%
%:%348=120%:%
%:%349=121%:%
%:%350=121%:%
%:%351=122%:%
%:%352=122%:%
%:%353=123%:%
%:%354=123%:%
%:%355=123%:%
%:%356=124%:%
%:%362=124%:%
%:%365=125%:%
%:%366=126%:%
%:%367=126%:%
%:%368=127%:%
%:%369=128%:%
%:%372=129%:%
%:%376=129%:%
%:%377=129%:%
%:%378=129%:%
%:%383=129%:%
%:%386=130%:%
%:%387=131%:%
%:%388=131%:%
%:%391=132%:%
%:%396=133%:%