### 2024-11-01 code/output/document/Names.tex

%
\begin{isabellebody}%
\setisabellecontext{Names}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Names and generic extensions%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Names\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ Forcing{\isacharunderscore}{\kern0pt}Data\isanewline
\ \ \ \ Interface\isanewline
\ \ \ \ Recursion{\isacharunderscore}{\kern0pt}Thms\isanewline
\ \ \ \ Synthetic{\isacharunderscore}{\kern0pt}Definition\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ SepReplace\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isasymRightarrow}i{\isacharcomma}{\kern0pt}\ i{\isasymRightarrow}\ o{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}SepReplace{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}b{\isacharcomma}{\kern0pt}Q{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}y\ {\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ y{\isacharequal}{\kern0pt}b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{syntax}\isamarkupfalse%
\isanewline
\ \ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}SepReplace{\isachardoublequoteclose}\ \ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ pttrn{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ o{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharbraceleft}{\kern0pt}{\isacharunderscore}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isacharslash}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\isacommand{translations}\isamarkupfalse%
\isanewline
\ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}b\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ Q{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharequal}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isachardoublequoteopen}CONST\ SepReplace{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ b{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ Q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Sep{\isacharunderscore}{\kern0pt}and{\isacharunderscore}{\kern0pt}Replace{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x{\isasymin}{\isacharbraceleft}{\kern0pt}y{\isasymin}A{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}SepReplace{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ SepReplace{\isacharunderscore}{\kern0pt}subset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A{\isasymsubseteq}A{\isacharprime}{\kern0pt}{\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}b\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ Q{\isacharbraceright}{\kern0pt}{\isasymsubseteq}{\isacharbraceleft}{\kern0pt}b\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ Q{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}SepReplace{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ SepReplace{\isacharunderscore}{\kern0pt}iff\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y{\isasymin}{\isacharbraceleft}{\kern0pt}b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isasymin}A{\isachardot}{\kern0pt}\ y{\isacharequal}{\kern0pt}b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharampersand}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}SepReplace{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ SepReplace{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}implies\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}\ x\ {\isachardot}{\kern0pt}\ x\ {\isasymin}A\ {\isasymLongrightarrow}\ b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ b{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharequal}{\kern0pt}{\isacharbraceleft}{\kern0pt}b{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}SepReplace{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ SepReplace{\isacharunderscore}{\kern0pt}pred{\isacharunderscore}{\kern0pt}implies\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymlongrightarrow}\ b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ b{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}b{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharequal}{\kern0pt}{\isacharbraceleft}{\kern0pt}b{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}A{\isacharcomma}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ \ {\isacharparenleft}{\kern0pt}force\ simp\ add{\isacharcolon}{\kern0pt}SepReplace{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{The well-founded relation \isa{ed}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{lemma}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}sing\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetD{\isacharbrackleft}{\kern0pt}OF\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}subset{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}simp{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ecloseE\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ {\isachardoublequoteopen}x\ {\isasymin}\ A\ {\isasymor}\ {\isacharparenleft}{\kern0pt}{\isasymexists}\ B\ {\isasymin}\ A\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ rule{\isacharcolon}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}induct{\isacharunderscore}{\kern0pt}down{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}\ y{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}\ y\ z{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}y\ {\isasymin}\ A\ {\isasymor}\ {\isacharparenleft}{\kern0pt}{\isasymexists}B{\isasymin}A{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}inA{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}exB{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}B{\isasymin}A{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ inA\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{2}}\ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ exB\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ B\ \isakeyword{where}\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}B{\isasymin}A{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{2}}\ ecloseD{\isacharbrackleft}{\kern0pt}of\ y\ B\ z{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}singE\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isacharequal}{\kern0pt}\ a\ {\isasymor}\ x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}blast\ dest{\isacharcolon}{\kern0pt}\ ecloseE{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}sing\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}x{\isacharequal}{\kern0pt}a{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}lt{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}eclose{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}singE\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}sing\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}trans\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ z{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Pair{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ecloseD{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isacharcomma}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ ecloseD{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharcomma}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
term\isa{ed} is the well-founded relation on which \isa{val} is defined.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ ed\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}ed{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ edrel\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Rrel{\isacharparenleft}{\kern0pt}ed{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ edI{\isacharbrackleft}{\kern0pt}intro{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}t{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ ed{\isacharparenleft}{\kern0pt}t{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ edD{\isacharbrackleft}{\kern0pt}dest{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}ed{\isacharparenleft}{\kern0pt}t{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}ed{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ s\ \isakeyword{where}\ {\isachardoublequoteopen}y{\isasymin}s{\isachardoublequoteclose}\ {\isachardoublequoteopen}s{\isasymin}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Pair{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}s{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}s{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}lt\ \isacommand{by}\isamarkupfalse%
\ blast{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}trans\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ edrel{\isacharunderscore}{\kern0pt}dest\ {\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}\ a{\isasymin}\ A{\isachardot}{\kern0pt}\ {\isasymexists}\ b\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}ed{\isacharunderscore}{\kern0pt}def\ edrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ edrelD\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}\ a{\isasymin}\ A{\isachardot}{\kern0pt}\ {\isasymexists}\ b\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}\ {\isasymand}\ a\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}ed{\isacharunderscore}{\kern0pt}def\ edrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ edrelI\ {\isacharbrackleft}{\kern0pt}intro{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}A\ {\isasymLongrightarrow}\ y{\isasymin}A\ {\isasymLongrightarrow}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}ed{\isacharunderscore}{\kern0pt}def\ edrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ edrel{\isacharunderscore}{\kern0pt}trans{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y{\isasymin}A\ {\isasymLongrightarrow}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ edrelI{\isacharcomma}{\kern0pt}\ auto\ simp\ add{\isacharcolon}{\kern0pt}Transset{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ Pair{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}trans{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y{\isasymin}A\ {\isasymLongrightarrow}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x{\isasymin}A{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ Transset{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ Pair{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}edrel\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ relation{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}edrel\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymsubseteq}A{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ edrel{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}memrel{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ trancl{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ z\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}z{\isasymin}edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ y\ \isakeyword{where}\isanewline
\ \ \ \ Eq{\isadigit{1}}{\isacharcolon}{\kern0pt}\ \ \ {\isachardoublequoteopen}x{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isacharequal}{\kern0pt}{\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}domain{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ edrelD\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ u\ v\ \isakeyword{where}\isanewline
\ \ \ \ Eq{\isadigit{2}}{\isacharcolon}{\kern0pt}\ \ \ {\isachardoublequoteopen}x{\isasymin}u{\isachardoublequoteclose}\ {\isachardoublequoteopen}u{\isasymin}v{\isachardoublequoteclose}\ {\isachardoublequoteopen}v{\isasymin}y{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}def\ Pair{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ Eq{\isadigit{1}}\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ Eq{\isadigit{3}}{\isacharcolon}{\kern0pt}\ \ \ {\isachardoublequoteopen}x{\isasymin}eclose{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}eclose{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}u{\isasymin}eclose{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v{\isasymin}eclose{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ {\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharminus}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\ ecloseD{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ {\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\ ecloseD{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all\ add{\isacharcolon}{\kern0pt}arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}trancl{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ Eq{\isadigit{2}}\ \isakeyword{and}\ Eq{\isadigit{3}}\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}u{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}r{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}u{\isacharcomma}{\kern0pt}v{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}r{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}v{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}r{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ r{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}r{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ trancl{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ {\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ trancl{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ Eq{\isadigit{1}}\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}{\isacharquery}{\kern0pt}r{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}edrel\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}wf{\isacharparenleft}{\kern0pt}edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}subset\ {\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}trancl{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ edrel{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}memrel\isanewline
\ \ \ \ wf{\isacharunderscore}{\kern0pt}trancl\ wf{\isacharunderscore}{\kern0pt}Memrel\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}induction{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ {\isasymlbrakk}{\isasymAnd}y{\isachardot}{\kern0pt}\ \ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}induct\ rule{\isacharcolon}{\kern0pt}\ wf{\isacharunderscore}{\kern0pt}induct{\isadigit{2}}{\isacharbrackleft}{\kern0pt}OF\ wf{\isacharunderscore}{\kern0pt}edrel{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ {\isacharcomma}{\kern0pt}of\ a\ {\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{2}}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}edrel\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{3}}\ x{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isacharbrackleft}{\kern0pt}of\ x{\isacharbrackright}{\kern0pt}\ edrelI\ domain{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ Transset{\isacharunderscore}{\kern0pt}eclose\ {\isadigit{3}}{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ dom{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}eclose{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ edrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ ed{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ edrel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ eclose{\isacharunderscore}{\kern0pt}sing\ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ed{\isacharunderscore}{\kern0pt}eclose\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}z{\isasymrangle}\ {\isasymin}\ edrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}drule\ edrelD{\isacharcomma}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}domain{\isacharunderscore}{\kern0pt}def\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}eclose\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}z{\isasymrangle}\ {\isasymin}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ trancl{\isacharunderscore}{\kern0pt}induct{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ ed{\isacharunderscore}{\kern0pt}eclose\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ restrict{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}z\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasyminter}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymtimes}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}intro\ equalityI\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}ec{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}\ y\ {\isachardot}{\kern0pt}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}ez{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}rr{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}ec{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasyminter}\ {\isacharquery}{\kern0pt}ez\ {\isasymtimes}\ {\isacharquery}{\kern0pt}ez{\isachardoublequoteclose}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ y\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ yr{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}y\ {\isasymin}\ {\isacharquery}{\kern0pt}rr{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ yr\ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}\ {\isasymin}\ {\isacharquery}{\kern0pt}rr{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymin}\ {\isacharquery}{\kern0pt}ez{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ {\isacharquery}{\kern0pt}ez{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}\ {\isasymin}\ {\isacharquery}{\kern0pt}ec{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isacharequal}{\kern0pt}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ edrelD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}ec{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}\ y\ {\isachardot}{\kern0pt}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}ez{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}rr{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}ec{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasyminter}\ {\isacharquery}{\kern0pt}ez\ {\isasymtimes}\ {\isacharquery}{\kern0pt}ez{\isachardoublequoteclose}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ y\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ yr{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}y\ {\isasymin}\ edrel{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}ez{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ {\isachardoublequoteopen}a\ {\isasymin}\ {\isacharquery}{\kern0pt}ez{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ {\isacharquery}{\kern0pt}ez{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isacharequal}{\kern0pt}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ edrelD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ this\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}sing\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ this\ {\isacartoucheopen}a{\isasymin}domain{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}\ {\isasymin}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ {\isacharquery}{\kern0pt}rr{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}subset\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}z\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ \ {\isachardoublequoteopen}tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}intro\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}\ x\ {\isachardot}{\kern0pt}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ y\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ \ {\isachardoublequoteopen}y\ {\isasymin}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}z{\isasymrangle}\ {\isasymin}\ {\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}downD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}eclose\ eclose{\isacharunderscore}{\kern0pt}sing\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ upairM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ y\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ singletonM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Rep{\isacharunderscore}{\kern0pt}simp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Replace{\isacharparenleft}{\kern0pt}u{\isacharcomma}{\kern0pt}{\isasymlambda}\ y\ z\ {\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ u{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Values and check-names%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ Hcheck\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Hcheck{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ \ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isasymlangle}f{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ z{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ check\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ transrec{\isacharparenleft}{\kern0pt}x\ {\isacharcomma}{\kern0pt}\ Hcheck{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ checkD{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ wfrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hcheck{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}def\ transrec{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ rcheck\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Hcheck{\isacharunderscore}{\kern0pt}trancl{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Hcheck{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharequal}{\kern0pt}\ Hcheck{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Hcheck{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ restrict{\isacharunderscore}{\kern0pt}trans{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}trancl{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hcheck{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ checkD\ wf{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}trancl\ Hcheck{\isacharunderscore}{\kern0pt}trancl\ \isacommand{unfolding}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ aux{\isacharunderscore}{\kern0pt}def{\isacharunderscore}{\kern0pt}check{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ y\ {\isasymLongrightarrow}\isanewline
\ \ wfrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hcheck{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ wfrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hcheck{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ wfrec{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\ eclose{\isacharunderscore}{\kern0pt}sing{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}check\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isasymlangle}check{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isachardot}{\kern0pt}\ w\ {\isasymin}\ y{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}y{\isachardot}{\kern0pt}\ Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ wfr{\isacharcolon}{\kern0pt}\ \ \ {\isachardoublequoteopen}{\isasymforall}w\ {\isachardot}{\kern0pt}\ wf{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}Memrel\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}\ Hcheck{\isacharparenleft}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isasymin}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hcheck{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ wfrec{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ y\ {\isachardoublequoteopen}Hcheck{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ checkD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Hcheck{\isacharparenleft}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isasymin}y{\isachardot}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hcheck{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ under{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}eclose\ arg{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Hcheck{\isacharparenleft}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isasymin}y{\isachardot}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ aux{\isacharunderscore}{\kern0pt}def{\isacharunderscore}{\kern0pt}check\ checkD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ Hcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}checkS\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ n\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isachardot}{\kern0pt}\ i\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}check\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isachardot}{\kern0pt}\ i\ {\isasymin}\ n{\isacharbraceright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}check{\isacharbrackleft}{\kern0pt}of\ n{\isacharcomma}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}Memrel{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ordinal{\isachardot}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}type\ field{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}subset\ assms\ eclose{\isacharunderscore}{\kern0pt}least{\isacharbrackleft}{\kern0pt}OF\ trans{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ subset{\isacharunderscore}{\kern0pt}trans\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ Hv\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Hv{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}y\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ y{\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ x\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The funcion \isa{val} interprets a name in \isa{M}
according to a (generic) filter \isa{G}. Note the definition
in terms of the well-founded recursor.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ val\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ wfrec{\isacharparenleft}{\kern0pt}edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isasymtau}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymtau}\ {\isacharcomma}{\kern0pt}Hv{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ aux{\isacharunderscore}{\kern0pt}def{\isacharunderscore}{\kern0pt}val{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}z\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}wfrec{\isacharparenleft}{\kern0pt}edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}Hv{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}Hv{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x\ {\isachardot}{\kern0pt}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ arg{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}sing\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}edrel\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}wf{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}edrel\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}subset\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}Hv{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}Hv{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}restr\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}z{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}Hv{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ restrict{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}eq\ wfrec{\isacharunderscore}{\kern0pt}restr{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
The next lemma provides the usual recursive expresion for the definition of term\isa{val}.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}val{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ \ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}x\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}{\isasymtau}\ {\isachardot}{\kern0pt}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isasymtau}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}f{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}z{\isasymin}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}Hv{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}{\isasymtau}{\isachardot}{\kern0pt}\ wf{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}edrel\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ wfrec\ {\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Hv{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Hv{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymlambda}z{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}Hv{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ dom{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Hv{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymlambda}z{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ aux{\isacharunderscore}{\kern0pt}def{\isacharunderscore}{\kern0pt}val\ val{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ Hv{\isacharunderscore}{\kern0pt}def\ SepReplace{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymsubseteq}y\ {\isasymLongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}val{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Check-names are the canonical names for elements of the
ground model. Here we show that this is the case.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ valcheck\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}one\ {\isasymin}\ G\ {\isasymLongrightarrow}\ \ one\ {\isasymin}\ P\ {\isasymLongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ rule{\isacharcolon}{\kern0pt}eps{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}\ y{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \ \ \ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isasymlangle}check{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isasymrangle}\ {\isachardot}{\kern0pt}\ w\ {\isasymin}\ y{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}C{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}check\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isasymrangle}\ {\isachardot}{\kern0pt}\ w\ {\isasymin}\ y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ \ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}C{\isacharparenright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ \ {\isasymlangle}t{\isacharcomma}{\kern0pt}\ p{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}C\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}C{\isacharparenright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isasymexists}w{\isasymin}y{\isachardot}{\kern0pt}\ t{\isacharequal}{\kern0pt}check{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}check{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ w{\isasymin}y\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}A{\isasymtimes}P{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}A\ {\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ \ Q{\isacharparenleft}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}A{\isasymtimes}P{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}{\isasymtau}\ {\isachardot}{\kern0pt}\ edrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isasymtau}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}f{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}z{\isasymin}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ wfR\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}wf{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymtau}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ wf{\isacharunderscore}{\kern0pt}edrel{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ A{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\ \isacommand{fix}\isamarkupfalse%
\ t\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}t\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ A\ {\isasymtimes}\ P\ {\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ t\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}if\ t\ {\isasymin}\ {\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharbraceright}{\kern0pt}\ then\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ else\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ dom{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}eclose\ H\ if{\isacharunderscore}{\kern0pt}P\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}t\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ A\ {\isasymtimes}\ P\ {\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}f{\isacharbackquote}{\kern0pt}\ t{\isachardoublequoteclose}\ \ \isakeyword{for}\ t\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ P\ {\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ {\isacharquery}{\kern0pt}n\ {\isasymand}\ p\ {\isasymin}\ G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ def{\isacharunderscore}{\kern0pt}val{\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharbackquote}{\kern0pt}t\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}n\ {\isasymand}\ p{\isasymin}G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Hv{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ SepReplace{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}implies{\isacharcomma}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}Eq{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharparenleft}{\kern0pt}if\ t{\isasymin}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharbraceright}{\kern0pt}\ then\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ else\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}n\ {\isasymand}\ p{\isasymin}G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{2}}{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}n\ {\isasymand}\ p{\isasymin}G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ dom{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}edrel{\isacharunderscore}{\kern0pt}eclose\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}if\ t{\isasymin}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharbraceright}{\kern0pt}\ then\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ else\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ {\isacharparenleft}{\kern0pt}if\ t{\isasymin}{\isacharquery}{\kern0pt}r{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharbraceright}{\kern0pt}\ then\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ else\ {\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}n\ {\isasymand}\ p{\isasymin}G{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}n\ {\isasymand}\ p{\isasymin}G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}A{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}n\ {\isasymand}\ p{\isasymin}G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}\ \ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}A{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isasymand}\ p{\isasymin}G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}alt\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}A{\isasymtimes}P{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}A\ {\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isasyminter}G\ {\isachardot}{\kern0pt}\ \ Q{\isacharparenleft}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}only{\isacharunderscore}{\kern0pt}names{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}name{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}name{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ t{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}name{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ {\isacharquery}{\kern0pt}name\ {\isasymand}\ p\ {\isasymin}\ F{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ t{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}{\isacharbraceleft}{\kern0pt}y{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}name{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ {\isacharquery}{\kern0pt}name\ {\isasymand}\ p\ {\isasymin}\ F{\isacharbraceright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Sep{\isacharunderscore}{\kern0pt}and{\isacharunderscore}{\kern0pt}Replace\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ t{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}{\isacharbraceleft}{\kern0pt}y{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ {\isasymtau}\ {\isasymand}\ p\ {\isasymin}\ F{\isacharbraceright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ t{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ {\isasymtau}\ {\isasymand}\ p\ {\isasymin}\ F{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Sep{\isacharunderscore}{\kern0pt}and{\isacharunderscore}{\kern0pt}Replace\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ {\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}val{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}only{\isacharunderscore}{\kern0pt}pairs{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}name{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}only{\isacharunderscore}{\kern0pt}names\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}name\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}times{\isacharunderscore}{\kern0pt}range{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}range{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}only{\isacharunderscore}{\kern0pt}pairs{\isacharbrackleft}{\kern0pt}THEN\ equalityD{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ val{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ {\isasymtau}\ {\isachardot}{\kern0pt}\ {\isasymexists}t\ p{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}\ p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}range{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}times{\isacharunderscore}{\kern0pt}P{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}only{\isacharunderscore}{\kern0pt}names{\isacharbrackleft}{\kern0pt}of\ F\ {\isasymtau}{\isacharbrackright}{\kern0pt}\ val{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}{\isasymtau}{\isachardot}{\kern0pt}\ {\isasymexists}t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isachardoublequoteclose}\ F{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ GenExt\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}GenExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isasymequiv}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymtau}\ {\isasymin}\ M{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}elem{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ {\isasympi}\ {\isasymLongrightarrow}\ p{\isasymin}G\ {\isasymLongrightarrow}\ p{\isasymin}P\ {\isasymLongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ {\isasympi}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtheta}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtheta}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ {\isasympi}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ \ {\isasymlangle}t{\isacharcomma}{\kern0pt}\ p{\isasymrangle}{\isasymin}{\isasympi}\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ def{\isacharunderscore}{\kern0pt}val{\isacharparenright}{\kern0pt}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}{\isasymtheta}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}asm{\isacharparenright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}val{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}{\isasymtheta}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ \ {\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isasympi}\ {\isasymand}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}asm{\isacharparenright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}val{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasympi}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}G{\isachardot}{\kern0pt}\ \ {\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isasympi}\ {\isasymand}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtheta}\ p\ \isakeyword{where}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isasympi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtheta}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ THEN\ conjunct{\isadigit{1}}{\isacharcomma}{\kern0pt}\ simplified{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ transitivity\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ GenExtD{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}{\isasymtau}{\isasymin}M{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}GenExt{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ GenExtI{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ GenExt{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\ \isacommand{fix}\isamarkupfalse%
\ vc\ y\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}vc\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}y\ {\isasymin}\ vc{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ c\ \isakeyword{where}\ {\isachardoublequoteopen}c{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}y\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtheta}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymtheta}{\isasymin}domain{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ trans{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}c{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}trans\ GenExtI\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ {\isacharequal}{\kern0pt}\ Transset{\isacharunderscore}{\kern0pt}intf{\isacharbrackleft}{\kern0pt}OF\ Transset{\isacharunderscore}{\kern0pt}MG{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ n\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ n{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{0}}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ def{\isacharunderscore}{\kern0pt}check{\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}succ\ x{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}one\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ P{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}M\ subsetD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ tuples{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ singletonM\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ def{\isacharunderscore}{\kern0pt}checkS\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ PHcheck\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}PHcheck{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ p{\isasymin}M\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}fy{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}fy{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}fy{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}Hcheck\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}hc{\isacharparenright}{\kern0pt}\ \ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}PHcheck{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}hc{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}one\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}\ transitivity{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}PHcheck{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}Hcheck{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Replace{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}PHcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}f{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}z{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ tuples{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transitivity\ that\ apply{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}y\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ z{\isacharcomma}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ one{\isasymrangle}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isacharbraceleft}{\kern0pt}y\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ z{\isacharcomma}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ one{\isasymrangle}\ {\isasymand}\ y{\isasymin}M\ {\isasymand}\ f{\isacharbackquote}{\kern0pt}x{\isasymin}M{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}z{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}f{\isasymin}M{\isacartoucheclose}\ transitivity\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Hcheck{\isacharunderscore}{\kern0pt}def\ PHcheck{\isacharunderscore}{\kern0pt}def\ RepFun{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ PHcheck{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharcomma}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}o{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ PHcheck{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ z\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ u\ {\isasymin}\ nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}PHcheck{\isacharunderscore}{\kern0pt}fm\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ z\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ u\ {\isasymin}\ nat\ {\isacharsemicolon}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isasymrbrakk}\isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ PHcheck{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}u{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Internalizations{\isachardot}{\kern0pt}nth{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ PHcheck{\isacharunderscore}{\kern0pt}def\ PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}hc{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Replace{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}o{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}hc{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ z\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ u\ {\isasymin}\ nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ z\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ u\ {\isasymin}\ nat\ {\isacharsemicolon}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isasymrbrakk}\isanewline
\ \ \ \ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}u{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}Replace{\isacharunderscore}{\kern0pt}fm\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}Hcheck\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}X{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}wfrec{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharcomma}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{8}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}c{\isacharcomma}{\kern0pt}b{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}d{\isacharcomma}{\kern0pt}e{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}a{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}c{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}d{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}e{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{for}\ a\ b\ c\ d\ e\ y\ x\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}X{\isasymin}M{\isacartoucheclose}\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{8}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharbrackleft}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}wfrec{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ x\ y\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ that\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm\ {\isacartoucheopen}X{\isasymin}M{\isacartoucheclose}\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}f{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{8}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ satsf{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}f{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}y{\isasymin}M{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isacharampersand}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ x\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ {\isadigit{1}}\ {\isacartoucheopen}X{\isasymin}M{\isacartoucheclose}\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}pair{\isacharunderscore}{\kern0pt}abs{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ artyf{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{4}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Replace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ pair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ restriction{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x\ z{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ replacement{\isacharunderscore}{\kern0pt}ax\ {\isadigit{1}}\ artyf\ {\isacartoucheopen}X{\isasymin}M{\isacartoucheclose}\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x\ z{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isasymexists}y{\isasymin}M{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isacharampersand}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ repl{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}of\ M\ {\isacharquery}{\kern0pt}f\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}one{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}X{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ satsf\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}pair{\isacharunderscore}{\kern0pt}abs{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ repl{\isacharunderscore}{\kern0pt}PHcheck\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}f{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}PHcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{4}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ upair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}f{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ replacement{\isacharunderscore}{\kern0pt}ax\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}f{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{unfolding}\isamarkupfalse%
\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}def\ univalent{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ univ{\isacharunderscore}{\kern0pt}PHcheck\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ z{\isasymin}M\ {\isacharsemicolon}{\kern0pt}\ f{\isasymin}M\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}PHcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ PHcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}Hcheck\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}relation{\isadigit{2}}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Hcheck{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}z{\isacharsemicolon}{\kern0pt}\ PHcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ z\ f\ x\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{unfolding}\isamarkupfalse%
\ PHcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}PHcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}hc{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ hc\ {\isacharequal}{\kern0pt}\ Replace{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}PHcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}hc{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ z\ f\ hc\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ Replace{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ univ{\isacharunderscore}{\kern0pt}PHcheck\ {\isadigit{1}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}PHcheck\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}def\ Hcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ PHcheck{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}z{\isasymin}M\ {\isacharsemicolon}{\kern0pt}\ f{\isasymin}M\ {\isacharsemicolon}{\kern0pt}\ x{\isasymin}z{\isacharsemicolon}{\kern0pt}\ PHcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ PHcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Hcheck{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymforall}y{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymforall}g{\isasymin}M{\isachardot}{\kern0pt}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Hcheck{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Replace{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}PHcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}f{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ f\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ repl{\isacharunderscore}{\kern0pt}PHcheck\ \ PHcheck{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}of\ y\ f{\isacharbrackright}{\kern0pt}\ univ{\isacharunderscore}{\kern0pt}PHcheck\isanewline
\ \ \ \ \ \ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}PHcheck\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}rcheck\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}M\ {\isasymLongrightarrow}\ wf{\isacharparenleft}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}trancl{\isacharbrackleft}{\kern0pt}OF\ wf{\isacharunderscore}{\kern0pt}Memrel{\isacharbrackright}{\kern0pt}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ trans{\isacharunderscore}{\kern0pt}rcheck\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}M\ {\isasymLongrightarrow}\ trans{\isacharparenleft}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ trans{\isacharunderscore}{\kern0pt}trancl\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}rcheck\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}M\ {\isasymLongrightarrow}\ relation{\isacharparenleft}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}trancl\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}M\ {\isasymLongrightarrow}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ transrec{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}Hcheck{\isacharbrackleft}{\kern0pt}of\ x{\isacharbrackright}{\kern0pt}\ check{\isacharunderscore}{\kern0pt}trancl\ wf{\isacharunderscore}{\kern0pt}rcheck\ trans{\isacharunderscore}{\kern0pt}rcheck\ relation{\isacharunderscore}{\kern0pt}rcheck\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ Hcheck{\isacharunderscore}{\kern0pt}closed\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}Hcheck\ trans{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ x\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ Hcheck{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}singleton\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}singleton{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}c{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ empty{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}cons{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}c{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}trivial{\isacharparenright}{\kern0pt}\ singleton{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}s{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}singleton{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}s{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ s\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ nonempty\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ singleton{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}singleton{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ cons{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ singleton{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ x\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ y\ {\isasymin}\ nat\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ singleton{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ singleton{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharsemicolon}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isacharsemicolon}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ i\ {\isasymin}\ nat{\isacharsemicolon}{\kern0pt}\ j{\isasymin}nat\ {\isacharsemicolon}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isasymrbrakk}\isanewline
\ \ \ \ \ \ \ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}singleton{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ singleton{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}def\ singleton{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data\ \isakeyword{begin}\isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}rcheck\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}r{\isasymin}M{\isachardot}{\kern0pt}\ tran{\isacharunderscore}{\kern0pt}closure{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}ec{\isasymin}M{\isachardot}{\kern0pt}\ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}ec{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}s{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}singleton{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}s{\isacharparenright}{\kern0pt}\ {\isasymand}\ \ is{\isacharunderscore}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}s{\isacharcomma}{\kern0pt}ec{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}M\ {\isacharsemicolon}{\kern0pt}\ r{\isasymin}M\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ r\ {\isacharequal}{\kern0pt}\ rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}rcheck{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ singletonM\ trancl{\isacharunderscore}{\kern0pt}closed\ Memrel{\isacharunderscore}{\kern0pt}closed\ eclose{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ rcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}rch{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}rcheck{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ is{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\isanewline
\ \ \ \ \ \ \ \ trans{\isacharunderscore}{\kern0pt}closure{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}rcheck{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ rcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}check\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}rch{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}rch{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rch{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ z\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ rcheck{\isacharunderscore}{\kern0pt}abs\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}trancl\ is{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}trancl\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ wfrec{\isacharunderscore}{\kern0pt}Hcheck{\isacharbrackleft}{\kern0pt}of\ x{\isacharbrackright}{\kern0pt}\ wf{\isacharunderscore}{\kern0pt}rcheck\ trans{\isacharunderscore}{\kern0pt}rcheck\ relation{\isacharunderscore}{\kern0pt}rcheck\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ Hcheck{\isacharunderscore}{\kern0pt}closed\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}Hcheck\ trans{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}rcheck{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ x\ z\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ Hcheck{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ check{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}rcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}x{\isasymin}nat{\isacharsemicolon}{\kern0pt}o{\isasymin}nat{\isacharsemicolon}{\kern0pt}z{\isasymin}nat{\isasymrbrakk}\ {\isasymLongrightarrow}\ check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ one{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}o{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}z\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}a{\isadigit{0}}\ a{\isadigit{1}}\ a{\isadigit{2}}\ a{\isadigit{3}}\ a{\isadigit{4}}{\isachardot}{\kern0pt}\ {\isasymlbrakk}\ a{\isadigit{0}}{\isasymin}M{\isacharsemicolon}{\kern0pt}\ a{\isadigit{1}}{\isasymin}M{\isacharsemicolon}{\kern0pt}\ a{\isadigit{2}}{\isasymin}M{\isacharsemicolon}{\kern0pt}\ a{\isadigit{3}}{\isasymin}M{\isacharsemicolon}{\kern0pt}\ a{\isadigit{4}}{\isasymin}M\ {\isasymrbrakk}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}\ a{\isadigit{1}}{\isacharcomma}{\kern0pt}\ a{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isadigit{0}}{\isacharcomma}{\kern0pt}a{\isadigit{1}}{\isacharcomma}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}a{\isadigit{3}}{\isacharcomma}{\kern0pt}a{\isadigit{4}}{\isacharcomma}{\kern0pt}r{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}r{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ r\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hcheck{\isacharparenleft}{\kern0pt}one{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}r{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ r\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ assms\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}def\ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ rcheck{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ rcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}replacement{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}P{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{3}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ rcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ trans{\isacharunderscore}{\kern0pt}closure{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Replace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ finite{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ restriction{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}case{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ quasinat{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Memrel{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ singleton{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ fm{\isacharunderscore}{\kern0pt}defs\ iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}P{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ Transset{\isacharunderscore}{\kern0pt}intf{\isacharbrackleft}{\kern0pt}of\ M\ x\ P{\isacharbrackright}{\kern0pt}\ trans{\isacharunderscore}{\kern0pt}M\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm\ check{\isacharunderscore}{\kern0pt}abs\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ Repl{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}one{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ is{\isacharunderscore}{\kern0pt}check\ check{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}check\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ p{\isasymin}M\ {\isacharsemicolon}{\kern0pt}\ y{\isasymin}M\ {\isasymrbrakk}\ \ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}c{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}c{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ y\ {\isacharequal}{\kern0pt}\ {\isasymlangle}check{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}abs\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ tuples{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}one\ {\isasymin}\ G\ {\isasymLongrightarrow}\ M\ {\isasymsubseteq}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ GenExtI\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ subsetI{\isacharcomma}{\kern0pt}\ subst\ valcheck\ {\isacharbrackleft}{\kern0pt}of\ G{\isacharcomma}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
The name for the generic filter%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ G{\isacharunderscore}{\kern0pt}dot\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}G{\isacharunderscore}{\kern0pt}dot\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}check{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isachardot}{\kern0pt}\ p{\isasymin}P{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}dot{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}G{\isacharunderscore}{\kern0pt}dot\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}is{\isacharunderscore}{\kern0pt}pcheck\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ {\isasymexists}ch{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}check{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}ch{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}ch{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}pcheck{\isacharunderscore}{\kern0pt}fm\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}pcheck{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharquery}{\kern0pt}is{\isacharunderscore}{\kern0pt}pcheck{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ x\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm\ that\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}is{\isacharunderscore}{\kern0pt}pcheck{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ y\ {\isacharequal}{\kern0pt}\ {\isasymlangle}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isasymrangle}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ x\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ check{\isacharunderscore}{\kern0pt}abs\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}pcheck{\isacharunderscore}{\kern0pt}fm{\isasymin}formula{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}pcheck{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}{\isadigit{3}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ rcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ trans{\isacharunderscore}{\kern0pt}closure{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Replace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ finite{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ restriction{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}case{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ quasinat{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Memrel{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ singleton{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ fm{\isacharunderscore}{\kern0pt}defs\ iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ tuples{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}check{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ \isakeyword{for}\ p\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}dot{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Repl{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}pcheck{\isacharunderscore}{\kern0pt}fm\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}one{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}dot\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}G\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}one\ {\isasymin}\ G{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}G{\isacharunderscore}{\kern0pt}dot{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ equalityI\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}G{\isacharunderscore}{\kern0pt}dot{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtheta}\ p\ \isakeyword{where}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ G{\isacharunderscore}{\kern0pt}dot{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymtheta}\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}dot{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair\ G{\isacharunderscore}{\kern0pt}dot{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}one{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}G{\isasymsubseteq}P{\isacartoucheclose}\ \isacommand{show}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ G{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ valcheck\ P{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ p\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}check{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ G{\isacharunderscore}{\kern0pt}dot{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ \isakeyword{for}\ q\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}dot{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}G{\isasymsubseteq}P{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}check{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}G{\isacharunderscore}{\kern0pt}dot{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}elem\ G{\isacharunderscore}{\kern0pt}dot{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}G{\isasymsubseteq}P{\isacartoucheclose}\ {\isacartoucheopen}one{\isasymin}G{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}G{\isacharunderscore}{\kern0pt}dot{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}sub{\isacharunderscore}{\kern0pt}M\ valcheck\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Gen{\isacharunderscore}{\kern0pt}Ext\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}G\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}one\ {\isasymin}\ G{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ \ {\isachardoublequoteopen}G\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ val{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}dot\ GenExtI{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ G{\isacharbrackright}{\kern0pt}\ G{\isacharunderscore}{\kern0pt}dot{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ fst{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}M\ {\isasymLongrightarrow}\ fst{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ M\ {\isasymand}\ snd{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}{\isasymexists}a{\isachardot}{\kern0pt}\ {\isasymexists}b{\isachardot}{\kern0pt}\ p\ {\isacharequal}{\kern0pt}\ {\isasymlangle}a{\isacharcomma}{\kern0pt}\ b{\isasymrangle}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}fst{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ M\ {\isasymand}\ snd{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ fst{\isacharunderscore}{\kern0pt}def\ snd{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ {\isachardoublequoteopen}p\ {\isacharequal}{\kern0pt}\ {\isasymlangle}a{\isacharcomma}{\kern0pt}\ b{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}fst{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}snd{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ b{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ fst{\isacharunderscore}{\kern0pt}def\ snd{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}a{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isacartoucheopen}p\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ Pair{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force\ intro{\isacharcolon}{\kern0pt}Transset{\isacharunderscore}{\kern0pt}M{\isacharbrackleft}{\kern0pt}OF\ trans{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ \ {\isacartoucheopen}p{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}M{\isacharbrackleft}{\kern0pt}OF\ trans{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ of\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ p{\isacharbrackright}{\kern0pt}\ Transset{\isacharunderscore}{\kern0pt}M{\isacharbrackleft}{\kern0pt}OF\ trans{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ of\ {\isachardoublequoteopen}b{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isacartoucheopen}p\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ Pair{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\ {\isacharequal}{\kern0pt}\ forcing{\isacharunderscore}{\kern0pt}data\ {\isacharplus}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ G\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ generic\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtI\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}nonempty{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}G{\isasymnoteq}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isasymsubseteq}P{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}dense\ {\isacartoucheopen}P{\isasymsubseteq}P{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}G\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic\ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Names.thy%:%
%:%11=1%:%
%:%27=3%:%
%:%28=3%:%
%:%29=4%:%
%:%30=5%:%
%:%31=6%:%
%:%32=7%:%
%:%33=8%:%
%:%34=9%:%
%:%39=9%:%
%:%42=10%:%
%:%43=11%:%
%:%44=11%:%
%:%45=12%:%
%:%46=13%:%
%:%47=14%:%
%:%48=15%:%
%:%49=15%:%
%:%50=16%:%
%:%51=17%:%
%:%52=17%:%
%:%53=18%:%
%:%54=19%:%
%:%55=20%:%
%:%56=20%:%
%:%59=21%:%
%:%63=21%:%
%:%64=21%:%
%:%69=21%:%
%:%72=22%:%
%:%73=23%:%
%:%74=23%:%
%:%77=24%:%
%:%81=24%:%
%:%82=24%:%
%:%87=24%:%
%:%90=25%:%
%:%91=26%:%
%:%92=26%:%
%:%95=27%:%
%:%99=27%:%
%:%100=27%:%
%:%105=27%:%
%:%108=28%:%
%:%109=29%:%
%:%110=29%:%
%:%111=30%:%
%:%114=31%:%
%:%118=31%:%
%:%119=31%:%
%:%124=31%:%
%:%127=32%:%
%:%128=33%:%
%:%129=33%:%
%:%130=34%:%
%:%133=35%:%
%:%137=35%:%
%:%138=35%:%
%:%152=37%:%
%:%162=39%:%
%:%163=39%:%
%:%166=40%:%
%:%170=40%:%
%:%171=40%:%
%:%176=40%:%
%:%179=41%:%
%:%180=42%:%
%:%181=42%:%
%:%182=43%:%
%:%183=44%:%
%:%186=45%:%
%:%190=45%:%
%:%191=45%:%
%:%192=46%:%
%:%193=46%:%
%:%194=47%:%
%:%195=47%:%
%:%196=48%:%
%:%197=48%:%
%:%198=49%:%
%:%199=49%:%
%:%200=50%:%
%:%201=50%:%
%:%202=50%:%
%:%203=51%:%
%:%204=51%:%
%:%205=52%:%
%:%206=52%:%
%:%207=53%:%
%:%208=53%:%
%:%209=54%:%
%:%210=54%:%
%:%211=55%:%
%:%212=55%:%
%:%213=56%:%
%:%214=56%:%
%:%215=56%:%
%:%216=57%:%
%:%217=57%:%
%:%218=58%:%
%:%219=58%:%
%:%220=59%:%
%:%221=59%:%
%:%222=60%:%
%:%223=60%:%
%:%224=60%:%
%:%225=60%:%
%:%226=61%:%
%:%227=61%:%
%:%228=62%:%
%:%229=62%:%
%:%230=63%:%
%:%231=63%:%
%:%232=63%:%
%:%233=64%:%
%:%234=64%:%
%:%235=65%:%
%:%236=65%:%
%:%237=66%:%
%:%238=66%:%
%:%239=66%:%
%:%240=66%:%
%:%241=67%:%
%:%242=67%:%
%:%243=68%:%
%:%249=68%:%
%:%252=69%:%
%:%253=70%:%
%:%254=70%:%
%:%257=71%:%
%:%261=71%:%
%:%262=71%:%
%:%267=71%:%
%:%270=72%:%
%:%271=73%:%
%:%272=73%:%
%:%273=74%:%
%:%274=75%:%
%:%281=76%:%
%:%282=76%:%
%:%283=77%:%
%:%284=77%:%
%:%285=78%:%
%:%286=78%:%
%:%287=79%:%
%:%288=79%:%
%:%289=79%:%
%:%290=80%:%
%:%291=80%:%
%:%292=81%:%
%:%293=81%:%
%:%294=82%:%
%:%295=82%:%
%:%296=83%:%
%:%297=83%:%
%:%298=84%:%
%:%304=84%:%
%:%307=85%:%
%:%308=86%:%
%:%309=86%:%
%:%310=87%:%
%:%311=88%:%
%:%318=89%:%
%:%319=89%:%
%:%320=90%:%
%:%321=90%:%
%:%322=91%:%
%:%323=91%:%
%:%324=92%:%
%:%325=92%:%
%:%326=92%:%
%:%327=93%:%
%:%328=93%:%
%:%329=94%:%
%:%330=94%:%
%:%331=95%:%
%:%332=95%:%
%:%333=96%:%
%:%334=96%:%
%:%335=97%:%
%:%336=97%:%
%:%337=98%:%
%:%347=100%:%
%:%349=101%:%
%:%350=101%:%
%:%351=102%:%
%:%352=103%:%
%:%353=104%:%
%:%354=105%:%
%:%355=105%:%
%:%356=106%:%
%:%357=107%:%
%:%358=108%:%
%:%359=109%:%
%:%360=110%:%
%:%361=110%:%
%:%364=111%:%
%:%368=111%:%
%:%369=111%:%
%:%375=111%:%
%:%378=112%:%
%:%379=113%:%
%:%380=113%:%
%:%383=114%:%
%:%387=114%:%
%:%388=114%:%
%:%394=114%:%
%:%397=115%:%
%:%398=116%:%
%:%399=117%:%
%:%400=117%:%
%:%401=118%:%
%:%402=119%:%
%:%409=120%:%
%:%410=120%:%
%:%411=121%:%
%:%412=121%:%
%:%413=122%:%
%:%414=122%:%
%:%415=122%:%
%:%416=123%:%
%:%417=123%:%
%:%418=124%:%
%:%419=124%:%
%:%420=124%:%
%:%421=124%:%
%:%422=125%:%
%:%423=125%:%
%:%424=126%:%
%:%425=126%:%
%:%426=127%:%
%:%427=127%:%
%:%428=127%:%
%:%429=128%:%
%:%430=128%:%
%:%431=129%:%
%:%432=129%:%
%:%433=130%:%
%:%434=130%:%
%:%435=130%:%
%:%436=131%:%
%:%442=131%:%
%:%445=132%:%
%:%446=133%:%
%:%447=133%:%
%:%450=134%:%
%:%454=134%:%
%:%455=134%:%
%:%460=134%:%
%:%463=135%:%
%:%464=136%:%
%:%465=136%:%
%:%468=137%:%
%:%472=137%:%
%:%473=137%:%
%:%478=137%:%
%:%481=138%:%
%:%482=139%:%
%:%483=139%:%
%:%486=140%:%
%:%490=140%:%
%:%491=140%:%
%:%496=140%:%
%:%499=141%:%
%:%500=142%:%
%:%501=142%:%
%:%504=143%:%
%:%508=143%:%
%:%509=143%:%
%:%514=143%:%
%:%517=144%:%
%:%518=145%:%
%:%519=145%:%
%:%522=146%:%
%:%526=146%:%
%:%527=146%:%
%:%532=146%:%
%:%535=147%:%
%:%536=148%:%
%:%537=148%:%
%:%540=149%:%
%:%544=149%:%
%:%545=149%:%
%:%550=149%:%
%:%553=150%:%
%:%554=151%:%
%:%555=151%:%
%:%558=152%:%
%:%562=152%:%
%:%563=152%:%
%:%568=152%:%
%:%571=153%:%
%:%572=154%:%
%:%573=154%:%
%:%580=155%:%
%:%581=155%:%
%:%582=156%:%
%:%583=156%:%
%:%584=157%:%
%:%585=157%:%
%:%586=158%:%
%:%587=159%:%
%:%588=159%:%
%:%589=159%:%
%:%590=160%:%
%:%591=161%:%
%:%592=161%:%
%:%593=162%:%
%:%594=162%:%
%:%595=163%:%
%:%596=163%:%
%:%597=163%:%
%:%598=164%:%
%:%599=165%:%
%:%600=165%:%
%:%601=165%:%
%:%602=166%:%
%:%603=166%:%
%:%604=166%:%
%:%605=167%:%
%:%606=168%:%
%:%607=168%:%
%:%608=169%:%
%:%609=169%:%
%:%610=170%:%
%:%611=171%:%
%:%612=171%:%
%:%613=171%:%
%:%614=172%:%
%:%615=173%:%
%:%616=173%:%
%:%617=174%:%
%:%618=174%:%
%:%619=174%:%
%:%620=175%:%
%:%621=176%:%
%:%622=176%:%
%:%623=177%:%
%:%624=177%:%
%:%625=177%:%
%:%626=177%:%
%:%627=178%:%
%:%633=178%:%
%:%636=179%:%
%:%637=180%:%
%:%638=180%:%
%:%641=181%:%
%:%645=181%:%
%:%646=181%:%
%:%647=182%:%
%:%648=183%:%
%:%649=183%:%
%:%654=183%:%
%:%657=184%:%
%:%658=185%:%
%:%659=185%:%
%:%660=186%:%
%:%661=187%:%
%:%668=188%:%
%:%669=188%:%
%:%670=189%:%
%:%671=189%:%
%:%672=190%:%
%:%673=190%:%
%:%674=190%:%
%:%675=190%:%
%:%676=190%:%
%:%677=191%:%
%:%678=191%:%
%:%679=192%:%
%:%680=192%:%
%:%681=193%:%
%:%682=193%:%
%:%683=193%:%
%:%684=193%:%
%:%685=193%:%
%:%686=194%:%
%:%687=194%:%
%:%688=195%:%
%:%689=195%:%
%:%690=196%:%
%:%691=196%:%
%:%692=197%:%
%:%693=197%:%
%:%694=198%:%
%:%695=198%:%
%:%696=198%:%
%:%697=199%:%
%:%703=199%:%
%:%706=200%:%
%:%707=201%:%
%:%708=201%:%
%:%715=202%:%
%:%716=202%:%
%:%717=203%:%
%:%718=203%:%
%:%719=204%:%
%:%720=204%:%
%:%721=205%:%
%:%722=205%:%
%:%723=206%:%
%:%724=206%:%
%:%725=207%:%
%:%726=207%:%
%:%727=208%:%
%:%728=208%:%
%:%729=209%:%
%:%730=209%:%
%:%731=210%:%
%:%732=210%:%
%:%733=211%:%
%:%739=211%:%
%:%742=212%:%
%:%743=213%:%
%:%744=213%:%
%:%747=214%:%
%:%751=214%:%
%:%752=214%:%
%:%757=214%:%
%:%760=215%:%
%:%761=216%:%
%:%762=216%:%
%:%765=217%:%
%:%769=217%:%
%:%770=217%:%
%:%775=217%:%
%:%778=218%:%
%:%779=219%:%
%:%780=220%:%
%:%781=220%:%
%:%782=221%:%
%:%783=222%:%
%:%790=223%:%
%:%791=223%:%
%:%792=224%:%
%:%793=224%:%
%:%794=225%:%
%:%795=225%:%
%:%796=226%:%
%:%797=226%:%
%:%798=227%:%
%:%799=227%:%
%:%800=228%:%
%:%801=228%:%
%:%802=229%:%
%:%803=229%:%
%:%804=229%:%
%:%805=230%:%
%:%806=230%:%
%:%807=231%:%
%:%808=231%:%
%:%809=232%:%
%:%810=232%:%
%:%811=233%:%
%:%812=233%:%
%:%813=233%:%
%:%814=233%:%
%:%815=234%:%
%:%816=234%:%
%:%817=235%:%
%:%818=235%:%
%:%819=235%:%
%:%820=236%:%
%:%821=236%:%
%:%822=237%:%
%:%823=237%:%
%:%824=238%:%
%:%825=238%:%
%:%826=239%:%
%:%827=239%:%
%:%828=240%:%
%:%829=240%:%
%:%830=241%:%
%:%831=241%:%
%:%832=242%:%
%:%833=242%:%
%:%834=242%:%
%:%835=243%:%
%:%836=243%:%
%:%837=243%:%
%:%838=244%:%
%:%839=244%:%
%:%840=245%:%
%:%841=245%:%
%:%842=246%:%
%:%843=246%:%
%:%844=246%:%
%:%845=246%:%
%:%846=247%:%
%:%847=247%:%
%:%848=248%:%
%:%849=248%:%
%:%850=249%:%
%:%851=249%:%
%:%852=249%:%
%:%853=249%:%
%:%854=250%:%
%:%855=250%:%
%:%856=251%:%
%:%857=251%:%
%:%858=252%:%
%:%859=252%:%
%:%860=252%:%
%:%861=253%:%
%:%862=253%:%
%:%863=254%:%
%:%864=254%:%
%:%865=254%:%
%:%866=255%:%
%:%872=255%:%
%:%875=256%:%
%:%876=257%:%
%:%877=257%:%
%:%878=258%:%
%:%879=259%:%
%:%886=260%:%
%:%887=260%:%
%:%888=261%:%
%:%889=261%:%
%:%890=262%:%
%:%891=262%:%
%:%892=263%:%
%:%893=263%:%
%:%894=264%:%
%:%895=264%:%
%:%896=265%:%
%:%897=265%:%
%:%898=265%:%
%:%899=265%:%
%:%900=266%:%
%:%901=266%:%
%:%902=267%:%
%:%903=267%:%
%:%904=267%:%
%:%905=267%:%
%:%906=268%:%
%:%912=268%:%
%:%915=269%:%
%:%916=270%:%
%:%917=271%:%
%:%918=271%:%
%:%919=272%:%
%:%920=273%:%
%:%921=274%:%
%:%922=274%:%
%:%925=275%:%
%:%929=275%:%
%:%930=275%:%
%:%935=275%:%
%:%938=276%:%
%:%939=277%:%
%:%940=277%:%
%:%943=278%:%
%:%947=278%:%
%:%948=278%:%
%:%953=278%:%
%:%956=279%:%
%:%957=280%:%
%:%958=280%:%
%:%961=281%:%
%:%965=281%:%
%:%966=281%:%
%:%971=281%:%
%:%974=282%:%
%:%975=283%:%
%:%983=285%:%
%:%993=286%:%
%:%994=286%:%
%:%995=287%:%
%:%996=288%:%
%:%997=289%:%
%:%998=289%:%
%:%999=290%:%
%:%1000=291%:%
%:%1001=292%:%
%:%1002=293%:%
%:%1003=293%:%
%:%1004=294%:%
%:%1005=295%:%
%:%1006=296%:%
%:%1007=297%:%
%:%1008=297%:%
%:%1009=298%:%
%:%1012=299%:%
%:%1016=299%:%
%:%1017=299%:%
%:%1023=299%:%
%:%1026=300%:%
%:%1027=301%:%
%:%1028=301%:%
%:%1029=302%:%
%:%1030=303%:%
%:%1031=304%:%
%:%1032=305%:%
%:%1033=306%:%
%:%1034=306%:%
%:%1035=307%:%
%:%1038=308%:%
%:%1042=308%:%
%:%1043=308%:%
%:%1044=309%:%
%:%1045=309%:%
%:%1046=309%:%
%:%1051=309%:%
%:%1054=310%:%
%:%1055=311%:%
%:%1056=311%:%
%:%1059=312%:%
%:%1063=312%:%
%:%1064=312%:%
%:%1065=312%:%
%:%1066=312%:%
%:%1071=312%:%
%:%1074=313%:%
%:%1075=314%:%
%:%1076=315%:%
%:%1077=315%:%
%:%1078=316%:%
%:%1081=317%:%
%:%1085=317%:%
%:%1086=317%:%
%:%1087=317%:%
%:%1092=317%:%
%:%1095=318%:%
%:%1096=319%:%
%:%1097=320%:%
%:%1098=320%:%
%:%1100=322%:%
%:%1103=323%:%
%:%1107=323%:%
%:%1108=323%:%
%:%1113=323%:%
%:%1116=324%:%
%:%1117=325%:%
%:%1118=325%:%
%:%1125=326%:%
%:%1126=326%:%
%:%1127=327%:%
%:%1128=327%:%
%:%1129=328%:%
%:%1130=329%:%
%:%1131=329%:%
%:%1132=330%:%
%:%1133=330%:%
%:%1134=330%:%
%:%1135=331%:%
%:%1136=331%:%
%:%1137=332%:%
%:%1138=332%:%
%:%1139=333%:%
%:%1140=333%:%
%:%1141=333%:%
%:%1142=334%:%
%:%1143=334%:%
%:%1144=335%:%
%:%1145=335%:%
%:%1146=336%:%
%:%1147=336%:%
%:%1148=336%:%
%:%1149=337%:%
%:%1150=337%:%
%:%1151=338%:%
%:%1152=338%:%
%:%1153=339%:%
%:%1154=339%:%
%:%1155=339%:%
%:%1156=340%:%
%:%1157=340%:%
%:%1158=340%:%
%:%1159=340%:%
%:%1160=340%:%
%:%1161=341%:%
%:%1167=341%:%
%:%1170=342%:%
%:%1171=343%:%
%:%1172=344%:%
%:%1173=344%:%
%:%1174=345%:%
%:%1175=346%:%
%:%1176=347%:%
%:%1183=348%:%
%:%1184=348%:%
%:%1185=349%:%
%:%1186=349%:%
%:%1187=350%:%
%:%1188=350%:%
%:%1189=350%:%
%:%1190=351%:%
%:%1191=351%:%
%:%1192=351%:%
%:%1193=352%:%
%:%1194=352%:%
%:%1195=353%:%
%:%1196=353%:%
%:%1197=353%:%
%:%1198=354%:%
%:%1199=354%:%
%:%1200=354%:%
%:%1201=355%:%
%:%1202=355%:%
%:%1203=355%:%
%:%1204=355%:%
%:%1205=356%:%
%:%1211=356%:%
%:%1214=357%:%
%:%1215=358%:%
%:%1216=358%:%
%:%1217=359%:%
%:%1218=360%:%
%:%1225=361%:%
%:%1226=361%:%
%:%1227=362%:%
%:%1228=362%:%
%:%1229=363%:%
%:%1230=363%:%
%:%1231=363%:%
%:%1232=364%:%
%:%1233=364%:%
%:%1234=365%:%
%:%1235=365%:%
%:%1236=365%:%
%:%1237=365%:%
%:%1238=366%:%
%:%1244=366%:%
%:%1247=367%:%
%:%1248=368%:%
%:%1249=368%:%
%:%1250=369%:%
%:%1251=370%:%
%:%1253=372%:%
%:%1254=373%:%
%:%1255=374%:%
%:%1257=376%:%
%:%1258=376%:%
%:%1259=377%:%
%:%1260=378%:%
%:%1261=379%:%
%:%1262=380%:%
%:%1263=380%:%
%:%1264=381%:%
%:%1265=382%:%
%:%1272=383%:%
%:%1273=383%:%
%:%1274=384%:%
%:%1275=384%:%
%:%1276=385%:%
%:%1277=385%:%
%:%1278=385%:%
%:%1279=385%:%
%:%1280=386%:%
%:%1281=386%:%
%:%1282=387%:%
%:%1283=387%:%
%:%1284=387%:%
%:%1285=387%:%
%:%1286=388%:%
%:%1287=388%:%
%:%1288=389%:%
%:%1289=389%:%
%:%1290=389%:%
%:%1291=389%:%
%:%1292=390%:%
%:%1293=390%:%
%:%1294=390%:%
%:%1295=391%:%
%:%1296=391%:%
%:%1297=391%:%
%:%1298=391%:%
%:%1299=392%:%
%:%1300=392%:%
%:%1301=393%:%
%:%1302=393%:%
%:%1303=394%:%
%:%1304=394%:%
%:%1305=394%:%
%:%1306=395%:%
%:%1307=395%:%
%:%1308=395%:%
%:%1309=396%:%
%:%1310=396%:%
%:%1311=397%:%
%:%1312=397%:%
%:%1313=397%:%
%:%1314=398%:%
%:%1315=398%:%
%:%1316=398%:%
%:%1317=398%:%
%:%1318=399%:%
%:%1328=401%:%
%:%1330=403%:%
%:%1331=403%:%
%:%1338=404%:%
%:%1339=404%:%
%:%1340=405%:%
%:%1341=405%:%
%:%1342=406%:%
%:%1343=407%:%
%:%1344=407%:%
%:%1345=408%:%
%:%1346=409%:%
%:%1347=409%:%
%:%1348=409%:%
%:%1349=409%:%
%:%1350=410%:%
%:%1351=410%:%
%:%1352=411%:%
%:%1353=411%:%
%:%1354=411%:%
%:%1355=411%:%
%:%1356=412%:%
%:%1357=412%:%
%:%1358=413%:%
%:%1359=413%:%
%:%1360=414%:%
%:%1361=414%:%
%:%1362=414%:%
%:%1363=415%:%
%:%1364=415%:%
%:%1365=416%:%
%:%1366=416%:%
%:%1367=417%:%
%:%1368=417%:%
%:%1369=417%:%
%:%1370=418%:%
%:%1371=418%:%
%:%1372=419%:%
%:%1373=419%:%
%:%1374=419%:%
%:%1375=419%:%
%:%1376=420%:%
%:%1382=420%:%
%:%1385=421%:%
%:%1386=422%:%
%:%1387=422%:%
%:%1390=423%:%
%:%1394=423%:%
%:%1395=423%:%
%:%1404=425%:%
%:%1405=426%:%
%:%1407=428%:%
%:%1408=428%:%
%:%1415=429%:%
%:%1416=429%:%
%:%1417=430%:%
%:%1418=430%:%
%:%1419=431%:%
%:%1420=431%:%
%:%1421=431%:%
%:%1422=432%:%
%:%1423=432%:%
%:%1424=433%:%
%:%1425=433%:%
%:%1426=434%:%
%:%1427=434%:%
%:%1428=434%:%
%:%1429=435%:%
%:%1430=435%:%
%:%1431=436%:%
%:%1432=436%:%
%:%1433=437%:%
%:%1434=437%:%
%:%1435=438%:%
%:%1436=438%:%
%:%1437=439%:%
%:%1438=439%:%
%:%1439=440%:%
%:%1440=440%:%
%:%1441=440%:%
%:%1442=441%:%
%:%1443=441%:%
%:%1444=442%:%
%:%1445=442%:%
%:%1446=443%:%
%:%1447=443%:%
%:%1448=443%:%
%:%1449=444%:%
%:%1450=444%:%
%:%1451=445%:%
%:%1452=445%:%
%:%1453=446%:%
%:%1454=446%:%
%:%1455=447%:%
%:%1456=447%:%
%:%1457=448%:%
%:%1458=448%:%
%:%1459=449%:%
%:%1460=449%:%
%:%1461=449%:%
%:%1462=450%:%
%:%1463=450%:%
%:%1464=451%:%
%:%1470=451%:%
%:%1473=452%:%
%:%1474=453%:%
%:%1475=453%:%
%:%1476=454%:%
%:%1483=455%:%
%:%1484=455%:%
%:%1485=456%:%
%:%1486=456%:%
%:%1487=457%:%
%:%1488=458%:%
%:%1489=459%:%
%:%1490=459%:%
%:%1491=460%:%
%:%1492=461%:%
%:%1493=461%:%
%:%1494=462%:%
%:%1495=463%:%
%:%1496=463%:%
%:%1497=464%:%
%:%1498=464%:%
%:%1499=464%:%
%:%1500=465%:%
%:%1501=465%:%
%:%1502=465%:%
%:%1503=466%:%
%:%1504=466%:%
%:%1505=467%:%
%:%1506=467%:%
%:%1507=467%:%
%:%1508=468%:%
%:%1509=468%:%
%:%1510=469%:%
%:%1511=469%:%
%:%1512=469%:%
%:%1513=470%:%
%:%1514=470%:%
%:%1515=470%:%
%:%1516=471%:%
%:%1517=471%:%
%:%1518=472%:%
%:%1519=472%:%
%:%1520=473%:%
%:%1521=473%:%
%:%1522=474%:%
%:%1523=474%:%
%:%1524=475%:%
%:%1525=475%:%
%:%1526=476%:%
%:%1527=476%:%
%:%1528=477%:%
%:%1529=477%:%
%:%1530=478%:%
%:%1531=478%:%
%:%1532=479%:%
%:%1533=479%:%
%:%1534=480%:%
%:%1535=480%:%
%:%1536=481%:%
%:%1537=481%:%
%:%1538=482%:%
%:%1539=482%:%
%:%1540=483%:%
%:%1541=483%:%
%:%1542=484%:%
%:%1543=484%:%
%:%1544=485%:%
%:%1545=485%:%
%:%1546=486%:%
%:%1547=486%:%
%:%1548=487%:%
%:%1549=487%:%
%:%1550=488%:%
%:%1551=488%:%
%:%1552=488%:%
%:%1553=489%:%
%:%1554=489%:%
%:%1555=490%:%
%:%1556=490%:%
%:%1557=491%:%
%:%1558=491%:%
%:%1559=492%:%
%:%1560=492%:%
%:%1561=493%:%
%:%1562=493%:%
%:%1563=494%:%
%:%1564=495%:%
%:%1565=495%:%
%:%1566=496%:%
%:%1567=496%:%
%:%1568=497%:%
%:%1569=497%:%
%:%1570=498%:%
%:%1571=498%:%
%:%1572=499%:%
%:%1573=499%:%
%:%1574=500%:%
%:%1575=500%:%
%:%1576=501%:%
%:%1577=501%:%
%:%1578=502%:%
%:%1579=502%:%
%:%1580=503%:%
%:%1586=503%:%
%:%1589=504%:%
%:%1590=505%:%
%:%1591=505%:%
%:%1592=506%:%
%:%1595=507%:%
%:%1599=507%:%
%:%1600=507%:%
%:%1601=507%:%
%:%1606=507%:%
%:%1609=508%:%
%:%1610=509%:%
%:%1611=509%:%
%:%1612=510%:%
%:%1619=511%:%
%:%1620=511%:%
%:%1621=512%:%
%:%1622=512%:%
%:%1623=513%:%
%:%1624=513%:%
%:%1625=513%:%
%:%1626=514%:%
%:%1627=514%:%
%:%1628=515%:%
%:%1629=515%:%
%:%1630=516%:%
%:%1631=516%:%
%:%1632=516%:%
%:%1633=517%:%
%:%1634=517%:%
%:%1635=518%:%
%:%1636=518%:%
%:%1637=519%:%
%:%1638=519%:%
%:%1639=520%:%
%:%1640=520%:%
%:%1641=521%:%
%:%1642=521%:%
%:%1643=522%:%
%:%1644=522%:%
%:%1645=522%:%
%:%1646=523%:%
%:%1647=523%:%
%:%1648=524%:%
%:%1649=524%:%
%:%1650=525%:%
%:%1651=525%:%
%:%1652=525%:%
%:%1653=526%:%
%:%1654=526%:%
%:%1655=527%:%
%:%1656=527%:%
%:%1657=527%:%
%:%1658=528%:%
%:%1664=528%:%
%:%1667=529%:%
%:%1668=530%:%
%:%1669=530%:%
%:%1676=531%:%
%:%1677=531%:%
%:%1678=532%:%
%:%1679=532%:%
%:%1680=533%:%
%:%1681=534%:%
%:%1682=534%:%
%:%1683=534%:%
%:%1684=535%:%
%:%1685=535%:%
%:%1686=536%:%
%:%1687=536%:%
%:%1688=537%:%
%:%1689=537%:%
%:%1690=537%:%
%:%1691=538%:%
%:%1692=538%:%
%:%1693=539%:%
%:%1694=539%:%
%:%1695=539%:%
%:%1696=540%:%
%:%1697=540%:%
%:%1698=541%:%
%:%1699=541%:%
%:%1700=542%:%
%:%1701=542%:%
%:%1702=542%:%
%:%1703=543%:%
%:%1709=543%:%
%:%1712=544%:%
%:%1713=545%:%
%:%1714=545%:%
%:%1717=546%:%
%:%1721=546%:%
%:%1722=546%:%
%:%1723=547%:%
%:%1724=547%:%
%:%1729=547%:%
%:%1732=548%:%
%:%1733=549%:%
%:%1734=549%:%
%:%1737=550%:%
%:%1741=550%:%
%:%1742=550%:%
%:%1743=551%:%
%:%1744=551%:%
%:%1749=551%:%
%:%1752=552%:%
%:%1753=553%:%
%:%1754=553%:%
%:%1755=554%:%
%:%1756=555%:%
%:%1757=556%:%
%:%1758=557%:%
%:%1759=558%:%
%:%1760=558%:%
%:%1767=559%:%
%:%1768=559%:%
%:%1769=560%:%
%:%1770=560%:%
%:%1771=561%:%
%:%1772=562%:%
%:%1773=562%:%
%:%1774=563%:%
%:%1775=563%:%
%:%1776=563%:%
%:%1777=564%:%
%:%1778=564%:%
%:%1779=565%:%
%:%1780=565%:%
%:%1781=566%:%
%:%1782=566%:%
%:%1783=567%:%
%:%1784=567%:%
%:%1785=568%:%
%:%1786=568%:%
%:%1787=569%:%
%:%1788=569%:%
%:%1789=569%:%
%:%1790=570%:%
%:%1796=570%:%
%:%1799=571%:%
%:%1800=572%:%
%:%1801=572%:%
%:%1804=573%:%
%:%1808=573%:%
%:%1809=573%:%
%:%1814=573%:%
%:%1817=574%:%
%:%1818=575%:%
%:%1819=575%:%
%:%1822=576%:%
%:%1826=576%:%
%:%1827=576%:%
%:%1832=576%:%
%:%1835=577%:%
%:%1836=578%:%
%:%1837=578%:%
%:%1838=579%:%
%:%1839=580%:%
%:%1846=581%:%
%:%1847=581%:%
%:%1848=582%:%
%:%1849=582%:%
%:%1850=583%:%
%:%1851=583%:%
%:%1852=584%:%
%:%1853=584%:%
%:%1854=584%:%
%:%1855=585%:%
%:%1856=585%:%
%:%1857=585%:%
%:%1858=586%:%
%:%1859=586%:%
%:%1860=587%:%
%:%1861=587%:%
%:%1862=588%:%
%:%1863=588%:%
%:%1864=589%:%
%:%1865=589%:%
%:%1866=590%:%
%:%1867=590%:%
%:%1868=590%:%
%:%1869=591%:%
%:%1875=591%:%
%:%1878=592%:%
%:%1879=593%:%
%:%1880=594%:%
%:%1881=594%:%
%:%1882=595%:%
%:%1885=596%:%
%:%1889=596%:%
%:%1890=596%:%
%:%1895=596%:%
%:%1898=597%:%
%:%1899=598%:%
%:%1900=598%:%
%:%1901=599%:%
%:%1904=600%:%
%:%1908=600%:%
%:%1909=600%:%
%:%1914=600%:%
%:%1917=601%:%
%:%1918=602%:%
%:%1919=602%:%
%:%1926=603%:%
%:%1927=603%:%
%:%1928=604%:%
%:%1929=604%:%
%:%1930=604%:%
%:%1931=605%:%
%:%1932=605%:%
%:%1933=606%:%
%:%1934=606%:%
%:%1935=606%:%
%:%1936=607%:%
%:%1937=607%:%
%:%1938=607%:%
%:%1939=608%:%
%:%1940=608%:%
%:%1941=609%:%
%:%1942=609%:%
%:%1943=610%:%
%:%1944=610%:%
%:%1945=610%:%
%:%1946=611%:%
%:%1947=611%:%
%:%1948=612%:%
%:%1949=612%:%
%:%1950=613%:%
%:%1951=613%:%
%:%1952=613%:%
%:%1953=614%:%
%:%1954=614%:%
%:%1955=615%:%
%:%1956=615%:%
%:%1957=616%:%
%:%1958=616%:%
%:%1959=616%:%
%:%1960=616%:%
%:%1961=617%:%
%:%1967=617%:%
%:%1970=618%:%
%:%1971=619%:%
%:%1972=619%:%
%:%1973=620%:%
%:%1974=621%:%
%:%1975=621%:%
%:%1976=622%:%
%:%1977=623%:%
%:%1978=624%:%
%:%1981=625%:%
%:%1985=625%:%
%:%1986=625%:%
%:%1987=626%:%
%:%1988=626%:%
%:%1989=627%:%
%:%1990=627%:%
%:%1991=628%:%
%:%1992=628%:%
%:%1993=628%:%
%:%1994=628%:%
%:%1995=628%:%
%:%1996=629%:%
%:%1997=629%:%
%:%1998=630%:%
%:%1999=630%:%
%:%2000=631%:%
%:%2001=631%:%
%:%2002=631%:%
%:%2003=631%:%
%:%2004=632%:%
%:%2005=632%:%
%:%2006=633%:%
%:%2007=633%:%
%:%2008=634%:%
%:%2009=634%:%
%:%2010=634%:%
%:%2011=635%:%
%:%2012=635%:%
%:%2013=636%:%
%:%2014=636%:%
%:%2015=637%:%
%:%2016=637%:%
%:%2017=637%:%
%:%2018=638%:%
%:%2019=638%:%
%:%2020=639%:%
%:%2021=639%:%
%:%2022=640%:%
%:%2023=640%:%
%:%2024=640%:%
%:%2025=641%:%
%:%2026=641%:%
%:%2027=641%:%
%:%2028=641%:%
%:%2029=641%:%
%:%2030=642%:%
%:%2036=642%:%
%:%2039=643%:%
%:%2040=644%:%
%:%2041=645%:%
%:%2042=645%:%
%:%2043=646%:%
%:%2044=647%:%
%:%2045=648%:%
%:%2046=649%:%
%:%2047=649%:%
%:%2048=650%:%
%:%2049=651%:%
%:%2050=652%:%
%:%2051=653%:%
%:%2052=653%:%
%:%2055=654%:%
%:%2059=654%:%
%:%2060=654%:%
%:%2065=654%:%
%:%2068=655%:%
%:%2069=656%:%
%:%2070=656%:%
%:%2071=657%:%
%:%2072=658%:%
%:%2073=659%:%
%:%2074=660%:%
%:%2081=661%:%
%:%2082=661%:%
%:%2083=662%:%
%:%2084=662%:%
%:%2085=663%:%
%:%2086=663%:%
%:%2087=664%:%
%:%2088=664%:%
%:%2089=664%:%
%:%2090=665%:%
%:%2091=665%:%
%:%2092=666%:%
%:%2093=666%:%
%:%2094=667%:%
%:%2095=667%:%
%:%2096=668%:%
%:%2097=668%:%
%:%2098=669%:%
%:%2099=669%:%
%:%2100=670%:%
%:%2101=670%:%
%:%2102=671%:%
%:%2103=671%:%
%:%2104=672%:%
%:%2105=672%:%
%:%2106=673%:%
%:%2112=673%:%
%:%2115=674%:%
%:%2116=677%:%
%:%2117=678%:%
%:%2118=678%:%
%:%2119=679%:%
%:%2120=680%:%
%:%2121=681%:%
%:%2122=682%:%
%:%2123=683%:%
%:%2124=683%:%
%:%2125=684%:%
%:%2128=685%:%
%:%2132=685%:%
%:%2133=685%:%
%:%2138=685%:%
%:%2141=686%:%
%:%2142=687%:%
%:%2143=687%:%
%:%2144=688%:%
%:%2146=690%:%
%:%2149=691%:%
%:%2153=691%:%
%:%2154=691%:%
%:%2155=691%:%
%:%2160=691%:%
%:%2163=692%:%
%:%2164=695%:%
%:%2165=696%:%
%:%2166=696%:%
%:%2167=697%:%
%:%2168=698%:%
%:%2169=699%:%
%:%2170=700%:%
%:%2171=700%:%
%:%2172=701%:%
%:%2175=702%:%
%:%2179=702%:%
%:%2180=702%:%
%:%2185=702%:%
%:%2188=703%:%
%:%2189=704%:%
%:%2190=704%:%
%:%2191=705%:%
%:%2193=707%:%
%:%2196=708%:%
%:%2200=708%:%
%:%2201=708%:%
%:%2202=708%:%
%:%2203=709%:%
%:%2204=709%:%
%:%2209=709%:%
%:%2212=710%:%
%:%2213=711%:%
%:%2214=712%:%
%:%2215=713%:%
%:%2216=713%:%
%:%2217=714%:%
%:%2218=715%:%
%:%2219=716%:%
%:%2220=717%:%
%:%2227=718%:%
%:%2228=718%:%
%:%2229=719%:%
%:%2230=719%:%
%:%2231=720%:%
%:%2232=721%:%
%:%2233=722%:%
%:%2234=723%:%
%:%2235=723%:%
%:%2236=723%:%
%:%2237=724%:%
%:%2238=724%:%
%:%2239=724%:%
%:%2241=726%:%
%:%2242=727%:%
%:%2243=728%:%
%:%2244=728%:%
%:%2245=728%:%
%:%2246=729%:%
%:%2247=729%:%
%:%2248=730%:%
%:%2249=731%:%
%:%2250=732%:%
%:%2251=732%:%
%:%2252=733%:%
%:%2253=734%:%
%:%2254=735%:%
%:%2255=735%:%
%:%2256=735%:%
%:%2257=736%:%
%:%2258=736%:%
%:%2259=737%:%
%:%2260=737%:%
%:%2261=738%:%
%:%2262=739%:%
%:%2263=740%:%
%:%2264=740%:%
%:%2265=741%:%
%:%2266=741%:%
%:%2267=742%:%
%:%2268=742%:%
%:%2269=743%:%
%:%2270=743%:%
%:%2271=743%:%
%:%2272=744%:%
%:%2273=744%:%
%:%2274=745%:%
%:%2275=745%:%
%:%2276=746%:%
%:%2277=747%:%
%:%2278=747%:%
%:%2279=747%:%
%:%2280=748%:%
%:%2281=748%:%
%:%2282=749%:%
%:%2283=749%:%
%:%2284=749%:%
%:%2285=749%:%
%:%2286=750%:%
%:%2292=750%:%
%:%2295=751%:%
%:%2296=752%:%
%:%2297=752%:%
%:%2298=753%:%
%:%2299=754%:%
%:%2300=755%:%
%:%2301=756%:%
%:%2308=757%:%
%:%2309=757%:%
%:%2310=758%:%
%:%2311=758%:%
%:%2312=759%:%
%:%2313=759%:%
%:%2314=760%:%
%:%2315=761%:%
%:%2316=761%:%
%:%2317=762%:%
%:%2318=762%:%
%:%2319=763%:%
%:%2320=763%:%
%:%2321=764%:%
%:%2322=764%:%
%:%2323=764%:%
%:%2324=765%:%
%:%2325=765%:%
%:%2326=766%:%
%:%2327=766%:%
%:%2328=766%:%
%:%2329=766%:%
%:%2330=766%:%
%:%2331=767%:%
%:%2337=767%:%
%:%2340=768%:%
%:%2341=769%:%
%:%2342=769%:%
%:%2345=770%:%
%:%2349=770%:%
%:%2350=770%:%
%:%2351=770%:%
%:%2356=770%:%
%:%2359=771%:%
%:%2360=772%:%
%:%2361=772%:%
%:%2362=773%:%
%:%2369=774%:%
%:%2370=774%:%
%:%2371=775%:%
%:%2372=775%:%
%:%2373=776%:%
%:%2374=777%:%
%:%2375=777%:%
%:%2376=777%:%
%:%2377=777%:%
%:%2378=778%:%
%:%2379=778%:%
%:%2380=779%:%
%:%2381=780%:%
%:%2382=780%:%
%:%2383=780%:%
%:%2384=781%:%
%:%2385=781%:%
%:%2386=782%:%
%:%2387=782%:%
%:%2388=783%:%
%:%2389=783%:%
%:%2390=783%:%
%:%2391=784%:%
%:%2397=784%:%
%:%2400=785%:%
%:%2401=786%:%
%:%2402=786%:%
%:%2403=787%:%
%:%2406=788%:%
%:%2410=788%:%
%:%2411=788%:%
%:%2412=788%:%
%:%2417=788%:%
%:%2420=789%:%
%:%2421=790%:%
%:%2422=790%:%
%:%2423=791%:%
%:%2430=792%:%
%:%2431=792%:%
%:%2432=793%:%
%:%2433=793%:%
%:%2434=794%:%
%:%2435=794%:%
%:%2436=795%:%
%:%2437=796%:%
%:%2438=796%:%
%:%2439=797%:%
%:%2440=797%:%
%:%2441=797%:%
%:%2442=797%:%
%:%2443=797%:%
%:%2444=798%:%
%:%2450=798%:%
%:%2453=799%:%
%:%2454=800%:%
%:%2455=800%:%
%:%2458=801%:%
%:%2462=801%:%
%:%2463=801%:%
%:%2464=801%:%
%:%2470=801%:%
%:%2473=802%:%
%:%2474=803%:%
%:%2475=803%:%
%:%2478=804%:%
%:%2482=804%:%
%:%2483=804%:%
%:%2484=804%:%
%:%2490=804%:%
%:%2493=805%:%
%:%2494=806%:%
%:%2495=806%:%
%:%2498=807%:%
%:%2502=807%:%
%:%2503=807%:%
%:%2504=807%:%
%:%2510=807%:%
%:%2513=808%:%
%:%2514=809%:%
%:%2515=809%:%
%:%2518=810%:%
%:%2522=810%:%
%:%2523=810%:%
%:%2524=811%:%
%:%2525=811%:%
%:%2526=812%:%
%:%2527=813%:%
%:%2528=813%:%
%:%2533=813%:%
%:%2536=814%:%
%:%2537=815%:%
%:%2538=815%:%
%:%2539=816%:%
%:%2540=817%:%
%:%2541=818%:%
%:%2542=818%:%
%:%2543=819%:%
%:%2544=820%:%
%:%2545=821%:%
%:%2546=822%:%
%:%2547=822%:%
%:%2550=823%:%
%:%2554=823%:%
%:%2555=823%:%
%:%2556=823%:%
%:%2557=823%:%
%:%2562=823%:%
%:%2565=824%:%
%:%2566=825%:%
%:%2567=825%:%
%:%2568=826%:%
%:%2569=827%:%
%:%2570=828%:%
%:%2571=829%:%
%:%2572=829%:%
%:%2575=830%:%
%:%2579=830%:%
%:%2580=830%:%
%:%2581=830%:%
%:%2586=830%:%
%:%2589=831%:%
%:%2590=832%:%
%:%2591=832%:%
%:%2592=833%:%
%:%2594=835%:%
%:%2597=836%:%
%:%2601=836%:%
%:%2602=836%:%
%:%2603=836%:%
%:%2608=836%:%
%:%2611=837%:%
%:%2612=838%:%
%:%2613=838%:%
%:%2614=839%:%
%:%2615=840%:%
%:%2616=841%:%
%:%2617=841%:%
%:%2618=842%:%
%:%2619=843%:%
%:%2620=844%:%
%:%2621=845%:%
%:%2622=846%:%
%:%2623=846%:%
%:%2624=847%:%
%:%2627=848%:%
%:%2631=848%:%
%:%2632=848%:%
%:%2633=849%:%
%:%2634=849%:%
%:%2635=849%:%
%:%2640=849%:%
%:%2643=850%:%
%:%2644=851%:%
%:%2645=851%:%
%:%2646=852%:%
%:%2647=853%:%
%:%2648=854%:%
%:%2649=855%:%
%:%2652=856%:%
%:%2656=856%:%
%:%2657=856%:%
%:%2658=857%:%
%:%2659=857%:%
%:%2660=858%:%
%:%2665=858%:%
%:%2670=859%:%
%:%2675=860%:%
%:%2676=860%:%
%:%2681=860%:%
%:%2684=861%:%
%:%2685=862%:%
%:%2686=862%:%
%:%2687=863%:%
%:%2688=864%:%
%:%2689=865%:%
%:%2690=866%:%
%:%2691=866%:%
%:%2692=867%:%
%:%2693=868%:%
%:%2694=869%:%
%:%2695=870%:%
%:%2702=871%:%
%:%2703=871%:%
%:%2704=872%:%
%:%2705=872%:%
%:%2706=873%:%
%:%2707=874%:%
%:%2708=874%:%
%:%2709=874%:%
%:%2710=875%:%
%:%2711=875%:%
%:%2712=875%:%
%:%2713=876%:%
%:%2714=876%:%
%:%2715=876%:%
%:%2716=877%:%
%:%2717=877%:%
%:%2718=878%:%
%:%2719=878%:%
%:%2720=879%:%
%:%2721=880%:%
%:%2722=880%:%
%:%2723=881%:%
%:%2729=881%:%
%:%2732=882%:%
%:%2733=883%:%
%:%2734=884%:%
%:%2735=884%:%
%:%2736=885%:%
%:%2737=886%:%
%:%2738=887%:%
%:%2739=888%:%
%:%2740=889%:%
%:%2741=889%:%
%:%2742=890%:%
%:%2745=891%:%
%:%2749=891%:%
%:%2750=891%:%
%:%2751=891%:%
%:%2756=891%:%
%:%2759=892%:%
%:%2760=893%:%
%:%2761=893%:%
%:%2762=894%:%
%:%2763=895%:%
%:%2764=896%:%
%:%2765=897%:%
%:%2772=898%:%
%:%2773=898%:%
%:%2774=899%:%
%:%2775=899%:%
%:%2776=900%:%
%:%2778=902%:%
%:%2779=903%:%
%:%2780=903%:%
%:%2781=903%:%
%:%2782=904%:%
%:%2783=904%:%
%:%2784=905%:%
%:%2785=905%:%
%:%2786=906%:%
%:%2787=907%:%
%:%2788=907%:%
%:%2789=907%:%
%:%2790=908%:%
%:%2791=908%:%
%:%2792=909%:%
%:%2793=909%:%
%:%2794=909%:%
%:%2795=910%:%
%:%2796=910%:%
%:%2797=910%:%
%:%2798=911%:%
%:%2804=911%:%
%:%2807=912%:%
%:%2808=913%:%
%:%2809=914%:%
%:%2810=914%:%
%:%2811=915%:%
%:%2818=916%:%
%:%2819=916%:%
%:%2820=917%:%
%:%2821=917%:%
%:%2822=918%:%
%:%2823=918%:%
%:%2824=919%:%
%:%2825=920%:%
%:%2826=921%:%
%:%2827=922%:%
%:%2828=922%:%
%:%2829=923%:%
%:%2830=923%:%
%:%2831=924%:%
%:%2832=924%:%
%:%2833=925%:%
%:%2834=925%:%
%:%2835=925%:%
%:%2836=926%:%
%:%2837=926%:%
%:%2838=927%:%
%:%2839=927%:%
%:%2840=927%:%
%:%2841=928%:%
%:%2842=928%:%
%:%2843=929%:%
%:%2849=929%:%
%:%2852=930%:%
%:%2853=931%:%
%:%2854=931%:%
%:%2857=932%:%
%:%2861=932%:%
%:%2862=932%:%
%:%2863=932%:%
%:%2868=932%:%
%:%2871=933%:%
%:%2872=934%:%
%:%2873=935%:%
%:%2874=935%:%
%:%2877=936%:%
%:%2881=936%:%
%:%2882=936%:%
%:%2883=937%:%
%:%2884=937%:%
%:%2893=939%:%
%:%2895=940%:%
%:%2896=940%:%
%:%2897=941%:%
%:%2898=942%:%
%:%2899=943%:%
%:%2900=944%:%
%:%2901=944%:%
%:%2902=945%:%
%:%2909=946%:%
%:%2910=946%:%
%:%2911=947%:%
%:%2912=947%:%
%:%2913=948%:%
%:%2914=948%:%
%:%2915=949%:%
%:%2916=949%:%
%:%2917=950%:%
%:%2918=950%:%
%:%2919=950%:%
%:%2920=951%:%
%:%2921=951%:%
%:%2922=952%:%
%:%2923=952%:%
%:%2924=953%:%
%:%2925=953%:%
%:%2926=953%:%
%:%2927=954%:%
%:%2928=954%:%
%:%2929=955%:%
%:%2930=955%:%
%:%2931=955%:%
%:%2932=956%:%
%:%2933=956%:%
%:%2934=957%:%
%:%2935=957%:%
%:%2936=958%:%
%:%2937=958%:%
%:%2938=959%:%
%:%2939=960%:%
%:%2940=961%:%
%:%2941=962%:%
%:%2942=962%:%
%:%2943=963%:%
%:%2944=963%:%
%:%2945=964%:%
%:%2946=964%:%
%:%2947=965%:%
%:%2948=965%:%
%:%2949=966%:%
%:%2950=966%:%
%:%2951=966%:%
%:%2952=967%:%
%:%2953=967%:%
%:%2954=968%:%
%:%2955=968%:%
%:%2956=969%:%
%:%2957=969%:%
%:%2958=970%:%
%:%2959=970%:%
%:%2960=971%:%
%:%2961=971%:%
%:%2962=972%:%
%:%2968=972%:%
%:%2971=973%:%
%:%2972=974%:%
%:%2973=975%:%
%:%2974=975%:%
%:%2975=976%:%
%:%2976=977%:%
%:%2977=978%:%
%:%2984=979%:%
%:%2985=979%:%
%:%2986=980%:%
%:%2987=980%:%
%:%2988=981%:%
%:%2989=981%:%
%:%2990=982%:%
%:%2991=982%:%
%:%2992=982%:%
%:%2993=983%:%
%:%2994=983%:%
%:%2995=983%:%
%:%2996=984%:%
%:%2997=984%:%
%:%2998=985%:%
%:%2999=985%:%
%:%3000=985%:%
%:%3001=986%:%
%:%3002=987%:%
%:%3003=987%:%
%:%3004=987%:%
%:%3005=988%:%
%:%3006=988%:%
%:%3007=989%:%
%:%3008=989%:%
%:%3009=990%:%
%:%3010=990%:%
%:%3011=991%:%
%:%3012=991%:%
%:%3013=992%:%
%:%3014=992%:%
%:%3015=992%:%
%:%3016=992%:%
%:%3017=993%:%
%:%3018=993%:%
%:%3019=994%:%
%:%3020=994%:%
%:%3021=995%:%
%:%3022=995%:%
%:%3023=995%:%
%:%3024=996%:%
%:%3025=996%:%
%:%3026=997%:%
%:%3027=997%:%
%:%3028=998%:%
%:%3029=998%:%
%:%3030=998%:%
%:%3031=999%:%
%:%3037=999%:%
%:%3040=1000%:%
%:%3041=1001%:%
%:%3042=1002%:%
%:%3043=1002%:%
%:%3044=1003%:%
%:%3045=1004%:%
%:%3048=1005%:%
%:%3052=1005%:%
%:%3053=1005%:%
%:%3054=1006%:%
%:%3055=1006%:%
%:%3060=1006%:%
%:%3063=1007%:%
%:%3064=1008%:%
%:%3065=1009%:%
%:%3066=1009%:%
%:%3073=1010%:%
%:%3074=1010%:%
%:%3075=1011%:%
%:%3076=1011%:%
%:%3077=1012%:%
%:%3078=1012%:%
%:%3079=1013%:%
%:%3080=1013%:%
%:%3081=1013%:%
%:%3082=1013%:%
%:%3083=1013%:%
%:%3084=1014%:%
%:%3085=1014%:%
%:%3086=1015%:%
%:%3087=1015%:%
%:%3088=1016%:%
%:%3089=1016%:%
%:%3090=1017%:%
%:%3091=1017%:%
%:%3092=1017%:%
%:%3093=1018%:%
%:%3094=1018%:%
%:%3095=1019%:%
%:%3096=1019%:%
%:%3097=1019%:%
%:%3098=1019%:%
%:%3099=1020%:%
%:%3100=1020%:%
%:%3101=1021%:%
%:%3102=1021%:%
%:%3103=1022%:%
%:%3104=1022%:%
%:%3105=1022%:%
%:%3106=1023%:%
%:%3107=1023%:%
%:%3108=1024%:%
%:%3109=1024%:%
%:%3110=1024%:%
%:%3111=1025%:%
%:%3112=1025%:%
%:%3113=1025%:%
%:%3114=1026%:%
%:%3115=1026%:%
%:%3116=1027%:%
%:%3117=1027%:%
%:%3118=1028%:%
%:%3119=1028%:%
%:%3120=1028%:%
%:%3121=1029%:%
%:%3122=1029%:%
%:%3123=1030%:%
%:%3124=1030%:%
%:%3125=1030%:%
%:%3126=1031%:%
%:%3132=1031%:%
%:%3135=1032%:%
%:%3136=1033%:%
%:%3137=1033%:%
%:%3138=1034%:%
%:%3139=1035%:%
%:%3140=1035%:%
%:%3141=1036%:%
%:%3142=1037%:%
%:%3143=1038%:%
%:%3144=1039%:%
%:%3145=1040%:%
%:%3146=1040%:%
%:%3147=1041%:%
%:%3154=1042%:%
%:%3155=1042%:%
%:%3156=1043%:%
%:%3157=1043%:%
%:%3158=1044%:%
%:%3159=1044%:%
%:%3160=1044%:%
%:%3161=1045%:%
%:%3162=1045%:%
%:%3163=1046%:%
%:%3164=1046%:%
%:%3165=1047%:%
%:%3166=1047%:%
%:%3167=1047%:%
%:%3168=1048%:%
%:%3169=1048%:%
%:%3170=1048%:%
%:%3171=1048%:%
%:%3172=1049%:%
%:%3178=1049%:%
%:%3181=1050%:%
%:%3182=1051%:%
%:%3183=1051%:%
%:%3190=1052%:%
%:%3191=1052%:%
%:%3192=1053%:%
%:%3193=1053%:%
%:%3194=1053%:%
%:%3195=1054%:%
%:%3196=1054%:%
%:%3197=1055%:%
%:%3198=1055%:%
%:%3199=1056%:%
%:%3200=1056%:%
%:%3201=1056%:%
%:%3202=1056%:%
%:%3203=1057%:%
%:%3209=1057%:%
%:%3212=1058%:%
%:%3213=1059%:%
%:%3214=1059%:%
%:%3221=1060%:%