### 2024-11-01 code/output/document/Nat_Miscellanea.tex

%
\begin{isabellebody}%
\setisabellecontext{Nat{\isacharunderscore}{\kern0pt}Miscellanea}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Auxiliary results on arithmetic%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Nat{\isacharunderscore}{\kern0pt}Miscellanea\ \isakeyword{imports}\ ZF\ \isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
Most of these results will get used at some point for the
calculation of arities.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemmas}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}succI\ {\isacharequal}{\kern0pt}\ \ Ord{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}mem{\isacharunderscore}{\kern0pt}iff\ {\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}OF\ nat{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}Ord{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}succD\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ \ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ n\ {\isasymin}\ m{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}drule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ltI{\isacharcomma}{\kern0pt}auto\ elim{\isacharcolon}{\kern0pt}ltD{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in\ {\isacharequal}{\kern0pt}\ \ ltD\ {\isacharbrackleft}{\kern0pt}OF\ nat{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharunderscore}{\kern0pt}le{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymin}\ m\ {\isasymLongrightarrow}\ n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}drule\ ltI{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}n{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ltI{\isacharunderscore}{\kern0pt}neg\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ j\ {\isasymle}\ x\ {\isasymLongrightarrow}\ j\ {\isasymnoteq}\ x\ {\isasymLongrightarrow}\ j\ {\isacharless}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ le{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}pred{\isacharunderscore}{\kern0pt}eq\ \ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymnoteq}\ {\isadigit{0}}\ \ {\isasymLongrightarrow}\ succ{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ elim{\isacharcolon}{\kern0pt}\ natE{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}ltI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ n\ {\isasymLongrightarrow}\ j\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ succ{\isacharunderscore}{\kern0pt}leE{\isacharbrackleft}{\kern0pt}OF\ leI{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}In\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymin}\ n\ {\isasymLongrightarrow}\ j\ {\isasymin}\ n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ succ{\isacharunderscore}{\kern0pt}ltI{\isacharbrackleft}{\kern0pt}THEN\ ltD{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ auto\ intro{\isacharcolon}{\kern0pt}\ ltI{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}leD\ {\isacharequal}{\kern0pt}\ succ{\isacharunderscore}{\kern0pt}leE{\isacharbrackleft}{\kern0pt}OF\ leI{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ succpred{\isacharunderscore}{\kern0pt}leI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ \ n\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ elim{\isacharcolon}{\kern0pt}\ natE{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ succpred{\isacharunderscore}{\kern0pt}n{\isadigit{0}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ p\ {\isasymLongrightarrow}\ p{\isasymnoteq}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ funcI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ A\ {\isasymrightarrow}\ B\ {\isasymLongrightarrow}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ b{\isacharequal}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ a\ {\isasymLongrightarrow}\ {\isasymlangle}a{\isacharcomma}{\kern0pt}\ b{\isasymrangle}\ {\isasymin}\ f{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all\ add{\isacharcolon}{\kern0pt}\ apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ natEin\ {\isacharequal}{\kern0pt}\ natE\ {\isacharbrackleft}{\kern0pt}OF\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}in\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymle}\ y\ \ {\isasymLongrightarrow}\ x\ {\isasymin}\ y{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ dest{\isacharcolon}{\kern0pt}ltD{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iffn\ {\isacharequal}{\kern0pt}\ \ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff\ {\isacharbrackleft}{\kern0pt}OF\ nat{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}Ord\ nat{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}Ord{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isasymLongrightarrow}\ n\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}n{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}le\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ m{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}n{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}leD{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}k{\isacharparenright}{\kern0pt}{\isasymLongrightarrow}\ \ i\ {\isasymunion}\ j\ {\isasymle}\ k\ {\isasymLongrightarrow}\ i\ {\isasymle}\ k{\isachardoublequoteclose}\ \ \ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{1}}{\isacharbrackleft}{\kern0pt}THEN\ conjunct{\isadigit{1}}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}leD{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}k{\isacharparenright}{\kern0pt}{\isasymLongrightarrow}\ \ i\ {\isasymunion}\ j\ {\isasymle}k\ {\isasymLongrightarrow}\ j\ {\isasymle}\ k{\isachardoublequoteclose}\ \ \ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{1}}{\isacharbrackleft}{\kern0pt}THEN\ conjunct{\isadigit{2}}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ gt{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ i\ {\isasymin}\ n\ {\isasymLongrightarrow}\ i\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ i\ {\isasymnoteq}\ {\isadigit{1}}\ {\isasymLongrightarrow}\ {\isadigit{1}}{\isacharless}{\kern0pt}i{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}erule\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}auto\ intro{\isacharcolon}{\kern0pt}\ Ord{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymle}\ m\ {\isasymLongrightarrow}\ pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ pred{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}n{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}le{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}erule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}m{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymle}\ m\ {\isasymLongrightarrow}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pred{\isadigit{2}}{\isacharunderscore}{\kern0pt}Un{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymle}\ j{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymle}\ j{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}m\ {\isasymunion}\ n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ pred{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}j{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ le{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt\ pred{\isacharunderscore}{\kern0pt}mono\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ i\ {\isasymle}\ j\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ i\ {\isasymunion}\ j\ {\isacharequal}{\kern0pt}\ j{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}absorb{\isadigit{1}}{\isacharcomma}{\kern0pt}erule\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ i\ {\isasymle}\ j\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ j\ {\isasymunion}\ i\ {\isacharequal}{\kern0pt}\ j{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}absorb{\isadigit{2}}{\isacharcomma}{\kern0pt}erule\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}un{\isacharunderscore}{\kern0pt}max\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ i\ {\isasymunion}\ j\ {\isacharequal}{\kern0pt}\ max{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}def\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ leI\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}max{\isacharunderscore}{\kern0pt}ty\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}Ord{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}max{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}nat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymnot}\ p{\isasymle}\ q\ {\isasymLongrightarrow}\ q\ {\isasymle}\ p{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ ltE{\isacharcomma}{\kern0pt}rule\ not{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}lt{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{1}}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}auto{\isacharcomma}{\kern0pt}drule\ ltI{\isacharbrackleft}{\kern0pt}of\ q\ p{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}auto{\isacharcomma}{\kern0pt}erule\ leI{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ {\isacharequal}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}un{\isacharunderscore}{\kern0pt}max\ nat{\isacharunderscore}{\kern0pt}max{\isacharunderscore}{\kern0pt}ty\ max{\isacharunderscore}{\kern0pt}def\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}succ\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}nat\ {\isasymLongrightarrow}\ x{\isasymle}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}pred\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}nat\ {\isasymLongrightarrow}\ pred{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymle}x{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}le{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ le{\isacharunderscore}{\kern0pt}succ{\isacharbrackright}{\kern0pt}\ pred{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}eq\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}compat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}o\ {\isasymle}\ p\ {\isasymLongrightarrow}\ q\ {\isasymle}\ r\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}o{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ o\ {\isasymunion}\ q\ {\isasymle}\ p\ {\isasymunion}\ r{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}of\ q\ r\ {\isachardoublequoteopen}p{\isasymunion}r{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}le{\isacharbrackright}{\kern0pt}\ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}of\ o\ p\ {\isachardoublequoteopen}p{\isasymunion}r{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{1}}{\isacharunderscore}{\kern0pt}le{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}le\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymle}\ r\ {\isasymLongrightarrow}\ q\ {\isasymle}\ r\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Ord{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ p\ {\isasymunion}\ q\ {\isasymle}\ r{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}leI{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}o\ {\isasymle}\ r\ {\isasymLongrightarrow}\ p\ {\isasymle}\ r\ {\isasymLongrightarrow}\ q\ {\isasymle}\ r\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Ord{\isacharparenleft}{\kern0pt}o{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ o\ {\isasymunion}\ p\ {\isasymunion}\ q\ {\isasymle}\ r{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymle}m{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}p\ {\isacharless}{\kern0pt}\ n{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}p{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}p\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}p\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}diff{\isacharunderscore}{\kern0pt}inverse{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ less{\isacharunderscore}{\kern0pt}diff{\isacharunderscore}{\kern0pt}conv{\isacharbrackleft}{\kern0pt}of\ n\ p\ {\isachardoublequoteopen}m\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ p{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}Un{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ y\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x\ {\isasymunion}\ Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ y\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}x\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Arith{\isachardot}{\kern0pt}pred{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymunion}\ y{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib\ pred{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}natI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}j\ {\isasymle}\ n\ {\isasymLongrightarrow}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ j{\isasymin}nat{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}drule\ ltD{\isacharcomma}{\kern0pt}rule\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}rule\ nat{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}of\ n{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}natE\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isasymin}nat\ {\isasymLongrightarrow}\ j\ {\isacharless}{\kern0pt}\ n\ {\isasymLongrightarrow}\ \ j{\isasymin}n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltE{\isacharbrackleft}{\kern0pt}of\ j\ n{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}simp{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}cancel\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ diff{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharunderscore}{\kern0pt}lemma\ leI\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ leD\ {\isacharcolon}{\kern0pt}\ \isakeyword{assumes}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymle}\ n{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ n\ {\isacharbar}{\kern0pt}\ j\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ leE{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}j{\isasymle}n{\isacartoucheclose}{\isacharcomma}{\kern0pt}of\ {\isachardoublequoteopen}j{\isacharless}{\kern0pt}n\ {\isacharbar}{\kern0pt}\ j\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Some results in ordinal arithmetic%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The following results are auxiliary to the proof of 
wellfoundedness of the relation \isa{frecR}%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}cong\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymle}\ y{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{shows}\ {\isachardoublequoteopen}max{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymle}\ max{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}y\ {\isasymle}\ z{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z\ {\isasymle}\ y{\isachardoublequoteclose}\ \ \isacommand{using}\isamarkupfalse%
\ assms\ not{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}lt\ leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}commutes\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}max{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ max{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ Un{\isacharunderscore}{\kern0pt}commute\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}cong{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymle}\ y{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}max{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymle}\ max{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ x\ {\isasymunion}\ z\ {\isasymle}\ y\ {\isasymunion}\ z{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}Ord\ Ord{\isacharunderscore}{\kern0pt}Un\ Un{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ \ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isasymle}y{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \ subset{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}le\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}D{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ {\isachardoublequoteopen}w\ {\isacharless}{\kern0pt}\ z{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}max{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}w{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ max{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}z{\isasymle}y{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isacharless}{\kern0pt}\ \ x\ {\isasymunion}\ w{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}lt{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}w{\isacharless}{\kern0pt}z{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ assms\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isacharless}{\kern0pt}\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ x\ w\ w{\isacharbrackright}{\kern0pt}\ lt{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}refl\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ y\ {\isasymunion}\ z{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ max{\isacharunderscore}{\kern0pt}commutes\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ assms\ leI\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}leD{\isadigit{2}}\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}D{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}w\ {\isacharequal}{\kern0pt}\ y\ {\isasymor}\ w\ {\isacharequal}{\kern0pt}\ z{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ y{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}max{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}w{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ max{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}w{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ z\ {\isasymunion}\ y{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}lt{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isacharless}{\kern0pt}y{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ y{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ w{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}w\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}d{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}w\ {\isacharequal}{\kern0pt}\ z{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ c\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ a\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ d\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}y\ {\isacharless}{\kern0pt}w{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isacharless}{\kern0pt}y{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isasymle}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le{\isacharbrackleft}{\kern0pt}OF\ assms{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ assms{\isacharparenleft}{\kern0pt}{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}w{\isacharequal}{\kern0pt}z{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}max{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \ \isacommand{unfolding}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x\ {\isasymunion}\ w{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ max{\isacharunderscore}{\kern0pt}commutes\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}iff\ assms\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ oadd{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymalpha}\ {\isacharless}{\kern0pt}\ {\isasymbeta}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isacharless}{\kern0pt}n{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymalpha}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ x\ {\isacharless}{\kern0pt}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}{\isasymbeta}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ y{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}{\isasymbeta}{\isacharequal}{\kern0pt}{\isadigit{0}}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}s{\isacharparenright}{\kern0pt}\ {\isasymgamma}\ \isakeyword{where}\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}{\isasymgamma}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymbeta}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}{\isasymgamma}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}cases{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharcomma}{\kern0pt}of\ {\isacharquery}{\kern0pt}thesis{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ cases\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{0}}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymalpha}{\isacharless}{\kern0pt}{\isasymbeta}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ s\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymalpha}{\isasymle}{\isasymgamma}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymalpha}{\isacharless}{\kern0pt}{\isasymbeta}{\isacartoucheclose}\ \isacommand{using}\isamarkupfalse%
\ leI\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymalpha}\ {\isasymle}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymgamma}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ omult{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}{\isasymalpha}{\isasymle}{\isasymgamma}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymalpha}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ x\ {\isacharless}{\kern0pt}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymgamma}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ oadd{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}x{\isacharless}{\kern0pt}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymbeta}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymbeta}{\isacharequal}{\kern0pt}succ{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ omult{\isacharunderscore}{\kern0pt}succ\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymalpha}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ x\ {\isacharless}{\kern0pt}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymbeta}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ oadd{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ lt{\isacharunderscore}{\kern0pt}trans{\isadigit{2}}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ l\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharless}{\kern0pt}n{\isacartoucheclose}\ lt{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ l\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ {\isasymbeta}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Limit{\isacharunderscore}{\kern0pt}has{\isacharunderscore}{\kern0pt}succ\ {\isacartoucheopen}{\isasymalpha}{\isacharless}{\kern0pt}{\isasymbeta}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymalpha}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ x\ {\isacharless}{\kern0pt}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymalpha}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ oadd{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ le{\isacharunderscore}{\kern0pt}refl{\isacharbrackleft}{\kern0pt}OF\ Ord{\isacharunderscore}{\kern0pt}omult{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}x{\isacharless}{\kern0pt}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ omult{\isacharunderscore}{\kern0pt}succ\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymalpha}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ x\ {\isacharless}{\kern0pt}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}succ{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ {\isasymbeta}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymalpha}\ {\isacharplus}{\kern0pt}{\isacharplus}{\kern0pt}\ x\ {\isacharless}{\kern0pt}\ n\ {\isacharasterisk}{\kern0pt}{\isacharasterisk}{\kern0pt}\ {\isasymbeta}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}trans\ omult{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isadigit{0}}{\isacharless}{\kern0pt}n{\isacartoucheclose}\ \ \isacommand{by}\isamarkupfalse%
\ auto\ \ \ \ \ \ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ oadd{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}{\isasymbeta}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ lt{\isacharunderscore}{\kern0pt}trans{\isadigit{2}}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Nat_Miscellanea.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%37=4%:%
%:%38=5%:%
%:%40=6%:%
%:%41=6%:%
%:%42=7%:%
%:%43=8%:%
%:%44=8%:%
%:%47=9%:%
%:%51=9%:%
%:%52=9%:%
%:%57=9%:%
%:%60=10%:%
%:%61=11%:%
%:%62=11%:%
%:%63=12%:%
%:%64=13%:%
%:%65=13%:%
%:%68=14%:%
%:%72=14%:%
%:%73=14%:%
%:%78=14%:%
%:%81=15%:%
%:%82=16%:%
%:%83=16%:%
%:%86=17%:%
%:%90=17%:%
%:%91=17%:%
%:%96=17%:%
%:%99=18%:%
%:%100=19%:%
%:%101=19%:%
%:%104=20%:%
%:%108=20%:%
%:%109=20%:%
%:%114=20%:%
%:%117=21%:%
%:%118=22%:%
%:%119=22%:%
%:%122=23%:%
%:%126=23%:%
%:%127=23%:%
%:%132=23%:%
%:%135=24%:%
%:%136=25%:%
%:%137=25%:%
%:%140=26%:%
%:%144=26%:%
%:%145=26%:%
%:%150=26%:%
%:%153=27%:%
%:%154=28%:%
%:%155=28%:%
%:%158=29%:%
%:%162=29%:%
%:%163=29%:%
%:%168=29%:%
%:%171=30%:%
%:%172=31%:%
%:%173=31%:%
%:%174=32%:%
%:%175=33%:%
%:%176=33%:%
%:%179=34%:%
%:%183=34%:%
%:%184=34%:%
%:%189=34%:%
%:%192=35%:%
%:%193=36%:%
%:%194=36%:%
%:%197=37%:%
%:%201=37%:%
%:%202=37%:%
%:%207=37%:%
%:%210=38%:%
%:%211=39%:%
%:%212=40%:%
%:%213=40%:%
%:%216=41%:%
%:%220=41%:%
%:%221=41%:%
%:%226=41%:%
%:%229=42%:%
%:%230=43%:%
%:%231=43%:%
%:%232=44%:%
%:%233=45%:%
%:%234=45%:%
%:%237=46%:%
%:%241=46%:%
%:%242=46%:%
%:%247=46%:%
%:%250=47%:%
%:%251=48%:%
%:%252=48%:%
%:%253=49%:%
%:%254=50%:%
%:%255=50%:%
%:%258=51%:%
%:%262=51%:%
%:%263=51%:%
%:%268=51%:%
%:%271=52%:%
%:%272=53%:%
%:%273=53%:%
%:%276=54%:%
%:%280=54%:%
%:%281=54%:%
%:%286=54%:%
%:%289=55%:%
%:%290=56%:%
%:%291=56%:%
%:%294=57%:%
%:%298=57%:%
%:%299=57%:%
%:%304=57%:%
%:%307=58%:%
%:%308=59%:%
%:%309=59%:%
%:%312=60%:%
%:%316=60%:%
%:%317=60%:%
%:%322=60%:%
%:%325=61%:%
%:%326=62%:%
%:%327=62%:%
%:%330=63%:%
%:%334=63%:%
%:%335=63%:%
%:%340=63%:%
%:%343=64%:%
%:%344=65%:%
%:%345=65%:%
%:%348=66%:%
%:%352=66%:%
%:%353=66%:%
%:%358=66%:%
%:%361=67%:%
%:%362=68%:%
%:%363=68%:%
%:%366=69%:%
%:%370=69%:%
%:%371=69%:%
%:%376=69%:%
%:%379=70%:%
%:%380=71%:%
%:%381=71%:%
%:%382=72%:%
%:%383=73%:%
%:%386=74%:%
%:%390=74%:%
%:%391=74%:%
%:%392=74%:%
%:%397=74%:%
%:%400=75%:%
%:%401=76%:%
%:%402=76%:%
%:%403=77%:%
%:%406=78%:%
%:%410=78%:%
%:%411=78%:%
%:%416=78%:%
%:%419=79%:%
%:%420=80%:%
%:%421=80%:%
%:%422=81%:%
%:%425=82%:%
%:%429=82%:%
%:%430=82%:%
%:%435=82%:%
%:%438=83%:%
%:%439=84%:%
%:%440=84%:%
%:%443=85%:%
%:%447=85%:%
%:%448=85%:%
%:%449=86%:%
%:%450=86%:%
%:%455=86%:%
%:%458=87%:%
%:%459=88%:%
%:%460=88%:%
%:%463=89%:%
%:%467=89%:%
%:%468=89%:%
%:%469=89%:%
%:%474=89%:%
%:%477=90%:%
%:%478=91%:%
%:%479=91%:%
%:%482=92%:%
%:%486=92%:%
%:%487=92%:%
%:%492=92%:%
%:%495=93%:%
%:%496=94%:%
%:%497=94%:%
%:%498=95%:%
%:%499=96%:%
%:%500=96%:%
%:%502=96%:%
%:%506=96%:%
%:%507=96%:%
%:%514=96%:%
%:%515=97%:%
%:%516=97%:%
%:%519=98%:%
%:%523=98%:%
%:%524=98%:%
%:%525=99%:%
%:%526=99%:%
%:%531=99%:%
%:%534=100%:%
%:%535=101%:%
%:%536=101%:%
%:%539=102%:%
%:%543=102%:%
%:%544=102%:%
%:%545=103%:%
%:%546=104%:%
%:%547=104%:%
%:%552=104%:%
%:%555=105%:%
%:%556=106%:%
%:%557=106%:%
%:%559=108%:%
%:%562=109%:%
%:%566=109%:%
%:%567=109%:%
%:%568=109%:%
%:%573=109%:%
%:%576=110%:%
%:%577=111%:%
%:%578=111%:%
%:%580=113%:%
%:%583=114%:%
%:%587=114%:%
%:%588=114%:%
%:%589=114%:%
%:%594=114%:%
%:%597=115%:%
%:%598=116%:%
%:%599=116%:%
%:%600=117%:%
%:%601=118%:%
%:%608=119%:%
%:%609=119%:%
%:%610=120%:%
%:%611=120%:%
%:%612=121%:%
%:%613=121%:%
%:%614=122%:%
%:%615=122%:%
%:%616=122%:%
%:%617=123%:%
%:%618=123%:%
%:%619=124%:%
%:%620=124%:%
%:%621=125%:%
%:%622=125%:%
%:%623=125%:%
%:%624=126%:%
%:%630=126%:%
%:%633=127%:%
%:%634=128%:%
%:%635=128%:%
%:%636=129%:%
%:%637=130%:%
%:%640=131%:%
%:%644=131%:%
%:%645=131%:%
%:%646=131%:%
%:%651=131%:%
%:%654=132%:%
%:%655=133%:%
%:%656=133%:%
%:%659=134%:%
%:%663=134%:%
%:%664=134%:%
%:%669=134%:%
%:%672=135%:%
%:%673=136%:%
%:%674=136%:%
%:%677=137%:%
%:%681=137%:%
%:%682=137%:%
%:%687=137%:%
%:%690=138%:%
%:%691=139%:%
%:%692=139%:%
%:%693=140%:%
%:%694=141%:%
%:%697=142%:%
%:%701=142%:%
%:%702=142%:%
%:%703=142%:%
%:%708=142%:%
%:%711=143%:%
%:%712=144%:%
%:%713=144%:%
%:%714=145%:%
%:%717=146%:%
%:%721=146%:%
%:%722=146%:%
%:%723=146%:%
%:%737=148%:%
%:%749=149%:%
%:%750=150%:%
%:%752=152%:%
%:%753=152%:%
%:%754=153%:%
%:%757=154%:%
%:%761=154%:%
%:%762=154%:%
%:%763=155%:%
%:%764=155%:%
%:%765=156%:%
%:%766=156%:%
%:%767=157%:%
%:%768=157%:%
%:%769=157%:%
%:%770=158%:%
%:%771=158%:%
%:%772=158%:%
%:%773=158%:%
%:%774=159%:%
%:%775=159%:%
%:%776=160%:%
%:%777=160%:%
%:%778=161%:%
%:%779=161%:%
%:%780=161%:%
%:%781=161%:%
%:%782=161%:%
%:%783=162%:%
%:%784=162%:%
%:%785=162%:%
%:%786=163%:%
%:%787=163%:%
%:%788=163%:%
%:%789=163%:%
%:%790=164%:%
%:%796=164%:%
%:%799=165%:%
%:%800=166%:%
%:%801=166%:%
%:%802=167%:%
%:%803=168%:%
%:%806=169%:%
%:%810=169%:%
%:%811=169%:%
%:%812=169%:%
%:%817=169%:%
%:%820=170%:%
%:%821=171%:%
%:%822=171%:%
%:%823=172%:%
%:%824=173%:%
%:%831=174%:%
%:%832=174%:%
%:%833=175%:%
%:%834=175%:%
%:%835=176%:%
%:%836=176%:%
%:%837=177%:%
%:%838=177%:%
%:%839=177%:%
%:%840=178%:%
%:%841=178%:%
%:%842=178%:%
%:%843=179%:%
%:%844=179%:%
%:%845=179%:%
%:%846=180%:%
%:%852=180%:%
%:%855=181%:%
%:%856=182%:%
%:%857=182%:%
%:%858=183%:%
%:%859=184%:%
%:%866=185%:%
%:%867=185%:%
%:%868=186%:%
%:%869=186%:%
%:%870=187%:%
%:%871=187%:%
%:%872=187%:%
%:%873=187%:%
%:%874=188%:%
%:%875=188%:%
%:%876=189%:%
%:%877=189%:%
%:%878=189%:%
%:%879=189%:%
%:%880=190%:%
%:%881=190%:%
%:%882=191%:%
%:%883=191%:%
%:%884=191%:%
%:%885=191%:%
%:%886=192%:%
%:%887=192%:%
%:%888=193%:%
%:%889=193%:%
%:%890=193%:%
%:%891=193%:%
%:%892=194%:%
%:%898=194%:%
%:%901=195%:%
%:%902=196%:%
%:%903=196%:%
%:%904=197%:%
%:%905=198%:%
%:%912=199%:%
%:%913=199%:%
%:%914=200%:%
%:%915=200%:%
%:%916=201%:%
%:%917=201%:%
%:%918=201%:%
%:%919=201%:%
%:%920=202%:%
%:%921=202%:%
%:%922=203%:%
%:%923=203%:%
%:%924=204%:%
%:%925=204%:%
%:%926=204%:%
%:%927=205%:%
%:%928=205%:%
%:%929=205%:%
%:%930=205%:%
%:%931=206%:%
%:%932=206%:%
%:%933=207%:%
%:%934=207%:%
%:%935=208%:%
%:%936=208%:%
%:%937=208%:%
%:%938=209%:%
%:%939=209%:%
%:%940=209%:%
%:%941=209%:%
%:%942=210%:%
%:%943=210%:%
%:%944=211%:%
%:%945=211%:%
%:%946=211%:%
%:%947=211%:%
%:%948=212%:%
%:%949=212%:%
%:%950=213%:%
%:%951=213%:%
%:%952=214%:%
%:%953=214%:%
%:%954=215%:%
%:%955=215%:%
%:%956=216%:%
%:%957=216%:%
%:%958=217%:%
%:%959=217%:%
%:%960=218%:%
%:%961=218%:%
%:%962=218%:%
%:%963=218%:%
%:%964=218%:%
%:%965=219%:%
%:%966=219%:%
%:%967=220%:%
%:%968=220%:%
%:%969=221%:%
%:%970=221%:%
%:%971=222%:%
%:%972=222%:%
%:%973=223%:%
%:%974=223%:%
%:%975=223%:%
%:%976=224%:%
%:%977=224%:%
%:%978=225%:%
%:%979=225%:%
%:%980=225%:%
%:%981=225%:%
%:%982=225%:%
%:%983=226%:%
%:%984=226%:%
%:%985=227%:%
%:%986=227%:%
%:%987=227%:%
%:%988=227%:%
%:%989=228%:%
%:%990=228%:%
%:%991=228%:%
%:%992=228%:%
%:%993=228%:%
%:%994=229%:%
%:%995=229%:%
%:%996=230%:%
%:%997=230%:%
%:%998=231%:%
%:%999=231%:%
%:%1000=232%:%
%:%1001=232%:%
%:%1002=233%:%
%:%1003=233%:%
%:%1004=233%:%
%:%1005=233%:%
%:%1006=234%:%
%:%1007=234%:%
%:%1008=235%:%
%:%1014=235%:%
%:%1017=236%:%
%:%1018=237%:%
%:%1019=237%:%
%:%1020=238%:%
%:%1021=239%:%
%:%1028=240%:%
%:%1029=240%:%
%:%1030=241%:%
%:%1031=241%:%
%:%1032=242%:%
%:%1033=242%:%
%:%1034=242%:%
%:%1035=243%:%
%:%1036=243%:%
%:%1037=243%:%
%:%1038=244%:%
%:%1039=244%:%
%:%1040=245%:%
%:%1041=245%:%
%:%1042=246%:%
%:%1043=246%:%
%:%1044=246%:%
%:%1045=246%:%
%:%1046=246%:%
%:%1047=247%:%
%:%1048=247%:%
%:%1049=248%:%
%:%1050=248%:%
%:%1051=249%:%
%:%1052=249%:%
%:%1053=250%:%
%:%1054=250%:%
%:%1055=250%:%
%:%1056=250%:%
%:%1057=250%:%
%:%1058=251%:%
%:%1059=251%:%
%:%1060=252%:%
%:%1061=252%:%
%:%1062=252%:%
%:%1063=252%:%
%:%1064=253%:%
%:%1065=253%:%
%:%1066=254%:%
%:%1067=254%:%
%:%1068=254%:%
%:%1069=254%:%
%:%1070=255%:%
%:%1071=255%:%
%:%1072=256%:%
%:%1073=256%:%
%:%1074=256%:%
%:%1075=256%:%
%:%1076=257%:%
%:%1077=257%:%
%:%1078=258%:%
%:%1079=258%:%
%:%1080=258%:%
%:%1081=259%:%
%:%1082=259%:%
%:%1083=260%:%
%:%1084=260%:%
%:%1085=260%:%
%:%1086=260%:%
%:%1087=261%:%
%:%1088=261%:%
%:%1089=262%:%
%:%1090=262%:%
%:%1091=263%:%
%:%1092=263%:%
%:%1093=263%:%
%:%1094=263%:%
%:%1095=264%:%
%:%1096=264%:%
%:%1097=265%:%
%:%1098=265%:%
%:%1099=265%:%
%:%1100=265%:%
%:%1101=266%:%
%:%1102=266%:%
%:%1103=267%:%
%:%1104=267%:%
%:%1105=267%:%
%:%1106=268%:%
%:%1107=268%:%
%:%1108=269%:%
%:%1109=269%:%
%:%1110=269%:%
%:%1111=269%:%
%:%1112=270%:%
%:%1113=270%:%
%:%1114=271%:%
%:%1115=271%:%
%:%1116=271%:%
%:%1117=272%:%
%:%1118=272%:%
%:%1119=273%:%
%:%1120=273%:%
%:%1121=273%:%
%:%1122=273%:%
%:%1123=274%:%
%:%1124=274%:%
%:%1125=274%:%
%:%1126=274%:%
%:%1127=274%:%
%:%1128=275%:%
%:%1129=275%:%
%:%1130=276%:%
%:%1136=276%:%
%:%1145=277%:%