### 2024-11-01 code/output/document/NotAC_Binmap.tex

%
\begin{isabellebody}%
\setisabellecontext{NotAC{\isacharunderscore}{\kern0pt}Binmap}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ NotAC{\isacharunderscore}{\kern0pt}Binmap\isanewline
\ \ \isakeyword{imports}\ Fn{\isacharunderscore}{\kern0pt}Perm{\isacharunderscore}{\kern0pt}Filter\ HS{\isacharunderscore}{\kern0pt}Forces\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ \isakeyword{begin}\ \isanewline
\isanewline
\isacommand{interpretation}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\ {\isachardoublequoteopen}Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ Fn{\isachardot}{\kern0pt}\ {\isasymexists}F\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}\ {\isasymand}\ F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row\ \isakeyword{where}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ G{\isachardot}{\kern0pt}\ p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}binmap{\isacharprime}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isacharbraceright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}\ {\isadigit{0}}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ binmap\ \isakeyword{where}\ {\isachardoublequoteopen}binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ f\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}\ {\isasymin}\ nat\ {\isasymtimes}\ Fn{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}A{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ F\ \isakeyword{where}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}F\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}\ {\isasymin}\ nat\ {\isasymtimes}\ Fn{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ F\ \isakeyword{where}\ mFH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}F\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}check{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ mFH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ fn{\isacharcomma}{\kern0pt}\ N{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ \isanewline
\ \ \ \ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ fn\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ N\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ k\ l\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ \isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ k\ l\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}k{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ rcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}tran{\isacharunderscore}{\kern0pt}closure{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm\ {\isacharbrackleft}{\kern0pt}\isakeyword{where}\ i{\isacharequal}{\kern0pt}{\isadigit{7}}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Hcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Replace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ PHcheck{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{1}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}trans\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ k\ l\ x\ n\ env\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nat\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}l{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}F\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}\ {\isasymand}\ F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymexists}Fnm\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}nm\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}checkm\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}F\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}zero\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ zero\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymand}\ F\ {\isasymin}\ Fn\ {\isasymand}\ m\ {\isasymin}\ nat\ {\isasymand}\ checkm\ {\isacharequal}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}checkm{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}\ {\isasymand}\ nm\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymand}\ F{\isacharbackquote}{\kern0pt}nm\ {\isacharequal}{\kern0pt}\ Fnm\ {\isasymand}\ Fnm\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}empty{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}check{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ check{\isacharunderscore}{\kern0pt}abs\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{8}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ F\ m{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ F\ m{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ F\ m{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ F\ m{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ nth{\isacharunderscore}{\kern0pt}type\ apply{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ X\ \isakeyword{where}\ {\isachardoublequoteopen}X\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ Fn{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ Fn{\isacharcomma}{\kern0pt}\ nat{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}X\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ X{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}notation{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}check{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ check{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ cartprod{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}X\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ X{\isacharunderscore}{\kern0pt}def\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ transM\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ blast{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}X\ {\isasymin}\ M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ checknatdom{\isacharcomma}{\kern0pt}\ fn{\isacharcomma}{\kern0pt}\ N{\isacharparenright}{\kern0pt}\ {\isasymequiv}\isanewline
\ \ \ \ Forall{\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ y{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}cartprod{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}checknatdom{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ fn{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ \isanewline
\ \ \ \ \ \ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ fn{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ N{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ k\ l\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ k\ l\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}k{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ assms\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}of\ {\isadigit{0}}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{1}}{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}cartprod{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{1}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{1}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ succ{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}distrib{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{1}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ i\ j\ k\ l\ m\ y\ n\ env\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}l{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharcomma}{\kern0pt}\ l{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ y\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}v\ {\isasymin}\ M{\isachardot}{\kern0pt}\ v\ {\isasymin}\ y\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymexists}A\ {\isasymin}\ M{\isachardot}{\kern0pt}\ A\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ Fn\ {\isasymand}\ v\ {\isasymin}\ A{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}F\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ {\isasymexists}m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}\ {\isasymand}\ F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Forall{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Iff{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ assms\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}cartprod{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ assms\ nth{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ cartprod{\isacharunderscore}{\kern0pt}abs{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ nth{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ assms\ nth{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ nth{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ assms\ nth{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}member{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ length{\isacharunderscore}{\kern0pt}type\ assms\ nth{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{9}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}v\ {\isasymin}\ M{\isachardot}{\kern0pt}\ v\ {\isasymin}\ y\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ball{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\ domain{\isacharunderscore}{\kern0pt}closed\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ transM\ nth{\isacharunderscore}{\kern0pt}type\ length{\isacharunderscore}{\kern0pt}type\ assms\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ transM\ nth{\isacharunderscore}{\kern0pt}type\ length{\isacharunderscore}{\kern0pt}type\ assms\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}binmap{\isacharprime}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ X\ \isakeyword{where}\ {\isachardoublequoteopen}X\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ Fn{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardot}{\kern0pt}\ {\isasymexists}n\ {\isasymin}\ M{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymand}\ x\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}X\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ Fn{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharcomma}{\kern0pt}\ nat{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}A{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ X{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}A\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}notation{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}of\ {\isadigit{1}}\ {\isadigit{0}}\ {\isadigit{2}}\ {\isadigit{3}}\ {\isadigit{4}}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ M{\isacharunderscore}{\kern0pt}powerset{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ cartprod{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}X\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}X\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}A{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}X\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}B{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ X{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}B\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}X\ {\isasymin}\ M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}B\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}\ {\isadigit{0}}\ {\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ cartprod{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \ \ \ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}pauto\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ f\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}A\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}B{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}p\ m{\isachardot}{\kern0pt}\ p\ {\isasymin}\ Fn\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ p\ m\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ l\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ domin{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isadigit{1}}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isasymrangle}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isasymrangle}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isasymrangle}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n{\isacharprime}{\kern0pt}{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}m{\isacharprime}{\kern0pt}{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}l\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isasymlangle}{\isasymlangle}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}\ {\isasymin}\ p\ {\isasymand}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymlangle}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isasymrangle}\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ H\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isasymrangle}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isadigit{1}}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms\ domin\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymlongleftrightarrow}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}not{\isadigit{0}}{\isacharunderscore}{\kern0pt}indom\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}f\ x{\isachardot}{\kern0pt}\ function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ f\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ apply{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ indom\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}m\ p{\isachardot}{\kern0pt}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasymin}\ Fn\ {\isasymLongrightarrow}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isasymLongrightarrow}\ p\ {\isasymin}\ Fn\ {\isasymLongrightarrow}\ {\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ m\ p\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ apply{\isacharunderscore}{\kern0pt}not{\isadigit{0}}{\isacharunderscore}{\kern0pt}indom{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasymin}\ Fn{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ l\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ assms\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n{\isacharprime}{\kern0pt}{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}m{\isacharprime}{\kern0pt}{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}l{\isacharprime}{\kern0pt}\ {\isasymin}\ {\isadigit{2}}{\isachardot}{\kern0pt}\ {\isasymlangle}{\isasymlangle}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isasymrangle}\ {\isasymin}\ p\ {\isasymand}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharprime}{\kern0pt}{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}\ \isakeyword{where}\ n{\isacharprime}{\kern0pt}H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}A\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ F{\isachargreater}{\kern0pt}\ {\isasymin}\ nat\ {\isasymtimes}\ Fn{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isacharbraceright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ nat\ {\isasymtimes}\ Fn{\isacharcomma}{\kern0pt}\ p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ nat\ {\isasymtimes}\ Fn{\isacharcomma}{\kern0pt}\ p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ m\ F{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ check{\isacharunderscore}{\kern0pt}fixpoint{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ m\ F{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ check{\isacharunderscore}{\kern0pt}fixpoint{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ nat\ {\isasymtimes}\ Fn{\isacharcomma}{\kern0pt}\ p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ m\ p{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharcomma}{\kern0pt}\ rule\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ H{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ apply{\isacharunderscore}{\kern0pt}not{\isadigit{0}}{\isacharunderscore}{\kern0pt}indom{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ m\ p{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ m\ p{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymexists}y{\isasymin}Fn{\isachardot}{\kern0pt}\ p\ {\isacharequal}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ m\ p{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}p\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ H{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ indom{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ apply{\isacharunderscore}{\kern0pt}not{\isadigit{0}}{\isacharunderscore}{\kern0pt}indom{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ p\ {\isasymin}\ Fn{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}bij\ assms\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ surj{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}check{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}symmetric\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}symmetric{\isacharparenleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Fix{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}n{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ sym{\isacharparenleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ F\ \isacommand{assume}\isamarkupfalse%
\ Fin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F\ {\isasymin}\ Fix{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}n{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ f\ \isakeyword{where}\ fH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ F{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fix{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}pauto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ fH\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ fH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}F\ {\isasymin}\ sym{\isacharparenleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ fH\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ assms\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ symmetric{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ sym{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroup{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}n{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{1}}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isachargreater}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ not{\isacharunderscore}{\kern0pt}emptyI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ range{\isacharunderscore}{\kern0pt}def\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ assms\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ H\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ HS{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharprime}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ HS{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}symmetric{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharprime}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pauto{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ F\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}F\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}{\isadigit{0}}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}one{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}pnauto\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ F\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}F\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}binmap{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ f\ \isakeyword{where}\ fH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F\ {\isacharequal}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}binmap{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ binmap{\isacharprime}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ rule\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ binmap{\isacharprime}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ vin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ binmap{\isacharprime}{\kern0pt}\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ mH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isachargreater}{\kern0pt}\ {\isasymin}\ binmap{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}{\isadigit{0}}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ fH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pauto{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ assms\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}pauto\ fH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ binmap{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}m\ {\isasymin}\ nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mH\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ fH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ f\ \isakeyword{where}\ fH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F\ {\isacharequal}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ binmap{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ mH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ fH\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}m\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}m{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ mH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ fH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}pauto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ bij{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}bij\ {\isacharbrackleft}{\kern0pt}of\ f\ nat\ nat{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ fH\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ mH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ right{\isacharunderscore}{\kern0pt}inverse{\isacharunderscore}{\kern0pt}bij{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fH\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ mH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ eq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}{\isadigit{0}}{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ rule\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pair{\isacharunderscore}{\kern0pt}relI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}m{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ bij{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}bij\ {\isacharbrackleft}{\kern0pt}of\ f\ nat\ nat{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ fH\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ mH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isachargreater}{\kern0pt}\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ eq\ pauto{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}binmap{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}sym{\isacharparenleft}{\kern0pt}binmap{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}pnauto\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}binmap{\isacharprime}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ HS{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ symmetric{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ sym{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}subgroup{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ not{\isacharunderscore}{\kern0pt}emptyI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ inj{\isacharunderscore}{\kern0pt}def\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ sym{\isacharunderscore}{\kern0pt}binmap{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Fix{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cons\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}finite{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymnotin}\ A{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}finite{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ n\ f\ \isakeyword{where}\ nfH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ g\ \isakeyword{where}\ {\isachardoublequoteopen}g\ {\isasymequiv}\ f\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymnotin}\ f{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ functionH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\ g{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ assms\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ domH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ cons{\isacharunderscore}{\kern0pt}def\ assms\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ ranH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ gpiH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}g\ {\isasymin}\ cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ g{\isacharunderscore}{\kern0pt}def\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ functionH\ domH\ ranH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ appAH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}a{\isachardot}{\kern0pt}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ g{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ disjI{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ g{\isacharunderscore}{\kern0pt}def\ functionH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ appAH{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}a{\isachardot}{\kern0pt}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ g{\isacharbackquote}{\kern0pt}a\ {\isasymin}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ a\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ ain\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ A{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ appAH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ ain\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ mH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ f{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ mH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}a\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}f{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ m{\isacartoucheclose}\ {\isacartoucheopen}g{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\ \ \ \ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ appxH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}def\ functionH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ appneq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}a{\isachardot}{\kern0pt}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ g{\isacharbackquote}{\kern0pt}a\ {\isasymnoteq}\ g{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ appAH{\isacharprime}{\kern0pt}\ appxH\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}n\ {\isasymin}\ n{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ rule\ mem{\isacharunderscore}{\kern0pt}irrefl{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ appneq{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}a\ b{\isachardot}{\kern0pt}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ b\ {\isasymin}\ A\ {\isasymLongrightarrow}\ g{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}b\ {\isasymLongrightarrow}\ a\ {\isacharequal}{\kern0pt}\ b{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ appAH\ nfH\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}a\ b{\isachardot}{\kern0pt}\ a\ {\isasymin}\ cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ b\ {\isasymin}\ cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ g{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}b\ {\isasymLongrightarrow}\ a\ {\isacharequal}{\kern0pt}\ b{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ appneq\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ appneq{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}cons{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ inj{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ gpiH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ g{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nfH\ Un{\isacharunderscore}{\kern0pt}closed\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ G\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ G{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{thm}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\isanewline
\ \ filter{\isacharunderscore}{\kern0pt}def\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}Fn{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ dense{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ forcing{\isacharunderscore}{\kern0pt}notion{\isachardot}{\kern0pt}dense{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\ forcing{\isacharunderscore}{\kern0pt}notion{\isacharunderscore}{\kern0pt}axioms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}G\ {\isasyminter}\ Fn\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ assms\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ pH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ G{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}leqD{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ pH\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ pH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ G\ \ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ vin{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ m\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ {\isasymlangle}m{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}v\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ def{\isacharunderscore}{\kern0pt}val{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ vin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ p\ \isakeyword{where}\ mpH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ mpH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mpH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ p\ \isakeyword{where}\ mpH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}check{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mpH\ M{\isacharunderscore}{\kern0pt}genericD\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ m\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymexists}p\ {\isasymin}\ Fn{\isachardot}{\kern0pt}\ {\isasymlangle}m{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mpH\ M{\isacharunderscore}{\kern0pt}genericD\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}val{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}m\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mpH\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sym{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ G\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ def{\isacharunderscore}{\kern0pt}val{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ assms\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n\ m{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}value\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ p\ q\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}q{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}q{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ v\ \isakeyword{where}\ vH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ v{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ vH{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ v{\isachargreater}{\kern0pt}\ {\isasymin}\ q{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ papp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ vH\ assms\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ qapp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}q{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ vH\ assms\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}q{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}m{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ papp\ qapp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}forces\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ p\ n\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ G\ {\isasymlongrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Member{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ subst\ SymExt{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ assms\ transM\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ assms\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ G{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m\ {\isasymand}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ assms\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharunderscore}{\kern0pt}forces\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ p\ n\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n\ {\isacharcomma}{\kern0pt}m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ G\ {\isasymlongrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ {\isacharparenleft}{\kern0pt}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ G\ {\isasymlongrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ G\ \isakeyword{where}\ GH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymnot}{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ listin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ assms\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ assms\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ listin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapin\ GH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m\ {\isasymin}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m\ {\isasymand}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ assms\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq\ GH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ qH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}q\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ domin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ qH\ apply{\isacharunderscore}{\kern0pt}{\isadigit{0}}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}G{\isachardot}{\kern0pt}\ {\isasymforall}q{\isasymin}G{\isachardot}{\kern0pt}\ compat{\isacharunderscore}{\kern0pt}in{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GH\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ r\ \isakeyword{where}\ rH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}r{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}r{\isacharcomma}{\kern0pt}\ q{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ G{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ qH\ GH\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ q{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}value\ domin\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ v{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}r{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ qH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ rH\ GH\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}value\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}False{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ v{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}distinct{\isacharunderscore}{\kern0pt}helper\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ m\ p\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymnoteq}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ False\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A{\isachardot}{\kern0pt}\ A\ {\isasymsubseteq}\ nat\ {\isasymLongrightarrow}\ Finite{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}a\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ a\ {\isasymnotin}\ A{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ A{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}A\ {\isacharequal}{\kern0pt}\ nat{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}Finite\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equalityI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}finite{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ nat\ {\isasymtimes}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ finitedom\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ n{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ X{\isacharequal}{\kern0pt}n\ \isakeyword{in}\ lepoll{\isacharunderscore}{\kern0pt}Finite{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ lepoll{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ not{\isacharunderscore}{\kern0pt}emptyE{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ nat{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}Finite{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ range{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}A\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Finite{\isacharunderscore}{\kern0pt}RepFun{\isacharcomma}{\kern0pt}\ rule\ finitedom{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}range{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}A\ {\isacharequal}{\kern0pt}\ range{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}x\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ x\ {\isasymnotin}\ range{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ H{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ mH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymnotin}\ range{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ q\ \isakeyword{where}\ {\isachardoublequoteopen}q\ {\isasymequiv}\ p\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ q{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ cons{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ cons{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ domfin{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}finite{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cons{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ mH\ Fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ qinFn\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}q\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ q{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}def\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ mH\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}def\ assms\ Un{\isacharunderscore}{\kern0pt}closed\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ mH\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ mH\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{7}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ mH\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domfin\ q{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ mH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ appn\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}q{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ q{\isacharunderscore}{\kern0pt}def\ qinFn\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ appn{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}q{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ q{\isacharunderscore}{\kern0pt}def\ qinFn\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \ \ \ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ listin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}G{\isachardot}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G\ {\isasymlongrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ qinFn\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ p{\isacharequal}{\kern0pt}p\ \isakeyword{in}\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ qinFn\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\ q{\isacharunderscore}{\kern0pt}def\ listin\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ eqH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapin\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G\ {\isasymlongrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}{\isadigit{1}}{\isacharunderscore}{\kern0pt}forces{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ mH\ qinFn\ appn\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ memH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ allI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ G{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ listin\ mH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G\ {\isasymlongrightarrow}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharunderscore}{\kern0pt}forces{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ mH\ qinFn\ appn{\isacharprime}{\kern0pt}\ q{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymnotin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ allI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ G{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ listin\ mH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ neqH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ memH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}existence\ qinFn\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ False\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ eqH\ neqH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}distinct\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ n\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymnoteq}\ m{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ Neg{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ ForcesHS{\isacharunderscore}{\kern0pt}Neg{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}distinct{\isacharunderscore}{\kern0pt}helper\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}distinct\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ G\ n\ m\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymnoteq}\ m{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ listin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ Neg{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}distinct{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}H{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isadigit{0}}{\isasymin}H\ \ {\isasymlongrightarrow}\ \ SymExt{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ listin\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ Neg{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ local{\isachardot}{\kern0pt}generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isacharquery}{\kern0pt}A{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ listin\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ rowneq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/NotAC_Binmap.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%18=3%:%
%:%21=4%:%
%:%22=5%:%
%:%23=5%:%
%:%24=6%:%
%:%25=7%:%
%:%26=7%:%
%:%29=8%:%
%:%33=8%:%
%:%34=8%:%
%:%35=8%:%
%:%40=8%:%
%:%43=9%:%
%:%44=10%:%
%:%45=10%:%
%:%46=11%:%
%:%47=11%:%
%:%48=12%:%
%:%49=12%:%
%:%50=13%:%
%:%51=13%:%
%:%52=14%:%
%:%53=15%:%
%:%54=16%:%
%:%55=16%:%
%:%56=17%:%
%:%57=18%:%
%:%58=19%:%
%:%61=20%:%
%:%66=21%:%
%:%67=21%:%
%:%68=22%:%
%:%69=22%:%
%:%70=23%:%
%:%71=23%:%
%:%72=24%:%
%:%73=24%:%
%:%74=24%:%
%:%75=25%:%
%:%76=25%:%
%:%77=26%:%
%:%78=26%:%
%:%79=27%:%
%:%80=27%:%
%:%81=27%:%
%:%82=28%:%
%:%83=28%:%
%:%84=29%:%
%:%85=29%:%
%:%86=30%:%
%:%87=30%:%
%:%88=31%:%
%:%89=31%:%
%:%90=32%:%
%:%91=32%:%
%:%92=32%:%
%:%93=33%:%
%:%94=33%:%
%:%95=34%:%
%:%96=34%:%
%:%97=35%:%
%:%98=35%:%
%:%99=35%:%
%:%100=36%:%
%:%101=36%:%
%:%102=37%:%
%:%103=37%:%
%:%104=37%:%
%:%105=38%:%
%:%106=38%:%
%:%107=39%:%
%:%108=39%:%
%:%109=40%:%
%:%115=40%:%
%:%118=41%:%
%:%119=42%:%
%:%120=42%:%
%:%121=43%:%
%:%130=52%:%
%:%131=53%:%
%:%132=54%:%
%:%133=54%:%
%:%134=55%:%
%:%135=56%:%
%:%136=57%:%
%:%139=58%:%
%:%140=59%:%
%:%144=59%:%
%:%145=59%:%
%:%146=60%:%
%:%147=60%:%
%:%148=61%:%
%:%149=61%:%
%:%150=62%:%
%:%151=62%:%
%:%152=63%:%
%:%153=63%:%
%:%154=64%:%
%:%160=64%:%
%:%163=65%:%
%:%164=66%:%
%:%165=66%:%
%:%166=67%:%
%:%167=68%:%
%:%168=69%:%
%:%171=70%:%
%:%172=71%:%
%:%176=71%:%
%:%177=71%:%
%:%178=72%:%
%:%179=72%:%
%:%180=73%:%
%:%181=73%:%
%:%182=74%:%
%:%183=74%:%
%:%184=75%:%
%:%185=75%:%
%:%186=76%:%
%:%187=76%:%
%:%188=77%:%
%:%189=77%:%
%:%190=78%:%
%:%191=78%:%
%:%192=79%:%
%:%193=79%:%
%:%194=80%:%
%:%195=80%:%
%:%196=81%:%
%:%197=81%:%
%:%198=82%:%
%:%199=82%:%
%:%200=83%:%
%:%201=83%:%
%:%202=84%:%
%:%203=84%:%
%:%204=85%:%
%:%205=85%:%
%:%206=86%:%
%:%207=86%:%
%:%208=87%:%
%:%209=87%:%
%:%210=88%:%
%:%211=88%:%
%:%212=89%:%
%:%213=89%:%
%:%214=90%:%
%:%215=90%:%
%:%216=91%:%
%:%217=91%:%
%:%218=92%:%
%:%219=92%:%
%:%220=93%:%
%:%221=93%:%
%:%222=94%:%
%:%223=94%:%
%:%224=95%:%
%:%225=95%:%
%:%226=96%:%
%:%227=96%:%
%:%228=97%:%
%:%229=97%:%
%:%230=98%:%
%:%236=98%:%
%:%239=99%:%
%:%240=100%:%
%:%241=100%:%
%:%242=101%:%
%:%243=102%:%
%:%244=103%:%
%:%245=104%:%
%:%246=105%:%
%:%249=106%:%
%:%250=107%:%
%:%254=107%:%
%:%255=107%:%
%:%256=108%:%
%:%257=108%:%
%:%258=109%:%
%:%259=110%:%
%:%260=110%:%
%:%261=111%:%
%:%262=111%:%
%:%263=112%:%
%:%264=112%:%
%:%265=113%:%
%:%266=113%:%
%:%267=114%:%
%:%268=114%:%
%:%269=115%:%
%:%270=115%:%
%:%271=116%:%
%:%272=116%:%
%:%273=117%:%
%:%274=117%:%
%:%275=118%:%
%:%276=118%:%
%:%277=119%:%
%:%278=119%:%
%:%279=120%:%
%:%280=120%:%
%:%281=121%:%
%:%282=121%:%
%:%283=122%:%
%:%284=122%:%
%:%285=123%:%
%:%286=123%:%
%:%287=124%:%
%:%288=124%:%
%:%289=125%:%
%:%290=125%:%
%:%291=126%:%
%:%292=126%:%
%:%293=127%:%
%:%294=127%:%
%:%295=128%:%
%:%296=128%:%
%:%297=129%:%
%:%298=129%:%
%:%299=130%:%
%:%300=130%:%
%:%301=131%:%
%:%302=131%:%
%:%303=132%:%
%:%304=132%:%
%:%305=133%:%
%:%306=133%:%
%:%307=134%:%
%:%308=134%:%
%:%309=135%:%
%:%310=135%:%
%:%311=136%:%
%:%312=136%:%
%:%313=137%:%
%:%314=137%:%
%:%315=138%:%
%:%316=138%:%
%:%317=139%:%
%:%318=139%:%
%:%319=140%:%
%:%320=140%:%
%:%321=141%:%
%:%322=141%:%
%:%323=142%:%
%:%324=142%:%
%:%325=143%:%
%:%326=143%:%
%:%327=144%:%
%:%328=144%:%
%:%329=145%:%
%:%330=145%:%
%:%335=145%:%
%:%338=146%:%
%:%339=147%:%
%:%340=147%:%
%:%341=148%:%
%:%342=149%:%
%:%343=150%:%
%:%350=151%:%
%:%351=151%:%
%:%352=152%:%
%:%353=152%:%
%:%354=153%:%
%:%355=154%:%
%:%356=154%:%
%:%357=155%:%
%:%358=155%:%
%:%359=156%:%
%:%360=156%:%
%:%361=157%:%
%:%362=157%:%
%:%363=158%:%
%:%364=158%:%
%:%365=159%:%
%:%366=159%:%
%:%367=160%:%
%:%368=160%:%
%:%369=161%:%
%:%370=161%:%
%:%371=162%:%
%:%372=162%:%
%:%373=163%:%
%:%374=163%:%
%:%375=164%:%
%:%376=164%:%
%:%377=165%:%
%:%378=165%:%
%:%379=166%:%
%:%380=166%:%
%:%381=167%:%
%:%382=167%:%
%:%383=168%:%
%:%384=168%:%
%:%385=169%:%
%:%386=169%:%
%:%387=170%:%
%:%388=171%:%
%:%389=171%:%
%:%390=172%:%
%:%391=172%:%
%:%392=173%:%
%:%393=173%:%
%:%394=174%:%
%:%395=174%:%
%:%396=175%:%
%:%397=175%:%
%:%398=176%:%
%:%399=176%:%
%:%400=177%:%
%:%401=177%:%
%:%402=178%:%
%:%403=179%:%
%:%404=179%:%
%:%405=179%:%
%:%406=179%:%
%:%407=179%:%
%:%408=180%:%
%:%414=180%:%
%:%417=181%:%
%:%418=182%:%
%:%419=182%:%
%:%420=183%:%
%:%423=186%:%
%:%424=187%:%
%:%425=188%:%
%:%426=188%:%
%:%427=189%:%
%:%428=190%:%
%:%429=191%:%
%:%432=192%:%
%:%436=192%:%
%:%437=192%:%
%:%438=193%:%
%:%439=193%:%
%:%440=194%:%
%:%441=194%:%
%:%442=195%:%
%:%443=195%:%
%:%444=196%:%
%:%445=196%:%
%:%446=197%:%
%:%447=197%:%
%:%452=197%:%
%:%455=198%:%
%:%456=199%:%
%:%457=199%:%
%:%458=200%:%
%:%459=201%:%
%:%460=202%:%
%:%463=203%:%
%:%467=203%:%
%:%468=203%:%
%:%469=204%:%
%:%470=204%:%
%:%471=205%:%
%:%472=205%:%
%:%473=206%:%
%:%474=206%:%
%:%475=207%:%
%:%476=207%:%
%:%477=208%:%
%:%478=208%:%
%:%479=209%:%
%:%480=209%:%
%:%481=210%:%
%:%482=210%:%
%:%483=211%:%
%:%484=211%:%
%:%485=212%:%
%:%486=212%:%
%:%487=213%:%
%:%488=213%:%
%:%489=214%:%
%:%490=214%:%
%:%491=215%:%
%:%492=215%:%
%:%493=216%:%
%:%494=216%:%
%:%495=217%:%
%:%496=217%:%
%:%497=218%:%
%:%498=218%:%
%:%499=219%:%
%:%500=219%:%
%:%501=220%:%
%:%502=220%:%
%:%503=221%:%
%:%504=221%:%
%:%505=222%:%
%:%506=222%:%
%:%507=223%:%
%:%508=223%:%
%:%509=224%:%
%:%510=224%:%
%:%511=225%:%
%:%512=225%:%
%:%513=226%:%
%:%514=226%:%
%:%515=227%:%
%:%516=227%:%
%:%517=228%:%
%:%523=228%:%
%:%526=229%:%
%:%527=230%:%
%:%528=230%:%
%:%529=231%:%
%:%530=232%:%
%:%531=233%:%
%:%532=234%:%
%:%533=235%:%
%:%536=236%:%
%:%537=237%:%
%:%541=237%:%
%:%542=237%:%
%:%543=238%:%
%:%544=238%:%
%:%545=239%:%
%:%546=239%:%
%:%547=240%:%
%:%548=240%:%
%:%549=241%:%
%:%550=241%:%
%:%551=242%:%
%:%552=242%:%
%:%553=243%:%
%:%554=243%:%
%:%555=244%:%
%:%556=244%:%
%:%557=245%:%
%:%558=245%:%
%:%559=246%:%
%:%560=246%:%
%:%561=247%:%
%:%562=247%:%
%:%563=248%:%
%:%564=248%:%
%:%565=249%:%
%:%566=249%:%
%:%567=250%:%
%:%568=250%:%
%:%569=251%:%
%:%570=251%:%
%:%571=252%:%
%:%572=252%:%
%:%573=253%:%
%:%574=253%:%
%:%575=254%:%
%:%576=254%:%
%:%577=255%:%
%:%578=255%:%
%:%579=256%:%
%:%580=256%:%
%:%581=257%:%
%:%582=257%:%
%:%583=258%:%
%:%584=258%:%
%:%585=259%:%
%:%586=259%:%
%:%587=260%:%
%:%588=260%:%
%:%589=261%:%
%:%590=261%:%
%:%591=262%:%
%:%592=262%:%
%:%593=263%:%
%:%594=263%:%
%:%595=264%:%
%:%596=264%:%
%:%597=265%:%
%:%598=265%:%
%:%599=266%:%
%:%600=266%:%
%:%601=267%:%
%:%602=267%:%
%:%603=268%:%
%:%604=268%:%
%:%605=269%:%
%:%606=269%:%
%:%607=270%:%
%:%608=270%:%
%:%609=271%:%
%:%610=271%:%
%:%611=272%:%
%:%612=272%:%
%:%613=273%:%
%:%614=273%:%
%:%615=274%:%
%:%616=274%:%
%:%621=274%:%
%:%624=275%:%
%:%625=276%:%
%:%626=276%:%
%:%633=277%:%
%:%634=277%:%
%:%635=278%:%
%:%636=278%:%
%:%637=279%:%
%:%638=280%:%
%:%639=280%:%
%:%640=281%:%
%:%641=282%:%
%:%642=283%:%
%:%643=283%:%
%:%644=284%:%
%:%645=284%:%
%:%646=285%:%
%:%647=285%:%
%:%648=286%:%
%:%649=286%:%
%:%650=287%:%
%:%651=287%:%
%:%652=288%:%
%:%653=288%:%
%:%654=289%:%
%:%655=289%:%
%:%656=290%:%
%:%657=290%:%
%:%658=291%:%
%:%659=291%:%
%:%660=292%:%
%:%661=292%:%
%:%662=293%:%
%:%663=293%:%
%:%664=294%:%
%:%665=294%:%
%:%666=295%:%
%:%667=295%:%
%:%668=296%:%
%:%669=296%:%
%:%670=297%:%
%:%671=298%:%
%:%672=298%:%
%:%673=299%:%
%:%674=299%:%
%:%675=300%:%
%:%676=300%:%
%:%677=301%:%
%:%678=301%:%
%:%679=302%:%
%:%680=302%:%
%:%681=303%:%
%:%682=303%:%
%:%683=304%:%
%:%684=304%:%
%:%685=305%:%
%:%686=305%:%
%:%687=306%:%
%:%688=306%:%
%:%689=307%:%
%:%690=307%:%
%:%691=308%:%
%:%692=308%:%
%:%693=309%:%
%:%694=309%:%
%:%695=310%:%
%:%696=310%:%
%:%697=311%:%
%:%698=311%:%
%:%699=312%:%
%:%700=312%:%
%:%701=313%:%
%:%702=313%:%
%:%703=314%:%
%:%704=315%:%
%:%705=315%:%
%:%706=315%:%
%:%707=315%:%
%:%708=315%:%
%:%709=316%:%
%:%710=317%:%
%:%711=317%:%
%:%712=318%:%
%:%713=318%:%
%:%714=319%:%
%:%715=319%:%
%:%716=320%:%
%:%717=320%:%
%:%718=321%:%
%:%719=321%:%
%:%720=322%:%
%:%721=322%:%
%:%722=323%:%
%:%723=323%:%
%:%724=324%:%
%:%725=325%:%
%:%726=325%:%
%:%727=325%:%
%:%728=325%:%
%:%729=325%:%
%:%730=326%:%
%:%731=326%:%
%:%732=326%:%
%:%733=326%:%
%:%734=326%:%
%:%735=327%:%
%:%736=327%:%
%:%737=327%:%
%:%738=328%:%
%:%739=328%:%
%:%740=329%:%
%:%741=329%:%
%:%742=330%:%
%:%748=330%:%
%:%751=331%:%
%:%752=332%:%
%:%753=332%:%
%:%754=333%:%
%:%755=334%:%
%:%756=335%:%
%:%759=336%:%
%:%760=337%:%
%:%764=337%:%
%:%765=337%:%
%:%766=338%:%
%:%767=338%:%
%:%768=339%:%
%:%769=339%:%
%:%770=340%:%
%:%771=340%:%
%:%772=341%:%
%:%773=341%:%
%:%778=341%:%
%:%781=342%:%
%:%782=343%:%
%:%783=343%:%
%:%784=344%:%
%:%785=345%:%
%:%786=346%:%
%:%793=347%:%
%:%794=347%:%
%:%795=348%:%
%:%796=349%:%
%:%797=349%:%
%:%798=350%:%
%:%799=350%:%
%:%800=351%:%
%:%801=351%:%
%:%802=352%:%
%:%803=352%:%
%:%804=353%:%
%:%805=354%:%
%:%806=354%:%
%:%807=354%:%
%:%808=354%:%
%:%809=355%:%
%:%810=355%:%
%:%811=355%:%
%:%812=356%:%
%:%813=356%:%
%:%814=357%:%
%:%815=357%:%
%:%816=358%:%
%:%817=358%:%
%:%818=358%:%
%:%819=358%:%
%:%820=359%:%
%:%821=360%:%
%:%822=360%:%
%:%823=361%:%
%:%824=361%:%
%:%825=362%:%
%:%826=362%:%
%:%827=363%:%
%:%828=363%:%
%:%829=364%:%
%:%830=364%:%
%:%831=365%:%
%:%832=365%:%
%:%833=366%:%
%:%834=366%:%
%:%835=367%:%
%:%836=367%:%
%:%837=368%:%
%:%838=368%:%
%:%839=369%:%
%:%840=369%:%
%:%841=370%:%
%:%842=370%:%
%:%843=370%:%
%:%844=371%:%
%:%845=371%:%
%:%846=372%:%
%:%847=372%:%
%:%848=373%:%
%:%849=373%:%
%:%850=373%:%
%:%851=374%:%
%:%852=374%:%
%:%853=375%:%
%:%854=375%:%
%:%855=376%:%
%:%856=376%:%
%:%857=377%:%
%:%858=377%:%
%:%859=378%:%
%:%860=378%:%
%:%861=378%:%
%:%862=379%:%
%:%863=379%:%
%:%864=380%:%
%:%865=380%:%
%:%866=381%:%
%:%867=381%:%
%:%868=382%:%
%:%869=382%:%
%:%870=382%:%
%:%871=383%:%
%:%872=383%:%
%:%873=384%:%
%:%874=384%:%
%:%875=384%:%
%:%876=384%:%
%:%877=385%:%
%:%878=385%:%
%:%879=385%:%
%:%880=385%:%
%:%881=385%:%
%:%882=386%:%
%:%883=386%:%
%:%884=386%:%
%:%885=386%:%
%:%886=386%:%
%:%887=387%:%
%:%888=387%:%
%:%889=388%:%
%:%890=388%:%
%:%891=388%:%
%:%892=389%:%
%:%893=389%:%
%:%894=390%:%
%:%895=390%:%
%:%896=391%:%
%:%897=391%:%
%:%898=392%:%
%:%899=392%:%
%:%900=393%:%
%:%901=393%:%
%:%902=394%:%
%:%903=394%:%
%:%904=395%:%
%:%905=395%:%
%:%906=396%:%
%:%907=396%:%
%:%908=397%:%
%:%909=397%:%
%:%910=398%:%
%:%911=398%:%
%:%912=399%:%
%:%913=399%:%
%:%914=399%:%
%:%915=400%:%
%:%916=400%:%
%:%917=401%:%
%:%918=401%:%
%:%919=402%:%
%:%920=402%:%
%:%921=403%:%
%:%922=403%:%
%:%923=403%:%
%:%924=403%:%
%:%925=404%:%
%:%926=404%:%
%:%927=405%:%
%:%928=406%:%
%:%929=406%:%
%:%930=407%:%
%:%931=407%:%
%:%932=408%:%
%:%933=408%:%
%:%934=409%:%
%:%935=409%:%
%:%936=410%:%
%:%937=410%:%
%:%938=411%:%
%:%939=412%:%
%:%940=412%:%
%:%941=413%:%
%:%942=413%:%
%:%943=414%:%
%:%944=414%:%
%:%945=415%:%
%:%946=415%:%
%:%947=416%:%
%:%948=416%:%
%:%949=417%:%
%:%950=417%:%
%:%951=418%:%
%:%952=418%:%
%:%953=419%:%
%:%954=419%:%
%:%955=420%:%
%:%956=420%:%
%:%957=421%:%
%:%958=421%:%
%:%959=422%:%
%:%960=422%:%
%:%961=423%:%
%:%962=423%:%
%:%963=424%:%
%:%964=424%:%
%:%965=424%:%
%:%966=424%:%
%:%967=425%:%
%:%968=425%:%
%:%969=425%:%
%:%970=426%:%
%:%971=426%:%
%:%972=427%:%
%:%973=427%:%
%:%974=428%:%
%:%975=428%:%
%:%976=428%:%
%:%977=429%:%
%:%978=429%:%
%:%979=430%:%
%:%980=430%:%
%:%981=431%:%
%:%982=431%:%
%:%983=432%:%
%:%984=432%:%
%:%985=432%:%
%:%986=432%:%
%:%987=433%:%
%:%988=433%:%
%:%989=433%:%
%:%990=433%:%
%:%991=433%:%
%:%992=434%:%
%:%993=434%:%
%:%994=434%:%
%:%995=434%:%
%:%996=434%:%
%:%997=435%:%
%:%998=435%:%
%:%999=436%:%
%:%1000=437%:%
%:%1001=437%:%
%:%1002=438%:%
%:%1003=438%:%
%:%1004=439%:%
%:%1005=439%:%
%:%1006=440%:%
%:%1007=440%:%
%:%1008=441%:%
%:%1009=441%:%
%:%1010=441%:%
%:%1011=442%:%
%:%1012=442%:%
%:%1013=443%:%
%:%1014=443%:%
%:%1015=444%:%
%:%1016=444%:%
%:%1017=445%:%
%:%1018=445%:%
%:%1019=445%:%
%:%1020=446%:%
%:%1021=446%:%
%:%1022=447%:%
%:%1023=447%:%
%:%1024=447%:%
%:%1025=448%:%
%:%1026=448%:%
%:%1027=449%:%
%:%1028=449%:%
%:%1029=450%:%
%:%1030=450%:%
%:%1031=451%:%
%:%1032=451%:%
%:%1033=452%:%
%:%1034=452%:%
%:%1035=453%:%
%:%1036=453%:%
%:%1037=454%:%
%:%1038=454%:%
%:%1039=455%:%
%:%1040=455%:%
%:%1041=456%:%
%:%1042=456%:%
%:%1043=457%:%
%:%1044=457%:%
%:%1045=458%:%
%:%1046=458%:%
%:%1047=458%:%
%:%1048=459%:%
%:%1049=459%:%
%:%1050=460%:%
%:%1051=460%:%
%:%1052=461%:%
%:%1053=461%:%
%:%1054=462%:%
%:%1055=462%:%
%:%1056=463%:%
%:%1057=463%:%
%:%1058=464%:%
%:%1059=464%:%
%:%1060=465%:%
%:%1061=465%:%
%:%1062=466%:%
%:%1063=466%:%
%:%1064=467%:%
%:%1065=467%:%
%:%1066=468%:%
%:%1067=468%:%
%:%1068=469%:%
%:%1069=469%:%
%:%1070=470%:%
%:%1071=470%:%
%:%1072=471%:%
%:%1073=471%:%
%:%1074=472%:%
%:%1075=472%:%
%:%1076=473%:%
%:%1077=473%:%
%:%1078=474%:%
%:%1079=474%:%
%:%1080=475%:%
%:%1081=475%:%
%:%1082=476%:%
%:%1083=476%:%
%:%1084=477%:%
%:%1085=477%:%
%:%1086=478%:%
%:%1087=478%:%
%:%1088=479%:%
%:%1089=479%:%
%:%1090=480%:%
%:%1091=480%:%
%:%1092=481%:%
%:%1093=481%:%
%:%1094=482%:%
%:%1095=482%:%
%:%1096=483%:%
%:%1097=483%:%
%:%1098=483%:%
%:%1099=484%:%
%:%1100=484%:%
%:%1101=485%:%
%:%1102=485%:%
%:%1103=486%:%
%:%1104=486%:%
%:%1105=487%:%
%:%1106=487%:%
%:%1107=488%:%
%:%1108=488%:%
%:%1109=489%:%
%:%1110=489%:%
%:%1111=490%:%
%:%1112=490%:%
%:%1113=491%:%
%:%1114=491%:%
%:%1115=492%:%
%:%1116=492%:%
%:%1117=492%:%
%:%1118=492%:%
%:%1119=493%:%
%:%1125=493%:%
%:%1128=494%:%
%:%1129=495%:%
%:%1130=495%:%
%:%1131=496%:%
%:%1132=497%:%
%:%1133=498%:%
%:%1140=499%:%
%:%1141=499%:%
%:%1142=500%:%
%:%1143=500%:%
%:%1144=501%:%
%:%1145=501%:%
%:%1146=502%:%
%:%1147=502%:%
%:%1148=502%:%
%:%1149=503%:%
%:%1150=503%:%
%:%1151=503%:%
%:%1152=504%:%
%:%1153=504%:%
%:%1154=505%:%
%:%1155=505%:%
%:%1156=506%:%
%:%1157=506%:%
%:%1158=507%:%
%:%1159=507%:%
%:%1160=508%:%
%:%1161=508%:%
%:%1162=509%:%
%:%1163=509%:%
%:%1164=510%:%
%:%1165=510%:%
%:%1166=510%:%
%:%1167=510%:%
%:%1168=510%:%
%:%1169=511%:%
%:%1170=511%:%
%:%1171=511%:%
%:%1172=512%:%
%:%1173=512%:%
%:%1174=513%:%
%:%1175=513%:%
%:%1176=514%:%
%:%1177=514%:%
%:%1178=515%:%
%:%1179=515%:%
%:%1180=516%:%
%:%1181=517%:%
%:%1182=517%:%
%:%1183=518%:%
%:%1184=518%:%
%:%1185=519%:%
%:%1186=519%:%
%:%1187=520%:%
%:%1188=520%:%
%:%1189=521%:%
%:%1190=521%:%
%:%1191=522%:%
%:%1192=522%:%
%:%1193=523%:%
%:%1194=523%:%
%:%1195=524%:%
%:%1196=524%:%
%:%1197=525%:%
%:%1198=525%:%
%:%1199=526%:%
%:%1200=526%:%
%:%1201=527%:%
%:%1202=527%:%
%:%1203=528%:%
%:%1204=528%:%
%:%1205=529%:%
%:%1206=529%:%
%:%1207=530%:%
%:%1208=530%:%
%:%1209=531%:%
%:%1215=531%:%
%:%1218=532%:%
%:%1219=533%:%
%:%1220=533%:%
%:%1221=534%:%
%:%1222=535%:%
%:%1223=536%:%
%:%1226=537%:%
%:%1230=537%:%
%:%1231=537%:%
%:%1232=538%:%
%:%1233=538%:%
%:%1234=539%:%
%:%1235=539%:%
%:%1236=540%:%
%:%1237=540%:%
%:%1238=541%:%
%:%1239=541%:%
%:%1240=542%:%
%:%1241=542%:%
%:%1242=543%:%
%:%1243=543%:%
%:%1244=544%:%
%:%1245=544%:%
%:%1246=545%:%
%:%1247=545%:%
%:%1248=546%:%
%:%1249=546%:%
%:%1250=547%:%
%:%1251=547%:%
%:%1252=548%:%
%:%1253=548%:%
%:%1254=549%:%
%:%1255=549%:%
%:%1260=549%:%
%:%1263=550%:%
%:%1264=551%:%
%:%1265=551%:%
%:%1266=552%:%
%:%1269=553%:%
%:%1270=554%:%
%:%1274=554%:%
%:%1275=554%:%
%:%1276=555%:%
%:%1277=555%:%
%:%1278=556%:%
%:%1279=556%:%
%:%1280=557%:%
%:%1281=557%:%
%:%1282=558%:%
%:%1283=558%:%
%:%1288=558%:%
%:%1291=559%:%
%:%1292=560%:%
%:%1293=560%:%
%:%1294=561%:%
%:%1295=562%:%
%:%1296=563%:%
%:%1299=564%:%
%:%1303=564%:%
%:%1304=564%:%
%:%1305=565%:%
%:%1306=565%:%
%:%1307=566%:%
%:%1308=566%:%
%:%1313=566%:%
%:%1316=567%:%
%:%1317=568%:%
%:%1318=568%:%
%:%1319=569%:%
%:%1320=570%:%
%:%1321=571%:%
%:%1328=572%:%
%:%1329=572%:%
%:%1330=573%:%
%:%1331=573%:%
%:%1332=574%:%
%:%1333=574%:%
%:%1334=575%:%
%:%1335=576%:%
%:%1336=576%:%
%:%1337=576%:%
%:%1338=576%:%
%:%1339=577%:%
%:%1340=578%:%
%:%1341=578%:%
%:%1342=579%:%
%:%1343=579%:%
%:%1344=580%:%
%:%1345=580%:%
%:%1346=580%:%
%:%1347=581%:%
%:%1348=581%:%
%:%1349=582%:%
%:%1350=582%:%
%:%1351=583%:%
%:%1352=583%:%
%:%1353=583%:%
%:%1354=584%:%
%:%1355=584%:%
%:%1356=585%:%
%:%1357=585%:%
%:%1358=586%:%
%:%1359=586%:%
%:%1360=587%:%
%:%1361=587%:%
%:%1362=587%:%
%:%1363=588%:%
%:%1364=588%:%
%:%1365=589%:%
%:%1366=589%:%
%:%1367=590%:%
%:%1368=590%:%
%:%1369=590%:%
%:%1370=591%:%
%:%1371=591%:%
%:%1372=592%:%
%:%1373=592%:%
%:%1374=593%:%
%:%1375=593%:%
%:%1376=593%:%
%:%1377=594%:%
%:%1378=594%:%
%:%1379=595%:%
%:%1380=595%:%
%:%1381=596%:%
%:%1382=596%:%
%:%1383=597%:%
%:%1384=597%:%
%:%1385=598%:%
%:%1386=598%:%
%:%1387=599%:%
%:%1388=599%:%
%:%1389=600%:%
%:%1390=601%:%
%:%1391=601%:%
%:%1392=601%:%
%:%1393=601%:%
%:%1394=602%:%
%:%1395=603%:%
%:%1396=603%:%
%:%1397=604%:%
%:%1398=604%:%
%:%1399=605%:%
%:%1400=605%:%
%:%1401=605%:%
%:%1402=605%:%
%:%1403=605%:%
%:%1404=606%:%
%:%1405=607%:%
%:%1406=607%:%
%:%1407=607%:%
%:%1408=607%:%
%:%1409=608%:%
%:%1410=608%:%
%:%1411=608%:%
%:%1412=608%:%
%:%1413=608%:%
%:%1414=609%:%
%:%1415=610%:%
%:%1416=610%:%
%:%1417=610%:%
%:%1418=610%:%
%:%1419=611%:%
%:%1420=611%:%
%:%1421=611%:%
%:%1422=612%:%
%:%1423=612%:%
%:%1424=613%:%
%:%1425=613%:%
%:%1426=614%:%
%:%1427=614%:%
%:%1428=615%:%
%:%1429=615%:%
%:%1430=616%:%
%:%1431=616%:%
%:%1432=617%:%
%:%1433=617%:%
%:%1434=617%:%
%:%1435=618%:%
%:%1436=618%:%
%:%1437=619%:%
%:%1438=619%:%
%:%1439=620%:%
%:%1440=620%:%
%:%1441=621%:%
%:%1442=621%:%
%:%1443=621%:%
%:%1444=621%:%
%:%1445=622%:%
%:%1446=623%:%
%:%1447=623%:%
%:%1448=624%:%
%:%1449=624%:%
%:%1450=625%:%
%:%1451=625%:%
%:%1452=626%:%
%:%1453=626%:%
%:%1454=627%:%
%:%1455=627%:%
%:%1456=628%:%
%:%1457=628%:%
%:%1458=629%:%
%:%1459=629%:%
%:%1460=630%:%
%:%1461=630%:%
%:%1462=631%:%
%:%1463=631%:%
%:%1464=631%:%
%:%1465=632%:%
%:%1466=632%:%
%:%1467=633%:%
%:%1468=633%:%
%:%1469=634%:%
%:%1470=634%:%
%:%1471=634%:%
%:%1472=635%:%
%:%1473=635%:%
%:%1474=636%:%
%:%1475=636%:%
%:%1476=637%:%
%:%1482=637%:%
%:%1485=638%:%
%:%1486=639%:%
%:%1487=639%:%
%:%1490=640%:%
%:%1494=640%:%
%:%1495=640%:%
%:%1496=641%:%
%:%1497=641%:%
%:%1498=642%:%
%:%1499=642%:%
%:%1504=642%:%
%:%1507=643%:%
%:%1508=644%:%
%:%1509=644%:%
%:%1512=645%:%
%:%1516=645%:%
%:%1517=645%:%
%:%1518=646%:%
%:%1519=646%:%
%:%1520=647%:%
%:%1521=647%:%
%:%1522=648%:%
%:%1523=648%:%
%:%1524=649%:%
%:%1525=649%:%
%:%1526=650%:%
%:%1527=650%:%
%:%1528=651%:%
%:%1529=651%:%
%:%1530=652%:%
%:%1531=652%:%
%:%1532=653%:%
%:%1533=653%:%
%:%1534=654%:%
%:%1535=654%:%
%:%1536=655%:%
%:%1537=655%:%
%:%1538=656%:%
%:%1539=656%:%
%:%1540=657%:%
%:%1541=657%:%
%:%1542=658%:%
%:%1543=658%:%
%:%1544=659%:%
%:%1545=659%:%
%:%1546=660%:%
%:%1547=660%:%
%:%1548=661%:%
%:%1549=661%:%
%:%1550=662%:%
%:%1551=662%:%
%:%1552=663%:%
%:%1553=663%:%
%:%1554=664%:%
%:%1555=664%:%
%:%1556=665%:%
%:%1557=665%:%
%:%1558=666%:%
%:%1559=666%:%
%:%1560=667%:%
%:%1561=667%:%
%:%1562=668%:%
%:%1563=668%:%
%:%1564=669%:%
%:%1565=669%:%
%:%1570=669%:%
%:%1573=670%:%
%:%1574=671%:%
%:%1575=671%:%
%:%1576=672%:%
%:%1577=673%:%
%:%1578=674%:%
%:%1585=675%:%
%:%1586=675%:%
%:%1587=676%:%
%:%1588=677%:%
%:%1589=677%:%
%:%1590=677%:%
%:%1591=677%:%
%:%1592=678%:%
%:%1593=679%:%
%:%1594=679%:%
%:%1595=680%:%
%:%1596=681%:%
%:%1597=681%:%
%:%1598=682%:%
%:%1599=682%:%
%:%1600=683%:%
%:%1601=683%:%
%:%1602=684%:%
%:%1603=684%:%
%:%1604=684%:%
%:%1605=685%:%
%:%1606=685%:%
%:%1607=686%:%
%:%1608=686%:%
%:%1609=687%:%
%:%1610=687%:%
%:%1611=688%:%
%:%1612=689%:%
%:%1613=689%:%
%:%1614=690%:%
%:%1615=690%:%
%:%1616=691%:%
%:%1617=691%:%
%:%1618=692%:%
%:%1619=692%:%
%:%1620=692%:%
%:%1621=693%:%
%:%1622=693%:%
%:%1623=694%:%
%:%1624=694%:%
%:%1625=695%:%
%:%1626=696%:%
%:%1627=696%:%
%:%1628=697%:%
%:%1629=697%:%
%:%1630=698%:%
%:%1631=698%:%
%:%1632=699%:%
%:%1633=699%:%
%:%1634=700%:%
%:%1635=701%:%
%:%1636=701%:%
%:%1637=702%:%
%:%1638=702%:%
%:%1639=703%:%
%:%1640=703%:%
%:%1641=704%:%
%:%1642=704%:%
%:%1643=705%:%
%:%1644=705%:%
%:%1645=706%:%
%:%1646=706%:%
%:%1647=707%:%
%:%1648=708%:%
%:%1649=708%:%
%:%1650=709%:%
%:%1651=709%:%
%:%1652=710%:%
%:%1653=710%:%
%:%1654=711%:%
%:%1655=711%:%
%:%1656=712%:%
%:%1657=712%:%
%:%1658=713%:%
%:%1659=713%:%
%:%1660=714%:%
%:%1661=714%:%
%:%1662=715%:%
%:%1663=716%:%
%:%1664=716%:%
%:%1665=717%:%
%:%1666=717%:%
%:%1667=718%:%
%:%1668=718%:%
%:%1669=719%:%
%:%1670=719%:%
%:%1671=720%:%
%:%1672=720%:%
%:%1673=720%:%
%:%1674=720%:%
%:%1675=720%:%
%:%1676=721%:%
%:%1677=721%:%
%:%1678=721%:%
%:%1679=721%:%
%:%1680=722%:%
%:%1681=722%:%
%:%1682=722%:%
%:%1683=722%:%
%:%1684=723%:%
%:%1685=723%:%
%:%1686=723%:%
%:%1687=723%:%
%:%1688=723%:%
%:%1689=724%:%
%:%1690=724%:%
%:%1691=724%:%
%:%1692=724%:%
%:%1693=725%:%
%:%1694=725%:%
%:%1695=725%:%
%:%1696=725%:%
%:%1697=725%:%
%:%1698=726%:%
%:%1699=726%:%
%:%1700=727%:%
%:%1701=728%:%
%:%1702=728%:%
%:%1703=729%:%
%:%1704=729%:%
%:%1705=730%:%
%:%1706=730%:%
%:%1707=731%:%
%:%1708=731%:%
%:%1709=732%:%
%:%1710=733%:%
%:%1711=733%:%
%:%1712=734%:%
%:%1713=734%:%
%:%1714=735%:%
%:%1715=735%:%
%:%1716=736%:%
%:%1717=736%:%
%:%1718=737%:%
%:%1719=737%:%
%:%1720=738%:%
%:%1721=738%:%
%:%1722=739%:%
%:%1723=740%:%
%:%1724=740%:%
%:%1725=741%:%
%:%1726=741%:%
%:%1727=742%:%
%:%1728=742%:%
%:%1729=743%:%
%:%1730=744%:%
%:%1731=744%:%
%:%1732=745%:%
%:%1733=745%:%
%:%1734=746%:%
%:%1735=746%:%
%:%1736=747%:%
%:%1737=747%:%
%:%1738=748%:%
%:%1739=748%:%
%:%1740=749%:%
%:%1741=749%:%
%:%1742=750%:%
%:%1743=750%:%
%:%1744=750%:%
%:%1745=751%:%
%:%1746=751%:%
%:%1747=752%:%
%:%1748=752%:%
%:%1749=753%:%
%:%1750=753%:%
%:%1751=754%:%
%:%1752=755%:%
%:%1753=755%:%
%:%1754=755%:%
%:%1755=756%:%
%:%1756=756%:%
%:%1757=757%:%
%:%1758=757%:%
%:%1759=758%:%
%:%1760=758%:%
%:%1761=759%:%
%:%1762=759%:%
%:%1763=760%:%
%:%1764=760%:%
%:%1765=761%:%
%:%1766=761%:%
%:%1767=762%:%
%:%1773=762%:%
%:%1776=763%:%
%:%1777=764%:%
%:%1778=764%:%
%:%1779=765%:%
%:%1780=766%:%
%:%1781=767%:%
%:%1788=768%:%
%:%1789=768%:%
%:%1790=769%:%
%:%1791=769%:%
%:%1792=770%:%
%:%1793=771%:%
%:%1794=772%:%
%:%1795=772%:%
%:%1796=773%:%
%:%1797=773%:%
%:%1798=774%:%
%:%1799=774%:%
%:%1800=775%:%
%:%1801=775%:%
%:%1802=776%:%
%:%1803=776%:%
%:%1804=777%:%
%:%1805=777%:%
%:%1806=778%:%
%:%1807=779%:%
%:%1808=779%:%
%:%1809=779%:%
%:%1810=779%:%
%:%1811=779%:%
%:%1812=780%:%
%:%1813=780%:%
%:%1814=780%:%
%:%1815=780%:%
%:%1816=781%:%
%:%1817=782%:%
%:%1818=782%:%
%:%1819=783%:%
%:%1820=783%:%
%:%1821=784%:%
%:%1822=784%:%
%:%1823=785%:%
%:%1824=785%:%
%:%1825=786%:%
%:%1826=786%:%
%:%1827=787%:%
%:%1828=787%:%
%:%1829=788%:%
%:%1830=788%:%
%:%1831=789%:%
%:%1837=789%:%
%:%1840=790%:%
%:%1841=791%:%
%:%1842=791%:%
%:%1843=792%:%
%:%1844=793%:%
%:%1845=794%:%
%:%1852=795%:%
%:%1853=795%:%
%:%1854=796%:%
%:%1855=796%:%
%:%1856=797%:%
%:%1857=797%:%
%:%1858=798%:%
%:%1859=798%:%
%:%1860=799%:%
%:%1861=799%:%
%:%1862=800%:%
%:%1863=800%:%
%:%1864=801%:%
%:%1865=801%:%
%:%1866=802%:%
%:%1867=802%:%
%:%1868=803%:%
%:%1869=803%:%
%:%1870=803%:%
%:%1871=804%:%
%:%1872=804%:%
%:%1873=805%:%
%:%1874=805%:%
%:%1875=806%:%
%:%1876=807%:%
%:%1877=807%:%
%:%1878=808%:%
%:%1879=808%:%
%:%1880=809%:%
%:%1881=809%:%
%:%1882=810%:%
%:%1883=810%:%
%:%1884=811%:%
%:%1885=811%:%
%:%1886=811%:%
%:%1887=811%:%
%:%1888=811%:%
%:%1889=812%:%
%:%1890=812%:%
%:%1891=812%:%
%:%1892=813%:%
%:%1893=813%:%
%:%1894=814%:%
%:%1895=814%:%
%:%1896=815%:%
%:%1897=815%:%
%:%1898=816%:%
%:%1899=816%:%
%:%1900=817%:%
%:%1901=817%:%
%:%1902=818%:%
%:%1903=818%:%
%:%1904=819%:%
%:%1905=819%:%
%:%1906=819%:%
%:%1907=819%:%
%:%1908=819%:%
%:%1909=820%:%
%:%1910=820%:%
%:%1911=820%:%
%:%1912=821%:%
%:%1913=821%:%
%:%1914=822%:%
%:%1915=822%:%
%:%1916=823%:%
%:%1917=823%:%
%:%1918=824%:%
%:%1919=824%:%
%:%1920=825%:%
%:%1921=825%:%
%:%1922=826%:%
%:%1923=826%:%
%:%1924=826%:%
%:%1925=827%:%
%:%1926=827%:%
%:%1927=828%:%
%:%1928=828%:%
%:%1929=829%:%
%:%1930=829%:%
%:%1931=830%:%
%:%1932=830%:%
%:%1933=831%:%
%:%1934=831%:%
%:%1935=831%:%
%:%1936=832%:%
%:%1937=832%:%
%:%1938=833%:%
%:%1939=833%:%
%:%1940=833%:%
%:%1941=834%:%
%:%1942=834%:%
%:%1943=835%:%
%:%1944=835%:%
%:%1945=836%:%
%:%1946=836%:%
%:%1947=837%:%
%:%1948=837%:%
%:%1949=838%:%
%:%1950=838%:%
%:%1951=839%:%
%:%1952=839%:%
%:%1953=840%:%
%:%1954=840%:%
%:%1955=841%:%
%:%1961=841%:%
%:%1964=842%:%
%:%1965=843%:%
%:%1966=843%:%
%:%1967=844%:%
%:%1968=845%:%
%:%1969=846%:%
%:%1972=847%:%
%:%1973=848%:%
%:%1977=848%:%
%:%1978=848%:%
%:%1979=849%:%
%:%1980=849%:%
%:%1981=850%:%
%:%1982=850%:%
%:%1983=851%:%
%:%1984=851%:%
%:%1985=852%:%
%:%1986=852%:%
%:%1987=853%:%
%:%1988=853%:%
%:%1989=854%:%
%:%1990=854%:%
%:%1991=855%:%
%:%1992=855%:%
%:%1993=856%:%
%:%1994=856%:%
%:%1995=857%:%
%:%1996=857%:%
%:%1997=858%:%
%:%1998=858%:%
%:%1999=859%:%
%:%2000=859%:%
%:%2001=860%:%
%:%2002=860%:%
%:%2003=861%:%
%:%2004=861%:%
%:%2005=862%:%
%:%2006=862%:%
%:%2007=863%:%
%:%2008=863%:%
%:%2013=863%:%
%:%2016=864%:%
%:%2017=865%:%
%:%2018=865%:%
%:%2019=866%:%
%:%2020=867%:%
%:%2021=868%:%
%:%2028=869%:%
%:%2029=869%:%
%:%2030=870%:%
%:%2031=870%:%
%:%2032=870%:%
%:%2033=870%:%
%:%2034=871%:%
%:%2035=871%:%
%:%2036=871%:%
%:%2037=871%:%
%:%2038=871%:%
%:%2039=872%:%
%:%2040=873%:%
%:%2041=873%:%
%:%2042=874%:%
%:%2043=874%:%
%:%2044=875%:%
%:%2045=875%:%
%:%2046=876%:%
%:%2047=876%:%
%:%2048=877%:%
%:%2049=878%:%
%:%2050=878%:%
%:%2051=879%:%
%:%2052=879%:%
%:%2053=880%:%
%:%2054=880%:%
%:%2055=881%:%
%:%2056=881%:%
%:%2057=882%:%
%:%2058=883%:%
%:%2059=883%:%
%:%2060=883%:%
%:%2061=883%:%
%:%2062=884%:%
%:%2068=884%:%
%:%2071=885%:%
%:%2072=886%:%
%:%2073=886%:%
%:%2074=887%:%
%:%2075=888%:%
%:%2076=889%:%
%:%2079=890%:%
%:%2083=890%:%
%:%2084=890%:%
%:%2085=891%:%
%:%2086=891%:%
%:%2087=892%:%
%:%2088=892%:%
%:%2089=893%:%
%:%2090=893%:%
%:%2091=894%:%
%:%2092=894%:%
%:%2093=895%:%
%:%2094=895%:%
%:%2095=896%:%
%:%2096=896%:%
%:%2097=897%:%
%:%2098=897%:%
%:%2099=898%:%
%:%2100=898%:%
%:%2101=899%:%
%:%2102=899%:%
%:%2103=900%:%
%:%2104=900%:%
%:%2105=901%:%
%:%2106=901%:%
%:%2107=902%:%
%:%2108=902%:%
%:%2109=903%:%
%:%2110=903%:%
%:%2111=904%:%
%:%2112=904%:%
%:%2113=905%:%
%:%2114=905%:%
%:%2115=906%:%
%:%2116=906%:%
%:%2121=906%:%
%:%2124=907%:%
%:%2125=908%:%
%:%2126=908%:%
%:%2127=909%:%
%:%2128=910%:%
%:%2129=911%:%
%:%2136=912%:%
%:%2137=912%:%
%:%2138=913%:%
%:%2139=913%:%
%:%2140=914%:%
%:%2141=914%:%
%:%2142=914%:%
%:%2143=915%:%
%:%2144=915%:%
%:%2145=916%:%
%:%2146=917%:%
%:%2147=917%:%
%:%2148=918%:%
%:%2149=918%:%
%:%2150=919%:%
%:%2151=919%:%
%:%2152=920%:%
%:%2153=920%:%
%:%2154=921%:%
%:%2155=921%:%
%:%2156=922%:%
%:%2157=922%:%
%:%2158=923%:%
%:%2159=923%:%
%:%2160=924%:%
%:%2161=924%:%
%:%2162=925%:%
%:%2163=926%:%
%:%2164=926%:%
%:%2165=927%:%
%:%2166=927%:%
%:%2167=928%:%
%:%2168=928%:%
%:%2169=929%:%
%:%2170=929%:%
%:%2171=930%:%
%:%2172=931%:%
%:%2173=931%:%
%:%2174=932%:%
%:%2175=932%:%
%:%2176=932%:%
%:%2177=933%:%
%:%2178=934%:%
%:%2179=934%:%
%:%2180=934%:%
%:%2181=935%:%
%:%2182=935%:%
%:%2183=936%:%
%:%2184=936%:%
%:%2185=937%:%
%:%2186=937%:%
%:%2187=937%:%
%:%2188=938%:%
%:%2189=938%:%
%:%2190=939%:%
%:%2191=939%:%
%:%2192=940%:%
%:%2193=940%:%
%:%2194=941%:%
%:%2195=941%:%
%:%2196=942%:%
%:%2197=942%:%
%:%2198=943%:%
%:%2199=943%:%
%:%2200=944%:%
%:%2201=944%:%
%:%2202=944%:%
%:%2203=945%:%
%:%2204=945%:%
%:%2205=946%:%
%:%2206=946%:%
%:%2207=947%:%
%:%2208=948%:%
%:%2209=948%:%
%:%2210=949%:%
%:%2211=949%:%
%:%2212=950%:%
%:%2213=950%:%
%:%2214=951%:%
%:%2215=951%:%
%:%2216=952%:%
%:%2217=953%:%
%:%2218=953%:%
%:%2219=954%:%
%:%2220=954%:%
%:%2221=954%:%
%:%2222=955%:%
%:%2223=955%:%
%:%2224=955%:%
%:%2225=956%:%
%:%2226=956%:%
%:%2227=956%:%
%:%2228=957%:%
%:%2229=957%:%
%:%2230=957%:%
%:%2231=958%:%
%:%2232=958%:%
%:%2233=959%:%
%:%2234=959%:%
%:%2235=960%:%
%:%2236=960%:%
%:%2237=960%:%
%:%2238=960%:%
%:%2239=960%:%
%:%2240=961%:%
%:%2241=962%:%
%:%2242=962%:%
%:%2243=963%:%
%:%2244=963%:%
%:%2245=964%:%
%:%2246=964%:%
%:%2247=965%:%
%:%2248=965%:%
%:%2249=965%:%
%:%2250=965%:%
%:%2251=965%:%
%:%2252=966%:%
%:%2258=966%:%
%:%2261=967%:%
%:%2262=968%:%
%:%2263=968%:%
%:%2264=969%:%
%:%2265=970%:%
%:%2266=971%:%
%:%2273=972%:%
%:%2274=972%:%
%:%2275=973%:%
%:%2276=973%:%
%:%2277=974%:%
%:%2278=974%:%
%:%2279=975%:%
%:%2280=975%:%
%:%2281=976%:%
%:%2282=976%:%
%:%2283=977%:%
%:%2284=977%:%
%:%2285=978%:%
%:%2286=978%:%
%:%2287=979%:%
%:%2288=979%:%
%:%2289=980%:%
%:%2290=981%:%
%:%2291=981%:%
%:%2292=981%:%
%:%2293=981%:%
%:%2294=982%:%
%:%2295=983%:%
%:%2296=983%:%
%:%2297=983%:%
%:%2298=984%:%
%:%2299=984%:%
%:%2300=985%:%
%:%2301=985%:%
%:%2302=986%:%
%:%2303=986%:%
%:%2304=987%:%
%:%2305=987%:%
%:%2306=988%:%
%:%2307=988%:%
%:%2308=989%:%
%:%2309=989%:%
%:%2310=990%:%
%:%2311=990%:%
%:%2312=991%:%
%:%2313=992%:%
%:%2314=992%:%
%:%2315=993%:%
%:%2316=993%:%
%:%2317=994%:%
%:%2318=994%:%
%:%2319=995%:%
%:%2320=995%:%
%:%2321=996%:%
%:%2322=996%:%
%:%2323=997%:%
%:%2324=998%:%
%:%2325=998%:%
%:%2326=999%:%
%:%2327=999%:%
%:%2328=1000%:%
%:%2329=1000%:%
%:%2330=1001%:%
%:%2331=1001%:%
%:%2332=1001%:%
%:%2333=1001%:%
%:%2334=1001%:%
%:%2335=1002%:%
%:%2336=1002%:%
%:%2337=1002%:%
%:%2338=1003%:%
%:%2339=1003%:%
%:%2340=1004%:%
%:%2341=1004%:%
%:%2342=1005%:%
%:%2343=1005%:%
%:%2344=1006%:%
%:%2345=1006%:%
%:%2346=1006%:%
%:%2347=1006%:%
%:%2348=1007%:%
%:%2349=1008%:%
%:%2350=1008%:%
%:%2351=1009%:%
%:%2352=1010%:%
%:%2353=1010%:%
%:%2354=1010%:%
%:%2355=1010%:%
%:%2356=1011%:%
%:%2357=1011%:%
%:%2358=1011%:%
%:%2359=1011%:%
%:%2360=1012%:%
%:%2361=1012%:%
%:%2362=1012%:%
%:%2363=1013%:%
%:%2364=1013%:%
%:%2365=1014%:%
%:%2366=1014%:%
%:%2367=1015%:%
%:%2368=1015%:%
%:%2369=1016%:%
%:%2370=1016%:%
%:%2371=1017%:%
%:%2372=1018%:%
%:%2373=1018%:%
%:%2374=1019%:%
%:%2375=1019%:%
%:%2376=1020%:%
%:%2377=1020%:%
%:%2378=1021%:%
%:%2379=1021%:%
%:%2380=1022%:%
%:%2381=1022%:%
%:%2382=1023%:%
%:%2383=1023%:%
%:%2384=1024%:%
%:%2385=1024%:%
%:%2386=1025%:%
%:%2387=1025%:%
%:%2388=1026%:%
%:%2389=1026%:%
%:%2390=1027%:%
%:%2391=1027%:%
%:%2392=1028%:%
%:%2393=1028%:%
%:%2394=1029%:%
%:%2395=1029%:%
%:%2396=1030%:%
%:%2397=1030%:%
%:%2398=1031%:%
%:%2399=1031%:%
%:%2400=1032%:%
%:%2401=1032%:%
%:%2402=1033%:%
%:%2403=1033%:%
%:%2404=1034%:%
%:%2405=1034%:%
%:%2406=1035%:%
%:%2407=1035%:%
%:%2408=1036%:%
%:%2409=1036%:%
%:%2410=1037%:%
%:%2411=1037%:%
%:%2412=1038%:%
%:%2413=1038%:%
%:%2414=1039%:%
%:%2415=1039%:%
%:%2416=1040%:%
%:%2417=1041%:%
%:%2418=1041%:%
%:%2419=1042%:%
%:%2420=1042%:%
%:%2421=1043%:%
%:%2422=1043%:%
%:%2423=1044%:%
%:%2424=1044%:%
%:%2425=1045%:%
%:%2426=1046%:%
%:%2427=1046%:%
%:%2428=1047%:%
%:%2429=1047%:%
%:%2430=1048%:%
%:%2431=1048%:%
%:%2432=1049%:%
%:%2433=1049%:%
%:%2434=1050%:%
%:%2435=1051%:%
%:%2436=1051%:%
%:%2437=1052%:%
%:%2438=1052%:%
%:%2439=1053%:%
%:%2440=1053%:%
%:%2441=1054%:%
%:%2442=1054%:%
%:%2443=1054%:%
%:%2444=1055%:%
%:%2445=1055%:%
%:%2446=1056%:%
%:%2447=1056%:%
%:%2448=1057%:%
%:%2449=1058%:%
%:%2450=1058%:%
%:%2451=1059%:%
%:%2452=1059%:%
%:%2453=1060%:%
%:%2454=1060%:%
%:%2455=1061%:%
%:%2456=1061%:%
%:%2457=1062%:%
%:%2458=1062%:%
%:%2459=1063%:%
%:%2460=1063%:%
%:%2461=1064%:%
%:%2462=1064%:%
%:%2463=1065%:%
%:%2464=1065%:%
%:%2465=1066%:%
%:%2466=1066%:%
%:%2467=1067%:%
%:%2468=1067%:%
%:%2469=1068%:%
%:%2470=1068%:%
%:%2471=1069%:%
%:%2472=1069%:%
%:%2473=1069%:%
%:%2474=1070%:%
%:%2475=1070%:%
%:%2476=1071%:%
%:%2477=1071%:%
%:%2478=1072%:%
%:%2479=1073%:%
%:%2480=1073%:%
%:%2481=1074%:%
%:%2482=1074%:%
%:%2483=1075%:%
%:%2484=1075%:%
%:%2485=1076%:%
%:%2486=1076%:%
%:%2487=1077%:%
%:%2488=1077%:%
%:%2489=1077%:%
%:%2490=1078%:%
%:%2491=1078%:%
%:%2492=1079%:%
%:%2493=1079%:%
%:%2494=1080%:%
%:%2495=1080%:%
%:%2496=1081%:%
%:%2497=1081%:%
%:%2498=1082%:%
%:%2499=1082%:%
%:%2500=1083%:%
%:%2501=1083%:%
%:%2502=1084%:%
%:%2503=1085%:%
%:%2504=1085%:%
%:%2505=1086%:%
%:%2506=1086%:%
%:%2507=1087%:%
%:%2508=1087%:%
%:%2509=1088%:%
%:%2510=1088%:%
%:%2511=1089%:%
%:%2512=1089%:%
%:%2513=1089%:%
%:%2514=1090%:%
%:%2515=1090%:%
%:%2516=1091%:%
%:%2517=1091%:%
%:%2518=1092%:%
%:%2519=1092%:%
%:%2520=1093%:%
%:%2521=1093%:%
%:%2522=1094%:%
%:%2523=1094%:%
%:%2524=1095%:%
%:%2525=1095%:%
%:%2526=1096%:%
%:%2527=1096%:%
%:%2528=1096%:%
%:%2529=1097%:%
%:%2530=1097%:%
%:%2531=1097%:%
%:%2532=1098%:%
%:%2533=1099%:%
%:%2534=1099%:%
%:%2535=1100%:%
%:%2536=1100%:%
%:%2537=1101%:%
%:%2538=1101%:%
%:%2539=1102%:%
%:%2540=1103%:%
%:%2541=1103%:%
%:%2542=1103%:%
%:%2543=1104%:%
%:%2544=1104%:%
%:%2545=1105%:%
%:%2546=1105%:%
%:%2547=1106%:%
%:%2553=1106%:%
%:%2556=1107%:%
%:%2557=1108%:%
%:%2558=1108%:%
%:%2559=1109%:%
%:%2560=1110%:%
%:%2561=1111%:%
%:%2564=1112%:%
%:%2565=1113%:%
%:%2569=1113%:%
%:%2570=1113%:%
%:%2571=1114%:%
%:%2572=1114%:%
%:%2573=1115%:%
%:%2574=1115%:%
%:%2575=1116%:%
%:%2576=1116%:%
%:%2577=1117%:%
%:%2578=1117%:%
%:%2579=1118%:%
%:%2580=1118%:%
%:%2585=1118%:%
%:%2588=1119%:%
%:%2589=1120%:%
%:%2590=1120%:%
%:%2591=1121%:%
%:%2592=1122%:%
%:%2593=1123%:%
%:%2600=1124%:%
%:%2601=1124%:%
%:%2602=1125%:%
%:%2603=1126%:%
%:%2604=1126%:%
%:%2605=1127%:%
%:%2606=1127%:%
%:%2607=1128%:%
%:%2608=1128%:%
%:%2609=1129%:%
%:%2610=1130%:%
%:%2611=1130%:%
%:%2612=1131%:%
%:%2613=1131%:%
%:%2614=1132%:%
%:%2615=1132%:%
%:%2616=1133%:%
%:%2617=1133%:%
%:%2618=1134%:%
%:%2619=1134%:%
%:%2620=1134%:%
%:%2621=1135%:%
%:%2622=1135%:%
%:%2623=1136%:%
%:%2624=1136%:%
%:%2625=1137%:%
%:%2626=1137%:%
%:%2627=1138%:%
%:%2628=1138%:%
%:%2629=1139%:%
%:%2630=1139%:%
%:%2631=1140%:%
%:%2632=1140%:%
%:%2633=1141%:%
%:%2634=1141%:%
%:%2635=1141%:%
%:%2636=1142%:%
%:%2637=1142%:%
%:%2638=1143%:%
%:%2639=1143%:%
%:%2640=1144%:%
%:%2641=1144%:%
%:%2642=1144%:%
%:%2643=1145%:%
%:%2644=1145%:%
%:%2645=1146%:%
%:%2646=1146%:%
%:%2647=1147%:%
%:%2648=1147%:%
%:%2649=1148%:%
%:%2650=1148%:%
%:%2651=1149%:%
%:%2652=1149%:%
%:%2653=1149%:%
%:%2654=1150%:%
%:%2655=1150%:%
%:%2656=1151%:%
%:%2657=1151%:%
%:%2658=1152%:%
%:%2659=1152%:%
%:%2660=1153%:%
%:%2661=1153%:%
%:%2662=1154%:%
%:%2663=1154%:%
%:%2664=1155%:%
%:%2665=1155%:%
%:%2666=1156%:%
%:%2667=1156%:%
%:%2668=1157%:%
%:%2674=1157%:%
%:%2677=1158%:%
%:%2678=1159%:%
%:%2679=1159%:%
%:%2686=1160%:%