### 2024-11-01 code/output/document/NotAC_Inj.tex

%
\begin{isabellebody}%
\setisabellecontext{NotAC{\isacharunderscore}{\kern0pt}Inj}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ NotAC{\isacharunderscore}{\kern0pt}Inj\ \isanewline
\ \ \isakeyword{imports}\ Utilities\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans\ \isakeyword{begin}\isanewline
\isanewline
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}x{\isasymin}A\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ f{\isacharcomma}{\kern0pt}\ r{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\ Implies{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}separation\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ f\ r\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ x{\isasymin}A\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}y{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ x{\isasymin}A\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}p\ {\isasymin}\ M{\isachardot}{\kern0pt}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ f{\isacharcomma}{\kern0pt}\ r{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}g\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}mx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}par\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharparenright}{\kern0pt}\ {\isasymand}\ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}par{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}par{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ r{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}obase{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\ Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}pred{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}separation\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ r\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}a{\isachardot}{\kern0pt}\ {\isasymexists}x{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}g{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}mx{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}par{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharparenright}{\kern0pt}\ {\isasymand}\ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}par{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}par{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}a{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}g\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}mx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}par\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharparenright}{\kern0pt}\ {\isasymand}\ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}par{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}par{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}a{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ obase{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ r{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pred{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Memrel{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ bijection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ injection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ surjection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}equals{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}x{\isasymin}A\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}g\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}my\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pxr\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}my{\isacharparenright}{\kern0pt}\ {\isasymand}\ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}pxr{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}pxr{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}my{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ r{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}equals{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}obase{\isacharunderscore}{\kern0pt}equals{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\ \isanewline
\ \ \ \ \ \ \ \ \ \ Implies\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ Neg{\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}pred{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}equals{\isacharunderscore}{\kern0pt}separation\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ r\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}separation\ {\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ x{\isasymin}A\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}y{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}g{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}my{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}pxr{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}my{\isacharparenright}{\kern0pt}\ {\isasymand}\ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}pxr{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}pxr{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}my{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x{\isasymin}A\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}g\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}my\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}pxr\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}my{\isacharparenright}{\kern0pt}\ {\isasymand}\ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}pxr{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}pxr{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}my{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}a{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ obase{\isacharunderscore}{\kern0pt}equals{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ r{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}equals{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ obase{\isacharunderscore}{\kern0pt}equals{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pred{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Memrel{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ bijection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ injection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ surjection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ omap{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}g\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}mx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}par\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}par{\isacharparenright}{\kern0pt}\ {\isasymand}\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}par{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ z{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ r{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ omap{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}omap{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\ Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}pred{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ omap{\isacharunderscore}{\kern0pt}replacement\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ r\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymlambda}a\ z{\isachardot}{\kern0pt}\ {\isasymexists}x{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}g{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}mx{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}par{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}par{\isacharparenright}{\kern0pt}\ {\isasymand}\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}par{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}a\ z{\isachardot}{\kern0pt}\ {\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}g\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}mx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}par\ {\isasymin}\ M{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ ordinal{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ membership{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ pred{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}par{\isacharparenright}{\kern0pt}\ {\isasymand}\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}par{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}mx{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}a\ z{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ omap{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ z{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ r{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ omap{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ omap{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto\ assms\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ replacement{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ omap{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pred{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Memrel{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ bijection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ order{\isacharunderscore}{\kern0pt}isomorphism{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ bijection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ injection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ surjection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ mordertype\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}ordertype{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ordertype{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ mbasic\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ordertype{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ well{\isacharunderscore}{\kern0pt}ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}separation\ obase{\isacharunderscore}{\kern0pt}separation\ obase{\isacharunderscore}{\kern0pt}equals{\isacharunderscore}{\kern0pt}separation\ omap{\isacharunderscore}{\kern0pt}replacement\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wellorder{\isacharunderscore}{\kern0pt}induces{\isacharunderscore}{\kern0pt}injection\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ r\ A\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}nat\ {\isasymlesssim}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}wellordered{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}f\ {\isasymin}\ M{\isachardot}{\kern0pt}\ f\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}f{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}i{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymand}\ f\ {\isasymin}\ ord{\isacharunderscore}{\kern0pt}iso{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ r{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ Memrel{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ M{\isacharunderscore}{\kern0pt}ordertype{\isachardot}{\kern0pt}ordertype{\isacharunderscore}{\kern0pt}exists{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mordertype\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ f\ i\ \isakeyword{where}\ fiH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isasymin}\ ord{\isacharunderscore}{\kern0pt}iso{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ r{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ Memrel{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ fbij{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ord{\isacharunderscore}{\kern0pt}iso{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}A\ {\isasymapprox}\ i{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ eqpoll{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}A\ {\isasymlesssim}\ i{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ eqpoll{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}lepoll\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nat\ {\isasymlesssim}\ i{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Y{\isacharequal}{\kern0pt}{\isachardoublequoteopen}A{\isachardoublequoteclose}\ \isakeyword{in}\ lepoll{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbar}{\kern0pt}nat{\isacharbar}{\kern0pt}\ {\isasymle}\ i{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ lepoll{\isacharunderscore}{\kern0pt}cardinal{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fiH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nat\ {\isasymle}\ i{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Card{\isacharunderscore}{\kern0pt}nat\ Card{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nat\ {\isasymsubseteq}\ i{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltD{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}nat\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ltI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ converse{\isacharunderscore}{\kern0pt}closed\ fiH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}bij\ fbij\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nat{\isacharparenright}{\kern0pt}\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}i\ \isakeyword{in}\ restrict{\isacharunderscore}{\kern0pt}inj{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fbij\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ {\isacartoucheopen}nat\ {\isasymsubseteq}\ i{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}force{\isacharcomma}{\kern0pt}\ rule\ restrict{\isacharunderscore}{\kern0pt}closed{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isacartoucheclose}\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \ \ \isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/NotAC_Inj.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%18=3%:%
%:%21=4%:%
%:%22=5%:%
%:%23=5%:%
%:%24=6%:%
%:%25=7%:%
%:%26=8%:%
%:%27=8%:%
%:%28=9%:%
%:%29=10%:%
%:%30=11%:%
%:%31=12%:%
%:%34=13%:%
%:%38=13%:%
%:%39=13%:%
%:%44=13%:%
%:%47=14%:%
%:%48=15%:%
%:%49=15%:%
%:%50=16%:%
%:%51=17%:%
%:%52=17%:%
%:%53=18%:%
%:%54=19%:%
%:%55=20%:%
%:%58=21%:%
%:%59=22%:%
%:%63=22%:%
%:%64=22%:%
%:%65=23%:%
%:%66=23%:%
%:%67=24%:%
%:%68=24%:%
%:%69=25%:%
%:%70=25%:%
%:%71=26%:%
%:%72=26%:%
%:%73=27%:%
%:%74=27%:%
%:%75=28%:%
%:%76=28%:%
%:%77=29%:%
%:%78=29%:%
%:%79=30%:%
%:%80=30%:%
%:%81=31%:%
%:%82=31%:%
%:%83=32%:%
%:%84=32%:%
%:%85=33%:%
%:%86=33%:%
%:%87=34%:%
%:%88=34%:%
%:%89=35%:%
%:%90=35%:%
%:%91=36%:%
%:%97=36%:%
%:%100=37%:%
%:%101=38%:%
%:%102=38%:%
%:%103=39%:%
%:%104=40%:%
%:%105=41%:%
%:%106=42%:%
%:%108=44%:%
%:%111=45%:%
%:%115=45%:%
%:%116=45%:%
%:%121=45%:%
%:%124=46%:%
%:%125=47%:%
%:%126=47%:%
%:%130=51%:%
%:%131=52%:%
%:%132=53%:%
%:%133=53%:%
%:%134=54%:%
%:%135=55%:%
%:%136=56%:%
%:%138=58%:%
%:%141=59%:%
%:%142=60%:%
%:%146=60%:%
%:%147=60%:%
%:%149=62%:%
%:%150=63%:%
%:%151=63%:%
%:%152=64%:%
%:%153=64%:%
%:%154=65%:%
%:%155=65%:%
%:%156=66%:%
%:%157=66%:%
%:%158=67%:%
%:%159=67%:%
%:%160=68%:%
%:%161=68%:%
%:%162=69%:%
%:%163=69%:%
%:%164=70%:%
%:%165=70%:%
%:%166=71%:%
%:%167=71%:%
%:%168=72%:%
%:%169=72%:%
%:%170=73%:%
%:%171=73%:%
%:%172=74%:%
%:%173=74%:%
%:%174=75%:%
%:%175=75%:%
%:%176=76%:%
%:%177=76%:%
%:%178=77%:%
%:%179=77%:%
%:%180=78%:%
%:%181=78%:%
%:%182=79%:%
%:%188=79%:%
%:%191=80%:%
%:%192=81%:%
%:%193=82%:%
%:%194=82%:%
%:%195=83%:%
%:%196=84%:%
%:%197=85%:%
%:%198=86%:%
%:%201=89%:%
%:%204=90%:%
%:%208=90%:%
%:%209=90%:%
%:%214=90%:%
%:%217=91%:%
%:%218=92%:%
%:%219=92%:%
%:%225=98%:%
%:%226=99%:%
%:%227=100%:%
%:%228=100%:%
%:%229=101%:%
%:%230=102%:%
%:%231=103%:%
%:%234=106%:%
%:%237=107%:%
%:%238=108%:%
%:%242=108%:%
%:%243=108%:%
%:%246=111%:%
%:%247=112%:%
%:%248=112%:%
%:%249=113%:%
%:%250=113%:%
%:%251=114%:%
%:%252=114%:%
%:%253=115%:%
%:%254=115%:%
%:%255=116%:%
%:%256=116%:%
%:%257=117%:%
%:%258=117%:%
%:%259=118%:%
%:%260=118%:%
%:%261=119%:%
%:%262=119%:%
%:%263=120%:%
%:%264=120%:%
%:%265=121%:%
%:%266=121%:%
%:%267=122%:%
%:%268=122%:%
%:%269=123%:%
%:%270=123%:%
%:%271=124%:%
%:%272=124%:%
%:%273=125%:%
%:%274=125%:%
%:%275=126%:%
%:%276=126%:%
%:%277=127%:%
%:%278=127%:%
%:%279=128%:%
%:%285=128%:%
%:%288=129%:%
%:%289=130%:%
%:%290=130%:%
%:%291=131%:%
%:%292=132%:%
%:%293=133%:%
%:%294=134%:%
%:%296=136%:%
%:%299=137%:%
%:%303=137%:%
%:%304=137%:%
%:%309=137%:%
%:%312=138%:%
%:%313=139%:%
%:%314=139%:%
%:%320=145%:%
%:%321=146%:%
%:%322=147%:%
%:%323=147%:%
%:%324=148%:%
%:%325=149%:%
%:%326=150%:%
%:%329=153%:%
%:%332=154%:%
%:%333=155%:%
%:%337=155%:%
%:%338=155%:%
%:%340=157%:%
%:%341=158%:%
%:%342=158%:%
%:%343=159%:%
%:%344=159%:%
%:%345=160%:%
%:%346=160%:%
%:%347=161%:%
%:%348=161%:%
%:%349=162%:%
%:%350=162%:%
%:%351=163%:%
%:%352=163%:%
%:%353=164%:%
%:%354=164%:%
%:%355=165%:%
%:%356=165%:%
%:%357=166%:%
%:%358=166%:%
%:%359=167%:%
%:%360=167%:%
%:%361=168%:%
%:%362=168%:%
%:%363=169%:%
%:%364=169%:%
%:%365=170%:%
%:%366=170%:%
%:%367=171%:%
%:%368=171%:%
%:%369=172%:%
%:%370=172%:%
%:%371=173%:%
%:%372=173%:%
%:%373=174%:%
%:%379=174%:%
%:%382=175%:%
%:%383=176%:%
%:%384=176%:%
%:%387=177%:%
%:%391=177%:%
%:%392=177%:%
%:%393=178%:%
%:%394=178%:%
%:%395=179%:%
%:%396=179%:%
%:%397=180%:%
%:%398=180%:%
%:%399=181%:%
%:%400=181%:%
%:%401=182%:%
%:%402=182%:%
%:%407=182%:%
%:%410=183%:%
%:%411=184%:%
%:%412=184%:%
%:%413=185%:%
%:%414=186%:%
%:%415=187%:%
%:%422=188%:%
%:%423=188%:%
%:%424=189%:%
%:%425=189%:%
%:%426=190%:%
%:%427=190%:%
%:%428=191%:%
%:%429=191%:%
%:%430=192%:%
%:%431=192%:%
%:%432=193%:%
%:%433=193%:%
%:%434=193%:%
%:%435=194%:%
%:%436=194%:%
%:%437=195%:%
%:%438=195%:%
%:%439=195%:%
%:%440=195%:%
%:%441=195%:%
%:%442=196%:%
%:%443=196%:%
%:%444=196%:%
%:%445=196%:%
%:%446=196%:%
%:%447=197%:%
%:%448=197%:%
%:%449=197%:%
%:%450=197%:%
%:%451=197%:%
%:%452=198%:%
%:%453=198%:%
%:%454=198%:%
%:%455=199%:%
%:%456=199%:%
%:%457=200%:%
%:%458=200%:%
%:%459=201%:%
%:%460=201%:%
%:%461=202%:%
%:%462=202%:%
%:%463=202%:%
%:%464=203%:%
%:%465=203%:%
%:%466=204%:%
%:%467=204%:%
%:%468=205%:%
%:%469=205%:%
%:%470=206%:%
%:%471=206%:%
%:%472=206%:%
%:%473=207%:%
%:%474=207%:%
%:%475=208%:%
%:%476=208%:%
%:%477=209%:%
%:%478=209%:%
%:%479=209%:%
%:%480=210%:%
%:%481=210%:%
%:%482=211%:%
%:%483=211%:%
%:%484=212%:%
%:%485=212%:%
%:%486=213%:%
%:%487=213%:%
%:%488=214%:%
%:%489=214%:%
%:%490=215%:%
%:%491=216%:%
%:%492=216%:%
%:%493=216%:%
%:%494=216%:%
%:%495=217%:%
%:%496=217%:%
%:%497=217%:%
%:%498=217%:%
%:%499=218%:%
%:%500=219%:%
%:%501=219%:%
%:%502=219%:%
%:%503=220%:%
%:%504=220%:%
%:%505=221%:%
%:%506=221%:%
%:%507=222%:%
%:%508=222%:%
%:%509=223%:%
%:%510=223%:%
%:%511=223%:%
%:%512=224%:%
%:%513=224%:%
%:%514=225%:%
%:%515=225%:%
%:%516=226%:%
%:%517=226%:%
%:%518=227%:%
%:%519=227%:%
%:%520=228%:%
%:%521=228%:%
%:%522=229%:%
%:%528=229%:%
%:%531=230%:%
%:%532=231%:%
%:%533=231%:%
%:%540=232%:%