### 2024-11-01 code/output/document/NotAC_Proof.tex

%
\begin{isabellebody}%
\setisabellecontext{NotAC{\isacharunderscore}{\kern0pt}Proof}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ NotAC{\isacharunderscore}{\kern0pt}Proof\ \isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ NotAC{\isacharunderscore}{\kern0pt}Binmap\ \isanewline
\ \ \ \ NotAC{\isacharunderscore}{\kern0pt}Inj\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ \isakeyword{begin}\ \isanewline
\isanewline
\isacommand{interpretation}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\ {\isachardoublequoteopen}Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ no{\isacharunderscore}{\kern0pt}injection\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ F{\isacharprime}{\kern0pt}\ p{\isadigit{0}}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}F{\isacharprime}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isadigit{0}}\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}p{\isadigit{0}}{\isacharcomma}{\kern0pt}\ injection{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ False\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ {\isasymphi}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}{\isasymphi}{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}G\ a\ b\ c{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isacharcomma}{\kern0pt}\ c{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isacharcomma}{\kern0pt}\ c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ a{\isacharbackquote}{\kern0pt}b\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ G\ a\ b\ c\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isacharcomma}{\kern0pt}\ c{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isanewline
\ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic\ \ {\isachardoublequoteopen}Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter{\isachardoublequoteclose}\ G\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}axioms\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}axioms\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isacharcomma}{\kern0pt}\ c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ a{\isacharbackquote}{\kern0pt}b\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymexists}d\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ a{\isacharbackquote}{\kern0pt}b\ {\isacharequal}{\kern0pt}\ d\ {\isasymand}\ d\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ M{\isacharunderscore}{\kern0pt}basic{\isachardot}{\kern0pt}apply{\isacharunderscore}{\kern0pt}abs{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans{\isachardot}{\kern0pt}mbasic\ SymExt{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{6}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ G\ \isakeyword{where}\ GH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isadigit{0}}\ {\isasymin}\ G{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}existence\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic\ \ {\isachardoublequoteopen}Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter{\isachardoublequoteclose}\ G\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}axioms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}axioms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ GH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}subset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A\ B\ C{\isachardot}{\kern0pt}\ A\ {\isasymsubseteq}\ C\ {\isasymLongrightarrow}\ B\ {\isasymsubseteq}\ C\ {\isasymLongrightarrow}\ A\ {\isasymunion}\ B\ {\isasymsubseteq}\ C{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ F\ \isakeyword{where}\ {\isachardoublequoteopen}F\ {\isasymequiv}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{0}}inFn\ {\isacharcolon}{\kern0pt}{\isachardoublequoteopen}p{\isadigit{0}}\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}genericD\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ e\ \isakeyword{where}\ eH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Fix{\isacharparenleft}{\kern0pt}e{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ sym{\isacharparenleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}e\ {\isasymsubseteq}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}finite{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ HS{\isacharunderscore}{\kern0pt}iff\ symmetric{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ e{\isacharunderscore}{\kern0pt}Finite\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}implies{\isacharunderscore}{\kern0pt}Finite\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ listin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ mapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ injection{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ GH\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ injection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ GH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}injection{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}bijection{\isacharunderscore}{\kern0pt}fm\ mapin\ F{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}injection{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nat{\isacharcomma}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ F{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nat\ {\isasymand}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ assms\ GH\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ binmap{\isacharunderscore}{\kern0pt}eq\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ Finj{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ M{\isacharunderscore}{\kern0pt}basic{\isachardot}{\kern0pt}injection{\isacharunderscore}{\kern0pt}abs{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans{\isachardot}{\kern0pt}mbasic\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GH\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}SymExt\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ binmap{\isacharunderscore}{\kern0pt}eq\ assms\ SymExt{\isacharunderscore}{\kern0pt}def\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ F{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}F\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ range{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ inj{\isacharunderscore}{\kern0pt}bij{\isacharunderscore}{\kern0pt}range{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Finj\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ rangeinfinite{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymnot}Finite{\isacharparenleft}{\kern0pt}range{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}Finite\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ B{\isacharequal}{\kern0pt}{\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ eqpoll{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}Finite{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ eqpoll{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}Finite{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}\ n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isasymnot}Finite{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}RangeIndexes{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ {\isasymnot}\ Finite{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}RangeIndexes{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ n\ {\isasymin}\ {\isacharquery}{\kern0pt}RangeIndexes\ {\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Finite{\isacharunderscore}{\kern0pt}RepFun{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ n\ {\isasymin}\ {\isacharquery}{\kern0pt}RangeIndexes\ {\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ range{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n\ x\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ F{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ m\ \isakeyword{where}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Finj\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ binmap{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}m{\isasymin}nat{\isachardot}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ F{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}range{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ False\ \isacommand{using}\isamarkupfalse%
\ rangeinfinite\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}RangeIndexes\ {\isacharminus}{\kern0pt}\ e\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymnot}Finite{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}RangeIndexes\ {\isacharminus}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Finite{\isacharunderscore}{\kern0pt}{\isadigit{0}}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ Diff{\isacharunderscore}{\kern0pt}Finite\ {\isacharbrackleft}{\kern0pt}of\ e\ {\isacharquery}{\kern0pt}RangeIndexes{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ e{\isacharunderscore}{\kern0pt}Finite\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ \isakeyword{where}\ nH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymnotin}\ e{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ {\isacharquery}{\kern0pt}RangeIndexes{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ e{\isacharunderscore}{\kern0pt}Finite\ eH\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ i\ \isakeyword{where}\ iH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}i{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ F{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Finj\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ iH{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}i\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Finj\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ binmapin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ mapin\ GH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ rowin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nH\ assms\ GH\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ nH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ listin{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\ SymExt{\isacharunderscore}{\kern0pt}def\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ iH\ transM\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}SymExt\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapeq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ assms\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq\ nH\ GH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}i\ {\isasymin}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Finj\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ iH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ appin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}i\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapin\ binmapin\ SymExt{\isacharunderscore}{\kern0pt}trans\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}\ {\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}{\isasymphi}{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GH\ listin{\isacharprime}{\kern0pt}\ iH\ iH{\isacharprime}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}q\ {\isasymin}\ G{\isachardot}{\kern0pt}\ ForcesHS{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ GH\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ iH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ nH\ assms\ GH\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapeq\ GH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ pH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ pinFn\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}def\ M{\isacharunderscore}{\kern0pt}genericD\ pH\ assms\ GH\ \isacommand{by}\isamarkupfalse%
\ blast\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}Finite\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}implies{\isacharunderscore}{\kern0pt}Finite\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}A\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Fn{\isacharunderscore}{\kern0pt}def\ pinFn\ GH\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ a\ b\ c{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Finite{\isacharunderscore}{\kern0pt}RepFun{\isacharcomma}{\kern0pt}\ rule\ domain{\isacharunderscore}{\kern0pt}Finite{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}A\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}e\ {\isasymunion}\ domain{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}n{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}Finite{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Finite{\isacharunderscore}{\kern0pt}Un\ e{\isacharunderscore}{\kern0pt}Finite\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}B\ {\isasymsubseteq}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ eH\ assms\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\ nH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\ \ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ ex{\isacharunderscore}{\kern0pt}remain{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A{\isachardot}{\kern0pt}\ A\ {\isasymsubseteq}\ nat\ {\isasymLongrightarrow}\ Finite{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}a\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ a\ {\isasymnotin}\ A{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ A{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}A\ {\isacharequal}{\kern0pt}\ nat{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}Finite\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equalityI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}\ \isakeyword{where}\ n{\isacharprime}{\kern0pt}H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymnotin}\ {\isacharquery}{\kern0pt}B{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ ex{\isacharunderscore}{\kern0pt}remain\ {\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}B{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}B\ {\isasymsubseteq}\ nat{\isacartoucheclose}\ {\isacartoucheopen}Finite{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ f\ \isakeyword{where}\ {\isachardoublequoteopen}f\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ nat{\isacharcomma}{\kern0pt}\ x\ {\isasymnoteq}\ n\ {\isasymand}\ x\ {\isasymnoteq}\ n{\isacharprime}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymunion}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ n{\isachargreater}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ fn\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ fn{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ fx\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ x\ {\isasymnoteq}\ n\ {\isasymLongrightarrow}\ x\ {\isasymnoteq}\ n{\isacharprime}{\kern0pt}\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ finj\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}a\ b{\isachardot}{\kern0pt}\ a\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ b\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ a\ {\isasymnoteq}\ b\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}a\ {\isasymnoteq}\ f{\isacharbackquote}{\kern0pt}b{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ a\ b{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}a\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ a\ b{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}b\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fn\ fn{\isacharprime}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fn\ fn{\isacharprime}{\kern0pt}\ fx\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ a\ b{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}b\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fn\ fn{\isacharprime}{\kern0pt}\ fx\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fn\ fn{\isacharprime}{\kern0pt}\ fx\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ {\isasympsi}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasympsi}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Or{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ And{\isacharparenleft}{\kern0pt}Neg{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Neg{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Or{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Equal{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Equal{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}v{\isachardot}{\kern0pt}\ v\ {\isasymin}\ M\ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharcomma}{\kern0pt}\ nat{\isacharcomma}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ f{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}b\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}a\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ x\ {\isasymin}\ nat\ {\isasymand}\ x\ {\isacharequal}{\kern0pt}\ a\ {\isasymand}\ x\ {\isacharequal}{\kern0pt}\ b\ {\isasymand}\ x\ {\isasymnoteq}\ n\ {\isasymand}\ x\ {\isasymnoteq}\ n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ \isanewline
\ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ n\ {\isasymand}\ b\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymor}\ \isanewline
\ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}\ {\isasymand}\ b\ {\isacharequal}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasympsi}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Or{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}disjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Exists{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}Or{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}disjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}And{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ nH\ n{\isacharprime}{\kern0pt}H\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ nat\ {\isasymtimes}\ nat{\isachardot}{\kern0pt}\ x\ {\isasymin}\ f\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}f\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ nH\ n{\isacharprime}{\kern0pt}H\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ nat\ {\isasymtimes}\ nat{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasympsi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ nat\ {\isasymtimes}\ nat{\isachardot}{\kern0pt}\ x\ {\isasymin}\ f\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ cartprod{\isacharunderscore}{\kern0pt}closed\ transM\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ separation{\isacharunderscore}{\kern0pt}notation{\isacharcomma}{\kern0pt}\ rule\ separation{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasympsi}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ nH\ n{\isacharprime}{\kern0pt}H\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ cartprod{\isacharunderscore}{\kern0pt}closed\ transM\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ fin{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ surj{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ relation{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ nH\ n{\isacharprime}{\kern0pt}H\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ finj\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fx\ fn\ fn{\isacharprime}{\kern0pt}\ nH\ n{\isacharprime}{\kern0pt}H\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ {\isasympi}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymequiv}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}a\ b\ c{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymLongrightarrow}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ c{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ b\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ a\ b\ c\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ c{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}m{\isasymin}nat{\isachardot}{\kern0pt}\ {\isasymexists}l{\isasymin}{\isadigit{2}}{\isachardot}{\kern0pt}\ {\isasymlangle}{\isasymlangle}n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ c{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymlangle}f\ {\isacharbackquote}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ m{\isasymrangle}{\isacharcomma}{\kern0pt}\ l{\isasymrangle}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Fn{\isacharunderscore}{\kern0pt}permE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ pinFn\ fin\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ y\ z\ \isakeyword{where}\ xyzH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}z{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ c{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}z\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}z\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xyzH\ fn\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xyzH\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \ \ \ \ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}z\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xyzH\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z\ {\isasymnoteq}\ n\ {\isasymand}\ z\ {\isasymnoteq}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}z\ {\isasymnoteq}\ n{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}z\ {\isacharequal}{\kern0pt}\ z{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ fx\ xyzH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ z{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xyzH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isasymin}\ p\ {\isasymand}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ c{\isachargreater}{\kern0pt}\ {\isasymin}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xyzH\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ compatH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}compat{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ compat{\isacharunderscore}{\kern0pt}def\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}p\ {\isasymunion}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymin}\ Fn{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}p\ {\isasymunion}\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}leq{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ fin\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}subset\ {\isacharbrackleft}{\kern0pt}of\ f\ p{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\ fin\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Un{\isacharunderscore}{\kern0pt}closed\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ fin\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}function{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}insert\ function{\isacharunderscore}{\kern0pt}Fn{\isacharunderscore}{\kern0pt}perm\ {\isacharbrackleft}{\kern0pt}of\ f\ p{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}def\ fin\ pinFn\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}def\ pinFn\ fin\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force{\isacharcomma}{\kern0pt}\ blast{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ finite{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ domain{\isacharunderscore}{\kern0pt}closed\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ pinFn\ fin\ Fn{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}subset{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ Fn{\isacharunderscore}{\kern0pt}def\ fin\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ fin\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharunderscore}{\kern0pt}subset\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ piinfix\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ Fix{\isacharparenleft}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Fix{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}f\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasympi}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}n{\isacharcomma}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ nH\ n{\isacharprime}{\kern0pt}H\ eH\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fin\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ fnneq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isasymnoteq}\ n{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ccontr{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fn\ n{\isacharprime}{\kern0pt}H\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ pi{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}F{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}F{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ piinfix\ eH\ sym{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapeq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ pi{\isacharunderscore}{\kern0pt}preserves{\isacharunderscore}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ check{\isacharunderscore}{\kern0pt}fixpoint{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasympi}{\isacharunderscore}{\kern0pt}def\ fin\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto\ iH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ {\isasympi}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}pauto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nH\ fin\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ symmetry{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\ {\isasympi}{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ fin\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ listin\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ iH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nH\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}{\isasymphi}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ pH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ piForces{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapeq\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ qH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}q{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}q{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymin}\ Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymin}\ Fn{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ compatH\ compat{\isacharunderscore}{\kern0pt}def\ {\isasympi}{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}eq\ fin\ pinFn\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ listin{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ iH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ listin{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ iH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nH\ fin\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ p{\isacharequal}{\kern0pt}{\isachardoublequoteopen}p{\isachardoublequoteclose}\ \isakeyword{in}\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ qH\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharcomma}{\kern0pt}\ rule\ list{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ listin{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ ForcesH{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}H{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isasymand}\ q{\isasymin}H\ \ {\isasymlongrightarrow}\ \ SymExt{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\ listin{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\ qH\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}{\isasymphi}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}q{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ p{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p{\isachardoublequoteclose}\ \isakeyword{in}\ HS{\isacharunderscore}{\kern0pt}strengthening{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasympi}{\isacharunderscore}{\kern0pt}def\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto\ fin\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pinFn\ qH\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ iH\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transM\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fin\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ nH\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ piForces\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ ForcesH{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}H{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isasymand}\ q{\isasymin}H\ \ {\isasymlongrightarrow}\ \ SymExt{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing{\isacharunderscore}{\kern0pt}HS{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}perm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto\ {\isasympi}{\isacharunderscore}{\kern0pt}def\ fin\ qH\ {\isasymphi}{\isacharunderscore}{\kern0pt}def\ listin{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}{\isasymphi}{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ H\ \isakeyword{where}\ HH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymin}\ H{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}existence\ qH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ F\ \isakeyword{where}\ {\isachardoublequoteopen}F\ {\isasymequiv}\ val{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesH{\isadigit{1}}\ HH\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}val{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isacharquery}{\kern0pt}A{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ sats{\isacharunderscore}{\kern0pt}{\isasymphi}{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HH\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ listin{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ Fieq{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}i\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ F{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ subst\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HH\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nH\ HH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}F{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ForcesH{\isadigit{2}}\ HH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ F{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}val{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isacharquery}{\kern0pt}A{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ sats{\isacharunderscore}{\kern0pt}{\isasymphi}{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HH\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ listin{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ Fieq{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}i\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isakeyword{in}\ iffD{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ F{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ subst\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HH\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fin\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ nH\ HH\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ roweq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Fieq{\isadigit{1}}\ Fieq{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ rowneq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}H{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}distinct{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HH\ nH\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HH\ nH\ fin\ nat{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}def\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fnneq\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ roweq\ rowneq\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ no{\isacharunderscore}{\kern0pt}wellorder\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ r\ G\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}wellordered{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ False\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic\ \ {\isachardoublequoteopen}Fn{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}enum{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isachardoublequoteclose}\ {\isachardoublequoteopen}Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter{\isachardoublequoteclose}\ G\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}axioms\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}axioms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ f\ \isakeyword{where}\ {\isachardoublequoteopen}f\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}n{\isacharcomma}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ fpi{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ nat\ {\isasymrightarrow}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ relation{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ binmap{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}m\ n{\isachardot}{\kern0pt}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymnoteq}\ n\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}m\ {\isasymnoteq}\ f{\isacharbackquote}{\kern0pt}n{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ m\ n\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymnoteq}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ fm{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}m\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ fpi\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ fn{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ assms{\isadigit{1}}\ fpi\ Pi{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ binmap{\isacharunderscore}{\kern0pt}row{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ binmap{\isacharunderscore}{\kern0pt}row{\isacharunderscore}{\kern0pt}distinct{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}m\ {\isasymnoteq}\ f{\isacharbackquote}{\kern0pt}n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ fm\ fn\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fpi\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ natle{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}nat\ {\isasymlesssim}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ lepoll{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}g{\isasymin}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ g\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans{\isachardot}{\kern0pt}wellorder{\isacharunderscore}{\kern0pt}induces{\isacharunderscore}{\kern0pt}injection{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans\ natle\ assms\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ binmap{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ g\ \isakeyword{where}\ gH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}g\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ g{\isacharprime}{\kern0pt}\ \isakeyword{where}\ g{\isacharprime}{\kern0pt}H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ g{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ {\isachardoublequoteopen}g{\isacharprime}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ checknatH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ valcheck{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}contains{\isacharunderscore}{\kern0pt}max\ assms\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ binmap{\isacharprime}{\kern0pt}H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ binmap{\isacharunderscore}{\kern0pt}eq\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ mapeq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ g{\isacharprime}{\kern0pt}H\ checknatH\ binmap{\isacharprime}{\kern0pt}H\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ injection{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ sats{\isacharunderscore}{\kern0pt}injection{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}SymExt\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ SymExt{\isacharunderscore}{\kern0pt}def\ gH\ binmap{\isacharunderscore}{\kern0pt}eq\ assms\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ M{\isacharunderscore}{\kern0pt}basic{\isachardot}{\kern0pt}injection{\isacharunderscore}{\kern0pt}abs{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}SymExt\ gH\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ SymExt{\isacharunderscore}{\kern0pt}def\ gH\ binmap{\isacharunderscore}{\kern0pt}eq\ assms\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans{\isachardot}{\kern0pt}mbasic\ SymExt{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}trans\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}p\ {\isasymin}\ G{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}ForcesHS{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ injection{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}truth{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\ g{\isacharprime}{\kern0pt}H\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}injection{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mapeq\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}ForcesHS{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ injection{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}check{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ binmap{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ False\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ F{\isacharprime}{\kern0pt}{\isacharequal}{\kern0pt}g{\isacharprime}{\kern0pt}\ \isakeyword{in}\ no{\isacharunderscore}{\kern0pt}injection{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ g{\isacharprime}{\kern0pt}H\ M{\isacharunderscore}{\kern0pt}genericD\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isanewline
\ \ \isanewline
\isanewline
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/NotAC_Proof.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=9%:%
%:%31=10%:%
%:%35=10%:%
%:%36=10%:%
%:%37=11%:%
%:%38=11%:%
%:%43=11%:%
%:%46=12%:%
%:%47=13%:%
%:%48=13%:%
%:%49=14%:%
%:%50=15%:%
%:%51=16%:%
%:%52=17%:%
%:%59=18%:%
%:%60=18%:%
%:%61=19%:%
%:%62=20%:%
%:%63=20%:%
%:%64=21%:%
%:%65=22%:%
%:%66=22%:%
%:%67=23%:%
%:%68=23%:%
%:%69=24%:%
%:%70=24%:%
%:%71=25%:%
%:%72=25%:%
%:%73=26%:%
%:%74=27%:%
%:%75=27%:%
%:%76=28%:%
%:%77=28%:%
%:%78=29%:%
%:%79=29%:%
%:%80=30%:%
%:%81=30%:%
%:%82=31%:%
%:%83=31%:%
%:%84=32%:%
%:%85=32%:%
%:%86=33%:%
%:%87=33%:%
%:%88=34%:%
%:%89=34%:%
%:%90=35%:%
%:%91=35%:%
%:%92=36%:%
%:%93=36%:%
%:%94=37%:%
%:%95=37%:%
%:%96=38%:%
%:%97=38%:%
%:%98=39%:%
%:%99=40%:%
%:%100=40%:%
%:%101=41%:%
%:%102=41%:%
%:%103=42%:%
%:%104=42%:%
%:%105=43%:%
%:%106=43%:%
%:%107=44%:%
%:%108=44%:%
%:%109=45%:%
%:%110=45%:%
%:%111=46%:%
%:%112=46%:%
%:%113=47%:%
%:%114=47%:%
%:%115=48%:%
%:%116=48%:%
%:%117=49%:%
%:%118=49%:%
%:%119=50%:%
%:%120=50%:%
%:%121=51%:%
%:%122=51%:%
%:%123=52%:%
%:%124=52%:%
%:%125=53%:%
%:%126=53%:%
%:%127=54%:%
%:%128=54%:%
%:%129=55%:%
%:%130=55%:%
%:%131=56%:%
%:%132=56%:%
%:%133=57%:%
%:%134=57%:%
%:%135=58%:%
%:%136=58%:%
%:%137=59%:%
%:%138=59%:%
%:%139=60%:%
%:%140=61%:%
%:%141=61%:%
%:%142=62%:%
%:%143=62%:%
%:%144=63%:%
%:%145=63%:%
%:%146=64%:%
%:%147=65%:%
%:%148=65%:%
%:%149=66%:%
%:%150=66%:%
%:%151=67%:%
%:%152=67%:%
%:%153=68%:%
%:%154=68%:%
%:%155=69%:%
%:%156=69%:%
%:%157=70%:%
%:%158=70%:%
%:%159=71%:%
%:%160=71%:%
%:%161=72%:%
%:%162=72%:%
%:%163=73%:%
%:%164=73%:%
%:%165=74%:%
%:%166=74%:%
%:%167=75%:%
%:%168=75%:%
%:%169=76%:%
%:%170=76%:%
%:%171=77%:%
%:%172=78%:%
%:%173=78%:%
%:%174=78%:%
%:%175=79%:%
%:%176=80%:%
%:%177=80%:%
%:%178=81%:%
%:%179=82%:%
%:%180=82%:%
%:%181=82%:%
%:%182=82%:%
%:%183=83%:%
%:%184=84%:%
%:%185=84%:%
%:%186=85%:%
%:%187=85%:%
%:%188=86%:%
%:%189=86%:%
%:%190=87%:%
%:%191=87%:%
%:%192=87%:%
%:%193=88%:%
%:%194=88%:%
%:%195=89%:%
%:%196=89%:%
%:%197=90%:%
%:%198=91%:%
%:%199=91%:%
%:%200=92%:%
%:%201=92%:%
%:%202=93%:%
%:%203=93%:%
%:%204=94%:%
%:%205=94%:%
%:%206=94%:%
%:%207=95%:%
%:%208=95%:%
%:%209=96%:%
%:%210=96%:%
%:%211=97%:%
%:%212=97%:%
%:%213=98%:%
%:%214=99%:%
%:%215=99%:%
%:%216=100%:%
%:%217=100%:%
%:%218=101%:%
%:%219=101%:%
%:%220=102%:%
%:%221=102%:%
%:%222=103%:%
%:%223=103%:%
%:%224=104%:%
%:%225=104%:%
%:%226=105%:%
%:%227=105%:%
%:%228=106%:%
%:%229=106%:%
%:%230=107%:%
%:%231=107%:%
%:%232=107%:%
%:%233=108%:%
%:%234=108%:%
%:%235=109%:%
%:%236=109%:%
%:%237=110%:%
%:%238=110%:%
%:%239=110%:%
%:%240=111%:%
%:%241=111%:%
%:%242=112%:%
%:%243=112%:%
%:%244=113%:%
%:%245=113%:%
%:%246=114%:%
%:%247=114%:%
%:%248=115%:%
%:%249=115%:%
%:%250=116%:%
%:%251=116%:%
%:%252=117%:%
%:%253=117%:%
%:%254=117%:%
%:%255=118%:%
%:%256=118%:%
%:%257=119%:%
%:%258=119%:%
%:%259=120%:%
%:%260=120%:%
%:%261=121%:%
%:%262=121%:%
%:%263=122%:%
%:%264=122%:%
%:%265=123%:%
%:%266=123%:%
%:%267=124%:%
%:%268=125%:%
%:%269=125%:%
%:%270=126%:%
%:%271=126%:%
%:%272=127%:%
%:%273=127%:%
%:%274=128%:%
%:%275=128%:%
%:%276=129%:%
%:%277=129%:%
%:%278=129%:%
%:%279=130%:%
%:%280=130%:%
%:%281=131%:%
%:%282=131%:%
%:%283=132%:%
%:%284=132%:%
%:%285=133%:%
%:%286=133%:%
%:%287=134%:%
%:%288=134%:%
%:%289=135%:%
%:%290=135%:%
%:%291=136%:%
%:%292=136%:%
%:%293=137%:%
%:%294=137%:%
%:%295=138%:%
%:%296=138%:%
%:%297=139%:%
%:%298=139%:%
%:%299=140%:%
%:%300=140%:%
%:%301=140%:%
%:%302=141%:%
%:%303=141%:%
%:%304=142%:%
%:%305=142%:%
%:%306=143%:%
%:%307=143%:%
%:%308=144%:%
%:%309=144%:%
%:%310=145%:%
%:%311=145%:%
%:%312=146%:%
%:%313=146%:%
%:%314=147%:%
%:%315=147%:%
%:%316=147%:%
%:%317=147%:%
%:%318=147%:%
%:%319=148%:%
%:%320=148%:%
%:%321=148%:%
%:%322=149%:%
%:%323=149%:%
%:%324=150%:%
%:%325=150%:%
%:%326=151%:%
%:%327=151%:%
%:%328=152%:%
%:%329=152%:%
%:%330=152%:%
%:%331=153%:%
%:%332=153%:%
%:%333=153%:%
%:%334=154%:%
%:%335=154%:%
%:%336=154%:%
%:%337=154%:%
%:%338=154%:%
%:%339=155%:%
%:%340=155%:%
%:%341=156%:%
%:%342=157%:%
%:%343=157%:%
%:%344=157%:%
%:%345=158%:%
%:%346=158%:%
%:%347=159%:%
%:%348=159%:%
%:%349=160%:%
%:%350=160%:%
%:%351=161%:%
%:%352=161%:%
%:%353=162%:%
%:%354=162%:%
%:%355=163%:%
%:%356=163%:%
%:%357=164%:%
%:%358=164%:%
%:%359=165%:%
%:%360=165%:%
%:%361=166%:%
%:%362=166%:%
%:%363=166%:%
%:%364=166%:%
%:%365=166%:%
%:%366=167%:%
%:%367=167%:%
%:%368=167%:%
%:%369=167%:%
%:%370=167%:%
%:%371=168%:%
%:%372=168%:%
%:%373=168%:%
%:%374=169%:%
%:%375=169%:%
%:%376=170%:%
%:%377=170%:%
%:%378=171%:%
%:%379=171%:%
%:%380=172%:%
%:%381=173%:%
%:%382=173%:%
%:%383=174%:%
%:%384=174%:%
%:%385=175%:%
%:%386=175%:%
%:%387=176%:%
%:%388=176%:%
%:%389=177%:%
%:%390=178%:%
%:%391=178%:%
%:%392=179%:%
%:%393=179%:%
%:%394=180%:%
%:%395=180%:%
%:%396=181%:%
%:%397=181%:%
%:%398=182%:%
%:%399=182%:%
%:%400=183%:%
%:%401=183%:%
%:%402=184%:%
%:%403=184%:%
%:%404=185%:%
%:%405=185%:%
%:%406=186%:%
%:%407=186%:%
%:%408=187%:%
%:%409=188%:%
%:%410=188%:%
%:%411=188%:%
%:%412=189%:%
%:%413=189%:%
%:%414=190%:%
%:%415=190%:%
%:%416=191%:%
%:%417=191%:%
%:%418=192%:%
%:%419=192%:%
%:%420=193%:%
%:%421=193%:%
%:%422=194%:%
%:%423=195%:%
%:%424=195%:%
%:%425=196%:%
%:%426=196%:%
%:%427=197%:%
%:%428=197%:%
%:%429=198%:%
%:%430=198%:%
%:%431=199%:%
%:%432=199%:%
%:%433=200%:%
%:%434=200%:%
%:%435=201%:%
%:%436=202%:%
%:%437=202%:%
%:%438=203%:%
%:%439=203%:%
%:%440=204%:%
%:%441=204%:%
%:%442=205%:%
%:%443=205%:%
%:%444=206%:%
%:%445=206%:%
%:%446=206%:%
%:%447=207%:%
%:%448=207%:%
%:%449=208%:%
%:%450=208%:%
%:%451=209%:%
%:%452=210%:%
%:%453=210%:%
%:%454=211%:%
%:%455=211%:%
%:%456=212%:%
%:%457=212%:%
%:%458=213%:%
%:%459=213%:%
%:%460=214%:%
%:%461=214%:%
%:%462=214%:%
%:%463=215%:%
%:%464=215%:%
%:%465=216%:%
%:%466=216%:%
%:%467=217%:%
%:%468=217%:%
%:%469=218%:%
%:%470=218%:%
%:%471=219%:%
%:%472=219%:%
%:%473=220%:%
%:%474=220%:%
%:%475=221%:%
%:%476=221%:%
%:%477=222%:%
%:%478=222%:%
%:%479=223%:%
%:%480=223%:%
%:%481=224%:%
%:%482=224%:%
%:%483=225%:%
%:%484=225%:%
%:%485=225%:%
%:%486=225%:%
%:%487=226%:%
%:%488=227%:%
%:%489=227%:%
%:%490=227%:%
%:%491=227%:%
%:%492=228%:%
%:%493=229%:%
%:%494=229%:%
%:%495=230%:%
%:%496=230%:%
%:%497=231%:%
%:%498=231%:%
%:%499=232%:%
%:%500=233%:%
%:%501=233%:%
%:%502=234%:%
%:%503=234%:%
%:%504=235%:%
%:%505=235%:%
%:%506=236%:%
%:%507=236%:%
%:%508=237%:%
%:%509=237%:%
%:%510=238%:%
%:%511=238%:%
%:%512=239%:%
%:%513=239%:%
%:%514=240%:%
%:%515=240%:%
%:%516=241%:%
%:%517=241%:%
%:%518=242%:%
%:%519=242%:%
%:%520=243%:%
%:%521=243%:%
%:%522=243%:%
%:%523=243%:%
%:%524=243%:%
%:%525=244%:%
%:%526=244%:%
%:%527=244%:%
%:%528=245%:%
%:%529=245%:%
%:%530=246%:%
%:%531=246%:%
%:%532=247%:%
%:%533=248%:%
%:%534=248%:%
%:%535=249%:%
%:%536=249%:%
%:%537=250%:%
%:%538=250%:%
%:%539=251%:%
%:%540=252%:%
%:%541=252%:%
%:%542=253%:%
%:%543=253%:%
%:%544=254%:%
%:%545=254%:%
%:%546=255%:%
%:%547=255%:%
%:%548=256%:%
%:%549=256%:%
%:%550=257%:%
%:%551=257%:%
%:%552=258%:%
%:%553=258%:%
%:%554=259%:%
%:%555=259%:%
%:%556=260%:%
%:%557=260%:%
%:%558=261%:%
%:%559=261%:%
%:%560=262%:%
%:%561=262%:%
%:%562=263%:%
%:%563=264%:%
%:%564=264%:%
%:%565=265%:%
%:%566=266%:%
%:%567=266%:%
%:%568=267%:%
%:%569=267%:%
%:%570=268%:%
%:%571=268%:%
%:%572=269%:%
%:%573=269%:%
%:%574=270%:%
%:%575=270%:%
%:%576=271%:%
%:%577=271%:%
%:%578=272%:%
%:%579=272%:%
%:%580=273%:%
%:%581=273%:%
%:%582=274%:%
%:%583=274%:%
%:%584=275%:%
%:%585=275%:%
%:%586=276%:%
%:%587=276%:%
%:%588=277%:%
%:%589=277%:%
%:%590=278%:%
%:%591=279%:%
%:%592=279%:%
%:%593=280%:%
%:%594=280%:%
%:%595=281%:%
%:%596=281%:%
%:%597=282%:%
%:%598=282%:%
%:%599=283%:%
%:%600=283%:%
%:%601=284%:%
%:%602=284%:%
%:%603=285%:%
%:%604=285%:%
%:%605=286%:%
%:%606=286%:%
%:%607=287%:%
%:%608=287%:%
%:%609=288%:%
%:%610=288%:%
%:%611=289%:%
%:%612=289%:%
%:%613=290%:%
%:%614=290%:%
%:%615=291%:%
%:%616=291%:%
%:%617=292%:%
%:%618=293%:%
%:%619=293%:%
%:%622=296%:%
%:%623=297%:%
%:%624=298%:%
%:%625=298%:%
%:%626=299%:%
%:%627=300%:%
%:%628=300%:%
%:%629=301%:%
%:%632=304%:%
%:%633=305%:%
%:%634=305%:%
%:%635=306%:%
%:%636=306%:%
%:%637=307%:%
%:%638=307%:%
%:%639=308%:%
%:%640=308%:%
%:%641=309%:%
%:%642=309%:%
%:%643=310%:%
%:%644=310%:%
%:%645=311%:%
%:%646=311%:%
%:%647=312%:%
%:%648=312%:%
%:%649=313%:%
%:%650=313%:%
%:%651=314%:%
%:%652=314%:%
%:%653=315%:%
%:%654=315%:%
%:%655=316%:%
%:%656=316%:%
%:%657=317%:%
%:%658=317%:%
%:%659=318%:%
%:%660=318%:%
%:%661=319%:%
%:%662=319%:%
%:%663=320%:%
%:%664=320%:%
%:%665=321%:%
%:%666=321%:%
%:%667=322%:%
%:%668=322%:%
%:%669=323%:%
%:%670=323%:%
%:%671=324%:%
%:%672=324%:%
%:%673=324%:%
%:%674=325%:%
%:%675=325%:%
%:%676=326%:%
%:%677=326%:%
%:%678=327%:%
%:%679=327%:%
%:%680=328%:%
%:%681=328%:%
%:%682=329%:%
%:%683=329%:%
%:%684=330%:%
%:%685=330%:%
%:%686=331%:%
%:%687=331%:%
%:%688=332%:%
%:%689=332%:%
%:%690=333%:%
%:%691=333%:%
%:%692=334%:%
%:%693=334%:%
%:%694=335%:%
%:%695=335%:%
%:%696=336%:%
%:%697=336%:%
%:%698=337%:%
%:%699=337%:%
%:%700=338%:%
%:%701=338%:%
%:%702=339%:%
%:%703=339%:%
%:%704=340%:%
%:%705=340%:%
%:%706=341%:%
%:%707=341%:%
%:%708=341%:%
%:%709=342%:%
%:%710=342%:%
%:%711=343%:%
%:%712=343%:%
%:%713=344%:%
%:%714=344%:%
%:%715=345%:%
%:%716=345%:%
%:%717=346%:%
%:%718=346%:%
%:%719=347%:%
%:%720=347%:%
%:%721=348%:%
%:%722=348%:%
%:%723=349%:%
%:%724=349%:%
%:%725=350%:%
%:%726=350%:%
%:%727=351%:%
%:%728=351%:%
%:%729=352%:%
%:%730=352%:%
%:%731=353%:%
%:%732=353%:%
%:%733=354%:%
%:%734=354%:%
%:%735=355%:%
%:%736=355%:%
%:%737=356%:%
%:%738=356%:%
%:%739=357%:%
%:%740=357%:%
%:%741=358%:%
%:%742=358%:%
%:%743=359%:%
%:%744=360%:%
%:%745=360%:%
%:%746=361%:%
%:%747=362%:%
%:%748=362%:%
%:%749=363%:%
%:%750=363%:%
%:%751=364%:%
%:%752=364%:%
%:%753=365%:%
%:%754=365%:%
%:%755=366%:%
%:%756=367%:%
%:%757=367%:%
%:%758=368%:%
%:%759=368%:%
%:%760=369%:%
%:%761=369%:%
%:%762=370%:%
%:%763=370%:%
%:%764=371%:%
%:%765=371%:%
%:%766=371%:%
%:%767=372%:%
%:%768=372%:%
%:%769=373%:%
%:%770=373%:%
%:%771=374%:%
%:%772=375%:%
%:%773=375%:%
%:%774=376%:%
%:%775=376%:%
%:%776=377%:%
%:%777=377%:%
%:%778=378%:%
%:%779=378%:%
%:%780=378%:%
%:%781=378%:%
%:%782=378%:%
%:%783=379%:%
%:%784=379%:%
%:%785=379%:%
%:%786=379%:%
%:%787=379%:%
%:%788=380%:%
%:%789=380%:%
%:%790=380%:%
%:%791=380%:%
%:%792=380%:%
%:%793=381%:%
%:%794=381%:%
%:%795=382%:%
%:%796=382%:%
%:%797=383%:%
%:%798=383%:%
%:%799=383%:%
%:%800=384%:%
%:%801=384%:%
%:%802=385%:%
%:%803=385%:%
%:%804=386%:%
%:%805=386%:%
%:%806=386%:%
%:%807=386%:%
%:%808=386%:%
%:%809=387%:%
%:%810=387%:%
%:%811=387%:%
%:%812=387%:%
%:%813=387%:%
%:%814=388%:%
%:%815=388%:%
%:%816=389%:%
%:%817=389%:%
%:%818=390%:%
%:%819=390%:%
%:%820=390%:%
%:%821=390%:%
%:%822=390%:%
%:%823=391%:%
%:%824=391%:%
%:%825=391%:%
%:%826=391%:%
%:%827=391%:%
%:%828=392%:%
%:%829=392%:%
%:%830=392%:%
%:%831=392%:%
%:%832=392%:%
%:%833=393%:%
%:%834=393%:%
%:%835=393%:%
%:%836=393%:%
%:%837=393%:%
%:%838=394%:%
%:%839=394%:%
%:%840=394%:%
%:%841=394%:%
%:%842=394%:%
%:%843=395%:%
%:%844=395%:%
%:%845=396%:%
%:%846=396%:%
%:%847=397%:%
%:%848=397%:%
%:%849=398%:%
%:%850=398%:%
%:%851=398%:%
%:%852=399%:%
%:%853=399%:%
%:%854=400%:%
%:%855=400%:%
%:%856=401%:%
%:%857=401%:%
%:%858=402%:%
%:%859=402%:%
%:%860=403%:%
%:%861=403%:%
%:%862=404%:%
%:%863=404%:%
%:%864=405%:%
%:%865=405%:%
%:%866=406%:%
%:%867=406%:%
%:%868=407%:%
%:%869=407%:%
%:%870=408%:%
%:%871=408%:%
%:%872=409%:%
%:%873=409%:%
%:%874=410%:%
%:%875=410%:%
%:%876=411%:%
%:%877=411%:%
%:%878=412%:%
%:%879=412%:%
%:%880=413%:%
%:%881=413%:%
%:%882=414%:%
%:%883=414%:%
%:%884=415%:%
%:%885=415%:%
%:%886=416%:%
%:%887=416%:%
%:%888=417%:%
%:%889=417%:%
%:%890=418%:%
%:%891=418%:%
%:%892=419%:%
%:%893=419%:%
%:%894=420%:%
%:%895=420%:%
%:%896=421%:%
%:%897=421%:%
%:%898=422%:%
%:%899=422%:%
%:%900=423%:%
%:%901=423%:%
%:%902=424%:%
%:%903=424%:%
%:%904=425%:%
%:%905=425%:%
%:%906=426%:%
%:%907=426%:%
%:%908=427%:%
%:%909=427%:%
%:%910=428%:%
%:%911=428%:%
%:%912=429%:%
%:%913=429%:%
%:%914=430%:%
%:%915=430%:%
%:%916=431%:%
%:%917=431%:%
%:%918=432%:%
%:%919=432%:%
%:%920=433%:%
%:%921=433%:%
%:%922=434%:%
%:%923=434%:%
%:%924=435%:%
%:%925=435%:%
%:%926=436%:%
%:%927=436%:%
%:%928=437%:%
%:%929=437%:%
%:%930=438%:%
%:%931=438%:%
%:%932=439%:%
%:%933=439%:%
%:%934=440%:%
%:%935=440%:%
%:%936=441%:%
%:%937=442%:%
%:%938=442%:%
%:%939=443%:%
%:%940=443%:%
%:%941=444%:%
%:%942=444%:%
%:%943=445%:%
%:%944=445%:%
%:%945=446%:%
%:%946=446%:%
%:%947=447%:%
%:%948=447%:%
%:%949=448%:%
%:%950=448%:%
%:%951=449%:%
%:%952=449%:%
%:%953=450%:%
%:%954=450%:%
%:%955=451%:%
%:%956=451%:%
%:%957=452%:%
%:%958=452%:%
%:%959=453%:%
%:%960=453%:%
%:%961=454%:%
%:%962=454%:%
%:%963=455%:%
%:%964=456%:%
%:%965=456%:%
%:%966=457%:%
%:%967=457%:%
%:%968=458%:%
%:%969=458%:%
%:%970=459%:%
%:%971=459%:%
%:%972=460%:%
%:%973=461%:%
%:%974=461%:%
%:%975=462%:%
%:%976=462%:%
%:%977=463%:%
%:%978=463%:%
%:%979=464%:%
%:%980=465%:%
%:%981=465%:%
%:%982=466%:%
%:%983=466%:%
%:%984=467%:%
%:%985=467%:%
%:%986=468%:%
%:%987=468%:%
%:%988=469%:%
%:%989=469%:%
%:%990=470%:%
%:%991=470%:%
%:%992=471%:%
%:%993=471%:%
%:%994=472%:%
%:%995=472%:%
%:%996=473%:%
%:%997=473%:%
%:%998=474%:%
%:%999=474%:%
%:%1000=475%:%
%:%1001=476%:%
%:%1002=476%:%
%:%1003=477%:%
%:%1004=477%:%
%:%1005=478%:%
%:%1006=478%:%
%:%1007=479%:%
%:%1008=479%:%
%:%1009=480%:%
%:%1010=480%:%
%:%1011=481%:%
%:%1012=481%:%
%:%1013=482%:%
%:%1014=482%:%
%:%1015=483%:%
%:%1016=483%:%
%:%1017=484%:%
%:%1018=484%:%
%:%1019=485%:%
%:%1020=485%:%
%:%1021=486%:%
%:%1022=486%:%
%:%1023=487%:%
%:%1024=487%:%
%:%1025=488%:%
%:%1026=488%:%
%:%1027=489%:%
%:%1028=489%:%
%:%1029=490%:%
%:%1030=490%:%
%:%1031=490%:%
%:%1032=491%:%
%:%1033=491%:%
%:%1034=492%:%
%:%1035=492%:%
%:%1036=493%:%
%:%1037=494%:%
%:%1038=494%:%
%:%1039=495%:%
%:%1040=495%:%
%:%1041=496%:%
%:%1042=496%:%
%:%1043=497%:%
%:%1044=498%:%
%:%1045=498%:%
%:%1046=499%:%
%:%1047=499%:%
%:%1048=500%:%
%:%1049=500%:%
%:%1050=501%:%
%:%1051=501%:%
%:%1052=502%:%
%:%1053=502%:%
%:%1054=503%:%
%:%1055=503%:%
%:%1056=504%:%
%:%1057=504%:%
%:%1058=505%:%
%:%1059=505%:%
%:%1060=506%:%
%:%1061=506%:%
%:%1062=507%:%
%:%1063=507%:%
%:%1064=508%:%
%:%1065=509%:%
%:%1066=509%:%
%:%1067=510%:%
%:%1068=510%:%
%:%1069=511%:%
%:%1070=511%:%
%:%1071=512%:%
%:%1072=512%:%
%:%1073=513%:%
%:%1074=513%:%
%:%1075=514%:%
%:%1076=514%:%
%:%1077=515%:%
%:%1078=515%:%
%:%1079=516%:%
%:%1080=516%:%
%:%1081=517%:%
%:%1082=517%:%
%:%1083=518%:%
%:%1084=518%:%
%:%1085=519%:%
%:%1086=519%:%
%:%1087=520%:%
%:%1088=521%:%
%:%1089=521%:%
%:%1090=522%:%
%:%1091=522%:%
%:%1092=523%:%
%:%1093=523%:%
%:%1094=524%:%
%:%1095=524%:%
%:%1096=525%:%
%:%1097=525%:%
%:%1098=526%:%
%:%1099=526%:%
%:%1100=527%:%
%:%1101=527%:%
%:%1102=528%:%
%:%1103=528%:%
%:%1104=529%:%
%:%1105=529%:%
%:%1106=530%:%
%:%1107=530%:%
%:%1108=531%:%
%:%1109=531%:%
%:%1110=532%:%
%:%1111=532%:%
%:%1112=532%:%
%:%1113=533%:%
%:%1114=533%:%
%:%1115=534%:%
%:%1116=534%:%
%:%1117=535%:%
%:%1118=535%:%
%:%1119=536%:%
%:%1120=536%:%
%:%1121=537%:%
%:%1122=537%:%
%:%1123=538%:%
%:%1124=538%:%
%:%1125=539%:%
%:%1126=539%:%
%:%1127=540%:%
%:%1128=541%:%
%:%1129=541%:%
%:%1130=542%:%
%:%1131=542%:%
%:%1132=543%:%
%:%1133=543%:%
%:%1134=544%:%
%:%1135=544%:%
%:%1136=545%:%
%:%1137=545%:%
%:%1138=546%:%
%:%1139=546%:%
%:%1140=547%:%
%:%1141=547%:%
%:%1142=548%:%
%:%1143=548%:%
%:%1144=549%:%
%:%1145=549%:%
%:%1146=550%:%
%:%1147=550%:%
%:%1148=551%:%
%:%1149=551%:%
%:%1150=552%:%
%:%1151=552%:%
%:%1152=553%:%
%:%1153=553%:%
%:%1154=554%:%
%:%1155=554%:%
%:%1156=555%:%
%:%1157=555%:%
%:%1158=556%:%
%:%1159=556%:%
%:%1160=557%:%
%:%1161=557%:%
%:%1162=558%:%
%:%1163=558%:%
%:%1164=559%:%
%:%1165=559%:%
%:%1166=559%:%
%:%1167=560%:%
%:%1168=560%:%
%:%1169=561%:%
%:%1170=561%:%
%:%1171=562%:%
%:%1172=562%:%
%:%1173=563%:%
%:%1174=563%:%
%:%1175=564%:%
%:%1176=564%:%
%:%1177=565%:%
%:%1178=565%:%
%:%1179=566%:%
%:%1180=566%:%
%:%1181=567%:%
%:%1182=568%:%
%:%1183=568%:%
%:%1184=569%:%
%:%1185=569%:%
%:%1186=570%:%
%:%1187=570%:%
%:%1188=571%:%
%:%1189=572%:%
%:%1190=572%:%
%:%1191=573%:%
%:%1192=574%:%
%:%1193=574%:%
%:%1194=575%:%
%:%1195=575%:%
%:%1196=576%:%
%:%1197=576%:%
%:%1198=577%:%
%:%1199=577%:%
%:%1200=577%:%
%:%1201=578%:%
%:%1202=578%:%
%:%1203=579%:%
%:%1204=579%:%
%:%1205=580%:%
%:%1206=580%:%
%:%1207=581%:%
%:%1208=581%:%
%:%1209=582%:%
%:%1210=582%:%
%:%1211=583%:%
%:%1212=583%:%
%:%1213=584%:%
%:%1214=584%:%
%:%1215=584%:%
%:%1216=585%:%
%:%1217=585%:%
%:%1218=586%:%
%:%1219=586%:%
%:%1220=587%:%
%:%1221=587%:%
%:%1222=588%:%
%:%1223=588%:%
%:%1224=589%:%
%:%1225=589%:%
%:%1226=590%:%
%:%1227=590%:%
%:%1228=591%:%
%:%1229=591%:%
%:%1230=592%:%
%:%1231=593%:%
%:%1232=593%:%
%:%1233=594%:%
%:%1234=594%:%
%:%1235=595%:%
%:%1236=595%:%
%:%1237=596%:%
%:%1238=596%:%
%:%1239=596%:%
%:%1240=597%:%
%:%1241=597%:%
%:%1242=598%:%
%:%1243=598%:%
%:%1244=599%:%
%:%1245=599%:%
%:%1246=600%:%
%:%1247=600%:%
%:%1248=601%:%
%:%1249=601%:%
%:%1250=602%:%
%:%1251=602%:%
%:%1252=603%:%
%:%1253=603%:%
%:%1254=603%:%
%:%1255=604%:%
%:%1256=604%:%
%:%1257=605%:%
%:%1258=605%:%
%:%1259=606%:%
%:%1260=606%:%
%:%1261=607%:%
%:%1262=607%:%
%:%1263=608%:%
%:%1264=608%:%
%:%1265=609%:%
%:%1266=609%:%
%:%1267=610%:%
%:%1268=610%:%
%:%1269=611%:%
%:%1270=611%:%
%:%1271=612%:%
%:%1272=613%:%
%:%1273=613%:%
%:%1274=613%:%
%:%1275=613%:%
%:%1276=614%:%
%:%1277=615%:%
%:%1278=615%:%
%:%1279=616%:%
%:%1280=616%:%
%:%1281=617%:%
%:%1282=617%:%
%:%1283=618%:%
%:%1284=618%:%
%:%1285=619%:%
%:%1286=619%:%
%:%1287=620%:%
%:%1288=620%:%
%:%1289=621%:%
%:%1290=621%:%
%:%1291=622%:%
%:%1292=622%:%
%:%1293=623%:%
%:%1294=623%:%
%:%1295=624%:%
%:%1296=625%:%
%:%1297=625%:%
%:%1298=626%:%
%:%1299=626%:%
%:%1300=627%:%
%:%1301=627%:%
%:%1302=628%:%
%:%1308=628%:%
%:%1311=629%:%
%:%1312=630%:%
%:%1313=630%:%
%:%1314=631%:%
%:%1315=632%:%
%:%1316=633%:%
%:%1323=634%:%
%:%1324=634%:%
%:%1325=635%:%
%:%1326=636%:%
%:%1327=636%:%
%:%1328=637%:%
%:%1329=637%:%
%:%1330=638%:%
%:%1331=638%:%
%:%1332=639%:%
%:%1333=639%:%
%:%1334=640%:%
%:%1335=640%:%
%:%1336=641%:%
%:%1337=641%:%
%:%1338=642%:%
%:%1339=642%:%
%:%1340=643%:%
%:%1341=643%:%
%:%1342=644%:%
%:%1343=644%:%
%:%1344=645%:%
%:%1345=645%:%
%:%1346=646%:%
%:%1347=646%:%
%:%1348=647%:%
%:%1349=647%:%
%:%1350=648%:%
%:%1351=649%:%
%:%1352=650%:%
%:%1353=650%:%
%:%1354=651%:%
%:%1355=651%:%
%:%1356=652%:%
%:%1357=652%:%
%:%1358=653%:%
%:%1359=653%:%
%:%1360=654%:%
%:%1361=654%:%
%:%1362=655%:%
%:%1363=655%:%
%:%1364=656%:%
%:%1365=656%:%
%:%1366=657%:%
%:%1367=657%:%
%:%1368=658%:%
%:%1369=658%:%
%:%1370=659%:%
%:%1371=659%:%
%:%1372=660%:%
%:%1373=660%:%
%:%1374=661%:%
%:%1375=661%:%
%:%1376=662%:%
%:%1377=662%:%
%:%1378=663%:%
%:%1379=664%:%
%:%1380=664%:%
%:%1381=665%:%
%:%1382=665%:%
%:%1383=666%:%
%:%1384=666%:%
%:%1385=667%:%
%:%1386=667%:%
%:%1387=668%:%
%:%1388=669%:%
%:%1389=669%:%
%:%1390=670%:%
%:%1391=670%:%
%:%1392=671%:%
%:%1393=671%:%
%:%1394=672%:%
%:%1395=672%:%
%:%1396=673%:%
%:%1397=673%:%
%:%1398=673%:%
%:%1399=674%:%
%:%1400=674%:%
%:%1401=675%:%
%:%1402=675%:%
%:%1403=676%:%
%:%1404=676%:%
%:%1405=677%:%
%:%1406=677%:%
%:%1407=677%:%
%:%1408=678%:%
%:%1409=678%:%
%:%1410=679%:%
%:%1411=679%:%
%:%1412=680%:%
%:%1413=680%:%
%:%1414=681%:%
%:%1415=681%:%
%:%1416=681%:%
%:%1417=681%:%
%:%1418=681%:%
%:%1419=682%:%
%:%1420=683%:%
%:%1421=683%:%
%:%1422=684%:%
%:%1423=684%:%
%:%1424=685%:%
%:%1425=685%:%
%:%1426=686%:%
%:%1427=686%:%
%:%1428=687%:%
%:%1429=687%:%
%:%1430=688%:%
%:%1431=688%:%
%:%1432=689%:%
%:%1433=689%:%
%:%1434=690%:%
%:%1435=690%:%
%:%1436=691%:%
%:%1437=691%:%
%:%1438=691%:%
%:%1439=691%:%
%:%1440=692%:%
%:%1441=693%:%
%:%1442=693%:%
%:%1443=693%:%
%:%1444=693%:%
%:%1445=693%:%
%:%1446=694%:%
%:%1447=694%:%
%:%1448=695%:%
%:%1449=695%:%
%:%1450=696%:%
%:%1451=696%:%
%:%1452=697%:%
%:%1453=697%:%
%:%1454=698%:%
%:%1455=698%:%
%:%1456=699%:%
%:%1457=699%:%
%:%1458=700%:%
%:%1459=700%:%
%:%1460=701%:%
%:%1461=702%:%
%:%1462=702%:%
%:%1463=703%:%
%:%1464=703%:%
%:%1465=704%:%
%:%1466=704%:%
%:%1467=705%:%
%:%1468=706%:%
%:%1469=706%:%
%:%1470=707%:%
%:%1471=707%:%
%:%1472=708%:%
%:%1473=708%:%
%:%1474=709%:%
%:%1475=709%:%
%:%1476=710%:%
%:%1477=710%:%
%:%1478=711%:%
%:%1479=711%:%
%:%1480=712%:%
%:%1481=712%:%
%:%1482=713%:%
%:%1483=713%:%
%:%1484=714%:%
%:%1485=714%:%
%:%1486=715%:%
%:%1487=715%:%
%:%1488=715%:%
%:%1489=716%:%
%:%1490=716%:%
%:%1491=717%:%
%:%1492=717%:%
%:%1493=718%:%
%:%1494=718%:%
%:%1495=719%:%
%:%1496=719%:%
%:%1497=720%:%
%:%1498=720%:%
%:%1499=721%:%
%:%1500=721%:%
%:%1501=722%:%
%:%1502=722%:%
%:%1503=723%:%
%:%1504=723%:%
%:%1505=724%:%
%:%1506=724%:%
%:%1507=724%:%
%:%1508=724%:%
%:%1509=725%:%
%:%1510=725%:%
%:%1511=725%:%
%:%1512=726%:%
%:%1513=726%:%
%:%1514=727%:%
%:%1515=727%:%
%:%1516=728%:%
%:%1517=728%:%
%:%1518=729%:%
%:%1524=729%:%
%:%1527=730%:%
%:%1528=731%:%
%:%1529=732%:%
%:%1530=733%:%
%:%1531=734%:%
%:%1532=735%:%
%:%1533=736%:%
%:%1534=737%:%
%:%1535=737%:%
%:%1542=738%:%