### 2024-11-01 code/output/document/Ordinals_In_MG.tex

%
\begin{isabellebody}%
\setisabellecontext{Ordinals{\isacharunderscore}{\kern0pt}In{\isacharunderscore}{\kern0pt}MG}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Ordinals in generic extensions%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Ordinals{\isacharunderscore}{\kern0pt}In{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ Forcing{\isacharunderscore}{\kern0pt}Theorems\ Relative{\isacharunderscore}{\kern0pt}Univ\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}val{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ rule{\isacharcolon}{\kern0pt}ed{\isacharunderscore}{\kern0pt}induction{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}Q{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}\ x{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ u{\isasymin}{\isacharbraceleft}{\kern0pt}t{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ \ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}x\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}val\ \isacommand{unfolding}\isamarkupfalse%
\ Sep{\isacharunderscore}{\kern0pt}and{\isacharunderscore}{\kern0pt}Replace\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}u{\isasymin}{\isacharbraceleft}{\kern0pt}t{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ \ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}x\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ y\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}{\isacharbrackleft}{\kern0pt}OF\ that{\isacharbrackright}{\kern0pt}\ rank{\isacharunderscore}{\kern0pt}ed{\isacharbrackleft}{\kern0pt}OF\ that{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ intro{\isacharcolon}{\kern0pt}lt{\isacharunderscore}{\kern0pt}trans{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymUnion}u{\isasymin}{\isacharbraceleft}{\kern0pt}t{\isasymin}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P\ {\isachardot}{\kern0pt}\ \ {\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}x\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ UN{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}MG{\isacharunderscore}{\kern0pt}iff{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymalpha}\ {\isasymin}\ M\ {\isasymlongleftrightarrow}\ {\isasymalpha}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymalpha}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isasymalpha}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharbrackleft}{\kern0pt}THEN\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G{\isacharcomma}{\kern0pt}\ THEN\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}MG{\isacharbrackright}{\kern0pt}\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymalpha}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymalpha}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isasymle}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymalpha}\ {\isasymle}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymalpha}\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymalpha}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ cons{\isacharunderscore}{\kern0pt}closed\ transitivity{\isacharbrackleft}{\kern0pt}of\ {\isasymalpha}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ rank{\isacharunderscore}{\kern0pt}closed\ \isacommand{unfolding}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\ \ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Ordinals_In_MG.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%36=5%:%
%:%39=6%:%
%:%40=7%:%
%:%41=7%:%
%:%42=8%:%
%:%43=9%:%
%:%44=10%:%
%:%45=10%:%
%:%52=11%:%
%:%53=11%:%
%:%54=12%:%
%:%55=12%:%
%:%56=13%:%
%:%57=13%:%
%:%58=14%:%
%:%59=14%:%
%:%60=14%:%
%:%61=14%:%
%:%62=15%:%
%:%63=15%:%
%:%64=16%:%
%:%65=16%:%
%:%66=17%:%
%:%67=17%:%
%:%68=17%:%
%:%69=18%:%
%:%70=18%:%
%:%71=19%:%
%:%72=19%:%
%:%73=20%:%
%:%74=20%:%
%:%75=20%:%
%:%76=21%:%
%:%77=21%:%
%:%78=21%:%
%:%79=22%:%
%:%80=22%:%
%:%81=23%:%
%:%82=23%:%
%:%83=24%:%
%:%84=24%:%
%:%85=25%:%
%:%86=25%:%
%:%87=25%:%
%:%88=26%:%
%:%94=26%:%
%:%97=27%:%
%:%98=28%:%
%:%99=28%:%
%:%100=29%:%
%:%101=30%:%
%:%108=31%:%
%:%109=31%:%
%:%110=32%:%
%:%111=32%:%
%:%112=33%:%
%:%113=33%:%
%:%114=33%:%
%:%115=34%:%
%:%116=34%:%
%:%117=35%:%
%:%118=35%:%
%:%119=36%:%
%:%120=36%:%
%:%121=37%:%
%:%122=37%:%
%:%123=38%:%
%:%124=38%:%
%:%125=38%:%
%:%126=39%:%
%:%127=39%:%
%:%128=40%:%
%:%129=40%:%
%:%130=41%:%
%:%131=41%:%
%:%132=41%:%
%:%133=42%:%
%:%134=42%:%
%:%135=43%:%
%:%136=43%:%
%:%137=44%:%
%:%138=44%:%
%:%139=44%:%
%:%140=45%:%
%:%141=45%:%
%:%142=46%:%
%:%143=46%:%
%:%144=46%:%
%:%145=46%:%
%:%146=47%:%
%:%147=47%:%
%:%148=48%:%
%:%149=48%:%
%:%150=49%:%
%:%151=49%:%
%:%152=50%:%
%:%153=50%:%
%:%154=50%:%
%:%155=51%:%
%:%161=51%:%
%:%164=52%:%
%:%165=53%:%
%:%166=53%:%
%:%169=54%:%
%:%174=55%:%