### 2024-11-01 code/output/document/P_Names.tex

%
\begin{isabellebody}%
\setisabellecontext{P{\isacharunderscore}{\kern0pt}Names}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}Names\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\isanewline
\ \ \ \ Utilities{\isacharunderscore}{\kern0pt}M\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data\isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ HP{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}HP{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ X{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Pow{\isacharparenleft}{\kern0pt}X\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ transrec{\isadigit{2}}{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ HP{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ transrec{\isadigit{2}}{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ transrec{\isadigit{2}}{\isacharunderscore}{\kern0pt}succ\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ HP{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ transrec{\isadigit{2}}{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ HP{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}lim\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ transrec{\isadigit{2}}{\isacharunderscore}{\kern0pt}Limit\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isacommand{definition}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}names\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ M\ {\isachardot}{\kern0pt}\ {\isasymexists}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymmu}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}ord\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}elems{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ helper\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isakeyword{in}\ trans{\isacharunderscore}{\kern0pt}induct{\isadigit{3}}{\isacharunderscore}{\kern0pt}raw{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ a\ \isacommand{assume}\isamarkupfalse%
\ \isanewline
\ \ \ \ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}y{\isasymin}a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b{\isacharless}{\kern0pt}y{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isacharless}{\kern0pt}\ a\ {\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}lim\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isasymin}\ a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ c\ \isakeyword{where}\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}c\ {\isacharless}{\kern0pt}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}c\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ lt{\isacharunderscore}{\kern0pt}trans\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}c{\isacharless}{\kern0pt}a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isasymexists}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ \ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}elems{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ bp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \ \ \isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}works\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ ah\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ Q\ \isakeyword{where}\ {\isachardoublequoteopen}Q\ {\isasymequiv}\ {\isasymlambda}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ah\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}a\ \isakeyword{in}\ LeastI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ ah\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Q{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}def\ Q{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}elems{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}elems{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bp\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}namesI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}elems{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}works\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in\ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}elems{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ {\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ helper\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ \isanewline
\ \ \ \ \ \ \ \ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}a\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ trans{\isacharunderscore}{\kern0pt}induct{\isadigit{3}}{\isacharunderscore}{\kern0pt}raw{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}{\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ \ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ transM\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{4}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}b{\isacharless}{\kern0pt}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}{\isadigit{2}}\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ a\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}b{\isasymin}a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}c{\isacharless}{\kern0pt}b{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}b{\isacharless}{\kern0pt}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isacharless}{\kern0pt}\ a\ {\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}lim\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isasymin}\ a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ c\ \isakeyword{where}\ cp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}c\ {\isacharless}{\kern0pt}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}c\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ lt{\isacharunderscore}{\kern0pt}trans\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}b{\isacharless}{\kern0pt}a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ cp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ helper\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bp\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ oex{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}elems{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ bp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}range\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}lt\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{thm}\isamarkupfalse%
\ Least{\isacharunderscore}{\kern0pt}le\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}works\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}b\ {\isacharless}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ oex{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}c\ {\isacharless}{\kern0pt}\ b{\isachardot}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ \ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}elems{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ{\isacharparenright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ c\ \isakeyword{where}\ cp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}c\ {\isacharless}{\kern0pt}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ oex{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}c\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}lt\ cp\ bp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymle}\ c{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ Least{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ cp\ lt{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{3}}\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}induct\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymforall}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}x{\isasymin}P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isasymin}P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ R\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}R\ {\isasymequiv}\ {\isasymlambda}a{\isachardot}{\kern0pt}\ {\isasymforall}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}a{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}b\ {\isasymin}\ a{\isachardot}{\kern0pt}\ R{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ R{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ R{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}b{\isasymin}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymforall}z{\isasymin}P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ b\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}z\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ z\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}z\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ assms{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}R{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ eps{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ R{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ set{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}A\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}a{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}A{\isacharequal}{\kern0pt}{\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ blast\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ F\ \isakeyword{where}\ {\isachardoublequoteopen}F\ {\isasymequiv}\ {\isasymlambda}x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ Limit{\isacharparenleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ ap\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bp\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ex{\isacharunderscore}{\kern0pt}larger{\isacharunderscore}{\kern0pt}limit\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}c\ {\isacharless}{\kern0pt}\ b{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}lim\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}c\ {\isacharless}{\kern0pt}\ b{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}a\ \isakeyword{in}\ OUN{\isacharunderscore}{\kern0pt}RepFunI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ ap\ bp\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ Limit{\isacharparenleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}b\ \isakeyword{in}\ LeastI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ bp\ Limit{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ I\ \isakeyword{where}\ {\isachardoublequoteopen}I\ {\isasymequiv}\ {\isasymUnion}{\isacharbraceleft}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ ilim\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}I{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ I{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Limit{\isacharunderscore}{\kern0pt}Union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assm\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}\ x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymle}\ I{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ I{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}x\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}F\ \isakeyword{in}\ UN{\isacharunderscore}{\kern0pt}upper{\isacharunderscore}{\kern0pt}le{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Limit{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}I{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}{\isachardoublequoteopen}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ j{\isacharequal}{\kern0pt}I\ \isakeyword{in}\ leE{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ I{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}c\ {\isacharless}{\kern0pt}\ I{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ OUN{\isacharunderscore}{\kern0pt}RepFunI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}I{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}lim\ ilim\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}I\ \isakeyword{in}\ exI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ilim\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ M\ {\isasymand}\ x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ aH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ xsubset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xsubset\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}namesI\ aH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymand}\ x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}\ x\ {\isasymin}\ M\ {\isasymand}\ x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymtimes}\ P\ {\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ aH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ set{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ vin\ ypH\ aH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ypH\ vin\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharunderscore}{\kern0pt}namesI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Limit{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ aH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pow{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M\ {\isacharequal}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ auto\ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}zero\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pow{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M\ {\isacharequal}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Least{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}induct\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymforall}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isachardot}{\kern0pt}\ {\isasymforall}\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymLongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}y{\isasymin}P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}x{\isasymin}P{\isacharunderscore}{\kern0pt}names{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymlongrightarrow}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}lt\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ p\ \isacommand{assume}\isamarkupfalse%
\ p{\isadigit{3}}{\isadigit{0}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ p{\isadigit{3}}{\isadigit{0}}\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ p{\isadigit{3}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymlongrightarrow}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M\ {\isasymlongrightarrow}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isakeyword{in}\ eps{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
{\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}y{\isasymin}x{\isachardot}{\kern0pt}\ y\ {\isasymin}\ M\ {\isasymlongrightarrow}\ check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ one{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}check\ \isacommand{by}\isamarkupfalse%
\ assumption\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{define}\isamarkupfalse%
\ A\ \isakeyword{where}\ {\isachardoublequoteopen}A\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ asubset{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}A\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ subsetI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assm\ transM\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}A\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}x\ {\isasymnoteq}\ {\isadigit{0}}{\isacartoucheclose}\ \isacommand{unfolding}\isamarkupfalse%
\ A{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}\ a{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}\ z\ {\isasymin}\ A{\isachardot}{\kern0pt}\ z\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ set{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}set\ asubset\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ ap\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymforall}\ z\ {\isasymin}\ A{\isachardot}{\kern0pt}\ z\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}{\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ A{\isacharunderscore}{\kern0pt}def\ p{\isadigit{1}}\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}set{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}set{\isacharunderscore}{\kern0pt}succ\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}names{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ exI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ap\ Limit{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/P_Names.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=10%:%
%:%30=11%:%
%:%31=12%:%
%:%32=13%:%
%:%33=13%:%
%:%34=14%:%
%:%35=15%:%
%:%36=16%:%
%:%37=16%:%
%:%39=16%:%
%:%43=16%:%
%:%44=16%:%
%:%45=16%:%
%:%46=16%:%
%:%53=16%:%
%:%54=17%:%
%:%55=18%:%
%:%56=18%:%
%:%57=19%:%
%:%60=20%:%
%:%64=20%:%
%:%65=20%:%
%:%66=21%:%
%:%67=21%:%
%:%68=22%:%
%:%69=22%:%
%:%70=22%:%
%:%71=22%:%
%:%72=22%:%
%:%73=23%:%
%:%74=23%:%
%:%75=23%:%
%:%76=23%:%
%:%77=23%:%
%:%78=24%:%
%:%84=24%:%
%:%87=25%:%
%:%88=26%:%
%:%89=26%:%
%:%90=27%:%
%:%93=28%:%
%:%97=28%:%
%:%98=28%:%
%:%99=28%:%
%:%100=28%:%
%:%105=28%:%
%:%108=29%:%
%:%109=30%:%
%:%110=30%:%
%:%111=31%:%
%:%112=32%:%
%:%113=32%:%
%:%114=33%:%
%:%115=34%:%
%:%116=35%:%
%:%117=35%:%
%:%120=36%:%
%:%124=36%:%
%:%125=36%:%
%:%126=36%:%
%:%131=36%:%
%:%134=37%:%
%:%135=38%:%
%:%136=38%:%
%:%137=39%:%
%:%144=40%:%
%:%145=40%:%
%:%146=41%:%
%:%147=41%:%
%:%148=42%:%
%:%149=42%:%
%:%150=42%:%
%:%151=43%:%
%:%152=43%:%
%:%153=44%:%
%:%154=45%:%
%:%155=45%:%
%:%156=46%:%
%:%157=46%:%
%:%158=46%:%
%:%159=47%:%
%:%160=47%:%
%:%161=48%:%
%:%162=48%:%
%:%163=49%:%
%:%164=49%:%
%:%165=49%:%
%:%166=50%:%
%:%167=51%:%
%:%168=52%:%
%:%169=53%:%
%:%170=53%:%
%:%171=53%:%
%:%172=53%:%
%:%173=53%:%
%:%174=54%:%
%:%175=54%:%
%:%176=54%:%
%:%177=54%:%
%:%178=55%:%
%:%179=55%:%
%:%180=55%:%
%:%181=55%:%
%:%182=55%:%
%:%183=56%:%
%:%184=56%:%
%:%185=56%:%
%:%186=56%:%
%:%187=56%:%
%:%188=57%:%
%:%189=57%:%
%:%190=57%:%
%:%191=57%:%
%:%192=57%:%
%:%193=58%:%
%:%194=58%:%
%:%195=58%:%
%:%196=58%:%
%:%197=58%:%
%:%198=59%:%
%:%199=59%:%
%:%200=60%:%
%:%201=60%:%
%:%202=60%:%
%:%203=60%:%
%:%204=60%:%
%:%205=61%:%
%:%211=61%:%
%:%214=62%:%
%:%215=63%:%
%:%216=63%:%
%:%223=64%:%
%:%224=64%:%
%:%225=65%:%
%:%226=65%:%
%:%227=66%:%
%:%228=66%:%
%:%229=66%:%
%:%230=66%:%
%:%231=66%:%
%:%232=67%:%
%:%233=67%:%
%:%234=67%:%
%:%235=67%:%
%:%236=67%:%
%:%237=68%:%
%:%238=68%:%
%:%239=68%:%
%:%240=68%:%
%:%241=68%:%
%:%242=69%:%
%:%243=69%:%
%:%244=69%:%
%:%245=69%:%
%:%246=69%:%
%:%247=70%:%
%:%253=70%:%
%:%256=71%:%
%:%257=72%:%
%:%258=72%:%
%:%259=73%:%
%:%266=74%:%
%:%267=74%:%
%:%268=75%:%
%:%269=75%:%
%:%270=76%:%
%:%271=76%:%
%:%272=76%:%
%:%273=76%:%
%:%274=76%:%
%:%275=77%:%
%:%276=78%:%
%:%277=78%:%
%:%278=79%:%
%:%279=79%:%
%:%280=79%:%
%:%281=79%:%
%:%282=79%:%
%:%283=80%:%
%:%284=80%:%
%:%285=80%:%
%:%286=81%:%
%:%287=81%:%
%:%288=81%:%
%:%289=81%:%
%:%290=82%:%
%:%291=82%:%
%:%292=82%:%
%:%293=82%:%
%:%294=82%:%
%:%295=83%:%
%:%296=83%:%
%:%297=83%:%
%:%298=83%:%
%:%299=83%:%
%:%300=84%:%
%:%306=84%:%
%:%309=85%:%
%:%310=86%:%
%:%311=86%:%
%:%318=87%:%
%:%319=87%:%
%:%320=88%:%
%:%321=88%:%
%:%322=89%:%
%:%323=89%:%
%:%324=89%:%
%:%325=89%:%
%:%326=89%:%
%:%327=90%:%
%:%328=90%:%
%:%329=90%:%
%:%330=90%:%
%:%331=91%:%
%:%332=91%:%
%:%333=91%:%
%:%334=91%:%
%:%335=91%:%
%:%336=92%:%
%:%337=92%:%
%:%338=92%:%
%:%339=92%:%
%:%340=92%:%
%:%341=93%:%
%:%342=93%:%
%:%343=93%:%
%:%344=93%:%
%:%345=94%:%
%:%351=94%:%
%:%354=95%:%
%:%355=96%:%
%:%356=96%:%
%:%359=97%:%
%:%363=97%:%
%:%364=97%:%
%:%365=97%:%
%:%366=97%:%
%:%371=97%:%
%:%374=98%:%
%:%375=99%:%
%:%376=99%:%
%:%383=100%:%
%:%384=100%:%
%:%385=101%:%
%:%386=101%:%
%:%387=102%:%
%:%388=102%:%
%:%389=102%:%
%:%390=102%:%
%:%391=102%:%
%:%392=103%:%
%:%393=103%:%
%:%394=103%:%
%:%395=104%:%
%:%396=104%:%
%:%397=104%:%
%:%398=105%:%
%:%404=105%:%
%:%407=106%:%
%:%408=107%:%
%:%409=107%:%
%:%412=108%:%
%:%416=108%:%
%:%417=108%:%
%:%418=108%:%
%:%419=108%:%
%:%424=108%:%
%:%427=109%:%
%:%428=110%:%
%:%429=110%:%
%:%436=111%:%
%:%437=111%:%
%:%438=112%:%
%:%439=112%:%
%:%440=113%:%
%:%441=113%:%
%:%442=113%:%
%:%443=113%:%
%:%444=113%:%
%:%445=114%:%
%:%446=114%:%
%:%447=114%:%
%:%448=114%:%
%:%449=114%:%
%:%450=115%:%
%:%456=115%:%
%:%459=116%:%
%:%460=117%:%
%:%461=117%:%
%:%462=118%:%
%:%469=119%:%
%:%470=119%:%
%:%471=120%:%
%:%472=120%:%
%:%473=121%:%
%:%474=121%:%
%:%475=122%:%
%:%476=123%:%
%:%477=123%:%
%:%478=124%:%
%:%479=124%:%
%:%480=124%:%
%:%481=125%:%
%:%482=125%:%
%:%483=126%:%
%:%484=126%:%
%:%485=127%:%
%:%486=127%:%
%:%487=128%:%
%:%488=128%:%
%:%489=128%:%
%:%490=128%:%
%:%491=128%:%
%:%492=129%:%
%:%493=129%:%
%:%494=129%:%
%:%495=129%:%
%:%496=130%:%
%:%497=130%:%
%:%498=130%:%
%:%499=130%:%
%:%500=130%:%
%:%501=131%:%
%:%502=131%:%
%:%503=131%:%
%:%504=131%:%
%:%505=132%:%
%:%506=132%:%
%:%507=132%:%
%:%508=132%:%
%:%509=133%:%
%:%510=133%:%
%:%511=133%:%
%:%512=133%:%
%:%513=133%:%
%:%514=134%:%
%:%515=134%:%
%:%516=134%:%
%:%517=134%:%
%:%518=134%:%
%:%519=135%:%
%:%520=135%:%
%:%521=135%:%
%:%522=136%:%
%:%523=136%:%
%:%524=136%:%
%:%525=137%:%
%:%526=137%:%
%:%527=138%:%
%:%528=138%:%
%:%529=138%:%
%:%530=139%:%
%:%531=140%:%
%:%532=140%:%
%:%533=141%:%
%:%534=141%:%
%:%535=142%:%
%:%536=142%:%
%:%537=143%:%
%:%538=143%:%
%:%539=143%:%
%:%540=143%:%
%:%541=143%:%
%:%542=144%:%
%:%543=144%:%
%:%544=144%:%
%:%545=144%:%
%:%546=145%:%
%:%547=145%:%
%:%548=145%:%
%:%549=145%:%
%:%550=145%:%
%:%551=146%:%
%:%552=146%:%
%:%553=146%:%
%:%554=146%:%
%:%555=146%:%
%:%556=147%:%
%:%557=147%:%
%:%558=147%:%
%:%559=147%:%
%:%560=147%:%
%:%561=148%:%
%:%562=148%:%
%:%563=148%:%
%:%564=148%:%
%:%565=148%:%
%:%566=149%:%
%:%567=149%:%
%:%568=150%:%
%:%569=150%:%
%:%570=151%:%
%:%571=151%:%
%:%572=152%:%
%:%573=152%:%
%:%574=152%:%
%:%575=152%:%
%:%576=152%:%
%:%577=153%:%
%:%583=153%:%
%:%586=154%:%
%:%587=155%:%
%:%588=155%:%
%:%589=156%:%
%:%596=157%:%
%:%597=157%:%
%:%598=158%:%
%:%599=158%:%
%:%600=159%:%
%:%601=159%:%
%:%602=159%:%
%:%603=159%:%
%:%604=159%:%
%:%605=160%:%
%:%606=160%:%
%:%607=160%:%
%:%608=160%:%
%:%609=160%:%
%:%610=161%:%
%:%611=161%:%
%:%612=161%:%
%:%613=161%:%
%:%614=161%:%
%:%615=162%:%
%:%616=162%:%
%:%617=162%:%
%:%618=162%:%
%:%619=162%:%
%:%620=163%:%
%:%621=163%:%
%:%622=163%:%
%:%623=163%:%
%:%624=163%:%
%:%625=163%:%
%:%626=164%:%
%:%632=164%:%
%:%635=165%:%
%:%636=166%:%
%:%637=166%:%
%:%638=167%:%
%:%645=168%:%
%:%646=168%:%
%:%647=169%:%
%:%648=169%:%
%:%649=170%:%
%:%650=170%:%
%:%651=170%:%
%:%652=170%:%
%:%653=171%:%
%:%654=171%:%
%:%655=171%:%
%:%656=171%:%
%:%657=171%:%
%:%658=172%:%
%:%664=172%:%
%:%667=173%:%
%:%668=174%:%
%:%669=174%:%
%:%670=175%:%
%:%677=176%:%
%:%678=176%:%
%:%679=177%:%
%:%680=177%:%
%:%681=178%:%
%:%682=178%:%
%:%683=178%:%
%:%684=178%:%
%:%685=178%:%
%:%686=179%:%
%:%687=179%:%
%:%688=179%:%
%:%689=179%:%
%:%690=179%:%
%:%691=180%:%
%:%697=180%:%
%:%700=181%:%
%:%701=182%:%
%:%702=182%:%
%:%703=183%:%
%:%704=184%:%
%:%705=184%:%
%:%712=185%:%
%:%713=185%:%
%:%714=186%:%
%:%715=186%:%
%:%716=187%:%
%:%717=187%:%
%:%718=187%:%
%:%719=187%:%
%:%720=187%:%
%:%721=188%:%
%:%722=188%:%
%:%723=188%:%
%:%724=188%:%
%:%725=189%:%
%:%726=189%:%
%:%727=189%:%
%:%728=190%:%
%:%729=190%:%
%:%730=191%:%
%:%731=191%:%
%:%732=191%:%
%:%733=192%:%
%:%734=192%:%
%:%735=192%:%
%:%736=192%:%
%:%737=192%:%
%:%738=193%:%
%:%739=193%:%
%:%740=193%:%
%:%741=194%:%
%:%742=194%:%
%:%743=194%:%
%:%744=195%:%
%:%745=195%:%
%:%746=195%:%
%:%747=195%:%
%:%748=195%:%
%:%749=196%:%
%:%750=196%:%
%:%751=196%:%
%:%752=196%:%
%:%753=197%:%
%:%754=197%:%
%:%755=198%:%
%:%756=198%:%
%:%757=199%:%
%:%758=199%:%
%:%759=200%:%
%:%760=200%:%
%:%761=200%:%
%:%762=201%:%
%:%763=201%:%
%:%764=201%:%
%:%765=201%:%
%:%766=201%:%
%:%767=202%:%
%:%773=202%:%
%:%776=203%:%
%:%777=204%:%
%:%778=204%:%
%:%779=205%:%
%:%780=206%:%
%:%787=207%:%
%:%788=207%:%
%:%789=208%:%
%:%790=208%:%
%:%791=209%:%
%:%792=209%:%
%:%793=210%:%
%:%794=210%:%
%:%795=211%:%
%:%796=211%:%
%:%797=212%:%
%:%798=212%:%
%:%799=213%:%
%:%800=213%:%
%:%801=214%:%
%:%802=214%:%
%:%803=215%:%
%:%804=215%:%
%:%805=216%:%
%:%806=216%:%
%:%807=217%:%
%:%808=217%:%
%:%809=217%:%
%:%810=218%:%
%:%811=218%:%
%:%812=219%:%
%:%813=219%:%
%:%814=219%:%
%:%815=220%:%
%:%816=220%:%
%:%817=220%:%
%:%818=220%:%
%:%819=220%:%
%:%820=221%:%
%:%821=221%:%
%:%822=221%:%
%:%823=221%:%
%:%824=221%:%
%:%825=222%:%
%:%826=222%:%
%:%827=223%:%
%:%828=223%:%
%:%829=223%:%
%:%830=223%:%
%:%831=223%:%
%:%832=224%:%
%:%833=224%:%
%:%834=225%:%
%:%835=225%:%
%:%836=225%:%
%:%837=226%:%
%:%838=226%:%
%:%839=226%:%
%:%840=227%:%
%:%841=227%:%
%:%842=227%:%
%:%843=227%:%
%:%844=227%:%
%:%845=227%:%
%:%846=228%:%
%:%852=228%:%
%:%855=229%:%
%:%856=230%:%
%:%857=230%:%
%:%858=231%:%
%:%861=232%:%
%:%865=232%:%
%:%866=232%:%
%:%867=232%:%
%:%868=233%:%
%:%869=233%:%
%:%870=234%:%
%:%871=234%:%
%:%872=235%:%
%:%873=235%:%
%:%874=236%:%
%:%875=236%:%
%:%876=236%:%
%:%877=237%:%
%:%878=237%:%
%:%879=238%:%
%:%880=238%:%
%:%881=238%:%
%:%882=239%:%
%:%883=239%:%
%:%884=239%:%
%:%885=239%:%
%:%886=239%:%
%:%887=240%:%
%:%888=240%:%
%:%889=240%:%
%:%890=240%:%
%:%891=240%:%
%:%892=241%:%
%:%893=241%:%
%:%894=241%:%
%:%895=241%:%
%:%896=241%:%
%:%897=242%:%
%:%898=242%:%
%:%899=242%:%
%:%900=242%:%
%:%901=242%:%
%:%902=243%:%
%:%903=243%:%
%:%904=244%:%
%:%905=244%:%
%:%906=244%:%
%:%907=244%:%
%:%908=245%:%
%:%909=245%:%
%:%910=245%:%
%:%911=245%:%
%:%912=245%:%
%:%913=246%:%
%:%914=246%:%
%:%915=246%:%
%:%916=247%:%
%:%917=247%:%
%:%918=248%:%
%:%919=248%:%
%:%920=249%:%
%:%921=249%:%
%:%922=249%:%
%:%923=250%:%
%:%924=250%:%
%:%925=251%:%
%:%926=251%:%
%:%927=252%:%
%:%928=252%:%
%:%929=252%:%
%:%930=253%:%
%:%931=253%:%
%:%932=254%:%
%:%933=254%:%
%:%934=255%:%
%:%935=255%:%
%:%936=255%:%
%:%937=256%:%
%:%938=256%:%
%:%939=256%:%
%:%940=257%:%
%:%941=257%:%
%:%942=258%:%
%:%943=258%:%
%:%944=259%:%
%:%945=259%:%
%:%946=260%:%
%:%947=260%:%
%:%948=260%:%
%:%949=261%:%
%:%950=261%:%
%:%951=261%:%
%:%952=262%:%
%:%953=262%:%
%:%954=263%:%
%:%955=263%:%
%:%956=264%:%
%:%957=264%:%
%:%958=265%:%
%:%959=265%:%
%:%960=265%:%
%:%961=266%:%
%:%962=266%:%
%:%963=266%:%
%:%964=267%:%
%:%965=267%:%
%:%966=268%:%
%:%967=268%:%
%:%968=268%:%
%:%969=269%:%
%:%970=269%:%
%:%971=269%:%
%:%972=269%:%
%:%973=269%:%
%:%974=270%:%
%:%975=270%:%
%:%976=271%:%
%:%977=271%:%
%:%978=271%:%
%:%979=272%:%
%:%980=272%:%
%:%981=272%:%
%:%982=272%:%
%:%983=272%:%
%:%984=273%:%
%:%985=273%:%
%:%986=274%:%
%:%987=274%:%
%:%988=274%:%
%:%989=275%:%
%:%990=275%:%
%:%991=276%:%
%:%992=276%:%
%:%993=276%:%
%:%994=277%:%
%:%1000=277%:%
%:%1003=278%:%
%:%1004=279%:%
%:%1005=279%:%
%:%1012=280%:%
%:%1013=280%:%
%:%1014=281%:%
%:%1015=281%:%
%:%1016=282%:%
%:%1017=282%:%
%:%1018=282%:%
%:%1019=283%:%
%:%1020=283%:%
%:%1021=284%:%
%:%1022=284%:%
%:%1023=285%:%
%:%1024=285%:%
%:%1025=285%:%
%:%1026=286%:%
%:%1027=286%:%
%:%1028=287%:%
%:%1029=287%:%
%:%1030=288%:%
%:%1031=288%:%
%:%1032=289%:%
%:%1033=289%:%
%:%1034=290%:%
%:%1035=290%:%
%:%1036=290%:%
%:%1037=291%:%
%:%1038=291%:%
%:%1039=291%:%
%:%1040=291%:%
%:%1041=291%:%
%:%1042=292%:%
%:%1043=292%:%
%:%1044=292%:%
%:%1045=292%:%
%:%1046=292%:%
%:%1047=293%:%
%:%1048=293%:%
%:%1049=293%:%
%:%1050=293%:%
%:%1051=293%:%
%:%1052=294%:%
%:%1053=294%:%
%:%1054=295%:%
%:%1055=295%:%
%:%1056=295%:%
%:%1057=295%:%
%:%1058=295%:%
%:%1059=296%:%
%:%1060=296%:%
%:%1061=297%:%
%:%1062=297%:%
%:%1063=298%:%
%:%1064=298%:%
%:%1065=298%:%
%:%1066=298%:%
%:%1067=299%:%
%:%1068=299%:%
%:%1069=299%:%
%:%1070=300%:%
%:%1071=300%:%
%:%1072=300%:%
%:%1073=301%:%
%:%1074=301%:%
%:%1075=302%:%
%:%1076=302%:%
%:%1077=303%:%
%:%1078=303%:%
%:%1079=303%:%
%:%1080=304%:%
%:%1081=304%:%
%:%1082=304%:%
%:%1083=304%:%
%:%1084=304%:%
%:%1085=305%:%
%:%1086=305%:%
%:%1087=305%:%
%:%1088=305%:%
%:%1089=305%:%
%:%1090=306%:%
%:%1091=306%:%
%:%1092=306%:%
%:%1093=306%:%
%:%1094=306%:%
%:%1095=307%:%
%:%1096=307%:%
%:%1097=308%:%
%:%1098=308%:%
%:%1099=308%:%
%:%1100=308%:%
%:%1101=308%:%
%:%1102=309%:%
%:%1103=309%:%
%:%1104=309%:%
%:%1105=310%:%
%:%1106=310%:%
%:%1107=311%:%
%:%1108=311%:%
%:%1109=312%:%
%:%1110=312%:%
%:%1111=313%:%
%:%1117=313%:%
%:%1120=314%:%
%:%1121=315%:%
%:%1122=315%:%
%:%1129=316%:%
%:%1130=316%:%
%:%1131=317%:%
%:%1132=317%:%
%:%1133=317%:%
%:%1134=317%:%
%:%1135=318%:%
%:%1136=318%:%
%:%1137=318%:%
%:%1138=318%:%
%:%1139=319%:%
%:%1140=319%:%
%:%1141=319%:%
%:%1142=319%:%
%:%1143=319%:%
%:%1144=320%:%
%:%1145=320%:%
%:%1146=320%:%
%:%1147=321%:%
%:%1148=321%:%
%:%1149=321%:%
%:%1150=322%:%
%:%1151=322%:%
%:%1152=322%:%
%:%1153=323%:%
%:%1159=323%:%
%:%1162=324%:%
%:%1163=325%:%
%:%1164=325%:%
%:%1171=326%:%
%:%1172=326%:%
%:%1173=327%:%
%:%1174=327%:%
%:%1175=327%:%
%:%1176=327%:%
%:%1177=328%:%
%:%1178=328%:%
%:%1179=328%:%
%:%1180=328%:%
%:%1181=329%:%
%:%1182=329%:%
%:%1183=329%:%
%:%1184=329%:%
%:%1185=329%:%
%:%1186=330%:%
%:%1187=330%:%
%:%1188=330%:%
%:%1189=331%:%
%:%1190=331%:%
%:%1191=332%:%
%:%1192=332%:%
%:%1193=333%:%
%:%1194=333%:%
%:%1195=334%:%
%:%1201=334%:%
%:%1204=335%:%
%:%1205=336%:%
%:%1206=336%:%
%:%1207=337%:%
%:%1210=338%:%
%:%1214=338%:%
%:%1215=338%:%
%:%1216=339%:%
%:%1217=339%:%
%:%1218=340%:%
%:%1219=340%:%
%:%1220=340%:%
%:%1221=341%:%
%:%1222=342%:%
%:%1223=343%:%
%:%1224=343%:%
%:%1225=343%:%
%:%1226=343%:%
%:%1227=344%:%
%:%1228=344%:%
%:%1229=345%:%
%:%1230=345%:%
%:%1231=345%:%
%:%1232=346%:%
%:%1233=347%:%
%:%1234=347%:%
%:%1235=347%:%
%:%1236=348%:%
%:%1237=348%:%
%:%1238=349%:%
%:%1239=349%:%
%:%1240=349%:%
%:%1241=350%:%
%:%1242=350%:%
%:%1243=350%:%
%:%1244=350%:%
%:%1245=350%:%
%:%1246=351%:%
%:%1247=351%:%
%:%1248=351%:%
%:%1249=351%:%
%:%1250=351%:%
%:%1251=352%:%
%:%1252=352%:%
%:%1253=352%:%
%:%1254=352%:%
%:%1255=352%:%
%:%1256=353%:%
%:%1257=353%:%
%:%1258=354%:%
%:%1259=354%:%
%:%1260=354%:%
%:%1261=354%:%
%:%1262=354%:%
%:%1263=355%:%
%:%1269=355%:%
%:%1272=356%:%
%:%1273=357%:%
%:%1274=358%:%
%:%1275=358%:%
%:%1278=359%:%
%:%1282=359%:%
%:%1283=359%:%
%:%1284=360%:%
%:%1285=360%:%
%:%1286=361%:%
%:%1287=361%:%
%:%1288=362%:%
%:%1289=362%:%
%:%1290=362%:%
%:%1291=363%:%
%:%1292=363%:%
%:%1293=363%:%
%:%1294=363%:%
%:%1295=364%:%
%:%1296=364%:%
%:%1297=364%:%
%:%1298=365%:%
%:%1299=365%:%
%:%1300=366%:%
%:%1301=366%:%
%:%1302=367%:%
%:%1303=367%:%
%:%1304=367%:%
%:%1305=367%:%
%:%1306=367%:%
%:%1307=368%:%
%:%1308=368%:%
%:%1309=368%:%
%:%1310=368%:%
%:%1311=368%:%
%:%1312=369%:%
%:%1313=369%:%
%:%1314=370%:%
%:%1315=370%:%
%:%1316=371%:%
%:%1317=371%:%
%:%1318=372%:%
%:%1319=372%:%
%:%1320=372%:%
%:%1321=373%:%
%:%1322=373%:%
%:%1323=374%:%
%:%1324=374%:%
%:%1325=374%:%
%:%1326=375%:%
%:%1327=375%:%
%:%1328=375%:%
%:%1329=375%:%
%:%1330=375%:%
%:%1331=376%:%
%:%1332=376%:%
%:%1333=376%:%
%:%1334=377%:%
%:%1335=377%:%
%:%1336=377%:%
%:%1337=378%:%
%:%1338=378%:%
%:%1339=378%:%
%:%1340=378%:%
%:%1341=379%:%
%:%1342=379%:%
%:%1343=379%:%
%:%1344=380%:%
%:%1345=380%:%
%:%1346=380%:%
%:%1347=381%:%
%:%1348=381%:%
%:%1349=381%:%
%:%1350=382%:%
%:%1351=382%:%
%:%1352=382%:%
%:%1353=383%:%
%:%1354=383%:%
%:%1355=383%:%
%:%1356=384%:%
%:%1357=384%:%
%:%1358=384%:%
%:%1359=385%:%
%:%1360=385%:%
%:%1361=385%:%
%:%1362=386%:%
%:%1363=386%:%
%:%1364=386%:%
%:%1365=386%:%
%:%1366=387%:%
%:%1367=387%:%
%:%1368=388%:%
%:%1369=388%:%
%:%1370=389%:%
%:%1371=389%:%
%:%1372=389%:%
%:%1373=390%:%
%:%1374=390%:%
%:%1375=391%:%
%:%1381=391%:%
%:%1384=392%:%
%:%1385=393%:%
%:%1386=393%:%
%:%1393=394%:%