### 2024-11-01 code/output/document/Pairing_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Pairing{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Axiom of Pairing in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Pairing{\isacharunderscore}{\kern0pt}Axiom\ \isakeyword{imports}\ Names\ \isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}Upair\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}one\ {\isasymin}\ G\ {\isasymLongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isasymtau}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharcomma}{\kern0pt}\ rule\ trans{\isacharcomma}{\kern0pt}\ subst\ def{\isacharunderscore}{\kern0pt}val{\isacharcomma}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ Sep{\isacharunderscore}{\kern0pt}and{\isacharunderscore}{\kern0pt}Replace{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pairing{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}upair{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ y\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}one{\isasymin}G{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}G{\isasymsubseteq}P{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isakeyword{and}\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}one{\isasymin}G{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}one{\isasymin}P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ subsetD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}one{\isasymin}M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtau}\ {\isasymrho}\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isadigit{0}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymrho}\ {\isasymin}\ M{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}one{\isasymin}M{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtau}{\isacharcomma}{\kern0pt}one{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isasymin}M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isasymtau}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}one{\isasymrangle}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymsigma}\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ GenExtI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}Upair\ assms\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{0}}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ upair{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Pairing_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%33=2%:%
%:%36=3%:%
%:%37=4%:%
%:%38=4%:%
%:%39=5%:%
%:%40=6%:%
%:%41=7%:%
%:%42=7%:%
%:%43=8%:%
%:%46=9%:%
%:%50=9%:%
%:%51=9%:%
%:%56=9%:%
%:%59=10%:%
%:%60=11%:%
%:%61=11%:%
%:%62=12%:%
%:%63=13%:%
%:%70=14%:%
%:%71=14%:%
%:%72=15%:%
%:%73=15%:%
%:%74=16%:%
%:%75=16%:%
%:%76=17%:%
%:%77=17%:%
%:%78=17%:%
%:%79=17%:%
%:%80=18%:%
%:%81=18%:%
%:%82=19%:%
%:%83=19%:%
%:%84=19%:%
%:%85=19%:%
%:%86=20%:%
%:%87=20%:%
%:%88=21%:%
%:%89=21%:%
%:%90=21%:%
%:%91=21%:%
%:%92=22%:%
%:%93=22%:%
%:%94=23%:%
%:%95=23%:%
%:%96=23%:%
%:%97=23%:%
%:%98=24%:%
%:%99=24%:%
%:%100=25%:%
%:%101=25%:%
%:%102=26%:%
%:%103=26%:%
%:%104=27%:%
%:%105=28%:%
%:%106=28%:%
%:%107=28%:%
%:%108=29%:%
%:%109=29%:%
%:%110=30%:%
%:%111=30%:%
%:%112=30%:%
%:%113=30%:%
%:%114=31%:%
%:%115=31%:%
%:%116=32%:%
%:%117=32%:%
%:%118=32%:%
%:%119=32%:%
%:%120=33%:%
%:%121=33%:%
%:%122=34%:%
%:%123=34%:%
%:%124=34%:%
%:%125=34%:%
%:%126=35%:%
%:%127=35%:%
%:%128=36%:%
%:%129=36%:%
%:%130=36%:%
%:%131=36%:%
%:%132=37%:%
%:%133=37%:%
%:%134=38%:%
%:%135=38%:%
%:%136=38%:%
%:%137=39%:%
%:%138=39%:%
%:%139=40%:%
%:%140=40%:%
%:%141=40%:%
%:%142=40%:%
%:%143=40%:%
%:%144=41%:%
%:%150=41%:%
%:%153=42%:%
%:%154=43%:%
%:%155=43%:%
%:%162=44%:%