### 2024-11-01 code/output/document/Powerset_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Powerset{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Powerset Axiom in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Powerset{\isacharunderscore}{\kern0pt}Axiom\isanewline
\ \ \isakeyword{imports}\ Renaming{\isacharunderscore}{\kern0pt}Auto\ Separation{\isacharunderscore}{\kern0pt}Axiom\ Pairing{\isacharunderscore}{\kern0pt}Axiom\ Union{\isacharunderscore}{\kern0pt}Axiom\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{simple{\isacharunderscore}{\kern0pt}rename}\isamarkupfalse%
\ {\isachardoublequoteopen}perm{\isacharunderscore}{\kern0pt}pow{\isachardoublequoteclose}\ \isakeyword{src}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}ss{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}fs{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{tgt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}fs{\isacharcomma}{\kern0pt}ss{\isacharcomma}{\kern0pt}sp{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}inter{\isacharunderscore}{\kern0pt}Transset{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}b\ {\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}b\ {\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{context}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\ \ \isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ name{\isacharunderscore}{\kern0pt}components{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}{\isasymin}{\isasymtheta}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymtheta}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ \ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}{\isacharless}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Pair{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}a{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}B{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}l{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}o{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymchi}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}sq\ {\isasymin}A{\isasymtimes}B\ {\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}sq{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}fst{\isacharparenleft}{\kern0pt}sq{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{6}}{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{7}}{\isasymin}nat{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymphi}{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{6}}{\isacharbackquote}{\kern0pt}{\isadigit{7}}{\isacharbackquote}{\kern0pt}perm{\isacharunderscore}{\kern0pt}pow{\isacharunderscore}{\kern0pt}fn{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}B{\isasymin}M{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}A{\isasymtimes}B\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}\ formula{\isacartoucheclose}\ {\isacartoucheopen}{\isadigit{6}}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isadigit{7}}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharprime}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{7}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ perm{\isacharunderscore}{\kern0pt}pow{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ perm{\isacharunderscore}{\kern0pt}pow{\isacharunderscore}{\kern0pt}thm\ \ arity{\isacharunderscore}{\kern0pt}ren\ ren{\isacharunderscore}{\kern0pt}tc\ Nil{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharprime}{\kern0pt}\ {\isasymin}\ formula{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{5}}{\isachardoublequoteclose}\ \ \ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{5}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ pred{\isacharunderscore}{\kern0pt}le\ arity{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ sp\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}A{\isasymtimes}B\ {\isasymin}\ M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}sp\ {\isasymin}\ A{\isasymtimes}B{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}fst{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}snd{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}\ {\isasymin}\ B{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ fst{\isacharunderscore}{\kern0pt}type\ snd{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sp\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}fst{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}snd{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}B{\isasymin}M{\isacartoucheclose}\ transitivity\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ inM\ {\isacharequal}{\kern0pt}\ {\isacartoucheopen}A{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}B{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}l{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}o{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isasymchi}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \ \ {\isacartoucheopen}sp{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}fst{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}snd{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\ {\isacartoucheopen}sp\ {\isasymin}\ M{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharprime}{\kern0pt}\ {\isasymin}\ formula{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}sp{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}\ {\isasymlongleftrightarrow}\ M{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}sp{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}env{\isadigit{0}}{\isacharat}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isasymTurnstile}{\isacharunderscore}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}{\isasympsi}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}p{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ M\ {\isacharquery}{\kern0pt}env{\isadigit{0}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ inM\ {\isacartoucheopen}sp\ {\isasymin}\ A{\isasymtimes}B{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}sp{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ inM\ {\isacartoucheopen}{\isasymphi}\ {\isasymin}\ formula{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}fst{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ren{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isadigit{6}}\ {\isadigit{7}}\ {\isacharquery}{\kern0pt}env{\isadigit{2}}\ M\ {\isacharquery}{\kern0pt}env{\isadigit{1}}\ perm{\isacharunderscore}{\kern0pt}pow{\isacharunderscore}{\kern0pt}fn{\isacharbrackright}{\kern0pt}\ perm{\isacharunderscore}{\kern0pt}pow{\isacharunderscore}{\kern0pt}thm\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ perm{\isacharunderscore}{\kern0pt}pow{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}sp{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharcomma}{\kern0pt}p{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}fst{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}sp{\isasymin}A{\isasymtimes}B\ {\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}sp{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}l{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharcomma}{\kern0pt}p{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\ {\isacartoucheopen}A{\isasymtimes}B{\isasymin}M{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymphi}{\isacharprime}{\kern0pt}\ {\isasymin}\ formula{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasympsi}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ assms\ {\isacartoucheopen}A{\isasymtimes}B{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ A{\isasymtimes}B\ {\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ l{\isacharcomma}{\kern0pt}\ o{\isacharcomma}{\kern0pt}\ {\isasymchi}{\isacharcomma}{\kern0pt}\ p{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}ax\ separation{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pow{\isacharunderscore}{\kern0pt}inter{\isacharunderscore}{\kern0pt}MG{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtau}\ \isakeyword{where}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pow{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtau}{\isasymin}M{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ cartprod{\isacharunderscore}{\kern0pt}closed\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}Q\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ power{\isacharunderscore}{\kern0pt}ax\ {\isacartoucheopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P\ {\isasymin}\ M{\isacartoucheclose}\ \isacommand{obtain}\isamarkupfalse%
\ Q\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}powerset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isacharcomma}{\kern0pt}Q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Q\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ power{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}Q\ {\isasymLongrightarrow}\ z{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isasymin}Pow{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ a{\isasymin}M{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P\ {\isasymin}\ M{\isacartoucheclose}\ powerset{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isachardoublequoteclose}\ Q{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}Q{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}Q{\isasymin}M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}{\isasympi}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}Q{\isasymtimes}{\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\isanewline
\ \ \ \ {\isacharquery}{\kern0pt}b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}Q{\isasymin}M{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasympi}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ transitivity\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasympi}{\isasymin}M{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}b\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymsubseteq}\ {\isacharquery}{\kern0pt}b{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ c\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}c\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymchi}\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymchi}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymtheta}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}sp\ {\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isasymtimes}P\ {\isachardot}{\kern0pt}\ snd{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}sp{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{6}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}at\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isacartoucheclose}\ {\isacartoucheopen}{\isasymchi}\ {\isasymin}\ M{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ sats{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}\ {\isasymin}\ {\isacharquery}{\kern0pt}Q{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharquery}{\kern0pt}b{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ generic\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}elem\ {\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}{\isasymtheta}\ one\ {\isacharquery}{\kern0pt}{\isasympi}\ G{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}intro\ equalityI\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymsigma}\ p\ \isakeyword{where}\isanewline
\ \ \ \ \ \ \ \ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}{\isasymin}{\isacharquery}{\kern0pt}{\isasymtheta}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharless}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}{\isasymin}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymtheta}\ {\isasymin}\ M{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ name{\isacharunderscore}{\kern0pt}components{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}\ {\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}c{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymchi}\ {\isasymin}\ M{\isacartoucheclose}\ generic\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ x{\isacartoucheclose}\ {\isacartoucheopen}{\isasymsigma}{\isasymin}M{\isacartoucheclose}\ \ {\isacartoucheopen}{\isasymchi}{\isasymin}M{\isacartoucheclose}\ GenExtI\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}c{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ c{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}c\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtau}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymsigma}\ \isakeyword{where}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}c{\isacartoucheclose}\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}c{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymsigma}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isasymin}\ {\isasymtau}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtau}{\isasymin}M{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ name{\isacharunderscore}{\kern0pt}components{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymchi}\ {\isasymin}\ M{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}p\ {\isasymtturnstile}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic\ truth{\isacharunderscore}{\kern0pt}lemma{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}{\isasymchi}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbrackright}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic\ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}{\isasymsigma}{\isacharcomma}{\kern0pt}p{\isachargreater}{\kern0pt}{\isasymin}{\isacharquery}{\kern0pt}{\isasymtheta}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymsigma}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ x{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}elem\ {\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymtheta}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharquery}{\kern0pt}b{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}c{\isasymin}{\isacharquery}{\kern0pt}b{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}{\isacharquery}{\kern0pt}b\ {\isachardot}{\kern0pt}\ x{\isasymsubseteq}a\ {\isacharampersand}{\kern0pt}\ x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}{\isacharquery}{\kern0pt}b\ {\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharampersand}{\kern0pt}\ x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}{\isacharquery}{\kern0pt}b\ {\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasyminter}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}b{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}{\isacharquery}{\kern0pt}b\ {\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}inter{\isacharunderscore}{\kern0pt}Transset\ Transset{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}b{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ GenExtI\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{end}\isamarkupfalse%
\ \isanewline
\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\ \isakeyword{begin}\isanewline
\isanewline
\isacommand{interpretation}\isamarkupfalse%
\ mgtriv{\isacharcolon}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}trivial\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ generic\ Union{\isacharunderscore}{\kern0pt}MG\ pairing{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ transitivity{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}trivial{\isacharunderscore}{\kern0pt}def\ M{\isacharunderscore}{\kern0pt}trans{\isacharunderscore}{\kern0pt}def\ M{\isacharunderscore}{\kern0pt}trivial{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharsemicolon}{\kern0pt}\ blast{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{theorem}\isamarkupfalse%
\ power{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}power{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ power{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ rallI{\isacharcomma}{\kern0pt}\ simp\ only{\isacharcolon}{\kern0pt}setclass{\isacharunderscore}{\kern0pt}iff\ rex{\isacharunderscore}{\kern0pt}setclass{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}bex{\isacharparenright}{\kern0pt}\isanewline
\ \ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ a\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Pow{\isacharunderscore}{\kern0pt}inter{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}powerset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ a{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ mgtriv{\isachardot}{\kern0pt}powerset{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isachardot}{\kern0pt}\ powerset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ a{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Powerset_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%35=4%:%
%:%40=5%:%
%:%45=6%:%
%:%46=6%:%
%:%51=6%:%
%:%54=7%:%
%:%55=8%:%
%:%56=8%:%
%:%57=9%:%
%:%58=10%:%
%:%59=11%:%
%:%60=12%:%
%:%63=13%:%
%:%67=13%:%
%:%68=13%:%
%:%69=13%:%
%:%70=14%:%
%:%71=14%:%
%:%76=14%:%
%:%79=15%:%
%:%80=16%:%
%:%81=16%:%
%:%82=17%:%
%:%83=18%:%
%:%84=18%:%
%:%85=19%:%
%:%86=20%:%
%:%93=21%:%
%:%94=21%:%
%:%95=22%:%
%:%96=22%:%
%:%97=22%:%
%:%98=23%:%
%:%99=24%:%
%:%100=24%:%
%:%101=24%:%
%:%102=25%:%
%:%103=25%:%
%:%104=25%:%
%:%105=26%:%
%:%106=26%:%
%:%107=27%:%
%:%108=27%:%
%:%109=27%:%
%:%110=28%:%
%:%111=28%:%
%:%112=28%:%
%:%113=29%:%
%:%114=29%:%
%:%115=30%:%
%:%116=30%:%
%:%117=30%:%
%:%118=31%:%
%:%119=31%:%
%:%120=32%:%
%:%121=32%:%
%:%122=33%:%
%:%123=33%:%
%:%124=33%:%
%:%125=34%:%
%:%131=34%:%
%:%134=35%:%
%:%135=36%:%
%:%136=36%:%
%:%137=37%:%
%:%138=38%:%
%:%139=39%:%
%:%140=40%:%
%:%141=41%:%
%:%142=42%:%
%:%149=43%:%
%:%150=43%:%
%:%151=44%:%
%:%152=44%:%
%:%153=44%:%
%:%154=45%:%
%:%155=45%:%
%:%156=46%:%
%:%157=46%:%
%:%158=46%:%
%:%159=47%:%
%:%160=48%:%
%:%161=48%:%
%:%162=48%:%
%:%163=49%:%
%:%164=49%:%
%:%165=50%:%
%:%166=50%:%
%:%167=51%:%
%:%168=51%:%
%:%169=52%:%
%:%170=52%:%
%:%171=53%:%
%:%172=53%:%
%:%173=54%:%
%:%174=54%:%
%:%175=55%:%
%:%176=55%:%
%:%177=56%:%
%:%178=56%:%
%:%179=57%:%
%:%180=57%:%
%:%181=57%:%
%:%182=58%:%
%:%183=58%:%
%:%184=59%:%
%:%185=59%:%
%:%186=60%:%
%:%187=60%:%
%:%188=61%:%
%:%189=61%:%
%:%190=62%:%
%:%191=62%:%
%:%192=63%:%
%:%193=63%:%
%:%194=63%:%
%:%195=64%:%
%:%196=64%:%
%:%197=65%:%
%:%198=65%:%
%:%199=65%:%
%:%200=66%:%
%:%201=66%:%
%:%202=67%:%
%:%203=67%:%
%:%204=68%:%
%:%205=68%:%
%:%206=69%:%
%:%207=69%:%
%:%208=70%:%
%:%209=70%:%
%:%210=71%:%
%:%211=72%:%
%:%212=72%:%
%:%213=73%:%
%:%214=73%:%
%:%215=74%:%
%:%216=74%:%
%:%217=74%:%
%:%218=75%:%
%:%219=75%:%
%:%220=75%:%
%:%221=76%:%
%:%222=76%:%
%:%223=77%:%
%:%224=77%:%
%:%225=78%:%
%:%226=78%:%
%:%227=78%:%
%:%228=79%:%
%:%229=79%:%
%:%230=80%:%
%:%231=81%:%
%:%232=81%:%
%:%233=82%:%
%:%234=82%:%
%:%235=82%:%
%:%236=83%:%
%:%237=83%:%
%:%238=84%:%
%:%239=84%:%
%:%240=85%:%
%:%241=85%:%
%:%242=86%:%
%:%243=86%:%
%:%244=87%:%
%:%245=87%:%
%:%246=87%:%
%:%247=88%:%
%:%248=89%:%
%:%249=89%:%
%:%250=90%:%
%:%251=90%:%
%:%252=90%:%
%:%253=90%:%
%:%254=91%:%
%:%255=92%:%
%:%256=92%:%
%:%257=93%:%
%:%258=93%:%
%:%259=94%:%
%:%260=94%:%
%:%261=95%:%
%:%262=95%:%
%:%263=95%:%
%:%264=96%:%
%:%265=96%:%
%:%266=96%:%
%:%267=97%:%
%:%268=97%:%
%:%269=98%:%
%:%270=98%:%
%:%271=99%:%
%:%272=99%:%
%:%273=99%:%
%:%274=100%:%
%:%275=100%:%
%:%276=101%:%
%:%277=101%:%
%:%278=102%:%
%:%279=102%:%
%:%280=103%:%
%:%281=103%:%
%:%282=104%:%
%:%283=104%:%
%:%284=105%:%
%:%285=105%:%
%:%286=105%:%
%:%287=105%:%
%:%288=106%:%
%:%294=106%:%
%:%297=107%:%
%:%298=108%:%
%:%299=108%:%
%:%300=109%:%
%:%301=110%:%
%:%302=111%:%
%:%303=112%:%
%:%310=113%:%
%:%311=113%:%
%:%312=114%:%
%:%313=114%:%
%:%314=114%:%
%:%315=115%:%
%:%316=116%:%
%:%317=116%:%
%:%318=116%:%
%:%319=117%:%
%:%320=117%:%
%:%321=118%:%
%:%322=118%:%
%:%323=119%:%
%:%324=119%:%
%:%325=120%:%
%:%326=120%:%
%:%327=121%:%
%:%328=121%:%
%:%329=122%:%
%:%330=122%:%
%:%331=123%:%
%:%332=123%:%
%:%333=124%:%
%:%334=124%:%
%:%335=125%:%
%:%336=125%:%
%:%337=125%:%
%:%338=126%:%
%:%339=127%:%
%:%340=127%:%
%:%341=127%:%
%:%342=128%:%
%:%343=128%:%
%:%344=128%:%
%:%345=129%:%
%:%346=129%:%
%:%347=130%:%
%:%348=130%:%
%:%349=130%:%
%:%350=131%:%
%:%351=131%:%
%:%352=132%:%
%:%353=132%:%
%:%354=133%:%
%:%355=133%:%
%:%356=134%:%
%:%357=134%:%
%:%358=135%:%
%:%359=135%:%
%:%360=136%:%
%:%361=136%:%
%:%362=137%:%
%:%363=137%:%
%:%364=138%:%
%:%365=138%:%
%:%366=139%:%
%:%367=139%:%
%:%368=139%:%
%:%369=139%:%
%:%370=140%:%
%:%371=140%:%
%:%372=141%:%
%:%373=141%:%
%:%374=142%:%
%:%375=143%:%
%:%376=143%:%
%:%377=144%:%
%:%378=145%:%
%:%379=145%:%
%:%380=146%:%
%:%381=146%:%
%:%382=147%:%
%:%383=147%:%
%:%384=148%:%
%:%385=148%:%
%:%386=149%:%
%:%387=149%:%
%:%388=150%:%
%:%389=150%:%
%:%390=151%:%
%:%391=151%:%
%:%392=151%:%
%:%393=152%:%
%:%394=152%:%
%:%395=153%:%
%:%396=153%:%
%:%397=154%:%
%:%398=154%:%
%:%399=155%:%
%:%400=155%:%
%:%401=156%:%
%:%402=156%:%
%:%403=156%:%
%:%404=157%:%
%:%405=158%:%
%:%406=158%:%
%:%407=158%:%
%:%408=159%:%
%:%409=159%:%
%:%410=160%:%
%:%411=160%:%
%:%412=161%:%
%:%413=161%:%
%:%414=161%:%
%:%415=162%:%
%:%416=162%:%
%:%417=163%:%
%:%418=163%:%
%:%419=164%:%
%:%420=164%:%
%:%421=165%:%
%:%422=165%:%
%:%423=166%:%
%:%424=166%:%
%:%425=167%:%
%:%426=167%:%
%:%427=168%:%
%:%428=168%:%
%:%429=169%:%
%:%430=169%:%
%:%431=170%:%
%:%432=170%:%
%:%433=171%:%
%:%434=171%:%
%:%435=172%:%
%:%436=172%:%
%:%437=173%:%
%:%438=173%:%
%:%439=174%:%
%:%440=174%:%
%:%441=175%:%
%:%442=175%:%
%:%443=176%:%
%:%444=176%:%
%:%445=177%:%
%:%446=177%:%
%:%447=177%:%
%:%448=178%:%
%:%449=179%:%
%:%450=179%:%
%:%451=180%:%
%:%452=180%:%
%:%453=181%:%
%:%454=181%:%
%:%455=181%:%
%:%456=182%:%
%:%457=182%:%
%:%458=183%:%
%:%459=183%:%
%:%460=183%:%
%:%461=184%:%
%:%462=184%:%
%:%463=184%:%
%:%464=185%:%
%:%465=185%:%
%:%466=186%:%
%:%467=186%:%
%:%468=187%:%
%:%469=187%:%
%:%470=188%:%
%:%471=188%:%
%:%472=189%:%
%:%473=189%:%
%:%474=190%:%
%:%475=190%:%
%:%476=191%:%
%:%477=191%:%
%:%478=192%:%
%:%479=192%:%
%:%480=193%:%
%:%481=193%:%
%:%482=194%:%
%:%483=194%:%
%:%484=195%:%
%:%485=195%:%
%:%486=195%:%
%:%487=196%:%
%:%488=196%:%
%:%489=197%:%
%:%490=197%:%
%:%491=198%:%
%:%492=198%:%
%:%493=198%:%
%:%494=199%:%
%:%495=199%:%
%:%496=200%:%
%:%497=200%:%
%:%498=201%:%
%:%499=201%:%
%:%500=202%:%
%:%501=202%:%
%:%502=203%:%
%:%503=203%:%
%:%504=204%:%
%:%505=204%:%
%:%506=205%:%
%:%507=205%:%
%:%508=206%:%
%:%509=206%:%
%:%510=207%:%
%:%511=207%:%
%:%512=208%:%
%:%513=209%:%
%:%514=209%:%
%:%515=210%:%
%:%516=210%:%
%:%517=211%:%
%:%518=211%:%
%:%519=211%:%
%:%520=212%:%
%:%521=212%:%
%:%522=212%:%
%:%523=213%:%
%:%524=213%:%
%:%525=214%:%
%:%526=214%:%
%:%527=215%:%
%:%528=215%:%
%:%529=215%:%
%:%530=216%:%
%:%531=216%:%
%:%532=216%:%
%:%533=217%:%
%:%534=217%:%
%:%535=218%:%
%:%536=218%:%
%:%537=219%:%
%:%538=219%:%
%:%539=220%:%
%:%540=220%:%
%:%541=220%:%
%:%542=221%:%
%:%543=221%:%
%:%544=222%:%
%:%545=222%:%
%:%546=223%:%
%:%547=223%:%
%:%548=224%:%
%:%549=224%:%
%:%550=225%:%
%:%551=225%:%
%:%552=226%:%
%:%553=226%:%
%:%554=227%:%
%:%555=227%:%
%:%556=228%:%
%:%557=228%:%
%:%558=229%:%
%:%559=229%:%
%:%560=229%:%
%:%561=230%:%
%:%562=230%:%
%:%563=231%:%
%:%564=231%:%
%:%565=231%:%
%:%566=232%:%
%:%567=232%:%
%:%568=233%:%
%:%569=233%:%
%:%570=234%:%
%:%571=234%:%
%:%572=235%:%
%:%573=235%:%
%:%574=236%:%
%:%575=236%:%
%:%576=236%:%
%:%577=237%:%
%:%578=237%:%
%:%579=238%:%
%:%580=238%:%
%:%581=238%:%
%:%582=238%:%
%:%583=239%:%
%:%584=239%:%
%:%585=240%:%
%:%586=240%:%
%:%587=241%:%
%:%588=241%:%
%:%589=241%:%
%:%590=242%:%
%:%591=242%:%
%:%592=243%:%
%:%593=243%:%
%:%594=244%:%
%:%595=244%:%
%:%596=244%:%
%:%597=245%:%
%:%598=245%:%
%:%599=246%:%
%:%600=246%:%
%:%601=247%:%
%:%602=247%:%
%:%603=247%:%
%:%604=248%:%
%:%605=248%:%
%:%606=249%:%
%:%607=249%:%
%:%608=250%:%
%:%609=250%:%
%:%610=251%:%
%:%611=251%:%
%:%612=252%:%
%:%613=252%:%
%:%614=252%:%
%:%615=253%:%
%:%616=253%:%
%:%617=254%:%
%:%618=254%:%
%:%619=254%:%
%:%620=255%:%
%:%621=255%:%
%:%622=256%:%
%:%623=256%:%
%:%624=257%:%
%:%625=257%:%
%:%626=258%:%
%:%627=258%:%
%:%628=258%:%
%:%629=259%:%
%:%630=259%:%
%:%631=260%:%
%:%632=260%:%
%:%633=261%:%
%:%634=261%:%
%:%635=262%:%
%:%636=262%:%
%:%637=262%:%
%:%638=263%:%
%:%639=263%:%
%:%640=264%:%
%:%641=264%:%
%:%642=264%:%
%:%643=265%:%
%:%644=265%:%
%:%645=265%:%
%:%646=265%:%
%:%647=266%:%
%:%653=266%:%
%:%656=267%:%
%:%657=267%:%
%:%658=268%:%
%:%659=269%:%
%:%660=270%:%
%:%661=270%:%
%:%662=271%:%
%:%663=272%:%
%:%664=272%:%
%:%667=273%:%
%:%671=273%:%
%:%672=273%:%
%:%673=274%:%
%:%674=274%:%
%:%675=274%:%
%:%680=274%:%
%:%683=275%:%
%:%684=276%:%
%:%685=277%:%
%:%686=277%:%
%:%689=278%:%
%:%693=278%:%
%:%694=278%:%
%:%695=279%:%
%:%696=279%:%
%:%697=280%:%
%:%697=282%:%
%:%698=283%:%
%:%699=283%:%
%:%700=284%:%
%:%701=284%:%
%:%702=285%:%
%:%703=285%:%
%:%704=286%:%
%:%705=286%:%
%:%706=286%:%
%:%707=287%:%
%:%708=287%:%
%:%709=288%:%
%:%710=288%:%
%:%711=289%:%
%:%712=289%:%
%:%713=289%:%
%:%714=290%:%
%:%715=290%:%
%:%716=291%:%
%:%717=291%:%
%:%718=291%:%
%:%719=292%:%
%:%720=292%:%
%:%721=293%:%
%:%722=293%:%
%:%723=293%:%
%:%724=294%:%
%:%725=294%:%
%:%726=294%:%
%:%727=295%:%
%:%728=295%:%
%:%729=296%:%
%:%730=296%:%
%:%731=297%:%
%:%732=297%:%
%:%733=298%:%
%:%734=298%:%
%:%735=299%:%
%:%736=299%:%
%:%737=300%:%
%:%738=300%:%
%:%739=301%:%
%:%745=301%:%
%:%748=302%:%
%:%749=302%:%
%:%756=303%:%