### 2024-11-01 code/output/document/Proper_Extension.tex

%
\begin{isabellebody}%
\setisabellecontext{Proper{\isacharunderscore}{\kern0pt}Extension}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Separative notions and proper extensions%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Proper{\isacharunderscore}{\kern0pt}Extension\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ Names\isanewline
\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
The key ingredient to obtain a proper extension is to have
a \emph{separative preorder}:%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{locale}\isamarkupfalse%
\ separative{\isacharunderscore}{\kern0pt}notion\ {\isacharequal}{\kern0pt}\ forcing{\isacharunderscore}{\kern0pt}notion\ {\isacharplus}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ separative{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}P\ {\isasymLongrightarrow}\ {\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymexists}r{\isasymin}P{\isachardot}{\kern0pt}\ q\ {\isasympreceq}\ p\ {\isasymand}\ r\ {\isasympreceq}\ p\ {\isasymand}\ q\ {\isasymbottom}\ r{\isachardoublequoteclose}\isanewline
\isakeyword{begin}%
\begin{isamarkuptext}%
For separative preorders, the complement of every filter is
dense. Hence an $M$-generic filter can't belong to the ground model.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}complement{\isacharunderscore}{\kern0pt}dense{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{shows}\ {\isachardoublequoteopen}dense{\isacharparenleft}{\kern0pt}P\ {\isacharminus}{\kern0pt}\ G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ p\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}d{\isasymin}P\ {\isacharminus}{\kern0pt}\ G{\isachardot}{\kern0pt}\ d\ {\isasympreceq}\ p{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ assms\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ q\ r\ \isakeyword{where}\ {\isachardoublequoteopen}q\ {\isasympreceq}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasympreceq}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymbottom}\ r{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ separative{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ s\ \isakeyword{where}\ {\isachardoublequoteopen}s\ {\isasympreceq}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}s\ {\isasymnotin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}s\ {\isasymin}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}compat{\isacharbrackleft}{\kern0pt}of\ G\ q\ r{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ leq{\isacharunderscore}{\kern0pt}reflI\ \isacommand{unfolding}\isamarkupfalse%
\ Diff{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ ctm{\isacharunderscore}{\kern0pt}separative\ {\isacharequal}{\kern0pt}\ forcing{\isacharunderscore}{\kern0pt}data\ {\isacharplus}{\kern0pt}\ separative{\isacharunderscore}{\kern0pt}notion\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isakeyword{shows}\ {\isachardoublequoteopen}G\ {\isasymnotin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}G{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}P\ {\isacharminus}{\kern0pt}\ G\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ Diff{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymexists}q{\isasymin}G{\isachardot}{\kern0pt}\ q\ {\isasymin}\ P\ {\isacharminus}{\kern0pt}\ G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}P\ {\isacharminus}{\kern0pt}\ G{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Diff{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}False{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}complement{\isacharunderscore}{\kern0pt}dense{\isacharbrackleft}{\kern0pt}of\ G{\isacharbrackright}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}denseD{\isacharbrackleft}{\kern0pt}of\ G\ {\isachardoublequoteopen}P{\isacharminus}{\kern0pt}G{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\ %
\isamarkupcmt{need to put generic ==> filter in claset%
}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{theorem}\isamarkupfalse%
\ proper{\isacharunderscore}{\kern0pt}extension{\isacharcolon}{\kern0pt}\ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{shows}\ {\isachardoublequoteopen}M\ {\isasymnoteq}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ G{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Gen{\isacharunderscore}{\kern0pt}Ext{\isacharbrackleft}{\kern0pt}of\ G{\isacharbrackright}{\kern0pt}\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G{\isacharbrackleft}{\kern0pt}of\ G{\isacharbrackright}{\kern0pt}\ generic{\isacharunderscore}{\kern0pt}not{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Proper_Extension.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%32=6%:%
%:%41=8%:%
%:%42=9%:%
%:%44=11%:%
%:%45=11%:%
%:%46=12%:%
%:%47=13%:%
%:%49=15%:%
%:%50=16%:%
%:%52=18%:%
%:%53=18%:%
%:%54=19%:%
%:%61=20%:%
%:%62=20%:%
%:%63=21%:%
%:%64=21%:%
%:%65=22%:%
%:%66=22%:%
%:%67=23%:%
%:%68=23%:%
%:%69=24%:%
%:%70=24%:%
%:%71=25%:%
%:%72=25%:%
%:%73=26%:%
%:%74=26%:%
%:%75=27%:%
%:%76=27%:%
%:%77=28%:%
%:%78=28%:%
%:%79=29%:%
%:%80=29%:%
%:%81=30%:%
%:%82=30%:%
%:%83=31%:%
%:%84=31%:%
%:%85=32%:%
%:%86=32%:%
%:%87=33%:%
%:%88=33%:%
%:%89=34%:%
%:%90=34%:%
%:%91=35%:%
%:%92=35%:%
%:%93=36%:%
%:%94=36%:%
%:%95=36%:%
%:%96=37%:%
%:%97=37%:%
%:%98=38%:%
%:%99=38%:%
%:%100=39%:%
%:%101=39%:%
%:%102=40%:%
%:%103=40%:%
%:%104=40%:%
%:%105=40%:%
%:%106=40%:%
%:%107=41%:%
%:%108=41%:%
%:%109=42%:%
%:%115=42%:%
%:%118=43%:%
%:%119=44%:%
%:%120=44%:%
%:%121=45%:%
%:%122=46%:%
%:%123=46%:%
%:%124=47%:%
%:%125=48%:%
%:%126=49%:%
%:%127=49%:%
%:%134=50%:%
%:%135=50%:%
%:%136=51%:%
%:%137=51%:%
%:%138=52%:%
%:%139=52%:%
%:%140=53%:%
%:%141=53%:%
%:%142=54%:%
%:%143=54%:%
%:%144=54%:%
%:%145=55%:%
%:%146=55%:%
%:%147=56%:%
%:%148=56%:%
%:%149=57%:%
%:%150=57%:%
%:%151=57%:%
%:%152=58%:%
%:%153=58%:%
%:%154=59%:%
%:%155=59%:%
%:%156=60%:%
%:%157=60%:%
%:%158=61%:%
%:%159=61%:%
%:%160=62%:%
%:%161=62%:%
%:%162=63%:%
%:%163=63%:%
%:%164=63%:%
%:%165=63%:%
%:%166=64%:%
%:%172=64%:%
%:%175=65%:%
%:%176=66%:%
%:%177=66%:%
%:%180=67%:%
%:%184=67%:%
%:%185=67%:%
%:%186=68%:%
%:%187=68%:%
%:%192=68%:%
%:%195=69%:%
%:%196=70%:%
%:%197=70%:%
%:%200=71%:%
%:%205=72%:%