### 2024-11-01 code/output/document/RecFun_M_Memrel.tex

%
\begin{isabellebody}%
\setisabellecontext{RecFun{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}Memrel}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ RecFun{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}Memrel\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ ZF\ \isanewline
\ \ \ \ RecFun{\isacharunderscore}{\kern0pt}M\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{definition}\isamarkupfalse%
\ InEclose\ \isakeyword{where}\ {\isachardoublequoteopen}InEclose{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\isacommand{definition}\isamarkupfalse%
\ InEclose{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}InEclose{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ \isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Relation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}InEclose\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}Relation{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ InEclose{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Relation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ clarify\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ InEclose{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}n{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ {\isasymforall}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ x\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}n{\isachardot}{\kern0pt}\ {\isasymforall}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isakeyword{in}\ nat{\isacharunderscore}{\kern0pt}induct{\isacharcomma}{\kern0pt}\ assumption{\isacharcomma}{\kern0pt}\ force{\isacharcomma}{\kern0pt}\ rule\ allI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n\ x\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ Union{\isacharparenleft}{\kern0pt}Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ yH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ y{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yH\ transM\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ \isakeyword{where}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}Union\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}Union{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}InEclose\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ x\ \isakeyword{where}\ yxH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ InEclose{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}n{\isasymin}nat{\isachardot}{\kern0pt}\ Union{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}Union\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ \isakeyword{where}\ nH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}n{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ {\isasymforall}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}n{\isachardot}{\kern0pt}\ {\isasymforall}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ nat{\isacharunderscore}{\kern0pt}induct{\isacharcomma}{\kern0pt}\ assumption{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n\ \isacommand{assume}\isamarkupfalse%
\ nnat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}{\isadigit{0}}\ {\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ allI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ z\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}z\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}{\isadigit{0}}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ rel\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ zin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yxH\ transM\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ r{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Memrel{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ rel\ zin\ yxH\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n\ {\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ x{\isasymrangle}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ x{\isasymrangle}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ allI{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ z\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}z\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z\ {\isasymin}\ Union{\isacharparenleft}{\kern0pt}Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ w\ \isakeyword{where}\ wH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z\ {\isasymin}\ w{\isachardoublequoteclose}\ {\isachardoublequoteopen}w\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ rel\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}w{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isasymin}\ field{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w\ {\isasymin}\ field{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}trancl\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ win\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}w\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ zin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ transM\ wH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}w\ \isakeyword{in}\ rtrancl{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}trancl{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ win\ zin\ wH\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ trancl{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}rtrancl{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}rel{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yxH\ nH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ x\ \isakeyword{where}\ yxH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ trancl{\isacharunderscore}{\kern0pt}def\ comp{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}y\ \isakeyword{in}\ trancl{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yxH\ vin\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ z\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}z\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ z{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}Union{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A\ B\ n{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ A\ {\isasymsubseteq}\ B\ {\isasymLongrightarrow}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ A\ B\ n{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}n{\isachardot}{\kern0pt}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ nat{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Union{\isacharunderscore}{\kern0pt}mono{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ z\ w\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ w{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ yinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ winM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}w\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ zinw\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z\ {\isasymin}\ w{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ Rrel{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ n\ \isakeyword{where}\ nH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}Union\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}Union{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ subsetD{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ H{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ zinw\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ Union{\isacharparenleft}{\kern0pt}Union{\isacharcircum}{\kern0pt}n{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ iterates{\isacharunderscore}{\kern0pt}commute\ nH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ Union{\isacharcircum}{\kern0pt}{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}Union{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nH\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ w{\isachargreater}{\kern0pt}\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ yinM\ winM\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yxH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ preds{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ preds{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ preds{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ transM\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ eclose{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}InEclose\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}field{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}field{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ rule\ disjI{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ domainI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}Union{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}vimage{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ M\ {\isasymLongrightarrow}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ eclose{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ y\ z{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}y\ \isakeyword{in}\ vimageI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ y\ z{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ transM\ eclose{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}Gfm{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}Gfm{\isacharcomma}{\kern0pt}\ InEclose{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ \isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ Gfm\ i\ j\ k\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}Gfm\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}Gfm{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ is{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ Gfm\ i\ j\ k\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}Gfm\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Gfm{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{3}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}Gfm{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isasymunion}\ succ{\isacharparenleft}{\kern0pt}k{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ InEclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ le{\isacharunderscore}{\kern0pt}refl\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ H\ G\ Gfm\ x\ v\ i\ j\ k\ env\ a\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}Gfm\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Gfm{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{3}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}x\ g{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ HGeq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}h\ g\ x{\isachardot}{\kern0pt}\ h\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ {\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ h{\isacharbackquote}{\kern0pt}y\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ H{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{and}\ sats{\isacharunderscore}{\kern0pt}Gfm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}a{\isadigit{0}}\ a{\isadigit{1}}\ a{\isadigit{2}}\ env{\isachardot}{\kern0pt}\ a{\isadigit{0}}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ a{\isadigit{1}}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ a{\isadigit{2}}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ a{\isadigit{0}}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}\ a{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ Gfm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isadigit{0}}{\isacharcomma}{\kern0pt}\ a{\isadigit{1}}{\isacharcomma}{\kern0pt}\ a{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}Gfm{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ prelvimageeq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ M\ {\isasymLongrightarrow}\ prel{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ a{\isasymrangle}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ eclose{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ prel{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ a{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}v{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ prel{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ z\ \isakeyword{where}\ zH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ prel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ zH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ v\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ z\ \isakeyword{where}\ zH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}z\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transM\ zH\ assms{\isadigit{1}}\ eclose{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ InEclose{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ zH\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}\ {\isasymin}\ prel{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ a{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ vimageI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ prelI{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}H{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ prel{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}\ a{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ zH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}Gfm{\isacharcomma}{\kern0pt}\ InEclose{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ \isanewline
\ \ \ \ \ \ a\ {\isacharequal}{\kern0pt}\ a\ \isakeyword{and}\ \isanewline
\ \ \ \ \ \ G\ {\isacharequal}{\kern0pt}\ G\isanewline
\ \ \ \ \ \ \isakeyword{in}\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}InEclose\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ rule\ wf{\isacharunderscore}{\kern0pt}trancl{\isacharcomma}{\kern0pt}\ rule\ wf{\isacharunderscore}{\kern0pt}Memrel{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ rule\ trans{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ preds{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ eclose{\isacharunderscore}{\kern0pt}closed{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Relation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}InEclose\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}Gfm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ h\ g\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HGeq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ h\ g\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}vimage{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ h\ g\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}vimage{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ h\ g\ x\ y{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}vimage{\isacharunderscore}{\kern0pt}eq\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}M\ \isakeyword{in}\ ssubst{\isacharcomma}{\kern0pt}\ rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \ \isacommand{using}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}InEclose\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ H\ G\ Gfm\ x\ a\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}Gfm\ {\isasymin}\ formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Gfm{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{3}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}x\ g{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{and}\ HGeq{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}h\ g\ x{\isachardot}{\kern0pt}\ h\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ {\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ h{\isacharbackquote}{\kern0pt}y\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ H{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ a{\isachargreater}{\kern0pt}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \isakeyword{and}\ sats{\isacharunderscore}{\kern0pt}Gfm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}a{\isadigit{0}}\ a{\isadigit{1}}\ a{\isadigit{2}}\ env{\isachardot}{\kern0pt}\ a{\isadigit{0}}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ a{\isadigit{1}}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ a{\isadigit{2}}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ a{\isadigit{0}}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}a{\isadigit{2}}{\isacharcomma}{\kern0pt}\ a{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ Gfm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isadigit{0}}{\isacharcomma}{\kern0pt}\ a{\isadigit{1}}{\isacharcomma}{\kern0pt}\ a{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharcomma}{\kern0pt}\ rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}a\ \isakeyword{and}\ Gfm{\isacharequal}{\kern0pt}Gfm\ \isakeyword{in}\ Rrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ rule\ wf{\isacharunderscore}{\kern0pt}trancl{\isacharcomma}{\kern0pt}\ rule\ wf{\isacharunderscore}{\kern0pt}Memrel{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ rule\ trans{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Relation{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}InEclose{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ preds{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ eclose{\isacharunderscore}{\kern0pt}closed{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}Gfm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ h\ g\ x{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HGeq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ h\ g\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}vimage{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ h\ g\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}vimage{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ h\ g\ x\ y{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharunderscore}{\kern0pt}vimage{\isacharunderscore}{\kern0pt}eq\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}M\ \isakeyword{in}\ ssubst{\isacharcomma}{\kern0pt}\ rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}InEclose{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}trancl\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ y\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ z\ \isakeyword{where}\ zH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ singin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}\ {\isasymin}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pair{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ yinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ yzinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ zH\ assms\ transM\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ zinM\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ rtrancl{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}trancl{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ rtrancl{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}rtrancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ r{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}rtrancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Memrel{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ yinM\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ yinM\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ zH\ singin\ zinM\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ zH\ yzinM\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ y\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ z\ \isakeyword{where}\ zH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ eclose{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}Union{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isadigit{2}}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ z{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pair{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ zH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/RecFun_M_Memrel.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=8%:%
%:%28=9%:%
%:%29=10%:%
%:%30=10%:%
%:%31=11%:%
%:%32=12%:%
%:%33=13%:%
%:%34=13%:%
%:%35=14%:%
%:%38=15%:%
%:%42=15%:%
%:%43=15%:%
%:%44=16%:%
%:%45=16%:%
%:%46=17%:%
%:%47=17%:%
%:%48=18%:%
%:%49=18%:%
%:%50=19%:%
%:%51=19%:%
%:%52=20%:%
%:%53=20%:%
%:%54=21%:%
%:%55=21%:%
%:%56=22%:%
%:%57=22%:%
%:%58=23%:%
%:%59=23%:%
%:%64=23%:%
%:%67=24%:%
%:%68=25%:%
%:%69=25%:%
%:%76=26%:%
%:%77=26%:%
%:%78=27%:%
%:%79=27%:%
%:%80=28%:%
%:%81=28%:%
%:%82=29%:%
%:%83=29%:%
%:%84=29%:%
%:%85=30%:%
%:%86=30%:%
%:%87=30%:%
%:%88=30%:%
%:%89=31%:%
%:%90=31%:%
%:%91=31%:%
%:%92=31%:%
%:%93=32%:%
%:%94=32%:%
%:%95=32%:%
%:%96=32%:%
%:%97=32%:%
%:%98=33%:%
%:%99=33%:%
%:%100=33%:%
%:%101=33%:%
%:%102=33%:%
%:%103=34%:%
%:%104=34%:%
%:%105=35%:%
%:%106=35%:%
%:%107=35%:%
%:%108=36%:%
%:%109=36%:%
%:%110=36%:%
%:%111=36%:%
%:%112=36%:%
%:%113=37%:%
%:%114=37%:%
%:%115=37%:%
%:%116=37%:%
%:%117=37%:%
%:%118=38%:%
%:%119=38%:%
%:%120=39%:%
%:%121=39%:%
%:%122=39%:%
%:%123=40%:%
%:%124=40%:%
%:%125=40%:%
%:%126=41%:%
%:%127=41%:%
%:%128=42%:%
%:%129=42%:%
%:%130=43%:%
%:%131=43%:%
%:%132=44%:%
%:%138=44%:%
%:%141=45%:%
%:%142=46%:%
%:%143=46%:%
%:%144=47%:%
%:%151=48%:%
%:%152=48%:%
%:%153=49%:%
%:%154=49%:%
%:%155=49%:%
%:%156=50%:%
%:%157=50%:%
%:%158=50%:%
%:%159=50%:%
%:%160=50%:%
%:%161=51%:%
%:%162=51%:%
%:%163=51%:%
%:%164=51%:%
%:%165=51%:%
%:%166=52%:%
%:%167=52%:%
%:%168=52%:%
%:%169=52%:%
%:%170=53%:%
%:%171=54%:%
%:%172=54%:%
%:%173=55%:%
%:%174=55%:%
%:%175=56%:%
%:%176=56%:%
%:%177=56%:%
%:%178=57%:%
%:%179=57%:%
%:%180=58%:%
%:%181=58%:%
%:%182=59%:%
%:%183=59%:%
%:%184=59%:%
%:%185=60%:%
%:%186=60%:%
%:%187=60%:%
%:%188=60%:%
%:%189=61%:%
%:%190=61%:%
%:%191=61%:%
%:%192=61%:%
%:%193=61%:%
%:%194=62%:%
%:%195=63%:%
%:%196=63%:%
%:%197=64%:%
%:%198=64%:%
%:%199=65%:%
%:%200=65%:%
%:%201=66%:%
%:%202=66%:%
%:%203=67%:%
%:%204=67%:%
%:%205=68%:%
%:%206=68%:%
%:%207=69%:%
%:%208=69%:%
%:%209=70%:%
%:%210=70%:%
%:%211=71%:%
%:%212=71%:%
%:%213=72%:%
%:%214=72%:%
%:%215=73%:%
%:%216=73%:%
%:%217=74%:%
%:%218=74%:%
%:%219=74%:%
%:%220=75%:%
%:%221=75%:%
%:%222=75%:%
%:%223=75%:%
%:%224=76%:%
%:%225=76%:%
%:%226=76%:%
%:%227=76%:%
%:%228=77%:%
%:%229=77%:%
%:%230=77%:%
%:%231=77%:%
%:%232=77%:%
%:%233=78%:%
%:%234=78%:%
%:%235=78%:%
%:%236=78%:%
%:%237=79%:%
%:%238=79%:%
%:%239=79%:%
%:%240=79%:%
%:%241=79%:%
%:%242=80%:%
%:%243=80%:%
%:%244=80%:%
%:%245=80%:%
%:%246=81%:%
%:%247=81%:%
%:%248=81%:%
%:%249=81%:%
%:%250=81%:%
%:%251=82%:%
%:%252=82%:%
%:%253=83%:%
%:%254=83%:%
%:%255=84%:%
%:%256=84%:%
%:%257=85%:%
%:%258=85%:%
%:%259=86%:%
%:%260=86%:%
%:%261=87%:%
%:%262=87%:%
%:%263=88%:%
%:%264=88%:%
%:%265=89%:%
%:%266=89%:%
%:%267=90%:%
%:%268=91%:%
%:%269=91%:%
%:%270=91%:%
%:%271=91%:%
%:%272=91%:%
%:%273=92%:%
%:%274=92%:%
%:%275=93%:%
%:%276=93%:%
%:%277=93%:%
%:%278=94%:%
%:%279=94%:%
%:%280=94%:%
%:%281=94%:%
%:%282=94%:%
%:%283=95%:%
%:%284=95%:%
%:%285=95%:%
%:%286=96%:%
%:%287=96%:%
%:%288=97%:%
%:%289=97%:%
%:%290=97%:%
%:%291=97%:%
%:%292=98%:%
%:%293=98%:%
%:%294=99%:%
%:%295=99%:%
%:%296=99%:%
%:%297=100%:%
%:%298=100%:%
%:%299=100%:%
%:%300=100%:%
%:%301=101%:%
%:%302=101%:%
%:%303=101%:%
%:%304=102%:%
%:%305=102%:%
%:%306=103%:%
%:%307=103%:%
%:%308=104%:%
%:%309=104%:%
%:%310=105%:%
%:%311=105%:%
%:%312=106%:%
%:%313=106%:%
%:%314=107%:%
%:%315=108%:%
%:%316=108%:%
%:%317=109%:%
%:%318=109%:%
%:%319=110%:%
%:%320=110%:%
%:%321=111%:%
%:%322=111%:%
%:%323=112%:%
%:%324=112%:%
%:%325=113%:%
%:%326=113%:%
%:%327=114%:%
%:%328=115%:%
%:%329=115%:%
%:%330=116%:%
%:%331=116%:%
%:%332=117%:%
%:%333=118%:%
%:%334=118%:%
%:%335=118%:%
%:%336=118%:%
%:%337=119%:%
%:%338=119%:%
%:%339=119%:%
%:%340=119%:%
%:%341=120%:%
%:%342=120%:%
%:%343=120%:%
%:%344=120%:%
%:%345=121%:%
%:%346=122%:%
%:%347=122%:%
%:%348=122%:%
%:%349=122%:%
%:%350=123%:%
%:%351=123%:%
%:%352=123%:%
%:%353=123%:%
%:%354=123%:%
%:%355=124%:%
%:%356=124%:%
%:%357=124%:%
%:%358=125%:%
%:%359=125%:%
%:%360=126%:%
%:%361=126%:%
%:%362=127%:%
%:%363=127%:%
%:%364=128%:%
%:%365=128%:%
%:%366=129%:%
%:%367=129%:%
%:%368=129%:%
%:%369=129%:%
%:%370=129%:%
%:%371=130%:%
%:%372=130%:%
%:%373=130%:%
%:%374=130%:%
%:%375=131%:%
%:%376=131%:%
%:%377=131%:%
%:%378=132%:%
%:%379=132%:%
%:%380=133%:%
%:%381=133%:%
%:%382=134%:%
%:%383=134%:%
%:%384=135%:%
%:%385=135%:%
%:%386=135%:%
%:%387=136%:%
%:%388=136%:%
%:%389=137%:%
%:%390=137%:%
%:%391=138%:%
%:%392=138%:%
%:%393=139%:%
%:%394=139%:%
%:%395=140%:%
%:%396=140%:%
%:%397=140%:%
%:%398=140%:%
%:%399=140%:%
%:%400=141%:%
%:%406=141%:%
%:%409=142%:%
%:%410=143%:%
%:%411=143%:%
%:%414=144%:%
%:%418=144%:%
%:%419=144%:%
%:%420=145%:%
%:%421=145%:%
%:%422=146%:%
%:%423=146%:%
%:%424=147%:%
%:%425=147%:%
%:%426=148%:%
%:%427=148%:%
%:%428=149%:%
%:%429=149%:%
%:%430=150%:%
%:%431=150%:%
%:%436=150%:%
%:%439=151%:%
%:%440=152%:%
%:%441=152%:%
%:%444=153%:%
%:%448=153%:%
%:%449=153%:%
%:%450=154%:%
%:%451=154%:%
%:%452=155%:%
%:%453=155%:%
%:%454=156%:%
%:%455=156%:%
%:%456=157%:%
%:%457=157%:%
%:%458=158%:%
%:%459=158%:%
%:%460=159%:%
%:%461=159%:%
%:%462=160%:%
%:%463=160%:%
%:%464=161%:%
%:%465=161%:%
%:%466=162%:%
%:%467=162%:%
%:%472=162%:%
%:%475=163%:%
%:%476=164%:%
%:%477=164%:%
%:%480=165%:%
%:%484=165%:%
%:%485=165%:%
%:%486=166%:%
%:%487=166%:%
%:%488=167%:%
%:%489=167%:%
%:%490=168%:%
%:%491=168%:%
%:%492=169%:%
%:%493=169%:%
%:%494=170%:%
%:%495=170%:%
%:%500=170%:%
%:%503=171%:%
%:%504=172%:%
%:%505=172%:%
%:%506=173%:%
%:%507=174%:%
%:%508=174%:%
%:%509=175%:%
%:%510=176%:%
%:%511=176%:%
%:%512=177%:%
%:%513=178%:%
%:%514=179%:%
%:%515=179%:%
%:%516=180%:%
%:%517=181%:%
%:%518=182%:%
%:%521=183%:%
%:%522=184%:%
%:%526=184%:%
%:%527=184%:%
%:%528=185%:%
%:%529=185%:%
%:%530=186%:%
%:%531=186%:%
%:%532=187%:%
%:%533=187%:%
%:%538=187%:%
%:%541=188%:%
%:%542=189%:%
%:%543=189%:%
%:%544=190%:%
%:%545=191%:%
%:%546=192%:%
%:%549=193%:%
%:%550=194%:%
%:%554=194%:%
%:%555=194%:%
%:%556=195%:%
%:%557=195%:%
%:%558=196%:%
%:%559=196%:%
%:%560=197%:%
%:%561=197%:%
%:%562=198%:%
%:%563=198%:%
%:%564=199%:%
%:%565=199%:%
%:%566=200%:%
%:%567=200%:%
%:%568=201%:%
%:%569=201%:%
%:%570=202%:%
%:%571=202%:%
%:%572=203%:%
%:%573=203%:%
%:%578=203%:%
%:%581=204%:%
%:%582=205%:%
%:%583=205%:%
%:%584=206%:%
%:%585=207%:%
%:%586=208%:%
%:%587=209%:%
%:%588=210%:%
%:%589=211%:%
%:%590=212%:%
%:%591=213%:%
%:%592=214%:%
%:%593=215%:%
%:%594=216%:%
%:%595=217%:%
%:%602=218%:%
%:%603=218%:%
%:%604=219%:%
%:%605=220%:%
%:%606=220%:%
%:%607=221%:%
%:%608=221%:%
%:%609=222%:%
%:%610=222%:%
%:%611=222%:%
%:%612=223%:%
%:%613=223%:%
%:%614=223%:%
%:%615=223%:%
%:%616=224%:%
%:%617=224%:%
%:%618=224%:%
%:%619=224%:%
%:%620=224%:%
%:%621=225%:%
%:%622=225%:%
%:%623=225%:%
%:%624=225%:%
%:%625=225%:%
%:%626=226%:%
%:%627=226%:%
%:%628=226%:%
%:%629=226%:%
%:%630=226%:%
%:%631=227%:%
%:%632=227%:%
%:%633=228%:%
%:%634=228%:%
%:%635=228%:%
%:%636=229%:%
%:%637=229%:%
%:%638=229%:%
%:%639=229%:%
%:%640=230%:%
%:%641=230%:%
%:%642=230%:%
%:%643=231%:%
%:%644=231%:%
%:%645=232%:%
%:%646=232%:%
%:%647=233%:%
%:%648=233%:%
%:%649=234%:%
%:%650=234%:%
%:%651=234%:%
%:%652=235%:%
%:%653=235%:%
%:%654=236%:%
%:%655=236%:%
%:%656=237%:%
%:%657=237%:%
%:%658=238%:%
%:%659=238%:%
%:%660=239%:%
%:%661=239%:%
%:%662=240%:%
%:%663=240%:%
%:%664=241%:%
%:%665=241%:%
%:%666=242%:%
%:%667=242%:%
%:%668=243%:%
%:%669=243%:%
%:%670=243%:%
%:%671=244%:%
%:%672=244%:%
%:%673=244%:%
%:%674=245%:%
%:%675=245%:%
%:%676=246%:%
%:%677=247%:%
%:%678=247%:%
%:%679=248%:%
%:%680=248%:%
%:%681=249%:%
%:%682=250%:%
%:%683=251%:%
%:%684=252%:%
%:%685=252%:%
%:%686=253%:%
%:%687=253%:%
%:%688=254%:%
%:%689=254%:%
%:%690=255%:%
%:%691=255%:%
%:%692=256%:%
%:%693=256%:%
%:%694=257%:%
%:%695=257%:%
%:%696=258%:%
%:%697=258%:%
%:%698=259%:%
%:%699=259%:%
%:%700=260%:%
%:%701=260%:%
%:%702=261%:%
%:%703=261%:%
%:%704=262%:%
%:%705=262%:%
%:%706=263%:%
%:%707=263%:%
%:%708=264%:%
%:%709=264%:%
%:%710=265%:%
%:%711=265%:%
%:%712=266%:%
%:%713=266%:%
%:%714=267%:%
%:%715=267%:%
%:%716=268%:%
%:%717=268%:%
%:%718=269%:%
%:%719=269%:%
%:%720=270%:%
%:%721=270%:%
%:%722=271%:%
%:%723=271%:%
%:%724=272%:%
%:%725=273%:%
%:%726=273%:%
%:%727=273%:%
%:%728=273%:%
%:%729=273%:%
%:%730=273%:%
%:%731=274%:%
%:%737=274%:%
%:%740=275%:%
%:%741=276%:%
%:%742=276%:%
%:%743=277%:%
%:%744=278%:%
%:%745=279%:%
%:%746=280%:%
%:%747=281%:%
%:%748=282%:%
%:%749=283%:%
%:%750=284%:%
%:%751=285%:%
%:%752=286%:%
%:%755=287%:%
%:%756=288%:%
%:%760=288%:%
%:%761=288%:%
%:%762=289%:%
%:%763=289%:%
%:%764=290%:%
%:%765=290%:%
%:%766=291%:%
%:%767=291%:%
%:%768=292%:%
%:%769=292%:%
%:%770=293%:%
%:%771=293%:%
%:%772=294%:%
%:%773=294%:%
%:%774=295%:%
%:%775=295%:%
%:%776=296%:%
%:%777=296%:%
%:%778=297%:%
%:%779=297%:%
%:%780=298%:%
%:%781=298%:%
%:%782=299%:%
%:%783=299%:%
%:%784=300%:%
%:%785=300%:%
%:%786=301%:%
%:%787=301%:%
%:%788=302%:%
%:%789=302%:%
%:%790=303%:%
%:%791=303%:%
%:%792=304%:%
%:%793=304%:%
%:%794=305%:%
%:%795=305%:%
%:%796=306%:%
%:%797=306%:%
%:%798=307%:%
%:%799=307%:%
%:%800=308%:%
%:%801=308%:%
%:%802=309%:%
%:%803=309%:%
%:%804=310%:%
%:%805=310%:%
%:%806=311%:%
%:%807=311%:%
%:%808=312%:%
%:%809=312%:%
%:%810=313%:%
%:%816=313%:%
%:%819=314%:%
%:%820=315%:%
%:%821=315%:%
%:%822=316%:%
%:%823=317%:%
%:%824=318%:%
%:%831=319%:%
%:%832=319%:%
%:%833=320%:%
%:%834=320%:%
%:%835=320%:%
%:%836=320%:%
%:%837=321%:%
%:%838=321%:%
%:%839=321%:%
%:%840=321%:%
%:%841=322%:%
%:%842=322%:%
%:%843=322%:%
%:%844=322%:%
%:%845=323%:%
%:%846=323%:%
%:%847=323%:%
%:%848=323%:%
%:%849=324%:%
%:%850=324%:%
%:%851=324%:%
%:%852=324%:%
%:%853=324%:%
%:%854=325%:%
%:%855=325%:%
%:%856=326%:%
%:%857=326%:%
%:%858=327%:%
%:%859=327%:%
%:%860=328%:%
%:%861=328%:%
%:%862=329%:%
%:%863=329%:%
%:%864=330%:%
%:%865=330%:%
%:%866=331%:%
%:%867=331%:%
%:%868=332%:%
%:%869=332%:%
%:%870=333%:%
%:%871=333%:%
%:%872=334%:%
%:%873=334%:%
%:%874=335%:%
%:%875=335%:%
%:%876=336%:%
%:%877=336%:%
%:%878=337%:%
%:%884=337%:%
%:%887=338%:%
%:%888=339%:%
%:%889=339%:%
%:%890=340%:%
%:%891=341%:%
%:%892=342%:%
%:%899=343%:%
%:%900=343%:%
%:%901=344%:%
%:%902=344%:%
%:%903=344%:%
%:%904=344%:%
%:%905=345%:%
%:%906=345%:%
%:%907=346%:%
%:%908=346%:%
%:%909=347%:%
%:%910=347%:%
%:%911=348%:%
%:%912=348%:%
%:%913=349%:%
%:%914=349%:%
%:%915=350%:%
%:%916=350%:%
%:%917=351%:%
%:%918=351%:%
%:%919=352%:%
%:%925=352%:%
%:%928=353%:%
%:%929=354%:%
%:%930=354%:%
%:%937=355%:%