### 2024-11-01 code/output/document/Recursion_Thms.tex

%
\begin{isabellebody}%
\setisabellecontext{Recursion{\isacharunderscore}{\kern0pt}Thms}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Some enhanced theorems on recursion%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Recursion{\isacharunderscore}{\kern0pt}Thms\ \isakeyword{imports}\ ZF{\isachardot}{\kern0pt}Epsilon\ \isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
We prove results concerning definitions by well-founded
recursion on some relation \isa{R} and its transitive closure
\isa{R{\isacharcircum}{\kern0pt}{\isacharasterisk}{\kern0pt}}%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ fld{\isacharunderscore}{\kern0pt}restrict{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}r\ {\isasyminter}\ A{\isasymtimes}A{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasyminter}\ A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}force{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ fld{\isacharunderscore}{\kern0pt}restrict{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ A\ {\isasymsubseteq}\ B\ {\isasymLongrightarrow}\ r\ {\isasyminter}\ A{\isasymtimes}A\ {\isasymsubseteq}\ r\ {\isasyminter}\ B{\isasymtimes}B{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ fld{\isacharunderscore}{\kern0pt}restrict{\isacharunderscore}{\kern0pt}dom\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isasymsubseteq}\ A{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}r{\isasyminter}\ A{\isasymtimes}A\ {\isacharequal}{\kern0pt}\ r{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equalityI{\isacharcomma}{\kern0pt}blast{\isacharcomma}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ xr{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ r{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ xr\ assms\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}\ a\ b\ {\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ relation{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}\ {\isasymin}\ r{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}a{\isacharcomma}{\kern0pt}b{\isasymrangle}\ {\isasymin}\ r{\isasyminter}A{\isasymtimes}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ r{\isasyminter}A{\isasymtimes}A{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ xr\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}\ r\ {\isasyminter}\ A{\isasymtimes}A{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ r\ {\isasymLongrightarrow}\ x{\isasymin}\ r{\isasyminter}A{\isasymtimes}A{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}down\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}downD\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}a{\isasymrangle}\ {\isasymin}\ r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ tr{\isacharunderscore}{\kern0pt}down{\isacharunderscore}{\kern0pt}def\ vimage{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}down\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ tr{\isacharunderscore}{\kern0pt}down{\isacharunderscore}{\kern0pt}def\ vimage{\isacharunderscore}{\kern0pt}mono\ r{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}down{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymLongrightarrow}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharcomma}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}tr{\isacharunderscore}{\kern0pt}down{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}auto\ dest{\isacharcolon}{\kern0pt}\ underD{\isacharcomma}{\kern0pt}force\ simp\ add{\isacharcolon}{\kern0pt}\ underI\ r{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}trancl\ trancl{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rest{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ B{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}a\ {\isasymin}\ B{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}r{\isasyminter}B{\isasymtimes}B{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ equalityI\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ B{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ subsetD{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}a{\isasymrangle}\ {\isasymin}\ r{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ underD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}r{\isasyminter}B{\isasymtimes}B{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}B{\isacartoucheclose}\ {\isacartoucheopen}a{\isasymin}B{\isacartoucheclose}\ underI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ r\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ \ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}r\ {\isasyminter}\ B{\isasymtimes}B{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ vimage{\isacharunderscore}{\kern0pt}mono\ that\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}restr{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}r{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ r\ {\isasyminter}\ A{\isasymtimes}A\ {\isasymLongrightarrow}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}r{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}restr\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ rr{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ wfr{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}wf{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ {\isachardoublequoteopen}a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ A\ {\isasymLongrightarrow}\ wfrec{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ a\ arbitrary{\isacharcolon}{\kern0pt}A\ rule{\isacharcolon}{\kern0pt}wf{\isacharunderscore}{\kern0pt}induct{\isacharunderscore}{\kern0pt}raw{\isacharbrackleft}{\kern0pt}OF\ wfr{\isacharbrackright}{\kern0pt}\ {\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{1}}\ a{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ wfRa\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}wf{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}subset\ wfr\ wf{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ Int{\isacharunderscore}{\kern0pt}lower{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}down\ rr\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ A{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force\ simp\ add{\isacharcolon}{\kern0pt}\ subset{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ x{\isacharunderscore}{\kern0pt}a\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ A{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}down\ rr\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ b\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}r\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ tr{\isacharunderscore}{\kern0pt}down{\isacharunderscore}{\kern0pt}mono\ x{\isacharunderscore}{\kern0pt}a\ rr\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ A{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ subset{\isacharunderscore}{\kern0pt}trans\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}a{\isasymrangle}\ {\isasymin}\ r{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ x{\isacharunderscore}{\kern0pt}a\ \ underD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\ {\isacartoucheopen}tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ A{\isacartoucheclose}\ {\isacartoucheopen}x\ {\isasymin}\ A{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}wfrec{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymLongrightarrow}\ wfrec{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{1}}\ {\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymlambda}\ x\ {\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isachardot}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymlambda}\ x\ {\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isachardot}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ lam{\isacharunderscore}{\kern0pt}cong\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}wfrec{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymlambda}\ x\ {\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isachardot}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}wfrec{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymlambda}\ x\ {\isasymin}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isachardot}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Eq{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isadigit{1}}\ {\isacartoucheopen}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ A{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymlambda}\ x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}r{\isasyminter}A{\isasymtimes}A{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}\ {\isachardot}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ rest{\isacharunderscore}{\kern0pt}eq\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}A{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymlambda}\ x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasyminter}A\ {\isachardot}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fld{\isacharunderscore}{\kern0pt}restrict{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}A{\isacartoucheclose}\ {\isacartoucheopen}wf{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}on\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}tr{\isacharunderscore}{\kern0pt}down\ {\isacharequal}{\kern0pt}\ wfrec{\isacharunderscore}{\kern0pt}restr{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ subset{\isacharunderscore}{\kern0pt}refl{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}trans{\isacharunderscore}{\kern0pt}restr\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ wf{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ trans{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isasymsubseteq}A\ {\isasymLongrightarrow}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\isanewline
\ \ wfrec{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ a{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharbrackleft}{\kern0pt}A{\isacharbrackright}{\kern0pt}{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ a{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}tr{\isacharunderscore}{\kern0pt}down{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ A{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}auto\ simp\ add\ {\isacharcolon}{\kern0pt}\ wfrec{\isacharunderscore}{\kern0pt}restr\ tr{\isacharunderscore}{\kern0pt}down{\isacharunderscore}{\kern0pt}def\ trancl{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}r{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}trancl\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ field{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ intro{\isacharcolon}{\kern0pt}\ r{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}trancl\ dest{\isacharbang}{\kern0pt}{\isacharcolon}{\kern0pt}\ trancl{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}THEN\ subsetD{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ Rrel\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}z{\isasymin}A{\isasymtimes}A{\isachardot}{\kern0pt}\ {\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymand}\ R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ RrelI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ y\ {\isasymin}\ A\ {\isasymLongrightarrow}\ R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}mem{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}mem{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Memrel{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ Memrel{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}Rrel{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ relation{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ \ d{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}mono\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isasymsubseteq}\ B\ {\isasymLongrightarrow}\ Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}restr{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasyminter}\ B{\isasymtimes}B\ {\isacharequal}{\kern0pt}\ Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}A{\isasyminter}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}Memrel\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}field{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ A{\isachardoublequoteclose}\isanewline
\ \ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}mem\ field{\isacharunderscore}{\kern0pt}Rrel\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ restrict{\isacharunderscore}{\kern0pt}trancl{\isacharunderscore}{\kern0pt}Rrel{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}R{\isacharparenleft}{\kern0pt}w{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}w\isanewline
\ \ \ \ \ \ \ {\isacharequal}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}w{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}y{\isasymin}d{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isacharquery}{\kern0pt}s{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}w{\isasymin}d{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}y{\isasymin}d{\isacartoucheclose}\ assms\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}w{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}r{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}w{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}s{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ r{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}trancl{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}r{\isacharbrackright}{\kern0pt}\ relation{\isacharunderscore}{\kern0pt}Rrel{\isacharbrackleft}{\kern0pt}of\ R\ d{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymlangle}w{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}r{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w{\isasymin}{\isacharquery}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}w{\isasymin}{\isacharquery}{\kern0pt}s{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ vimage{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w{\isasymnotin}domain{\isacharparenleft}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ subsetD{\isacharbrackleft}{\kern0pt}OF\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharbrackleft}{\kern0pt}of\ R\ d{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}w{\isasymnotin}d{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w{\isasymnotin}domain{\isacharparenleft}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}s{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ subsetD{\isacharbrackleft}{\kern0pt}OF\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharbrackleft}{\kern0pt}of\ R\ d{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ of\ w{\isacharbrackright}{\kern0pt}\ field{\isacharunderscore}{\kern0pt}trancl{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}r{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ fieldI{\isadigit{1}}{\isacharbrackleft}{\kern0pt}of\ w\ y\ {\isacharquery}{\kern0pt}s{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}w\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}s{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}w\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}d{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}s{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharequal}{\kern0pt}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w{\isasymnotin}{\isacharquery}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}y{\isasymnotin}d{\isacartoucheclose}\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y{\isasymnotin}field{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}s{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ field{\isacharunderscore}{\kern0pt}trancl\ subsetD{\isacharbrackleft}{\kern0pt}OF\ field{\isacharunderscore}{\kern0pt}Rrel{\isacharbrackleft}{\kern0pt}of\ R\ d{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w{\isasymnotin}{\isacharquery}{\kern0pt}s{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ vimage{\isacharunderscore}{\kern0pt}singleton{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}w{\isasymnotin}{\isacharquery}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ restrict{\isacharunderscore}{\kern0pt}trans{\isacharunderscore}{\kern0pt}eq{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}w\ {\isasymin}\ y{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}w\isanewline
\ \ \ \ \ \ \ {\isacharequal}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}w{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ restrict{\isacharunderscore}{\kern0pt}trancl{\isacharunderscore}{\kern0pt}Rrel{\isacharbrackleft}{\kern0pt}of\ mem\ {\isacharbrackright}{\kern0pt}\ Rrel{\isacharunderscore}{\kern0pt}mem\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}trancl{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}\ f\ y\ {\isachardot}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}R{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}R{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ {\isachardoublequoteopen}wfrec{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}R{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}wfrec{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ {\isasymlambda}y\ f{\isachardot}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ {\isasymlambda}y\ f{\isachardot}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ trancl{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}r{\isacharbrackleft}{\kern0pt}OF\ relation{\isacharunderscore}{\kern0pt}trancl\ trans{\isacharunderscore}{\kern0pt}trancl{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Recursion_Thms.thy%:%
%:%11=1%:%
%:%27=3%:%
%:%28=3%:%
%:%37=5%:%
%:%38=6%:%
%:%41=10%:%
%:%42=10%:%
%:%45=11%:%
%:%49=11%:%
%:%50=11%:%
%:%55=11%:%
%:%58=12%:%
%:%59=13%:%
%:%60=13%:%
%:%63=14%:%
%:%67=14%:%
%:%68=14%:%
%:%73=14%:%
%:%76=15%:%
%:%77=16%:%
%:%78=16%:%
%:%79=17%:%
%:%80=18%:%
%:%87=19%:%
%:%88=19%:%
%:%89=20%:%
%:%90=20%:%
%:%91=20%:%
%:%92=21%:%
%:%93=21%:%
%:%94=22%:%
%:%95=22%:%
%:%96=22%:%
%:%97=22%:%
%:%98=23%:%
%:%99=23%:%
%:%100=23%:%
%:%101=24%:%
%:%102=24%:%
%:%103=25%:%
%:%104=25%:%
%:%105=26%:%
%:%106=26%:%
%:%107=26%:%
%:%108=26%:%
%:%109=27%:%
%:%110=27%:%
%:%111=28%:%
%:%112=28%:%
%:%113=28%:%
%:%114=28%:%
%:%115=29%:%
%:%121=29%:%
%:%124=30%:%
%:%125=31%:%
%:%126=31%:%
%:%127=32%:%
%:%128=33%:%
%:%129=34%:%
%:%130=34%:%
%:%133=35%:%
%:%137=35%:%
%:%138=35%:%
%:%143=35%:%
%:%146=36%:%
%:%147=37%:%
%:%148=37%:%
%:%151=38%:%
%:%155=38%:%
%:%156=38%:%
%:%161=38%:%
%:%164=39%:%
%:%165=40%:%
%:%166=40%:%
%:%169=41%:%
%:%173=41%:%
%:%174=41%:%
%:%179=41%:%
%:%182=42%:%
%:%183=43%:%
%:%184=43%:%
%:%185=44%:%
%:%186=45%:%
%:%193=46%:%
%:%194=46%:%
%:%195=47%:%
%:%196=47%:%
%:%197=48%:%
%:%198=48%:%
%:%199=49%:%
%:%200=49%:%
%:%201=50%:%
%:%202=50%:%
%:%203=50%:%
%:%204=50%:%
%:%205=51%:%
%:%206=51%:%
%:%207=52%:%
%:%208=52%:%
%:%209=52%:%
%:%210=52%:%
%:%211=53%:%
%:%212=53%:%
%:%213=54%:%
%:%214=54%:%
%:%215=54%:%
%:%216=54%:%
%:%217=55%:%
%:%218=55%:%
%:%219=56%:%
%:%220=56%:%
%:%221=57%:%
%:%222=57%:%
%:%223=58%:%
%:%224=58%:%
%:%225=58%:%
%:%226=59%:%
%:%232=59%:%
%:%235=60%:%
%:%236=61%:%
%:%237=61%:%
%:%240=62%:%
%:%244=62%:%
%:%245=62%:%
%:%250=62%:%
%:%253=63%:%
%:%254=64%:%
%:%255=64%:%
%:%256=65%:%
%:%257=66%:%
%:%264=67%:%
%:%265=67%:%
%:%266=68%:%
%:%267=68%:%
%:%268=69%:%
%:%269=69%:%
%:%270=70%:%
%:%271=70%:%
%:%272=70%:%
%:%273=71%:%
%:%274=71%:%
%:%275=72%:%
%:%276=72%:%
%:%277=72%:%
%:%278=73%:%
%:%279=73%:%
%:%280=74%:%
%:%281=74%:%
%:%282=74%:%
%:%283=75%:%
%:%284=75%:%
%:%285=76%:%
%:%286=76%:%
%:%287=77%:%
%:%288=77%:%
%:%289=78%:%
%:%290=78%:%
%:%291=79%:%
%:%292=79%:%
%:%293=79%:%
%:%294=80%:%
%:%295=80%:%
%:%296=81%:%
%:%297=81%:%
%:%298=81%:%
%:%299=82%:%
%:%300=82%:%
%:%301=83%:%
%:%302=83%:%
%:%303=84%:%
%:%304=84%:%
%:%305=84%:%
%:%306=85%:%
%:%307=85%:%
%:%308=86%:%
%:%309=86%:%
%:%310=86%:%
%:%311=86%:%
%:%312=87%:%
%:%313=87%:%
%:%314=87%:%
%:%315=87%:%
%:%316=88%:%
%:%317=88%:%
%:%318=89%:%
%:%319=89%:%
%:%320=89%:%
%:%321=90%:%
%:%322=90%:%
%:%323=91%:%
%:%324=91%:%
%:%325=92%:%
%:%326=92%:%
%:%327=92%:%
%:%328=93%:%
%:%329=93%:%
%:%330=94%:%
%:%331=94%:%
%:%332=95%:%
%:%333=95%:%
%:%334=95%:%
%:%335=96%:%
%:%336=97%:%
%:%337=97%:%
%:%338=98%:%
%:%339=98%:%
%:%340=98%:%
%:%341=99%:%
%:%342=99%:%
%:%343=100%:%
%:%344=100%:%
%:%345=101%:%
%:%346=101%:%
%:%347=101%:%
%:%348=102%:%
%:%349=102%:%
%:%350=102%:%
%:%351=103%:%
%:%352=103%:%
%:%353=104%:%
%:%354=104%:%
%:%355=104%:%
%:%356=105%:%
%:%357=105%:%
%:%358=105%:%
%:%359=106%:%
%:%360=106%:%
%:%361=107%:%
%:%362=107%:%
%:%363=107%:%
%:%364=108%:%
%:%365=108%:%
%:%366=108%:%
%:%367=109%:%
%:%368=109%:%
%:%369=109%:%
%:%370=109%:%
%:%371=110%:%
%:%372=110%:%
%:%373=110%:%
%:%374=110%:%
%:%375=111%:%
%:%381=111%:%
%:%384=112%:%
%:%385=113%:%
%:%386=113%:%
%:%387=114%:%
%:%388=115%:%
%:%389=115%:%
%:%390=116%:%
%:%393=117%:%
%:%397=117%:%
%:%398=117%:%
%:%403=117%:%
%:%406=118%:%
%:%407=119%:%
%:%408=120%:%
%:%409=120%:%
%:%412=121%:%
%:%416=121%:%
%:%417=121%:%
%:%422=121%:%
%:%425=122%:%
%:%426=123%:%
%:%427=123%:%
%:%428=124%:%
%:%429=125%:%
%:%430=126%:%
%:%431=127%:%
%:%432=127%:%
%:%435=128%:%
%:%439=128%:%
%:%440=128%:%
%:%441=128%:%
%:%446=128%:%
%:%449=129%:%
%:%450=130%:%
%:%451=130%:%
%:%454=131%:%
%:%458=131%:%
%:%459=131%:%
%:%465=131%:%
%:%468=132%:%
%:%469=133%:%
%:%470=133%:%
%:%473=134%:%
%:%477=134%:%
%:%478=134%:%
%:%479=134%:%
%:%484=134%:%
%:%487=135%:%
%:%488=136%:%
%:%489=136%:%
%:%492=137%:%
%:%496=137%:%
%:%497=137%:%
%:%498=137%:%
%:%503=137%:%
%:%506=138%:%
%:%507=139%:%
%:%508=139%:%
%:%511=140%:%
%:%515=140%:%
%:%516=140%:%
%:%517=140%:%
%:%522=140%:%
%:%525=141%:%
%:%526=142%:%
%:%527=142%:%
%:%530=143%:%
%:%534=143%:%
%:%535=143%:%
%:%536=143%:%
%:%541=143%:%
%:%544=144%:%
%:%545=145%:%
%:%546=146%:%
%:%547=146%:%
%:%548=147%:%
%:%551=148%:%
%:%555=148%:%
%:%556=148%:%
%:%557=148%:%
%:%562=148%:%
%:%565=149%:%
%:%566=150%:%
%:%567=150%:%
%:%568=151%:%
%:%569=152%:%
%:%570=153%:%
%:%577=154%:%
%:%578=154%:%
%:%579=155%:%
%:%580=155%:%
%:%581=156%:%
%:%582=156%:%
%:%583=157%:%
%:%584=157%:%
%:%585=158%:%
%:%586=158%:%
%:%587=159%:%
%:%588=159%:%
%:%589=160%:%
%:%590=160%:%
%:%591=161%:%
%:%592=161%:%
%:%593=162%:%
%:%594=162%:%
%:%595=162%:%
%:%596=163%:%
%:%597=163%:%
%:%598=164%:%
%:%599=164%:%
%:%600=165%:%
%:%601=165%:%
%:%602=165%:%
%:%603=166%:%
%:%604=166%:%
%:%605=167%:%
%:%606=167%:%
%:%607=168%:%
%:%608=168%:%
%:%609=168%:%
%:%610=169%:%
%:%611=169%:%
%:%612=170%:%
%:%613=170%:%
%:%614=170%:%
%:%615=171%:%
%:%616=171%:%
%:%617=172%:%
%:%618=172%:%
%:%619=173%:%
%:%620=173%:%
%:%621=174%:%
%:%622=174%:%
%:%623=175%:%
%:%624=175%:%
%:%625=175%:%
%:%626=176%:%
%:%627=176%:%
%:%628=176%:%
%:%629=177%:%
%:%630=177%:%
%:%631=178%:%
%:%632=178%:%
%:%633=179%:%
%:%634=179%:%
%:%635=180%:%
%:%636=180%:%
%:%637=181%:%
%:%638=181%:%
%:%639=182%:%
%:%640=182%:%
%:%641=182%:%
%:%642=183%:%
%:%643=183%:%
%:%644=183%:%
%:%645=183%:%
%:%646=184%:%
%:%647=184%:%
%:%648=185%:%
%:%649=185%:%
%:%650=186%:%
%:%651=186%:%
%:%652=187%:%
%:%653=187%:%
%:%654=188%:%
%:%655=188%:%
%:%656=189%:%
%:%657=189%:%
%:%658=190%:%
%:%659=190%:%
%:%660=191%:%
%:%661=191%:%
%:%662=191%:%
%:%663=192%:%
%:%664=192%:%
%:%665=193%:%
%:%666=193%:%
%:%667=193%:%
%:%668=194%:%
%:%669=194%:%
%:%670=195%:%
%:%671=195%:%
%:%672=196%:%
%:%673=196%:%
%:%674=196%:%
%:%675=197%:%
%:%676=197%:%
%:%677=198%:%
%:%678=198%:%
%:%679=199%:%
%:%680=199%:%
%:%681=199%:%
%:%682=200%:%
%:%683=200%:%
%:%684=201%:%
%:%685=201%:%
%:%686=201%:%
%:%687=202%:%
%:%693=202%:%
%:%696=203%:%
%:%697=204%:%
%:%698=204%:%
%:%699=205%:%
%:%700=206%:%
%:%701=207%:%
%:%704=208%:%
%:%708=208%:%
%:%709=208%:%
%:%710=208%:%
%:%715=208%:%
%:%718=209%:%
%:%719=210%:%
%:%720=210%:%
%:%721=211%:%
%:%722=212%:%
%:%729=213%:%
%:%730=213%:%
%:%731=214%:%
%:%732=214%:%
%:%733=215%:%
%:%734=215%:%
%:%735=215%:%
%:%736=216%:%
%:%737=216%:%
%:%738=217%:%
%:%739=217%:%
%:%740=218%:%
%:%741=218%:%
%:%742=218%:%
%:%743=219%:%
%:%744=219%:%
%:%745=220%:%
%:%746=220%:%
%:%747=221%:%
%:%748=221%:%
%:%749=221%:%
%:%750=221%:%
%:%751=222%:%
%:%752=222%:%
%:%753=223%:%
%:%754=223%:%
%:%755=223%:%
%:%756=224%:%
%:%762=224%:%
%:%767=225%:%
%:%772=226%:%