### 2024-11-01 code/output/document/Relative_Univ.tex

%
\begin{isabellebody}%
\setisabellecontext{Relative{\isacharunderscore}{\kern0pt}Univ}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Relativization of the cumulative hierarchy%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Relative{\isacharunderscore}{\kern0pt}Univ\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ {\isachardoublequoteopen}ZF{\isacharminus}{\kern0pt}Constructible{\isachardot}{\kern0pt}Rank{\isachardoublequoteclose}\isanewline
\ \ \ \ Internalizations\isanewline
\ \ \ \ Recursion{\isacharunderscore}{\kern0pt}Thms\isanewline
\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}trivial{\isacharparenright}{\kern0pt}\ powerset{\isacharunderscore}{\kern0pt}abs{\isacharprime}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}powerset{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ y\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isasymin}Pow{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ powerset{\isacharunderscore}{\kern0pt}abs\ assms\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}inter{\isacharunderscore}{\kern0pt}Transset{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}b\ {\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}b\ {\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}trivial{\isacharparenright}{\kern0pt}\ family{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ {\isasymforall}x{\isasymin}A{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymrbrakk}\isanewline
\ \ \ \ \ \ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}{\isasymUnion}x{\isasymin}A{\isachardot}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ RepFun{\isacharunderscore}{\kern0pt}closed\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ HVfrom\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ A\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}x{\isachardot}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isasymin}Pow{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}powapply\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}fy{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}fy{\isacharparenright}{\kern0pt}\ {\isasymand}\ powerset{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}fy{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}HVfrom\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}h{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}U{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}R{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ \ union{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}U{\isacharcomma}{\kern0pt}h{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ {\isasymand}\ big{\isacharunderscore}{\kern0pt}union{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}R{\isacharcomma}{\kern0pt}U{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}R{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}Vfrom\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}transrec{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}Vset\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Vset{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}z{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ empty{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Formula synthesis%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}f{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}{\isasymin}A{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}ipa{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}Collect{\isacharunderscore}{\kern0pt}def\ powerset{\isacharunderscore}{\kern0pt}def\ subset{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}closed\ assms\isanewline
\ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ \ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ ff{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ yy{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ zz{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}{\isasymin}A{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ nat{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}y\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}z\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}ff{\isacharcomma}{\kern0pt}yy{\isacharcomma}{\kern0pt}zz{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}is{\isacharunderscore}{\kern0pt}one{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ {\isacartoucheopen}nth{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ ff{\isacartoucheclose}{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}nth{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ yy{\isacartoucheclose}{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ {\isacartoucheopen}nth{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ zz{\isacartoucheclose}{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ Hrank\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Hrank{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}x{\isachardot}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ PHrank\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}fy{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ fun{\isacharunderscore}{\kern0pt}apply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}fy{\isacharparenright}{\kern0pt}\ {\isasymand}\ successor{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}fy{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}Hrank\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Hrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}hc{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharparenleft}{\kern0pt}{\isasymexists}R{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ big{\isacharunderscore}{\kern0pt}union{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}R{\isacharcomma}{\kern0pt}hc{\isacharparenright}{\kern0pt}\ {\isasymand}is{\isacharunderscore}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}R{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ rrank\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}rrank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}eclose{\isacharparenright}{\kern0pt}\ wf{\isacharunderscore}{\kern0pt}rrank\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ wf{\isacharparenleft}{\kern0pt}rrank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rrank{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ wf{\isacharunderscore}{\kern0pt}trancl{\isacharbrackleft}{\kern0pt}OF\ wf{\isacharunderscore}{\kern0pt}Memrel{\isacharbrackright}{\kern0pt}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}eclose{\isacharparenright}{\kern0pt}\ trans{\isacharunderscore}{\kern0pt}rrank\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ trans{\isacharparenleft}{\kern0pt}rrank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rrank{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ trans{\isacharunderscore}{\kern0pt}trancl\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}eclose{\isacharparenright}{\kern0pt}\ relation{\isacharunderscore}{\kern0pt}rrank\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ relation{\isacharparenleft}{\kern0pt}rrank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rrank{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}trancl\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}eclose{\isacharparenright}{\kern0pt}\ rrank{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}rrank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rrank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Absoluteness results%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{locale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}pow\ {\isacharequal}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}eclose\ {\isacharplus}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ power{\isacharunderscore}{\kern0pt}ax\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}power{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ powapply{\isacharunderscore}{\kern0pt}replacement\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ HVfrom{\isacharunderscore}{\kern0pt}replacement\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ transrec{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ PHrank{\isacharunderscore}{\kern0pt}replacement\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ is{\isacharunderscore}{\kern0pt}Hrank{\isacharunderscore}{\kern0pt}replacement\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ wfrec{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hrank{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rrank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isasymrbrakk}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ z\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlbrakk}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}h{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ is{\isacharunderscore}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}h{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}R{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ h\ {\isacharequal}{\kern0pt}\ A\ {\isasymunion}\ {\isasymUnion}R\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharcomma}{\kern0pt}\ R{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}abs\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HVfrom{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Replace{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}R{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ R{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ R\ {\isacharequal}{\kern0pt}\ Replace{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}univalent{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x\ y{\isachardot}{\kern0pt}\ {\isasymlbrakk}\ x{\isasymin}A{\isacharsemicolon}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}R{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ Replace{\isacharunderscore}{\kern0pt}abs\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ powapply{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}closed\ power{\isacharunderscore}{\kern0pt}ax\ \isacommand{unfolding}\isamarkupfalse%
\ power{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ RepFun{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}R{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ {\isachardoublequoteopen}Replace{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ RepFun{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}y{\isachardot}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}y\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ A{\isacharcomma}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ y\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}y\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ A{\isacharcomma}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ powapply{\isacharunderscore}{\kern0pt}closed\ transM{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ A{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ A{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}abs\ transM{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ A{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ RepFun{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}univalent{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}A\ {\isacharsemicolon}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}powapply{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ powapply{\isacharunderscore}{\kern0pt}replacement\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}x{\isachardot}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}a{\isasymin}Pow{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isasymin}Pow{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}y{\isasymin}x{\isachardoublequoteclose}\ \isakeyword{for}\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ assms\ transM{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}\ powapply{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}a{\isasymin}Pow{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ y{\isasymin}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ transM{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}\ \ RepFun{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}closed\ RepFun{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}HVfrom{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ relation{\isadigit{2}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}HVfrom{\isacharunderscore}{\kern0pt}def\ HVfrom{\isacharunderscore}{\kern0pt}def\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Replace{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply\ RepFun{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}powapply\ \isanewline
\ \ \ \ \ \ \ \ \ \ Union{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}closed\ RepFun{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ HVfrom{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}x{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymforall}g{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ M{\isacharparenleft}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ HVfrom{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ transrec{\isacharunderscore}{\kern0pt}HVfrom{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ transrec{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ rule{\isacharcolon}{\kern0pt}trans{\isacharunderscore}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}step\ i{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ A\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}i{\isachardot}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymlambda}x{\isasymin}i{\isachardot}{\kern0pt}\ Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}transrec{\isacharbrackleft}{\kern0pt}OF\ Vfrom{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ of\ A\ i{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ A\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}i{\isachardot}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}A{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}i{\isachardot}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ A\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}i{\isachardot}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ intro{\isacharcolon}{\kern0pt}transM{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ A\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}i{\isachardot}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isasymin}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}z{\isasymin}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}y{\isasymin}i{\isachardoublequoteclose}\ \isakeyword{for}\ y\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ intro{\isacharcolon}{\kern0pt}transM{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ step\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ A\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}i{\isachardot}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Pow{\isacharparenleft}{\kern0pt}transrec{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ transrec{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}transrec{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isasymlambda}y{\isachardot}{\kern0pt}\ transrec{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ i{\isacharcomma}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ HVfrom{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Vfrom{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}V{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ V\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Vfrom{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}HVfrom\ HVfrom{\isacharunderscore}{\kern0pt}closed\ HVfrom{\isacharunderscore}{\kern0pt}replacement\ \isanewline
\ \ \ \ transrec{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ i\ {\isachardoublequoteopen}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ transrec{\isacharunderscore}{\kern0pt}HVfrom\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Vfrom{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vfrom{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Vfrom{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}HVfrom\ HVfrom{\isacharunderscore}{\kern0pt}closed\ HVfrom{\isacharunderscore}{\kern0pt}replacement\ \isanewline
\ \ \ \ transrec{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ i\ {\isachardoublequoteopen}HVfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ transrec{\isacharunderscore}{\kern0pt}HVfrom\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Vset{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}V{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}Vset{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ V\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}Vset{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Vfrom{\isacharunderscore}{\kern0pt}abs\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Vset{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Vset{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vset{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Vfrom{\isacharunderscore}{\kern0pt}closed\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Vset{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Hrank{\isacharunderscore}{\kern0pt}trancl{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Hrank{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharequal}{\kern0pt}\ Hrank{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Hrank{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ restrict{\isacharunderscore}{\kern0pt}trans{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}trancl{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}rrank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hrank{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ wfrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}eclose{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hrank{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}def\ transrec{\isacharunderscore}{\kern0pt}def\ Hrank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ {\isasymlambda}y\ f{\isachardot}{\kern0pt}\ Hrank{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ wftrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ {\isasymlambda}y\ f{\isachardot}{\kern0pt}\ Hrank{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}y{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Hrank{\isacharunderscore}{\kern0pt}trancl\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ wfrec{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}r{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hrank{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ wfrec{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ trancl{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}r{\isacharbrackleft}{\kern0pt}OF\ relation{\isacharunderscore}{\kern0pt}trancl\ trans{\isacharunderscore}{\kern0pt}trancl{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ rrank{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ univ{\isacharunderscore}{\kern0pt}PHrank\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ univalent{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ PHrank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ PHrank{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymlbrakk}\ M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ z\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ PHrank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ PHrank{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ PHrank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Replace{\isacharunderscore}{\kern0pt}PHrank{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}hr{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}z{\isacharcomma}{\kern0pt}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}hr{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ hr\ {\isacharequal}{\kern0pt}\ Replace{\isacharparenleft}{\kern0pt}z{\isacharcomma}{\kern0pt}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x\ y{\isachardot}{\kern0pt}\ {\isasymlbrakk}x{\isasymin}z{\isacharsemicolon}{\kern0pt}\ PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{unfolding}\isamarkupfalse%
\ PHrank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}hr{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ univ{\isacharunderscore}{\kern0pt}PHrank\ Replace{\isacharunderscore}{\kern0pt}abs\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ RepFun{\isacharunderscore}{\kern0pt}PHrank{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}R{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ {\isachardoublequoteopen}Replace{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ RepFun{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}{\isasymlambda}y{\isachardot}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}z\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ A{\isacharcomma}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymand}\ z\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}z\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ A{\isacharcomma}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ PHrank{\isacharunderscore}{\kern0pt}closed\ transM{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ A{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ A{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ PHrank{\isacharunderscore}{\kern0pt}abs\ transM{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ A{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ RepFun{\isacharunderscore}{\kern0pt}PHrank{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}Replace{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlbrakk}\ x{\isasymin}A\ {\isacharsemicolon}{\kern0pt}\ PHrank{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ PHrank{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ univ{\isacharunderscore}{\kern0pt}PHrank\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ PHrank{\isacharunderscore}{\kern0pt}replacement\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}Hrank\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}relation{\isadigit{2}}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}Hrank{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Hrank{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Hrank{\isacharunderscore}{\kern0pt}def\ Hrank{\isacharunderscore}{\kern0pt}def\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Replace{\isacharunderscore}{\kern0pt}PHrank{\isacharunderscore}{\kern0pt}abs\ RepFun{\isacharunderscore}{\kern0pt}PHrank\ RepFun{\isacharunderscore}{\kern0pt}PHrank{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}PHrank{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}{\isasymUnion}y{\isasymin}x{\isachardot}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}y{\isasymin}x{\isachardoublequoteclose}\ \isakeyword{for}\ y\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ assms\ transM{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ y{\isasymin}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ transM{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}\ \ RepFun{\isacharunderscore}{\kern0pt}PHrank{\isacharunderscore}{\kern0pt}closed\ RepFun{\isacharunderscore}{\kern0pt}PHrank\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Hrank{\isacharunderscore}{\kern0pt}closed\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}x{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymforall}g{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ M{\isacharparenleft}{\kern0pt}Hrank{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Hrank{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ RepFun{\isacharunderscore}{\kern0pt}PHrank{\isacharunderscore}{\kern0pt}closed\ Union{\isacharunderscore}{\kern0pt}PHrank{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}trancl\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ relation{\isadigit{2}}{\isacharunderscore}{\kern0pt}Hrank\ is{\isacharunderscore}{\kern0pt}Hrank{\isacharunderscore}{\kern0pt}closed\ is{\isacharunderscore}{\kern0pt}Hrank{\isacharunderscore}{\kern0pt}replacement\ \isanewline
\ \ \ \ \ \ \ \ wf{\isacharunderscore}{\kern0pt}rrank\ relation{\isacharunderscore}{\kern0pt}rrank\ trans{\isacharunderscore}{\kern0pt}rrank\ rrank{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \ \ \ trans{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}rrank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ a\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}Hrank{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}Vset{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}i{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}V{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ ordinal{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymand}\ a{\isasymin}V{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}i{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}a{\isasymin}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vfrom{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}a{\isasymin}{\isacharquery}{\kern0pt}V{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Vset{\isacharunderscore}{\kern0pt}Ord{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}V{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Vfrom{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ordinal{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}Vfrom{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}i{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymand}\ a\ {\isasymin}\ {\isacharquery}{\kern0pt}V{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}rank\ Vfrom{\isacharunderscore}{\kern0pt}abs\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Relative_Univ.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%32=6%:%
%:%33=7%:%
%:%34=8%:%
%:%39=8%:%
%:%42=9%:%
%:%43=10%:%
%:%44=10%:%
%:%45=11%:%
%:%46=12%:%
%:%47=13%:%
%:%48=14%:%
%:%51=15%:%
%:%55=15%:%
%:%56=15%:%
%:%57=15%:%
%:%62=15%:%
%:%65=16%:%
%:%66=17%:%
%:%67=17%:%
%:%68=18%:%
%:%69=19%:%
%:%70=20%:%
%:%71=21%:%
%:%74=22%:%
%:%78=22%:%
%:%79=22%:%
%:%80=22%:%
%:%81=23%:%
%:%82=23%:%
%:%87=23%:%
%:%90=24%:%
%:%91=25%:%
%:%92=25%:%
%:%93=26%:%
%:%96=27%:%
%:%100=27%:%
%:%101=27%:%
%:%107=27%:%
%:%110=28%:%
%:%111=29%:%
%:%112=31%:%
%:%113=32%:%
%:%114=32%:%
%:%115=33%:%
%:%116=34%:%
%:%117=35%:%
%:%118=36%:%
%:%119=37%:%
%:%120=37%:%
%:%121=38%:%
%:%122=39%:%
%:%123=40%:%
%:%124=41%:%
%:%125=42%:%
%:%126=42%:%
%:%129=43%:%
%:%133=43%:%
%:%134=43%:%
%:%135=43%:%
%:%140=43%:%
%:%143=44%:%
%:%144=45%:%
%:%145=46%:%
%:%146=46%:%
%:%147=47%:%
%:%148=48%:%
%:%149=49%:%
%:%150=50%:%
%:%151=51%:%
%:%152=52%:%
%:%153=52%:%
%:%154=53%:%
%:%155=54%:%
%:%156=55%:%
%:%157=56%:%
%:%158=56%:%
%:%159=57%:%
%:%160=58%:%
%:%167=61%:%
%:%177=63%:%
%:%178=63%:%
%:%179=64%:%
%:%180=65%:%
%:%181=66%:%
%:%182=67%:%
%:%183=68%:%
%:%186=69%:%
%:%190=69%:%
%:%191=69%:%
%:%192=70%:%
%:%193=70%:%
%:%194=71%:%
%:%195=71%:%
%:%200=71%:%
%:%203=72%:%
%:%204=73%:%
%:%205=73%:%
%:%206=74%:%
%:%207=75%:%
%:%208=76%:%
%:%209=77%:%
%:%210=78%:%
%:%213=79%:%
%:%217=79%:%
%:%218=79%:%
%:%219=80%:%
%:%220=81%:%
%:%221=81%:%
%:%226=81%:%
%:%229=82%:%
%:%230=83%:%
%:%231=84%:%
%:%232=84%:%
%:%233=85%:%
%:%234=86%:%
%:%235=87%:%
%:%236=88%:%
%:%237=88%:%
%:%238=89%:%
%:%239=90%:%
%:%240=91%:%
%:%241=92%:%
%:%242=92%:%
%:%243=93%:%
%:%244=94%:%
%:%245=95%:%
%:%246=96%:%
%:%247=96%:%
%:%248=97%:%
%:%249=98%:%
%:%250=99%:%
%:%251=100%:%
%:%252=100%:%
%:%255=101%:%
%:%259=101%:%
%:%260=101%:%
%:%261=101%:%
%:%267=101%:%
%:%270=102%:%
%:%271=103%:%
%:%272=103%:%
%:%275=104%:%
%:%279=104%:%
%:%280=104%:%
%:%281=104%:%
%:%287=104%:%
%:%290=105%:%
%:%291=106%:%
%:%292=106%:%
%:%295=107%:%
%:%299=107%:%
%:%300=107%:%
%:%301=107%:%
%:%307=107%:%
%:%310=108%:%
%:%311=109%:%
%:%312=109%:%
%:%315=110%:%
%:%319=110%:%
%:%320=110%:%
%:%321=110%:%
%:%335=113%:%
%:%345=115%:%
%:%346=115%:%
%:%347=116%:%
%:%348=117%:%
%:%349=118%:%
%:%350=119%:%
%:%351=120%:%
%:%352=121%:%
%:%353=122%:%
%:%354=123%:%
%:%355=124%:%
%:%356=125%:%
%:%357=126%:%
%:%358=126%:%
%:%361=127%:%
%:%365=127%:%
%:%366=127%:%
%:%367=127%:%
%:%372=127%:%
%:%375=128%:%
%:%376=129%:%
%:%377=129%:%
%:%379=131%:%
%:%382=132%:%
%:%386=132%:%
%:%387=132%:%
%:%388=132%:%
%:%389=132%:%
%:%394=132%:%
%:%397=133%:%
%:%398=134%:%
%:%399=134%:%
%:%400=135%:%
%:%401=136%:%
%:%402=137%:%
%:%403=138%:%
%:%410=139%:%
%:%411=139%:%
%:%412=140%:%
%:%413=140%:%
%:%414=141%:%
%:%415=141%:%
%:%416=141%:%
%:%417=141%:%
%:%418=142%:%
%:%419=142%:%
%:%420=143%:%
%:%421=143%:%
%:%422=144%:%
%:%423=144%:%
%:%424=144%:%
%:%425=144%:%
%:%426=145%:%
%:%427=145%:%
%:%428=146%:%
%:%429=146%:%
%:%430=146%:%
%:%431=146%:%
%:%432=147%:%
%:%438=147%:%
%:%441=148%:%
%:%442=149%:%
%:%443=149%:%
%:%444=150%:%
%:%447=151%:%
%:%451=151%:%
%:%452=151%:%
%:%453=151%:%
%:%454=151%:%
%:%459=151%:%
%:%462=152%:%
%:%463=153%:%
%:%464=153%:%
%:%465=154%:%
%:%466=155%:%
%:%467=156%:%
%:%468=157%:%
%:%475=158%:%
%:%476=158%:%
%:%477=159%:%
%:%478=159%:%
%:%479=160%:%
%:%480=160%:%
%:%481=160%:%
%:%482=161%:%
%:%483=161%:%
%:%484=162%:%
%:%485=162%:%
%:%486=162%:%
%:%487=163%:%
%:%488=163%:%
%:%489=164%:%
%:%490=164%:%
%:%491=164%:%
%:%492=164%:%
%:%493=165%:%
%:%499=165%:%
%:%502=166%:%
%:%503=167%:%
%:%504=167%:%
%:%505=168%:%
%:%506=169%:%
%:%507=170%:%
%:%508=171%:%
%:%515=172%:%
%:%516=172%:%
%:%517=173%:%
%:%518=173%:%
%:%519=174%:%
%:%520=174%:%
%:%521=174%:%
%:%522=174%:%
%:%523=175%:%
%:%524=175%:%
%:%525=176%:%
%:%526=176%:%
%:%527=177%:%
%:%528=177%:%
%:%529=177%:%
%:%530=177%:%
%:%531=178%:%
%:%532=178%:%
%:%533=179%:%
%:%534=179%:%
%:%535=179%:%
%:%536=179%:%
%:%537=180%:%
%:%543=180%:%
%:%546=181%:%
%:%547=182%:%
%:%548=182%:%
%:%549=183%:%
%:%550=184%:%
%:%551=185%:%
%:%552=186%:%
%:%559=187%:%
%:%560=187%:%
%:%561=188%:%
%:%562=188%:%
%:%563=189%:%
%:%564=189%:%
%:%565=189%:%
%:%566=190%:%
%:%567=190%:%
%:%568=191%:%
%:%569=191%:%
%:%570=192%:%
%:%571=192%:%
%:%572=192%:%
%:%573=193%:%
%:%574=193%:%
%:%575=193%:%
%:%576=193%:%
%:%577=193%:%
%:%578=194%:%
%:%584=194%:%
%:%587=195%:%
%:%588=196%:%
%:%589=196%:%
%:%592=197%:%
%:%596=197%:%
%:%597=197%:%
%:%598=198%:%
%:%599=198%:%
%:%600=199%:%
%:%601=199%:%
%:%606=199%:%
%:%609=200%:%
%:%610=201%:%
%:%611=201%:%
%:%612=202%:%
%:%615=203%:%
%:%619=203%:%
%:%620=203%:%
%:%621=203%:%
%:%622=203%:%
%:%627=203%:%
%:%630=204%:%
%:%631=205%:%
%:%632=205%:%
%:%633=206%:%
%:%634=207%:%
%:%641=208%:%
%:%642=208%:%
%:%643=209%:%
%:%644=209%:%
%:%645=210%:%
%:%646=210%:%
%:%647=211%:%
%:%648=211%:%
%:%649=211%:%
%:%650=212%:%
%:%651=212%:%
%:%652=213%:%
%:%653=213%:%
%:%654=214%:%
%:%655=214%:%
%:%656=215%:%
%:%657=215%:%
%:%658=216%:%
%:%659=216%:%
%:%660=217%:%
%:%661=217%:%
%:%662=218%:%
%:%663=218%:%
%:%664=219%:%
%:%665=219%:%
%:%666=220%:%
%:%667=220%:%
%:%668=221%:%
%:%669=221%:%
%:%670=222%:%
%:%671=222%:%
%:%672=223%:%
%:%673=223%:%
%:%674=224%:%
%:%675=224%:%
%:%676=225%:%
%:%677=225%:%
%:%678=226%:%
%:%679=226%:%
%:%680=227%:%
%:%681=227%:%
%:%682=227%:%
%:%683=228%:%
%:%684=228%:%
%:%685=229%:%
%:%686=229%:%
%:%687=229%:%
%:%688=230%:%
%:%689=230%:%
%:%690=230%:%
%:%691=231%:%
%:%692=231%:%
%:%693=232%:%
%:%694=232%:%
%:%695=233%:%
%:%696=233%:%
%:%697=234%:%
%:%698=234%:%
%:%699=234%:%
%:%700=235%:%
%:%701=235%:%
%:%702=236%:%
%:%703=236%:%
%:%704=236%:%
%:%705=237%:%
%:%711=237%:%
%:%714=238%:%
%:%715=239%:%
%:%716=239%:%
%:%719=240%:%
%:%723=240%:%
%:%724=240%:%
%:%725=241%:%
%:%726=241%:%
%:%727=242%:%
%:%728=242%:%
%:%733=242%:%
%:%736=243%:%
%:%737=244%:%
%:%738=244%:%
%:%741=245%:%
%:%745=245%:%
%:%746=245%:%
%:%747=246%:%
%:%748=246%:%
%:%749=247%:%
%:%750=247%:%
%:%755=247%:%
%:%758=248%:%
%:%759=249%:%
%:%760=249%:%
%:%763=250%:%
%:%767=250%:%
%:%768=250%:%
%:%769=250%:%
%:%770=250%:%
%:%775=250%:%
%:%778=251%:%
%:%779=252%:%
%:%780=252%:%
%:%783=253%:%
%:%787=253%:%
%:%788=253%:%
%:%789=253%:%
%:%790=253%:%
%:%795=253%:%
%:%798=254%:%
%:%799=255%:%
%:%800=255%:%
%:%801=256%:%
%:%804=257%:%
%:%808=257%:%
%:%809=257%:%
%:%810=258%:%
%:%811=258%:%
%:%812=258%:%
%:%817=258%:%
%:%820=259%:%
%:%821=260%:%
%:%822=260%:%
%:%829=261%:%
%:%830=261%:%
%:%831=262%:%
%:%832=262%:%
%:%833=263%:%
%:%834=264%:%
%:%835=264%:%
%:%836=264%:%
%:%837=265%:%
%:%838=265%:%
%:%839=266%:%
%:%840=266%:%
%:%841=267%:%
%:%842=267%:%
%:%843=267%:%
%:%844=268%:%
%:%845=268%:%
%:%846=269%:%
%:%847=269%:%
%:%848=270%:%
%:%849=270%:%
%:%850=270%:%
%:%851=271%:%
%:%852=271%:%
%:%853=272%:%
%:%854=272%:%
%:%855=273%:%
%:%856=273%:%
%:%857=273%:%
%:%858=273%:%
%:%859=274%:%
%:%860=274%:%
%:%861=275%:%
%:%862=275%:%
%:%863=275%:%
%:%864=275%:%
%:%865=276%:%
%:%871=276%:%
%:%874=277%:%
%:%875=278%:%
%:%876=278%:%
%:%879=279%:%
%:%883=279%:%
%:%884=279%:%
%:%885=279%:%
%:%890=279%:%
%:%893=280%:%
%:%894=281%:%
%:%895=282%:%
%:%896=282%:%
%:%897=283%:%
%:%900=284%:%
%:%904=284%:%
%:%905=284%:%
%:%906=284%:%
%:%911=284%:%
%:%914=285%:%
%:%915=286%:%
%:%916=286%:%
%:%919=287%:%
%:%923=287%:%
%:%924=287%:%
%:%925=287%:%
%:%930=287%:%
%:%933=288%:%
%:%934=289%:%
%:%935=289%:%
%:%936=290%:%
%:%937=291%:%
%:%938=292%:%
%:%939=293%:%
%:%946=294%:%
%:%947=294%:%
%:%948=295%:%
%:%949=295%:%
%:%950=296%:%
%:%951=296%:%
%:%952=296%:%
%:%953=296%:%
%:%954=297%:%
%:%955=297%:%
%:%956=298%:%
%:%957=298%:%
%:%958=298%:%
%:%959=298%:%
%:%960=299%:%
%:%966=299%:%
%:%969=300%:%
%:%970=301%:%
%:%971=301%:%
%:%972=302%:%
%:%973=303%:%
%:%974=304%:%
%:%975=305%:%
%:%982=306%:%
%:%983=306%:%
%:%984=307%:%
%:%985=307%:%
%:%986=308%:%
%:%987=308%:%
%:%988=308%:%
%:%989=309%:%
%:%990=309%:%
%:%991=310%:%
%:%992=310%:%
%:%993=310%:%
%:%994=311%:%
%:%995=311%:%
%:%996=312%:%
%:%997=312%:%
%:%998=312%:%
%:%999=312%:%
%:%1000=313%:%
%:%1006=313%:%
%:%1009=314%:%
%:%1010=315%:%
%:%1011=315%:%
%:%1012=316%:%
%:%1013=317%:%
%:%1014=318%:%
%:%1015=319%:%
%:%1022=320%:%
%:%1023=320%:%
%:%1024=321%:%
%:%1025=321%:%
%:%1026=322%:%
%:%1027=322%:%
%:%1028=322%:%
%:%1029=322%:%
%:%1030=323%:%
%:%1031=323%:%
%:%1032=324%:%
%:%1033=324%:%
%:%1034=324%:%
%:%1035=324%:%
%:%1036=325%:%
%:%1042=325%:%
%:%1045=326%:%
%:%1046=327%:%
%:%1047=327%:%
%:%1048=328%:%
%:%1051=329%:%
%:%1055=329%:%
%:%1056=329%:%
%:%1057=330%:%
%:%1058=330%:%
%:%1059=330%:%
%:%1064=330%:%
%:%1067=331%:%
%:%1068=332%:%
%:%1069=333%:%
%:%1070=333%:%
%:%1071=334%:%
%:%1072=335%:%
%:%1073=336%:%
%:%1074=337%:%
%:%1081=338%:%
%:%1082=338%:%
%:%1083=339%:%
%:%1084=339%:%
%:%1085=340%:%
%:%1086=340%:%
%:%1087=340%:%
%:%1088=341%:%
%:%1089=341%:%
%:%1090=342%:%
%:%1091=342%:%
%:%1092=343%:%
%:%1093=343%:%
%:%1094=343%:%
%:%1095=344%:%
%:%1096=344%:%
%:%1097=344%:%
%:%1098=344%:%
%:%1099=344%:%
%:%1100=345%:%
%:%1106=345%:%
%:%1109=346%:%
%:%1110=347%:%
%:%1111=347%:%
%:%1112=348%:%
%:%1115=349%:%
%:%1119=349%:%
%:%1120=349%:%
%:%1121=349%:%
%:%1122=349%:%
%:%1127=349%:%
%:%1130=350%:%
%:%1131=351%:%
%:%1132=351%:%
%:%1135=352%:%
%:%1139=352%:%
%:%1140=352%:%
%:%1141=353%:%
%:%1142=353%:%
%:%1143=354%:%
%:%1144=355%:%
%:%1145=355%:%
%:%1150=355%:%
%:%1153=356%:%
%:%1154=357%:%
%:%1155=358%:%
%:%1156=358%:%
%:%1157=359%:%
%:%1158=360%:%
%:%1165=361%:%
%:%1166=361%:%
%:%1167=362%:%
%:%1168=362%:%
%:%1169=363%:%
%:%1170=363%:%
%:%1171=364%:%
%:%1172=364%:%
%:%1173=365%:%
%:%1174=365%:%
%:%1175=365%:%
%:%1176=366%:%
%:%1177=366%:%
%:%1178=366%:%
%:%1179=367%:%
%:%1180=367%:%
%:%1181=368%:%
%:%1182=368%:%
%:%1183=368%:%
%:%1184=369%:%
%:%1185=369%:%
%:%1186=370%:%
%:%1187=370%:%
%:%1188=371%:%
%:%1189=371%:%
%:%1190=371%:%
%:%1191=372%:%
%:%1192=372%:%
%:%1193=373%:%
%:%1194=373%:%
%:%1195=373%:%
%:%1196=374%:%
%:%1197=374%:%
%:%1198=374%:%
%:%1199=375%:%
%:%1200=375%:%
%:%1201=376%:%
%:%1202=376%:%
%:%1203=376%:%
%:%1204=377%:%
%:%1205=377%:%
%:%1206=378%:%
%:%1207=378%:%
%:%1208=378%:%
%:%1209=379%:%
%:%1215=379%:%
%:%1218=380%:%
%:%1219=381%:%
%:%1220=381%:%
%:%1227=382%:%