### 2024-11-01 code/output/document/Renaming.tex

%
\begin{isabellebody}%
\setisabellecontext{Renaming}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Renaming of variables in internalized formulas%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Renaming\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ Nat{\isacharunderscore}{\kern0pt}Miscellanea\isanewline
\ \ \ \ {\isachardoublequoteopen}ZF{\isacharminus}{\kern0pt}Constructible{\isachardot}{\kern0pt}Formula{\isachardoublequoteclose}\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ app{\isacharunderscore}{\kern0pt}nm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isasymin}n{\isasymrightarrow}m{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}x{\isasymin}n{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ apply{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ apply{\isacharunderscore}{\kern0pt}{\isadigit{0}}\ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fun\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Renaming of free variables%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ union{\isacharunderscore}{\kern0pt}fun\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}union{\isacharunderscore}{\kern0pt}fun{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymlambda}j\ {\isasymin}\ m\ {\isasymunion}\ p\ \ {\isachardot}{\kern0pt}\ if\ j{\isasymin}m\ then\ f{\isacharbackquote}{\kern0pt}j\ else\ g{\isacharbackquote}{\kern0pt}j{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}type{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}g\ {\isasymin}\ p\ {\isasymrightarrow}\ q{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}union{\isacharunderscore}{\kern0pt}fun{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ m\ {\isasymunion}\ p\ {\isasymrightarrow}\ n\ {\isasymunion}\ q{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}h{\isacharequal}{\kern0pt}{\isachardoublequoteopen}union{\isacharunderscore}{\kern0pt}fun{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ D{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n\ {\isasymunion}\ q{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x\ {\isasymin}\ m\ {\isasymunion}\ p{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}x\ {\isasymin}\ m{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ m\ {\isasymunion}\ p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def\ \ beta\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}f\ {\isasymin}\ m\ {\isasymrightarrow}\ n{\isacartoucheclose}\ {\isacartoucheopen}x{\isasymin}m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x\ {\isasymin}\ m\ {\isasymunion}\ p{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ p{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymnotin}m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ beta\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}g\ {\isasymin}\ p\ {\isasymrightarrow}\ q{\isacartoucheclose}\ {\isacartoucheopen}x{\isasymin}p{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ q{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}lam\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ x{\isasymin}\ {\isacharparenleft}{\kern0pt}m\ {\isasymunion}\ p{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharparenleft}{\kern0pt}n\ {\isasymunion}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}\ {\isacharquery}{\kern0pt}h{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ lamE{\isacharbrackleft}{\kern0pt}of\ x\ {\isachardoublequoteopen}m\ {\isasymunion}\ p{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}m\ {\isasymunion}\ p{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharparenleft}{\kern0pt}n\ {\isasymunion}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ D\ \isacommand{unfolding}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ B{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}h\ {\isasymsubseteq}\ {\isacharparenleft}{\kern0pt}m\ {\isasymunion}\ p{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharparenleft}{\kern0pt}n\ {\isasymunion}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m\ {\isasymunion}\ p\ {\isasymsubseteq}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}lam\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ A\ B\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ \ Pi{\isacharunderscore}{\kern0pt}iff\ {\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}action\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m\ {\isasymunion}\ p{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}\ i\ {\isachardot}{\kern0pt}\ i\ {\isasymin}\ m\ {\isasymlongrightarrow}\ \ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}\ j\ {\isachardot}{\kern0pt}\ j\ {\isasymin}\ p\ {\isasymlongrightarrow}\ nth{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}\ i\ {\isachardot}{\kern0pt}\ i\ {\isasymin}\ m\ {\isasymunion}\ p\ {\isasymlongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}union{\isacharunderscore}{\kern0pt}fun{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}h\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}union{\isacharunderscore}{\kern0pt}fun{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x\ {\isasymin}\ m\ {\isasymunion}\ p{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}x{\isasymin}m{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def\ \ beta\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\ {\isacartoucheopen}x{\isasymin}m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x\ {\isasymin}\ m\ {\isasymunion}\ p{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymnotin}m{\isachardoublequoteclose}\ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def\ beta\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\ {\isacartoucheopen}x{\isasymin}p{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}id{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ n\ {\isasymrightarrow}\ n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ id{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}action{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymAnd}\ j\ {\isachardot}{\kern0pt}\ j\ {\isacharless}{\kern0pt}\ n\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\ \isakeyword{for}\ j\ \isacommand{using}\isamarkupfalse%
\ that\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ sum\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymlambda}j\ {\isasymin}\ m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p\ \ {\isachardot}{\kern0pt}\ if\ j{\isacharless}{\kern0pt}m\ then\ f{\isacharbackquote}{\kern0pt}j\ else\ {\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}j{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}inl{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ m{\isasymrightarrow}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ m{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isasymle}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self{\isacharbrackleft}{\kern0pt}of\ m{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltI{\isacharbrackleft}{\kern0pt}of\ x\ m{\isacharbrackright}{\kern0pt}\ lt{\isacharunderscore}{\kern0pt}trans{\isadigit{2}}{\isacharbrackleft}{\kern0pt}of\ x\ m\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}inr{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}g{\isasymin}p{\isasymrightarrow}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}m\ {\isasymle}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ ltD\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ x{\isacharless}{\kern0pt}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymnot}\ x{\isacharless}{\kern0pt}m{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}action\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ m{\isasymrightarrow}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}g{\isasymin}p{\isasymrightarrow}q{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isadigit{2}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ m\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ j{\isachardot}{\kern0pt}\ j\ {\isacharless}{\kern0pt}\ p\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p\ {\isasymlongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}h\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}q{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isasymle}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymle}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymle}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self{\isacharbrackleft}{\kern0pt}of\ m{\isacharbrackright}{\kern0pt}\ \ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self{\isadigit{2}}{\isacharbrackleft}{\kern0pt}of\ n\ q{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}inverse{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}m{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ sum{\isacharunderscore}{\kern0pt}inl\ ltD\ apply{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ f\ m\ {\isacharunderscore}{\kern0pt}\ x{\isacharbrackright}{\kern0pt}\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharless}{\kern0pt}m{\isacartoucheclose}\ assms\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x{\isacharless}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltI\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{2}}\ {\isacartoucheopen}x{\isacharless}{\kern0pt}m{\isacartoucheclose}\ assms\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}append{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isadigit{2}}\ {\isacartoucheopen}x{\isacharless}{\kern0pt}m{\isacartoucheclose}\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}append{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isacharprime}{\kern0pt}{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}f{\isacharbackquote}{\kern0pt}x{\isacharless}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}f{\isacharbackquote}{\kern0pt}x\ {\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ x{\isacharbrackright}{\kern0pt}\ ltD\ {\isacartoucheopen}p{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isasymle}x{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymle}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}{\isasymnot}x{\isacharless}{\kern0pt}m{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymnot}x{\isacharless}{\kern0pt}m{\isacartoucheclose}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}{\isasymnot}\ x\ {\isacharless}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ sum{\isacharunderscore}{\kern0pt}inr\ that\ beta\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ assms\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}p{\isacharequal}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m\ {\isacharless}{\kern0pt}\ p{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}x{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isacartoucheclose}\ {\isacartoucheopen}m{\isasymle}x{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isasymin}p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}g{\isasymin}p{\isasymrightarrow}q{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymin}\ q{\isachardoublequoteclose}\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ q{\isachardoublequoteclose}\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltI\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharless}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymle}\ g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymnot}\ g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono{\isadigit{1}}{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ n{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}q{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self{\isadigit{2}}{\isacharbrackleft}{\kern0pt}of\ n{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ assms\ {\isacartoucheopen}{\isasymnot}\ x\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}env\ {\isacharat}{\kern0pt}\ env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharcomma}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}append{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ {\isacartoucheopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m\ {\isacharless}{\kern0pt}\ p{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ diff{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}inverse{\isadigit{2}}\ {\isacartoucheopen}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ nth{\isacharunderscore}{\kern0pt}append{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isacharprime}{\kern0pt}{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}{\isasymnot}\ g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}type\ \ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ m{\isasymrightarrow}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}g{\isasymin}p{\isasymrightarrow}q{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isacharparenleft}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}h\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}q{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isasymle}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymle}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymle}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self{\isacharbrackleft}{\kern0pt}of\ m{\isacharbrackright}{\kern0pt}\ \ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self{\isadigit{2}}{\isacharbrackleft}{\kern0pt}of\ n\ q{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}inverse{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ sum{\isacharunderscore}{\kern0pt}inl\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}f{\isasymin}m{\isasymrightarrow}n{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymle}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isacharless}{\kern0pt}\ n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}trans{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ \ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isasymle}x{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltI\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x{\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ sum{\isacharunderscore}{\kern0pt}inr\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ assms\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}p{\isacharequal}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m\ {\isacharless}{\kern0pt}\ p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}x{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isacartoucheclose}\ {\isacartoucheopen}m{\isasymle}x{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isasymin}p{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}g{\isasymin}p{\isasymrightarrow}q{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymin}\ q{\isachardoublequoteclose}\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}nat{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ q{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharless}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono{\isadigit{1}}{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}g{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ n{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}q{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{2}}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ \ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isasymle}x{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ D{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isasymle}x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ that\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{2}}\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}lam\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ x{\isasymin}\ {\isacharparenleft}{\kern0pt}m\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharparenleft}{\kern0pt}n\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}\ {\isacharquery}{\kern0pt}h{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ lamE{\isacharbrackleft}{\kern0pt}of\ x\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}m\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharparenleft}{\kern0pt}n\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ D\ \isacommand{unfolding}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ B{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharquery}{\kern0pt}h\ {\isasymsubseteq}\ {\isacharparenleft}{\kern0pt}m\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharparenleft}{\kern0pt}n\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ p\ {\isasymsubseteq}\ domain{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}lam\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ A\ B\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ \ Pi{\isacharunderscore}{\kern0pt}iff\ {\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}type{\isacharunderscore}{\kern0pt}id\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymrightarrow}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ length{\isacharunderscore}{\kern0pt}type\ id{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}type\ sum{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}type{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}aux{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ m{\isasymrightarrow}n{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isacharparenleft}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ id{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}type\ sum{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}action{\isacharunderscore}{\kern0pt}id\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymrightarrow}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymAnd}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}m\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}n\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}p\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ lenv\ {\isacharequal}{\kern0pt}\ id{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}action{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}p{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}env{\isadigit{1}}{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ lenv{\isacharunderscore}{\kern0pt}ty\ {\isacharequal}{\kern0pt}\ id{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}p{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ i\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}m{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}action{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}m{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}p{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}p{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}f{\isasymin}{\isacharquery}{\kern0pt}m{\isasymrightarrow}{\isacharquery}{\kern0pt}n{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \ \ lenv{\isacharunderscore}{\kern0pt}ty\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isacharprime}{\kern0pt}{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isacartoucheopen}env{\isadigit{1}}{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isadigit{1}}{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacharunderscore}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ \ assms{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ lenv\isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}i{\isacharless}{\kern0pt}{\isacharquery}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ {\isacharquery}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}m{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}action{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}aux\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ m{\isasymrightarrow}n{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ m{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ m\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymAnd}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}sum{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharprime}{\kern0pt}{\isacharat}{\kern0pt}env{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ length{\isacharunderscore}{\kern0pt}type\ id{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}type\ sum{\isacharunderscore}{\kern0pt}action{\isacharunderscore}{\kern0pt}id\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ sum{\isacharunderscore}{\kern0pt}id\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ sum{\isacharparenleft}{\kern0pt}{\isasymlambda}x{\isasymin}{\isadigit{1}}{\isachardot}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}id{\isadigit{0}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m{\isasymin}nat{\isasymLongrightarrow}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}m{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isadigit{0}}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}unfold\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}subst\ sum{\isacharunderscore}{\kern0pt}inl{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}idS\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isasymin}nat\ {\isasymLongrightarrow}\ q{\isasymin}nat\ {\isasymLongrightarrow}\ f{\isasymin}p{\isasymrightarrow}q\ {\isasymLongrightarrow}\ x\ {\isasymin}\ p\ {\isasymLongrightarrow}\ sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}unfold\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}subst\ sum{\isacharunderscore}{\kern0pt}inr{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ simp{\isacharunderscore}{\kern0pt}all\ add{\isacharcolon}{\kern0pt}ltI{\isacharcomma}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all\ add{\isacharcolon}{\kern0pt}\ app{\isacharunderscore}{\kern0pt}nm\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc{\isacharunderscore}{\kern0pt}aux\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}p\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ \ q\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ f\ {\isasymin}\ p\ {\isasymrightarrow}\ q\ {\isasymLongrightarrow}\ sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}p\ {\isasymrightarrow}\ {\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}q{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}unfold\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}rule\ sum{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ f\ {\isasymin}\ n\ {\isasymrightarrow}\ m\ {\isasymLongrightarrow}\ sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ssubst{\isacharbrackleft}{\kern0pt}of\ \ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymrightarrow}\ {\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ simp{\isacharcomma}{\kern0pt}rule\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc{\isacharunderscore}{\kern0pt}aux{\isacharcomma}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Renaming of formulas%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{consts}\isamarkupfalse%
\ \ \ ren\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\isanewline
\isacommand{primrec}\isamarkupfalse%
\isanewline
\ \ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymlambda}\ n\ {\isasymin}\ nat\ {\isachardot}{\kern0pt}\ {\isasymlambda}\ m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymlambda}f\ {\isasymin}\ n\ {\isasymrightarrow}\ m{\isachardot}{\kern0pt}\ Member\ {\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
{\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymlambda}\ n\ {\isasymin}\ nat\ {\isachardot}{\kern0pt}\ {\isasymlambda}\ m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymlambda}f\ {\isasymin}\ n\ {\isasymrightarrow}\ m{\isachardot}{\kern0pt}\ Equal\ {\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
{\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymlambda}\ n\ {\isasymin}\ nat\ {\isachardot}{\kern0pt}\ {\isasymlambda}\ m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymlambda}f\ {\isasymin}\ n\ {\isasymrightarrow}\ m{\isachardot}{\kern0pt}\ Nand\ {\isacharparenleft}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharcomma}{\kern0pt}\ ren{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
{\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymlambda}\ n\ {\isasymin}\ nat\ {\isachardot}{\kern0pt}\ {\isasymlambda}\ m\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ {\isasymlambda}f\ {\isasymin}\ n\ {\isasymrightarrow}\ m{\isachardot}{\kern0pt}\ Forall\ {\isacharparenleft}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}meml\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}l\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ l\ {\isasymLongrightarrow}\ x\ {\isasymin}\ l{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}rule\ subsetD{\isacharcomma}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharcomma}{\kern0pt}assumption{\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}memr\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}l\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ l\ {\isasymLongrightarrow}\ y\ {\isasymin}\ l{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}rule\ subsetD{\isacharcomma}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharcomma}{\kern0pt}assumption{\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}eql\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}l\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ l\ {\isasymLongrightarrow}\ x\ {\isasymin}\ l{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}rule\ subsetD{\isacharcomma}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharcomma}{\kern0pt}assumption{\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}eqr\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}l\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ l\ {\isasymLongrightarrow}\ y\ {\isasymin}\ l{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}rule\ subsetD{\isacharcomma}{\kern0pt}rule\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharcomma}{\kern0pt}assumption{\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ \ nand{\isacharunderscore}{\kern0pt}ar{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ q{\isasymin}formula\ {\isasymLongrightarrow}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymle}\ arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{1}}{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}simp{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ nand{\isacharunderscore}{\kern0pt}ar{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ q{\isasymin}formula\ {\isasymLongrightarrow}arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymle}\ arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}simp{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nand{\isacharunderscore}{\kern0pt}ar{\isadigit{1}}D\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ q{\isasymin}formula\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ n\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymle}\ n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ \ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{1}}{\isacharunderscore}{\kern0pt}le{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ nand{\isacharunderscore}{\kern0pt}ar{\isadigit{2}}D\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ q{\isasymin}formula\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ n\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymle}\ n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ \ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ Un{\isacharunderscore}{\kern0pt}upper{\isadigit{2}}{\isacharunderscore}{\kern0pt}le{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}tc\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\isanewline
\ \ {\isacharparenleft}{\kern0pt}{\isasymAnd}\ n\ m\ f\ {\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ f\ {\isasymin}\ n{\isasymrightarrow}m\ {\isasymLongrightarrow}\ \ ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f\ {\isasymin}\ formula{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ set{\isacharcolon}{\kern0pt}formula{\isacharcomma}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ app{\isacharunderscore}{\kern0pt}nm\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ren\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ {\isachardoublequoteopen}p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymAnd}\ n\ m\ f\ {\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ f\ {\isasymin}\ n{\isasymrightarrow}m\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymle}\ n\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isasymle}m{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct\ set{\isacharcolon}{\kern0pt}formula{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Member\ x\ y{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ m{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Member\ assms\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ arity{\isacharunderscore}{\kern0pt}meml\ apply{\isacharunderscore}{\kern0pt}funtype{\isacharcomma}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}arity{\isacharunderscore}{\kern0pt}memr\ apply{\isacharunderscore}{\kern0pt}funtype{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ Member\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt\ ltI{\isacharparenright}{\kern0pt}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Equal\ x\ y{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ m{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Equal\ assms\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ arity{\isacharunderscore}{\kern0pt}eql\ apply{\isacharunderscore}{\kern0pt}funtype{\isacharcomma}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}arity{\isacharunderscore}{\kern0pt}eqr\ apply{\isacharunderscore}{\kern0pt}funtype{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ Equal\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt\ ltI{\isacharparenright}{\kern0pt}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Nand\ p\ q{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isasymle}arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isasymle}arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ \ nand{\isacharunderscore}{\kern0pt}ar{\isadigit{1}}{\isacharcomma}{\kern0pt}simp{\isacharcomma}{\kern0pt}simp{\isacharcomma}{\kern0pt}simp{\isacharcomma}{\kern0pt}subst\ nand{\isacharunderscore}{\kern0pt}ar{\isadigit{2}}{\isacharcomma}{\kern0pt}simp{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isasymle}n{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isasymle}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Nand\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}simp{\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymle}\ m{\isachardoublequoteclose}\ \isakeyword{and}\ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}ren{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymle}\ m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Nand\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ Nand\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Forall\ p{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ Forall\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ Forall\ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymrightarrow}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ Forall\ \isacommand{have}\isamarkupfalse%
\ {\isadigit{3}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}simp{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymle}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ Forall\ {\isacartoucheopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isasymin}nat{\isacartoucheclose}\ {\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ Forall\ {\isadigit{2}}\ {\isadigit{3}}\ ren{\isacharunderscore}{\kern0pt}tc\ arity{\isacharunderscore}{\kern0pt}type\ pred{\isacharunderscore}{\kern0pt}le\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}forallE\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ formula\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ m\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ natE{\isacharcomma}{\kern0pt}erule\ arity{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}simp{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ env{\isacharunderscore}{\kern0pt}coincidence{\isacharunderscore}{\kern0pt}sum{\isacharunderscore}{\kern0pt}id\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymrho}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymrho}{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}f\ {\isasymin}\ n\ {\isasymrightarrow}\ m{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ n\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}a\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}j{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}g{\isacharequal}{\kern0pt}{\isachardoublequoteopen}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}j{\isasymin}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}g{\isacharbackquote}{\kern0pt}j{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ rule{\isacharcolon}{\kern0pt}natE{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}j{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ sum{\isacharunderscore}{\kern0pt}id{\isadigit{0}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}\ i{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}j{\isasymin}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isasymin}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}succD\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}i{\isasymin}m{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}f{\isasymin}n{\isasymrightarrow}m{\isacartoucheclose}\ apply{\isacharunderscore}{\kern0pt}type\ {\isacartoucheopen}i{\isasymin}n{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ {\isacartoucheopen}i{\isasymin}n{\isacartoucheclose}\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}f{\isacharbackquote}{\kern0pt}i{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}g{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ sum{\isacharunderscore}{\kern0pt}idS{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ \ {\isacartoucheopen}f{\isasymin}n{\isasymrightarrow}m{\isacartoucheclose}\ {\isacartoucheopen}i\ {\isasymin}\ n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ cases\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}g{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}j{\isacharequal}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ren\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ {\isachardoublequoteopen}{\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ \ n\ {\isasymin}\ nat\ {\isacharsemicolon}{\kern0pt}\ m\ {\isasymin}\ nat\ {\isacharsemicolon}{\kern0pt}\ {\isasymrho}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ {\isasymrho}{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isacharsemicolon}{\kern0pt}\ f\ {\isasymin}\ n\ {\isasymrightarrow}\ m\ {\isacharsemicolon}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ n\ {\isacharsemicolon}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isasymAnd}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ n\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}\ {\isasymin}\ formula{\isacartoucheclose}\isanewline
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}induct\ {\isasymphi}\ arbitrary{\isacharcolon}{\kern0pt}n\ m\ {\isasymrho}\ {\isasymrho}{\isacharprime}{\kern0pt}\ f{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Member\ x\ y{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f\ {\isacharequal}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Member\ assms\ arity{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Member\ arity{\isacharunderscore}{\kern0pt}meml\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Member\ arity{\isacharunderscore}{\kern0pt}memr\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ Member\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Equal\ x\ y{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}Equal{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f\ {\isacharequal}{\kern0pt}\ Equal{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Equal\ assms\ arity{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Equal\ arity{\isacharunderscore}{\kern0pt}eql\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Equal\ arity{\isacharunderscore}{\kern0pt}eqr\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ Equal\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Nand\ p\ q{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}Nand{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f\ {\isacharequal}{\kern0pt}\ Nand{\isacharparenleft}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharcomma}{\kern0pt}ren{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Nand\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymle}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Nand\ nand{\isacharunderscore}{\kern0pt}ar{\isadigit{1}}D\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isasymin}\ arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ i\ {\isasymin}\ n{\isachardoublequoteclose}\ \isakeyword{for}\ i\ \isacommand{using}\isamarkupfalse%
\ subsetD{\isacharbrackleft}{\kern0pt}OF\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymle}\ n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isasymin}\ arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ i\ \isacommand{using}\isamarkupfalse%
\ Nand\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isasymle}n{\isacartoucheclose}\ Nand\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymle}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Nand\ nand{\isacharunderscore}{\kern0pt}ar{\isadigit{2}}D\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isasymin}\ arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ i\ {\isasymin}\ n{\isachardoublequoteclose}\ \isakeyword{for}\ i\ \isacommand{using}\isamarkupfalse%
\ subsetD{\isacharbrackleft}{\kern0pt}OF\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymle}\ n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isasymin}\ arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ i\ \isacommand{using}\isamarkupfalse%
\ Nand\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}q{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}ren{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}q{\isacharparenright}{\kern0pt}{\isasymle}n{\isacartoucheclose}\ Nand\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ Nand\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Forall\ p{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{0}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}ren{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n{\isacharbackquote}{\kern0pt}m{\isacharbackquote}{\kern0pt}f\ {\isacharequal}{\kern0pt}\ Forall{\isacharparenleft}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Forall\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}sum{\isacharunderscore}{\kern0pt}id{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}g\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ sum{\isacharunderscore}{\kern0pt}id{\isacharunderscore}{\kern0pt}tc\ Forall\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Forall\ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ succpred{\isacharunderscore}{\kern0pt}leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Forall\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}\ j\ {\isachardot}{\kern0pt}j\ {\isacharless}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ {\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}g{\isacharbackquote}{\kern0pt}j{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ {\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}a{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ a\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ env{\isacharunderscore}{\kern0pt}coincidence{\isacharunderscore}{\kern0pt}sum{\isacharunderscore}{\kern0pt}id\ Forall\ ltD\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharquery}{\kern0pt}g{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}a{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ a\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ C{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Forall\ that\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}ren{\isacharparenleft}{\kern0pt}p{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharquery}{\kern0pt}g{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Forall{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}succ{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isasymin}nat{\isacartoucheclose}\ C{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ C{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isadigit{1}}\ {\isadigit{2}}\ A{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}a{\isasymin}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{using}\isamarkupfalse%
\ Forall\ {\isadigit{0}}\ {\isadigit{1}}\ {\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Renaming.thy%:%
%:%11=1%:%
%:%27=3%:%
%:%28=3%:%
%:%29=4%:%
%:%30=5%:%
%:%31=6%:%
%:%32=7%:%
%:%37=7%:%
%:%40=8%:%
%:%41=9%:%
%:%42=9%:%
%:%43=10%:%
%:%44=11%:%
%:%51=12%:%
%:%52=12%:%
%:%53=13%:%
%:%54=13%:%
%:%55=14%:%
%:%56=14%:%
%:%57=14%:%
%:%58=14%:%
%:%59=14%:%
%:%60=15%:%
%:%61=15%:%
%:%62=16%:%
%:%63=16%:%
%:%64=17%:%
%:%65=17%:%
%:%66=17%:%
%:%67=17%:%
%:%68=17%:%
%:%69=18%:%
%:%84=20%:%
%:%94=22%:%
%:%95=22%:%
%:%96=23%:%
%:%97=24%:%
%:%98=25%:%
%:%99=26%:%
%:%100=26%:%
%:%101=27%:%
%:%102=28%:%
%:%103=29%:%
%:%110=30%:%
%:%111=30%:%
%:%112=31%:%
%:%113=31%:%
%:%114=32%:%
%:%115=32%:%
%:%116=33%:%
%:%117=34%:%
%:%118=34%:%
%:%119=35%:%
%:%120=35%:%
%:%121=36%:%
%:%122=36%:%
%:%123=36%:%
%:%124=37%:%
%:%125=37%:%
%:%126=38%:%
%:%127=38%:%
%:%128=39%:%
%:%129=39%:%
%:%130=40%:%
%:%131=40%:%
%:%132=40%:%
%:%133=41%:%
%:%134=41%:%
%:%135=42%:%
%:%136=42%:%
%:%137=42%:%
%:%138=43%:%
%:%139=43%:%
%:%140=43%:%
%:%141=43%:%
%:%142=44%:%
%:%143=44%:%
%:%144=45%:%
%:%145=45%:%
%:%146=46%:%
%:%147=46%:%
%:%148=47%:%
%:%149=47%:%
%:%150=48%:%
%:%151=48%:%
%:%152=49%:%
%:%153=49%:%
%:%154=50%:%
%:%155=50%:%
%:%156=51%:%
%:%157=51%:%
%:%158=51%:%
%:%159=51%:%
%:%160=52%:%
%:%161=52%:%
%:%162=53%:%
%:%163=53%:%
%:%164=53%:%
%:%165=54%:%
%:%166=54%:%
%:%167=54%:%
%:%168=54%:%
%:%169=55%:%
%:%170=55%:%
%:%171=56%:%
%:%172=56%:%
%:%173=56%:%
%:%174=56%:%
%:%175=56%:%
%:%176=57%:%
%:%177=57%:%
%:%178=58%:%
%:%179=58%:%
%:%180=58%:%
%:%181=59%:%
%:%182=59%:%
%:%183=60%:%
%:%184=60%:%
%:%185=60%:%
%:%186=60%:%
%:%187=61%:%
%:%188=61%:%
%:%189=62%:%
%:%190=62%:%
%:%191=62%:%
%:%192=62%:%
%:%193=63%:%
%:%194=63%:%
%:%195=64%:%
%:%196=64%:%
%:%197=64%:%
%:%198=64%:%
%:%199=65%:%
%:%205=65%:%
%:%208=66%:%
%:%209=67%:%
%:%210=67%:%
%:%211=68%:%
%:%212=69%:%
%:%213=70%:%
%:%214=71%:%
%:%215=72%:%
%:%216=73%:%
%:%217=74%:%
%:%218=75%:%
%:%225=76%:%
%:%226=76%:%
%:%227=77%:%
%:%228=77%:%
%:%229=78%:%
%:%230=78%:%
%:%231=79%:%
%:%232=79%:%
%:%233=80%:%
%:%234=80%:%
%:%235=81%:%
%:%236=81%:%
%:%237=82%:%
%:%238=82%:%
%:%239=83%:%
%:%240=83%:%
%:%241=84%:%
%:%242=84%:%
%:%243=84%:%
%:%244=85%:%
%:%245=85%:%
%:%246=86%:%
%:%247=86%:%
%:%248=86%:%
%:%249=87%:%
%:%250=87%:%
%:%251=87%:%
%:%252=87%:%
%:%253=88%:%
%:%254=88%:%
%:%255=89%:%
%:%256=89%:%
%:%257=90%:%
%:%258=90%:%
%:%259=91%:%
%:%260=91%:%
%:%261=92%:%
%:%262=92%:%
%:%263=93%:%
%:%264=93%:%
%:%265=94%:%
%:%266=94%:%
%:%267=95%:%
%:%268=95%:%
%:%269=95%:%
%:%270=96%:%
%:%271=96%:%
%:%272=97%:%
%:%273=97%:%
%:%274=97%:%
%:%275=98%:%
%:%276=98%:%
%:%277=98%:%
%:%278=98%:%
%:%279=99%:%
%:%280=99%:%
%:%281=100%:%
%:%282=100%:%
%:%283=100%:%
%:%284=100%:%
%:%285=101%:%
%:%291=101%:%
%:%294=102%:%
%:%295=103%:%
%:%296=104%:%
%:%297=104%:%
%:%298=105%:%
%:%299=106%:%
%:%302=107%:%
%:%306=107%:%
%:%307=107%:%
%:%308=107%:%
%:%309=107%:%
%:%314=107%:%
%:%317=108%:%
%:%318=109%:%
%:%319=109%:%
%:%320=110%:%
%:%321=111%:%
%:%328=112%:%
%:%329=112%:%
%:%330=113%:%
%:%331=113%:%
%:%332=113%:%
%:%333=113%:%
%:%334=114%:%
%:%340=114%:%
%:%343=115%:%
%:%344=116%:%
%:%345=117%:%
%:%346=117%:%
%:%347=118%:%
%:%348=119%:%
%:%349=120%:%
%:%350=121%:%
%:%351=121%:%
%:%352=122%:%
%:%353=123%:%
%:%354=124%:%
%:%361=125%:%
%:%362=125%:%
%:%363=126%:%
%:%364=126%:%
%:%365=127%:%
%:%366=127%:%
%:%367=128%:%
%:%368=128%:%
%:%369=128%:%
%:%370=129%:%
%:%371=129%:%
%:%372=130%:%
%:%373=130%:%
%:%374=131%:%
%:%375=131%:%
%:%376=131%:%
%:%377=132%:%
%:%378=132%:%
%:%379=133%:%
%:%380=133%:%
%:%381=134%:%
%:%382=134%:%
%:%383=134%:%
%:%384=135%:%
%:%385=135%:%
%:%386=136%:%
%:%387=136%:%
%:%388=136%:%
%:%389=136%:%
%:%390=137%:%
%:%396=137%:%
%:%399=138%:%
%:%400=139%:%
%:%401=139%:%
%:%402=140%:%
%:%403=141%:%
%:%404=142%:%
%:%411=143%:%
%:%412=143%:%
%:%413=144%:%
%:%414=144%:%
%:%415=145%:%
%:%416=145%:%
%:%417=146%:%
%:%418=146%:%
%:%419=147%:%
%:%420=147%:%
%:%421=148%:%
%:%422=148%:%
%:%423=149%:%
%:%424=149%:%
%:%425=150%:%
%:%426=150%:%
%:%427=150%:%
%:%428=151%:%
%:%429=151%:%
%:%430=152%:%
%:%431=152%:%
%:%432=153%:%
%:%433=153%:%
%:%434=153%:%
%:%435=154%:%
%:%436=154%:%
%:%437=155%:%
%:%438=155%:%
%:%439=155%:%
%:%440=155%:%
%:%441=156%:%
%:%447=156%:%
%:%450=157%:%
%:%451=158%:%
%:%452=159%:%
%:%453=159%:%
%:%454=160%:%
%:%455=161%:%
%:%456=162%:%
%:%457=163%:%
%:%458=164%:%
%:%459=165%:%
%:%460=166%:%
%:%461=167%:%
%:%462=168%:%
%:%463=169%:%
%:%464=170%:%
%:%465=171%:%
%:%466=172%:%
%:%473=173%:%
%:%474=173%:%
%:%475=174%:%
%:%476=174%:%
%:%477=175%:%
%:%478=175%:%
%:%479=176%:%
%:%480=176%:%
%:%481=177%:%
%:%482=177%:%
%:%483=177%:%
%:%484=178%:%
%:%485=178%:%
%:%486=179%:%
%:%487=179%:%
%:%488=179%:%
%:%489=179%:%
%:%490=180%:%
%:%491=180%:%
%:%492=181%:%
%:%493=181%:%
%:%494=182%:%
%:%495=182%:%
%:%496=183%:%
%:%497=183%:%
%:%498=184%:%
%:%499=184%:%
%:%500=185%:%
%:%501=185%:%
%:%502=185%:%
%:%503=186%:%
%:%504=186%:%
%:%505=187%:%
%:%506=187%:%
%:%507=188%:%
%:%508=188%:%
%:%509=188%:%
%:%510=189%:%
%:%511=189%:%
%:%512=190%:%
%:%513=190%:%
%:%514=191%:%
%:%515=191%:%
%:%516=191%:%
%:%517=192%:%
%:%518=192%:%
%:%519=193%:%
%:%520=193%:%
%:%521=194%:%
%:%522=195%:%
%:%523=195%:%
%:%524=195%:%
%:%525=196%:%
%:%526=196%:%
%:%527=197%:%
%:%528=197%:%
%:%529=198%:%
%:%530=198%:%
%:%531=198%:%
%:%532=199%:%
%:%533=199%:%
%:%534=200%:%
%:%535=200%:%
%:%536=201%:%
%:%537=201%:%
%:%538=201%:%
%:%539=202%:%
%:%540=202%:%
%:%541=203%:%
%:%542=203%:%
%:%543=203%:%
%:%544=204%:%
%:%545=204%:%
%:%546=204%:%
%:%547=204%:%
%:%548=205%:%
%:%549=205%:%
%:%550=206%:%
%:%551=206%:%
%:%552=207%:%
%:%553=207%:%
%:%554=208%:%
%:%555=208%:%
%:%556=208%:%
%:%557=209%:%
%:%558=209%:%
%:%559=210%:%
%:%560=210%:%
%:%561=211%:%
%:%562=211%:%
%:%563=211%:%
%:%564=212%:%
%:%565=212%:%
%:%566=213%:%
%:%567=213%:%
%:%568=214%:%
%:%569=214%:%
%:%570=215%:%
%:%571=215%:%
%:%572=215%:%
%:%573=216%:%
%:%574=216%:%
%:%575=217%:%
%:%576=217%:%
%:%577=218%:%
%:%578=218%:%
%:%579=218%:%
%:%580=219%:%
%:%581=219%:%
%:%582=219%:%
%:%583=219%:%
%:%584=219%:%
%:%585=220%:%
%:%586=220%:%
%:%587=221%:%
%:%588=221%:%
%:%589=221%:%
%:%590=222%:%
%:%591=222%:%
%:%592=223%:%
%:%593=223%:%
%:%594=223%:%
%:%595=223%:%
%:%596=224%:%
%:%597=224%:%
%:%598=225%:%
%:%599=225%:%
%:%600=226%:%
%:%601=226%:%
%:%602=227%:%
%:%603=228%:%
%:%604=228%:%
%:%605=229%:%
%:%606=229%:%
%:%607=230%:%
%:%608=230%:%
%:%609=231%:%
%:%610=231%:%
%:%611=231%:%
%:%612=232%:%
%:%613=232%:%
%:%614=233%:%
%:%615=233%:%
%:%616=234%:%
%:%617=234%:%
%:%618=234%:%
%:%619=235%:%
%:%620=235%:%
%:%621=236%:%
%:%622=236%:%
%:%623=237%:%
%:%624=237%:%
%:%625=238%:%
%:%626=239%:%
%:%627=239%:%
%:%628=240%:%
%:%629=240%:%
%:%630=241%:%
%:%631=241%:%
%:%632=242%:%
%:%633=242%:%
%:%634=243%:%
%:%635=243%:%
%:%636=244%:%
%:%637=244%:%
%:%638=245%:%
%:%639=245%:%
%:%640=246%:%
%:%641=246%:%
%:%642=246%:%
%:%643=247%:%
%:%644=247%:%
%:%645=248%:%
%:%646=248%:%
%:%647=248%:%
%:%648=249%:%
%:%649=249%:%
%:%650=249%:%
%:%651=249%:%
%:%652=250%:%
%:%653=250%:%
%:%654=251%:%
%:%655=251%:%
%:%656=251%:%
%:%657=251%:%
%:%658=252%:%
%:%664=252%:%
%:%667=253%:%
%:%668=254%:%
%:%669=254%:%
%:%670=255%:%
%:%671=256%:%
%:%672=257%:%
%:%679=258%:%
%:%680=258%:%
%:%681=259%:%
%:%682=259%:%
%:%683=260%:%
%:%684=260%:%
%:%685=261%:%
%:%686=261%:%
%:%687=262%:%
%:%688=262%:%
%:%689=262%:%
%:%690=263%:%
%:%691=263%:%
%:%692=264%:%
%:%693=264%:%
%:%694=264%:%
%:%695=264%:%
%:%696=265%:%
%:%697=265%:%
%:%698=265%:%
%:%699=266%:%
%:%700=266%:%
%:%701=267%:%
%:%702=267%:%
%:%703=267%:%
%:%704=268%:%
%:%705=268%:%
%:%706=268%:%
%:%707=269%:%
%:%708=269%:%
%:%709=270%:%
%:%710=270%:%
%:%711=270%:%
%:%712=271%:%
%:%713=271%:%
%:%714=271%:%
%:%715=271%:%
%:%716=271%:%
%:%717=272%:%
%:%718=272%:%
%:%719=273%:%
%:%720=273%:%
%:%721=273%:%
%:%722=273%:%
%:%723=274%:%
%:%724=274%:%
%:%725=275%:%
%:%726=275%:%
%:%727=275%:%
%:%728=275%:%
%:%729=276%:%
%:%730=276%:%
%:%731=277%:%
%:%732=277%:%
%:%733=277%:%
%:%734=277%:%
%:%735=277%:%
%:%736=278%:%
%:%737=278%:%
%:%738=278%:%
%:%739=279%:%
%:%740=279%:%
%:%741=280%:%
%:%742=280%:%
%:%743=280%:%
%:%744=280%:%
%:%745=280%:%
%:%746=281%:%
%:%747=281%:%
%:%748=282%:%
%:%749=282%:%
%:%750=283%:%
%:%751=283%:%
%:%752=283%:%
%:%753=284%:%
%:%754=284%:%
%:%755=285%:%
%:%756=285%:%
%:%757=285%:%
%:%758=285%:%
%:%759=286%:%
%:%760=286%:%
%:%761=286%:%
%:%762=286%:%
%:%763=286%:%
%:%764=287%:%
%:%765=287%:%
%:%766=288%:%
%:%767=288%:%
%:%768=288%:%
%:%769=289%:%
%:%770=289%:%
%:%771=289%:%
%:%772=289%:%
%:%773=289%:%
%:%774=290%:%
%:%775=290%:%
%:%776=291%:%
%:%777=291%:%
%:%778=291%:%
%:%779=291%:%
%:%780=292%:%
%:%781=292%:%
%:%782=293%:%
%:%783=293%:%
%:%784=293%:%
%:%785=293%:%
%:%786=294%:%
%:%787=294%:%
%:%788=295%:%
%:%789=295%:%
%:%790=295%:%
%:%791=295%:%
%:%792=295%:%
%:%793=296%:%
%:%794=296%:%
%:%795=297%:%
%:%796=298%:%
%:%797=298%:%
%:%798=299%:%
%:%799=299%:%
%:%800=300%:%
%:%801=300%:%
%:%802=301%:%
%:%803=301%:%
%:%804=301%:%
%:%805=301%:%
%:%806=301%:%
%:%807=302%:%
%:%808=302%:%
%:%809=303%:%
%:%810=303%:%
%:%811=304%:%
%:%812=304%:%
%:%813=304%:%
%:%814=304%:%
%:%815=304%:%
%:%816=305%:%
%:%817=305%:%
%:%818=305%:%
%:%819=305%:%
%:%820=305%:%
%:%821=306%:%
%:%822=306%:%
%:%823=307%:%
%:%824=307%:%
%:%825=307%:%
%:%826=307%:%
%:%827=307%:%
%:%828=308%:%
%:%829=308%:%
%:%830=309%:%
%:%831=309%:%
%:%832=309%:%
%:%833=310%:%
%:%834=310%:%
%:%835=311%:%
%:%836=311%:%
%:%837=311%:%
%:%838=311%:%
%:%839=312%:%
%:%840=312%:%
%:%841=313%:%
%:%842=313%:%
%:%843=313%:%
%:%844=313%:%
%:%845=314%:%
%:%846=314%:%
%:%847=315%:%
%:%848=315%:%
%:%849=315%:%
%:%850=315%:%
%:%851=316%:%
%:%857=316%:%
%:%860=317%:%
%:%861=318%:%
%:%862=318%:%
%:%863=319%:%
%:%864=320%:%
%:%865=321%:%
%:%866=322%:%
%:%867=323%:%
%:%868=324%:%
%:%869=325%:%
%:%870=326%:%
%:%873=327%:%
%:%877=327%:%
%:%878=327%:%
%:%879=328%:%
%:%880=328%:%
%:%885=328%:%
%:%888=329%:%
%:%889=330%:%
%:%890=330%:%
%:%891=331%:%
%:%892=332%:%
%:%893=333%:%
%:%894=334%:%
%:%895=335%:%
%:%896=336%:%
%:%897=337%:%
%:%900=338%:%
%:%904=338%:%
%:%905=338%:%
%:%906=339%:%
%:%907=339%:%
%:%912=339%:%
%:%915=340%:%
%:%916=341%:%
%:%917=341%:%
%:%918=342%:%
%:%919=343%:%
%:%920=344%:%
%:%921=345%:%
%:%922=346%:%
%:%923=347%:%
%:%924=348%:%
%:%925=349%:%
%:%932=350%:%
%:%933=350%:%
%:%934=351%:%
%:%935=351%:%
%:%936=352%:%
%:%937=352%:%
%:%938=352%:%
%:%939=353%:%
%:%940=353%:%
%:%941=353%:%
%:%942=353%:%
%:%943=354%:%
%:%944=354%:%
%:%945=354%:%
%:%946=354%:%
%:%947=355%:%
%:%948=355%:%
%:%949=356%:%
%:%950=356%:%
%:%951=357%:%
%:%952=357%:%
%:%953=358%:%
%:%954=358%:%
%:%955=359%:%
%:%956=359%:%
%:%957=360%:%
%:%958=360%:%
%:%959=361%:%
%:%960=361%:%
%:%961=362%:%
%:%962=363%:%
%:%963=364%:%
%:%964=365%:%
%:%965=365%:%
%:%966=366%:%
%:%967=366%:%
%:%968=367%:%
%:%969=367%:%
%:%970=367%:%
%:%971=368%:%
%:%972=368%:%
%:%973=369%:%
%:%979=369%:%
%:%982=370%:%
%:%983=371%:%
%:%984=371%:%
%:%985=372%:%
%:%986=373%:%
%:%987=374%:%
%:%988=375%:%
%:%989=376%:%
%:%990=377%:%
%:%991=378%:%
%:%992=379%:%
%:%993=380%:%
%:%994=381%:%
%:%995=382%:%
%:%998=383%:%
%:%1002=383%:%
%:%1003=383%:%
%:%1004=384%:%
%:%1005=384%:%
%:%1010=384%:%
%:%1013=385%:%
%:%1014=386%:%
%:%1015=387%:%
%:%1016=387%:%
%:%1017=388%:%
%:%1018=389%:%
%:%1019=390%:%
%:%1020=391%:%
%:%1021=391%:%
%:%1024=392%:%
%:%1028=392%:%
%:%1029=392%:%
%:%1034=392%:%
%:%1037=393%:%
%:%1038=394%:%
%:%1039=394%:%
%:%1042=395%:%
%:%1046=395%:%
%:%1047=395%:%
%:%1048=396%:%
%:%1053=396%:%
%:%1056=397%:%
%:%1057=398%:%
%:%1058=398%:%
%:%1059=399%:%
%:%1062=400%:%
%:%1066=400%:%
%:%1067=400%:%
%:%1072=400%:%
%:%1075=401%:%
%:%1076=402%:%
%:%1077=402%:%
%:%1078=403%:%
%:%1081=404%:%
%:%1085=404%:%
%:%1086=404%:%
%:%1087=405%:%
%:%1101=407%:%
%:%1111=409%:%
%:%1112=409%:%
%:%1113=410%:%
%:%1114=410%:%
%:%1115=411%:%
%:%1116=412%:%
%:%1117=413%:%
%:%1118=414%:%
%:%1119=415%:%
%:%1120=416%:%
%:%1121=417%:%
%:%1122=418%:%
%:%1123=419%:%
%:%1124=420%:%
%:%1125=421%:%
%:%1126=422%:%
%:%1127=423%:%
%:%1128=423%:%
%:%1131=424%:%
%:%1135=424%:%
%:%1136=424%:%
%:%1141=424%:%
%:%1144=425%:%
%:%1145=425%:%
%:%1148=426%:%
%:%1152=426%:%
%:%1153=426%:%
%:%1158=426%:%
%:%1161=427%:%
%:%1162=427%:%
%:%1165=428%:%
%:%1169=428%:%
%:%1170=428%:%
%:%1175=428%:%
%:%1178=429%:%
%:%1179=429%:%
%:%1182=430%:%
%:%1186=430%:%
%:%1187=430%:%
%:%1192=430%:%
%:%1195=431%:%
%:%1196=431%:%
%:%1199=432%:%
%:%1203=432%:%
%:%1204=432%:%
%:%1209=432%:%
%:%1212=433%:%
%:%1213=433%:%
%:%1216=434%:%
%:%1220=434%:%
%:%1221=434%:%
%:%1226=434%:%
%:%1229=435%:%
%:%1230=436%:%
%:%1231=436%:%
%:%1234=437%:%
%:%1238=437%:%
%:%1239=437%:%
%:%1244=437%:%
%:%1247=438%:%
%:%1248=438%:%
%:%1251=439%:%
%:%1255=439%:%
%:%1256=439%:%
%:%1261=439%:%
%:%1264=440%:%
%:%1265=441%:%
%:%1266=442%:%
%:%1267=442%:%
%:%1268=443%:%
%:%1271=444%:%
%:%1275=444%:%
%:%1276=444%:%
%:%1281=444%:%
%:%1284=445%:%
%:%1285=446%:%
%:%1286=447%:%
%:%1287=447%:%
%:%1288=448%:%
%:%1289=449%:%
%:%1290=450%:%
%:%1293=451%:%
%:%1297=451%:%
%:%1298=451%:%
%:%1299=452%:%
%:%1300=452%:%
%:%1301=453%:%
%:%1302=453%:%
%:%1303=454%:%
%:%1304=454%:%
%:%1305=454%:%
%:%1306=455%:%
%:%1307=455%:%
%:%1308=455%:%
%:%1309=456%:%
%:%1310=456%:%
%:%1311=456%:%
%:%1312=456%:%
%:%1313=456%:%
%:%1314=457%:%
%:%1315=457%:%
%:%1316=458%:%
%:%1317=458%:%
%:%1318=459%:%
%:%1319=459%:%
%:%1320=459%:%
%:%1321=460%:%
%:%1322=460%:%
%:%1323=460%:%
%:%1324=461%:%
%:%1325=461%:%
%:%1326=461%:%
%:%1327=461%:%
%:%1328=461%:%
%:%1329=462%:%
%:%1330=462%:%
%:%1331=463%:%
%:%1332=463%:%
%:%1333=464%:%
%:%1334=464%:%
%:%1335=464%:%
%:%1336=465%:%
%:%1337=466%:%
%:%1338=466%:%
%:%1339=467%:%
%:%1340=467%:%
%:%1341=467%:%
%:%1342=468%:%
%:%1343=468%:%
%:%1344=469%:%
%:%1345=469%:%
%:%1346=470%:%
%:%1347=470%:%
%:%1348=470%:%
%:%1349=471%:%
%:%1350=471%:%
%:%1351=471%:%
%:%1352=472%:%
%:%1353=472%:%
%:%1354=472%:%
%:%1355=472%:%
%:%1356=472%:%
%:%1357=473%:%
%:%1358=473%:%
%:%1359=474%:%
%:%1360=474%:%
%:%1361=475%:%
%:%1362=475%:%
%:%1363=475%:%
%:%1364=475%:%
%:%1365=476%:%
%:%1366=476%:%
%:%1367=476%:%
%:%1368=476%:%
%:%1369=477%:%
%:%1370=477%:%
%:%1371=477%:%
%:%1372=477%:%
%:%1373=478%:%
%:%1374=478%:%
%:%1375=478%:%
%:%1376=478%:%
%:%1377=479%:%
%:%1378=479%:%
%:%1379=480%:%
%:%1380=480%:%
%:%1381=480%:%
%:%1382=480%:%
%:%1383=480%:%
%:%1384=481%:%
%:%1390=481%:%
%:%1393=482%:%
%:%1394=483%:%
%:%1395=483%:%
%:%1398=484%:%
%:%1402=484%:%
%:%1403=484%:%
%:%1408=484%:%
%:%1411=485%:%
%:%1412=486%:%
%:%1413=486%:%
%:%1414=487%:%
%:%1415=488%:%
%:%1416=489%:%
%:%1417=490%:%
%:%1418=491%:%
%:%1419=492%:%
%:%1426=493%:%
%:%1427=493%:%
%:%1428=494%:%
%:%1429=494%:%
%:%1430=495%:%
%:%1431=495%:%
%:%1432=495%:%
%:%1433=495%:%
%:%1434=496%:%
%:%1435=496%:%
%:%1436=496%:%
%:%1437=496%:%
%:%1438=496%:%
%:%1439=497%:%
%:%1440=497%:%
%:%1441=497%:%
%:%1442=498%:%
%:%1443=498%:%
%:%1444=499%:%
%:%1445=499%:%
%:%1446=500%:%
%:%1447=500%:%
%:%1448=500%:%
%:%1449=500%:%
%:%1450=500%:%
%:%1451=501%:%
%:%1452=501%:%
%:%1453=502%:%
%:%1454=502%:%
%:%1455=503%:%
%:%1456=503%:%
%:%1457=503%:%
%:%1458=503%:%
%:%1459=504%:%
%:%1460=504%:%
%:%1461=504%:%
%:%1462=504%:%
%:%1463=504%:%
%:%1464=505%:%
%:%1465=505%:%
%:%1466=505%:%
%:%1467=505%:%
%:%1468=506%:%
%:%1469=506%:%
%:%1470=506%:%
%:%1471=506%:%
%:%1472=506%:%
%:%1473=507%:%
%:%1474=507%:%
%:%1475=507%:%
%:%1476=507%:%
%:%1477=508%:%
%:%1478=508%:%
%:%1479=508%:%
%:%1480=508%:%
%:%1481=508%:%
%:%1482=509%:%
%:%1483=509%:%
%:%1484=509%:%
%:%1485=509%:%
%:%1486=509%:%
%:%1487=510%:%
%:%1488=510%:%
%:%1489=510%:%
%:%1490=511%:%
%:%1491=511%:%
%:%1492=511%:%
%:%1493=512%:%
%:%1494=512%:%
%:%1495=512%:%
%:%1496=512%:%
%:%1497=513%:%
%:%1498=513%:%
%:%1499=513%:%
%:%1500=513%:%
%:%1501=513%:%
%:%1502=514%:%
%:%1503=514%:%
%:%1504=515%:%
%:%1505=515%:%
%:%1506=515%:%
%:%1507=515%:%
%:%1508=516%:%
%:%1514=516%:%
%:%1517=517%:%
%:%1518=518%:%
%:%1519=518%:%
%:%1520=519%:%
%:%1521=520%:%
%:%1522=521%:%
%:%1525=524%:%
%:%1528=525%:%
%:%1532=525%:%
%:%1533=525%:%
%:%1534=526%:%
%:%1535=526%:%
%:%1536=527%:%
%:%1537=527%:%
%:%1538=528%:%
%:%1539=528%:%
%:%1540=528%:%
%:%1541=528%:%
%:%1542=529%:%
%:%1543=529%:%
%:%1544=530%:%
%:%1545=530%:%
%:%1546=530%:%
%:%1547=530%:%
%:%1548=531%:%
%:%1549=531%:%
%:%1550=532%:%
%:%1551=532%:%
%:%1552=532%:%
%:%1553=532%:%
%:%1554=533%:%
%:%1555=533%:%
%:%1556=534%:%
%:%1557=534%:%
%:%1558=534%:%
%:%1559=534%:%
%:%1560=535%:%
%:%1561=535%:%
%:%1562=536%:%
%:%1563=536%:%
%:%1564=537%:%
%:%1565=537%:%
%:%1566=537%:%
%:%1567=537%:%
%:%1568=538%:%
%:%1569=538%:%
%:%1570=539%:%
%:%1571=539%:%
%:%1572=539%:%
%:%1573=539%:%
%:%1574=540%:%
%:%1575=540%:%
%:%1576=541%:%
%:%1577=541%:%
%:%1578=541%:%
%:%1579=541%:%
%:%1580=542%:%
%:%1581=542%:%
%:%1582=542%:%
%:%1583=542%:%
%:%1584=542%:%
%:%1585=543%:%
%:%1586=543%:%
%:%1587=544%:%
%:%1588=544%:%
%:%1589=545%:%
%:%1590=545%:%
%:%1591=545%:%
%:%1592=545%:%
%:%1593=546%:%
%:%1594=546%:%
%:%1595=547%:%
%:%1596=547%:%
%:%1597=547%:%
%:%1598=547%:%
%:%1599=548%:%
%:%1600=548%:%
%:%1601=548%:%
%:%1602=549%:%
%:%1603=549%:%
%:%1604=549%:%
%:%1605=549%:%
%:%1606=550%:%
%:%1607=550%:%
%:%1608=550%:%
%:%1609=551%:%
%:%1610=551%:%
%:%1611=551%:%
%:%1612=551%:%
%:%1613=552%:%
%:%1614=552%:%
%:%1615=552%:%
%:%1616=553%:%
%:%1617=553%:%
%:%1618=553%:%
%:%1619=553%:%
%:%1620=554%:%
%:%1621=554%:%
%:%1622=554%:%
%:%1623=554%:%
%:%1624=555%:%
%:%1625=555%:%
%:%1626=555%:%
%:%1627=556%:%
%:%1628=556%:%
%:%1629=556%:%
%:%1630=556%:%
%:%1631=557%:%
%:%1632=557%:%
%:%1633=557%:%
%:%1634=558%:%
%:%1635=558%:%
%:%1636=558%:%
%:%1637=558%:%
%:%1638=559%:%
%:%1639=559%:%
%:%1640=559%:%
%:%1641=560%:%
%:%1642=560%:%
%:%1643=560%:%
%:%1644=560%:%
%:%1645=561%:%
%:%1646=561%:%
%:%1647=562%:%
%:%1648=562%:%
%:%1649=562%:%
%:%1650=562%:%
%:%1651=563%:%
%:%1652=563%:%
%:%1653=564%:%
%:%1654=564%:%
%:%1655=565%:%
%:%1656=565%:%
%:%1657=566%:%
%:%1658=566%:%
%:%1659=566%:%
%:%1660=567%:%
%:%1661=567%:%
%:%1662=567%:%
%:%1663=567%:%
%:%1664=568%:%
%:%1665=568%:%
%:%1666=568%:%
%:%1667=569%:%
%:%1668=569%:%
%:%1669=569%:%
%:%1670=570%:%
%:%1671=570%:%
%:%1672=570%:%
%:%1673=570%:%
%:%1674=571%:%
%:%1675=571%:%
%:%1676=571%:%
%:%1677=572%:%
%:%1678=572%:%
%:%1679=572%:%
%:%1680=573%:%
%:%1681=573%:%
%:%1682=574%:%
%:%1683=575%:%
%:%1684=575%:%
%:%1685=576%:%
%:%1686=576%:%
%:%1687=576%:%
%:%1688=576%:%
%:%1689=577%:%
%:%1690=577%:%
%:%1691=578%:%
%:%1692=578%:%
%:%1693=578%:%
%:%1694=579%:%
%:%1695=579%:%
%:%1696=579%:%
%:%1697=579%:%
%:%1698=580%:%
%:%1699=580%:%
%:%1700=581%:%
%:%1701=581%:%
%:%1702=581%:%
%:%1703=581%:%
%:%1704=581%:%
%:%1705=582%:%
%:%1711=582%:%
%:%1716=583%:%
%:%1721=584%:%