### 2024-11-01 code/output/document/Renaming_Auto.tex

%
\begin{isabellebody}%
\setisabellecontext{Renaming{\isacharunderscore}{\kern0pt}Auto}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Renaming{\isacharunderscore}{\kern0pt}Auto\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ Renaming\isanewline
\ \ \ \ Utils\isanewline
\ \ \ \ ZF{\isachardot}{\kern0pt}Finite\isanewline
\ \ \ \ ZF{\isachardot}{\kern0pt}List\isanewline
\ \ \isakeyword{keywords}\ {\isachardoublequoteopen}rename{\isachardoublequoteclose}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ thy{\isacharunderscore}{\kern0pt}decl\ {\isacharpercent}{\kern0pt}\ {\isachardoublequoteopen}ML{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ {\isachardoublequoteopen}simple{\isacharunderscore}{\kern0pt}rename{\isachardoublequoteclose}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ thy{\isacharunderscore}{\kern0pt}decl\ {\isacharpercent}{\kern0pt}\ {\isachardoublequoteopen}ML{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ {\isachardoublequoteopen}src{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{and}\ {\isachardoublequoteopen}tgt{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{abbrevs}\ {\isachardoublequoteopen}simple{\isacharunderscore}{\kern0pt}rename{\isachardoublequoteclose}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isachardoublequoteclose}\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ app{\isacharunderscore}{\kern0pt}fun\ {\isacharequal}{\kern0pt}\ apply{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}succI\ {\isacharequal}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{ML{\isacharunderscore}{\kern0pt}file}\isamarkupfalse%
\ {\isacartoucheopen}renaming{\isachardot}{\kern0pt}ML{\isacartoucheclose}\isanewline
\isacommand{ML}\isamarkupfalse%
{\isacartoucheopen}\isanewline
\ \ fun\ renaming{\isacharunderscore}{\kern0pt}def\ mk{\isacharunderscore}{\kern0pt}ren\ name\ from\ to\ ctxt\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ let\ val\ to\ {\isacharequal}{\kern0pt}\ to\ {\isacharbar}{\kern0pt}{\isachargreater}{\kern0pt}\ Syntax{\isachardot}{\kern0pt}read{\isacharunderscore}{\kern0pt}term\ ctxt\isanewline
\ \ \ \ \ \ \ \ val\ from\ {\isacharequal}{\kern0pt}\ from\ {\isacharbar}{\kern0pt}{\isachargreater}{\kern0pt}\ Syntax{\isachardot}{\kern0pt}read{\isacharunderscore}{\kern0pt}term\ ctxt\isanewline
\ \ \ \ \ \ \ \ val\ {\isacharparenleft}{\kern0pt}tc{\isacharunderscore}{\kern0pt}lemma{\isacharcomma}{\kern0pt}action{\isacharunderscore}{\kern0pt}lemma{\isacharcomma}{\kern0pt}fvs{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ mk{\isacharunderscore}{\kern0pt}ren\ from\ to\ ctxt\isanewline
\ \ \ \ \ \ \ \ val\ {\isacharparenleft}{\kern0pt}tc{\isacharunderscore}{\kern0pt}lemma{\isacharcomma}{\kern0pt}action{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Renaming{\isachardot}{\kern0pt}fix{\isacharunderscore}{\kern0pt}vars\ tc{\isacharunderscore}{\kern0pt}lemma\ fvs\ ctxt{\isacharcomma}{\kern0pt}\ Renaming{\isachardot}{\kern0pt}fix{\isacharunderscore}{\kern0pt}vars\ action{\isacharunderscore}{\kern0pt}lemma\ fvs\ ctxt{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ val\ ren{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}name\ {\isacharequal}{\kern0pt}\ Binding{\isachardot}{\kern0pt}name\ {\isacharparenleft}{\kern0pt}name\ {\isacharcircum}{\kern0pt}\ {\isachardoublequote}{\kern0pt}{\isacharunderscore}{\kern0pt}fn{\isachardoublequote}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ val\ ren{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def\ {\isacharequal}{\kern0pt}\ \ Binding{\isachardot}{\kern0pt}name\ {\isacharparenleft}{\kern0pt}name\ {\isacharcircum}{\kern0pt}\ {\isachardoublequote}{\kern0pt}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def{\isachardoublequote}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ val\ ren{\isacharunderscore}{\kern0pt}thm\ {\isacharequal}{\kern0pt}\ Binding{\isachardot}{\kern0pt}name\ {\isacharparenleft}{\kern0pt}name\ {\isacharcircum}{\kern0pt}\ {\isachardoublequote}{\kern0pt}{\isacharunderscore}{\kern0pt}thm{\isachardoublequote}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ in\isanewline
\ \ \ \ \ \ Local{\isacharunderscore}{\kern0pt}Theory{\isachardot}{\kern0pt}note\ \ \ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}thm{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}tc{\isacharunderscore}{\kern0pt}lemma{\isacharcomma}{\kern0pt}action{\isacharunderscore}{\kern0pt}lemma{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ ctxt\ {\isacharbar}{\kern0pt}{\isachargreater}{\kern0pt}\ snd\ {\isacharbar}{\kern0pt}{\isachargreater}{\kern0pt}\isanewline
\ \ \ \ \ \ Local{\isacharunderscore}{\kern0pt}Theory{\isachardot}{\kern0pt}define\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}name{\isacharcomma}{\kern0pt}\ NoSyn{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ r{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbar}{\kern0pt}{\isachargreater}{\kern0pt}\ snd\ \ \ \ \ \ \isanewline
\ \ end{\isacharsemicolon}{\kern0pt}\isanewline
{\isacartoucheclose}\isanewline
\isanewline
\isacommand{ML}\isamarkupfalse%
{\isacartoucheopen}\isanewline
local\isanewline
\isanewline
\ \ val\ ren{\isacharunderscore}{\kern0pt}parser\ {\isacharequal}{\kern0pt}\ Parse{\isachardot}{\kern0pt}position\ {\isacharparenleft}{\kern0pt}Parse{\isachardot}{\kern0pt}string\ {\isacharminus}{\kern0pt}{\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Parse{\isachardot}{\kern0pt}{\isachardollar}{\kern0pt}{\isachardollar}{\kern0pt}{\isachardollar}{\kern0pt}\ {\isachardoublequote}{\kern0pt}src{\isachardoublequote}{\kern0pt}\ {\isacharbar}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharminus}{\kern0pt}\ Parse{\isachardot}{\kern0pt}string\ {\isacharminus}{\kern0pt}{\isacharminus}{\kern0pt}{\isacharbar}{\kern0pt}\ Parse{\isachardot}{\kern0pt}{\isachardollar}{\kern0pt}{\isachardollar}{\kern0pt}{\isachardollar}{\kern0pt}\ {\isachardoublequote}{\kern0pt}tgt{\isachardoublequote}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharminus}{\kern0pt}\ Parse{\isachardot}{\kern0pt}string{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharsemicolon}{\kern0pt}\isanewline
\isanewline
\ \ val\ {\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ Outer{\isacharunderscore}{\kern0pt}Syntax{\isachardot}{\kern0pt}local{\isacharunderscore}{\kern0pt}theory\ \isactrlcommandUNDERSCOREkeyword {\isasymopen}rename{\isasymclose}\ {\isachardoublequote}{\kern0pt}ML\ setup\ for\ synthetic\ definitions{\isachardoublequote}{\kern0pt}\isanewline
\ \ \ \ \ {\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}parser\ {\isachargreater}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isacharparenleft}{\kern0pt}fn\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}name{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}from{\isacharcomma}{\kern0pt}to{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}{\isachargreater}{\kern0pt}\ renaming{\isacharunderscore}{\kern0pt}def\ Renaming{\isachardot}{\kern0pt}sum{\isacharunderscore}{\kern0pt}rename\ name\ from\ to\ {\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\isanewline
\ \ val\ {\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ Outer{\isacharunderscore}{\kern0pt}Syntax{\isachardot}{\kern0pt}local{\isacharunderscore}{\kern0pt}theory\ \isactrlcommandUNDERSCOREkeyword {\isasymopen}simple{\isacharunderscore}{\kern0pt}rename{\isasymclose}\ {\isachardoublequote}{\kern0pt}ML\ setup\ for\ synthetic\ definitions{\isachardoublequote}{\kern0pt}\isanewline
\ \ \ \ \ {\isacharparenleft}{\kern0pt}ren{\isacharunderscore}{\kern0pt}parser\ {\isachargreater}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isacharparenleft}{\kern0pt}fn\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}name{\isacharcomma}{\kern0pt}{\isacharparenleft}{\kern0pt}from{\isacharcomma}{\kern0pt}to{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}{\isachargreater}{\kern0pt}\ renaming{\isacharunderscore}{\kern0pt}def\ Renaming{\isachardot}{\kern0pt}ren{\isacharunderscore}{\kern0pt}thm\ name\ from\ to\ {\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\isanewline
in\isanewline
end\isanewline
{\isacartoucheclose}%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Renaming_Auto.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%16=6%:%
%:%17=7%:%
%:%18=8%:%
%:%19=9%:%
%:%20=10%:%
%:%21=11%:%
%:%22=12%:%
%:%27=12%:%
%:%30=13%:%
%:%31=14%:%
%:%32=14%:%
%:%33=15%:%
%:%34=15%:%
%:%37=16%:%
%:%42=17%:%
%:%43=17%:%
%:%44=18%:%
%:%45=18%:%
%:%59=32%:%
%:%60=33%:%
%:%61=34%:%
%:%62=34%:%
%:%83=50%:%
%:%92=51%:%