### 2024-11-01 code/output/document/Replacement_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Replacement{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Axiom of Replacement in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Replacement{\isacharunderscore}{\kern0pt}Axiom\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ Least\ Relative{\isacharunderscore}{\kern0pt}Univ\ Separation{\isacharunderscore}{\kern0pt}Axiom\ Renaming{\isacharunderscore}{\kern0pt}Auto\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{rename}\isamarkupfalse%
\ {\isachardoublequoteopen}renrep{\isadigit{1}}{\isachardoublequoteclose}\ \isakeyword{src}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{tgt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}V{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{definition}\isamarkupfalse%
\ renrep{\isacharunderscore}{\kern0pt}fn\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}renrep{\isacharunderscore}{\kern0pt}fn{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ sum{\isacharparenleft}{\kern0pt}renrep{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}{\isadigit{8}}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ renrep\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}renrep{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{8}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}renrep{\isacharunderscore}{\kern0pt}fn{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ renrep{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}renrep{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ renrep{\isacharunderscore}{\kern0pt}def\ renrep{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renrep{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ renrep{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ ren{\isacharunderscore}{\kern0pt}tc\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}renrep{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}\ {\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}renrep{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{8}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ \ renrep{\isacharunderscore}{\kern0pt}def\ renrep{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renrep{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ renrep{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ arity{\isacharunderscore}{\kern0pt}ren\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ renrep{\isacharunderscore}{\kern0pt}sats\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}V\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymalpha}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ renrep{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}V{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ \ renrep{\isacharunderscore}{\kern0pt}def\ renrep{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renrep{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ren{\isacharcomma}{\kern0pt}insert\ assms{\isacharcomma}{\kern0pt}\ auto\ simp\ add{\isacharcolon}{\kern0pt}renrep{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ M{\isacharcomma}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ renrep{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}simplified{\isacharcomma}{\kern0pt}\isakeyword{where}\ p{\isacharequal}{\kern0pt}p\ \isakeyword{and}\ {\isasymalpha}{\isacharequal}{\kern0pt}{\isasymalpha}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{rename}\isamarkupfalse%
\ {\isachardoublequoteopen}renpbdy{\isadigit{1}}{\isachardoublequoteclose}\ \isakeyword{src}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{tgt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{definition}\isamarkupfalse%
\ renpbdy{\isacharunderscore}{\kern0pt}fn\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}renpbdy{\isacharunderscore}{\kern0pt}fn{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ sum{\isacharparenleft}{\kern0pt}renpbdy{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}{\isadigit{7}}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ renpbdy\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}renpbdy{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}renpbdy{\isacharunderscore}{\kern0pt}fn{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\isanewline
\ \ renpbdy{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ renpbdy{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ renpbdy{\isacharunderscore}{\kern0pt}def\ renpbdy{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renpbdy{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ \ renpbdy{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ ren{\isacharunderscore}{\kern0pt}tc\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ arity{\isacharunderscore}{\kern0pt}renpbdy{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}renpbdy{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ renpbdy{\isacharunderscore}{\kern0pt}def\ renpbdy{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renpbdy{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ \ renpbdy{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ arity{\isacharunderscore}{\kern0pt}ren\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\isanewline
\ \ sats{\isacharunderscore}{\kern0pt}renpbdy{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ renpbdy{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ renpbdy{\isacharunderscore}{\kern0pt}def\ renpbdy{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renpbdy{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ren{\isacharcomma}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ renpbdy{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ M{\isacharcomma}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ renpbdy{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}simplified{\isacharcomma}{\kern0pt}\isakeyword{where}\ {\isasymalpha}{\isacharequal}{\kern0pt}{\isasymalpha}\ \isakeyword{and}\ x{\isacharequal}{\kern0pt}x{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{rename}\isamarkupfalse%
\ {\isachardoublequoteopen}renbody{\isadigit{1}}{\isachardoublequoteclose}\ \isakeyword{src}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{tgt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{definition}\isamarkupfalse%
\ renbody{\isacharunderscore}{\kern0pt}fn\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}renbody{\isacharunderscore}{\kern0pt}fn{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ sum{\isacharparenleft}{\kern0pt}renbody{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharcomma}{\kern0pt}id{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ renbody\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}renbody{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}renbody{\isacharunderscore}{\kern0pt}fn{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\isanewline
\ \ renbody{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ renbody{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ renbody{\isacharunderscore}{\kern0pt}def\ renbody{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renbody{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ \ renbody{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ ren{\isacharunderscore}{\kern0pt}tc\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ arity{\isacharunderscore}{\kern0pt}renbody{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{5}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ arity{\isacharparenleft}{\kern0pt}renbody{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ renbody{\isacharunderscore}{\kern0pt}def\ renbody{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renbody{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ \ renbody{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ arity{\isacharunderscore}{\kern0pt}ren\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\isanewline
\ \ sats{\isacharunderscore}{\kern0pt}renbody{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{5}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ renbody{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ renbody{\isacharunderscore}{\kern0pt}def\ renbody{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\ renbody{\isadigit{1}}{\isacharunderscore}{\kern0pt}fn{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ sats{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ren{\isacharcomma}{\kern0pt}\ auto\ simp\ add{\isacharcolon}{\kern0pt}renbody{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ M{\isacharcomma}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ renbody{\isadigit{1}}{\isacharunderscore}{\kern0pt}thm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharbrackleft}{\kern0pt}\isakeyword{where}\ {\isasymalpha}{\isacharequal}{\kern0pt}{\isasymalpha}\ \isakeyword{and}\ m{\isacharequal}{\kern0pt}m{\isacharcomma}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{context}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pow{\isacharunderscore}{\kern0pt}inter{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}powerset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ y\ {\isacharequal}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}{\isasymalpha}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtau}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}V{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Vset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isasymtau}{\isasymin}V\ {\isasymand}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms{\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ is{\isacharunderscore}{\kern0pt}Vset{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ nonempty{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ is{\isacharunderscore}{\kern0pt}Vset{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\ is{\isacharunderscore}{\kern0pt}Vset{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ nonempty{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ nonempty{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ length{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}THEN\ nat{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}Ord{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ blast{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ renrep{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ renrep{\isacharunderscore}{\kern0pt}type\ definability{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}le\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize{\isacharunderscore}{\kern0pt}notc}\isamarkupfalse%
\ {\isachardoublequoteopen}prebody{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ sats{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}forces{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}renrep{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ new{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}defs\ {\isacharequal}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ is{\isacharunderscore}{\kern0pt}transrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ mem{\isacharunderscore}{\kern0pt}eclose{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ finite{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}wfrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \ Memrel{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ eclose{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}iterates{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ iterates{\isacharunderscore}{\kern0pt}MH{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}case{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ quasinat{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pre{\isacharunderscore}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ restriction{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymalpha}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtau}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}V{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Vset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isasymtau}{\isasymin}V\ {\isasymand}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ sats{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto\ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymalpha}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{6}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}HVfrom{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}powapply{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ new{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\isanewline
\ \ \ \ arity{\isacharunderscore}{\kern0pt}renrep{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}forces{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ arity{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}le{\isacharbrackleft}{\kern0pt}simplified{\isacharbrackright}{\kern0pt}\ pred{\isacharunderscore}{\kern0pt}le\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}renpbdy{\isacharparenleft}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymalpha}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{5}}\ \ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ new{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ arity{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm\ pred{\isacharunderscore}{\kern0pt}le\ \ arity{\isacharunderscore}{\kern0pt}renpbdy{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}renpbdy{\isacharparenleft}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ fst{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isasymin}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ body{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ renbody{\isacharparenleft}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymphi}{\isasymin}formula\ {\isasymLongrightarrow}\ \ body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}renpbdy{\isacharparenleft}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ sats{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}renbody{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ symmetric{\isacharbrackright}{\kern0pt}\ arity{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}renpbdy{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}renpbdy{\isacharparenleft}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ fst{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isasymin}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}renpbdy{\isacharbrackleft}{\kern0pt}OF\ arity{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm\ {\isacharunderscore}{\kern0pt}\ prebody{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ of\ concl{\isacharcolon}{\kern0pt}M{\isacharcomma}{\kern0pt}\ symmetric{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}lemma{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymexists}t\ p{\isachardot}{\kern0pt}\ x{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtau}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}V{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Vset{\isacharparenleft}{\kern0pt}{\isasymlambda}a{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isasymtau}\ {\isasymin}\ V\ {\isasymand}\ {\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtturnstile}\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ sats{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ x\ {\isasymalpha}\ m\ nenv{\isacharbrackright}{\kern0pt}\ sats{\isacharunderscore}{\kern0pt}renpbdy{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ x\ {\isasymalpha}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ sats{\isacharunderscore}{\kern0pt}prebody{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ fst{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isasymin}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}simp{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Replace{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ c{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ v{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}v{\isachardot}{\kern0pt}\ x{\isasymin}c{\isacharcomma}{\kern0pt}\ v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}R\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isasymlambda}\ x\ v\ {\isachardot}{\kern0pt}\ v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasympi}{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}{\isacharprime}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isasymin}M{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasympi}{\isasymin}M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ domain{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}c\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}val{\isacharbrackleft}{\kern0pt}of\ G\ {\isacharquery}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G{\isacharbrackleft}{\kern0pt}OF\ generic{\isacharbrackright}{\kern0pt}\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val\isanewline
\ \ \ \ \ \ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}prod{\isacharbrackleft}{\kern0pt}OF\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P{\isacharcomma}{\kern0pt}\ of\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ nenv\ \isakeyword{where}\ {\isachardoublequoteopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isacharequal}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ map{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ f\ \isakeyword{where}\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ {\isasymalpha}{\isasymin}M\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtau}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymtau}\ {\isasymin}\ Vset{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}\ {\isasymtturnstile}\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isasymequiv}\ {\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isakeyword{for}\ {\isasymrho}p\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ {\isasymalpha}{\isasymin}M\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtau}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}V{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Vset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}V{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isasymtau}{\isasymin}V\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}\ {\isasymtturnstile}\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ {\isasymalpha}{\isasymin}M\ {\isasymand}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isakeyword{for}\ {\isasymrho}p\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Vset{\isacharunderscore}{\kern0pt}abs\ Vset{\isacharunderscore}{\kern0pt}closed\ Ord{\isacharunderscore}{\kern0pt}Least{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}\ {\isasymalpha}{\isachardot}{\kern0pt}\ {\isasymalpha}{\isasymin}M\ {\isasymand}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ simp\ del{\isacharcolon}{\kern0pt}setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymrho}p\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ Least{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}{\isasymalpha}{\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}f{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymrho}p\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ least{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isasymlambda}{\isasymalpha}{\isachardot}{\kern0pt}\ {\isasymalpha}{\isasymin}M\ {\isasymand}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ least{\isacharunderscore}{\kern0pt}conj\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymrho}p\ \isacommand{unfolding}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ QQ\ \isakeyword{where}\ {\isachardoublequoteopen}QQ{\isasymequiv}{\isacharquery}{\kern0pt}Q{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isadigit{1}}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}{\isasymalpha}{\isachardot}{\kern0pt}\ QQ{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}f{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymrho}p\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ QQ{\isacharunderscore}{\kern0pt}def\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ assms\ {\isacartoucheopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasympi}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymrho}p{\isasymin}{\isacharquery}{\kern0pt}{\isasympi}\ {\isasymLongrightarrow}\ {\isasymexists}t\ p{\isachardot}{\kern0pt}\ {\isasymrho}p{\isacharequal}{\kern0pt}{\isasymlangle}t{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymrho}p\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ body{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}M\ {\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}{\isasymrho}p{\isasymin}{\isacharquery}{\kern0pt}{\isasympi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymrho}p{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymalpha}{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymalpha}\ {\isasymrho}p\ m\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ body{\isacharunderscore}{\kern0pt}lemma{\isacharbrackleft}{\kern0pt}of\ {\isasymrho}p\ {\isasymalpha}\ m\ nenv\ {\isasymphi}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isacharequal}{\kern0pt}{\isachardoublequoteopen}least{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ {\isasymrho}p\ m\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ asm{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymrho}p{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymrho}p{\isasymin}{\isacharquery}{\kern0pt}{\isasympi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ inM\ {\isacharequal}{\kern0pt}\ this\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ body\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ body{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isasymAnd}{\isasymalpha}{\isachardot}{\kern0pt}\ {\isasymalpha}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtau}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}V{\isasymin}M{\isachardot}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Vset{\isacharparenleft}{\kern0pt}{\isasymlambda}a{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymalpha}{\isacharcomma}{\kern0pt}\ V{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isasymtau}\ {\isasymin}\ V\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}snd{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}\ {\isasymtturnstile}\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}{\isasymalpha}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}\ m{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ body{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ inM\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M\ {\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm\ {\isasymlongleftrightarrow}\ least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ QQ{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}OF\ body{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ of\ {\isadigit{1}}{\isacharbrackright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ QQ{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}m{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm\ {\isasymlongleftrightarrow}\ least{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ QQ{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}{\isasymrho}p{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymrho}p{\isasymin}{\isacharquery}{\kern0pt}{\isasympi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymrho}p\ m\ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}\ {\isasymlambda}{\isasymrho}p\ m{\isachardot}{\kern0pt}\ M\ {\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}m{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ intro{\isacharcolon}{\kern0pt}unique{\isacharunderscore}{\kern0pt}least{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{3}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}nenv{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{5}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ body{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \ new{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}defs\ least{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}forces\ arity{\isacharunderscore}{\kern0pt}renrep\ arity{\isacharunderscore}{\kern0pt}renbody\ arity{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}fm{\isacharprime}{\kern0pt}\ nonempty\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ pred{\isacharunderscore}{\kern0pt}Un\ Un{\isacharunderscore}{\kern0pt}assoc{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}\ Un{\isacharunderscore}{\kern0pt}assoc{\isacharbrackleft}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{1}}\ pred{\isacharunderscore}{\kern0pt}Un{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharcomma}{\kern0pt}\ rule\ pred{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ auto\ intro{\isacharcolon}{\kern0pt}leI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isasymin}formula{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ inM\ {\isacharequal}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasympi}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ Y\ \isakeyword{where}\ {\isachardoublequoteopen}Y{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}m{\isasymin}M{\isachardot}{\kern0pt}\ m\ {\isasymin}\ Y\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymrho}p{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymrho}p\ {\isasymin}\ {\isacharquery}{\kern0pt}{\isasympi}\ {\isasymand}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymrho}p{\isacharcomma}{\kern0pt}m{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ replacement{\isacharunderscore}{\kern0pt}ax{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}least{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}QQ{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}f{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}f{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasympi}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ {\isacartoucheopen}{\isacharunderscore}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharunderscore}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharunderscore}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}f{\isacharunderscore}{\kern0pt}fm\ {\isasymlongleftrightarrow}\ least{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}{\isasymrho}p{\isacharparenright}{\kern0pt}{\isasymin}Y{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}{\isasymrho}p{\isasymin}{\isacharquery}{\kern0pt}{\isasympi}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymrho}p\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasympi}{\isasymin}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarsimp{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}y{\isasymin}Y{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}Y{\isasymin}M{\isacartoucheclose}\ separation{\isacharunderscore}{\kern0pt}ax\ sats{\isacharunderscore}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm\ trans{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ separation{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}y{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}ordinal{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}y{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ separation{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymUnion}\ {\isacharbraceleft}{\kern0pt}y{\isasymin}Y{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}sup\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vset{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}sup{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Vset{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ singletonM\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vset{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}sup{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M{\isacharbraceright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}blast\ intro{\isacharcolon}{\kern0pt}GenExtI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}c{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}c{\isacartoucheclose}\ {\isacartoucheopen}{\isasympi}{\isacharprime}{\kern0pt}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymrho}\ p\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isasympi}{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymrho}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair{\isacharprime}{\kern0pt}{\isacharbrackleft}{\kern0pt}of\ {\isasympi}{\isacharprime}{\kern0pt}\ x\ G{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymsigma}\ \isakeyword{where}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymsigma}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}nenv{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ q\ \isakeyword{where}\ {\isachardoublequoteopen}q{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymtturnstile}\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ truth{\isacharunderscore}{\kern0pt}lemma{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ generic{\isacharcomma}{\kern0pt}\ symmetric{\isacharcomma}{\kern0pt}\ of\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isasympi}{\isacharprime}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}{\isasympi}\ {\isasymLongrightarrow}\ f{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharparenright}{\kern0pt}{\isasymin}Y{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharparenright}{\kern0pt}{\isasymin}Y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic\ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymalpha}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}rank{\isacharparenleft}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymsigma}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymalpha}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ rank{\isacharunderscore}{\kern0pt}closed\ cons{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ Vset{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Vset{\isacharunderscore}{\kern0pt}Ord{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}q\ {\isasymtturnstile}\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ del{\isacharcolon}{\kern0pt}\ Vset{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymmu}\ {\isasymalpha}{\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ f{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtau}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymtau}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ Vset{\isacharparenleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymtturnstile}\ {\isasymphi}\ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ LeastI{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isasymlambda}\ {\isasymalpha}{\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymalpha}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharquery}{\kern0pt}{\isasymalpha}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}q{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}{\isasymrho}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}nenv{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ truth{\isacharunderscore}{\kern0pt}lemma{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ generic{\isacharcomma}{\kern0pt}\ of\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymrho}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}c{\isacartoucheclose}\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isasymTurnstile}\ {\isasymphi}{\isacartoucheclose}\ {\isacartoucheopen}env\ {\isacharequal}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymtau}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymrho}{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}x{\isacartoucheclose}\isanewline
\ \ \ \ \ \ {\isacartoucheopen}univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}x{\isasymin}c{\isacartoucheclose}\ {\isacartoucheopen}v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v{\isacharequal}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtI{\isacharbrackleft}{\kern0pt}of\ {\isasymtau}\ G{\isacharbrackright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtau}\ {\isasymin}\ Vset{\isacharparenleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}f{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \ {\isacartoucheopen}f{\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymrho}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isacharparenright}{\kern0pt}{\isasymin}Y{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ Vset{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}sup{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Vset{\isacharunderscore}{\kern0pt}Ord{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}iff\ lt{\isacharunderscore}{\kern0pt}Union{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtau}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}prod{\isacharbrackleft}{\kern0pt}of\ one\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vset{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}sup{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbrackright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}val{\isacharbrackleft}{\kern0pt}of\ G\ {\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G{\isacharbrackleft}{\kern0pt}OF\ generic{\isacharbrackright}{\kern0pt}\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ del{\isacharcolon}{\kern0pt}\ Vset{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}v{\isacharequal}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isasymin}Vset{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}sup{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ M{\isacharbraceright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}one{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}v{\isachardot}{\kern0pt}\ x{\isasymin}c{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}repl{\isasymsubseteq}{\isacharquery}{\kern0pt}big{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}repl\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}v{\isasymin}{\isacharquery}{\kern0pt}big{\isachardot}{\kern0pt}\ {\isasymexists}x{\isasymin}c{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharquery}{\kern0pt}rhs{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}intro\ equalityI\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v{\isasymin}{\isacharquery}{\kern0pt}repl{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}repl{\isasymsubseteq}{\isacharquery}{\kern0pt}big{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ {\isachardoublequoteopen}x{\isasymin}c{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v{\isasymin}{\isacharquery}{\kern0pt}big{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ subsetD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharquery}{\kern0pt}rhs{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ ReplaceI{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isasymlambda}\ x\ v{\isachardot}{\kern0pt}\ {\isasymexists}x{\isasymin}c{\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v{\isasymin}{\isacharquery}{\kern0pt}rhs{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}v{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}c{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ GenExtI{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharcomma}{\kern0pt}of\ G{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ {\isacartoucheopen}univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isakeyword{for}\ y\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v{\isasymin}{\isacharquery}{\kern0pt}repl{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ReplaceI{\isacharbrackleft}{\kern0pt}of\ {\isacharquery}{\kern0pt}R\ x\ v\ c{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasympsi}\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isasymin}c{\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{for}\ v\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}nth{\isacharunderscore}{\kern0pt}concat{\isacharbrackleft}{\kern0pt}of\ v\ c\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ env{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ inMG{\isacharequal}{\kern0pt}\ {\isacartoucheopen}nth{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}x{\isasymin}c{\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}x{\isasymin}c{\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}x{\isasymin}c{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ inMG\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ inMG\isanewline
\ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}c{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ inMG\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}x{\isasymin}c{\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}of\ {\isasymphi}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ lt{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ le{\isacharunderscore}{\kern0pt}refl{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}v{\isasymin}{\isacharquery}{\kern0pt}big{\isachardot}{\kern0pt}\ {\isasymexists}x{\isasymin}c{\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}v{\isasymin}{\isacharquery}{\kern0pt}big{\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ GenExtI{\isacharcomma}{\kern0pt}\ OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}big{\isacharunderscore}{\kern0pt}name{\isasymin}M{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\ \isakeyword{and}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}c{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}big{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}v{\isasymin}{\isacharquery}{\kern0pt}big{\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{theorem}\isamarkupfalse%
\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymlambda}x\ v{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}R{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x\ y\ {\isachardot}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ A\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}Y{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}v\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ A{\isacharcomma}{\kern0pt}\ v{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}x{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}y{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymforall}z{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ y\ {\isacharequal}{\kern0pt}\ z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}R{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}A{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\ {\isacartoucheopen}A{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}Y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Replace{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isasymin}\ {\isacharquery}{\kern0pt}Y\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ b\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}A{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}x{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}b\ {\isasymin}\ {\isacharquery}{\kern0pt}Y{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ transitivity{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isasymin}\ {\isacharquery}{\kern0pt}Y{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}{\isasymexists}x{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ that\isanewline
\ \ \ \ \ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}b{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ {\isadigit{1}}\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}z{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ b\ {\isacharequal}{\kern0pt}\ z{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ReplaceI{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isasymlambda}\ x\ y{\isachardot}{\kern0pt}\ y{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}b{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ b\ {\isasymin}\ {\isacharquery}{\kern0pt}Y\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}Y{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isacharparenleft}{\kern0pt}{\isasymexists}Y{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymforall}b{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ b\ {\isasymin}\ Y\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymand}\ {\isacharquery}{\kern0pt}R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}def\ univalent{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Replacement_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%36=5%:%
%:%41=6%:%
%:%46=7%:%
%:%47=7%:%
%:%52=7%:%
%:%55=8%:%
%:%56=9%:%
%:%57=9%:%
%:%58=10%:%
%:%59=11%:%
%:%60=12%:%
%:%61=12%:%
%:%62=13%:%
%:%63=14%:%
%:%64=15%:%
%:%65=16%:%
%:%66=16%:%
%:%67=17%:%
%:%68=18%:%
%:%71=19%:%
%:%75=19%:%
%:%76=19%:%
%:%77=20%:%
%:%78=20%:%
%:%79=21%:%
%:%80=21%:%
%:%85=21%:%
%:%88=22%:%
%:%89=23%:%
%:%90=23%:%
%:%91=24%:%
%:%92=25%:%
%:%95=26%:%
%:%99=26%:%
%:%100=26%:%
%:%101=27%:%
%:%102=27%:%
%:%103=28%:%
%:%104=28%:%
%:%109=28%:%
%:%112=29%:%
%:%113=30%:%
%:%114=30%:%
%:%115=31%:%
%:%116=32%:%
%:%117=33%:%
%:%118=34%:%
%:%119=35%:%
%:%122=36%:%
%:%126=36%:%
%:%127=36%:%
%:%128=37%:%
%:%129=37%:%
%:%130=38%:%
%:%135=38%:%
%:%140=39%:%
%:%145=40%:%
%:%146=40%:%
%:%151=40%:%
%:%154=41%:%
%:%155=42%:%
%:%156=42%:%
%:%157=43%:%
%:%158=44%:%
%:%159=45%:%
%:%160=45%:%
%:%161=46%:%
%:%162=47%:%
%:%163=48%:%
%:%164=49%:%
%:%165=50%:%
%:%166=50%:%
%:%167=51%:%
%:%170=52%:%
%:%174=52%:%
%:%175=52%:%
%:%176=53%:%
%:%177=53%:%
%:%178=54%:%
%:%179=54%:%
%:%184=54%:%
%:%187=55%:%
%:%188=56%:%
%:%189=56%:%
%:%192=57%:%
%:%196=57%:%
%:%197=57%:%
%:%198=58%:%
%:%199=58%:%
%:%200=59%:%
%:%201=59%:%
%:%206=59%:%
%:%209=60%:%
%:%210=61%:%
%:%211=61%:%
%:%212=62%:%
%:%213=63%:%
%:%216=64%:%
%:%220=64%:%
%:%221=64%:%
%:%222=65%:%
%:%223=65%:%
%:%224=66%:%
%:%229=66%:%
%:%232=67%:%
%:%235=68%:%
%:%240=69%:%
%:%241=69%:%
%:%246=69%:%
%:%249=70%:%
%:%250=71%:%
%:%251=71%:%
%:%252=72%:%
%:%253=73%:%
%:%254=74%:%
%:%255=74%:%
%:%256=75%:%
%:%257=76%:%
%:%258=77%:%
%:%259=78%:%
%:%260=78%:%
%:%261=79%:%
%:%264=80%:%
%:%268=80%:%
%:%269=80%:%
%:%270=81%:%
%:%271=81%:%
%:%272=82%:%
%:%273=82%:%
%:%278=82%:%
%:%281=83%:%
%:%282=84%:%
%:%283=84%:%
%:%284=85%:%
%:%287=86%:%
%:%291=86%:%
%:%292=86%:%
%:%293=87%:%
%:%294=87%:%
%:%295=88%:%
%:%296=88%:%
%:%301=88%:%
%:%304=89%:%
%:%305=90%:%
%:%306=90%:%
%:%307=91%:%
%:%308=92%:%
%:%311=93%:%
%:%315=93%:%
%:%316=93%:%
%:%317=94%:%
%:%318=94%:%
%:%319=95%:%
%:%324=95%:%
%:%327=96%:%
%:%328=97%:%
%:%329=97%:%
%:%330=98%:%
%:%331=99%:%
%:%332=100%:%
%:%333=100%:%
%:%334=101%:%
%:%335=102%:%
%:%336=103%:%
%:%337=104%:%
%:%340=105%:%
%:%344=105%:%
%:%345=105%:%
%:%346=105%:%
%:%351=105%:%
%:%354=106%:%
%:%355=107%:%
%:%356=108%:%
%:%357=108%:%
%:%358=109%:%
%:%359=110%:%
%:%360=111%:%
%:%361=112%:%
%:%362=113%:%
%:%365=114%:%
%:%369=114%:%
%:%370=114%:%
%:%371=115%:%
%:%372=115%:%
%:%373=116%:%
%:%374=116%:%
%:%375=117%:%
%:%376=117%:%
%:%377=118%:%
%:%378=118%:%
%:%379=119%:%
%:%380=119%:%
%:%381=120%:%
%:%382=120%:%
%:%383=121%:%
%:%384=121%:%
%:%385=122%:%
%:%386=122%:%
%:%387=123%:%
%:%388=123%:%
%:%389=124%:%
%:%390=124%:%
%:%391=125%:%
%:%392=125%:%
%:%393=126%:%
%:%394=126%:%
%:%395=127%:%
%:%396=127%:%
%:%397=128%:%
%:%398=128%:%
%:%399=129%:%
%:%400=129%:%
%:%401=130%:%
%:%402=130%:%
%:%403=131%:%
%:%404=131%:%
%:%405=131%:%
%:%406=132%:%
%:%407=132%:%
%:%408=133%:%
%:%409=133%:%
%:%410=134%:%
%:%411=134%:%
%:%412=134%:%
%:%413=135%:%
%:%419=135%:%
%:%422=136%:%
%:%425=137%:%
%:%430=138%:%
%:%431=138%:%
%:%436=138%:%
%:%439=139%:%
%:%440=140%:%
%:%441=140%:%
%:%442=141%:%
%:%443=142%:%
%:%444=143%:%
%:%451=144%:%
%:%452=144%:%
%:%453=145%:%
%:%454=145%:%
%:%455=146%:%
%:%456=146%:%
%:%457=146%:%
%:%458=147%:%
%:%459=147%:%
%:%460=148%:%
%:%461=148%:%
%:%462=149%:%
%:%463=149%:%
%:%464=149%:%
%:%465=150%:%
%:%466=150%:%
%:%467=150%:%
%:%468=150%:%
%:%469=150%:%
%:%470=151%:%
%:%476=151%:%
%:%479=152%:%
%:%480=153%:%
%:%481=153%:%
%:%482=154%:%
%:%483=155%:%
%:%484=156%:%
%:%485=157%:%
%:%486=157%:%
%:%487=158%:%
%:%488=159%:%
%:%489=160%:%
%:%490=161%:%
%:%491=162%:%
%:%494=163%:%
%:%498=163%:%
%:%499=163%:%
%:%500=163%:%
%:%501=163%:%
%:%506=163%:%
%:%509=164%:%
%:%510=165%:%
%:%511=166%:%
%:%512=166%:%
%:%513=167%:%
%:%514=168%:%
%:%515=169%:%
%:%516=170%:%
%:%519=171%:%
%:%523=171%:%
%:%524=171%:%
%:%525=172%:%
%:%526=172%:%
%:%527=173%:%
%:%528=173%:%
%:%533=173%:%
%:%536=174%:%
%:%537=175%:%
%:%538=176%:%
%:%539=176%:%
%:%540=177%:%
%:%541=178%:%
%:%542=179%:%
%:%543=180%:%
%:%544=180%:%
%:%547=181%:%
%:%551=181%:%
%:%552=181%:%
%:%553=181%:%
%:%554=182%:%
%:%555=182%:%
%:%560=182%:%
%:%563=183%:%
%:%564=184%:%
%:%565=184%:%
%:%566=185%:%
%:%567=186%:%
%:%568=187%:%
%:%569=188%:%
%:%572=189%:%
%:%576=189%:%
%:%577=189%:%
%:%578=190%:%
%:%579=190%:%
%:%580=191%:%
%:%581=191%:%
%:%586=191%:%
%:%589=192%:%
%:%590=193%:%
%:%591=193%:%
%:%592=194%:%
%:%593=195%:%
%:%594=196%:%
%:%595=197%:%
%:%596=198%:%
%:%599=199%:%
%:%603=199%:%
%:%604=199%:%
%:%605=199%:%
%:%606=200%:%
%:%607=200%:%
%:%612=200%:%
%:%615=201%:%
%:%616=202%:%
%:%617=202%:%
%:%618=203%:%
%:%619=204%:%
%:%620=205%:%
%:%621=206%:%
%:%622=206%:%
%:%625=207%:%
%:%629=207%:%
%:%630=207%:%
%:%631=207%:%
%:%636=207%:%
%:%639=208%:%
%:%640=209%:%
%:%641=209%:%
%:%642=210%:%
%:%643=211%:%
%:%644=212%:%
%:%645=213%:%
%:%646=214%:%
%:%647=215%:%
%:%650=216%:%
%:%654=216%:%
%:%655=216%:%
%:%656=217%:%
%:%657=217%:%
%:%658=218%:%
%:%659=218%:%
%:%664=218%:%
%:%667=219%:%
%:%668=220%:%
%:%669=220%:%
%:%670=221%:%
%:%671=222%:%
%:%672=223%:%
%:%673=224%:%
%:%674=225%:%
%:%675=226%:%
%:%678=227%:%
%:%682=227%:%
%:%683=227%:%
%:%684=228%:%
%:%685=229%:%
%:%686=229%:%
%:%691=229%:%
%:%694=230%:%
%:%695=231%:%
%:%696=231%:%
%:%697=232%:%
%:%698=233%:%
%:%699=234%:%
%:%700=235%:%
%:%701=236%:%
%:%702=237%:%
%:%705=238%:%
%:%709=238%:%
%:%710=238%:%
%:%711=239%:%
%:%712=240%:%
%:%713=240%:%
%:%718=240%:%
%:%721=241%:%
%:%722=242%:%
%:%723=242%:%
%:%724=243%:%
%:%725=244%:%
%:%726=245%:%
%:%727=246%:%
%:%728=247%:%
%:%729=248%:%
%:%736=249%:%
%:%737=249%:%
%:%738=250%:%
%:%739=250%:%
%:%740=251%:%
%:%741=251%:%
%:%742=252%:%
%:%743=252%:%
%:%744=253%:%
%:%745=253%:%
%:%746=253%:%
%:%747=254%:%
%:%748=254%:%
%:%749=255%:%
%:%750=255%:%
%:%751=256%:%
%:%752=256%:%
%:%753=256%:%
%:%754=257%:%
%:%755=257%:%
%:%756=258%:%
%:%757=258%:%
%:%758=259%:%
%:%759=259%:%
%:%760=260%:%
%:%761=260%:%
%:%762=261%:%
%:%763=261%:%
%:%764=262%:%
%:%765=262%:%
%:%766=263%:%
%:%767=263%:%
%:%768=263%:%
%:%769=264%:%
%:%770=264%:%
%:%771=265%:%
%:%772=265%:%
%:%773=265%:%
%:%774=266%:%
%:%775=266%:%
%:%776=267%:%
%:%777=268%:%
%:%778=268%:%
%:%779=269%:%
%:%780=270%:%
%:%781=270%:%
%:%782=270%:%
%:%783=271%:%
%:%784=271%:%
%:%785=272%:%
%:%786=272%:%
%:%787=273%:%
%:%788=273%:%
%:%789=274%:%
%:%790=274%:%
%:%791=274%:%
%:%792=274%:%
%:%793=275%:%
%:%794=275%:%
%:%795=276%:%
%:%796=276%:%
%:%797=277%:%
%:%798=277%:%
%:%799=278%:%
%:%800=278%:%
%:%801=279%:%
%:%802=279%:%
%:%803=279%:%
%:%804=279%:%
%:%805=280%:%
%:%806=280%:%
%:%807=281%:%
%:%808=281%:%
%:%809=282%:%
%:%810=282%:%
%:%811=283%:%
%:%812=283%:%
%:%813=283%:%
%:%814=284%:%
%:%815=284%:%
%:%816=285%:%
%:%817=285%:%
%:%818=286%:%
%:%819=286%:%
%:%820=287%:%
%:%821=287%:%
%:%822=288%:%
%:%823=288%:%
%:%824=289%:%
%:%825=289%:%
%:%826=290%:%
%:%827=290%:%
%:%828=291%:%
%:%829=291%:%
%:%830=292%:%
%:%831=292%:%
%:%832=293%:%
%:%833=293%:%
%:%834=294%:%
%:%835=295%:%
%:%836=295%:%
%:%837=295%:%
%:%838=296%:%
%:%839=296%:%
%:%840=297%:%
%:%841=297%:%
%:%842=298%:%
%:%843=298%:%
%:%844=299%:%
%:%845=299%:%
%:%846=300%:%
%:%847=300%:%
%:%848=301%:%
%:%849=301%:%
%:%850=302%:%
%:%851=302%:%
%:%853=304%:%
%:%854=304%:%
%:%855=305%:%
%:%856=305%:%
%:%857=306%:%
%:%858=306%:%
%:%859=307%:%
%:%860=307%:%
%:%861=307%:%
%:%862=308%:%
%:%863=308%:%
%:%864=309%:%
%:%865=309%:%
%:%866=310%:%
%:%867=310%:%
%:%868=311%:%
%:%869=311%:%
%:%870=312%:%
%:%871=312%:%
%:%872=312%:%
%:%873=313%:%
%:%874=313%:%
%:%875=314%:%
%:%876=314%:%
%:%877=315%:%
%:%878=315%:%
%:%879=315%:%
%:%880=316%:%
%:%881=316%:%
%:%882=316%:%
%:%883=317%:%
%:%884=317%:%
%:%885=317%:%
%:%886=318%:%
%:%887=318%:%
%:%888=318%:%
%:%889=319%:%
%:%890=320%:%
%:%891=320%:%
%:%892=321%:%
%:%893=321%:%
%:%894=322%:%
%:%895=322%:%
%:%896=323%:%
%:%897=323%:%
%:%898=324%:%
%:%899=325%:%
%:%900=325%:%
%:%901=325%:%
%:%902=326%:%
%:%903=326%:%
%:%904=326%:%
%:%905=327%:%
%:%906=327%:%
%:%907=328%:%
%:%908=328%:%
%:%909=329%:%
%:%910=329%:%
%:%911=330%:%
%:%912=330%:%
%:%913=331%:%
%:%914=332%:%
%:%915=332%:%
%:%916=333%:%
%:%917=333%:%
%:%918=333%:%
%:%919=334%:%
%:%920=334%:%
%:%921=335%:%
%:%922=336%:%
%:%923=336%:%
%:%924=337%:%
%:%925=337%:%
%:%926=338%:%
%:%927=338%:%
%:%928=339%:%
%:%929=339%:%
%:%930=340%:%
%:%931=340%:%
%:%932=341%:%
%:%933=341%:%
%:%934=342%:%
%:%935=343%:%
%:%936=343%:%
%:%937=344%:%
%:%938=344%:%
%:%939=345%:%
%:%940=345%:%
%:%941=346%:%
%:%942=346%:%
%:%943=346%:%
%:%944=347%:%
%:%945=347%:%
%:%946=348%:%
%:%947=348%:%
%:%948=349%:%
%:%949=349%:%
%:%950=349%:%
%:%951=350%:%
%:%952=350%:%
%:%953=351%:%
%:%954=351%:%
%:%955=352%:%
%:%956=352%:%
%:%957=352%:%
%:%958=353%:%
%:%959=353%:%
%:%960=354%:%
%:%961=354%:%
%:%962=355%:%
%:%963=355%:%
%:%964=355%:%
%:%965=356%:%
%:%966=356%:%
%:%967=357%:%
%:%968=357%:%
%:%969=358%:%
%:%970=358%:%
%:%971=359%:%
%:%972=359%:%
%:%973=360%:%
%:%974=360%:%
%:%975=361%:%
%:%976=361%:%
%:%977=362%:%
%:%978=362%:%
%:%979=363%:%
%:%980=363%:%
%:%981=364%:%
%:%982=364%:%
%:%983=365%:%
%:%984=365%:%
%:%985=366%:%
%:%986=366%:%
%:%987=367%:%
%:%988=367%:%
%:%989=367%:%
%:%990=368%:%
%:%991=368%:%
%:%992=369%:%
%:%993=369%:%
%:%994=370%:%
%:%995=370%:%
%:%996=371%:%
%:%997=371%:%
%:%998=372%:%
%:%999=372%:%
%:%1000=372%:%
%:%1001=373%:%
%:%1002=373%:%
%:%1003=374%:%
%:%1004=374%:%
%:%1005=375%:%
%:%1006=375%:%
%:%1007=376%:%
%:%1008=376%:%
%:%1009=377%:%
%:%1010=377%:%
%:%1011=378%:%
%:%1012=378%:%
%:%1013=379%:%
%:%1014=379%:%
%:%1015=380%:%
%:%1016=380%:%
%:%1017=381%:%
%:%1018=381%:%
%:%1019=382%:%
%:%1020=382%:%
%:%1021=383%:%
%:%1022=383%:%
%:%1023=383%:%
%:%1024=383%:%
%:%1025=384%:%
%:%1026=384%:%
%:%1027=385%:%
%:%1028=385%:%
%:%1029=386%:%
%:%1030=386%:%
%:%1031=386%:%
%:%1032=387%:%
%:%1033=387%:%
%:%1034=388%:%
%:%1035=388%:%
%:%1036=388%:%
%:%1037=389%:%
%:%1038=389%:%
%:%1039=390%:%
%:%1040=390%:%
%:%1041=391%:%
%:%1042=391%:%
%:%1043=391%:%
%:%1044=392%:%
%:%1045=392%:%
%:%1046=393%:%
%:%1047=393%:%
%:%1048=394%:%
%:%1049=394%:%
%:%1050=395%:%
%:%1051=395%:%
%:%1052=395%:%
%:%1053=396%:%
%:%1054=396%:%
%:%1055=397%:%
%:%1056=397%:%
%:%1057=398%:%
%:%1058=398%:%
%:%1059=398%:%
%:%1060=399%:%
%:%1061=399%:%
%:%1062=400%:%
%:%1063=400%:%
%:%1064=401%:%
%:%1065=401%:%
%:%1066=401%:%
%:%1067=402%:%
%:%1068=402%:%
%:%1069=403%:%
%:%1070=403%:%
%:%1071=404%:%
%:%1072=404%:%
%:%1073=404%:%
%:%1074=405%:%
%:%1075=405%:%
%:%1076=405%:%
%:%1077=406%:%
%:%1078=406%:%
%:%1079=406%:%
%:%1080=406%:%
%:%1081=407%:%
%:%1082=407%:%
%:%1083=408%:%
%:%1084=408%:%
%:%1085=409%:%
%:%1086=410%:%
%:%1087=410%:%
%:%1088=411%:%
%:%1089=411%:%
%:%1090=412%:%
%:%1091=412%:%
%:%1092=412%:%
%:%1093=412%:%
%:%1094=413%:%
%:%1095=413%:%
%:%1096=414%:%
%:%1097=414%:%
%:%1098=415%:%
%:%1099=415%:%
%:%1100=415%:%
%:%1101=416%:%
%:%1102=416%:%
%:%1103=417%:%
%:%1104=417%:%
%:%1105=418%:%
%:%1106=418%:%
%:%1107=419%:%
%:%1108=419%:%
%:%1109=420%:%
%:%1110=420%:%
%:%1111=421%:%
%:%1112=421%:%
%:%1113=422%:%
%:%1114=422%:%
%:%1115=423%:%
%:%1116=423%:%
%:%1117=424%:%
%:%1118=424%:%
%:%1119=425%:%
%:%1120=425%:%
%:%1121=426%:%
%:%1122=426%:%
%:%1123=427%:%
%:%1124=427%:%
%:%1125=428%:%
%:%1126=428%:%
%:%1127=429%:%
%:%1128=429%:%
%:%1129=430%:%
%:%1130=430%:%
%:%1131=431%:%
%:%1132=431%:%
%:%1133=432%:%
%:%1134=432%:%
%:%1135=433%:%
%:%1136=433%:%
%:%1137=434%:%
%:%1138=434%:%
%:%1139=434%:%
%:%1140=435%:%
%:%1141=435%:%
%:%1142=436%:%
%:%1143=436%:%
%:%1144=437%:%
%:%1145=437%:%
%:%1146=438%:%
%:%1147=438%:%
%:%1148=438%:%
%:%1149=439%:%
%:%1150=439%:%
%:%1151=440%:%
%:%1152=440%:%
%:%1153=441%:%
%:%1154=441%:%
%:%1155=442%:%
%:%1156=442%:%
%:%1157=443%:%
%:%1158=443%:%
%:%1159=444%:%
%:%1160=445%:%
%:%1161=445%:%
%:%1162=446%:%
%:%1163=446%:%
%:%1164=446%:%
%:%1165=447%:%
%:%1166=447%:%
%:%1167=448%:%
%:%1168=448%:%
%:%1169=448%:%
%:%1170=449%:%
%:%1171=449%:%
%:%1172=449%:%
%:%1173=450%:%
%:%1174=450%:%
%:%1175=451%:%
%:%1176=451%:%
%:%1177=451%:%
%:%1178=452%:%
%:%1179=452%:%
%:%1180=453%:%
%:%1181=453%:%
%:%1182=454%:%
%:%1183=454%:%
%:%1184=455%:%
%:%1185=455%:%
%:%1186=456%:%
%:%1187=456%:%
%:%1188=457%:%
%:%1189=457%:%
%:%1190=458%:%
%:%1191=458%:%
%:%1192=459%:%
%:%1193=459%:%
%:%1194=460%:%
%:%1195=461%:%
%:%1196=462%:%
%:%1197=462%:%
%:%1198=463%:%
%:%1199=463%:%
%:%1200=464%:%
%:%1201=464%:%
%:%1202=465%:%
%:%1203=465%:%
%:%1204=466%:%
%:%1205=466%:%
%:%1206=467%:%
%:%1207=467%:%
%:%1208=468%:%
%:%1209=468%:%
%:%1210=469%:%
%:%1211=469%:%
%:%1212=470%:%
%:%1213=470%:%
%:%1214=471%:%
%:%1215=471%:%
%:%1216=472%:%
%:%1217=472%:%
%:%1218=473%:%
%:%1219=473%:%
%:%1220=473%:%
%:%1221=474%:%
%:%1222=475%:%
%:%1223=475%:%
%:%1224=476%:%
%:%1225=476%:%
%:%1226=477%:%
%:%1227=477%:%
%:%1228=478%:%
%:%1229=478%:%
%:%1230=479%:%
%:%1231=479%:%
%:%1232=480%:%
%:%1233=480%:%
%:%1234=481%:%
%:%1235=481%:%
%:%1236=482%:%
%:%1237=482%:%
%:%1238=483%:%
%:%1239=483%:%
%:%1240=484%:%
%:%1241=484%:%
%:%1242=485%:%
%:%1243=486%:%
%:%1244=486%:%
%:%1245=487%:%
%:%1246=487%:%
%:%1247=488%:%
%:%1248=488%:%
%:%1249=489%:%
%:%1250=489%:%
%:%1251=490%:%
%:%1252=490%:%
%:%1253=491%:%
%:%1254=491%:%
%:%1255=492%:%
%:%1256=492%:%
%:%1257=493%:%
%:%1258=493%:%
%:%1259=494%:%
%:%1260=494%:%
%:%1261=495%:%
%:%1262=495%:%
%:%1263=496%:%
%:%1264=496%:%
%:%1265=497%:%
%:%1266=497%:%
%:%1267=498%:%
%:%1268=498%:%
%:%1269=499%:%
%:%1270=499%:%
%:%1271=499%:%
%:%1272=500%:%
%:%1273=500%:%
%:%1274=501%:%
%:%1275=501%:%
%:%1276=501%:%
%:%1277=502%:%
%:%1278=502%:%
%:%1279=503%:%
%:%1280=503%:%
%:%1281=504%:%
%:%1282=504%:%
%:%1283=505%:%
%:%1284=505%:%
%:%1285=505%:%
%:%1286=506%:%
%:%1287=506%:%
%:%1288=507%:%
%:%1289=507%:%
%:%1290=507%:%
%:%1291=508%:%
%:%1292=508%:%
%:%1293=509%:%
%:%1294=509%:%
%:%1295=509%:%
%:%1296=510%:%
%:%1302=510%:%
%:%1305=511%:%
%:%1306=512%:%
%:%1307=512%:%
%:%1308=513%:%
%:%1309=514%:%
%:%1310=515%:%
%:%1311=516%:%
%:%1318=517%:%
%:%1319=517%:%
%:%1320=518%:%
%:%1321=518%:%
%:%1322=519%:%
%:%1323=519%:%
%:%1324=520%:%
%:%1325=520%:%
%:%1326=521%:%
%:%1327=521%:%
%:%1328=522%:%
%:%1329=522%:%
%:%1330=523%:%
%:%1331=524%:%
%:%1332=524%:%
%:%1333=525%:%
%:%1334=525%:%
%:%1335=526%:%
%:%1336=526%:%
%:%1337=526%:%
%:%1338=527%:%
%:%1339=527%:%
%:%1340=528%:%
%:%1341=528%:%
%:%1342=529%:%
%:%1343=529%:%
%:%1344=529%:%
%:%1345=530%:%
%:%1346=530%:%
%:%1347=531%:%
%:%1348=531%:%
%:%1349=532%:%
%:%1350=532%:%
%:%1351=533%:%
%:%1352=533%:%
%:%1353=534%:%
%:%1354=534%:%
%:%1355=534%:%
%:%1356=535%:%
%:%1357=535%:%
%:%1358=536%:%
%:%1359=536%:%
%:%1360=537%:%
%:%1361=537%:%
%:%1362=538%:%
%:%1363=538%:%
%:%1364=539%:%
%:%1365=539%:%
%:%1366=539%:%
%:%1367=540%:%
%:%1368=540%:%
%:%1369=541%:%
%:%1370=541%:%
%:%1371=542%:%
%:%1372=542%:%
%:%1373=543%:%
%:%1374=543%:%
%:%1375=543%:%
%:%1376=544%:%
%:%1377=544%:%
%:%1378=545%:%
%:%1379=545%:%
%:%1380=546%:%
%:%1381=546%:%
%:%1382=547%:%
%:%1383=547%:%
%:%1384=548%:%
%:%1385=548%:%
%:%1386=548%:%
%:%1387=549%:%
%:%1388=549%:%
%:%1389=550%:%
%:%1390=550%:%
%:%1391=551%:%
%:%1392=551%:%
%:%1393=552%:%
%:%1394=552%:%
%:%1395=553%:%
%:%1396=553%:%
%:%1397=554%:%
%:%1398=554%:%
%:%1399=555%:%
%:%1400=555%:%
%:%1401=556%:%
%:%1402=556%:%
%:%1403=557%:%
%:%1404=557%:%
%:%1405=558%:%
%:%1406=558%:%
%:%1407=558%:%
%:%1408=558%:%
%:%1409=559%:%
%:%1410=559%:%
%:%1411=560%:%
%:%1417=560%:%
%:%1420=561%:%
%:%1421=562%:%
%:%1422=562%:%
%:%1425=563%:%
%:%1430=564%:%