### 2024-11-01 code/output/document/Separation_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Separation{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Axiom of Separation in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Separation{\isacharunderscore}{\kern0pt}Axiom\isanewline
\ \ \isakeyword{imports}\ Forcing{\isacharunderscore}{\kern0pt}Theorems\ Separation{\isacharunderscore}{\kern0pt}Rename\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ G{\isacharunderscore}{\kern0pt}generic\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ map{\isacharunderscore}{\kern0pt}val\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ env\ {\isacharequal}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}induct\ env{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ Nil\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Nil{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Nil{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Cons\ a\ l{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a{\isacharprime}{\kern0pt}\ l{\isacharprime}{\kern0pt}\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}l{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l{\isacharequal}{\kern0pt}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}a{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}a{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}a{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ GenExtD\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ \ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}c{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \ \isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ {\isasympi}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasymchi}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharequal}{\kern0pt}{\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}ren{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ phi\ {\isacharequal}{\kern0pt}\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymchi}{\isasymin}formula{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ phi\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ phi\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ arity{\isacharunderscore}{\kern0pt}forces{\isacharunderscore}{\kern0pt}le\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ arity{\isacharunderscore}{\kern0pt}forces\ leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}{\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}\ {\isasymin}\ formula{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}rensep{\isacharbrackleft}{\kern0pt}OF\ definability{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymchi}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \ definability{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymchi}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ type{\isacharunderscore}{\kern0pt}rensep\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}pred{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{5}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasympsi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ length{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ pred{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ pred{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form\ {\isasymin}\ formula{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{5}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ arity{\isacharunderscore}{\kern0pt}forces\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ definability\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharunderscore}{\kern0pt}all\ flip{\isacharcolon}{\kern0pt}setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ nenv\ \isakeyword{where}\ {\isachardoublequoteopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isacharequal}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ map{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}of\ {\isadigit{1}}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ le{\isacharunderscore}{\kern0pt}refl{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}nenv{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymtheta}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}abs{\isadigit{2}}{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\ \ \ \ \isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M\ {\isacharequal}{\kern0pt}\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P\ {\isasymin}\ M{\isacartoucheclose}\ \ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ u\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}u\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form\ {\isasymin}\ formula{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasympsi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}nenv\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ u\ {\isacharequal}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharquery}{\kern0pt}Pl{\isadigit{1}}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ transitivity{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymtheta}{\isasymin}M\ {\isasymLongrightarrow}\ p{\isasymin}P\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharquery}{\kern0pt}Pl{\isadigit{1}}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ F\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isakeyword{for}\ {\isasymtheta}\ p\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ p\ {\isasymtheta}\ \isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtheta}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ transitivity{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ in{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}{\isasymtheta}\ {\isasymin}\ M{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacartoucheclose}\ {\isacartoucheopen}u\ {\isasymin}\ M{\isacartoucheclose}\ {\isacartoucheopen}nenv{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}env{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}p{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharquery}{\kern0pt}Pl{\isadigit{1}}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}env{\isacharequal}{\kern0pt}{\isachardoublequoteopen}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ u{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasymchi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isasymin}\ formula{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ phi\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{1}}{\isacharprime}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ \ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharquery}{\kern0pt}Pl{\isadigit{1}}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}env\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ Eq{\isadigit{1}}{\isacharprime}{\kern0pt}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isasymin}\ formula{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ {\isacartoucheopen}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ u{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}env\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ sepren{\isacharunderscore}{\kern0pt}action{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}nenv{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharcomma}{\kern0pt}\ \ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}nenv{\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ app{\isacharunderscore}{\kern0pt}assoc\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ phi\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isasymin}formula{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ M{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{6}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isasymin}formula{\isacartoucheclose}\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ phi\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ F\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ a{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isasymle}\ {\isadigit{1}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}nenv{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ P\ {\isasymLongrightarrow}\ {\isacharquery}{\kern0pt}{\isasymchi}{\isasymin}formula\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}\ nenv{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymforall}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ G\ {\isasymlongrightarrow}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ F\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymchi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ a{\isadigit{1}}\ {\isacartoucheopen}{\isasymtheta}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ F\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing\ {\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{2}}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ one{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasymchi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ in{\isacharunderscore}{\kern0pt}M{\isacharprime}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharquery}{\kern0pt}Pl{\isadigit{1}}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}new{\isacharunderscore}{\kern0pt}form{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ F\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ Eq{\isadigit{1}}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ u\ {\isacharequal}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ F\ {\isasymlongrightarrow}\ M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Equivalence{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u{\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P\ {\isasymLongrightarrow}\ u\ {\isasymin}\ M\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ u\ {\isacharequal}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ F\ {\isasymlongrightarrow}\ M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isakeyword{for}\ u\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}env\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ map{\isacharunderscore}{\kern0pt}nenv{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ nenv{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ map{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}distrib\ append{\isadigit{1}}{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ aux{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ u\ {\isacharequal}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}p{\isasymin}G\ {\isasymlongrightarrow}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharparenleft}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}vals{\isacharparenleft}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isacharunderscore}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isakeyword{if}\ {\isachardoublequoteopen}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}u\ {\isasymin}\ M{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isachardoublequoteclose}\ \isakeyword{for}\ u\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Equivalence{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ OF\ that{\isacharbrackright}{\kern0pt}\ generic\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtheta}{\isasymin}M\ {\isasymLongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymtheta}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtheta}{\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymtheta}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ GenExtI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtheta}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExtI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ u{\isacharequal}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}p{\isasymin}G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isasymin}nth{\isacharparenleft}{\kern0pt}{\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ {\isasymand}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharquery}{\kern0pt}vals{\isacharparenleft}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}u\ {\isasymin}\ M{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isachardoublequoteclose}\ \isakeyword{for}\ u\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ aux{\isacharbrackleft}{\kern0pt}OF\ that{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ nth{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}{\isasymtheta}{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ {\isasymtheta}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}concat{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ that\ GenExtI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ u{\isacharequal}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}p{\isasymin}G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharquery}{\kern0pt}vals{\isacharparenleft}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}u\ {\isasymin}\ M{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isachardoublequoteclose}\ \isakeyword{for}\ u\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}u{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymtimes}P\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ u\ {\isacharequal}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}p\ {\isasymin}\ G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharquery}{\kern0pt}vals{\isacharparenleft}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}transitivity{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}u{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymtimes}P\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\isanewline
\ \ \ \ \ {\isacharbraceleft}{\kern0pt}u{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymtimes}P\ {\isachardot}{\kern0pt}\ {\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ u\ {\isacharequal}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}p\ {\isasymin}\ G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}vals{\isacharparenleft}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}n{\isasymsubseteq}{\isacharquery}{\kern0pt}m{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}mono\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ first{\isacharunderscore}{\kern0pt}incl{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ \ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\ \isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}{\isasympsi}{\isasymin}formula{\isacartoucheclose}\ \ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympsi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ in{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}nenv\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}n{\isasymin}M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}ax\ leI\ separation{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ generic\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}G{\isasymsubseteq}P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ filter{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isasymexists}q{\isasymin}P\ {\isachardot}{\kern0pt}\ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}p\ {\isasymin}\ G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}name\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isacharbraceleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharcomma}{\kern0pt}\ {\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ t{\isacharparenright}{\kern0pt}\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ t{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}t{\isasymin}M\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}{\isasymtheta}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymlangle}t{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}p\ {\isasymin}\ G\ {\isasymlongrightarrow}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ {\isasymlongleftrightarrow}\ \isanewline
\ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ t{\isacharparenright}{\kern0pt}\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ t{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}{\isasymTurnstile}\ {\isasymphi}\ {\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ t\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymin}M{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}transitivity{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isacharbraceleft}{\kern0pt}x\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}c\ {\isacharcomma}{\kern0pt}\ {\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ x\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymsubseteq}\ {\isacharbraceleft}{\kern0pt}x\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}c\ {\isacharcomma}{\kern0pt}\ {\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ x\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isanewline
\ \ \ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}{\isacharbraceleft}{\kern0pt}x\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}c\ {\isacharcomma}{\kern0pt}\ {\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ x\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ x\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\ \ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ {\isasymexists}t{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}x\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}t{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Sep{\isacharunderscore}{\kern0pt}and{\isacharunderscore}{\kern0pt}Replace\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isacharbraceleft}{\kern0pt}x\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ x{\isasymin}c\ {\isacharcomma}{\kern0pt}\ {\isasymexists}q{\isasymin}P{\isachardot}{\kern0pt}\ x\ {\isasymin}\ c\ {\isasymand}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymand}\ q\ {\isasymin}\ G{\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ SepReplace{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}c{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}G{\isasymsubseteq}P{\isacartoucheclose}\ G{\isacharunderscore}{\kern0pt}nonempty\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}m{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}c{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}m\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{4}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}c{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}{\isasymtheta}{\isachardot}{\kern0pt}\ {\isasymexists}q{\isasymin}G{\isachardot}{\kern0pt}\ {\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}{\isasympi}\ {\isasymand}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtheta}\ q\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}{\isasympi}{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}{\isasympi}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtheta}{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ trans{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}{\isasympi}{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}nenv\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ \ Eq{\isadigit{4}}\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}x{\isacartoucheclose}\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c{\isacartoucheclose}\ {\isacartoucheopen}x\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ nth\ {\isacartoucheopen}{\isasymtheta}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{5}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtheta}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ \ Eq{\isadigit{5}}\ {\isacartoucheopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}nenv\ {\isasymin}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheclose}\ {\isacartoucheopen}env\ {\isacharequal}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheclose}\ map{\isacharunderscore}{\kern0pt}nenv\ \isanewline
\ \ \ \ \ \ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}r{\isasymin}G{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}r{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ truth{\isacharunderscore}{\kern0pt}lemma\ \ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ r\ \isakeyword{where}\ \ \ \ \ \ \isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}r{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}r{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isakeyword{and}\ {\isacartoucheopen}q{\isasymin}G{\isacartoucheclose}\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}p{\isasymin}G{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasympreceq}q{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasympreceq}r{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}def\ compat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}r{\isasymin}G{\isacartoucheclose}\ \ {\isacartoucheopen}q{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}G{\isasymsubseteq}P{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}p{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}transitivity{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}{\isasymtheta}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ \ {\isacartoucheopen}p{\isasympreceq}r{\isacartoucheclose}\ {\isacartoucheopen}nenv\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ {\isacartoucheopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}r{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}one{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ forces{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ strengthening{\isacharunderscore}{\kern0pt}lemma\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ {\isacartoucheopen}{\isasymtheta}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}nenv\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p\ {\isasymin}\ F\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ definition{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}forcing\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ {\isacartoucheopen}{\isasymtheta}{\isasymin}M{\isacartoucheclose}\ \ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{6}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymexists}{\isasymtheta}{\isacharprime}{\kern0pt}{\isasymin}M{\isachardot}{\kern0pt}\ {\isasymexists}p{\isacharprime}{\kern0pt}{\isasymin}P{\isachardot}{\kern0pt}\ \ {\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}{\isasymtheta}{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymforall}F{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymand}\ p{\isacharprime}{\kern0pt}\ {\isasymin}\ F\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ M{\isacharbrackleft}{\kern0pt}F{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ \ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymtheta}{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isacharquery}{\kern0pt}{\isasymchi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasympi}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}{\isasympi}{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}transitivity{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}{\isasympi}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymtheta}{\isasymin}M{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ \ {\isacartoucheopen}p{\isasymin}M{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ tuples{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtheta}{\isasymin}M{\isacartoucheclose}\ Eq{\isadigit{6}}\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharquery}{\kern0pt}Pl{\isadigit{1}}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isasympi}{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ nenv\ {\isasymTurnstile}\ {\isacharquery}{\kern0pt}{\isasympsi}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Equivalence\ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}domain{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isasymtimes}P{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isasymin}{\isacharquery}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}elem{\isacharbrackleft}{\kern0pt}of\ {\isasymtheta}\ p{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}x{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}m\ first{\isacharunderscore}{\kern0pt}incl\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}c{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isasymin}c{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}c{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ GenExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isakeyword{and}\ {\isacartoucheopen}x{\isasymin}c{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharparenleft}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}c{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ \ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ \ {\isasymphi}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ phi\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ arity{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}\ \ \ \isanewline
\ \ \ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\ \ \ \ \ \ \isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}c{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}n{\isasymin}M{\isacartoucheclose}\ GenExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{theorem}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymlambda}x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ c\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}c{\isasymin}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ nenv\ \isakeyword{where}\ \ {\isachardoublequoteopen}nenv{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}env\ {\isacharequal}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}nenv{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ GenExt{\isacharunderscore}{\kern0pt}def\ map{\isacharunderscore}{\kern0pt}val{\isacharbrackleft}{\kern0pt}of\ env{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymphi}\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Eq{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}x{\isasymin}c{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}iff\ rev{\isacharunderscore}{\kern0pt}bexI\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Collect{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Separation_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%35=4%:%
%:%38=5%:%
%:%39=6%:%
%:%40=6%:%
%:%41=7%:%
%:%42=8%:%
%:%43=9%:%
%:%44=9%:%
%:%45=10%:%
%:%46=11%:%
%:%49=12%:%
%:%53=12%:%
%:%54=12%:%
%:%55=13%:%
%:%56=13%:%
%:%57=14%:%
%:%58=14%:%
%:%59=15%:%
%:%60=15%:%
%:%61=15%:%
%:%62=16%:%
%:%63=16%:%
%:%64=16%:%
%:%65=16%:%
%:%66=17%:%
%:%67=17%:%
%:%68=18%:%
%:%69=18%:%
%:%70=19%:%
%:%71=19%:%
%:%72=19%:%
%:%73=20%:%
%:%74=21%:%
%:%75=22%:%
%:%76=22%:%
%:%77=23%:%
%:%78=23%:%
%:%79=24%:%
%:%80=24%:%
%:%81=24%:%
%:%82=24%:%
%:%83=25%:%
%:%89=25%:%
%:%92=26%:%
%:%93=27%:%
%:%94=28%:%
%:%95=28%:%
%:%96=29%:%
%:%97=30%:%
%:%98=31%:%
%:%99=32%:%
%:%100=33%:%
%:%107=34%:%
%:%108=34%:%
%:%109=35%:%
%:%110=35%:%
%:%111=36%:%
%:%112=36%:%
%:%113=37%:%
%:%114=37%:%
%:%115=37%:%
%:%116=38%:%
%:%117=38%:%
%:%118=39%:%
%:%119=39%:%
%:%120=40%:%
%:%121=40%:%
%:%122=41%:%
%:%123=41%:%
%:%124=42%:%
%:%125=42%:%
%:%126=43%:%
%:%127=43%:%
%:%128=43%:%
%:%129=44%:%
%:%130=44%:%
%:%131=45%:%
%:%132=45%:%
%:%133=46%:%
%:%134=46%:%
%:%135=46%:%
%:%136=47%:%
%:%137=47%:%
%:%138=48%:%
%:%139=48%:%
%:%140=49%:%
%:%141=49%:%
%:%142=49%:%
%:%143=50%:%
%:%144=50%:%
%:%145=51%:%
%:%146=51%:%
%:%147=52%:%
%:%148=52%:%
%:%149=52%:%
%:%150=53%:%
%:%151=53%:%
%:%152=54%:%
%:%153=54%:%
%:%154=55%:%
%:%155=55%:%
%:%156=56%:%
%:%157=56%:%
%:%158=57%:%
%:%159=57%:%
%:%160=58%:%
%:%161=58%:%
%:%162=59%:%
%:%163=59%:%
%:%164=60%:%
%:%165=60%:%
%:%166=61%:%
%:%167=62%:%
%:%168=62%:%
%:%169=63%:%
%:%170=63%:%
%:%171=64%:%
%:%172=64%:%
%:%173=65%:%
%:%174=65%:%
%:%175=66%:%
%:%176=66%:%
%:%177=67%:%
%:%178=67%:%
%:%179=68%:%
%:%180=68%:%
%:%181=69%:%
%:%182=69%:%
%:%183=70%:%
%:%184=70%:%
%:%185=70%:%
%:%186=71%:%
%:%187=71%:%
%:%188=72%:%
%:%189=72%:%
%:%190=73%:%
%:%191=73%:%
%:%192=74%:%
%:%193=74%:%
%:%194=75%:%
%:%195=75%:%
%:%196=76%:%
%:%197=76%:%
%:%198=76%:%
%:%199=77%:%
%:%200=77%:%
%:%201=78%:%
%:%202=78%:%
%:%203=79%:%
%:%204=79%:%
%:%205=80%:%
%:%206=80%:%
%:%207=81%:%
%:%208=81%:%
%:%209=82%:%
%:%210=82%:%
%:%211=83%:%
%:%212=83%:%
%:%213=84%:%
%:%214=84%:%
%:%215=85%:%
%:%216=85%:%
%:%217=86%:%
%:%218=86%:%
%:%219=87%:%
%:%220=87%:%
%:%221=88%:%
%:%222=88%:%
%:%223=89%:%
%:%224=89%:%
%:%225=90%:%
%:%226=90%:%
%:%228=92%:%
%:%229=93%:%
%:%230=93%:%
%:%231=94%:%
%:%232=94%:%
%:%234=96%:%
%:%235=97%:%
%:%236=98%:%
%:%237=98%:%
%:%238=99%:%
%:%239=99%:%
%:%240=100%:%
%:%241=100%:%
%:%242=101%:%
%:%243=101%:%
%:%244=102%:%
%:%245=102%:%
%:%246=102%:%
%:%247=102%:%
%:%248=103%:%
%:%249=103%:%
%:%250=104%:%
%:%251=104%:%
%:%252=105%:%
%:%253=105%:%
%:%254=105%:%
%:%255=106%:%
%:%256=106%:%
%:%257=107%:%
%:%258=107%:%
%:%259=108%:%
%:%260=108%:%
%:%261=109%:%
%:%262=109%:%
%:%263=110%:%
%:%264=110%:%
%:%265=110%:%
%:%266=111%:%
%:%267=111%:%
%:%268=112%:%
%:%269=112%:%
%:%270=112%:%
%:%271=113%:%
%:%272=113%:%
%:%273=114%:%
%:%274=114%:%
%:%275=114%:%
%:%276=115%:%
%:%277=115%:%
%:%278=115%:%
%:%279=115%:%
%:%280=116%:%
%:%281=116%:%
%:%282=116%:%
%:%283=116%:%
%:%284=117%:%
%:%285=117%:%
%:%286=118%:%
%:%287=118%:%
%:%288=119%:%
%:%289=119%:%
%:%290=120%:%
%:%291=120%:%
%:%292=121%:%
%:%293=122%:%
%:%294=123%:%
%:%295=123%:%
%:%296=124%:%
%:%297=124%:%
%:%298=125%:%
%:%299=125%:%
%:%300=126%:%
%:%301=126%:%
%:%302=126%:%
%:%303=127%:%
%:%304=127%:%
%:%305=128%:%
%:%306=128%:%
%:%307=128%:%
%:%308=129%:%
%:%309=129%:%
%:%310=130%:%
%:%311=130%:%
%:%312=131%:%
%:%313=132%:%
%:%314=132%:%
%:%315=133%:%
%:%316=133%:%
%:%317=134%:%
%:%318=134%:%
%:%319=135%:%
%:%320=135%:%
%:%321=136%:%
%:%322=136%:%
%:%323=137%:%
%:%324=138%:%
%:%325=138%:%
%:%326=139%:%
%:%327=139%:%
%:%328=140%:%
%:%329=140%:%
%:%330=141%:%
%:%331=141%:%
%:%332=142%:%
%:%333=142%:%
%:%334=143%:%
%:%335=143%:%
%:%337=145%:%
%:%338=146%:%
%:%339=146%:%
%:%340=147%:%
%:%341=147%:%
%:%342=148%:%
%:%343=148%:%
%:%344=149%:%
%:%345=150%:%
%:%346=150%:%
%:%347=150%:%
%:%348=151%:%
%:%349=151%:%
%:%350=152%:%
%:%351=152%:%
%:%352=153%:%
%:%353=154%:%
%:%354=154%:%
%:%355=155%:%
%:%356=155%:%
%:%357=156%:%
%:%358=156%:%
%:%359=157%:%
%:%360=157%:%
%:%361=158%:%
%:%362=158%:%
%:%363=159%:%
%:%364=159%:%
%:%365=160%:%
%:%366=160%:%
%:%367=161%:%
%:%368=162%:%
%:%369=162%:%
%:%370=163%:%
%:%371=163%:%
%:%372=164%:%
%:%373=164%:%
%:%374=165%:%
%:%375=165%:%
%:%377=167%:%
%:%378=168%:%
%:%379=168%:%
%:%380=169%:%
%:%381=169%:%
%:%382=170%:%
%:%383=170%:%
%:%384=171%:%
%:%385=171%:%
%:%388=174%:%
%:%389=175%:%
%:%390=176%:%
%:%391=176%:%
%:%392=177%:%
%:%393=177%:%
%:%394=177%:%
%:%395=178%:%
%:%396=178%:%
%:%397=179%:%
%:%398=179%:%
%:%399=179%:%
%:%400=180%:%
%:%401=180%:%
%:%402=181%:%
%:%403=181%:%
%:%404=182%:%
%:%405=183%:%
%:%406=184%:%
%:%407=184%:%
%:%408=184%:%
%:%409=185%:%
%:%410=185%:%
%:%411=186%:%
%:%412=186%:%
%:%413=187%:%
%:%414=187%:%
%:%415=187%:%
%:%416=188%:%
%:%417=188%:%
%:%418=189%:%
%:%419=189%:%
%:%420=190%:%
%:%421=190%:%
%:%422=191%:%
%:%423=191%:%
%:%424=192%:%
%:%425=192%:%
%:%426=192%:%
%:%427=192%:%
%:%428=193%:%
%:%429=193%:%
%:%430=194%:%
%:%431=194%:%
%:%432=195%:%
%:%433=195%:%
%:%434=196%:%
%:%435=196%:%
%:%436=197%:%
%:%437=197%:%
%:%438=198%:%
%:%439=198%:%
%:%440=199%:%
%:%441=199%:%
%:%442=199%:%
%:%443=200%:%
%:%444=200%:%
%:%445=201%:%
%:%446=201%:%
%:%447=202%:%
%:%448=202%:%
%:%449=203%:%
%:%450=204%:%
%:%451=205%:%
%:%452=205%:%
%:%453=205%:%
%:%454=206%:%
%:%455=206%:%
%:%456=206%:%
%:%457=207%:%
%:%458=207%:%
%:%459=208%:%
%:%460=209%:%
%:%461=209%:%
%:%462=209%:%
%:%463=209%:%
%:%464=210%:%
%:%465=210%:%
%:%466=211%:%
%:%467=211%:%
%:%468=212%:%
%:%469=213%:%
%:%470=213%:%
%:%471=213%:%
%:%472=214%:%
%:%473=214%:%
%:%474=215%:%
%:%475=215%:%
%:%476=216%:%
%:%477=217%:%
%:%478=217%:%
%:%479=218%:%
%:%480=218%:%
%:%481=219%:%
%:%482=219%:%
%:%484=221%:%
%:%485=222%:%
%:%486=223%:%
%:%487=223%:%
%:%488=224%:%
%:%489=224%:%
%:%490=225%:%
%:%491=225%:%
%:%492=226%:%
%:%493=226%:%
%:%494=227%:%
%:%495=227%:%
%:%496=228%:%
%:%497=228%:%
%:%498=229%:%
%:%499=229%:%
%:%500=230%:%
%:%501=230%:%
%:%502=230%:%
%:%503=231%:%
%:%504=231%:%
%:%505=232%:%
%:%506=232%:%
%:%507=233%:%
%:%508=233%:%
%:%509=233%:%
%:%510=234%:%
%:%511=234%:%
%:%512=235%:%
%:%513=235%:%
%:%516=238%:%
%:%517=239%:%
%:%518=239%:%
%:%519=239%:%
%:%520=240%:%
%:%521=240%:%
%:%522=241%:%
%:%523=241%:%
%:%524=242%:%
%:%525=243%:%
%:%526=243%:%
%:%527=244%:%
%:%528=245%:%
%:%529=245%:%
%:%533=249%:%
%:%534=250%:%
%:%535=250%:%
%:%536=251%:%
%:%537=251%:%
%:%538=251%:%
%:%539=251%:%
%:%540=251%:%
%:%541=252%:%
%:%542=252%:%
%:%543=253%:%
%:%544=253%:%
%:%545=254%:%
%:%546=254%:%
%:%547=255%:%
%:%548=255%:%
%:%549=256%:%
%:%550=257%:%
%:%551=257%:%
%:%552=258%:%
%:%553=258%:%
%:%554=259%:%
%:%555=259%:%
%:%556=260%:%
%:%556=264%:%
%:%557=265%:%
%:%558=265%:%
%:%559=266%:%
%:%560=266%:%
%:%561=267%:%
%:%562=267%:%
%:%563=268%:%
%:%564=268%:%
%:%565=269%:%
%:%566=269%:%
%:%567=270%:%
%:%568=270%:%
%:%569=271%:%
%:%570=271%:%
%:%571=272%:%
%:%572=272%:%
%:%573=273%:%
%:%574=273%:%
%:%575=273%:%
%:%576=274%:%
%:%577=274%:%
%:%578=275%:%
%:%579=275%:%
%:%580=276%:%
%:%581=276%:%
%:%582=277%:%
%:%583=277%:%
%:%584=277%:%
%:%585=278%:%
%:%586=278%:%
%:%587=279%:%
%:%588=279%:%
%:%589=280%:%
%:%590=280%:%
%:%591=281%:%
%:%592=281%:%
%:%593=281%:%
%:%594=282%:%
%:%595=282%:%
%:%596=283%:%
%:%597=283%:%
%:%598=283%:%
%:%599=284%:%
%:%600=284%:%
%:%601=285%:%
%:%602=285%:%
%:%603=286%:%
%:%604=286%:%
%:%605=287%:%
%:%606=287%:%
%:%607=288%:%
%:%608=288%:%
%:%609=289%:%
%:%610=289%:%
%:%611=289%:%
%:%612=290%:%
%:%613=290%:%
%:%614=291%:%
%:%615=291%:%
%:%616=291%:%
%:%617=292%:%
%:%618=292%:%
%:%619=293%:%
%:%620=293%:%
%:%621=294%:%
%:%622=294%:%
%:%623=294%:%
%:%624=295%:%
%:%625=295%:%
%:%626=295%:%
%:%627=296%:%
%:%628=296%:%
%:%629=297%:%
%:%630=297%:%
%:%631=298%:%
%:%632=298%:%
%:%633=299%:%
%:%634=299%:%
%:%635=299%:%
%:%636=300%:%
%:%637=300%:%
%:%638=301%:%
%:%639=301%:%
%:%640=302%:%
%:%641=302%:%
%:%642=302%:%
%:%643=303%:%
%:%644=303%:%
%:%645=304%:%
%:%646=304%:%
%:%647=305%:%
%:%648=305%:%
%:%649=306%:%
%:%650=307%:%
%:%651=307%:%
%:%652=308%:%
%:%653=309%:%
%:%654=309%:%
%:%655=310%:%
%:%656=310%:%
%:%657=311%:%
%:%658=311%:%
%:%659=312%:%
%:%660=312%:%
%:%661=312%:%
%:%662=313%:%
%:%663=313%:%
%:%664=314%:%
%:%665=314%:%
%:%666=314%:%
%:%667=315%:%
%:%668=316%:%
%:%669=316%:%
%:%670=316%:%
%:%671=317%:%
%:%672=317%:%
%:%673=318%:%
%:%674=318%:%
%:%675=319%:%
%:%676=319%:%
%:%677=319%:%
%:%678=320%:%
%:%679=320%:%
%:%680=321%:%
%:%681=322%:%
%:%682=322%:%
%:%683=323%:%
%:%684=323%:%
%:%685=324%:%
%:%686=324%:%
%:%687=325%:%
%:%688=325%:%
%:%689=326%:%
%:%690=326%:%
%:%691=327%:%
%:%692=328%:%
%:%693=328%:%
%:%694=329%:%
%:%695=329%:%
%:%696=330%:%
%:%697=330%:%
%:%698=331%:%
%:%699=331%:%
%:%700=332%:%
%:%701=332%:%
%:%702=333%:%
%:%703=333%:%
%:%704=334%:%
%:%705=334%:%
%:%706=334%:%
%:%707=335%:%
%:%708=335%:%
%:%709=336%:%
%:%710=336%:%
%:%711=337%:%
%:%712=337%:%
%:%713=337%:%
%:%714=338%:%
%:%715=338%:%
%:%716=339%:%
%:%717=339%:%
%:%718=340%:%
%:%719=340%:%
%:%720=340%:%
%:%721=341%:%
%:%722=341%:%
%:%723=342%:%
%:%724=342%:%
%:%725=342%:%
%:%726=343%:%
%:%727=343%:%
%:%728=344%:%
%:%729=344%:%
%:%730=345%:%
%:%731=345%:%
%:%732=345%:%
%:%733=346%:%
%:%734=346%:%
%:%735=347%:%
%:%736=347%:%
%:%737=347%:%
%:%738=348%:%
%:%739=348%:%
%:%740=349%:%
%:%741=349%:%
%:%742=350%:%
%:%743=350%:%
%:%744=350%:%
%:%745=351%:%
%:%746=351%:%
%:%747=352%:%
%:%748=352%:%
%:%749=353%:%
%:%750=353%:%
%:%751=354%:%
%:%752=354%:%
%:%753=355%:%
%:%754=355%:%
%:%755=356%:%
%:%756=356%:%
%:%757=357%:%
%:%758=357%:%
%:%759=357%:%
%:%760=358%:%
%:%761=358%:%
%:%762=359%:%
%:%763=359%:%
%:%764=359%:%
%:%765=360%:%
%:%766=360%:%
%:%767=360%:%
%:%768=361%:%
%:%769=361%:%
%:%770=362%:%
%:%771=362%:%
%:%772=363%:%
%:%773=363%:%
%:%774=364%:%
%:%775=364%:%
%:%776=365%:%
%:%777=365%:%
%:%778=366%:%
%:%779=366%:%
%:%780=366%:%
%:%781=367%:%
%:%782=367%:%
%:%783=368%:%
%:%784=368%:%
%:%785=368%:%
%:%786=368%:%
%:%787=369%:%
%:%788=369%:%
%:%789=370%:%
%:%790=370%:%
%:%791=371%:%
%:%792=371%:%
%:%793=372%:%
%:%794=372%:%
%:%795=372%:%
%:%796=373%:%
%:%802=373%:%
%:%805=374%:%
%:%806=375%:%
%:%807=375%:%
%:%808=376%:%
%:%809=377%:%
%:%810=378%:%
%:%811=379%:%
%:%818=380%:%
%:%819=380%:%
%:%820=381%:%
%:%821=381%:%
%:%822=382%:%
%:%823=382%:%
%:%824=383%:%
%:%825=383%:%
%:%826=384%:%
%:%827=384%:%
%:%828=384%:%
%:%829=385%:%
%:%830=385%:%
%:%831=386%:%
%:%832=387%:%
%:%833=387%:%
%:%834=387%:%
%:%835=388%:%
%:%836=388%:%
%:%837=388%:%
%:%838=389%:%
%:%839=389%:%
%:%840=390%:%
%:%841=390%:%
%:%842=391%:%
%:%843=391%:%
%:%844=391%:%
%:%845=392%:%
%:%846=392%:%
%:%847=393%:%
%:%848=393%:%
%:%849=394%:%
%:%850=394%:%
%:%851=395%:%
%:%852=395%:%
%:%853=395%:%
%:%854=395%:%
%:%855=396%:%
%:%861=396%:%
%:%864=397%:%
%:%865=398%:%
%:%866=398%:%
%:%869=399%:%
%:%874=400%:%