### 2024-11-01 code/output/document/Separation_Rename.tex

%
\begin{isabellebody}%
\setisabellecontext{Separation{\isacharunderscore}{\kern0pt}Rename}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Auxiliary renamings for Separation%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Separation{\isacharunderscore}{\kern0pt}Rename\isanewline
\ \ \isakeyword{imports}\ Interface\ Renaming\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{lemmas}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}fun\ {\isacharequal}{\kern0pt}\ apply{\isacharunderscore}{\kern0pt}iff{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{1}}{\isacharbrackright}{\kern0pt}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}concat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ env{\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}{\isadigit{1}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}t{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ t{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}nth{\isacharunderscore}{\kern0pt}append{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}concat{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env{\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ nth{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ p{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}nth{\isacharunderscore}{\kern0pt}append{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}concat{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}env{\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ u\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ u{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}nth{\isacharunderscore}{\kern0pt}append{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ sep{\isacharunderscore}{\kern0pt}var\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{3}}{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{4}}{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isadigit{3}}{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isadigit{4}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{6}}{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isasymrangle}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ sep{\isacharunderscore}{\kern0pt}env\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}env{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymlambda}\ i\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}\ {\isachardot}{\kern0pt}\ i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ weak\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m\ {\isachardot}{\kern0pt}\ i\ {\isasymin}\ n{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ weakD\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}k{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}k{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}\ i\ {\isasymin}\ n\ {\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}k{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}equal\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}m{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}\ m{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isasymsubseteq}{\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}m{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}intro\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ i\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}i{\isasymin}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharequal}{\kern0pt}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ weakD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isasymle}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isachardoublequoteclose}\ {\isachardoublequoteopen}i{\isacharless}{\kern0pt}n{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self{\isadigit{2}}{\isacharbrackleft}{\kern0pt}of\ m\ i{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ ltI{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}i{\isasymin}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isacharless}{\kern0pt}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}i{\isasymin}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharequal}{\kern0pt}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymnotin}m{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltI\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ assms\ {\isacartoucheopen}x{\isacharequal}{\kern0pt}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isacartoucheclose}\ {\isacartoucheopen}i{\isacharless}{\kern0pt}n{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono{\isadigit{1}}{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}i{\isacharless}{\kern0pt}n{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}{\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}m{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltD\ DiffI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}m{\isasymsubseteq}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}intro\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}{\isacharparenleft}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymnotin}m{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ DiffD{\isadigit{1}}{\isacharbrackleft}{\kern0pt}of\ x\ {\isachardoublequoteopen}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isachardoublequoteclose}\ m{\isacharbrackright}{\kern0pt}\ DiffD{\isadigit{2}}{\isacharbrackleft}{\kern0pt}of\ x\ {\isachardoublequoteopen}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isachardoublequoteclose}\ m{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltI\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}OF\ add{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ m\ n{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ i\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ less{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}add{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}{\isacharcomma}{\kern0pt}of\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ add{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}i{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymnotin}m{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}x{\isacharless}{\kern0pt}m{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isasymle}x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacartoucheclose}\ \ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m\ {\isacharequal}{\kern0pt}\ \ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}cancel{\isadigit{2}}\ {\isacartoucheopen}m{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\ \ \ \isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharless}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacartoucheclose}\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m\ {\isasymin}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharequal}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}m{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}m{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltD\ add{\isacharunderscore}{\kern0pt}diff{\isacharunderscore}{\kern0pt}inverse{\isadigit{2}}{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}m{\isasymle}x{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}zero{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ weakening{\isacharunderscore}{\kern0pt}diff\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{7}}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}\ weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isacharbraceleft}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ i\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}i{\isasymin}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymnoteq}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}w{\isasymin}n{\isachardot}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ natify{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i{\isacharless}{\kern0pt}n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltI\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}i{\isasymin}n{\isacartoucheclose}\ {\isacartoucheopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymnoteq}n{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymnoteq}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}i{\isacharless}{\kern0pt}n{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isasymle}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ leD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ cases\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymle}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ succ{\isacharunderscore}{\kern0pt}leI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ leD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ cases\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymnoteq}n{\isacartoucheclose}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ b\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ natify{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}{\isasymforall}w{\isasymin}n{\isachardot}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ natify{\isacharparenleft}{\kern0pt}w{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}False{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ n\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ {\isasymforall}x{\isasymin}n{\isachardot}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ natify{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ y\ {\isasymin}\ n\ {\isasymLongrightarrow}\ succ{\isacharparenleft}{\kern0pt}natify{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isakeyword{for}\ y\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}del\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isasymin}j{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ j\ {\isasymor}\ x\ {\isasymin}\ weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}j{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}x{\isasymin}j{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ assms\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}j\ {\isasymle}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ j{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ False\ {\isacartoucheopen}j{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ ltI{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isasymin}j{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}j\ {\isacharless}{\kern0pt}\ {\isacharparenleft}{\kern0pt}j\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ n{\isacharparenright}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ j{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ j\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}mono\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}j{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}j{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ add{\isacharunderscore}{\kern0pt}diff{\isacharunderscore}{\kern0pt}inverse{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}j{\isasymle}x{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}j\ {\isacharless}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}j\ {\isacharparenright}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ j{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}inverse\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ add{\isacharunderscore}{\kern0pt}commute\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}j\ {\isasymin}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}j{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}j{\isacharparenright}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}j{\isacartoucheclose}\ RepFunI{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}j{\isasymin}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ add{\isacharunderscore}{\kern0pt}commute\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \ \isacommand{{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}env{\isacharunderscore}{\kern0pt}action{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}\ i\ {\isachardot}{\kern0pt}\ i\ {\isasymin}\ weak{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ nth{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}env{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ o{\isacharcomma}{\kern0pt}\ t{\isacharbrackright}{\kern0pt}\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}f{\isacharequal}{\kern0pt}{\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}env{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ EQ{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}\ {\isadigit{5}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}equal\ length{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}tgt{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}src{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}i\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ i\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ that\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymin}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}i\ {\isasymnotin}\ {\isadigit{5}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}i{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{3}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymnot}\ i\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{5}}\ {\isasymle}\ i{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{2}}\ {\isasymle}\ {\isadigit{5}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{2}}\ {\isasymle}\ i{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}trans{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isadigit{2}}{\isasymle}{\isadigit{5}}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ A\ {\isacartoucheopen}i\ {\isasymin}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}{\isadigit{2}}{\isasymle}i{\isacartoucheclose}\ A\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ C{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}\ {\isacharless}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ that\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ B{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isacharbackquote}{\kern0pt}i\ {\isacharequal}{\kern0pt}\ i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}env{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isadigit{3}}\ assms{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}\ {\isacharless}{\kern0pt}\ {\isadigit{7}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}mono\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}i\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isadigit{3}}\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ {\isadigit{5}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}mono{\isacharbrackleft}{\kern0pt}of\ i\ {\isachardoublequoteopen}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isadigit{5}}{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacharunderscore}{\kern0pt}\ {\isacartoucheopen}i\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le{\isacharbrackleft}{\kern0pt}THEN\ iffD{\isadigit{1}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ diff{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}inverse\ length{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}src{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}nth{\isacharparenleft}{\kern0pt}i{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}env{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}append{\isacharbrackleft}{\kern0pt}OF\ A{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isacartoucheopen}i{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isadigit{3}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}append{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env\ {\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}i{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}i{\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}tgt{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}append{\isacharbrackleft}{\kern0pt}OF\ assms{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacartoucheopen}i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ {\isacartoucheopen}{\isasymnot}\ i{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}\ {\isacharless}{\kern0pt}{\isadigit{7}}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}src{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}tgt{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ B\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ EQ\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}env{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}env{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}\ {\isasymrightarrow}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{7}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}h{\isacharequal}{\kern0pt}{\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}env{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}\ {\isacharequal}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isasymin}nat{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ D{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}env{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{7}}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{5}}{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharminus}{\kern0pt}{\isadigit{5}}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}env{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ ltI\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}\ {\isacharless}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}\ {\isasymin}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharminus}{\kern0pt}{\isadigit{5}}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymnotin}{\isadigit{5}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}x{\isacharless}{\kern0pt}{\isadigit{5}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{5}}{\isasymle}x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{7}}{\isasymle}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}mono\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharless}{\kern0pt}{\isadigit{7}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}\ {\isasymnotin}\ {\isadigit{7}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltI\ {\isacartoucheopen}x{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}\ {\isasymin}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacartoucheclose}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}{\isacharquery}{\kern0pt}h{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ x{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacartoucheclose}\ DiffI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}env{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ lam{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ \ {\isacharminus}{\kern0pt}{\isacharbar}{\kern0pt}{\isacharbar}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ consI\ ltD\ emptyI\ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}domain\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ \ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharminus}{\kern0pt}\ weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ le{\isacharunderscore}{\kern0pt}natE{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ C{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isacharequal}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymor}\ x{\isacharequal}{\kern0pt}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymor}\ x\ {\isasymle}\ {\isadigit{4}}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ D\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{7}}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ ltI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \ lt{\isacharunderscore}{\kern0pt}irrefl{\isacharbrackleft}{\kern0pt}of\ {\isacharunderscore}{\kern0pt}\ False{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ {\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ R{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ that\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ i\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}i{\isacharless}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharequal}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}i{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltI\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ RepFun{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}i\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}mono{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharequal}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}i{\isacartoucheclose}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isacharless}{\kern0pt}\ {\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}x{\isasymnotin}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}{\isasymnot}x\ {\isacharless}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymnotin}\ {\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isasymnotin}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymnotin}\ {\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ {\isacartoucheopen}{\isasymnot}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharless}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ blast\ \ \ \ \isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymnotin}\ {\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}{\isasymnot}{\isadigit{6}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharless}{\kern0pt}{\isadigit{5}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ E{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}x{\isasymnotin}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ C\ that\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ F{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharminus}{\kern0pt}\ {\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ A\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}{\isadigit{7}}\ {\isasymor}\ x{\isasymin}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{7}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}add{\isacharunderscore}{\kern0pt}del{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}x{\isasymin}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ asm{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}x{\isasymin}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}x{\isasymnotin}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}x{\isasymin}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{7}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{7}}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isasymsubseteq}{\isacharbraceleft}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{6}}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ weakening{\isacharunderscore}{\kern0pt}diff\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ \ {\isacartoucheopen}x{\isasymnotin}{\isacharquery}{\kern0pt}A{\isacartoucheclose}\ asm\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}{\isacharbraceleft}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{6}}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ \ subsetD\ DiffI\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ asm{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}x{\isasymin}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}x{\isasymnotin}{\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}{\isadigit{7}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isasymin}domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}{\isadigit{2}}\ {\isasymle}\ n{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{0}}{\isacharless}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ \ leD{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ {\isacartoucheopen}{\isadigit{2}}{\isasymle}n{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ lt{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharunderscore}{\kern0pt}lt\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharless}{\kern0pt}{\isadigit{5}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}x{\isacharless}{\kern0pt}{\isadigit{7}}{\isacartoucheclose}\ {\isacartoucheopen}x{\isasymnotin}{\isacharquery}{\kern0pt}A{\isacartoucheclose}\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ in{\isacharunderscore}{\kern0pt}n{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarsimp\ simp\ add{\isacharcolon}{\kern0pt}not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ auto\ simp\ add{\isacharcolon}{\kern0pt}lt{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarsimp\ simp\ add{\isacharcolon}{\kern0pt}not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le{\isacharcomma}{\kern0pt}\ auto\ simp\ add{\isacharcolon}{\kern0pt}lt{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}n{\isacharequal}{\kern0pt}{\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ ltD\ asm\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharless}{\kern0pt}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ \ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ {\isacartoucheopen}{\isasymnot}\ {\isadigit{2}}\ {\isasymle}\ n{\isacartoucheclose}\ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ n\ {\isacharless}{\kern0pt}{\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymnoteq}{\isadigit{0}}{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n{\isacharequal}{\kern0pt}{\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}le\ {\isacartoucheopen}n{\isacharless}{\kern0pt}{\isadigit{2}}{\isacartoucheclose}\ le{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymnotin}{\isacharquery}{\kern0pt}A{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}def\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltD\ asm\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}w{\isasymin}domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}w{\isasymin}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharminus}{\kern0pt}\ {\isacharquery}{\kern0pt}A{\isachardoublequoteclose}\ \isakeyword{for}\ w\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharminus}{\kern0pt}\ {\isacharquery}{\kern0pt}A\ {\isasymsubseteq}\ domain{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ F\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ FiniteFun{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun{\isacharbrackleft}{\kern0pt}OF\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}fin{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}domain{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}action\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}\ i\ {\isachardot}{\kern0pt}\ i\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}\ weak{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ nth{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subst\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}domain{\isacharbrackleft}{\kern0pt}OF\ length{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ i\ y\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}i{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Cons{\isacharparenleft}{\kern0pt}t{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}u{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}leq{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}p{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}leq{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}o{\isacharcomma}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}t{\isacharcomma}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}\ u{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}fun{\isacharbrackleft}{\kern0pt}OF\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}type{\isacharbrackright}{\kern0pt}\ assms\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ nth{\isacharunderscore}{\kern0pt}concat{\isadigit{2}}{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \ nth{\isacharunderscore}{\kern0pt}concat{\isadigit{3}}{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharcomma}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ rensep\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}rensep{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ union{\isacharunderscore}{\kern0pt}fun{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}sep{\isacharunderscore}{\kern0pt}env{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rensep{\isacharunderscore}{\kern0pt}aux\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}\ {\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ n\ {\isacharminus}{\kern0pt}\ {\isadigit{7}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{5}}{\isacharminus}{\kern0pt}{\isadigit{5}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ weak{\isacharunderscore}{\kern0pt}equal\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ \ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}\ {\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ n\ {\isacharminus}{\kern0pt}\ {\isadigit{7}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Diff{\isacharunderscore}{\kern0pt}partition\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rensep{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}rensep{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymrightarrow}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rensep{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ weak{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isasymunion}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n\ {\isacharminus}{\kern0pt}\ {\isadigit{7}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ rensep{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}type\ \ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}type\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ sep{\isacharunderscore}{\kern0pt}env{\isacharunderscore}{\kern0pt}type\ weak{\isacharunderscore}{\kern0pt}equal\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ rensep{\isacharunderscore}{\kern0pt}aux\ {\isacartoucheopen}n{\isasymin}nat{\isacartoucheclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ rensep{\isacharunderscore}{\kern0pt}action\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}\ i\ {\isachardot}{\kern0pt}\ i\ {\isacharless}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ nth{\isacharparenleft}{\kern0pt}rensep{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}tgt{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}env{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}src{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}m{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharminus}{\kern0pt}\ weak{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}p{\isacharequal}{\kern0pt}{\isachardoublequoteopen}weak{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}f{\isacharequal}{\kern0pt}{\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}var{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}g{\isacharequal}{\kern0pt}{\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}env{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}n{\isacharequal}{\kern0pt}{\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}src\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}tgt\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharquery}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ weak{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}\ length{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}src{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ weak{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Diff{\isacharunderscore}{\kern0pt}partition\ le{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}subset\ rensep{\isacharunderscore}{\kern0pt}aux\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}src{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ nth{\isacharparenleft}{\kern0pt}union{\isacharunderscore}{\kern0pt}fun{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}g{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}m{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}p{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ i{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}tgt{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ i\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}i{\isacharless}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ weak{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ simp\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ rensep{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ \ \isanewline
\ \ \ \ \ \ \ \ union{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}action{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharquery}{\kern0pt}src{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}tgt{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}length{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}src{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharminus}{\kern0pt}weak{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymunion}\ weak{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \ \ \ \ sep{\isacharunderscore}{\kern0pt}var{\isacharunderscore}{\kern0pt}action{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ \ \ sep{\isacharunderscore}{\kern0pt}env{\isacharunderscore}{\kern0pt}action{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isacharbrackright}{\kern0pt}\ that\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ rensep{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}ren\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}ren{\isacharparenleft}{\kern0pt}n{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ ren{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}rensep{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}rensep{\isacharcolon}{\kern0pt}\ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isakeyword{shows}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}ren{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}ren\ rensep{\isacharunderscore}{\kern0pt}type\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}rensep\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}ren{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ ren{\isacharunderscore}{\kern0pt}tc\ rensep{\isacharunderscore}{\kern0pt}type\ assms\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sepren{\isacharunderscore}{\kern0pt}action{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymphi}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}ren{\isacharparenleft}{\kern0pt}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}u{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}pi{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymphi}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}\ {\isacharbrackleft}{\kern0pt}t{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ u{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ leq{\isacharcomma}{\kern0pt}\ o{\isacharcomma}{\kern0pt}\ pi{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}p{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ \ \ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}p{\isacharcomma}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}o{\isacharcomma}{\kern0pt}t{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ env\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}pi{\isacharcomma}{\kern0pt}u{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ app{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}ren{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}ren{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}{\isasymphi}{\isasymin}formula{\isacartoucheclose}\ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ add{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ {\isadigit{7}}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ add{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ {\isadigit{7}}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ {\isadigit{2}}\ {\isadigit{1}}{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ rensep{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}OF\ length{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}{\isasymphi}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{7}}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ rensep{\isacharunderscore}{\kern0pt}action{\isacharbrackleft}{\kern0pt}OF\ {\isadigit{1}}{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}rule{\isacharunderscore}{\kern0pt}format{\isacharcomma}{\kern0pt}symmetric{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Separation_Rename.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%35=4%:%
%:%38=5%:%
%:%39=6%:%
%:%40=6%:%
%:%41=7%:%
%:%42=8%:%
%:%43=8%:%
%:%46=9%:%
%:%50=9%:%
%:%51=9%:%
%:%56=9%:%
%:%59=10%:%
%:%60=11%:%
%:%61=11%:%
%:%64=12%:%
%:%68=12%:%
%:%69=12%:%
%:%74=12%:%
%:%77=13%:%
%:%78=14%:%
%:%79=14%:%
%:%82=15%:%
%:%86=15%:%
%:%87=15%:%
%:%92=15%:%
%:%95=16%:%
%:%96=17%:%
%:%97=17%:%
%:%98=18%:%
%:%99=19%:%
%:%100=20%:%
%:%101=21%:%
%:%102=21%:%
%:%103=22%:%
%:%104=23%:%
%:%105=24%:%
%:%106=25%:%
%:%107=25%:%
%:%108=26%:%
%:%109=27%:%
%:%110=28%:%
%:%111=28%:%
%:%112=29%:%
%:%113=30%:%
%:%116=31%:%
%:%120=31%:%
%:%121=31%:%
%:%122=31%:%
%:%123=31%:%
%:%128=31%:%
%:%131=32%:%
%:%132=33%:%
%:%133=33%:%
%:%134=34%:%
%:%135=35%:%
%:%142=36%:%
%:%143=36%:%
%:%144=37%:%
%:%145=37%:%
%:%146=38%:%
%:%147=38%:%
%:%148=39%:%
%:%149=39%:%
%:%150=40%:%
%:%151=40%:%
%:%152=41%:%
%:%153=41%:%
%:%154=42%:%
%:%155=42%:%
%:%156=43%:%
%:%157=44%:%
%:%158=44%:%
%:%159=44%:%
%:%160=45%:%
%:%161=45%:%
%:%162=46%:%
%:%163=46%:%
%:%164=47%:%
%:%165=47%:%
%:%166=47%:%
%:%167=48%:%
%:%168=48%:%
%:%169=49%:%
%:%170=49%:%
%:%171=50%:%
%:%172=50%:%
%:%173=50%:%
%:%174=51%:%
%:%175=51%:%
%:%176=52%:%
%:%177=52%:%
%:%178=53%:%
%:%179=53%:%
%:%180=53%:%
%:%181=54%:%
%:%182=54%:%
%:%183=55%:%
%:%184=55%:%
%:%185=56%:%
%:%186=56%:%
%:%187=57%:%
%:%188=57%:%
%:%189=57%:%
%:%190=58%:%
%:%191=58%:%
%:%192=59%:%
%:%193=59%:%
%:%194=60%:%
%:%195=60%:%
%:%196=60%:%
%:%197=61%:%
%:%198=61%:%
%:%199=62%:%
%:%200=62%:%
%:%201=63%:%
%:%202=63%:%
%:%203=64%:%
%:%204=64%:%
%:%205=65%:%
%:%206=65%:%
%:%207=66%:%
%:%208=66%:%
%:%209=67%:%
%:%210=67%:%
%:%211=68%:%
%:%212=68%:%
%:%213=69%:%
%:%214=69%:%
%:%215=69%:%
%:%216=70%:%
%:%217=70%:%
%:%218=71%:%
%:%219=71%:%
%:%220=72%:%
%:%221=72%:%
%:%222=72%:%
%:%223=73%:%
%:%224=73%:%
%:%225=74%:%
%:%226=74%:%
%:%227=75%:%
%:%228=76%:%
%:%229=76%:%
%:%230=76%:%
%:%231=77%:%
%:%232=77%:%
%:%233=78%:%
%:%234=78%:%
%:%235=78%:%
%:%236=78%:%
%:%237=79%:%
%:%238=79%:%
%:%239=80%:%
%:%240=80%:%
%:%241=80%:%
%:%242=80%:%
%:%243=81%:%
%:%244=81%:%
%:%245=82%:%
%:%246=82%:%
%:%247=82%:%
%:%248=82%:%
%:%249=83%:%
%:%250=83%:%
%:%251=84%:%
%:%252=84%:%
%:%253=85%:%
%:%254=85%:%
%:%255=85%:%
%:%256=86%:%
%:%257=86%:%
%:%258=86%:%
%:%259=86%:%
%:%260=87%:%
%:%261=87%:%
%:%262=88%:%
%:%263=88%:%
%:%264=89%:%
%:%265=89%:%
%:%266=89%:%
%:%267=90%:%
%:%268=90%:%
%:%269=91%:%
%:%270=91%:%
%:%271=92%:%
%:%272=92%:%
%:%273=92%:%
%:%274=93%:%
%:%275=93%:%
%:%276=94%:%
%:%277=94%:%
%:%278=95%:%
%:%279=95%:%
%:%280=95%:%
%:%281=96%:%
%:%287=96%:%
%:%290=97%:%
%:%291=98%:%
%:%292=98%:%
%:%293=99%:%
%:%296=100%:%
%:%300=100%:%
%:%301=100%:%
%:%302=100%:%
%:%307=100%:%
%:%310=101%:%
%:%311=102%:%
%:%312=102%:%
%:%313=103%:%
%:%314=104%:%
%:%317=105%:%
%:%321=105%:%
%:%322=105%:%
%:%323=105%:%
%:%324=106%:%
%:%325=106%:%
%:%326=107%:%
%:%327=107%:%
%:%328=108%:%
%:%329=108%:%
%:%330=109%:%
%:%331=109%:%
%:%332=110%:%
%:%333=110%:%
%:%334=111%:%
%:%335=111%:%
%:%336=112%:%
%:%337=112%:%
%:%338=112%:%
%:%339=113%:%
%:%340=113%:%
%:%341=114%:%
%:%342=114%:%
%:%343=114%:%
%:%344=114%:%
%:%345=115%:%
%:%346=115%:%
%:%347=116%:%
%:%348=116%:%
%:%349=116%:%
%:%350=116%:%
%:%351=117%:%
%:%352=117%:%
%:%353=118%:%
%:%354=118%:%
%:%355=119%:%
%:%356=119%:%
%:%357=119%:%
%:%358=120%:%
%:%359=120%:%
%:%360=120%:%
%:%361=121%:%
%:%362=121%:%
%:%363=122%:%
%:%364=122%:%
%:%365=123%:%
%:%366=123%:%
%:%367=123%:%
%:%368=123%:%
%:%369=124%:%
%:%370=124%:%
%:%371=125%:%
%:%372=125%:%
%:%373=126%:%
%:%374=126%:%
%:%375=127%:%
%:%376=127%:%
%:%377=127%:%
%:%378=127%:%
%:%379=128%:%
%:%380=128%:%
%:%381=129%:%
%:%382=129%:%
%:%383=130%:%
%:%384=130%:%
%:%385=130%:%
%:%386=131%:%
%:%387=131%:%
%:%388=131%:%
%:%389=132%:%
%:%390=132%:%
%:%391=133%:%
%:%392=133%:%
%:%393=134%:%
%:%394=134%:%
%:%395=134%:%
%:%396=134%:%
%:%397=135%:%
%:%398=135%:%
%:%399=136%:%
%:%400=136%:%
%:%401=137%:%
%:%402=137%:%
%:%403=138%:%
%:%404=138%:%
%:%405=138%:%
%:%406=138%:%
%:%407=139%:%
%:%408=139%:%
%:%409=140%:%
%:%410=140%:%
%:%411=141%:%
%:%412=141%:%
%:%413=141%:%
%:%414=142%:%
%:%415=142%:%
%:%416=143%:%
%:%417=143%:%
%:%418=143%:%
%:%419=143%:%
%:%420=144%:%
%:%421=144%:%
%:%422=144%:%
%:%423=144%:%
%:%424=145%:%
%:%425=145%:%
%:%426=146%:%
%:%427=146%:%
%:%428=146%:%
%:%429=146%:%
%:%430=147%:%
%:%431=147%:%
%:%432=148%:%
%:%433=148%:%
%:%434=148%:%
%:%435=148%:%
%:%436=149%:%
%:%437=149%:%
%:%438=150%:%
%:%439=150%:%
%:%440=151%:%
%:%441=151%:%
%:%444=154%:%
%:%445=155%:%
%:%446=155%:%
%:%447=156%:%
%:%453=156%:%
%:%456=157%:%
%:%457=158%:%
%:%458=158%:%
%:%459=159%:%
%:%460=160%:%
%:%467=161%:%
%:%468=161%:%
%:%469=162%:%
%:%470=162%:%
%:%471=163%:%
%:%472=163%:%
%:%473=163%:%
%:%474=163%:%
%:%475=164%:%
%:%476=164%:%
%:%477=165%:%
%:%478=165%:%
%:%479=166%:%
%:%480=166%:%
%:%481=167%:%
%:%482=167%:%
%:%483=167%:%
%:%484=168%:%
%:%485=168%:%
%:%486=169%:%
%:%487=169%:%
%:%488=170%:%
%:%489=170%:%
%:%490=170%:%
%:%491=171%:%
%:%492=171%:%
%:%493=172%:%
%:%494=172%:%
%:%495=173%:%
%:%496=173%:%
%:%497=174%:%
%:%498=174%:%
%:%499=175%:%
%:%500=175%:%
%:%501=176%:%
%:%502=176%:%
%:%503=177%:%
%:%504=177%:%
%:%505=177%:%
%:%506=178%:%
%:%507=178%:%
%:%508=179%:%
%:%509=179%:%
%:%510=179%:%
%:%511=179%:%
%:%512=180%:%
%:%513=180%:%
%:%514=181%:%
%:%515=181%:%
%:%516=182%:%
%:%517=182%:%
%:%518=183%:%
%:%519=183%:%
%:%520=183%:%
%:%521=184%:%
%:%522=184%:%
%:%523=184%:%
%:%524=184%:%
%:%525=185%:%
%:%531=185%:%
%:%534=186%:%
%:%535=187%:%
%:%536=188%:%
%:%537=188%:%
%:%538=189%:%
%:%539=190%:%
%:%540=191%:%
%:%541=192%:%
%:%542=193%:%
%:%549=194%:%
%:%550=194%:%
%:%551=195%:%
%:%552=195%:%
%:%553=196%:%
%:%554=196%:%
%:%555=197%:%
%:%556=197%:%
%:%557=198%:%
%:%558=198%:%
%:%559=199%:%
%:%560=199%:%
%:%561=200%:%
%:%562=200%:%
%:%563=200%:%
%:%564=201%:%
%:%565=201%:%
%:%566=202%:%
%:%567=202%:%
%:%568=203%:%
%:%569=203%:%
%:%570=204%:%
%:%571=205%:%
%:%572=205%:%
%:%573=206%:%
%:%574=206%:%
%:%575=207%:%
%:%576=207%:%
%:%577=208%:%
%:%578=208%:%
%:%579=208%:%
%:%580=209%:%
%:%581=209%:%
%:%582=210%:%
%:%583=210%:%
%:%584=210%:%
%:%585=210%:%
%:%586=211%:%
%:%587=211%:%
%:%588=212%:%
%:%589=212%:%
%:%590=213%:%
%:%591=213%:%
%:%592=213%:%
%:%593=214%:%
%:%594=214%:%
%:%595=214%:%
%:%596=214%:%
%:%597=214%:%
%:%598=215%:%
%:%599=215%:%
%:%600=216%:%
%:%601=216%:%
%:%602=216%:%
%:%603=216%:%
%:%604=217%:%
%:%605=217%:%
%:%606=218%:%
%:%607=218%:%
%:%608=218%:%
%:%609=219%:%
%:%610=219%:%
%:%611=220%:%
%:%612=220%:%
%:%613=220%:%
%:%614=220%:%
%:%615=221%:%
%:%616=221%:%
%:%617=222%:%
%:%618=222%:%
%:%619=222%:%
%:%620=222%:%
%:%621=223%:%
%:%622=223%:%
%:%623=224%:%
%:%624=224%:%
%:%625=225%:%
%:%626=225%:%
%:%627=226%:%
%:%628=226%:%
%:%629=227%:%
%:%630=227%:%
%:%631=228%:%
%:%632=228%:%
%:%633=228%:%
%:%634=228%:%
%:%635=229%:%
%:%636=229%:%
%:%637=230%:%
%:%638=230%:%
%:%639=230%:%
%:%640=231%:%
%:%641=231%:%
%:%642=232%:%
%:%643=232%:%
%:%644=233%:%
%:%645=233%:%
%:%646=233%:%
%:%647=234%:%
%:%648=234%:%
%:%649=235%:%
%:%650=235%:%
%:%651=236%:%
%:%652=236%:%
%:%653=236%:%
%:%654=237%:%
%:%655=237%:%
%:%656=238%:%
%:%657=238%:%
%:%658=239%:%
%:%659=239%:%
%:%660=239%:%
%:%661=240%:%
%:%662=240%:%
%:%663=240%:%
%:%664=240%:%
%:%665=240%:%
%:%666=241%:%
%:%667=241%:%
%:%668=242%:%
%:%669=242%:%
%:%670=242%:%
%:%671=242%:%
%:%672=242%:%
%:%673=243%:%
%:%679=243%:%
%:%682=244%:%
%:%683=245%:%
%:%684=245%:%
%:%685=246%:%
%:%686=247%:%
%:%693=248%:%
%:%694=248%:%
%:%695=249%:%
%:%696=249%:%
%:%697=250%:%
%:%698=250%:%
%:%699=251%:%
%:%700=251%:%
%:%701=251%:%
%:%702=252%:%
%:%703=252%:%
%:%704=253%:%
%:%705=254%:%
%:%706=254%:%
%:%707=255%:%
%:%708=255%:%
%:%709=256%:%
%:%710=256%:%
%:%711=257%:%
%:%712=257%:%
%:%713=257%:%
%:%714=257%:%
%:%715=258%:%
%:%716=258%:%
%:%717=259%:%
%:%718=259%:%
%:%719=259%:%
%:%720=260%:%
%:%721=260%:%
%:%722=261%:%
%:%723=261%:%
%:%724=261%:%
%:%725=261%:%
%:%726=262%:%
%:%727=262%:%
%:%728=263%:%
%:%729=263%:%
%:%730=263%:%
%:%731=264%:%
%:%732=264%:%
%:%733=264%:%
%:%734=264%:%
%:%735=264%:%
%:%736=265%:%
%:%737=265%:%
%:%738=265%:%
%:%739=265%:%
%:%740=265%:%
%:%741=266%:%
%:%742=266%:%
%:%743=266%:%
%:%744=266%:%
%:%745=266%:%
%:%746=267%:%
%:%747=267%:%
%:%748=267%:%
%:%749=267%:%
%:%750=267%:%
%:%751=268%:%
%:%752=268%:%
%:%753=268%:%
%:%754=268%:%
%:%755=268%:%
%:%756=269%:%
%:%757=269%:%
%:%758=269%:%
%:%759=269%:%
%:%760=269%:%
%:%761=270%:%
%:%762=270%:%
%:%763=271%:%
%:%764=271%:%
%:%765=271%:%
%:%766=271%:%
%:%767=271%:%
%:%768=271%:%
%:%769=272%:%
%:%775=272%:%
%:%778=273%:%
%:%779=274%:%
%:%780=274%:%
%:%781=275%:%
%:%782=276%:%
%:%785=277%:%
%:%789=277%:%
%:%790=277%:%
%:%791=278%:%
%:%792=278%:%
%:%793=278%:%
%:%798=278%:%
%:%801=279%:%
%:%802=280%:%
%:%803=280%:%
%:%804=281%:%
%:%805=282%:%
%:%812=283%:%
%:%813=283%:%
%:%814=284%:%
%:%815=284%:%
%:%816=285%:%
%:%817=285%:%
%:%818=286%:%
%:%819=286%:%
%:%820=287%:%
%:%821=287%:%
%:%822=288%:%
%:%823=288%:%
%:%824=289%:%
%:%825=289%:%
%:%826=289%:%
%:%827=289%:%
%:%828=290%:%
%:%829=290%:%
%:%830=291%:%
%:%831=291%:%
%:%832=291%:%
%:%833=292%:%
%:%834=292%:%
%:%835=292%:%
%:%836=292%:%
%:%837=293%:%
%:%838=293%:%
%:%839=293%:%
%:%840=293%:%
%:%841=294%:%
%:%842=294%:%
%:%843=295%:%
%:%844=295%:%
%:%845=296%:%
%:%846=296%:%
%:%847=297%:%
%:%848=297%:%
%:%849=298%:%
%:%850=299%:%
%:%851=299%:%
%:%852=300%:%
%:%853=300%:%
%:%854=300%:%
%:%855=301%:%
%:%856=301%:%
%:%857=302%:%
%:%858=302%:%
%:%859=302%:%
%:%860=302%:%
%:%861=303%:%
%:%862=303%:%
%:%863=304%:%
%:%864=304%:%
%:%865=304%:%
%:%866=305%:%
%:%867=305%:%
%:%868=306%:%
%:%869=306%:%
%:%870=307%:%
%:%871=307%:%
%:%872=307%:%
%:%873=307%:%
%:%874=308%:%
%:%875=308%:%
%:%876=309%:%
%:%877=309%:%
%:%878=310%:%
%:%879=310%:%
%:%880=310%:%
%:%881=310%:%
%:%882=311%:%
%:%883=311%:%
%:%884=311%:%
%:%885=311%:%
%:%886=311%:%
%:%887=312%:%
%:%888=312%:%
%:%889=313%:%
%:%890=313%:%
%:%891=314%:%
%:%892=314%:%
%:%893=315%:%
%:%894=315%:%
%:%895=316%:%
%:%896=316%:%
%:%897=316%:%
%:%898=317%:%
%:%899=317%:%
%:%900=318%:%
%:%901=318%:%
%:%902=318%:%
%:%903=318%:%
%:%904=319%:%
%:%905=319%:%
%:%906=320%:%
%:%907=320%:%
%:%908=321%:%
%:%909=321%:%
%:%910=321%:%
%:%911=322%:%
%:%912=322%:%
%:%913=323%:%
%:%914=323%:%
%:%915=324%:%
%:%916=324%:%
%:%917=325%:%
%:%918=325%:%
%:%919=326%:%
%:%920=326%:%
%:%921=327%:%
%:%922=327%:%
%:%923=328%:%
%:%924=328%:%
%:%925=329%:%
%:%926=329%:%
%:%927=330%:%
%:%928=330%:%
%:%929=331%:%
%:%930=331%:%
%:%931=331%:%
%:%932=332%:%
%:%933=332%:%
%:%934=333%:%
%:%935=333%:%
%:%936=333%:%
%:%937=333%:%
%:%938=334%:%
%:%939=334%:%
%:%940=335%:%
%:%941=335%:%
%:%942=335%:%
%:%943=335%:%
%:%944=336%:%
%:%945=336%:%
%:%946=337%:%
%:%947=337%:%
%:%948=338%:%
%:%949=338%:%
%:%950=339%:%
%:%951=339%:%
%:%952=340%:%
%:%953=340%:%
%:%954=341%:%
%:%955=341%:%
%:%956=342%:%
%:%957=342%:%
%:%958=342%:%
%:%959=343%:%
%:%960=343%:%
%:%961=344%:%
%:%962=344%:%
%:%963=345%:%
%:%964=345%:%
%:%965=346%:%
%:%966=346%:%
%:%967=346%:%
%:%968=346%:%
%:%969=347%:%
%:%970=347%:%
%:%971=348%:%
%:%972=348%:%
%:%973=349%:%
%:%974=349%:%
%:%975=350%:%
%:%976=350%:%
%:%977=351%:%
%:%978=351%:%
%:%979=352%:%
%:%980=352%:%
%:%981=353%:%
%:%982=353%:%
%:%983=353%:%
%:%984=354%:%
%:%985=354%:%
%:%986=355%:%
%:%987=355%:%
%:%988=356%:%
%:%989=356%:%
%:%990=357%:%
%:%991=357%:%
%:%992=358%:%
%:%993=358%:%
%:%994=359%:%
%:%995=359%:%
%:%996=360%:%
%:%997=360%:%
%:%998=361%:%
%:%999=361%:%
%:%1000=361%:%
%:%1001=362%:%
%:%1002=362%:%
%:%1003=362%:%
%:%1004=362%:%
%:%1005=363%:%
%:%1006=363%:%
%:%1007=364%:%
%:%1008=364%:%
%:%1009=365%:%
%:%1010=365%:%
%:%1011=366%:%
%:%1012=366%:%
%:%1013=366%:%
%:%1014=366%:%
%:%1015=367%:%
%:%1016=367%:%
%:%1017=368%:%
%:%1018=368%:%
%:%1019=368%:%
%:%1020=368%:%
%:%1021=369%:%
%:%1022=369%:%
%:%1023=370%:%
%:%1024=370%:%
%:%1025=370%:%
%:%1026=370%:%
%:%1027=371%:%
%:%1028=371%:%
%:%1029=371%:%
%:%1030=372%:%
%:%1031=372%:%
%:%1032=373%:%
%:%1033=373%:%
%:%1034=374%:%
%:%1035=374%:%
%:%1036=374%:%
%:%1037=375%:%
%:%1038=375%:%
%:%1039=376%:%
%:%1040=376%:%
%:%1041=377%:%
%:%1042=377%:%
%:%1043=378%:%
%:%1044=378%:%
%:%1045=379%:%
%:%1046=379%:%
%:%1047=380%:%
%:%1048=380%:%
%:%1049=380%:%
%:%1050=381%:%
%:%1051=381%:%
%:%1052=382%:%
%:%1053=382%:%
%:%1054=382%:%
%:%1055=383%:%
%:%1056=383%:%
%:%1057=384%:%
%:%1058=384%:%
%:%1059=384%:%
%:%1060=385%:%
%:%1066=385%:%
%:%1069=386%:%
%:%1070=387%:%
%:%1071=387%:%
%:%1072=388%:%
%:%1073=389%:%
%:%1076=390%:%
%:%1080=390%:%
%:%1081=390%:%
%:%1082=391%:%
%:%1083=391%:%
%:%1088=391%:%
%:%1091=392%:%
%:%1092=393%:%
%:%1093=393%:%
%:%1094=394%:%
%:%1095=395%:%
%:%1096=396%:%
%:%1097=397%:%
%:%1098=398%:%
%:%1101=399%:%
%:%1105=399%:%
%:%1106=399%:%
%:%1107=400%:%
%:%1108=400%:%
%:%1109=401%:%
%:%1110=401%:%
%:%1111=402%:%
%:%1112=402%:%
%:%1113=403%:%
%:%1114=403%:%
%:%1115=404%:%
%:%1116=404%:%
%:%1118=406%:%
%:%1119=407%:%
%:%1120=407%:%
%:%1121=408%:%
%:%1122=408%:%
%:%1123=409%:%
%:%1124=409%:%
%:%1125=410%:%
%:%1126=410%:%
%:%1127=411%:%
%:%1133=411%:%
%:%1136=412%:%
%:%1137=413%:%
%:%1138=413%:%
%:%1139=414%:%
%:%1140=415%:%
%:%1141=416%:%
%:%1142=417%:%
%:%1143=417%:%
%:%1144=418%:%
%:%1145=419%:%
%:%1152=420%:%
%:%1153=420%:%
%:%1154=421%:%
%:%1155=421%:%
%:%1156=422%:%
%:%1157=422%:%
%:%1158=423%:%
%:%1159=423%:%
%:%1160=423%:%
%:%1161=424%:%
%:%1162=424%:%
%:%1163=425%:%
%:%1164=425%:%
%:%1165=426%:%
%:%1166=426%:%
%:%1167=426%:%
%:%1168=427%:%
%:%1174=427%:%
%:%1177=428%:%
%:%1178=429%:%
%:%1179=429%:%
%:%1180=430%:%
%:%1181=431%:%
%:%1188=432%:%
%:%1189=432%:%
%:%1190=433%:%
%:%1191=433%:%
%:%1192=434%:%
%:%1193=434%:%
%:%1194=435%:%
%:%1195=435%:%
%:%1196=436%:%
%:%1197=436%:%
%:%1198=437%:%
%:%1199=437%:%
%:%1200=438%:%
%:%1201=438%:%
%:%1202=439%:%
%:%1203=439%:%
%:%1204=439%:%
%:%1205=439%:%
%:%1206=440%:%
%:%1212=440%:%
%:%1215=441%:%
%:%1216=442%:%
%:%1217=442%:%
%:%1218=443%:%
%:%1219=444%:%
%:%1226=445%:%
%:%1227=445%:%
%:%1228=446%:%
%:%1229=446%:%
%:%1230=447%:%
%:%1231=447%:%
%:%1232=448%:%
%:%1233=448%:%
%:%1234=449%:%
%:%1235=449%:%
%:%1236=450%:%
%:%1237=450%:%
%:%1238=451%:%
%:%1239=451%:%
%:%1240=452%:%
%:%1241=452%:%
%:%1242=453%:%
%:%1243=453%:%
%:%1244=454%:%
%:%1245=454%:%
%:%1246=455%:%
%:%1247=456%:%
%:%1248=457%:%
%:%1249=458%:%
%:%1250=458%:%
%:%1251=458%:%
%:%1252=459%:%
%:%1253=459%:%
%:%1254=460%:%
%:%1255=460%:%
%:%1256=461%:%
%:%1257=461%:%
%:%1258=462%:%
%:%1259=462%:%
%:%1260=463%:%
%:%1261=463%:%
%:%1262=464%:%
%:%1263=464%:%
%:%1264=464%:%
%:%1265=465%:%
%:%1266=465%:%
%:%1267=465%:%
%:%1268=466%:%
%:%1269=466%:%
%:%1270=466%:%
%:%1271=467%:%
%:%1272=468%:%
%:%1273=469%:%
%:%1274=470%:%
%:%1275=471%:%
%:%1276=471%:%
%:%1277=472%:%
%:%1278=472%:%
%:%1279=473%:%
%:%1280=473%:%
%:%1281=473%:%
%:%1282=473%:%
%:%1283=473%:%
%:%1284=474%:%
%:%1290=474%:%
%:%1293=475%:%
%:%1294=476%:%
%:%1295=476%:%
%:%1296=477%:%
%:%1297=478%:%
%:%1298=479%:%
%:%1299=479%:%
%:%1300=480%:%
%:%1301=481%:%
%:%1304=482%:%
%:%1308=482%:%
%:%1309=482%:%
%:%1310=483%:%
%:%1311=483%:%
%:%1312=484%:%
%:%1313=484%:%
%:%1318=484%:%
%:%1321=485%:%
%:%1322=486%:%
%:%1323=486%:%
%:%1324=487%:%
%:%1325=488%:%
%:%1328=489%:%
%:%1332=489%:%
%:%1333=489%:%
%:%1334=490%:%
%:%1335=490%:%
%:%1336=491%:%
%:%1337=491%:%
%:%1342=491%:%
%:%1345=492%:%
%:%1346=493%:%
%:%1347=493%:%
%:%1348=494%:%
%:%1349=495%:%
%:%1350=496%:%
%:%1351=497%:%
%:%1352=498%:%
%:%1359=499%:%
%:%1360=499%:%
%:%1361=500%:%
%:%1362=500%:%
%:%1363=501%:%
%:%1364=501%:%
%:%1365=502%:%
%:%1366=503%:%
%:%1367=504%:%
%:%1368=504%:%
%:%1369=505%:%
%:%1370=505%:%
%:%1371=506%:%
%:%1372=506%:%
%:%1373=506%:%
%:%1374=506%:%
%:%1375=507%:%
%:%1376=507%:%
%:%1377=508%:%
%:%1378=508%:%
%:%1379=509%:%
%:%1380=509%:%
%:%1381=510%:%
%:%1382=511%:%
%:%1383=512%:%
%:%1384=513%:%
%:%1385=514%:%
%:%1386=515%:%
%:%1387=516%:%
%:%1388=516%:%
%:%1389=517%:%
%:%1395=517%:%
%:%1400=518%:%
%:%1405=519%:%