### 2024-11-01 code/output/document/Succession_Poset.tex

%
\begin{isabellebody}%
\setisabellecontext{Succession{\isacharunderscore}{\kern0pt}Poset}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{A poset of successions%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Succession{\isacharunderscore}{\kern0pt}Poset\isanewline
\ \ \isakeyword{imports}\isanewline
\ \ \ \ Arities\ Proper{\isacharunderscore}{\kern0pt}Extension\ Synthetic{\isacharunderscore}{\kern0pt}Definition\isanewline
\ \ \ \ Names\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{The set of finite binary sequences%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
We implement the poset for adding one Cohen real, the set 
$2^{<\omega}$ of of finite binary sequences.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ seqspace\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}{\isachardoublequoteopen}{\isacharunderscore}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ {\isacharbrackleft}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isadigit{0}}{\isacharbrackright}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}seqspace{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymUnion}n{\isasymin}nat{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}n{\isasymrightarrow}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seqspaceI{\isacharbrackleft}{\kern0pt}intro{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n{\isasymin}nat\ {\isasymLongrightarrow}\ f{\isacharcolon}{\kern0pt}n{\isasymrightarrow}B\ {\isasymLongrightarrow}\ f{\isasymin}seqspace{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seqspaceD{\isacharbrackleft}{\kern0pt}dest{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f{\isasymin}seqspace{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}n{\isasymin}nat{\isachardot}{\kern0pt}\ f{\isacharcolon}{\kern0pt}n{\isasymrightarrow}B{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}type{\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}f\ {\isasymin}\ B{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}\ {\isasymLongrightarrow}\ {\isasymexists}n{\isasymin}nat{\isachardot}{\kern0pt}\ f{\isacharcolon}{\kern0pt}n{\isasymrightarrow}B{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ z{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}h{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ B{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}i\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}h{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymexists}om{\isasymin}A{\isachardot}{\kern0pt}\ omega{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}om{\isacharparenright}{\kern0pt}\ {\isasymand}\ n\ {\isasymin}\ om\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}funspace{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}A{\isacharcomma}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ B{\isacharcomma}{\kern0pt}\ z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ env\ {\isasymTurnstile}\ {\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}sqsprp{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}j{\isacharcomma}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}funspace{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimML
\isanewline
%
\endisadelimML
%
\isatagML
\isacommand{synthesize}\isamarkupfalse%
\ {\isachardoublequoteopen}seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isachardoublequoteclose}\ \isakeyword{from{\isacharunderscore}{\kern0pt}schematic}\ seqspace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto%
\endisatagML
{\isafoldML}%
%
\isadelimML
\isanewline
%
\endisadelimML
\ \isanewline
\isacommand{locale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}seqspace\ {\isacharequal}{\kern0pt}\ \ M{\isacharunderscore}{\kern0pt}trancl\ {\isacharplus}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ \isanewline
\ \ \ \ seqspace{\isacharunderscore}{\kern0pt}replacement{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isasymlambda}n\ z{\isachardot}{\kern0pt}\ n{\isasymin}nat\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}funspace{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}n{\isacharcomma}{\kern0pt}B{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}B{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}B{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}replacement{\isacharbrackleft}{\kern0pt}of\ B{\isacharbrackright}{\kern0pt}\ RepFun{\isacharunderscore}{\kern0pt}closed{\isadigit{2}}\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
\isanewline
\isanewline
\isacommand{sublocale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ {\isasymsubseteq}\ M{\isacharunderscore}{\kern0pt}seqspace\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}unfold{\isacharunderscore}{\kern0pt}locales{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ B\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{3}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}omega{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}typed{\isacharunderscore}{\kern0pt}function{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}B{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ B{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ replacement{\isacharunderscore}{\kern0pt}ax{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ {\isacartoucheopen}B{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}univalent{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ M{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ B{\isacharbrackright}{\kern0pt}\ {\isasymTurnstile}\ seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isakeyword{if}\ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\ \isakeyword{for}\ A\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{unfolding}\isamarkupfalse%
\ univalent{\isacharunderscore}{\kern0pt}def\ seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ blast\ dest{\isacharcolon}{\kern0pt}transitivity{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}n\ z{\isachardot}{\kern0pt}\ {\isasymexists}om{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ omega{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}om{\isacharparenright}{\kern0pt}\ {\isasymand}\ n\ {\isasymin}\ om\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}funspace{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ B{\isacharcomma}{\kern0pt}\ z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharbrackleft}{\kern0pt}of\ {\isadigit{0}}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isacharcomma}{\kern0pt}B{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharunderscore}{\kern0pt}\ {\isadigit{1}}\ {\isacharunderscore}{\kern0pt}\ {\isadigit{2}}\ B\ M{\isacharbrackright}{\kern0pt}\ \isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}rep{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}B{\isasymin}M{\isacartoucheclose}\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}n\ z{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}funspace{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ n{\isacharcomma}{\kern0pt}\ B{\isacharcomma}{\kern0pt}\ z{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymlambda}\ j\ {\isasymin}\ succ{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ if\ j\ {\isacharless}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ then\ f{\isacharbackquote}{\kern0pt}j\ else\ a{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isasymin}n{\isasymrightarrow}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}A{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isasymin}\ succ{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ A{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ equ{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fun\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ j\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}j{\isasymin}succ{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ equ\ {\isacartoucheopen}n{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}j{\isasymle}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltI\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}n{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{consider}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}lt{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}j{\isacharless}{\kern0pt}n{\isachardoublequoteclose}\ {\isacharbar}{\kern0pt}\ {\isacharparenleft}{\kern0pt}eq{\isacharparenright}{\kern0pt}\ {\isachardoublequoteopen}j{\isacharequal}{\kern0pt}n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ leD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}if\ j\ {\isacharless}{\kern0pt}\ n\ then\ f{\isacharbackquote}{\kern0pt}j\ else\ a{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ cases\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ lt\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}f{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}type\ ltD{\isacharbrackleft}{\kern0pt}OF\ lt{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ eq\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ equ\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ lam{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}type\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f{\isasymin}A{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}A{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymin}\ A{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}f{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isasymin}y{\isasymrightarrow}A{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}A{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isasymin}succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isasymrightarrow}A{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}y{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}domain\ {\isacharbrackleft}{\kern0pt}simp{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f{\isacharcolon}{\kern0pt}n{\isasymrightarrow}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}n\ {\isacharequal}{\kern0pt}\ a{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fun\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}seqspace\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ A{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ seqleR\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}seqleR{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ g\ {\isasymsubseteq}\ f{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ seqlerel\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}seqlerel{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Rrel{\isacharparenleft}{\kern0pt}{\isasymlambda}x\ y{\isachardot}{\kern0pt}\ y\ {\isasymsubseteq}\ x{\isacharcomma}{\kern0pt}A{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ seqle\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}seqle\ {\isasymequiv}\ seqlerel{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seqleI{\isacharbrackleft}{\kern0pt}intro{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymlangle}f{\isacharcomma}{\kern0pt}g{\isasymrangle}\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isasymtimes}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}\ {\isasymLongrightarrow}\ g\ {\isasymsubseteq}\ f\ \ {\isasymLongrightarrow}\ {\isasymlangle}f{\isacharcomma}{\kern0pt}g{\isasymrangle}\ {\isasymin}\ seqle{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ \ seqspace{\isacharunderscore}{\kern0pt}def\ seqle{\isacharunderscore}{\kern0pt}def\ seqlerel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seqleD{\isacharbrackleft}{\kern0pt}dest{\isacharbang}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}z\ {\isasymin}\ seqle\ {\isasymLongrightarrow}\ {\isasymexists}x\ y{\isachardot}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isasymtimes}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}\ {\isasymand}\ y\ {\isasymsubseteq}\ x\ {\isasymand}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqle{\isacharunderscore}{\kern0pt}def\ seqlerel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ upd{\isacharunderscore}{\kern0pt}leI\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f{\isasymin}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isasymin}{\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymlangle}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}f{\isasymrangle}{\isasymin}seqle{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isasymlangle}{\isacharquery}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isacharunderscore}{\kern0pt}{\isasymrangle}{\isasymin}{\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isasymlangle}{\isacharquery}{\kern0pt}f{\isacharcomma}{\kern0pt}\ f{\isasymrangle}\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}\ {\isasymtimes}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ \ {\isachardoublequoteopen}f\ {\isasymsubseteq}\ seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ f{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}f\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ n\ \isakeyword{where}\ \ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f\ {\isacharcolon}{\kern0pt}\ n\ {\isasymrightarrow}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \ \ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ {\isachardoublequoteopen}y{\isasymin}n{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharequal}{\kern0pt}{\isasymlangle}y{\isacharcomma}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}memberD{\isacharbrackleft}{\kern0pt}of\ f\ n\ {\isachardoublequoteopen}{\isasymlambda}{\isacharunderscore}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isadigit{2}}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}f{\isacharcolon}{\kern0pt}n{\isasymrightarrow}{\isadigit{2}}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}fun\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}def\ lam{\isacharunderscore}{\kern0pt}def\ \ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}auto\ intro{\isacharcolon}{\kern0pt}ltI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}seqle{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}preorder{\isacharunderscore}{\kern0pt}on{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isacharcomma}{\kern0pt}seqle{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ refl{\isacharunderscore}{\kern0pt}def\ trans{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ blast%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}seqle{\isacharunderscore}{\kern0pt}max{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isasymin}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}\ {\isasymLongrightarrow}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isasymrangle}\ {\isasymin}\ seqle{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}seqspace\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{interpretation}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}notion\ {\isachardoublequoteopen}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ {\isachardoublequoteopen}seqle{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ preorder{\isacharunderscore}{\kern0pt}on{\isacharunderscore}{\kern0pt}seqle\ zero{\isacharunderscore}{\kern0pt}seqle{\isacharunderscore}{\kern0pt}max\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}seqspace\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ unfold{\isacharunderscore}{\kern0pt}locales\ simp{\isacharunderscore}{\kern0pt}all%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ SEQle\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}\isakeyword{infixl}\ {\isachardoublequoteopen}{\isasympreceq}s{\isachardoublequoteclose}\ {\isadigit{5}}{\isadigit{0}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}x\ {\isasympreceq}s\ y\ {\isasymequiv}\ Leq{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ SEQIncompatible\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \ {\isacharparenleft}{\kern0pt}\isakeyword{infixl}\ {\isachardoublequoteopen}{\isasymbottom}s{\isachardoublequoteclose}\ {\isadigit{5}}{\isadigit{0}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}x\ {\isasymbottom}s\ y\ {\isasymequiv}\ Incompatible{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}separative{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}f{\isasymin}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymbottom}s\ seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f\ {\isasymbottom}s\ {\isacharquery}{\kern0pt}g{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}compat{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharcomma}{\kern0pt}\ {\isacharquery}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ h\ \isakeyword{where}\ {\isachardoublequoteopen}h\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f\ {\isasymsubseteq}\ h{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}g\ {\isasymsubseteq}\ h{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}f{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ {\isachardoublequoteopen}y{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}f{\isacharcolon}{\kern0pt}y{\isasymrightarrow}{\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}f{\isacharcolon}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isadigit{2}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}g{\isacharcolon}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}type\ \isacommand{by}\isamarkupfalse%
\ blast{\isacharplus}{\kern0pt}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}f{\isacharbackquote}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ {\isacharquery}{\kern0pt}f{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}g{\isacharbackquote}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ {\isacharquery}{\kern0pt}g{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}Pair\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isasymrangle}\ {\isasymin}\ h{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}y{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isasymrangle}\ {\isasymin}\ h{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}h\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isacartoucheclose}\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ n\ \isakeyword{where}\ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}h{\isacharcolon}{\kern0pt}n{\isasymrightarrow}{\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}False{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ fun{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}function{\isacharbrackleft}{\kern0pt}of\ h\ n\ {\isachardoublequoteopen}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ {\isadigit{2}}{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ seqspace{\isacharunderscore}{\kern0pt}def\ function{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}seqleR\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}seqleR{\isacharparenleft}{\kern0pt}Q{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ g\ {\isasymsubseteq}\ f{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ seqleR{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}seqleR{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}fg{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Exists{\isacharparenleft}{\kern0pt}Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}fg{\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ type{\isacharunderscore}{\kern0pt}seqleR{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}fg\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ seqleR{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}fg{\isacharparenright}{\kern0pt}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqleR{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}seqleR{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}fg\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ arity{\isacharparenleft}{\kern0pt}seqleR{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}fg{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}fg{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqleR{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm\ arity{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}basic{\isacharparenright}{\kern0pt}\ seqleR{\isacharunderscore}{\kern0pt}abs{\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}seqleR{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}seqleR{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqleR{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}seqleR{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ apply{\isacharunderscore}{\kern0pt}abs\ domain{\isacharunderscore}{\kern0pt}abs\ domain{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\ \ domain{\isacharunderscore}{\kern0pt}closed{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ relP\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}relP{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}r{\isacharcomma}{\kern0pt}xy{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}y{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ pair{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharcomma}{\kern0pt}xy{\isacharparenright}{\kern0pt}\ {\isasymand}\ r{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}ctm{\isacharparenright}{\kern0pt}\ seqleR{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}fg{\isasymin}nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}env{\isasymin}list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}seqleR{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}fg{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ relP{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}seqleR{\isacharcomma}{\kern0pt}nth{\isacharparenleft}{\kern0pt}fg{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ seqleR{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}seqleR{\isacharunderscore}{\kern0pt}def\ relP{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ trans{\isacharunderscore}{\kern0pt}M\ sats{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}fm\ pair{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}sats\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}basic{\isacharparenright}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}related{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}\ f\ g\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ rel{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}rel{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymAnd}z\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ relP{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}rel{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymand}\ rel{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ relP{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ assms\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}RRel\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}RRel{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}r{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymexists}A{\isadigit{2}}{\isacharbrackleft}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ cartprod{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}A{\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}Collect{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isadigit{2}}{\isacharcomma}{\kern0pt}\ relP{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}basic{\isacharparenright}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ f\ g\ {\isachardot}{\kern0pt}\ M{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ M{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ rel{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}rel{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}RRel{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}rel{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \ r\ {\isacharequal}{\kern0pt}\ Rrel{\isacharparenleft}{\kern0pt}rel{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}z{\isasymin}A{\isasymtimes}A{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\ transM{\isacharbrackleft}{\kern0pt}of\ z\ {\isachardoublequoteopen}A{\isasymtimes}A{\isachardoublequoteclose}{\isacharbrackright}{\kern0pt}\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}relP{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}rel{\isacharcomma}{\kern0pt}\ z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymand}\ rel{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}z{\isasymin}A{\isasymtimes}A{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ that\ is{\isacharunderscore}{\kern0pt}related{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}of\ rel\ is{\isacharunderscore}{\kern0pt}rel{\isacharcomma}{\kern0pt}OF\ assms{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isasymtimes}A{\isacharcomma}{\kern0pt}relP{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}rel{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Collect{\isacharparenleft}{\kern0pt}A{\isasymtimes}A{\isacharcomma}{\kern0pt}{\isasymlambda}z{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymand}\ rel{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}cong{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}A{\isasymtimes}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}A{\isasymtimes}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}relP{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}rel{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ A{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharbrackright}{\kern0pt}\ assms{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}\ assms{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}RRel{\isacharunderscore}{\kern0pt}def\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ is{\isacharunderscore}{\kern0pt}seqlerel\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}seqlerel{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}RRel{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}seqleR{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}\isakeyword{in}\ M{\isacharunderscore}{\kern0pt}basic{\isacharparenright}{\kern0pt}\ seqlerel{\isacharunderscore}{\kern0pt}abs\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}M{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}seqlerel{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}A{\isacharcomma}{\kern0pt}r{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ r\ {\isacharequal}{\kern0pt}\ Rrel{\isacharparenleft}{\kern0pt}seqleR{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}seqlerel{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}Rrel{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}OF\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}M{\isacharparenleft}{\kern0pt}r{\isacharparenright}{\kern0pt}{\isacartoucheclose}{\isacharcomma}{\kern0pt}of\ seqleR\ is{\isacharunderscore}{\kern0pt}seqleR{\isacharbrackright}{\kern0pt}\ seqleR{\isacharunderscore}{\kern0pt}abs\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ RrelP\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isasymRightarrow}i{\isasymRightarrow}o{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}RrelP{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}z{\isasymin}A{\isasymtimes}A{\isachardot}{\kern0pt}\ {\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymand}\ R{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}RrelP{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Rrel{\isacharparenleft}{\kern0pt}R{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ RrelP{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}closed{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A{\isasymin}M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ a{\isachardot}{\kern0pt}\ a\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ rel{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isasymin}formula{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ f\ g\ {\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ rel{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}rel{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}f{\isacharcomma}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}rel{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymAnd}\ a\ {\isachardot}{\kern0pt}\ a\ {\isasymin}\ M\ {\isasymLongrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}rel{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ relP{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}rel{\isacharcomma}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}Rrel{\isacharparenleft}{\kern0pt}rel{\isacharcomma}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}z{\isasymin}\ M\ {\isasymLongrightarrow}\ relP{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}rel{\isacharcomma}{\kern0pt}\ z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymand}\ rel{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ z\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharparenright}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}related{\isacharunderscore}{\kern0pt}abs{\isacharbrackleft}{\kern0pt}of\ rel\ is{\isacharunderscore}{\kern0pt}rel{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Collect{\isacharparenleft}{\kern0pt}A{\isasymtimes}A{\isacharcomma}{\kern0pt}{\isasymlambda}z{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymand}\ rel{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}{\isadigit{0}}p{\isacharbrackleft}{\kern0pt}of\ {\isachardoublequoteopen}rel{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}\ A\ z\ {\isachardot}{\kern0pt}\ relP{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}is{\isacharunderscore}{\kern0pt}rel{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlambda}\ z{\isachardot}{\kern0pt}{\isasymexists}x\ y{\isachardot}{\kern0pt}\ z\ {\isacharequal}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymand}\ rel{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ cartprod{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ seqle{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}seqle\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Rrel{\isacharunderscore}{\kern0pt}closed\ seqspace{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \ \ transitivity{\isacharbrackleft}{\kern0pt}OF\ {\isacharunderscore}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharbrackright}{\kern0pt}\ type{\isacharunderscore}{\kern0pt}seqleR{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ {\isadigit{0}}{\isacharbrackright}{\kern0pt}\ arity{\isacharunderscore}{\kern0pt}seqleR{\isacharunderscore}{\kern0pt}fm{\isacharbrackleft}{\kern0pt}of\ {\isadigit{0}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ seqleR{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}sats{\isacharbrackleft}{\kern0pt}of\ {\isadigit{0}}{\isacharbrackright}{\kern0pt}\ seqleR{\isacharunderscore}{\kern0pt}abs\ seqlerel{\isacharunderscore}{\kern0pt}abs\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ seqle{\isacharunderscore}{\kern0pt}def\ seqlerel{\isacharunderscore}{\kern0pt}def\ seqleR{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Cohen extension is proper%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
\isacommand{interpretation}\isamarkupfalse%
\ ctm{\isacharunderscore}{\kern0pt}separative\ {\isachardoublequoteopen}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ seqle\ {\isadigit{0}}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}unfold{\isacharunderscore}{\kern0pt}locales{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ f\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isacharquery}{\kern0pt}r{\isacharequal}{\kern0pt}{\isachardoublequoteopen}seq{\isacharunderscore}{\kern0pt}upd{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}q\ {\isasympreceq}s\ f\ {\isasymand}\ {\isacharquery}{\kern0pt}r\ {\isasympreceq}s\ f\ {\isasymand}\ {\isacharquery}{\kern0pt}q\ {\isasymbottom}s\ {\isacharquery}{\kern0pt}r{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ upd{\isacharunderscore}{\kern0pt}leI\ seqspace{\isacharunderscore}{\kern0pt}separative\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ calculation\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}q\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}r\ {\isasymin}\ {\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ seq{\isacharunderscore}{\kern0pt}upd{\isacharunderscore}{\kern0pt}type{\isacharbrackleft}{\kern0pt}of\ f\ {\isadigit{2}}{\isacharbrackright}{\kern0pt}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}q{\isasymin}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardot}{\kern0pt}\ {\isasymexists}r{\isasymin}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}{\isachardot}{\kern0pt}\ q\ {\isasympreceq}s\ f\ {\isasymand}\ r\ {\isasympreceq}s\ f\ {\isasymand}\ q\ {\isasymbottom}s\ r{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ bexI{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\ %
\isamarkupcmt{why the heck auto-tools don't solve this?%
}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{2}}{\isacharcircum}{\kern0pt}{\isacharless}{\kern0pt}{\isasymomega}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}M\ seqspace{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}seqle\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ seqle{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{{\isachardot}{\kern0pt}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ cohen{\isacharunderscore}{\kern0pt}extension{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}proper{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymexists}G{\isachardot}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ M\ {\isasymnoteq}\ GenExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ proper{\isacharunderscore}{\kern0pt}extension\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}existence\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}seqspace\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Succession_Poset.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%31=5%:%
%:%32=6%:%
%:%46=8%:%
%:%58=10%:%
%:%59=11%:%
%:%61=13%:%
%:%62=13%:%
%:%63=14%:%
%:%64=15%:%
%:%65=16%:%
%:%66=17%:%
%:%67=17%:%
%:%70=18%:%
%:%74=18%:%
%:%75=18%:%
%:%76=18%:%
%:%81=18%:%
%:%84=19%:%
%:%85=20%:%
%:%86=20%:%
%:%89=21%:%
%:%93=21%:%
%:%94=21%:%
%:%95=21%:%
%:%100=21%:%
%:%103=22%:%
%:%104=23%:%
%:%105=23%:%
%:%106=24%:%
%:%109=25%:%
%:%113=25%:%
%:%114=25%:%
%:%115=25%:%
%:%120=25%:%
%:%123=26%:%
%:%124=27%:%
%:%125=27%:%
%:%126=28%:%
%:%127=29%:%
%:%128=30%:%
%:%129=31%:%
%:%130=32%:%
%:%133=33%:%
%:%137=33%:%
%:%138=33%:%
%:%139=34%:%
%:%140=34%:%
%:%145=34%:%
%:%150=35%:%
%:%155=36%:%
%:%156=36%:%
%:%161=36%:%
%:%164=37%:%
%:%165=38%:%
%:%166=38%:%
%:%167=39%:%
%:%168=40%:%
%:%169=41%:%
%:%170=42%:%
%:%171=43%:%
%:%172=43%:%
%:%173=44%:%
%:%176=45%:%
%:%180=45%:%
%:%181=45%:%
%:%182=45%:%
%:%183=46%:%
%:%184=46%:%
%:%189=46%:%
%:%192=47%:%
%:%193=48%:%
%:%194=48%:%
%:%195=49%:%
%:%196=50%:%
%:%197=51%:%
%:%198=51%:%
%:%205=52%:%
%:%206=52%:%
%:%207=53%:%
%:%208=53%:%
%:%209=54%:%
%:%210=54%:%
%:%211=55%:%
%:%212=55%:%
%:%213=56%:%
%:%214=56%:%
%:%215=57%:%
%:%216=57%:%
%:%217=58%:%
%:%218=58%:%
%:%219=59%:%
%:%220=59%:%
%:%221=60%:%
%:%222=60%:%
%:%223=61%:%
%:%224=61%:%
%:%225=62%:%
%:%226=62%:%
%:%227=63%:%
%:%228=63%:%
%:%229=64%:%
%:%230=64%:%
%:%231=65%:%
%:%232=65%:%
%:%233=66%:%
%:%234=66%:%
%:%235=66%:%
%:%236=67%:%
%:%237=67%:%
%:%238=68%:%
%:%239=69%:%
%:%240=69%:%
%:%241=69%:%
%:%242=70%:%
%:%243=70%:%
%:%244=71%:%
%:%245=71%:%
%:%246=72%:%
%:%247=72%:%
%:%248=73%:%
%:%249=73%:%
%:%250=73%:%
%:%251=74%:%
%:%252=74%:%
%:%253=75%:%
%:%254=75%:%
%:%255=76%:%
%:%256=76%:%
%:%257=77%:%
%:%258=77%:%
%:%259=77%:%
%:%260=78%:%
%:%266=78%:%
%:%269=79%:%
%:%270=80%:%
%:%271=80%:%
%:%272=81%:%
%:%273=82%:%
%:%274=83%:%
%:%275=83%:%
%:%276=84%:%
%:%277=85%:%
%:%284=86%:%
%:%285=86%:%
%:%286=87%:%
%:%287=87%:%
%:%288=88%:%
%:%289=88%:%
%:%290=88%:%
%:%291=88%:%
%:%292=89%:%
%:%293=89%:%
%:%294=90%:%
%:%295=90%:%
%:%296=91%:%
%:%297=91%:%
%:%298=92%:%
%:%299=92%:%
%:%300=93%:%
%:%301=93%:%
%:%302=93%:%
%:%303=93%:%
%:%304=94%:%
%:%305=94%:%
%:%306=95%:%
%:%307=95%:%
%:%308=95%:%
%:%309=95%:%
%:%310=96%:%
%:%311=96%:%
%:%312=97%:%
%:%313=97%:%
%:%314=98%:%
%:%315=98%:%
%:%316=99%:%
%:%317=99%:%
%:%318=100%:%
%:%319=100%:%
%:%320=101%:%
%:%321=101%:%
%:%322=101%:%
%:%323=101%:%
%:%324=102%:%
%:%325=102%:%
%:%326=103%:%
%:%327=103%:%
%:%328=104%:%
%:%329=104%:%
%:%330=105%:%
%:%331=105%:%
%:%332=105%:%
%:%333=106%:%
%:%334=106%:%
%:%335=107%:%
%:%336=107%:%
%:%337=108%:%
%:%338=108%:%
%:%339=109%:%
%:%340=109%:%
%:%341=110%:%
%:%342=110%:%
%:%343=111%:%
%:%344=111%:%
%:%345=112%:%
%:%346=112%:%
%:%347=113%:%
%:%353=113%:%
%:%356=114%:%
%:%357=115%:%
%:%358=115%:%
%:%359=116%:%
%:%360=117%:%
%:%367=118%:%
%:%368=118%:%
%:%369=119%:%
%:%370=119%:%
%:%371=120%:%
%:%372=120%:%
%:%373=121%:%
%:%374=121%:%
%:%375=121%:%
%:%376=122%:%
%:%377=122%:%
%:%378=123%:%
%:%379=123%:%
%:%380=124%:%
%:%381=124%:%
%:%382=124%:%
%:%383=125%:%
%:%384=125%:%
%:%385=126%:%
%:%386=126%:%
%:%387=127%:%
%:%388=127%:%
%:%389=127%:%
%:%390=128%:%
%:%396=128%:%
%:%399=129%:%
%:%400=130%:%
%:%401=130%:%
%:%402=131%:%
%:%403=132%:%
%:%406=133%:%
%:%410=133%:%
%:%411=133%:%
%:%412=133%:%
%:%413=133%:%
%:%418=133%:%
%:%421=134%:%
%:%422=135%:%
%:%423=135%:%
%:%424=136%:%
%:%427=137%:%
%:%431=137%:%
%:%432=137%:%
%:%433=138%:%
%:%434=138%:%
%:%439=138%:%
%:%442=139%:%
%:%443=140%:%
%:%444=140%:%
%:%445=141%:%
%:%446=142%:%
%:%447=143%:%
%:%448=144%:%
%:%449=144%:%
%:%450=145%:%
%:%451=146%:%
%:%452=147%:%
%:%453=148%:%
%:%454=148%:%
%:%455=149%:%
%:%456=150%:%
%:%457=151%:%
%:%458=152%:%
%:%459=152%:%
%:%460=153%:%
%:%463=154%:%
%:%467=154%:%
%:%468=154%:%
%:%469=155%:%
%:%470=155%:%
%:%475=155%:%
%:%478=156%:%
%:%479=157%:%
%:%480=157%:%
%:%481=158%:%
%:%484=159%:%
%:%488=159%:%
%:%489=159%:%
%:%490=160%:%
%:%491=160%:%
%:%496=160%:%
%:%499=161%:%
%:%500=162%:%
%:%501=162%:%
%:%502=163%:%
%:%503=164%:%
%:%510=165%:%
%:%511=165%:%
%:%512=166%:%
%:%513=166%:%
%:%514=167%:%
%:%515=167%:%
%:%516=167%:%
%:%517=168%:%
%:%518=168%:%
%:%519=169%:%
%:%520=169%:%
%:%521=170%:%
%:%522=170%:%
%:%523=171%:%
%:%524=171%:%
%:%525=172%:%
%:%526=172%:%
%:%527=173%:%
%:%528=173%:%
%:%529=173%:%
%:%530=174%:%
%:%531=174%:%
%:%532=175%:%
%:%533=175%:%
%:%534=175%:%
%:%535=176%:%
%:%536=176%:%
%:%537=176%:%
%:%538=177%:%
%:%539=177%:%
%:%540=177%:%
%:%541=178%:%
%:%542=178%:%
%:%543=179%:%
%:%544=179%:%
%:%545=179%:%
%:%546=180%:%
%:%547=180%:%
%:%548=180%:%
%:%549=180%:%
%:%550=181%:%
%:%551=181%:%
%:%552=182%:%
%:%553=182%:%
%:%554=183%:%
%:%555=183%:%
%:%556=184%:%
%:%557=184%:%
%:%558=185%:%
%:%559=185%:%
%:%560=186%:%
%:%566=186%:%
%:%569=187%:%
%:%570=188%:%
%:%571=188%:%
%:%574=189%:%
%:%578=189%:%
%:%579=189%:%
%:%580=189%:%
%:%585=189%:%
%:%588=190%:%
%:%589=191%:%
%:%590=191%:%
%:%593=192%:%
%:%597=192%:%
%:%598=192%:%
%:%599=193%:%
%:%600=193%:%
%:%605=193%:%
%:%608=194%:%
%:%609=195%:%
%:%610=195%:%
%:%613=196%:%
%:%617=196%:%
%:%618=196%:%
%:%619=197%:%
%:%620=197%:%
%:%625=197%:%
%:%628=198%:%
%:%629=199%:%
%:%630=199%:%
%:%631=200%:%
%:%632=201%:%
%:%633=202%:%
%:%634=202%:%
%:%635=203%:%
%:%636=204%:%
%:%637=205%:%
%:%638=205%:%
%:%639=206%:%
%:%640=207%:%
%:%647=208%:%
%:%648=208%:%
%:%649=209%:%
%:%650=209%:%
%:%651=210%:%
%:%652=210%:%
%:%653=211%:%
%:%654=211%:%
%:%655=212%:%
%:%656=212%:%
%:%657=213%:%
%:%658=213%:%
%:%659=213%:%
%:%660=214%:%
%:%661=214%:%
%:%662=214%:%
%:%663=215%:%
%:%664=215%:%
%:%665=215%:%
%:%666=216%:%
%:%667=216%:%
%:%668=217%:%
%:%669=217%:%
%:%670=217%:%
%:%671=218%:%
%:%672=218%:%
%:%673=218%:%
%:%674=219%:%
%:%675=219%:%
%:%676=219%:%
%:%677=219%:%
%:%678=220%:%
%:%679=220%:%
%:%680=221%:%
%:%681=221%:%
%:%682=221%:%
%:%683=222%:%
%:%684=222%:%
%:%685=222%:%
%:%686=223%:%
%:%687=223%:%
%:%688=223%:%
%:%689=224%:%
%:%690=224%:%
%:%691=225%:%
%:%692=225%:%
%:%693=226%:%
%:%694=226%:%
%:%695=227%:%
%:%696=227%:%
%:%697=227%:%
%:%698=228%:%
%:%704=228%:%
%:%707=229%:%
%:%708=230%:%
%:%709=230%:%
%:%710=231%:%
%:%711=232%:%
%:%712=233%:%
%:%713=233%:%
%:%714=234%:%
%:%715=235%:%
%:%716=236%:%
%:%717=236%:%
%:%718=237%:%
%:%721=238%:%
%:%725=238%:%
%:%726=238%:%
%:%727=239%:%
%:%728=239%:%
%:%733=239%:%
%:%736=240%:%
%:%737=241%:%
%:%738=241%:%
%:%739=242%:%
%:%742=243%:%
%:%746=243%:%
%:%747=243%:%
%:%748=244%:%
%:%749=244%:%
%:%750=244%:%
%:%755=244%:%
%:%758=245%:%
%:%759=246%:%
%:%760=246%:%
%:%761=247%:%
%:%762=248%:%
%:%765=249%:%
%:%769=249%:%
%:%770=249%:%
%:%771=250%:%
%:%772=250%:%
%:%773=251%:%
%:%774=251%:%
%:%779=251%:%
%:%782=252%:%
%:%783=253%:%
%:%784=253%:%
%:%785=254%:%
%:%786=255%:%
%:%787=256%:%
%:%788=257%:%
%:%789=257%:%
%:%790=258%:%
%:%791=259%:%
%:%794=260%:%
%:%798=260%:%
%:%799=260%:%
%:%800=261%:%
%:%801=261%:%
%:%802=262%:%
%:%803=262%:%
%:%808=262%:%
%:%811=263%:%
%:%812=264%:%
%:%813=265%:%
%:%814=265%:%
%:%815=266%:%
%:%816=267%:%
%:%819=268%:%
%:%823=268%:%
%:%824=268%:%
%:%825=268%:%
%:%826=268%:%
%:%831=268%:%
%:%834=269%:%
%:%835=270%:%
%:%836=270%:%
%:%837=271%:%
%:%838=272%:%
%:%839=273%:%
%:%840=274%:%
%:%841=274%:%
%:%842=275%:%
%:%843=276%:%
%:%844=277%:%
%:%851=278%:%
%:%852=278%:%
%:%853=279%:%
%:%854=279%:%
%:%855=280%:%
%:%856=280%:%
%:%857=281%:%
%:%858=281%:%
%:%859=281%:%
%:%860=282%:%
%:%861=282%:%
%:%862=283%:%
%:%863=283%:%
%:%864=284%:%
%:%865=284%:%
%:%866=284%:%
%:%867=285%:%
%:%868=285%:%
%:%869=286%:%
%:%870=286%:%
%:%871=287%:%
%:%872=287%:%
%:%873=288%:%
%:%874=288%:%
%:%875=289%:%
%:%876=289%:%
%:%877=290%:%
%:%878=290%:%
%:%879=290%:%
%:%880=290%:%
%:%881=291%:%
%:%882=291%:%
%:%883=292%:%
%:%889=292%:%
%:%892=293%:%
%:%893=294%:%
%:%894=294%:%
%:%895=295%:%
%:%896=296%:%
%:%897=297%:%
%:%898=298%:%
%:%899=298%:%
%:%900=299%:%
%:%901=300%:%
%:%904=301%:%
%:%908=301%:%
%:%909=301%:%
%:%910=302%:%
%:%911=302%:%
%:%912=303%:%
%:%913=303%:%
%:%918=303%:%
%:%921=304%:%
%:%922=305%:%
%:%923=305%:%
%:%924=306%:%
%:%925=307%:%
%:%926=308%:%
%:%927=308%:%
%:%930=309%:%
%:%934=309%:%
%:%935=309%:%
%:%936=309%:%
%:%941=309%:%
%:%944=310%:%
%:%945=311%:%
%:%946=311%:%
%:%947=312%:%
%:%948=313%:%
%:%949=314%:%
%:%950=314%:%
%:%951=315%:%
%:%952=316%:%
%:%953=317%:%
%:%954=318%:%
%:%955=319%:%
%:%956=320%:%
%:%963=321%:%
%:%964=321%:%
%:%965=322%:%
%:%966=322%:%
%:%967=323%:%
%:%968=323%:%
%:%969=324%:%
%:%970=324%:%
%:%971=325%:%
%:%972=325%:%
%:%973=326%:%
%:%974=326%:%
%:%975=327%:%
%:%976=327%:%
%:%977=328%:%
%:%978=329%:%
%:%979=329%:%
%:%980=330%:%
%:%981=330%:%
%:%982=330%:%
%:%983=331%:%
%:%984=331%:%
%:%985=331%:%
%:%986=332%:%
%:%992=332%:%
%:%995=333%:%
%:%996=334%:%
%:%997=334%:%
%:%1000=335%:%
%:%1004=335%:%
%:%1005=335%:%
%:%1006=336%:%
%:%1007=337%:%
%:%1008=338%:%
%:%1009=338%:%
%:%1010=339%:%
%:%1011=339%:%
%:%1025=341%:%
%:%1035=343%:%
%:%1036=343%:%
%:%1043=344%:%
%:%1044=344%:%
%:%1045=345%:%
%:%1046=345%:%
%:%1047=346%:%
%:%1048=346%:%
%:%1049=347%:%
%:%1050=347%:%
%:%1051=348%:%
%:%1052=348%:%
%:%1053=349%:%
%:%1054=349%:%
%:%1055=350%:%
%:%1056=350%:%
%:%1057=350%:%
%:%1058=351%:%
%:%1059=351%:%
%:%1060=351%:%
%:%1061=352%:%
%:%1062=352%:%
%:%1063=353%:%
%:%1064=353%:%
%:%1065=353%:%
%:%1066=354%:%
%:%1067=354%:%
%:%1068=355%:%
%:%1069=355%:%
%:%1070=356%:%
%:%1071=356%:%
%:%1072=356%:%
%:%1073=356%:%
%:%1074=357%:%
%:%1075=357%:%
%:%1076=358%:%
%:%1077=358%:%
%:%1078=358%:%
%:%1079=358%:%
%:%1080=359%:%
%:%1081=359%:%
%:%1082=360%:%
%:%1083=360%:%
%:%1084=360%:%
%:%1085=360%:%
%:%1086=361%:%
%:%1092=361%:%
%:%1095=362%:%
%:%1096=363%:%
%:%1097=363%:%
%:%1100=364%:%
%:%1104=364%:%
%:%1105=364%:%
%:%1106=365%:%
%:%1107=365%:%
%:%1112=365%:%
%:%1115=366%:%
%:%1116=367%:%
%:%1117=367%:%
%:%1120=368%:%
%:%1125=369%:%