### 2024-11-01 code/output/document/SymExt_Definition.tex

%
\begin{isabellebody}%
\setisabellecontext{SymExt{\isacharunderscore}{\kern0pt}Definition}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}Definition\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\ \isanewline
\ \ \ \ HS{\isacharunderscore}{\kern0pt}Theorems\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\ \isanewline
\isakeyword{begin}\ \isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ SymExt\ \isakeyword{where}\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ HS\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
\isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic\ {\isacharequal}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\ {\isacharplus}{\kern0pt}\ G{\isacharunderscore}{\kern0pt}generic\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}SymExt\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M\ {\isasymsubseteq}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ check{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}HS\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ check{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ valcheck\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ assms\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ generic\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}GenExt\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ GenExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ GenExt{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}SymExt\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}\ Transset{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}def\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ ballI{\isacharsemicolon}{\kern0pt}\ rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ y\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x{\isacharprime}{\kern0pt}\ \isakeyword{where}\ x{\isacharprime}{\kern0pt}H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ x{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isacharprime}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y{\isacharprime}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val\ \isacommand{by}\isamarkupfalse%
\ blast\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y{\isacharprime}{\kern0pt}\ \isakeyword{where}\ y{\isacharprime}{\kern0pt}H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y{\isacharprime}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ y{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ y{\isacharprime}{\kern0pt}inHS\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y{\isacharprime}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ x{\isacharprime}{\kern0pt}H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ y{\isacharprime}{\kern0pt}H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nonempty{\isacharunderscore}{\kern0pt}SymExt\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isadigit{0}}\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ M{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}SymExt\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ex{\isacharunderscore}{\kern0pt}HS{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}list\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ l\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}l\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}l{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ l{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ Nil\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}Nil\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Cons\ h\ t{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}h\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}t{\isacharprime}{\kern0pt}{\isasymin}list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ map{\isacharparenleft}{\kern0pt}{\isasymlambda}a{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ t{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ t{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ t{\isacharprime}{\kern0pt}\ \isakeyword{where}\ t{\isacharprime}{\kern0pt}H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}t{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}HS{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}map{\isacharparenleft}{\kern0pt}{\isasymlambda}a{\isachardot}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ a{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ t{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ t{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ h{\isacharprime}{\kern0pt}\ \isakeyword{where}\ h{\isacharprime}{\kern0pt}H\ {\isacharcolon}{\kern0pt}{\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ h{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ h{\isachardoublequoteclose}\ {\isachardoublequoteopen}h{\isacharprime}{\kern0pt}\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}h{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ t{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ h{\isacharprime}{\kern0pt}H\ t{\isacharprime}{\kern0pt}H\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}trans\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ x\ {\isasymLongrightarrow}\ y\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}SymExt\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ Transset{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/SymExt_Definition.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=10%:%
%:%30=11%:%
%:%31=12%:%
%:%32=12%:%
%:%33=13%:%
%:%34=14%:%
%:%35=15%:%
%:%36=15%:%
%:%37=16%:%
%:%38=17%:%
%:%39=18%:%
%:%40=18%:%
%:%47=19%:%
%:%48=19%:%
%:%49=20%:%
%:%50=20%:%
%:%51=20%:%
%:%52=21%:%
%:%53=21%:%
%:%54=21%:%
%:%55=21%:%
%:%56=21%:%
%:%57=22%:%
%:%58=22%:%
%:%59=22%:%
%:%60=22%:%
%:%61=22%:%
%:%62=23%:%
%:%63=23%:%
%:%64=23%:%
%:%65=23%:%
%:%66=23%:%
%:%67=24%:%
%:%73=24%:%
%:%76=25%:%
%:%77=26%:%
%:%78=26%:%
%:%81=27%:%
%:%85=27%:%
%:%86=27%:%
%:%87=28%:%
%:%88=28%:%
%:%89=29%:%
%:%90=29%:%
%:%91=30%:%
%:%92=30%:%
%:%97=30%:%
%:%100=31%:%
%:%101=32%:%
%:%102=32%:%
%:%105=33%:%
%:%109=33%:%
%:%110=33%:%
%:%111=34%:%
%:%112=34%:%
%:%113=35%:%
%:%114=35%:%
%:%115=35%:%
%:%116=36%:%
%:%117=36%:%
%:%118=36%:%
%:%119=36%:%
%:%120=36%:%
%:%121=37%:%
%:%122=37%:%
%:%123=37%:%
%:%124=37%:%
%:%125=37%:%
%:%126=38%:%
%:%127=38%:%
%:%128=38%:%
%:%129=38%:%
%:%130=38%:%
%:%131=39%:%
%:%132=39%:%
%:%133=39%:%
%:%134=39%:%
%:%135=39%:%
%:%136=40%:%
%:%137=40%:%
%:%138=40%:%
%:%139=40%:%
%:%140=40%:%
%:%141=41%:%
%:%147=41%:%
%:%150=42%:%
%:%151=43%:%
%:%152=43%:%
%:%155=44%:%
%:%159=44%:%
%:%160=44%:%
%:%161=45%:%
%:%162=45%:%
%:%163=46%:%
%:%164=46%:%
%:%169=46%:%
%:%172=47%:%
%:%173=48%:%
%:%174=49%:%
%:%175=49%:%
%:%176=50%:%
%:%177=51%:%
%:%178=52%:%
%:%181=53%:%
%:%182=54%:%
%:%186=54%:%
%:%187=54%:%
%:%188=55%:%
%:%189=55%:%
%:%190=56%:%
%:%191=56%:%
%:%192=57%:%
%:%193=57%:%
%:%194=57%:%
%:%195=58%:%
%:%196=58%:%
%:%197=59%:%
%:%198=59%:%
%:%199=60%:%
%:%200=60%:%
%:%201=61%:%
%:%202=61%:%
%:%203=62%:%
%:%204=62%:%
%:%205=62%:%
%:%206=62%:%
%:%207=63%:%
%:%208=63%:%
%:%209=63%:%
%:%210=63%:%
%:%211=64%:%
%:%212=64%:%
%:%213=65%:%
%:%214=65%:%
%:%215=66%:%
%:%216=66%:%
%:%217=67%:%
%:%218=67%:%
%:%219=68%:%
%:%220=68%:%
%:%221=69%:%
%:%222=69%:%
%:%223=70%:%
%:%224=70%:%
%:%225=71%:%
%:%226=71%:%
%:%227=72%:%
%:%233=72%:%
%:%236=73%:%
%:%237=74%:%
%:%238=74%:%
%:%241=75%:%
%:%245=75%:%
%:%246=75%:%
%:%247=76%:%
%:%248=76%:%
%:%249=77%:%
%:%250=77%:%
%:%255=77%:%
%:%258=78%:%
%:%259=79%:%
%:%260=79%:%
%:%267=80%:%