### 2024-11-01 code/output/document/SymExt_Separation_Base.tex

%
\begin{isabellebody}%
\setisabellecontext{SymExt{\isacharunderscore}{\kern0pt}Separation{\isacharunderscore}{\kern0pt}Base}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}Separation{\isacharunderscore}{\kern0pt}Base\isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\ \isanewline
\ \ \ \ SymExt{\isacharunderscore}{\kern0pt}Definition\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ cartprod{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}cartprod{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ B{\isacharcomma}{\kern0pt}\ Z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymforall}u\ {\isasymin}\ M{\isachardot}{\kern0pt}\ u\ {\isasymin}\ Z\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ M{\isachardot}{\kern0pt}\ y\ {\isasymin}\ B\ {\isasymand}\ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ u{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ powerset{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}powerset{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ Z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymforall}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ x\ {\isasymin}\ Z\ {\isasymlongleftrightarrow}\ subset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}singleton{\isacharprime}{\kern0pt}\ \isakeyword{where}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}singleton{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ Z{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isasymforall}a\ {\isasymin}\ M{\isachardot}{\kern0pt}\ a\ {\isasymin}\ Z\ {\isasymlongleftrightarrow}\ a\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
\isacommand{lemma}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}singleton{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ Z\ {\isasymin}\ M\ {\isasymLongrightarrow}\ is{\isacharunderscore}{\kern0pt}singleton{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ Z{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ is{\isacharunderscore}{\kern0pt}singleton{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}singleton{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharcomma}{\kern0pt}\ rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ transM\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M\ \isakeyword{where}\ {\isachardoublequoteopen}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ v\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}x{\isachardot}{\kern0pt}\ {\isasymexists}P{\isachardot}{\kern0pt}\ xP\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ P{\isachargreater}{\kern0pt}\ {\isasymand}\ v\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isacommand{definition}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond\ \isakeyword{where}\ {\isachardoublequoteopen}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}v{\isacharcomma}{\kern0pt}\ xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymequiv}\isanewline
\ \ \ {\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}P\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}domx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}Ps\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}dPs\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}gi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}gu\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}guP\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}PguP\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ xP{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ domx{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}singleton{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ Ps{\isacharparenright}{\kern0pt}\ {\isasymand}\ cartprod{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}domx{\isacharcomma}{\kern0pt}\ Ps{\isacharcomma}{\kern0pt}\ dPs{\isacharparenright}{\kern0pt}\ {\isasymand}\ image{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ dPs{\isacharcomma}{\kern0pt}\ gi{\isacharparenright}{\kern0pt}\ {\isasymand}\ big{\isacharunderscore}{\kern0pt}union{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ gi{\isacharcomma}{\kern0pt}\ gu{\isacharparenright}{\kern0pt}\ {\isasymand}\ cartprod{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}gu{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ guP{\isacharparenright}{\kern0pt}\ {\isasymand}\ powerset{\isacharprime}{\kern0pt}{\isacharparenleft}{\kern0pt}guP{\isacharcomma}{\kern0pt}\ PguP{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isasymin}\ PguP{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}xP\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ v\ {\isasymin}\ M{\isachardot}{\kern0pt}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}v{\isacharcomma}{\kern0pt}\ xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}eq{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymexists}x\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}P\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}domx\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}Ps\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}dPs\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}gi\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}gu\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}guP\ {\isasymin}\ M{\isachardot}{\kern0pt}\ {\isasymexists}PguP\ {\isasymin}\ M{\isachardot}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ pair{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ xP{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ domx{\isacharparenright}{\kern0pt}\ {\isasymand}\ is{\isacharunderscore}{\kern0pt}singleton{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ Ps{\isacharparenright}{\kern0pt}\ {\isasymand}\ cartprod{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ domx{\isacharcomma}{\kern0pt}\ Ps{\isacharcomma}{\kern0pt}\ dPs{\isacharparenright}{\kern0pt}\ {\isasymand}\ image{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ g{\isacharcomma}{\kern0pt}\ dPs{\isacharcomma}{\kern0pt}\ gi{\isacharparenright}{\kern0pt}\ {\isasymand}\ big{\isacharunderscore}{\kern0pt}union{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ gi{\isacharcomma}{\kern0pt}\ gu{\isacharparenright}{\kern0pt}\ {\isasymand}\ cartprod{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ gu{\isacharcomma}{\kern0pt}\ P{\isacharcomma}{\kern0pt}\ guP{\isacharparenright}{\kern0pt}\ {\isasymand}\ powerset{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ guP{\isacharcomma}{\kern0pt}\ PguP{\isacharparenright}{\kern0pt}\ {\isasymand}\ v\ {\isasymin}\ PguP{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ cartprod{\isacharunderscore}{\kern0pt}def\ cartprod{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ powerset{\isacharunderscore}{\kern0pt}def\ powerset{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}singleton{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ clarsimp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ domain{\isacharunderscore}{\kern0pt}closed\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ cartprod{\isacharunderscore}{\kern0pt}closed\ image{\isacharunderscore}{\kern0pt}closed\ Union{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ clarsimp\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ v\ x\ P{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}a\ {\isasymin}\ Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}g\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ a\ {\isasymin}\ M{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}g\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ M{\isacharunderscore}{\kern0pt}powerset{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ domain{\isacharunderscore}{\kern0pt}closed\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ cartprod{\isacharunderscore}{\kern0pt}closed\ image{\isacharunderscore}{\kern0pt}closed\ Union{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ clarsimp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{schematic{\isacharunderscore}{\kern0pt}goal}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ {\isachardoublequoteopen}s\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}xP\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ s{\isachardoublequoteclose}\ \ \ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ \ \ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ xP{\isachardoublequoteclose}\ \ \ \ \isanewline
\ \ \ \ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \isakeyword{shows}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}s\ {\isacharequal}{\kern0pt}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}v\ {\isasymin}\ M{\isachardot}{\kern0pt}\ v\ {\isasymin}\ s\ {\isasymlongleftrightarrow}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharparenleft}{\kern0pt}v{\isacharcomma}{\kern0pt}\ xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ iff{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharcomma}{\kern0pt}\ rule\ ballI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}eq\ assms\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}eq\ assms\ transM\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}force{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ \ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}cond{\isacharunderscore}{\kern0pt}def\ cartprod{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ powerset{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}singleton{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}def\ subset{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}insert\ assms\ {\isacharsemicolon}{\kern0pt}\ {\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}rules\ {\isacharbar}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ \isanewline
\ \ {\isachardoublequoteopen}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\ Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{8}}{\isacharcomma}{\kern0pt}\ {\isadigit{7}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}domain{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{8}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Equal{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{5}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{8}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}image{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}big{\isacharunderscore}{\kern0pt}union{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{3}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Forall\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Exists\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ And{\isacharparenleft}{\kern0pt}Forall{\isacharparenleft}{\kern0pt}Iff{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Forall{\isacharparenleft}{\kern0pt}Implies{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{3}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Member{\isacharparenleft}{\kern0pt}{\isadigit{9}}{\isacharcomma}{\kern0pt}\ {\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \ {\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}xP\ {\isasymin}\ M\ {\isasymLongrightarrow}\ g\ {\isasymin}\ M\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}{\isasymexists}x{\isachardot}{\kern0pt}\ {\isasymexists}P{\isachardot}{\kern0pt}\ xP\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ P{\isachargreater}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}xP\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ P\ \isakeyword{where}\ xPH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}xP\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ P{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}xP{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ xPH\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ M{\isacharunderscore}{\kern0pt}powerset{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ xPH\ assms\ pair{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ domain{\isacharunderscore}{\kern0pt}closed\ cartprod{\isacharunderscore}{\kern0pt}closed\ image{\isacharunderscore}{\kern0pt}closed\ Union{\isacharunderscore}{\kern0pt}closed\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \ \isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ singleton{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base\ \isakeyword{where}\ {\isachardoublequoteopen}Hsep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}g{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base\ \isakeyword{where}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ wftrec{\isacharparenleft}{\kern0pt}Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharbraceleft}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ F\ \isakeyword{where}\ {\isachardoublequoteopen}F\ {\isasymequiv}\ {\isasymlambda}y\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ S\ \isakeyword{where}\ {\isachardoublequoteopen}S\ {\isasymequiv}\ {\isacharbraceleft}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}F{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ wftrec{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ wf{\isacharunderscore}{\kern0pt}trancl{\isacharcomma}{\kern0pt}rule\ wf{\isacharunderscore}{\kern0pt}Memrel{\isacharcomma}{\kern0pt}\ rule\ trans{\isacharunderscore}{\kern0pt}trancl{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\ F{\isacharunderscore}{\kern0pt}def\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ S{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ F{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ u\ \isakeyword{where}\ uH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}u{\isacharcomma}{\kern0pt}\ v{\isachargreater}{\kern0pt}\ {\isasymin}\ F{\isachardoublequoteclose}\ {\isachardoublequoteopen}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Fu\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}u\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}uH{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}lam{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}u\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ uH\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}trancl\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Fu\ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ S{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ S{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ uH\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ S{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ u\ \isakeyword{where}\ uH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ S{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ uin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isasymin}\ Memrel{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isacharcircum}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}Memrel{\isacharunderscore}{\kern0pt}trancl\ uH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ Fu{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}u\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ uin\ uH\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}u{\isacharcomma}{\kern0pt}\ F{\isacharbackquote}{\kern0pt}u{\isachargreater}{\kern0pt}\ {\isasymin}\ F{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}lam{\isacharcomma}{\kern0pt}\ subst\ domain{\isacharunderscore}{\kern0pt}lam{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}uin{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ F{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ uin\ uH\ Fu\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H{\isadigit{1}}\ H{\isadigit{2}}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ F{\isacharunderscore}{\kern0pt}def\ S{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm\ \isakeyword{where}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ s{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ p{\isacharcomma}{\kern0pt}\ s{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ h\ g\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}h\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymrightarrow}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}\ {\isasymrightarrow}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ h\ {\isacharbackquote}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ P{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Hsep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ h{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharparenleft}{\kern0pt}{\isasymlangle}x{\isacharcomma}{\kern0pt}\ P{\isasymrangle}{\isacharcomma}{\kern0pt}\ g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}lemma\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}a\ b\ c{\isachardot}{\kern0pt}\ b\ {\isacharequal}{\kern0pt}\ c\ {\isasymLongrightarrow}\ a\ {\isacharequal}{\kern0pt}\ b\ {\isasymlongleftrightarrow}\ a\ {\isacharequal}{\kern0pt}\ c{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ image{\isacharunderscore}{\kern0pt}lemma\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}f\ d\ v{\isachardot}{\kern0pt}\ function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ d\ {\isasymsubseteq}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ v\ {\isasymin}\ f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}d\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}a\ {\isasymin}\ d{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ f\ d\ v\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}d{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ aH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ v{\isachargreater}{\kern0pt}\ {\isasymin}\ f{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymin}\ d{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}a\ {\isacharequal}{\kern0pt}\ v{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}equality{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}aH{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isadigit{1}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}a{\isasymin}d{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ aH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ f\ d\ v\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ {\isasymsubseteq}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymexists}a{\isasymin}d{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ a{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ aH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ d{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachargreater}{\kern0pt}\ {\isasymin}\ f{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ aH\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ f\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ d{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ aH\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}h\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ h\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ h{\isacharbackquote}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ image{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Pi{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}h{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ B{\isacharequal}{\kern0pt}M\ \isakeyword{in}\ domain{\isacharunderscore}{\kern0pt}Pi{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ P{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharcomma}{\kern0pt}\ rule\ iff{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ y{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y\ {\isasymin}\ eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}yP\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ v\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}yP{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ g\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ image{\isacharunderscore}{\kern0pt}lemma{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Pi{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}eclose{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ B{\isacharequal}{\kern0pt}M\ \isakeyword{in}\ domain{\isacharunderscore}{\kern0pt}Pi{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}eclose\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ h\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ v\ {\isasymin}\ g\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}h\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ g\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ {\isacharbraceleft}{\kern0pt}P{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ force{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ env\ i\ j\ k\ x\ s\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}env\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}i\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}j\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}k\ {\isacharless}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}env{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}j{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}nth{\isacharparenleft}{\kern0pt}k{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ s{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}i{\isacharcomma}{\kern0pt}\ j{\isacharcomma}{\kern0pt}\ k{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ env{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ s\ {\isacharequal}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ G{\isacharequal}{\kern0pt}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M\ \isakeyword{in}\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{9}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ G\ {\isacharequal}{\kern0pt}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M\ \isakeyword{and}\ Gfm\ {\isacharequal}{\kern0pt}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm\ \isakeyword{in}\ \ memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{5}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}in\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ domH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ {\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ ranH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff\ assms\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymsubseteq}\ {\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ vin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ vin\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ yin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ {\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ domH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ vin\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ pin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ranH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ypH\ yin\ pin\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ def{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}subset\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}\ x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ z\ y\ p{\isacharcomma}{\kern0pt}\ subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ Pow{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ def{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x{\isadigit{0}}\ x\ {\isasympi}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isachardoublequoteclose}\ \ \ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ main\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isadigit{0}}\ {\isachardot}{\kern0pt}\ x{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}\ x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isadigit{0}}\ {\isachardot}{\kern0pt}\ x{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}\ x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ ed{\isacharunderscore}{\kern0pt}induction{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharcomma}{\kern0pt}\ rule\ ballI{\isacharcomma}{\kern0pt}\ rename{\isacharunderscore}{\kern0pt}tac\ a\ x{\isadigit{0}}\ x{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x{\isadigit{0}}\ x\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymAnd}y{\isadigit{0}}{\isachardot}{\kern0pt}\ ed{\isacharparenleft}{\kern0pt}y{\isadigit{0}}{\isacharcomma}{\kern0pt}\ x{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}\ {\isasymforall}\ y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ xpname\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}subset\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ pnautoeq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pn{\isacharunderscore}{\kern0pt}auto\ xpname\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ deleq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ Pow{\isacharparenleft}{\kern0pt}{\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasyminter}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ def{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ xsubset\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymsubseteq}\ {\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ deleq\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ domH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ {\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ \isacommand{assume}\isamarkupfalse%
\ yindom\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y{\isadigit{0}}\ \isakeyword{where}\ y{\isadigit{0}}H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y{\isadigit{0}}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ xsubset\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ ed{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ {\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ y{\isadigit{0}}H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymsubseteq}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ xpname\ relation{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ pnautoeq\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pnautoeq\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y\ {\isasymin}\ {\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ypH\ domH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ xpname\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}RepFun{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ deleq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ xpname\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ x{\isadigit{0}}\ {\isasympi}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}x{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assm{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bij{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}bij{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}def\ inj{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ assm{\isadigit{1}}\ H\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}converse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}converse{\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ O\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ eq{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ comp{\isacharunderscore}{\kern0pt}fun{\isacharunderscore}{\kern0pt}apply{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}type{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ id{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ left{\isacharunderscore}{\kern0pt}comp{\isacharunderscore}{\kern0pt}inverse{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}bij\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ y\ {\isacharequal}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}strong{\isacharunderscore}{\kern0pt}replacement{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}\ {\isasymlambda}x\ y{\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}\ {\isadigit{2}}{\isacharcomma}{\kern0pt}\ {\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}P{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ replacement{\isacharunderscore}{\kern0pt}ax{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ rule\ is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{4}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ rule\ le{\isacharunderscore}{\kern0pt}trans{\isacharcomma}{\kern0pt}\ rule\ arity{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}memrel{\isacharunderscore}{\kern0pt}wftrec{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}\ Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}Hsep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp\ del{\isacharcolon}{\kern0pt}FOL{\isacharunderscore}{\kern0pt}sats{\isacharunderscore}{\kern0pt}iff\ pair{\isacharunderscore}{\kern0pt}abs\ add{\isacharcolon}{\kern0pt}\ fm{\isacharunderscore}{\kern0pt}defs\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{3}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Un{\isacharunderscore}{\kern0pt}least{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}cong{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ x\ y{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ env\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharbrackright}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}P{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ i{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{and}\ j{\isacharequal}{\kern0pt}{\isadigit{2}}\ \isakeyword{and}\ k{\isacharequal}{\kern0pt}{\isadigit{1}}\ \isakeyword{in}\ sats{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{8}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ H\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}repfun{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}A\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ RepFun{\isacharunderscore}{\kern0pt}closed{\isacharcomma}{\kern0pt}\ rule\ strong{\isacharunderscore}{\kern0pt}replacement{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}base{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ transM\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{definition}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}Base\ \isakeyword{where}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ \ {\isasymequiv}\ {\isasymUnion}{\isacharbraceleft}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasyminter}\ HS{\isachardoublequoteclose}\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}separation{\isacharcomma}{\kern0pt}\ rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ Union{\isacharunderscore}{\kern0pt}closed{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}repfun{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ domain{\isacharunderscore}{\kern0pt}closed\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymsubseteq}\ HS\ {\isasymtimes}\ P\ {\isasymLongrightarrow}\ y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}simp{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}y{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}in{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}v\ {\isasymin}\ HS\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ v\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ x{\isadigit{0}}\ {\isasympi}\ x\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}x{\isadigit{0}}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ {\isasymG}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ HS\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}iff\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ HS\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymexists}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}base{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ HS{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bex{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ sep{\isacharunderscore}{\kern0pt}base{\isacharunderscore}{\kern0pt}closed{\isacharunderscore}{\kern0pt}under{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}domain{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}name\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iff{\isacharunderscore}{\kern0pt}flip{\isacharcomma}{\kern0pt}\ rule\ in{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ x\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}x{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ex{\isacharunderscore}{\kern0pt}separation{\isacharunderscore}{\kern0pt}base\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ X\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}X\ {\isasymsubseteq}\ HS{\isachardoublequoteclose}\ {\isachardoublequoteopen}X\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}S\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ X\ {\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ S{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ A\ \isakeyword{where}\ {\isachardoublequoteopen}A\ {\isasymequiv}\ X\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ B\ \isakeyword{where}\ {\isachardoublequoteopen}B\ {\isasymequiv}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Apname\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ A{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ cartprod{\isacharunderscore}{\kern0pt}closed{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ HS{\isacharunderscore}{\kern0pt}iff\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ pname\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ P{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ conjI{\isacharcomma}{\kern0pt}\ rule\ to{\isacharunderscore}{\kern0pt}rin{\isacharcomma}{\kern0pt}\ rule\ cartprod{\isacharunderscore}{\kern0pt}closed{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ HS{\isacharunderscore}{\kern0pt}iff\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}{\isasympi}{\isachardot}{\kern0pt}\ {\isasympi}\ {\isasymin}\ {\isasymG}\ {\isasymLongrightarrow}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ {\isasympi}\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ piin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ {\isasymG}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ pipauto\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}auto{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}group{\isacharunderscore}{\kern0pt}def\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pname\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharcomma}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ piin\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}def\ HS{\isacharunderscore}{\kern0pt}iff\ ypH\ Apname\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{2}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasympi}{\isacharbackquote}{\kern0pt}p\ {\isasymin}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}value{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pipauto\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ ypH\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ H{\isadigit{1}}\ H{\isadigit{2}}\ ypH\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ ypH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ ypname\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ypH\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}def\ HS{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}P{\isacharunderscore}{\kern0pt}names{\isacharcomma}{\kern0pt}\ P{\isacharunderscore}{\kern0pt}names{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj{\isacharcomma}{\kern0pt}\ rule\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}bij{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pipauto\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y{\isacharprime}{\kern0pt}\ \isakeyword{where}\ y{\isacharprime}{\kern0pt}H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isacharequal}{\kern0pt}\ Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y{\isacharprime}{\kern0pt}\ {\isasymin}\ P{\isacharunderscore}{\kern0pt}names{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ ypname\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ y{\isacharprime}{\kern0pt}in\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y{\isacharprime}{\kern0pt}\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ {\isasympi}{\isacharequal}{\kern0pt}{\isasympi}\ \isakeyword{in}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ y{\isacharprime}{\kern0pt}H\ ypH\ piin\ assms\ Apname\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ surj{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}surj{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ pipauto\ P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\ is{\isacharunderscore}{\kern0pt}P{\isacharunderscore}{\kern0pt}auto{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p{\isacharprime}{\kern0pt}\ \isakeyword{where}\ p{\isacharprime}{\kern0pt}H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p{\isacharprime}{\kern0pt}\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isacharequal}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ surj{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}y{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}\ {\isasympi}{\isacharbackquote}{\kern0pt}p{\isacharprime}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ y{\isacharprime}{\kern0pt}H\ p{\isacharprime}{\kern0pt}H\ y{\isacharprime}{\kern0pt}in\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}{\isasymlangle}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ y{\isacharcomma}{\kern0pt}\ {\isasympi}\ {\isacharbackquote}{\kern0pt}\ p{\isasymrangle}\ {\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isacharprime}{\kern0pt}H\ y{\isacharprime}{\kern0pt}H\ ypH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Pn{\isacharunderscore}{\kern0pt}auto{\isacharparenleft}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P\ {\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sym{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isasymG}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sym{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ symH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}symmetric{\isacharparenleft}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ symmetric{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymG}{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}{\isasymF}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P\ {\isasymin}\ HS{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ iffD{\isadigit{2}}{\isacharcomma}{\kern0pt}\ rule\ HS{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ pname\ symH\ sep{\isacharunderscore}{\kern0pt}Base{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ insymext\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}\ {\isasymin}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ SymExt{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ Xsubset{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}X\ {\isasymsubseteq}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ A{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharcomma}{\kern0pt}\ rule\ domain{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}sep{\isacharunderscore}{\kern0pt}Base{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}one\ \isakeyword{in}\ domainI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ subsetH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ X\ {\isacharbraceright}{\kern0pt}\ {\isasymsubseteq}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ x\ {\isasymin}\ X\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ xH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ X{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ {\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}\ def{\isacharunderscore}{\kern0pt}val{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ bexI{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ rule\ conjI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Xsubset\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}one\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}G\ one{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}P\ generic\ Xsubset\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ xH\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ sep{\isacharunderscore}{\kern0pt}Base{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetH{\isacharcomma}{\kern0pt}\ rule\ insymext{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/SymExt_Separation_Base.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%28=10%:%
%:%29=11%:%
%:%30=11%:%
%:%31=12%:%
%:%32=12%:%
%:%33=13%:%
%:%34=13%:%
%:%35=14%:%
%:%36=14%:%
%:%39=15%:%
%:%43=15%:%
%:%44=15%:%
%:%45=16%:%
%:%46=16%:%
%:%47=17%:%
%:%48=17%:%
%:%49=18%:%
%:%50=18%:%
%:%55=18%:%
%:%58=19%:%
%:%59=20%:%
%:%60=20%:%
%:%61=21%:%
%:%62=21%:%
%:%64=23%:%
%:%65=24%:%
%:%66=25%:%
%:%67=25%:%
%:%68=26%:%
%:%71=27%:%
%:%75=27%:%
%:%76=27%:%
%:%77=28%:%
%:%78=28%:%
%:%79=29%:%
%:%80=29%:%
%:%81=30%:%
%:%82=31%:%
%:%83=31%:%
%:%84=32%:%
%:%85=32%:%
%:%86=33%:%
%:%87=33%:%
%:%88=34%:%
%:%89=34%:%
%:%90=35%:%
%:%91=35%:%
%:%92=36%:%
%:%93=36%:%
%:%94=37%:%
%:%95=37%:%
%:%96=38%:%
%:%97=38%:%
%:%98=39%:%
%:%99=39%:%
%:%100=40%:%
%:%101=40%:%
%:%102=41%:%
%:%103=41%:%
%:%104=42%:%
%:%105=42%:%
%:%106=43%:%
%:%107=43%:%
%:%112=43%:%
%:%115=44%:%
%:%116=45%:%
%:%117=45%:%
%:%118=46%:%
%:%119=47%:%
%:%120=48%:%
%:%121=49%:%
%:%122=50%:%
%:%123=51%:%
%:%124=52%:%
%:%125=53%:%
%:%128=54%:%
%:%129=55%:%
%:%133=55%:%
%:%134=55%:%
%:%135=56%:%
%:%136=56%:%
%:%137=57%:%
%:%138=57%:%
%:%139=58%:%
%:%140=58%:%
%:%141=59%:%
%:%142=59%:%
%:%143=60%:%
%:%144=60%:%
%:%145=61%:%
%:%146=61%:%
%:%147=62%:%
%:%148=62%:%
%:%149=63%:%
%:%150=63%:%
%:%155=63%:%
%:%158=64%:%
%:%159=65%:%
%:%160=65%:%
%:%161=66%:%
%:%185=90%:%
%:%186=91%:%
%:%187=92%:%
%:%188=92%:%
%:%195=93%:%
%:%196=93%:%
%:%197=94%:%
%:%198=94%:%
%:%199=95%:%
%:%200=95%:%
%:%201=95%:%
%:%202=95%:%
%:%203=95%:%
%:%204=96%:%
%:%205=96%:%
%:%206=96%:%
%:%207=96%:%
%:%208=96%:%
%:%209=97%:%
%:%210=97%:%
%:%211=98%:%
%:%212=98%:%
%:%213=99%:%
%:%214=100%:%
%:%215=100%:%
%:%216=101%:%
%:%217=101%:%
%:%218=101%:%
%:%219=101%:%
%:%220=102%:%
%:%221=103%:%
%:%222=103%:%
%:%223=104%:%
%:%224=104%:%
%:%225=105%:%
%:%226=105%:%
%:%227=106%:%
%:%228=106%:%
%:%229=107%:%
%:%230=108%:%
%:%231=108%:%
%:%232=109%:%
%:%233=109%:%
%:%234=110%:%
%:%235=110%:%
%:%236=111%:%
%:%237=111%:%
%:%238=112%:%
%:%239=113%:%
%:%240=113%:%
%:%241=113%:%
%:%242=114%:%
%:%243=114%:%
%:%244=114%:%
%:%245=115%:%
%:%251=115%:%
%:%254=116%:%
%:%255=117%:%
%:%256=117%:%
%:%257=118%:%
%:%258=119%:%
%:%259=119%:%
%:%260=120%:%
%:%261=121%:%
%:%262=121%:%
%:%263=122%:%
%:%264=123%:%
%:%265=124%:%
%:%272=125%:%
%:%273=125%:%
%:%274=126%:%
%:%275=127%:%
%:%276=127%:%
%:%277=128%:%
%:%278=128%:%
%:%279=129%:%
%:%280=130%:%
%:%281=130%:%
%:%282=131%:%
%:%283=131%:%
%:%284=132%:%
%:%285=132%:%
%:%286=133%:%
%:%287=133%:%
%:%288=134%:%
%:%289=134%:%
%:%290=135%:%
%:%291=135%:%
%:%292=136%:%
%:%293=137%:%
%:%294=137%:%
%:%295=138%:%
%:%296=138%:%
%:%297=139%:%
%:%298=139%:%
%:%299=139%:%
%:%300=140%:%
%:%301=140%:%
%:%302=140%:%
%:%303=140%:%
%:%304=141%:%
%:%305=141%:%
%:%306=142%:%
%:%307=142%:%
%:%308=143%:%
%:%309=143%:%
%:%310=144%:%
%:%311=144%:%
%:%312=145%:%
%:%313=145%:%
%:%314=146%:%
%:%315=146%:%
%:%316=147%:%
%:%317=147%:%
%:%318=147%:%
%:%319=147%:%
%:%320=148%:%
%:%321=148%:%
%:%322=148%:%
%:%323=148%:%
%:%324=148%:%
%:%325=148%:%
%:%326=149%:%
%:%327=149%:%
%:%328=149%:%
%:%329=150%:%
%:%330=150%:%
%:%331=151%:%
%:%332=151%:%
%:%333=152%:%
%:%334=152%:%
%:%335=153%:%
%:%336=153%:%
%:%337=154%:%
%:%338=154%:%
%:%339=154%:%
%:%340=155%:%
%:%341=155%:%
%:%342=155%:%
%:%343=155%:%
%:%344=155%:%
%:%345=156%:%
%:%346=156%:%
%:%347=156%:%
%:%348=156%:%
%:%349=156%:%
%:%350=157%:%
%:%351=157%:%
%:%352=158%:%
%:%353=158%:%
%:%354=159%:%
%:%355=159%:%
%:%356=160%:%
%:%357=160%:%
%:%358=161%:%
%:%359=161%:%
%:%360=162%:%
%:%361=162%:%
%:%362=163%:%
%:%363=163%:%
%:%364=164%:%
%:%365=164%:%
%:%366=165%:%
%:%367=165%:%
%:%368=166%:%
%:%369=166%:%
%:%370=166%:%
%:%371=167%:%
%:%372=167%:%
%:%373=167%:%
%:%374=168%:%
%:%375=168%:%
%:%376=169%:%
%:%377=170%:%
%:%378=170%:%
%:%379=171%:%
%:%380=171%:%
%:%381=172%:%
%:%382=172%:%
%:%383=173%:%
%:%384=173%:%
%:%385=174%:%
%:%391=174%:%
%:%394=175%:%
%:%395=176%:%
%:%396=176%:%
%:%397=177%:%
%:%398=178%:%
%:%399=178%:%
%:%400=179%:%
%:%401=180%:%
%:%402=181%:%
%:%403=182%:%
%:%410=183%:%
%:%411=183%:%
%:%412=184%:%
%:%413=185%:%
%:%414=185%:%
%:%415=185%:%
%:%416=186%:%
%:%417=187%:%
%:%418=187%:%
%:%419=188%:%
%:%420=188%:%
%:%421=189%:%
%:%422=189%:%
%:%423=189%:%
%:%424=190%:%
%:%425=190%:%
%:%426=190%:%
%:%427=190%:%
%:%428=191%:%
%:%429=191%:%
%:%430=191%:%
%:%431=192%:%
%:%432=192%:%
%:%433=192%:%
%:%434=192%:%
%:%435=192%:%
%:%436=193%:%
%:%437=193%:%
%:%438=194%:%
%:%439=194%:%
%:%440=194%:%
%:%441=195%:%
%:%442=195%:%
%:%443=195%:%
%:%444=195%:%
%:%445=196%:%
%:%446=196%:%
%:%447=197%:%
%:%448=197%:%
%:%449=198%:%
%:%450=198%:%
%:%451=199%:%
%:%452=199%:%
%:%453=200%:%
%:%454=200%:%
%:%455=200%:%
%:%456=201%:%
%:%457=201%:%
%:%458=202%:%
%:%459=202%:%
%:%460=203%:%
%:%461=203%:%
%:%462=204%:%
%:%463=205%:%
%:%464=205%:%
%:%465=206%:%
%:%466=206%:%
%:%467=207%:%
%:%468=207%:%
%:%469=208%:%
%:%470=208%:%
%:%471=209%:%
%:%472=209%:%
%:%473=210%:%
%:%474=210%:%
%:%475=211%:%
%:%476=211%:%
%:%477=212%:%
%:%478=212%:%
%:%479=213%:%
%:%480=213%:%
%:%481=214%:%
%:%482=214%:%
%:%483=215%:%
%:%484=215%:%
%:%485=216%:%
%:%486=216%:%
%:%487=217%:%
%:%488=217%:%
%:%489=217%:%
%:%490=218%:%
%:%491=218%:%
%:%492=219%:%
%:%493=219%:%
%:%494=220%:%
%:%495=220%:%
%:%496=221%:%
%:%497=221%:%
%:%498=222%:%
%:%499=222%:%
%:%500=223%:%
%:%501=223%:%
%:%502=224%:%
%:%503=224%:%
%:%504=224%:%
%:%505=224%:%
%:%506=225%:%
%:%507=225%:%
%:%508=225%:%
%:%509=226%:%
%:%510=226%:%
%:%511=227%:%
%:%512=227%:%
%:%513=228%:%
%:%514=228%:%
%:%515=229%:%
%:%516=229%:%
%:%517=230%:%
%:%518=230%:%
%:%519=231%:%
%:%520=231%:%
%:%521=232%:%
%:%522=232%:%
%:%523=233%:%
%:%524=233%:%
%:%525=233%:%
%:%526=233%:%
%:%527=234%:%
%:%528=234%:%
%:%529=235%:%
%:%530=236%:%
%:%531=236%:%
%:%532=236%:%
%:%533=237%:%
%:%534=237%:%
%:%535=238%:%
%:%536=238%:%
%:%537=239%:%
%:%538=239%:%
%:%539=240%:%
%:%540=240%:%
%:%541=241%:%
%:%547=241%:%
%:%550=242%:%
%:%551=243%:%
%:%552=243%:%
%:%553=244%:%
%:%554=245%:%
%:%555=246%:%
%:%556=247%:%
%:%559=248%:%
%:%560=249%:%
%:%564=249%:%
%:%565=249%:%
%:%566=250%:%
%:%567=250%:%
%:%568=251%:%
%:%569=251%:%
%:%570=252%:%
%:%571=252%:%
%:%572=253%:%
%:%573=253%:%
%:%574=254%:%
%:%575=254%:%
%:%576=255%:%
%:%577=255%:%
%:%578=256%:%
%:%579=256%:%
%:%580=257%:%
%:%581=257%:%
%:%582=258%:%
%:%583=258%:%
%:%584=259%:%
%:%585=259%:%
%:%586=260%:%
%:%587=260%:%
%:%588=261%:%
%:%589=261%:%
%:%590=262%:%
%:%591=262%:%
%:%596=262%:%
%:%599=263%:%
%:%600=264%:%
%:%601=264%:%
%:%602=265%:%
%:%603=266%:%
%:%604=267%:%
%:%607=268%:%
%:%608=269%:%
%:%612=269%:%
%:%613=269%:%
%:%614=270%:%
%:%615=270%:%
%:%616=271%:%
%:%617=271%:%
%:%618=272%:%
%:%619=272%:%
%:%620=273%:%
%:%621=273%:%
%:%622=274%:%
%:%623=274%:%
%:%624=275%:%
%:%625=275%:%
%:%626=276%:%
%:%627=276%:%
%:%628=277%:%
%:%629=277%:%
%:%630=278%:%
%:%631=278%:%
%:%632=279%:%
%:%633=279%:%
%:%634=280%:%
%:%635=280%:%
%:%636=281%:%
%:%637=281%:%
%:%642=281%:%
%:%645=282%:%
%:%646=283%:%
%:%647=283%:%
%:%648=284%:%
%:%649=285%:%
%:%650=286%:%
%:%657=287%:%
%:%658=287%:%
%:%659=288%:%
%:%660=288%:%
%:%661=289%:%
%:%662=289%:%
%:%663=290%:%
%:%664=290%:%
%:%665=290%:%
%:%666=291%:%
%:%667=292%:%
%:%668=292%:%
%:%669=293%:%
%:%670=293%:%
%:%671=294%:%
%:%672=294%:%
%:%673=295%:%
%:%674=295%:%
%:%675=296%:%
%:%676=297%:%
%:%677=297%:%
%:%678=298%:%
%:%679=298%:%
%:%680=299%:%
%:%681=299%:%
%:%682=300%:%
%:%683=301%:%
%:%684=301%:%
%:%685=302%:%
%:%686=302%:%
%:%687=303%:%
%:%688=303%:%
%:%689=303%:%
%:%690=304%:%
%:%691=304%:%
%:%692=304%:%
%:%693=304%:%
%:%694=304%:%
%:%695=305%:%
%:%696=305%:%
%:%697=305%:%
%:%698=305%:%
%:%699=306%:%
%:%700=306%:%
%:%701=306%:%
%:%702=306%:%
%:%703=306%:%
%:%704=307%:%
%:%705=307%:%
%:%706=307%:%
%:%707=307%:%
%:%708=308%:%
%:%709=308%:%
%:%710=308%:%
%:%711=308%:%
%:%712=308%:%
%:%713=309%:%
%:%714=309%:%
%:%715=309%:%
%:%716=309%:%
%:%717=310%:%
%:%718=310%:%
%:%719=311%:%
%:%720=311%:%
%:%721=311%:%
%:%722=312%:%
%:%723=312%:%
%:%724=313%:%
%:%725=313%:%
%:%726=314%:%
%:%727=314%:%
%:%728=315%:%
%:%729=315%:%
%:%730=316%:%
%:%731=317%:%
%:%732=317%:%
%:%733=317%:%
%:%734=317%:%
%:%735=317%:%
%:%736=318%:%
%:%742=318%:%
%:%745=319%:%
%:%746=320%:%
%:%747=320%:%
%:%748=321%:%
%:%749=322%:%
%:%750=323%:%
%:%757=324%:%
%:%758=324%:%
%:%759=325%:%
%:%760=325%:%
%:%761=326%:%
%:%762=326%:%
%:%763=327%:%
%:%764=327%:%
%:%765=328%:%
%:%766=328%:%
%:%767=329%:%
%:%768=329%:%
%:%769=330%:%
%:%770=330%:%
%:%771=331%:%
%:%772=331%:%
%:%773=332%:%
%:%774=332%:%
%:%775=333%:%
%:%776=333%:%
%:%777=334%:%
%:%778=334%:%
%:%779=334%:%
%:%780=335%:%
%:%781=335%:%
%:%782=336%:%
%:%783=336%:%
%:%784=337%:%
%:%785=337%:%
%:%786=338%:%
%:%787=338%:%
%:%788=339%:%
%:%789=339%:%
%:%790=340%:%
%:%791=340%:%
%:%792=340%:%
%:%793=341%:%
%:%794=341%:%
%:%795=341%:%
%:%796=342%:%
%:%797=342%:%
%:%798=343%:%
%:%799=343%:%
%:%800=343%:%
%:%801=343%:%
%:%802=343%:%
%:%803=344%:%
%:%809=344%:%
%:%812=345%:%
%:%813=346%:%
%:%814=346%:%
%:%815=347%:%
%:%816=348%:%
%:%817=349%:%
%:%824=350%:%
%:%825=350%:%
%:%826=351%:%
%:%827=351%:%
%:%828=352%:%
%:%829=352%:%
%:%830=353%:%
%:%831=353%:%
%:%832=353%:%
%:%833=354%:%
%:%834=355%:%
%:%835=355%:%
%:%836=355%:%
%:%837=355%:%
%:%838=356%:%
%:%839=356%:%
%:%840=356%:%
%:%841=356%:%
%:%842=357%:%
%:%843=357%:%
%:%844=358%:%
%:%845=358%:%
%:%846=358%:%
%:%847=359%:%
%:%848=359%:%
%:%849=359%:%
%:%850=359%:%
%:%851=360%:%
%:%852=361%:%
%:%853=361%:%
%:%854=362%:%
%:%855=362%:%
%:%856=363%:%
%:%857=363%:%
%:%858=363%:%
%:%859=364%:%
%:%860=364%:%
%:%861=364%:%
%:%862=364%:%
%:%863=364%:%
%:%864=365%:%
%:%865=365%:%
%:%866=365%:%
%:%867=365%:%
%:%868=365%:%
%:%869=366%:%
%:%870=366%:%
%:%871=366%:%
%:%872=366%:%
%:%873=366%:%
%:%874=367%:%
%:%875=367%:%
%:%876=368%:%
%:%877=369%:%
%:%878=369%:%
%:%879=370%:%
%:%880=370%:%
%:%881=371%:%
%:%882=371%:%
%:%883=371%:%
%:%884=372%:%
%:%885=372%:%
%:%886=372%:%
%:%887=373%:%
%:%888=373%:%
%:%889=374%:%
%:%890=374%:%
%:%891=375%:%
%:%892=375%:%
%:%893=376%:%
%:%894=376%:%
%:%895=376%:%
%:%896=376%:%
%:%897=376%:%
%:%898=377%:%
%:%899=377%:%
%:%900=377%:%
%:%901=377%:%
%:%902=377%:%
%:%903=378%:%
%:%904=378%:%
%:%905=378%:%
%:%906=378%:%
%:%907=379%:%
%:%908=379%:%
%:%909=379%:%
%:%910=379%:%
%:%911=379%:%
%:%912=380%:%
%:%913=380%:%
%:%914=380%:%
%:%915=380%:%
%:%916=380%:%
%:%917=381%:%
%:%918=381%:%
%:%919=382%:%
%:%920=383%:%
%:%921=383%:%
%:%922=384%:%
%:%923=384%:%
%:%924=385%:%
%:%925=385%:%
%:%926=386%:%
%:%927=386%:%
%:%928=387%:%
%:%929=387%:%
%:%930=388%:%
%:%931=388%:%
%:%932=389%:%
%:%933=389%:%
%:%934=390%:%
%:%935=390%:%
%:%936=391%:%
%:%937=392%:%
%:%938=392%:%
%:%939=392%:%
%:%940=392%:%
%:%941=392%:%
%:%942=393%:%
%:%948=393%:%
%:%951=394%:%
%:%952=395%:%
%:%953=395%:%
%:%954=396%:%
%:%955=397%:%
%:%956=398%:%
%:%963=399%:%
%:%964=399%:%
%:%965=400%:%
%:%966=400%:%
%:%967=401%:%
%:%968=401%:%
%:%969=401%:%
%:%970=402%:%
%:%971=402%:%
%:%972=403%:%
%:%973=403%:%
%:%974=404%:%
%:%975=404%:%
%:%976=405%:%
%:%977=405%:%
%:%978=406%:%
%:%979=406%:%
%:%980=407%:%
%:%981=407%:%
%:%982=408%:%
%:%983=408%:%
%:%984=409%:%
%:%985=409%:%
%:%986=410%:%
%:%987=410%:%
%:%988=411%:%
%:%989=411%:%
%:%990=411%:%
%:%991=412%:%
%:%992=412%:%
%:%993=413%:%
%:%994=413%:%
%:%995=414%:%
%:%996=414%:%
%:%997=415%:%
%:%998=415%:%
%:%999=416%:%
%:%1000=416%:%
%:%1001=417%:%
%:%1002=417%:%
%:%1003=418%:%
%:%1004=418%:%
%:%1005=419%:%
%:%1006=419%:%
%:%1007=420%:%
%:%1008=420%:%
%:%1009=421%:%
%:%1010=421%:%
%:%1011=422%:%
%:%1012=422%:%
%:%1013=423%:%
%:%1014=424%:%
%:%1015=424%:%
%:%1016=425%:%
%:%1017=425%:%
%:%1018=426%:%
%:%1019=426%:%
%:%1020=427%:%
%:%1021=427%:%
%:%1022=428%:%
%:%1023=428%:%
%:%1024=428%:%
%:%1025=429%:%
%:%1026=429%:%
%:%1027=430%:%
%:%1028=430%:%
%:%1029=431%:%
%:%1030=431%:%
%:%1031=432%:%
%:%1032=432%:%
%:%1033=433%:%
%:%1034=433%:%
%:%1035=433%:%
%:%1036=434%:%
%:%1037=434%:%
%:%1038=435%:%
%:%1039=435%:%
%:%1040=436%:%
%:%1041=436%:%
%:%1042=437%:%
%:%1043=437%:%
%:%1044=437%:%
%:%1045=437%:%
%:%1046=437%:%
%:%1047=438%:%
%:%1048=439%:%
%:%1049=439%:%
%:%1050=439%:%
%:%1051=439%:%
%:%1052=439%:%
%:%1053=440%:%
%:%1059=440%:%
%:%1062=441%:%
%:%1063=442%:%
%:%1064=442%:%
%:%1065=443%:%
%:%1072=444%:%
%:%1073=444%:%
%:%1074=445%:%
%:%1075=445%:%
%:%1076=446%:%
%:%1077=446%:%
%:%1078=447%:%
%:%1079=447%:%
%:%1080=448%:%
%:%1081=448%:%
%:%1082=449%:%
%:%1083=449%:%
%:%1084=450%:%
%:%1085=450%:%
%:%1086=451%:%
%:%1087=451%:%
%:%1088=452%:%
%:%1089=452%:%
%:%1090=453%:%
%:%1091=453%:%
%:%1092=454%:%
%:%1093=454%:%
%:%1094=455%:%
%:%1095=455%:%
%:%1096=456%:%
%:%1097=456%:%
%:%1098=457%:%
%:%1099=457%:%
%:%1100=458%:%
%:%1101=458%:%
%:%1102=459%:%
%:%1103=459%:%
%:%1104=460%:%
%:%1105=460%:%
%:%1106=461%:%
%:%1107=461%:%
%:%1108=462%:%
%:%1109=462%:%
%:%1110=463%:%
%:%1111=463%:%
%:%1112=464%:%
%:%1118=464%:%
%:%1121=465%:%
%:%1122=466%:%
%:%1123=466%:%
%:%1124=467%:%
%:%1125=468%:%
%:%1126=469%:%
%:%1129=470%:%
%:%1130=471%:%
%:%1134=471%:%
%:%1135=471%:%
%:%1136=472%:%
%:%1137=472%:%
%:%1138=473%:%
%:%1139=473%:%
%:%1140=474%:%
%:%1141=474%:%
%:%1146=474%:%
%:%1149=475%:%
%:%1150=476%:%
%:%1151=476%:%
%:%1152=477%:%
%:%1153=478%:%
%:%1154=478%:%
%:%1155=479%:%
%:%1156=480%:%
%:%1157=481%:%
%:%1160=482%:%
%:%1161=483%:%
%:%1165=483%:%
%:%1166=483%:%
%:%1167=484%:%
%:%1168=484%:%
%:%1169=485%:%
%:%1170=485%:%
%:%1171=486%:%
%:%1172=486%:%
%:%1173=487%:%
%:%1174=487%:%
%:%1179=487%:%
%:%1182=488%:%
%:%1183=489%:%
%:%1184=489%:%
%:%1187=490%:%
%:%1191=490%:%
%:%1192=490%:%
%:%1193=491%:%
%:%1194=491%:%
%:%1195=492%:%
%:%1196=492%:%
%:%1197=493%:%
%:%1198=493%:%
%:%1199=494%:%
%:%1200=494%:%
%:%1201=495%:%
%:%1202=495%:%
%:%1203=496%:%
%:%1204=496%:%
%:%1205=497%:%
%:%1206=497%:%
%:%1207=498%:%
%:%1208=498%:%
%:%1213=498%:%
%:%1216=499%:%
%:%1217=500%:%
%:%1218=500%:%
%:%1221=501%:%
%:%1225=501%:%
%:%1226=501%:%
%:%1227=502%:%
%:%1228=502%:%
%:%1233=502%:%
%:%1236=503%:%
%:%1237=504%:%
%:%1238=504%:%
%:%1239=505%:%
%:%1240=506%:%
%:%1241=507%:%
%:%1248=508%:%
%:%1249=508%:%
%:%1250=509%:%
%:%1251=509%:%
%:%1252=510%:%
%:%1253=510%:%
%:%1254=510%:%
%:%1255=511%:%
%:%1256=511%:%
%:%1257=511%:%
%:%1258=512%:%
%:%1259=512%:%
%:%1260=513%:%
%:%1261=513%:%
%:%1262=514%:%
%:%1263=514%:%
%:%1264=515%:%
%:%1265=515%:%
%:%1266=516%:%
%:%1267=516%:%
%:%1268=517%:%
%:%1269=517%:%
%:%1270=518%:%
%:%1271=518%:%
%:%1272=518%:%
%:%1273=519%:%
%:%1274=519%:%
%:%1275=520%:%
%:%1276=520%:%
%:%1277=521%:%
%:%1278=521%:%
%:%1279=521%:%
%:%1280=521%:%
%:%1281=522%:%
%:%1287=522%:%
%:%1290=523%:%
%:%1291=524%:%
%:%1292=524%:%
%:%1293=525%:%
%:%1294=526%:%
%:%1295=527%:%
%:%1302=528%:%
%:%1303=528%:%
%:%1304=529%:%
%:%1305=529%:%
%:%1306=530%:%
%:%1307=530%:%
%:%1308=531%:%
%:%1309=532%:%
%:%1310=532%:%
%:%1311=533%:%
%:%1312=533%:%
%:%1313=534%:%
%:%1314=534%:%
%:%1315=535%:%
%:%1316=535%:%
%:%1317=536%:%
%:%1318=536%:%
%:%1319=537%:%
%:%1320=537%:%
%:%1321=538%:%
%:%1322=539%:%
%:%1323=539%:%
%:%1324=539%:%
%:%1325=540%:%
%:%1326=540%:%
%:%1327=541%:%
%:%1328=542%:%
%:%1329=542%:%
%:%1330=542%:%
%:%1331=543%:%
%:%1332=543%:%
%:%1333=544%:%
%:%1334=544%:%
%:%1335=545%:%
%:%1336=545%:%
%:%1337=546%:%
%:%1338=546%:%
%:%1339=547%:%
%:%1340=547%:%
%:%1341=548%:%
%:%1342=548%:%
%:%1343=549%:%
%:%1344=550%:%
%:%1345=550%:%
%:%1346=551%:%
%:%1347=551%:%
%:%1348=552%:%
%:%1349=552%:%
%:%1350=553%:%
%:%1351=553%:%
%:%1352=554%:%
%:%1353=554%:%
%:%1354=554%:%
%:%1355=555%:%
%:%1356=555%:%
%:%1357=556%:%
%:%1358=556%:%
%:%1359=557%:%
%:%1360=558%:%
%:%1361=558%:%
%:%1362=559%:%
%:%1363=559%:%
%:%1364=560%:%
%:%1365=560%:%
%:%1366=561%:%
%:%1367=561%:%
%:%1368=562%:%
%:%1369=562%:%
%:%1370=562%:%
%:%1371=563%:%
%:%1372=563%:%
%:%1373=564%:%
%:%1374=564%:%
%:%1375=564%:%
%:%1376=565%:%
%:%1377=565%:%
%:%1378=565%:%
%:%1379=565%:%
%:%1380=566%:%
%:%1381=567%:%
%:%1382=567%:%
%:%1383=568%:%
%:%1384=568%:%
%:%1385=569%:%
%:%1386=569%:%
%:%1387=570%:%
%:%1388=570%:%
%:%1389=571%:%
%:%1390=571%:%
%:%1391=572%:%
%:%1392=572%:%
%:%1393=573%:%
%:%1394=573%:%
%:%1395=574%:%
%:%1396=574%:%
%:%1397=575%:%
%:%1398=575%:%
%:%1399=576%:%
%:%1400=576%:%
%:%1401=577%:%
%:%1402=577%:%
%:%1403=578%:%
%:%1404=578%:%
%:%1405=579%:%
%:%1406=579%:%
%:%1407=579%:%
%:%1408=580%:%
%:%1409=580%:%
%:%1410=580%:%
%:%1411=580%:%
%:%1412=581%:%
%:%1413=582%:%
%:%1414=582%:%
%:%1415=582%:%
%:%1416=582%:%
%:%1417=583%:%
%:%1418=584%:%
%:%1419=584%:%
%:%1420=585%:%
%:%1421=585%:%
%:%1422=586%:%
%:%1423=586%:%
%:%1424=587%:%
%:%1425=587%:%
%:%1426=588%:%
%:%1427=588%:%
%:%1428=588%:%
%:%1429=589%:%
%:%1430=589%:%
%:%1431=589%:%
%:%1432=589%:%
%:%1433=590%:%
%:%1434=590%:%
%:%1435=591%:%
%:%1436=591%:%
%:%1437=592%:%
%:%1438=592%:%
%:%1439=593%:%
%:%1440=593%:%
%:%1441=594%:%
%:%1442=595%:%
%:%1443=595%:%
%:%1444=596%:%
%:%1445=596%:%
%:%1446=597%:%
%:%1447=597%:%
%:%1448=598%:%
%:%1449=598%:%
%:%1450=599%:%
%:%1451=599%:%
%:%1452=599%:%
%:%1453=600%:%
%:%1454=600%:%
%:%1455=600%:%
%:%1456=600%:%
%:%1457=601%:%
%:%1458=602%:%
%:%1459=602%:%
%:%1460=603%:%
%:%1461=603%:%
%:%1462=603%:%
%:%1463=604%:%
%:%1464=604%:%
%:%1465=604%:%
%:%1466=605%:%
%:%1467=605%:%
%:%1468=605%:%
%:%1469=606%:%
%:%1470=606%:%
%:%1471=607%:%
%:%1472=608%:%
%:%1473=608%:%
%:%1474=608%:%
%:%1475=608%:%
%:%1476=609%:%
%:%1477=609%:%
%:%1478=610%:%
%:%1479=611%:%
%:%1480=611%:%
%:%1481=611%:%
%:%1482=612%:%
%:%1483=612%:%
%:%1484=613%:%
%:%1485=613%:%
%:%1486=614%:%
%:%1487=615%:%
%:%1488=615%:%
%:%1489=615%:%
%:%1490=616%:%
%:%1491=616%:%
%:%1492=617%:%
%:%1493=617%:%
%:%1494=618%:%
%:%1495=618%:%
%:%1496=619%:%
%:%1497=620%:%
%:%1498=620%:%
%:%1499=621%:%
%:%1500=621%:%
%:%1501=622%:%
%:%1502=622%:%
%:%1503=623%:%
%:%1504=623%:%
%:%1505=624%:%
%:%1506=625%:%
%:%1507=625%:%
%:%1508=625%:%
%:%1509=626%:%
%:%1510=626%:%
%:%1511=626%:%
%:%1512=627%:%
%:%1513=628%:%
%:%1514=628%:%
%:%1515=629%:%
%:%1516=629%:%
%:%1517=630%:%
%:%1518=630%:%
%:%1519=631%:%
%:%1520=631%:%
%:%1521=632%:%
%:%1522=632%:%
%:%1523=633%:%
%:%1524=633%:%
%:%1525=634%:%
%:%1526=634%:%
%:%1527=635%:%
%:%1528=635%:%
%:%1529=636%:%
%:%1530=637%:%
%:%1531=637%:%
%:%1532=638%:%
%:%1533=638%:%
%:%1534=639%:%
%:%1535=639%:%
%:%1536=639%:%
%:%1537=640%:%
%:%1538=640%:%
%:%1539=640%:%
%:%1540=640%:%
%:%1541=641%:%
%:%1542=641%:%
%:%1543=641%:%
%:%1544=642%:%
%:%1545=642%:%
%:%1546=643%:%
%:%1547=643%:%
%:%1548=644%:%
%:%1549=644%:%
%:%1550=645%:%
%:%1551=645%:%
%:%1552=646%:%
%:%1553=646%:%
%:%1554=647%:%
%:%1555=647%:%
%:%1556=648%:%
%:%1557=648%:%
%:%1558=649%:%
%:%1559=649%:%
%:%1560=649%:%
%:%1561=650%:%
%:%1562=650%:%
%:%1563=651%:%
%:%1564=651%:%
%:%1565=652%:%
%:%1566=652%:%
%:%1567=653%:%
%:%1568=654%:%
%:%1569=654%:%
%:%1570=655%:%
%:%1571=655%:%
%:%1572=656%:%
%:%1573=656%:%
%:%1574=657%:%
%:%1575=657%:%
%:%1576=658%:%
%:%1582=658%:%
%:%1585=659%:%
%:%1586=660%:%
%:%1587=660%:%
%:%1594=661%:%