### 2024-11-01 code/output/document/Union_Axiom.tex

%
\begin{isabellebody}%
\setisabellecontext{Union{\isacharunderscore}{\kern0pt}Axiom}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{The Axiom of Unions in $M[G]$%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}Axiom\isanewline
\ \ \isakeyword{imports}\ Names\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isacommand{context}\isamarkupfalse%
\ forcing{\isacharunderscore}{\kern0pt}data\isanewline
\isakeyword{begin}\isanewline
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharcomma}{\kern0pt}i{\isacharbrackright}{\kern0pt}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body{\isacharparenleft}{\kern0pt}P{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}leq{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymtheta}p{\isacharparenright}{\kern0pt}\ {\isasymequiv}\ {\isacharparenleft}{\kern0pt}{\isasymexists}\ {\isasymsigma}{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ {\isasymexists}\ q{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}q{\isasymin}\ P{\isacharprime}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymlangle}{\isasymsigma}{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ {\isasymtau}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}{\kern0pt}{\isasymexists}\ r{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}r{\isasymin}P{\isacharprime}{\kern0pt}\ {\isasymand}\ {\isacharparenleft}{\kern0pt}{\isasymlangle}fst{\isacharparenleft}{\kern0pt}{\isasymtheta}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}r{\isasymrangle}\ {\isasymin}\ {\isasymsigma}\ {\isasymand}\ {\isasymlangle}snd{\isacharparenleft}{\kern0pt}{\isasymtheta}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}r{\isasymrangle}\ {\isasymin}\ leq{\isacharprime}{\kern0pt}\ {\isasymand}\ {\isasymlangle}snd{\isacharparenleft}{\kern0pt}{\isasymtheta}p{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ leq{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isasymequiv}\isanewline
\ \ \ \ Exists{\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ Exists{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ Exists\ {\isacharparenleft}{\kern0pt}\isanewline
\ \ \ \ Exists\ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{7}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ Exists\ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{6}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ Exists\ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{9}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ Exists\ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{4}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ Exists\ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{6}}{\isacharcomma}{\kern0pt}{\isadigit{2}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ Exists\ {\isacharparenleft}{\kern0pt}And{\isacharparenleft}{\kern0pt}pair{\isacharunderscore}{\kern0pt}fm{\isacharparenleft}{\kern0pt}{\isadigit{7}}{\isacharcomma}{\kern0pt}{\isadigit{5}}{\isacharcomma}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}Member{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharcomma}{\kern0pt}{\isadigit{1}}{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}type\ {\isacharbrackleft}{\kern0pt}TC{\isacharbrackright}{\kern0pt}{\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isasymin}formula{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{4}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ upair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ pair{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}auto\ simp\ add{\isacharcolon}{\kern0pt}\ nat{\isacharunderscore}{\kern0pt}simp{\isacharunderscore}{\kern0pt}union{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sats{\isacharunderscore}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}\ a\ {\isasymin}\ M{\isacharsemicolon}{\kern0pt}\ b\ {\isasymin}\ M\ {\isacharsemicolon}{\kern0pt}\ P{\isacharprime}{\kern0pt}\ {\isasymin}\ M\ {\isacharsemicolon}{\kern0pt}\ p\ {\isasymin}\ M\ {\isacharsemicolon}{\kern0pt}\ {\isasymtheta}\ {\isasymin}\ M\ {\isacharsemicolon}{\kern0pt}\ {\isasymtau}\ {\isasymin}\ M\ {\isacharsemicolon}{\kern0pt}\ leq{\isacharprime}{\kern0pt}\ {\isasymin}\ M\ {\isasymrbrakk}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}leq{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}P{\isacharprime}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}a{\isacharcomma}{\kern0pt}b{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\isanewline
\ \ \ \ \ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body{\isacharparenleft}{\kern0pt}P{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}leq{\isacharprime}{\kern0pt}{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharunderscore}{\kern0pt}def\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}def\ tuples{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ M{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ auto\ simp\ add\ {\isacharcolon}{\kern0pt}\ tuples{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domD\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ Transset{\isacharunderscore}{\kern0pt}M\ trans{\isacharunderscore}{\kern0pt}M\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ flip{\isacharcolon}{\kern0pt}\ setclass{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymequiv}\isanewline
\ \ \ \ {\isacharbraceleft}{\kern0pt}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P\ {\isachardot}{\kern0pt}\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}M\ {\isacharcolon}{\kern0pt}\ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}u\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ P\ {\isachardot}{\kern0pt}\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}\ x\ {\isachardot}{\kern0pt}\ sats{\isacharparenleft}{\kern0pt}M{\isacharcomma}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharcomma}{\kern0pt}{\isacharbrackleft}{\kern0pt}x{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}leq{\isacharbrackright}{\kern0pt}{\isacharat}{\kern0pt}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}leq{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}\ x\ {\isachardot}{\kern0pt}\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body{\isacharparenleft}{\kern0pt}P{\isacharcomma}{\kern0pt}leq{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtau}{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isasymin}M{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}d\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}closed\ Union{\isacharunderscore}{\kern0pt}closed\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}d\ {\isasymtimes}\ P\ {\isasymin}\ M{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ cartprod{\isacharunderscore}{\kern0pt}closed\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}arity{\isacharparenleft}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{6}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ arity{\isacharunderscore}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \ arity{\isacharunderscore}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}leq{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}P{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharcomma}{\kern0pt}leq{\isacharbrackright}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\ assms\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \ {\isacartoucheopen}arity{\isacharparenleft}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm{\isacharparenright}{\kern0pt}{\isasymle}{\isadigit{6}}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}separation{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}P{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ separation{\isacharunderscore}{\kern0pt}ax\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}d\ {\isasymtimes}\ P\ {\isasymin}\ M{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isacharbraceleft}{\kern0pt}\ u\ {\isasymin}\ {\isacharquery}{\kern0pt}d\ {\isasymtimes}\ P\ {\isachardot}{\kern0pt}\ {\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ \ separation{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymlongleftrightarrow}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{if}\ {\isachardoublequoteopen}x{\isasymin}\ {\isacharquery}{\kern0pt}d{\isasymtimes}P{\isachardoublequoteclose}\ \isakeyword{for}\ x\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}\ {\isacharquery}{\kern0pt}d{\isasymtimes}P{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isasymlangle}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymrangle}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Pair{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}eq\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x{\isasymin}{\isacharquery}{\kern0pt}d{\isasymtimes}P{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}d{\isasymin}M{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ mtrans\ fst{\isacharunderscore}{\kern0pt}type\ snd{\isacharunderscore}{\kern0pt}type\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}trans{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}{\isasymlangle}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ \ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}{\isasymlangle}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymrangle}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ sats{\isacharunderscore}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}fm\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ {\isacartoucheopen}{\isasymtau}{\isasymin}M{\isacartoucheclose}\ leq{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x\ {\isacharequal}{\kern0pt}\ {\isasymlangle}fst{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}snd{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isasymrangle}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharquery}{\kern0pt}Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ that\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ Collect{\isacharunderscore}{\kern0pt}cong\ A\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}MG{\isacharunderscore}{\kern0pt}Eq\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}a\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymUnion}\ a\ {\isacharequal}{\kern0pt}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isasymUnion}\ {\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ i\ \isakeyword{where}\ {\isachardoublequoteopen}i\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ i{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtau}\ {\isasymin}\ M{\isacartoucheclose}\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymsigma}\ q\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}q\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymsigma}{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ {\isasymtau}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ i{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair\ domD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}x\ {\isasymin}\ i{\isacartoucheclose}\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtheta}\ r\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}r\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}r{\isasymrangle}\ {\isasymin}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymtheta}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair\ domD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymlangle}{\isasymsigma}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}{\isasymtau}{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymtheta}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}q{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}r{\isasymin}G{\isacartoucheclose}\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\isanewline
\ \ \ \ \ \ A{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}p{\isacharcomma}{\kern0pt}r{\isasymrangle}\ {\isasymin}\ leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}p{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isasymin}\ P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymin}\ P{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ low{\isacharunderscore}{\kern0pt}bound{\isacharunderscore}{\kern0pt}filter\ filterD\ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}M{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ mtrans\ P{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}M\ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}trans{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ A\ {\isacartoucheopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}r{\isasymrangle}\ {\isasymin}\ {\isasymsigma}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymlangle}{\isasymsigma}{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ {\isasymtau}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymtheta}\ {\isasymin}\ M{\isacartoucheclose}\ {\isacartoucheopen}{\isasymtheta}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \ {\isacartoucheopen}{\isasymsigma}{\isasymin}M{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}P{\isacartoucheclose}\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}elem\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}x{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}a{\isacharequal}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isasymUnion}\ a\ {\isasymLongrightarrow}\ x\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtheta}\ p\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}p\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ elem{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}val{\isacharunderscore}{\kern0pt}pair\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isasymin}P{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ filterD\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}p{\isasymrangle}\ {\isasymin}\ Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymsigma}\ q\ r\ \isakeyword{where}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}{\isasymlangle}{\isasymsigma}{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ {\isasymtau}\ {\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}r{\isasymrangle}\ {\isasymin}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}q{\isasymin}P{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}p{\isacharcomma}{\kern0pt}r{\isasymrangle}\ {\isasymin}\ leq{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymlangle}p{\isacharcomma}{\kern0pt}q{\isasymrangle}\ {\isasymin}\ leq{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}body{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}p{\isasymin}G{\isacartoucheclose}\ {\isacartoucheopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r\ {\isasymin}\ G{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymin}\ G{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ filter{\isacharunderscore}{\kern0pt}leqD\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymlangle}{\isasymtheta}{\isacharcomma}{\kern0pt}r{\isasymrangle}\ {\isasymin}\ {\isasymsigma}{\isacartoucheclose}\ {\isacartoucheopen}{\isasymlangle}{\isasymsigma}{\isacharcomma}{\kern0pt}q{\isasymrangle}{\isasymin}{\isasymtau}{\isacartoucheclose}\ {\isacartoucheopen}q{\isasymin}P{\isacartoucheclose}\ {\isacartoucheopen}r{\isasymin}P{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymsigma}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ val{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}elem\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharplus}{\kern0pt}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isasymUnion}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtheta}{\isacharparenright}{\kern0pt}{\isacharequal}{\kern0pt}x{\isacartoucheclose}\ {\isacartoucheopen}a{\isacharequal}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{have}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}x\ {\isasymin}\ {\isasymUnion}\ a{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}a{\isacharequal}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ {\isasymUnion}\ a{\isachardoublequoteclose}\ \isakeyword{for}\ x\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ {\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ \isakeyword{assumes}\ {\isachardoublequoteopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{{\isacharbraceleft}{\kern0pt}}\isamarkupfalse%
\ \isacommand{fix}\isamarkupfalse%
\ a\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ mgtrans\ {\isacharcolon}{\kern0pt}\ M{\isacharunderscore}{\kern0pt}trans\ {\isachardoublequoteopen}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ transitivity{\isacharunderscore}{\kern0pt}MG\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}unfold{\isacharunderscore}{\kern0pt}locales{\isacharsemicolon}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymtau}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}a{\isacharequal}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ GenExtD\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}name{\isacharparenleft}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}{\isasympi}\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}M\ \isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}name{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}{\isasympi}{\isacharparenright}{\kern0pt}\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ {\isacharparenleft}{\kern0pt}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}{\kern0pt}U\ {\isasymin}\ {\isacharunderscore}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ GenExtI\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}a{\isasymin}{\isacharunderscore}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isacharquery}{\kern0pt}U{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isasymtau}\ {\isasymin}\ M{\isacartoucheclose}\ {\isacartoucheopen}filter{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharquery}{\kern0pt}U\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}a{\isacharequal}{\kern0pt}val{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}{\isasymtau}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}big{\isacharunderscore}{\kern0pt}union{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}{\isacharquery}{\kern0pt}U{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}MG{\isacharunderscore}{\kern0pt}Eq\ Union{\isacharunderscore}{\kern0pt}abs\ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacartoucheopen}{\isacharquery}{\kern0pt}U\ {\isasymin}\ M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}z{\isacharbrackleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardot}{\kern0pt}\ big{\isacharunderscore}{\kern0pt}union{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharcomma}{\kern0pt}a{\isacharcomma}{\kern0pt}z{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{{\isacharbraceright}{\kern0pt}}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Union{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}ax{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{theorem}\isamarkupfalse%
\ Union{\isacharunderscore}{\kern0pt}MG\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Union{\isacharunderscore}{\kern0pt}ax{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}M{\isacharbrackleft}{\kern0pt}G{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp\ add{\isacharcolon}{\kern0pt}M{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}MG{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{end}\isamarkupfalse%
\ \isanewline
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Forcing/Union_Axiom.thy%:%
%:%11=1%:%
%:%27=2%:%
%:%28=2%:%
%:%29=3%:%
%:%30=4%:%
%:%35=4%:%
%:%38=5%:%
%:%39=6%:%
%:%40=6%:%
%:%41=7%:%
%:%42=8%:%
%:%43=9%:%
%:%44=10%:%
%:%45=10%:%
%:%46=11%:%
%:%48=13%:%
%:%49=14%:%
%:%50=15%:%
%:%51=15%:%
%:%52=16%:%
%:%61=25%:%
%:%62=26%:%
%:%63=27%:%
%:%64=27%:%
%:%65=28%:%
%:%68=29%:%
%:%72=29%:%
%:%73=29%:%
%:%74=29%:%
%:%79=29%:%
%:%82=30%:%
%:%83=31%:%
%:%84=32%:%
%:%85=32%:%
%:%86=33%:%
%:%89=34%:%
%:%93=34%:%
%:%94=34%:%
%:%95=35%:%
%:%96=35%:%
%:%101=35%:%
%:%104=36%:%
%:%105=37%:%
%:%106=37%:%
%:%107=38%:%
%:%109=40%:%
%:%112=41%:%
%:%116=41%:%
%:%117=41%:%
%:%118=42%:%
%:%119=42%:%
%:%124=42%:%
%:%127=43%:%
%:%128=44%:%
%:%129=45%:%
%:%130=45%:%
%:%131=46%:%
%:%132=47%:%
%:%135=48%:%
%:%139=48%:%
%:%140=48%:%
%:%141=49%:%
%:%142=49%:%
%:%147=49%:%
%:%150=50%:%
%:%151=51%:%
%:%152=52%:%
%:%153=52%:%
%:%154=53%:%
%:%155=54%:%
%:%156=55%:%
%:%157=56%:%
%:%158=56%:%
%:%159=57%:%
%:%162=58%:%
%:%166=58%:%
%:%167=58%:%
%:%168=59%:%
%:%169=59%:%
%:%170=60%:%
%:%171=60%:%
%:%172=61%:%
%:%173=61%:%
%:%174=62%:%
%:%175=62%:%
%:%176=63%:%
%:%177=63%:%
%:%178=63%:%
%:%179=63%:%
%:%180=64%:%
%:%181=64%:%
%:%182=65%:%
%:%183=65%:%
%:%184=65%:%
%:%185=65%:%
%:%186=66%:%
%:%187=66%:%
%:%188=66%:%
%:%189=66%:%
%:%190=67%:%
%:%191=67%:%
%:%192=68%:%
%:%193=68%:%
%:%194=68%:%
%:%195=69%:%
%:%196=69%:%
%:%197=70%:%
%:%198=70%:%
%:%199=71%:%
%:%200=71%:%
%:%201=71%:%
%:%202=72%:%
%:%203=72%:%
%:%204=73%:%
%:%205=73%:%
%:%206=74%:%
%:%207=74%:%
%:%208=74%:%
%:%209=75%:%
%:%210=75%:%
%:%211=76%:%
%:%212=76%:%
%:%213=77%:%
%:%214=77%:%
%:%215=78%:%
%:%216=78%:%
%:%217=78%:%
%:%218=78%:%
%:%219=79%:%
%:%220=79%:%
%:%221=80%:%
%:%222=80%:%
%:%223=81%:%
%:%224=81%:%
%:%225=81%:%
%:%226=81%:%
%:%227=82%:%
%:%228=82%:%
%:%229=83%:%
%:%230=83%:%
%:%231=84%:%
%:%232=84%:%
%:%233=84%:%
%:%234=85%:%
%:%235=85%:%
%:%236=86%:%
%:%237=86%:%
%:%238=86%:%
%:%239=86%:%
%:%240=87%:%
%:%241=87%:%
%:%242=88%:%
%:%243=88%:%
%:%244=88%:%
%:%245=88%:%
%:%246=88%:%
%:%247=89%:%
%:%253=89%:%
%:%256=90%:%
%:%257=91%:%
%:%258=92%:%
%:%259=93%:%
%:%260=93%:%
%:%261=94%:%
%:%262=95%:%
%:%269=96%:%
%:%270=96%:%
%:%271=97%:%
%:%272=97%:%
%:%273=98%:%
%:%274=98%:%
%:%275=99%:%
%:%276=99%:%
%:%277=100%:%
%:%278=100%:%
%:%279=100%:%
%:%280=100%:%
%:%281=101%:%
%:%282=101%:%
%:%283=101%:%
%:%284=102%:%
%:%285=103%:%
%:%286=103%:%
%:%287=103%:%
%:%288=104%:%
%:%289=104%:%
%:%290=104%:%
%:%291=105%:%
%:%292=106%:%
%:%293=106%:%
%:%294=106%:%
%:%295=107%:%
%:%296=107%:%
%:%297=107%:%
%:%298=107%:%
%:%299=108%:%
%:%300=108%:%
%:%301=108%:%
%:%302=109%:%
%:%303=110%:%
%:%304=110%:%
%:%305=110%:%
%:%306=111%:%
%:%307=111%:%
%:%308=111%:%
%:%309=112%:%
%:%310=112%:%
%:%311=112%:%
%:%312=112%:%
%:%313=113%:%
%:%314=113%:%
%:%315=113%:%
%:%316=114%:%
%:%317=114%:%
%:%318=115%:%
%:%319=115%:%
%:%320=116%:%
%:%321=116%:%
%:%322=116%:%
%:%323=117%:%
%:%324=117%:%
%:%325=117%:%
%:%326=118%:%
%:%327=118%:%
%:%328=118%:%
%:%329=118%:%
%:%330=119%:%
%:%331=119%:%
%:%332=120%:%
%:%333=120%:%
%:%334=120%:%
%:%335=120%:%
%:%336=121%:%
%:%337=121%:%
%:%338=122%:%
%:%339=122%:%
%:%340=123%:%
%:%341=123%:%
%:%342=124%:%
%:%343=124%:%
%:%344=124%:%
%:%345=125%:%
%:%346=126%:%
%:%347=126%:%
%:%348=126%:%
%:%349=127%:%
%:%350=127%:%
%:%351=127%:%
%:%352=127%:%
%:%353=127%:%
%:%354=128%:%
%:%355=128%:%
%:%356=128%:%
%:%357=129%:%
%:%358=130%:%
%:%359=130%:%
%:%360=130%:%
%:%361=131%:%
%:%362=131%:%
%:%363=131%:%
%:%364=132%:%
%:%365=132%:%
%:%366=132%:%
%:%367=133%:%
%:%368=133%:%
%:%369=133%:%
%:%370=134%:%
%:%371=135%:%
%:%372=135%:%
%:%373=135%:%
%:%374=136%:%
%:%375=136%:%
%:%376=136%:%
%:%377=136%:%
%:%378=137%:%
%:%379=137%:%
%:%380=137%:%
%:%381=138%:%
%:%382=138%:%
%:%383=139%:%
%:%384=139%:%
%:%385=140%:%
%:%386=140%:%
%:%387=141%:%
%:%388=141%:%
%:%389=141%:%
%:%390=142%:%
%:%391=142%:%
%:%392=143%:%
%:%393=143%:%
%:%394=143%:%
%:%395=143%:%
%:%396=144%:%
%:%402=144%:%
%:%405=145%:%
%:%406=146%:%
%:%407=146%:%
%:%408=147%:%
%:%415=148%:%
%:%416=148%:%
%:%417=149%:%
%:%418=149%:%
%:%419=149%:%
%:%420=150%:%
%:%421=150%:%
%:%422=151%:%
%:%423=151%:%
%:%424=152%:%
%:%425=152%:%
%:%426=153%:%
%:%427=153%:%
%:%428=153%:%
%:%429=154%:%
%:%430=154%:%
%:%431=154%:%
%:%432=154%:%
%:%433=154%:%
%:%434=155%:%
%:%435=155%:%
%:%436=156%:%
%:%437=156%:%
%:%438=156%:%
%:%439=156%:%
%:%440=156%:%
%:%441=157%:%
%:%442=157%:%
%:%443=158%:%
%:%444=158%:%
%:%445=158%:%
%:%446=158%:%
%:%447=159%:%
%:%448=159%:%
%:%449=160%:%
%:%450=160%:%
%:%451=160%:%
%:%452=161%:%
%:%453=161%:%
%:%454=162%:%
%:%455=162%:%
%:%456=163%:%
%:%457=163%:%
%:%458=163%:%
%:%459=164%:%
%:%460=164%:%
%:%461=165%:%
%:%462=165%:%
%:%463=165%:%
%:%464=166%:%
%:%465=166%:%
%:%466=167%:%
%:%467=167%:%
%:%468=168%:%
%:%469=168%:%
%:%470=168%:%
%:%471=168%:%
%:%472=169%:%
%:%473=169%:%
%:%474=170%:%
%:%475=170%:%
%:%476=170%:%
%:%477=171%:%
%:%483=171%:%
%:%486=172%:%
%:%487=173%:%
%:%488=173%:%
%:%491=174%:%
%:%495=174%:%
%:%496=174%:%
%:%501=174%:%
%:%504=175%:%
%:%505=176%:%
%:%506=176%:%
%:%513=177%:%