### 2024-11-01 code/output/document/Utilities.tex

%
\begin{isabellebody}%
\setisabellecontext{Utilities}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Utilities\ \isanewline
\ \ \isakeyword{imports}\ \isanewline
\ \ \ \ ZF\ \isanewline
\ \ \ \ {\isachardoublequoteopen}Forcing{\isacharslash}{\kern0pt}Forcing{\isacharunderscore}{\kern0pt}Main{\isachardoublequoteclose}\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\ \isanewline
%
\endisadelimtheory
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}value\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ D\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ D\ {\isacharbraceright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ D{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ D\ {\isacharbraceright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ D\ {\isacharbraceright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}{\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ a\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ equal{\isacharunderscore}{\kern0pt}singleton{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ D\ {\isacharbraceright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ D\ {\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ D\ {\isacharbraceright}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}Base{\isachardot}{\kern0pt}apply{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharbraceleft}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}{\isachardot}{\kern0pt}\ y\ {\isasymin}\ D\ {\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}value{\isacharunderscore}{\kern0pt}in\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ A\ {\isasymrightarrow}\ B\ {\isasymLongrightarrow}\ a\ {\isasymin}\ A\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}a\ {\isasymin}\ B{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}domran\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ C\ {\isasymLongrightarrow}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ range{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ C{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ abh{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ adom\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isasymin}\ range{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ rangeI\ abh\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ range{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ adom\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ \ domain{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ range{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ abh\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C\ {\isacharbraceright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}x{\isachardot}{\kern0pt}\ {\isasymexists}y{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ u\ \isakeyword{where}\ i{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}u\ {\isasymin}\ C{\isachardoublequoteclose}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ split{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ u{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ h\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}fst{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ snd{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ split{\isacharunderscore}{\kern0pt}def\ i\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}fst{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ snd{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ u{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}u\ \isakeyword{and}\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}x{\isachardot}{\kern0pt}\ range{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ pair{\isachardot}{\kern0pt}Pair{\isacharunderscore}{\kern0pt}fst{\isacharunderscore}{\kern0pt}snd{\isacharunderscore}{\kern0pt}eq{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ relation{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}domran{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ i\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}fst{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isacharcomma}{\kern0pt}\ snd{\isacharparenleft}{\kern0pt}u{\isacharparenright}{\kern0pt}{\isachargreater}{\kern0pt}\ {\isasymin}\ C{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ i\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ h\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}relI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ F{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isacharparenright}{\kern0pt}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ eq{\isacharunderscore}{\kern0pt}flip\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isacharequal}{\kern0pt}\ B\ {\isasymLongrightarrow}\ B\ {\isacharequal}{\kern0pt}\ A{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}\isanewline
\ \ \ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x{\isachardot}{\kern0pt}\ {\isasymforall}y{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C\ {\isasymlongrightarrow}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ {\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C\ {\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharsemicolon}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}{\kern0pt}\ {\isasymforall}y{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C\ {\isasymlongrightarrow}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ C{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}x\ y{\isachardot}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymin}\ C\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharsemicolon}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ y\ \isakeyword{where}\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C{\isachardoublequoteclose}\ {\isachardoublequoteopen}v{\isacharequal}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ split{\isacharparenleft}{\kern0pt}G{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ C{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ p{\isadigit{2}}\ p{\isadigit{3}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \ \ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}{\kern0pt}\ {\isasymforall}y{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C\ {\isasymlongrightarrow}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ C{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}x\ y{\isachardot}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}\ y{\isasymrangle}\ {\isasymin}\ C\ {\isasymand}\ v\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg{\isacharsemicolon}{\kern0pt}\ auto{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ y\ \isakeyword{where}\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ C{\isachardoublequoteclose}\ {\isachardoublequoteopen}v{\isacharequal}{\kern0pt}G{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}rel{\isacharunderscore}{\kern0pt}arg\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ split{\isacharparenleft}{\kern0pt}F{\isacharcomma}{\kern0pt}\ {\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ {\isasymlangle}x{\isacharcomma}{\kern0pt}y{\isasymrangle}\ {\isasymin}\ C{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}y{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ p{\isadigit{2}}\ p{\isadigit{3}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ OUN{\isacharunderscore}{\kern0pt}UN{\isacharunderscore}{\kern0pt}equals\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isasymin}\ a\ {\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharsemicolon}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bp\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ OUN{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isasymin}\ a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ ltD\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}\ b\ {\isasymin}\ a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isasymin}\ a\ {\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ UN{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isasymin}\ a\ {\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ b\ \isakeyword{where}\ bp\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}b\ {\isasymin}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ UN{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isacharless}{\kern0pt}\ a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}b\ {\isasymin}\ a{\isacartoucheclose}\ {\isacartoucheopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ ltI\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ x\ {\isasymin}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}\ b\ {\isacharless}{\kern0pt}\ a{\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}lt\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ b\ {\isasymLongrightarrow}\ b\ {\isasymle}\ c\ {\isasymLongrightarrow}\ a\ {\isacharless}{\kern0pt}\ c{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}b\ \isakeyword{and}\ j{\isacharequal}{\kern0pt}c\ \isakeyword{in}\ leE{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ assumption\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}trans\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymle}\ b\ {\isasymLongrightarrow}\ b\ {\isacharless}{\kern0pt}\ c\ {\isasymLongrightarrow}\ a\ {\isacharless}{\kern0pt}\ c{\isachardoublequoteclose}\ \ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}a\ \isakeyword{and}\ j{\isacharequal}{\kern0pt}b\ \isakeyword{in}\ leE{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ assumption\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}trans\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}le\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ a\ {\isasymle}\ b\ {\isasymLongrightarrow}\ a\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}b\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}refl\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}lt\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ a\ {\isacharless}{\kern0pt}\ b\ {\isasymLongrightarrow}\ a\ {\isacharless}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}b\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}refl\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ex{\isacharunderscore}{\kern0pt}larger{\isacharunderscore}{\kern0pt}limit\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}b{\isachardot}{\kern0pt}\ a\ {\isacharless}{\kern0pt}\ b\ {\isasymand}\ Limit{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Card{\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assm\ csucc{\isacharunderscore}{\kern0pt}basic\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ cadd{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{3}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ Card{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Ord\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{4}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{5}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Card{\isacharparenleft}{\kern0pt}nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}subset{\isacharunderscore}{\kern0pt}into{\isacharunderscore}{\kern0pt}Card\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{6}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}nat\ {\isasymle}\ {\isacharparenleft}{\kern0pt}nat\ {\isasymoplus}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{3}}\ cadd{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}self\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{7}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}nat\ {\isasymle}\ {\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ cadd{\isacharunderscore}{\kern0pt}commute\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{8}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Card{\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ cadd{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{9}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}InfCard{\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ InfCard{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{7}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}{\isadigit{0}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Limit{\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ InfCard{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}Limit\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ {\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ Q\ {\isacharequal}{\kern0pt}\ {\isachardoublequoteopen}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ disjE{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat\ {\isasymor}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}iff\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}trans\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ csucc{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymoplus}\ nat{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ p{\isadigit{1}}{\isadigit{0}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ UN{\isacharunderscore}{\kern0pt}RepFunI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ b\ {\isasymLongrightarrow}\ x\ {\isasymin}\ F{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}a\ {\isasymin}\ b{\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ OUN{\isacharunderscore}{\kern0pt}RepFunI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isacharless}{\kern0pt}\ b\ {\isasymLongrightarrow}\ x\ {\isasymin}\ F{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ {\isacharparenleft}{\kern0pt}{\isasymUnion}a\ {\isacharless}{\kern0pt}\ b{\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ RepFun{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}y\ {\isasymin}\ B{\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ {\isasymforall}y\ {\isasymin}\ B{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ B{\isacharbraceright}{\kern0pt}\ {\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharsemicolon}{\kern0pt}\ rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ A{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}x{\isasymin}A{\isachardot}{\kern0pt}\ {\isasymexists}y{\isasymin}B{\isachardot}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ vp{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ yp{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}F{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ B{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ vp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}G{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ B{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{next}\isamarkupfalse%
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ v\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}G{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ x\ {\isasymin}\ B{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}x{\isasymin}B{\isachardot}{\kern0pt}\ {\isasymexists}y{\isasymin}A{\isachardot}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ x\ \isakeyword{where}\ vp{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ G{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ B{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ yp{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}G{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}y\ {\isasymin}\ A{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ F{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ vp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}v\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}F{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isachardot}{\kern0pt}\ y\ {\isasymin}\ A{\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yp\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}ball{\isacharunderscore}{\kern0pt}mp\ {\isacharcolon}{\kern0pt}\isanewline
\ \ {\isachardoublequoteopen}C\ {\isasymsubseteq}\ D\ {\isasymtimes}\ E\ {\isasymLongrightarrow}\ {\isasymforall}{\isacharless}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isachargreater}{\kern0pt}\ {\isasymin}\ C{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\isanewline
\ \ \ {\isasymLongrightarrow}\ {\isasymforall}d\ {\isasymin}\ D{\isachardot}{\kern0pt}\ {\isasymforall}e\ {\isasymin}\ E{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}{\isacharless}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isachargreater}{\kern0pt}\ {\isasymin}\ C{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}{\isasymlangle}d{\isacharcomma}{\kern0pt}e{\isasymrangle}{\isasymin}C{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}\ {\isasymforall}d{\isasymin}D{\isachardot}{\kern0pt}\ {\isasymforall}e{\isasymin}E{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ C{\isachardoublequoteclose}\ {\isachardoublequoteopen}C\ {\isasymsubseteq}\ D\ {\isasymtimes}\ E{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ d\ e\ \isakeyword{where}\ de\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ {\isasymin}\ D{\isachardoublequoteclose}\ {\isachardoublequoteopen}e\ {\isasymin}\ E{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ de\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymlambda}{\isasymlangle}d{\isacharcomma}{\kern0pt}e{\isasymrangle}{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ de\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}ball{\isacharunderscore}{\kern0pt}conj\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}C\ {\isasymsubseteq}\ D\ {\isasymtimes}\ E\ {\isasymLongrightarrow}\ {\isasymforall}{\isacharless}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isachargreater}{\kern0pt}\ {\isasymin}\ C{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}{\isacharless}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isachargreater}{\kern0pt}\ {\isasymin}\ C{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}{\isacharless}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isachargreater}{\kern0pt}\ {\isasymin}\ C{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymforall}{\isasymlangle}d{\isacharcomma}{\kern0pt}e{\isasymrangle}{\isasymin}C{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}{\isasymlangle}d{\isacharcomma}{\kern0pt}e{\isasymrangle}{\isasymin}C{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ C{\isachardoublequoteclose}\ {\isachardoublequoteopen}C\ {\isasymsubseteq}\ D\ {\isasymtimes}\ E{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ d\ e\ \isakeyword{where}\ de\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ {\isasymin}\ D{\isachardoublequoteclose}\ {\isachardoublequoteopen}e\ {\isasymin}\ E{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ p{\isadigit{2}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ de\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ {\isacharparenleft}{\kern0pt}{\isasymlambda}{\isasymlangle}d{\isacharcomma}{\kern0pt}e{\isasymrangle}{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}\ {\isasymand}\ Q{\isacharparenleft}{\kern0pt}d{\isacharcomma}{\kern0pt}\ e{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ de\ p{\isadigit{1}}\ p{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ inter{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isacharequal}{\kern0pt}\ B\ {\isasymLongrightarrow}\ A\ {\isasyminter}\ C\ {\isacharequal}{\kern0pt}\ B\ {\isasyminter}\ C{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}flip\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}A\ {\isasymlongleftrightarrow}\ B\ {\isasymLongrightarrow}\ B\ {\isasymlongleftrightarrow}\ A{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ empty{\isacharunderscore}{\kern0pt}or{\isacharunderscore}{\kern0pt}not\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}\ {\isasymlongrightarrow}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymlongrightarrow}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}nat{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ lt{\isacharunderscore}{\kern0pt}Ord\ ltI\ Ord{\isacharunderscore}{\kern0pt}nat\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ in{\isacharunderscore}{\kern0pt}dom{\isacharunderscore}{\kern0pt}or{\isacharunderscore}{\kern0pt}not\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ F\ x{\isachardot}{\kern0pt}\ function{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}F{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}F{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \isacommand{fix}\isamarkupfalse%
\ P\ F\ x\ \isanewline
\ \isacommand{assume}\isamarkupfalse%
\ assm\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}P{\isacharparenleft}{\kern0pt}{\isadigit{0}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}F{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharparenleft}{\kern0pt}F{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}F{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}F{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}def\ ZF{\isacharunderscore}{\kern0pt}Base{\isachardot}{\kern0pt}image{\isacharunderscore}{\kern0pt}def\ domain{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P{\isacharparenleft}{\kern0pt}F{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assm\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}forallI\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymAnd}A\ x{\isachardot}{\kern0pt}\ relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardot}{\kern0pt}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}clarify{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ A\ x\ v\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}{\isachardoublequoteopen}v\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ p\ \isakeyword{where}\ yph\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}v\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymlambda}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardot}{\kern0pt}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ yph\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}forallD\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymAnd}A\ x{\isachardot}{\kern0pt}\ relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardot}{\kern0pt}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ A\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}v{\isasymin}x{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymlambda}{\isasymlangle}y{\isacharcomma}{\kern0pt}p{\isasymrangle}{\isachardot}{\kern0pt}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenleft}{\kern0pt}v{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isasymlangle}y{\isacharcomma}{\kern0pt}\ p{\isasymrangle}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ y\ p\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}{\kern0pt}forallE\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymAnd}A\ Q\ x{\isachardot}{\kern0pt}\ relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymforall}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardot}{\kern0pt}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Q{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Q{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ A\ Q\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardot}{\kern0pt}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ r{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymAnd}y\ p{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x\ {\isasymLongrightarrow}\ A{\isacharparenleft}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Q{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}Q{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ r{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ pair{\isacharunderscore}{\kern0pt}forallD{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ converse{\isacharunderscore}{\kern0pt}apply\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ B{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ A{\isachardoublequoteclose}\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}A\ {\isasymsubseteq}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ \isacommand{unfolding}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x{\isachargreater}{\kern0pt}\ {\isasymin}\ f{\isachardoublequoteclose}\ \ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}f{\isacharbackquote}{\kern0pt}x{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ converseI{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}B{\isacharcomma}{\kern0pt}\ A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}bij\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymin}\ B\ {\isasymrightarrow}\ A{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}\ x{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}fun\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ image{\isacharunderscore}{\kern0pt}converse{\isacharunderscore}{\kern0pt}image\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ B{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ C\ {\isasymsubseteq}\ A\ {\isasymLongrightarrow}\ converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ C{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ C{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ B{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ {\isachardoublequoteopen}C{\isasymsubseteq}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}f\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ y\ \isakeyword{where}\ yh\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}C{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ x{\isachargreater}{\kern0pt}\ {\isasymin}\ converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ image{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}x{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ f{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ converseE\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ A\ {\isasymand}\ f{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ B{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ apply{\isacharunderscore}{\kern0pt}fun{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ z\ \isakeyword{where}\ zh{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z\ {\isasymin}\ C{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}z{\isacharcomma}{\kern0pt}\ y{\isachargreater}{\kern0pt}\ {\isasymin}\ f{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ yh\ image{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}z\ {\isasymin}\ A\ {\isasymand}\ f{\isacharbackquote}{\kern0pt}z\ {\isacharequal}{\kern0pt}\ y{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ B{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ apply{\isacharunderscore}{\kern0pt}fun{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isasymin}\ inj{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}inj\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x{\isacharequal}{\kern0pt}z{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ inj{\isacharunderscore}{\kern0pt}def\ \isacommand{using}\isamarkupfalse%
\ H\ H{\isadigit{2}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ C{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ zh\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ B{\isacharparenright}{\kern0pt}\ {\isachardoublequoteclose}\ {\isachardoublequoteopen}C{\isasymsubseteq}A{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ C{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x\ {\isasymin}\ f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}C{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}x\ \isakeyword{in}\ imageI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ \isacommand{unfolding}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isacommand{apply}\isamarkupfalse%
\ blast\ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ converse{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharparenleft}{\kern0pt}f{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}C{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}f{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \isakeyword{in}\ imageI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ converseI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ bij{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}fun\ assms\ \isacommand{unfolding}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isacommand{apply}\isamarkupfalse%
\ blast\ \isacommand{done}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ifT{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymnoteq}\ b\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}if\ P\ then\ a\ else\ b{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a\ {\isasymLongrightarrow}\ P{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymnot}P{\isachardoublequoteclose}\ \isakeyword{in}\ disjE{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ifF{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymnoteq}\ b\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}if\ P\ then\ a\ else\ b{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ b\ {\isasymLongrightarrow}\ {\isasymnot}P{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}P\ \isakeyword{and}\ Q{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymnot}P{\isachardoublequoteclose}\ \isakeyword{in}\ disjE{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ neq{\isacharunderscore}{\kern0pt}flip\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymnoteq}\ b\ {\isasymLongrightarrow}\ b\ {\isasymnoteq}\ a{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A\ P\ Q{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}\ x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isacharbraceright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ ex{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A\ P\ Q{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ bex{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A\ P\ Q{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ all{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A\ P\ Q{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ ball{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A\ P\ Q{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ A\ {\isasymLongrightarrow}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ P{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymforall}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ Q{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}disjI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q\ R\ S{\isachardot}{\kern0pt}\ P\ {\isasymlongleftrightarrow}\ Q\ {\isasymLongrightarrow}\ R\ {\isasymlongleftrightarrow}\ S\ {\isasymLongrightarrow}\ P\ {\isasymor}\ R\ {\isasymlongleftrightarrow}\ Q\ {\isasymor}\ S{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}conjI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q\ R\ S{\isachardot}{\kern0pt}\ P\ {\isasymlongleftrightarrow}\ Q\ {\isasymLongrightarrow}\ R\ {\isasymlongleftrightarrow}\ S\ {\isasymLongrightarrow}\ P\ {\isasymand}\ R\ {\isasymlongleftrightarrow}\ Q\ {\isasymand}\ S{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}conjI{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q\ R\ S{\isachardot}{\kern0pt}\ P\ {\isasymlongleftrightarrow}\ Q\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}Q\ {\isasymLongrightarrow}\ R\ {\isasymlongleftrightarrow}\ S{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}P\ {\isasymand}\ R\ {\isasymlongleftrightarrow}\ Q\ {\isasymand}\ S{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ iff{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q\ R\ S{\isachardot}{\kern0pt}\ P\ {\isasymlongleftrightarrow}\ Q\ {\isasymLongrightarrow}\ R\ {\isasymlongleftrightarrow}\ S\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}P\ {\isasymlongleftrightarrow}\ R{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}Q\ {\isasymlongleftrightarrow}\ S{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ imp{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q\ R\ S{\isachardot}{\kern0pt}\ P\ {\isasymlongleftrightarrow}\ Q\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}R\ {\isasymlongleftrightarrow}\ S{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}P\ {\isasymlongrightarrow}\ R{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}Q\ {\isasymlongrightarrow}\ S{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ imp{\isacharunderscore}{\kern0pt}iff{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q\ R\ S{\isachardot}{\kern0pt}\ P\ {\isasymlongleftrightarrow}\ Q\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}Q\ {\isasymLongrightarrow}\ R\ {\isasymlongleftrightarrow}\ S{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}P\ {\isasymlongrightarrow}\ R{\isacharparenright}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}Q\ {\isasymlongrightarrow}\ S{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ notnot{\isacharunderscore}{\kern0pt}iff\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}P\ Q{\isachardot}{\kern0pt}\ P\ {\isasymlongleftrightarrow}\ Q\ {\isasymLongrightarrow}\ {\isasymnot}P\ {\isasymlongleftrightarrow}\ {\isasymnot}Q{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}le{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ a\ {\isasymle}\ a\ {\isasymunion}\ b{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}iff\ le{\isacharunderscore}{\kern0pt}refl\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ b\ {\isasymle}\ a\ {\isasymunion}\ b{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}iff\ le{\isacharunderscore}{\kern0pt}refl\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ zero{\isacharunderscore}{\kern0pt}le\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}i{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isadigit{0}}\ {\isasymle}\ i{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ mp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymlambda}i{\isachardot}{\kern0pt}\ Ord{\isacharparenleft}{\kern0pt}i{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isadigit{0}}\ {\isasymle}\ i{\isachardoublequoteclose}\ \isakeyword{in}\ eps{\isacharunderscore}{\kern0pt}induct{\isacharcomma}{\kern0pt}\ rule\ impI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ i\ x{\isacharcomma}{\kern0pt}\ case{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}{\isasymexists}y{\isachardot}{\kern0pt}\ y\ {\isasymin}\ x{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ lt{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ i\ x\ y{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}y\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}lt{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Ord\ ltI\ \isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ auto{\isacharbrackleft}{\kern0pt}{\isadigit{2}}{\isacharbrackright}{\kern0pt}\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rename{\isacharunderscore}{\kern0pt}tac\ i\ x{\isacharcomma}{\kern0pt}\ rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}x\ \isakeyword{and}\ a{\isacharequal}{\kern0pt}{\isadigit{0}}\ \isakeyword{in}\ ssubst{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}un{\isacharunderscore}{\kern0pt}eq{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ b\ {\isasymle}\ a\ {\isasymLongrightarrow}\ a\ {\isasymunion}\ b\ {\isacharequal}{\kern0pt}\ a{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ leE{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Ord{\isacharunderscore}{\kern0pt}Un{\isacharunderscore}{\kern0pt}if{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Ord{\isacharunderscore}{\kern0pt}un{\isacharunderscore}{\kern0pt}eq{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ a\ {\isasymle}\ b\ {\isasymLongrightarrow}\ a\ {\isasymunion}\ b\ {\isacharequal}{\kern0pt}\ b{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ Un{\isacharunderscore}{\kern0pt}commute{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Ord{\isacharunderscore}{\kern0pt}un{\isacharunderscore}{\kern0pt}eq{\isadigit{1}}{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}le{\isacharprime}{\kern0pt}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ b\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ a\ {\isasymle}\ b\ {\isasymLongrightarrow}\ pred{\isacharparenleft}{\kern0pt}a{\isacharparenright}{\kern0pt}\ {\isasymle}\ pred{\isacharparenleft}{\kern0pt}b{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}b\ \isakeyword{in}\ natE{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subgoal{\isacharunderscore}{\kern0pt}tac\ {\isachardoublequoteopen}a\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ pred{\isacharunderscore}{\kern0pt}le\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}elem{\isacharunderscore}{\kern0pt}rank{\isacharunderscore}{\kern0pt}lt\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}y\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ p\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}\ {\isasymin}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}y{\isacharparenright}{\kern0pt}\ {\isacharless}{\kern0pt}\ rank{\isacharparenleft}{\kern0pt}x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}{\isachardoublequoteopen}rank{\isacharparenleft}{\kern0pt}{\isacharless}{\kern0pt}y{\isacharcomma}{\kern0pt}\ p{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ rank{\isacharunderscore}{\kern0pt}pair{\isadigit{1}}{\isacharcomma}{\kern0pt}\ rule\ rank{\isacharunderscore}{\kern0pt}lt{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ final{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}notation\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ l\ \isakeyword{assumes}\ lin\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}l\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ lnotnil\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}l\ {\isasymnoteq}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}l{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}\ a\ {\isasymin}\ A{\isachardot}{\kern0pt}\ l\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}l\ {\isasymnoteq}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}l{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}\ a\ {\isasymin}\ A{\isachardot}{\kern0pt}\ l\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lnotnil\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ lin\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ Nil\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Cons\ a\ l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}cases\ {\isachardoublequoteopen}l{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharbrackleft}{\kern0pt}{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}a\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}l{\isacharprime}{\kern0pt}a{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}a{\isasymin}A{\isachardot}{\kern0pt}\ l{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}a\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ l{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\ t\ \isakeyword{where}\ {\isachardoublequoteopen}t\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}l{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l{\isacharprime}{\kern0pt}\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}t{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}l{\isacharprime}{\kern0pt}a{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}aa{\isasymin}A{\isachardot}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}a\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}aa{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}a{\isacharcomma}{\kern0pt}\ l{\isacharprime}{\kern0pt}{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}t\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}a\ {\isasymin}\ A{\isacartoucheclose}\ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ length{\isadigit{1}}{\isacharunderscore}{\kern0pt}notation\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ l\ \isakeyword{assumes}\ {\isachardoublequoteopen}l\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}a\ {\isasymin}\ A{\isachardot}{\kern0pt}\ l\ {\isacharequal}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}l\ {\isasymnoteq}\ Nil{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ length{\isacharunderscore}{\kern0pt}is{\isacharunderscore}{\kern0pt}{\isadigit{0}}{\isacharunderscore}{\kern0pt}iff\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}l{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}\ a\ {\isasymin}\ A{\isachardot}{\kern0pt}\ l\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ final{\isacharunderscore}{\kern0pt}app{\isacharunderscore}{\kern0pt}notation\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ l{\isacharprime}{\kern0pt}\ a\ \isakeyword{where}\ H{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}l{\isacharprime}{\kern0pt}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isacharequal}{\kern0pt}\ l{\isacharprime}{\kern0pt}\ {\isacharat}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}l\ {\isacharequal}{\kern0pt}\ {\isacharbrackleft}{\kern0pt}a{\isacharbrackright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ H\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ not{\isacharunderscore}{\kern0pt}nil{\isacharunderscore}{\kern0pt}obtain{\isacharunderscore}{\kern0pt}hd{\isacharunderscore}{\kern0pt}tl\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ A\ l\ \isakeyword{assumes}\ {\isachardoublequoteopen}l\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l\ {\isasymnoteq}\ Nil{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}hd\ {\isasymin}\ A{\isachardot}{\kern0pt}\ {\isasymexists}tl\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ l\ {\isacharequal}{\kern0pt}\ Cons{\isacharparenleft}{\kern0pt}hd{\isacharcomma}{\kern0pt}\ tl{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}l\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isacommand{apply}\isamarkupfalse%
\ cases\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ append{\isacharunderscore}{\kern0pt}notation\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ l\ n\ \isakeyword{assumes}\ {\isachardoublequoteopen}l\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}l{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}l{\isadigit{2}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}l{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isasymand}\ l\ {\isacharequal}{\kern0pt}\ l{\isadigit{1}}\ {\isacharat}{\kern0pt}\ l{\isadigit{2}}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ P{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isasymforall}n\ {\isasymin}\ nat{\isachardot}{\kern0pt}\ n\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\kern0pt}\ {\isasymexists}l{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}l{\isadigit{2}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}l{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isasymand}\ l\ {\isacharequal}{\kern0pt}\ l{\isadigit{1}}\ {\isacharat}{\kern0pt}\ l{\isadigit{2}}\ {\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ mp{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ clarify\ \isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ {\isacartoucheopen}l\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}induct{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ Nil\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n\ {\isacharequal}{\kern0pt}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}Nil\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}Nil\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}Cons\ hd\ tl{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}case\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}n\ \isakeyword{in}\ natE{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ n{\isacharprime}{\kern0pt}\ \isacommand{assume}\isamarkupfalse%
\ assms{\isadigit{1}}{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isacharequal}{\kern0pt}\ succ{\isacharparenleft}{\kern0pt}n{\isacharprime}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}n\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}Cons{\isacharparenleft}{\kern0pt}hd{\isacharcomma}{\kern0pt}\ tl{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}hd\ {\isasymin}\ A\ {\isachardoublequoteclose}\ {\isachardoublequoteopen}tl\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymAnd}n{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}tl{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isasymexists}l{\isadigit{1}}{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}l{\isadigit{2}}{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}l{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isasymand}\ tl\ {\isacharequal}{\kern0pt}\ l{\isadigit{1}}\ {\isacharat}{\kern0pt}\ l{\isadigit{2}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}n{\isacharprime}{\kern0pt}\ {\isasymle}\ length{\isacharparenleft}{\kern0pt}tl{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}l{\isadigit{1}}{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}l{\isadigit{2}}{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}l{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}\ {\isasymand}\ tl\ {\isacharequal}{\kern0pt}\ l{\isadigit{1}}\ {\isacharat}{\kern0pt}\ l{\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ l{\isadigit{1}}\ l{\isadigit{2}}\ \isakeyword{where}\ H\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}l{\isadigit{1}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}l{\isadigit{2}}\ {\isasymin}\ list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}length{\isacharparenleft}{\kern0pt}l{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n{\isacharprime}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}tl\ {\isacharequal}{\kern0pt}\ l{\isadigit{1}}\ {\isacharat}{\kern0pt}\ l{\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}l{\isadigit{1}}{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymexists}l{\isadigit{2}}{\isasymin}list{\isacharparenleft}{\kern0pt}A{\isacharparenright}{\kern0pt}{\isachardot}{\kern0pt}\ length{\isacharparenleft}{\kern0pt}l{\isadigit{1}}{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ n\ {\isasymand}\ Cons{\isacharparenleft}{\kern0pt}hd{\isacharcomma}{\kern0pt}\ tl{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ l{\isadigit{1}}\ {\isacharat}{\kern0pt}\ l{\isadigit{2}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}Cons{\isacharparenleft}{\kern0pt}hd{\isacharcomma}{\kern0pt}\ l{\isadigit{1}}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}l{\isadigit{2}}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\ \isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isadigit{1}}\ H\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}memberI\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ A\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ range{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ B\ {\isasymLongrightarrow}\ f\ {\isasymin}\ Pi{\isacharparenleft}{\kern0pt}A{\isacharcomma}{\kern0pt}\ {\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ B{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffD{\isadigit{2}}{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ Pi{\isacharunderscore}{\kern0pt}iff{\isacharparenright}{\kern0pt}\ \isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ subsetI{\isacharparenright}{\kern0pt}\ \ \ \isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isasymin}\ f{\isachardoublequoteclose}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}A\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ {\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymsubseteq}\ B{\isachardoublequoteclose}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ abH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isasymin}\ A\ {\isasymand}\ b\ {\isasymin}\ B{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{apply}\isamarkupfalse%
\ auto\ \isacommand{done}\isamarkupfalse%
\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymtimes}\ B{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ abH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ domain{\isacharunderscore}{\kern0pt}Pi\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ A\ {\isasymrightarrow}\ B\ {\isasymLongrightarrow}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ A{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}helper\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}A\ B\ C\ f\ g{\isachardot}{\kern0pt}\ f\ {\isasymin}\ A\ {\isasymrightarrow}\ B\ {\isasymLongrightarrow}\ g\ {\isasymin}\ A\ {\isasymrightarrow}\ C\ {\isasymLongrightarrow}\ {\isasymforall}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}x\ {\isasymLongrightarrow}\ f\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ equality{\isacharunderscore}{\kern0pt}iffI{\isacharparenright}{\kern0pt}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\isanewline
\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ helper\ {\isacharcolon}{\kern0pt}\ \ {\isachardoublequoteopen}{\isasymAnd}A\ B\ f\ x{\isachardot}{\kern0pt}\ f\ {\isasymin}\ A\ {\isasymrightarrow}\ B\ {\isasymLongrightarrow}\ x\ {\isasymin}\ f\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}a\ {\isasymin}\ A{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachargreater}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule\ iffI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ A\ B\ f\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ A\ {\isasymrightarrow}\ B{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymin}\ f{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ b\ \isakeyword{where}\ abH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ b{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ Pi{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}b\ {\isacharequal}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ apply{\isacharunderscore}{\kern0pt}equality\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}a{\isasymin}A{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ abH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{next}\isamarkupfalse%
\ \isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ A\ B\ f\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ A\ {\isasymrightarrow}\ B{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymexists}a{\isasymin}A{\isachardot}{\kern0pt}\ x\ {\isacharequal}{\kern0pt}\ {\isasymlangle}a{\isacharcomma}{\kern0pt}\ f\ {\isacharbackquote}{\kern0pt}\ a{\isasymrangle}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ a\ \isakeyword{where}\ aH\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}a\ {\isasymin}\ A{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharequal}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachargreater}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachargreater}{\kern0pt}\ {\isasymin}\ f{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ function{\isacharunderscore}{\kern0pt}apply{\isacharunderscore}{\kern0pt}Pair{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ assms\ Pi{\isacharunderscore}{\kern0pt}iff\ aH\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymin}\ f{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ aH\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ A\ B\ C\ f\ g\ x\ \isacommand{assume}\isamarkupfalse%
\ assms\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}f\ {\isasymin}\ A\ {\isasymrightarrow}\ B{\isachardoublequoteclose}\ {\isachardoublequoteopen}g\ {\isasymin}\ A\ {\isasymrightarrow}\ C{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}x\ {\isasymin}\ A{\isachardot}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}x{\isachardoublequoteclose}\ \ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{1}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}x\ {\isasymin}\ f\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}a\ {\isasymin}\ A{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ f{\isacharbackquote}{\kern0pt}a{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ iffI{\isacharparenright}{\kern0pt}\ \isacommand{using}\isamarkupfalse%
\ helper\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{2}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymexists}a\ {\isasymin}\ A{\isachardot}{\kern0pt}\ {\isacharless}{\kern0pt}a{\isacharcomma}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}a{\isachargreater}{\kern0pt}\ {\isacharequal}{\kern0pt}\ x{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ H{\isadigit{3}}{\isacharcolon}{\kern0pt}{\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ g{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ helper\ assms\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}\ x\ {\isasymin}\ f\ {\isasymlongleftrightarrow}\ x\ {\isasymin}\ g\ {\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ H{\isadigit{1}}\ H{\isadigit{2}}\ H{\isadigit{3}}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ function{\isacharunderscore}{\kern0pt}eq\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}{\isasymAnd}f\ g{\isachardot}{\kern0pt}\ relation{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ relation{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ function{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ domain{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ {\isacharparenleft}{\kern0pt}{\isasymAnd}x{\isachardot}{\kern0pt}\ x\ {\isasymin}\ domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ f{\isacharbackquote}{\kern0pt}x\ {\isacharequal}{\kern0pt}\ g{\isacharbackquote}{\kern0pt}x{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ f\ {\isacharequal}{\kern0pt}\ g{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ A{\isacharequal}{\kern0pt}{\isachardoublequoteopen}domain{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ B{\isacharequal}{\kern0pt}{\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ C{\isacharequal}{\kern0pt}{\isachardoublequoteopen}range{\isacharparenleft}{\kern0pt}g{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ function{\isacharunderscore}{\kern0pt}eq{\isacharunderscore}{\kern0pt}helper{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ Pi{\isacharunderscore}{\kern0pt}memberI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ simp{\isacharunderscore}{\kern0pt}all\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ n\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ lt{\isacharunderscore}{\kern0pt}nat{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ rule\ ltI{\isacharcomma}{\kern0pt}\ simp{\isacharunderscore}{\kern0pt}all{\isacharparenright}{\kern0pt}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}nat\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ m\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymunion}\ m\ {\isasymin}\ nat{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}lt{\isadigit{1}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ n\ {\isacharless}{\kern0pt}\ m\ {\isasymLongrightarrow}\ n\ {\isacharless}{\kern0pt}\ m\ {\isasymunion}\ l{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}m\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}le{\isadigit{1}}\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ union{\isacharunderscore}{\kern0pt}lt{\isadigit{2}}\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}m{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ Ord{\isacharparenleft}{\kern0pt}l{\isacharparenright}{\kern0pt}\ {\isasymLongrightarrow}\ n\ {\isacharless}{\kern0pt}\ l\ {\isasymLongrightarrow}\ n\ {\isacharless}{\kern0pt}\ m\ {\isasymunion}\ l{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ b{\isacharequal}{\kern0pt}l\ \isakeyword{in}\ lt{\isacharunderscore}{\kern0pt}le{\isacharunderscore}{\kern0pt}lt{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{using}\isamarkupfalse%
\ max{\isacharunderscore}{\kern0pt}le{\isadigit{2}}\ \isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ add{\isacharunderscore}{\kern0pt}diff{\isacharunderscore}{\kern0pt}swap\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{fixes}\ a\ b\ c\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}a\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}b\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}c\ {\isasymin}\ nat{\isachardoublequoteclose}\ {\isachardoublequoteopen}c\ {\isasymle}\ b{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}a\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ b\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ c\ {\isacharequal}{\kern0pt}\ a\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ c{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}c\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharparenleft}{\kern0pt}a\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ b\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ c{\isacharparenright}{\kern0pt}\ {\isacharequal}{\kern0pt}\ {\isacharparenleft}{\kern0pt}a\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ b\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ c{\isacharparenright}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ c{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ b{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ add{\isacharunderscore}{\kern0pt}diff{\isacharunderscore}{\kern0pt}inverse{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ j{\isacharequal}{\kern0pt}b\ \isakeyword{in}\ le{\isacharunderscore}{\kern0pt}trans{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ a\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isacharparenleft}{\kern0pt}b\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ c{\isacharparenright}{\kern0pt}\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ c{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}a\ \isakeyword{and}\ j\ {\isacharequal}{\kern0pt}\ a\ \isakeyword{in}\ add{\isacharunderscore}{\kern0pt}left{\isacharunderscore}{\kern0pt}cancel{\isacharcomma}{\kern0pt}\ simp{\isacharcomma}{\kern0pt}\ simp\ add{\isacharcolon}{\kern0pt}assms{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}subst\ add{\isacharunderscore}{\kern0pt}diff{\isacharunderscore}{\kern0pt}inverse{\isadigit{2}}{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}\ {\isacharequal}{\kern0pt}\ c\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharparenleft}{\kern0pt}a\ {\isacharhash}{\kern0pt}{\isacharplus}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b\ {\isacharhash}{\kern0pt}{\isacharminus}{\kern0pt}\ c{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ i{\isacharequal}{\kern0pt}c\ \isakeyword{and}\ j{\isacharequal}{\kern0pt}c\ \isakeyword{in}\ add{\isacharunderscore}{\kern0pt}left{\isacharunderscore}{\kern0pt}cancel{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ le{\isacharunderscore}{\kern0pt}succ{\isacharunderscore}{\kern0pt}pred\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ {\isachardoublequoteopen}{\isasymAnd}n{\isachardot}{\kern0pt}\ n\ {\isasymin}\ nat\ {\isasymLongrightarrow}\ n\ {\isasymle}\ succ{\isacharparenleft}{\kern0pt}pred{\isacharparenleft}{\kern0pt}n{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ n{\isacharequal}{\kern0pt}n\ \isakeyword{in}\ natE{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ relation{\isacharunderscore}{\kern0pt}recfun\ {\isacharcolon}{\kern0pt}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}is{\isacharunderscore}{\kern0pt}recfun{\isacharparenleft}{\kern0pt}r{\isacharcomma}{\kern0pt}\ x{\isacharcomma}{\kern0pt}\ H{\isacharcomma}{\kern0pt}\ f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}relation{\isacharparenleft}{\kern0pt}f{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\isanewline
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isanewline
\ \ \isacommand{unfolding}\isamarkupfalse%
\ is{\isacharunderscore}{\kern0pt}recfun{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ a{\isacharequal}{\kern0pt}{\isachardoublequoteopen}{\isacharparenleft}{\kern0pt}{\isasymlambda}x{\isasymin}r\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isachardot}{\kern0pt}\ H{\isacharparenleft}{\kern0pt}x{\isacharcomma}{\kern0pt}\ restrict{\isacharparenleft}{\kern0pt}f{\isacharcomma}{\kern0pt}\ r\ {\isacharminus}{\kern0pt}{\isacharbackquote}{\kern0pt}{\isacharbackquote}{\kern0pt}\ {\isacharbraceleft}{\kern0pt}x{\isacharbraceright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{and}\ b{\isacharequal}{\kern0pt}f\ \isakeyword{in}\ ssubst{\isacharcomma}{\kern0pt}\ simp{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ relation{\isacharunderscore}{\kern0pt}lam{\isacharparenright}{\kern0pt}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/Utilities.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%14=4%:%
%:%15=5%:%
%:%20=5%:%
%:%23=6%:%
%:%24=7%:%
%:%25=7%:%
%:%32=8%:%
%:%33=8%:%
%:%34=9%:%
%:%35=9%:%
%:%36=9%:%
%:%37=10%:%
%:%38=10%:%
%:%39=10%:%
%:%40=10%:%
%:%41=11%:%
%:%42=11%:%
%:%43=11%:%
%:%44=12%:%
%:%45=12%:%
%:%46=13%:%
%:%47=13%:%
%:%48=14%:%
%:%49=14%:%
%:%50=14%:%
%:%51=14%:%
%:%52=15%:%
%:%53=15%:%
%:%54=16%:%
%:%55=16%:%
%:%56=16%:%
%:%57=17%:%
%:%58=17%:%
%:%59=17%:%
%:%60=17%:%
%:%61=17%:%
%:%62=18%:%
%:%68=18%:%
%:%71=19%:%
%:%72=20%:%
%:%73=20%:%
%:%76=21%:%
%:%80=21%:%
%:%81=21%:%
%:%86=21%:%
%:%89=22%:%
%:%90=23%:%
%:%91=23%:%
%:%92=24%:%
%:%99=25%:%
%:%100=25%:%
%:%101=26%:%
%:%102=26%:%
%:%103=27%:%
%:%104=27%:%
%:%105=27%:%
%:%106=27%:%
%:%107=27%:%
%:%108=28%:%
%:%109=28%:%
%:%110=28%:%
%:%111=28%:%
%:%112=28%:%
%:%113=28%:%
%:%114=29%:%
%:%115=29%:%
%:%116=29%:%
%:%117=29%:%
%:%118=30%:%
%:%119=30%:%
%:%120=30%:%
%:%121=30%:%
%:%122=30%:%
%:%123=31%:%
%:%124=31%:%
%:%125=31%:%
%:%126=31%:%
%:%127=31%:%
%:%128=32%:%
%:%134=32%:%
%:%137=33%:%
%:%138=34%:%
%:%139=34%:%
%:%140=35%:%
%:%147=36%:%
%:%148=36%:%
%:%149=37%:%
%:%150=37%:%
%:%151=38%:%
%:%152=38%:%
%:%153=38%:%
%:%154=38%:%
%:%155=39%:%
%:%156=39%:%
%:%157=39%:%
%:%158=39%:%
%:%159=39%:%
%:%160=40%:%
%:%161=40%:%
%:%162=41%:%
%:%163=41%:%
%:%164=42%:%
%:%165=42%:%
%:%166=43%:%
%:%167=43%:%
%:%168=43%:%
%:%169=44%:%
%:%170=44%:%
%:%171=44%:%
%:%172=44%:%
%:%173=44%:%
%:%174=45%:%
%:%175=45%:%
%:%176=45%:%
%:%177=45%:%
%:%178=45%:%
%:%179=46%:%
%:%185=46%:%
%:%188=47%:%
%:%189=48%:%
%:%190=48%:%
%:%193=49%:%
%:%197=49%:%
%:%198=49%:%
%:%199=49%:%
%:%200=49%:%
%:%206=49%:%
%:%209=50%:%
%:%210=51%:%
%:%211=51%:%
%:%213=51%:%
%:%217=51%:%
%:%218=51%:%
%:%225=51%:%
%:%226=52%:%
%:%227=53%:%
%:%228=53%:%
%:%229=54%:%
%:%231=56%:%
%:%234=57%:%
%:%238=57%:%
%:%239=57%:%
%:%240=58%:%
%:%241=58%:%
%:%242=59%:%
%:%243=59%:%
%:%244=59%:%
%:%245=60%:%
%:%246=61%:%
%:%247=62%:%
%:%248=63%:%
%:%249=63%:%
%:%250=63%:%
%:%251=64%:%
%:%252=64%:%
%:%253=65%:%
%:%254=65%:%
%:%255=65%:%
%:%256=65%:%
%:%257=66%:%
%:%258=66%:%
%:%259=66%:%
%:%260=66%:%
%:%261=66%:%
%:%262=67%:%
%:%263=67%:%
%:%264=67%:%
%:%265=67%:%
%:%266=67%:%
%:%267=68%:%
%:%268=68%:%
%:%269=68%:%
%:%270=69%:%
%:%271=69%:%
%:%272=70%:%
%:%273=70%:%
%:%274=71%:%
%:%275=71%:%
%:%276=71%:%
%:%277=72%:%
%:%278=72%:%
%:%279=73%:%
%:%280=73%:%
%:%281=73%:%
%:%282=74%:%
%:%283=75%:%
%:%284=76%:%
%:%285=77%:%
%:%286=77%:%
%:%287=77%:%
%:%288=78%:%
%:%289=78%:%
%:%290=79%:%
%:%291=79%:%
%:%292=79%:%
%:%293=79%:%
%:%294=79%:%
%:%295=80%:%
%:%296=80%:%
%:%297=80%:%
%:%298=80%:%
%:%299=80%:%
%:%300=81%:%
%:%301=81%:%
%:%302=81%:%
%:%303=81%:%
%:%304=81%:%
%:%305=82%:%
%:%306=82%:%
%:%307=82%:%
%:%308=83%:%
%:%309=83%:%
%:%310=84%:%
%:%311=84%:%
%:%312=85%:%
%:%313=85%:%
%:%314=85%:%
%:%315=86%:%
%:%321=86%:%
%:%324=87%:%
%:%325=88%:%
%:%326=89%:%
%:%327=89%:%
%:%330=90%:%
%:%334=90%:%
%:%335=90%:%
%:%336=91%:%
%:%337=91%:%
%:%338=92%:%
%:%339=92%:%
%:%340=92%:%
%:%341=93%:%
%:%342=93%:%
%:%343=93%:%
%:%344=93%:%
%:%345=93%:%
%:%346=94%:%
%:%347=94%:%
%:%348=94%:%
%:%349=94%:%
%:%350=94%:%
%:%351=95%:%
%:%352=95%:%
%:%353=95%:%
%:%354=95%:%
%:%355=95%:%
%:%356=96%:%
%:%357=96%:%
%:%358=96%:%
%:%359=96%:%
%:%360=96%:%
%:%361=97%:%
%:%362=97%:%
%:%363=98%:%
%:%364=98%:%
%:%365=98%:%
%:%366=99%:%
%:%367=99%:%
%:%368=99%:%
%:%369=99%:%
%:%370=99%:%
%:%371=100%:%
%:%372=100%:%
%:%373=100%:%
%:%374=100%:%
%:%375=100%:%
%:%376=101%:%
%:%377=101%:%
%:%378=101%:%
%:%379=101%:%
%:%380=101%:%
%:%381=102%:%
%:%382=102%:%
%:%383=102%:%
%:%384=102%:%
%:%385=103%:%
%:%391=103%:%
%:%394=104%:%
%:%395=105%:%
%:%396=105%:%
%:%399=106%:%
%:%403=106%:%
%:%404=106%:%
%:%405=107%:%
%:%406=107%:%
%:%407=108%:%
%:%408=108%:%
%:%409=108%:%
%:%414=108%:%
%:%417=109%:%
%:%418=110%:%
%:%419=110%:%
%:%422=111%:%
%:%426=111%:%
%:%427=111%:%
%:%428=112%:%
%:%429=112%:%
%:%430=113%:%
%:%431=113%:%
%:%432=113%:%
%:%437=113%:%
%:%440=114%:%
%:%441=115%:%
%:%442=115%:%
%:%445=116%:%
%:%449=116%:%
%:%450=116%:%
%:%451=117%:%
%:%452=117%:%
%:%453=118%:%
%:%454=118%:%
%:%455=119%:%
%:%456=119%:%
%:%457=119%:%
%:%462=119%:%
%:%465=120%:%
%:%466=121%:%
%:%467=121%:%
%:%470=122%:%
%:%474=122%:%
%:%475=122%:%
%:%476=123%:%
%:%477=123%:%
%:%478=123%:%
%:%483=123%:%
%:%486=124%:%
%:%487=125%:%
%:%488=126%:%
%:%489=126%:%
%:%496=127%:%
%:%497=127%:%
%:%498=128%:%
%:%499=128%:%
%:%500=129%:%
%:%501=129%:%
%:%502=129%:%
%:%503=129%:%
%:%504=130%:%
%:%505=130%:%
%:%506=130%:%
%:%507=130%:%
%:%508=130%:%
%:%509=131%:%
%:%510=131%:%
%:%511=131%:%
%:%512=131%:%
%:%513=132%:%
%:%514=132%:%
%:%515=132%:%
%:%516=133%:%
%:%517=133%:%
%:%518=133%:%
%:%519=133%:%
%:%520=134%:%
%:%521=134%:%
%:%522=134%:%
%:%523=134%:%
%:%524=134%:%
%:%525=135%:%
%:%526=135%:%
%:%527=135%:%
%:%528=135%:%
%:%529=135%:%
%:%530=136%:%
%:%531=136%:%
%:%532=136%:%
%:%533=136%:%
%:%534=137%:%
%:%535=137%:%
%:%536=137%:%
%:%537=137%:%
%:%538=138%:%
%:%539=138%:%
%:%540=138%:%
%:%541=139%:%
%:%542=139%:%
%:%543=139%:%
%:%544=139%:%
%:%545=139%:%
%:%546=140%:%
%:%547=140%:%
%:%548=141%:%
%:%549=141%:%
%:%550=142%:%
%:%551=142%:%
%:%552=143%:%
%:%553=143%:%
%:%554=144%:%
%:%555=144%:%
%:%556=144%:%
%:%557=145%:%
%:%558=145%:%
%:%559=146%:%
%:%560=146%:%
%:%561=147%:%
%:%562=147%:%
%:%563=147%:%
%:%564=148%:%
%:%565=148%:%
%:%566=148%:%
%:%567=149%:%
%:%568=149%:%
%:%569=150%:%
%:%570=150%:%
%:%571=151%:%
%:%572=151%:%
%:%573=151%:%
%:%574=151%:%
%:%575=151%:%
%:%576=152%:%
%:%577=152%:%
%:%578=153%:%
%:%579=153%:%
%:%580=153%:%
%:%581=153%:%
%:%582=153%:%
%:%583=154%:%
%:%589=154%:%
%:%592=155%:%
%:%593=156%:%
%:%594=157%:%
%:%595=157%:%
%:%598=158%:%
%:%602=158%:%
%:%603=158%:%
%:%608=158%:%
%:%611=159%:%
%:%612=159%:%
%:%615=160%:%
%:%619=160%:%
%:%620=160%:%
%:%625=160%:%
%:%628=161%:%
%:%629=162%:%
%:%630=162%:%
%:%632=164%:%
%:%635=165%:%
%:%639=165%:%
%:%640=165%:%
%:%641=166%:%
%:%642=166%:%
%:%643=167%:%
%:%644=167%:%
%:%645=167%:%
%:%646=168%:%
%:%647=168%:%
%:%648=168%:%
%:%649=168%:%
%:%650=169%:%
%:%651=169%:%
%:%652=169%:%
%:%653=169%:%
%:%654=169%:%
%:%655=170%:%
%:%656=170%:%
%:%657=170%:%
%:%658=170%:%
%:%659=170%:%
%:%660=171%:%
%:%661=171%:%
%:%662=171%:%
%:%663=171%:%
%:%664=171%:%
%:%665=172%:%
%:%666=172%:%
%:%667=173%:%
%:%668=173%:%
%:%669=173%:%
%:%670=174%:%
%:%671=174%:%
%:%672=174%:%
%:%673=174%:%
%:%674=175%:%
%:%675=175%:%
%:%676=175%:%
%:%677=175%:%
%:%678=175%:%
%:%679=176%:%
%:%680=176%:%
%:%681=176%:%
%:%682=176%:%
%:%683=176%:%
%:%684=177%:%
%:%685=177%:%
%:%686=177%:%
%:%687=177%:%
%:%688=177%:%
%:%689=178%:%
%:%695=178%:%
%:%698=179%:%
%:%699=180%:%
%:%700=181%:%
%:%701=181%:%
%:%702=182%:%
%:%703=183%:%
%:%710=184%:%
%:%711=184%:%
%:%712=185%:%
%:%713=185%:%
%:%714=186%:%
%:%715=186%:%
%:%716=187%:%
%:%717=187%:%
%:%718=187%:%
%:%719=187%:%
%:%720=188%:%
%:%721=188%:%
%:%722=188%:%
%:%723=188%:%
%:%724=188%:%
%:%725=189%:%
%:%726=189%:%
%:%727=189%:%
%:%728=189%:%
%:%729=189%:%
%:%730=190%:%
%:%731=190%:%
%:%732=190%:%
%:%733=190%:%
%:%734=190%:%
%:%735=191%:%
%:%741=191%:%
%:%744=192%:%
%:%745=193%:%
%:%746=193%:%
%:%747=194%:%
%:%754=195%:%
%:%755=195%:%
%:%756=196%:%
%:%757=196%:%
%:%758=197%:%
%:%759=197%:%
%:%760=198%:%
%:%761=198%:%
%:%762=198%:%
%:%763=198%:%
%:%764=199%:%
%:%765=199%:%
%:%766=199%:%
%:%767=199%:%
%:%768=199%:%
%:%769=200%:%
%:%770=200%:%
%:%771=200%:%
%:%772=200%:%
%:%773=200%:%
%:%774=201%:%
%:%775=201%:%
%:%776=201%:%
%:%777=201%:%
%:%778=201%:%
%:%779=202%:%
%:%785=202%:%
%:%788=203%:%
%:%789=204%:%
%:%790=204%:%
%:%792=204%:%
%:%796=204%:%
%:%797=204%:%
%:%804=204%:%
%:%805=205%:%
%:%806=206%:%
%:%807=206%:%
%:%809=206%:%
%:%813=206%:%
%:%814=206%:%
%:%821=206%:%
%:%822=207%:%
%:%823=208%:%
%:%824=208%:%
%:%827=209%:%
%:%831=209%:%
%:%832=209%:%
%:%837=209%:%
%:%840=210%:%
%:%841=211%:%
%:%842=211%:%
%:%845=212%:%
%:%849=212%:%
%:%850=212%:%
%:%851=212%:%
%:%856=212%:%
%:%859=213%:%
%:%860=214%:%
%:%861=214%:%
%:%868=215%:%
%:%869=215%:%
%:%870=216%:%
%:%871=216%:%
%:%872=217%:%
%:%873=217%:%
%:%874=218%:%
%:%875=218%:%
%:%876=219%:%
%:%877=219%:%
%:%878=220%:%
%:%879=220%:%
%:%880=221%:%
%:%881=221%:%
%:%882=221%:%
%:%883=221%:%
%:%884=221%:%
%:%885=222%:%
%:%886=222%:%
%:%887=223%:%
%:%888=223%:%
%:%889=224%:%
%:%890=224%:%
%:%891=224%:%
%:%892=224%:%
%:%893=224%:%
%:%894=225%:%
%:%895=225%:%
%:%896=225%:%
%:%897=225%:%
%:%898=225%:%
%:%899=226%:%
%:%900=226%:%
%:%901=227%:%
%:%907=227%:%
%:%910=228%:%
%:%911=229%:%
%:%912=229%:%
%:%913=230%:%
%:%920=231%:%
%:%921=231%:%
%:%922=232%:%
%:%923=232%:%
%:%924=232%:%
%:%925=233%:%
%:%926=233%:%
%:%927=233%:%
%:%928=233%:%
%:%929=233%:%
%:%930=234%:%
%:%931=234%:%
%:%932=234%:%
%:%933=234%:%
%:%934=234%:%
%:%935=235%:%
%:%936=235%:%
%:%937=235%:%
%:%938=235%:%
%:%939=235%:%
%:%940=236%:%
%:%946=236%:%
%:%949=237%:%
%:%950=238%:%
%:%951=238%:%
%:%952=239%:%
%:%959=240%:%
%:%960=240%:%
%:%961=241%:%
%:%962=241%:%
%:%963=241%:%
%:%964=242%:%
%:%965=242%:%
%:%966=243%:%
%:%967=243%:%
%:%968=244%:%
%:%969=244%:%
%:%970=244%:%
%:%971=245%:%
%:%972=245%:%
%:%973=245%:%
%:%974=245%:%
%:%975=245%:%
%:%976=246%:%
%:%977=246%:%
%:%978=247%:%
%:%984=247%:%
%:%987=248%:%
%:%988=249%:%
%:%989=249%:%
%:%990=250%:%
%:%997=251%:%
%:%998=251%:%
%:%999=252%:%
%:%1000=252%:%
%:%1001=252%:%
%:%1002=253%:%
%:%1003=253%:%
%:%1004=253%:%
%:%1005=254%:%
%:%1006=254%:%
%:%1007=255%:%
%:%1008=255%:%
%:%1009=256%:%
%:%1010=256%:%
%:%1011=257%:%
%:%1017=257%:%
%:%1020=258%:%
%:%1021=259%:%
%:%1022=259%:%
%:%1029=260%:%
%:%1030=260%:%
%:%1031=261%:%
%:%1032=261%:%
%:%1033=262%:%
%:%1034=262%:%
%:%1035=262%:%
%:%1036=262%:%
%:%1037=262%:%
%:%1038=262%:%
%:%1039=263%:%
%:%1040=263%:%
%:%1041=263%:%
%:%1042=264%:%
%:%1043=264%:%
%:%1044=264%:%
%:%1045=264%:%
%:%1046=264%:%
%:%1047=265%:%
%:%1048=265%:%
%:%1049=265%:%
%:%1050=266%:%
%:%1051=266%:%
%:%1052=266%:%
%:%1053=266%:%
%:%1054=266%:%
%:%1055=266%:%
%:%1056=267%:%
%:%1057=267%:%
%:%1058=267%:%
%:%1059=267%:%
%:%1060=268%:%
%:%1061=268%:%
%:%1062=268%:%
%:%1063=268%:%
%:%1064=268%:%
%:%1065=269%:%
%:%1066=269%:%
%:%1067=269%:%
%:%1068=269%:%
%:%1069=269%:%
%:%1070=270%:%
%:%1076=270%:%
%:%1079=271%:%
%:%1080=272%:%
%:%1081=272%:%
%:%1084=273%:%
%:%1088=273%:%
%:%1089=273%:%
%:%1090=274%:%
%:%1091=274%:%
%:%1092=275%:%
%:%1093=275%:%
%:%1094=276%:%
%:%1095=276%:%
%:%1096=276%:%
%:%1097=277%:%
%:%1098=277%:%
%:%1099=277%:%
%:%1100=277%:%
%:%1101=277%:%
%:%1102=278%:%
%:%1103=278%:%
%:%1104=278%:%
%:%1105=278%:%
%:%1106=278%:%
%:%1107=279%:%
%:%1108=279%:%
%:%1109=279%:%
%:%1110=280%:%
%:%1111=280%:%
%:%1112=281%:%
%:%1113=281%:%
%:%1114=281%:%
%:%1115=282%:%
%:%1116=282%:%
%:%1117=282%:%
%:%1118=282%:%
%:%1119=282%:%
%:%1120=283%:%
%:%1121=283%:%
%:%1122=283%:%
%:%1123=284%:%
%:%1124=284%:%
%:%1125=285%:%
%:%1126=285%:%
%:%1127=285%:%
%:%1128=286%:%
%:%1129=286%:%
%:%1130=286%:%
%:%1131=286%:%
%:%1132=287%:%
%:%1133=287%:%
%:%1134=287%:%
%:%1135=287%:%
%:%1136=287%:%
%:%1137=287%:%
%:%1138=288%:%
%:%1139=288%:%
%:%1140=288%:%
%:%1141=288%:%
%:%1142=288%:%
%:%1143=289%:%
%:%1144=289%:%
%:%1145=290%:%
%:%1146=290%:%
%:%1147=290%:%
%:%1148=291%:%
%:%1149=291%:%
%:%1150=291%:%
%:%1151=292%:%
%:%1152=292%:%
%:%1153=293%:%
%:%1154=293%:%
%:%1155=294%:%
%:%1156=294%:%
%:%1157=294%:%
%:%1158=294%:%
%:%1159=294%:%
%:%1160=294%:%
%:%1161=295%:%
%:%1162=295%:%
%:%1163=295%:%
%:%1164=296%:%
%:%1165=296%:%
%:%1166=297%:%
%:%1167=297%:%
%:%1168=298%:%
%:%1169=298%:%
%:%1170=299%:%
%:%1171=299%:%
%:%1172=299%:%
%:%1173=299%:%
%:%1174=299%:%
%:%1175=299%:%
%:%1176=300%:%
%:%1182=300%:%
%:%1185=301%:%
%:%1186=302%:%
%:%1187=302%:%
%:%1190=303%:%
%:%1194=303%:%
%:%1195=303%:%
%:%1196=303%:%
%:%1202=303%:%
%:%1205=304%:%
%:%1206=305%:%
%:%1207=305%:%
%:%1210=306%:%
%:%1214=306%:%
%:%1215=306%:%
%:%1216=306%:%
%:%1222=306%:%
%:%1225=307%:%
%:%1226=308%:%
%:%1227=308%:%
%:%1229=308%:%
%:%1233=308%:%
%:%1234=308%:%
%:%1241=308%:%
%:%1242=309%:%
%:%1243=310%:%
%:%1244=310%:%
%:%1246=310%:%
%:%1250=310%:%
%:%1251=310%:%
%:%1258=310%:%
%:%1259=311%:%
%:%1260=312%:%
%:%1261=312%:%
%:%1263=312%:%
%:%1267=312%:%
%:%1268=312%:%
%:%1275=312%:%
%:%1276=313%:%
%:%1277=313%:%
%:%1279=313%:%
%:%1283=313%:%
%:%1284=313%:%
%:%1291=313%:%
%:%1292=314%:%
%:%1293=314%:%
%:%1295=314%:%
%:%1299=314%:%
%:%1300=314%:%
%:%1307=314%:%
%:%1308=315%:%
%:%1309=315%:%
%:%1311=315%:%
%:%1315=315%:%
%:%1316=315%:%
%:%1323=315%:%
%:%1324=316%:%
%:%1325=316%:%
%:%1327=316%:%
%:%1331=316%:%
%:%1332=316%:%
%:%1339=316%:%
%:%1340=317%:%
%:%1341=317%:%
%:%1343=317%:%
%:%1347=317%:%
%:%1348=317%:%
%:%1355=317%:%
%:%1356=318%:%
%:%1357=318%:%
%:%1359=318%:%
%:%1363=318%:%
%:%1364=318%:%
%:%1371=318%:%
%:%1372=319%:%
%:%1373=319%:%
%:%1375=319%:%
%:%1379=319%:%
%:%1380=319%:%
%:%1387=319%:%
%:%1388=320%:%
%:%1389=320%:%
%:%1391=320%:%
%:%1395=320%:%
%:%1396=320%:%
%:%1403=320%:%
%:%1404=321%:%
%:%1405=321%:%
%:%1407=321%:%
%:%1411=321%:%
%:%1412=321%:%
%:%1419=321%:%
%:%1420=322%:%
%:%1421=322%:%
%:%1423=322%:%
%:%1427=322%:%
%:%1428=322%:%
%:%1435=322%:%
%:%1436=323%:%
%:%1437=324%:%
%:%1438=324%:%
%:%1441=325%:%
%:%1445=325%:%
%:%1446=325%:%
%:%1447=325%:%
%:%1454=325%:%
%:%1455=326%:%
%:%1456=327%:%
%:%1457=327%:%
%:%1460=328%:%
%:%1464=328%:%
%:%1465=328%:%
%:%1466=328%:%
%:%1473=328%:%
%:%1474=329%:%
%:%1475=330%:%
%:%1476=330%:%
%:%1479=331%:%
%:%1483=331%:%
%:%1484=331%:%
%:%1485=332%:%
%:%1486=332%:%
%:%1487=333%:%
%:%1488=333%:%
%:%1489=334%:%
%:%1490=334%:%
%:%1491=335%:%
%:%1492=335%:%
%:%1493=336%:%
%:%1494=336%:%
%:%1495=337%:%
%:%1496=337%:%
%:%1497=338%:%
%:%1498=338%:%
%:%1499=339%:%
%:%1500=339%:%
%:%1501=340%:%
%:%1502=340%:%
%:%1507=340%:%
%:%1510=341%:%
%:%1511=342%:%
%:%1512=342%:%
%:%1515=343%:%
%:%1519=343%:%
%:%1520=343%:%
%:%1521=344%:%
%:%1522=344%:%
%:%1523=345%:%
%:%1524=345%:%
%:%1525=346%:%
%:%1526=346%:%
%:%1531=346%:%
%:%1534=347%:%
%:%1535=348%:%
%:%1536=348%:%
%:%1539=349%:%
%:%1543=349%:%
%:%1544=349%:%
%:%1545=350%:%
%:%1546=350%:%
%:%1547=351%:%
%:%1548=351%:%
%:%1553=351%:%
%:%1556=352%:%
%:%1557=353%:%
%:%1558=353%:%
%:%1561=354%:%
%:%1565=354%:%
%:%1566=354%:%
%:%1567=355%:%
%:%1568=355%:%
%:%1569=356%:%
%:%1570=356%:%
%:%1571=357%:%
%:%1572=357%:%
%:%1573=358%:%
%:%1574=358%:%
%:%1579=358%:%
%:%1582=359%:%
%:%1583=360%:%
%:%1584=360%:%
%:%1591=361%:%
%:%1592=361%:%
%:%1593=362%:%
%:%1594=362%:%
%:%1595=363%:%
%:%1596=363%:%
%:%1597=363%:%
%:%1598=363%:%
%:%1599=364%:%
%:%1600=364%:%
%:%1601=364%:%
%:%1602=365%:%
%:%1603=365%:%
%:%1604=366%:%
%:%1605=366%:%
%:%1606=367%:%
%:%1607=367%:%
%:%1608=368%:%
%:%1614=368%:%
%:%1617=369%:%
%:%1618=370%:%
%:%1619=370%:%
%:%1620=371%:%
%:%1621=372%:%
%:%1624=373%:%
%:%1625=374%:%
%:%1629=374%:%
%:%1630=374%:%
%:%1631=375%:%
%:%1632=375%:%
%:%1633=375%:%
%:%1634=376%:%
%:%1635=376%:%
%:%1636=377%:%
%:%1637=377%:%
%:%1638=378%:%
%:%1639=378%:%
%:%1640=379%:%
%:%1641=379%:%
%:%1642=379%:%
%:%1643=379%:%
%:%1644=380%:%
%:%1645=380%:%
%:%1646=381%:%
%:%1647=381%:%
%:%1648=382%:%
%:%1649=382%:%
%:%1650=382%:%
%:%1651=383%:%
%:%1652=383%:%
%:%1653=383%:%
%:%1654=383%:%
%:%1655=383%:%
%:%1656=383%:%
%:%1657=384%:%
%:%1658=384%:%
%:%1659=385%:%
%:%1660=385%:%
%:%1661=386%:%
%:%1662=386%:%
%:%1663=386%:%
%:%1664=386%:%
%:%1665=387%:%
%:%1666=387%:%
%:%1667=387%:%
%:%1668=388%:%
%:%1669=388%:%
%:%1670=389%:%
%:%1671=389%:%
%:%1672=389%:%
%:%1673=389%:%
%:%1674=389%:%
%:%1675=390%:%
%:%1676=390%:%
%:%1677=391%:%
%:%1683=391%:%
%:%1686=392%:%
%:%1687=393%:%
%:%1688=393%:%
%:%1689=394%:%
%:%1690=395%:%
%:%1697=396%:%
%:%1698=396%:%
%:%1699=397%:%
%:%1700=397%:%
%:%1701=397%:%
%:%1702=397%:%
%:%1703=398%:%
%:%1704=398%:%
%:%1705=398%:%
%:%1706=398%:%
%:%1707=398%:%
%:%1708=399%:%
%:%1709=399%:%
%:%1710=399%:%
%:%1711=399%:%
%:%1712=400%:%
%:%1713=400%:%
%:%1714=400%:%
%:%1715=400%:%
%:%1716=400%:%
%:%1717=401%:%
%:%1718=401%:%
%:%1719=401%:%
%:%1720=401%:%
%:%1721=401%:%
%:%1722=402%:%
%:%1728=402%:%
%:%1731=403%:%
%:%1732=404%:%
%:%1733=404%:%
%:%1734=405%:%
%:%1735=406%:%
%:%1738=407%:%
%:%1739=408%:%
%:%1743=408%:%
%:%1744=408%:%
%:%1745=408%:%
%:%1746=408%:%
%:%1747=408%:%
%:%1748=409%:%
%:%1749=409%:%
%:%1754=409%:%
%:%1757=410%:%
%:%1758=411%:%
%:%1759=411%:%
%:%1760=412%:%
%:%1761=413%:%
%:%1764=414%:%
%:%1768=414%:%
%:%1769=414%:%
%:%1770=415%:%
%:%1771=415%:%
%:%1772=415%:%
%:%1773=415%:%
%:%1774=416%:%
%:%1775=416%:%
%:%1776=417%:%
%:%1777=417%:%
%:%1778=418%:%
%:%1779=418%:%
%:%1780=419%:%
%:%1781=419%:%
%:%1782=419%:%
%:%1783=419%:%
%:%1784=420%:%
%:%1785=420%:%
%:%1786=420%:%
%:%1787=421%:%
%:%1788=421%:%
%:%1789=422%:%
%:%1790=422%:%
%:%1791=423%:%
%:%1792=423%:%
%:%1793=424%:%
%:%1794=424%:%
%:%1795=425%:%
%:%1796=425%:%
%:%1797=426%:%
%:%1798=426%:%
%:%1799=426%:%
%:%1800=427%:%
%:%1801=427%:%
%:%1802=427%:%
%:%1803=427%:%
%:%1804=428%:%
%:%1805=428%:%
%:%1806=429%:%
%:%1807=429%:%
%:%1808=429%:%
%:%1809=430%:%
%:%1810=431%:%
%:%1811=432%:%
%:%1812=432%:%
%:%1813=432%:%
%:%1814=432%:%
%:%1815=433%:%
%:%1816=433%:%
%:%1817=433%:%
%:%1818=433%:%
%:%1819=433%:%
%:%1820=434%:%
%:%1821=434%:%
%:%1822=434%:%
%:%1823=434%:%
%:%1824=435%:%
%:%1825=435%:%
%:%1826=435%:%
%:%1827=436%:%
%:%1828=436%:%
%:%1829=437%:%
%:%1830=437%:%
%:%1831=438%:%
%:%1832=438%:%
%:%1833=438%:%
%:%1834=439%:%
%:%1835=439%:%
%:%1836=440%:%
%:%1842=440%:%
%:%1845=441%:%
%:%1846=442%:%
%:%1847=442%:%
%:%1850=443%:%
%:%1854=443%:%
%:%1855=443%:%
%:%1856=443%:%
%:%1857=443%:%
%:%1858=443%:%
%:%1859=444%:%
%:%1860=444%:%
%:%1861=445%:%
%:%1862=445%:%
%:%1863=445%:%
%:%1864=446%:%
%:%1865=446%:%
%:%1866=446%:%
%:%1867=446%:%
%:%1868=446%:%
%:%1869=447%:%
%:%1870=447%:%
%:%1871=447%:%
%:%1872=447%:%
%:%1873=447%:%
%:%1874=447%:%
%:%1875=448%:%
%:%1876=448%:%
%:%1877=448%:%
%:%1878=448%:%
%:%1879=448%:%
%:%1880=449%:%
%:%1886=449%:%
%:%1889=450%:%
%:%1890=451%:%
%:%1891=451%:%
%:%1894=452%:%
%:%1898=452%:%
%:%1899=452%:%
%:%1900=453%:%
%:%1901=453%:%
%:%1906=453%:%
%:%1909=454%:%
%:%1910=455%:%
%:%1911=455%:%
%:%1914=456%:%
%:%1918=456%:%
%:%1919=456%:%
%:%1920=457%:%
%:%1921=457%:%
%:%1922=458%:%
%:%1923=459%:%
%:%1924=459%:%
%:%1925=460%:%
%:%1926=460%:%
%:%1927=461%:%
%:%1928=461%:%
%:%1929=461%:%
%:%1930=462%:%
%:%1931=462%:%
%:%1932=462%:%
%:%1933=462%:%
%:%1934=462%:%
%:%1935=463%:%
%:%1936=463%:%
%:%1937=463%:%
%:%1938=463%:%
%:%1939=463%:%
%:%1940=464%:%
%:%1941=464%:%
%:%1942=464%:%
%:%1943=464%:%
%:%1944=464%:%
%:%1945=465%:%
%:%1946=465%:%
%:%1947=466%:%
%:%1948=466%:%
%:%1949=466%:%
%:%1950=467%:%
%:%1951=467%:%
%:%1952=467%:%
%:%1953=467%:%
%:%1954=468%:%
%:%1955=468%:%
%:%1956=468%:%
%:%1957=469%:%
%:%1958=469%:%
%:%1959=469%:%
%:%1960=469%:%
%:%1961=470%:%
%:%1962=470%:%
%:%1963=470%:%
%:%1964=470%:%
%:%1965=470%:%
%:%1966=471%:%
%:%1967=471%:%
%:%1968=472%:%
%:%1969=473%:%
%:%1970=473%:%
%:%1971=473%:%
%:%1972=474%:%
%:%1973=474%:%
%:%1974=474%:%
%:%1975=474%:%
%:%1976=474%:%
%:%1977=475%:%
%:%1978=475%:%
%:%1979=475%:%
%:%1980=475%:%
%:%1981=476%:%
%:%1982=476%:%
%:%1983=476%:%
%:%1984=476%:%
%:%1985=477%:%
%:%1986=477%:%
%:%1987=477%:%
%:%1988=477%:%
%:%1989=478%:%
%:%1995=478%:%
%:%1998=479%:%
%:%1999=480%:%
%:%2000=480%:%
%:%2003=481%:%
%:%2007=481%:%
%:%2008=481%:%
%:%2009=482%:%
%:%2010=482%:%
%:%2011=483%:%
%:%2012=483%:%
%:%2013=484%:%
%:%2014=484%:%
%:%2015=485%:%
%:%2016=485%:%
%:%2017=486%:%
%:%2023=486%:%
%:%2026=487%:%
%:%2027=488%:%
%:%2028=488%:%
%:%2031=489%:%
%:%2035=489%:%
%:%2036=489%:%
%:%2041=489%:%
%:%2044=490%:%
%:%2045=491%:%
%:%2046=491%:%
%:%2049=492%:%
%:%2053=492%:%
%:%2054=492%:%
%:%2059=492%:%
%:%2062=493%:%
%:%2063=494%:%
%:%2064=494%:%
%:%2067=495%:%
%:%2071=495%:%
%:%2072=495%:%
%:%2073=496%:%
%:%2074=496%:%
%:%2075=497%:%
%:%2076=497%:%
%:%2081=497%:%
%:%2084=498%:%
%:%2085=498%:%
%:%2088=499%:%
%:%2092=499%:%
%:%2093=499%:%
%:%2094=500%:%
%:%2095=500%:%
%:%2096=501%:%
%:%2097=501%:%
%:%2102=501%:%
%:%2105=502%:%
%:%2106=503%:%
%:%2107=504%:%
%:%2108=504%:%
%:%2109=505%:%
%:%2110=506%:%
%:%2111=507%:%
%:%2118=508%:%
%:%2119=508%:%
%:%2120=509%:%
%:%2121=509%:%
%:%2122=509%:%
%:%2123=510%:%
%:%2124=510%:%
%:%2125=510%:%
%:%2126=511%:%
%:%2127=511%:%
%:%2128=512%:%
%:%2129=512%:%
%:%2130=513%:%
%:%2131=513%:%
%:%2132=514%:%
%:%2133=514%:%
%:%2134=515%:%
%:%2135=515%:%
%:%2136=515%:%
%:%2137=516%:%
%:%2138=516%:%
%:%2139=517%:%
%:%2140=517%:%
%:%2141=518%:%
%:%2142=518%:%
%:%2143=519%:%
%:%2144=519%:%
%:%2145=520%:%
%:%2146=520%:%
%:%2147=520%:%
%:%2148=520%:%
%:%2149=521%:%
%:%2150=522%:%
%:%2151=522%:%
%:%2152=522%:%
%:%2153=523%:%
%:%2154=523%:%
%:%2155=524%:%
%:%2156=524%:%
%:%2157=525%:%
%:%2163=525%:%
%:%2166=526%:%
%:%2167=527%:%
%:%2168=527%:%
%:%2169=528%:%
%:%2172=529%:%
%:%2176=529%:%
%:%2177=529%:%
%:%2178=530%:%
%:%2179=530%:%
%:%2184=530%:%
%:%2187=531%:%
%:%2188=532%:%
%:%2189=532%:%
%:%2190=533%:%
%:%2191=534%:%
%:%2194=535%:%
%:%2195=536%:%
%:%2199=536%:%
%:%2200=536%:%
%:%2201=537%:%
%:%2202=537%:%
%:%2203=538%:%
%:%2204=538%:%
%:%2205=539%:%
%:%2206=539%:%
%:%2207=540%:%
%:%2213=540%:%
%:%2216=541%:%
%:%2219=542%:%
%:%2224=543%:%