### 2024-11-01 code/output/document/ZF_notAC.tex

%
\begin{isabellebody}%
\setisabellecontext{ZF{\isacharunderscore}{\kern0pt}notAC}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}notAC\ \isanewline
\ \ \isakeyword{imports}\ NotAC{\isacharunderscore}{\kern0pt}Proof\ \isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
\isanewline
\isanewline
\isacommand{theorem}\isamarkupfalse%
\ ZF{\isacharunderscore}{\kern0pt}notAC{\isacharunderscore}{\kern0pt}main{\isacharunderscore}{\kern0pt}theorem\ {\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ M\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}nat\ {\isasymapprox}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}M\ {\isasymTurnstile}\ ZF{\isachardoublequoteclose}\ {\isachardoublequoteopen}Transset{\isacharparenleft}{\kern0pt}M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymexists}N{\isachardot}{\kern0pt}\ N\ {\isasymTurnstile}\ ZF\ {\isasymand}\ {\isasymnot}{\isacharparenleft}{\kern0pt}{\isasymforall}A\ {\isasymin}\ N{\isachardot}{\kern0pt}\ {\isasymexists}R\ {\isasymin}\ N{\isachardot}{\kern0pt}\ wellordered{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}N{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ R{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharminus}{\kern0pt}\ \isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ enum\ \isakeyword{where}\ {\isachardoublequoteopen}enum\ {\isasymin}\ bij{\isacharparenleft}{\kern0pt}nat{\isacharcomma}{\kern0pt}\ M{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms\ eqpoll{\isacharunderscore}{\kern0pt}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{interpret}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm\ M\ enum\ \isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ctm{\isacharunderscore}{\kern0pt}def\ M{\isacharunderscore}{\kern0pt}ctm{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}satT\ assms\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{interpret}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\ Fn\ Fn{\isacharunderscore}{\kern0pt}leq\ {\isadigit{0}}\ M\ enum\ Fn{\isacharunderscore}{\kern0pt}perms\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter\ \ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ Fn{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{obtain}\isamarkupfalse%
\ G\ \isakeyword{where}\ GH{\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}M{\isacharunderscore}{\kern0pt}generic{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ generic{\isacharunderscore}{\kern0pt}filter{\isacharunderscore}{\kern0pt}existence\ zero{\isacharunderscore}{\kern0pt}in{\isacharunderscore}{\kern0pt}Fn\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{interpret}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic\ Fn\ Fn{\isacharunderscore}{\kern0pt}leq\ {\isadigit{0}}\ M\ enum\ Fn{\isacharunderscore}{\kern0pt}perms\ Fn{\isacharunderscore}{\kern0pt}perms{\isacharunderscore}{\kern0pt}filter\ G\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ M{\isacharunderscore}{\kern0pt}symmetric{\isacharunderscore}{\kern0pt}system{\isacharunderscore}{\kern0pt}axioms\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}def\ forcing{\isacharunderscore}{\kern0pt}data{\isacharunderscore}{\kern0pt}axioms\ G{\isacharunderscore}{\kern0pt}generic{\isacharunderscore}{\kern0pt}axioms{\isacharunderscore}{\kern0pt}def\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{define}\isamarkupfalse%
\ N\ \isakeyword{where}\ {\isachardoublequoteopen}N\ {\isasymequiv}\ SymExt{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ N{\isacharunderscore}{\kern0pt}ZF\ {\isacharcolon}{\kern0pt}\ {\isachardoublequoteopen}N\ {\isasymTurnstile}\ ZF{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ N{\isacharunderscore}{\kern0pt}def\ SymExt{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}ZF\ M{\isacharunderscore}{\kern0pt}ZF{\isacharunderscore}{\kern0pt}iff{\isacharunderscore}{\kern0pt}M{\isacharunderscore}{\kern0pt}satT\ \isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}A\ {\isasymin}\ N{\isachardot}{\kern0pt}\ {\isasymforall}R\ {\isasymin}\ N{\isachardot}{\kern0pt}\ {\isasymnot}wellordered{\isacharparenleft}{\kern0pt}{\isacharhash}{\kern0pt}{\isacharhash}{\kern0pt}N{\isacharcomma}{\kern0pt}\ A{\isacharcomma}{\kern0pt}\ R{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule{\isacharunderscore}{\kern0pt}tac\ x{\isacharequal}{\kern0pt}{\isachardoublequoteopen}binmap{\isacharparenleft}{\kern0pt}G{\isacharparenright}{\kern0pt}{\isachardoublequoteclose}\ \isakeyword{in}\ bexI{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ ballI{\isacharcomma}{\kern0pt}\ rule\ ccontr{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \ \isacommand{apply}\isamarkupfalse%
{\isacharparenleft}{\kern0pt}rule\ no{\isacharunderscore}{\kern0pt}wellorder{\isacharparenright}{\kern0pt}\isanewline
\ \ \ \ \isacommand{using}\isamarkupfalse%
\ GH\ N{\isacharunderscore}{\kern0pt}def\ SymExt{\isacharunderscore}{\kern0pt}def\ binmap{\isacharunderscore}{\kern0pt}eq\ binmap{\isacharprime}{\kern0pt}{\isacharunderscore}{\kern0pt}HS\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ auto\ \isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}{\kern0pt}thesis\ \isacommand{using}\isamarkupfalse%
\ N{\isacharunderscore}{\kern0pt}ZF\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/source/repos/ZF-notAC/code/ZF_notAC.thy%:%
%:%10=1%:%
%:%11=1%:%
%:%12=2%:%
%:%13=3%:%
%:%18=3%:%
%:%21=4%:%
%:%22=5%:%
%:%23=6%:%
%:%24=6%:%
%:%25=7%:%
%:%26=8%:%
%:%27=9%:%
%:%34=10%:%
%:%35=10%:%
%:%36=11%:%
%:%37=12%:%
%:%38=12%:%
%:%39=12%:%
%:%40=12%:%
%:%41=13%:%
%:%42=13%:%
%:%43=13%:%
%:%44=14%:%
%:%45=14%:%
%:%46=15%:%
%:%47=15%:%
%:%48=16%:%
%:%49=16%:%
%:%50=17%:%
%:%51=17%:%
%:%52=18%:%
%:%53=18%:%
%:%54=19%:%
%:%55=19%:%
%:%56=20%:%
%:%57=21%:%
%:%58=21%:%
%:%59=22%:%
%:%60=22%:%
%:%61=23%:%
%:%62=23%:%
%:%63=24%:%
%:%64=24%:%
%:%65=24%:%
%:%66=25%:%
%:%67=25%:%
%:%68=26%:%
%:%69=26%:%
%:%70=27%:%
%:%71=27%:%
%:%72=28%:%
%:%73=29%:%
%:%74=29%:%
%:%75=30%:%
%:%76=31%:%
%:%77=31%:%
%:%78=32%:%
%:%79=32%:%
%:%80=33%:%
%:%81=33%:%
%:%82=34%:%
%:%83=35%:%
%:%84=35%:%
%:%85=36%:%
%:%86=36%:%
%:%87=37%:%
%:%88=37%:%
%:%89=38%:%
%:%90=38%:%
%:%91=39%:%
%:%92=39%:%
%:%93=40%:%
%:%94=40%:%
%:%95=41%:%
%:%96=41%:%
%:%97=41%:%
%:%98=41%:%
%:%99=41%:%
%:%100=42%:%
%:%106=42%:%
%:%109=43%:%
%:%110=44%:%
%:%113=45%:%
%:%118=46%:%